### 2020-12-10 EDA_wages_hs2020.Rmd
Add calculating features

---
title: "EDA of Wages Tidied Data"
author: "Di Cook, Emi Tanaka, Dewi Lestari Amaliah"
date: "`r Sys.Date()`"
output: 
  bookdown::html_document2:
      keep_tex: yes
      number_sections: yes
      citation_package: biblatex
      toc: true
      toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

library(tidyverse)
library(brolgar)
library(visdat)
library(naniar)
library(skimr)
library(kableExtra)
library(patchwork)
library(gghighlight)
library(stickylabeller)
```

# Background

# Data Description

```{r data-load}
load("wages_hs2020.rda")
```

## Data Structures

```{r dat-structure, fig.cap = "Data structures of wages data"}
visdat::vis_dat(wages_hs2020, 
                warn_large_data = FALSE)
```

## Skim the data

```{r skim}

skim(wages_hs2020) %>%
  as_tibble() %>%
  kable(caption = "The summary of wages data") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

There are 5274 missing values observations in the mean of hourly wage variable or more than 5 percent from the total observations in the data set. This NA is not because of the subject is joining the survey late or out from the survey, but it is the missing value in the series.

# Calculating Features of Data

## Number of Observation per ID

```{r wages_tsibble}
wages <- as_tsibble(x = wages_hs2020,
                    key = id,
                    index = year,
                    regular = FALSE)
```

```{r}

wages_obs <- wages %>%
  features(mean_hourly_wage, n_obs) %>%
  arrange(n_obs)

ggplot(wages_obs, aes(x = n_obs)) +
  geom_bar() +
  xlab("Number of observations")

```

Learn: 
Different years of observations. The minimum years of observation is 5 years, while the maximum years of observation is 27 years.

```{r}
wages %>%
  features(year, list(min = min)) %>%
  ggplot(aes(x = min)) +
  geom_histogram(binwidth = 1) +
  xlab("First time in study")
```

Learn: Subject starts in the study at different year, most of them start in 1979.

```{r}
wages %>%
  features(year, list(max = max))%>%
  ggplot(aes(x = max)) +
  geom_histogram(binwidth = 1) +
  xlab("Last time in study")
```

Learn: Subject out from the study in the study at different year, most of them are still in the study till the study ends in 2016.

```{r}
wages_year_range <- wages %>%
  features(year, feat_ranges)

ggplot(wages_year_range,
       aes(x = range_diff)) +
  geom_histogram() +
  xlab("Range of year")
```

```{r}
index_regular(wages)
```

Note: the number of observation did not match with the range of year because the year is actually not sequential annually (not regular).

## Minimum, Median, and Maximum of each ID

```{r features}
wages_feature<- wages %>%
  features(mean_hourly_wage, 
           feat_three_num
           )
```

```{r p1}
p1 <- ggplot(wages_feature) +
  geom_density(aes(x = min), fill = "magenta", alpha = 0.5)
```

```{r p2}

wages_feature_long <- wages_feature %>%
  pivot_longer(c(min, med, max), names_to = "feature", values_to = "value")

p2 <- ggplot(wages_feature_long) +
  geom_density(aes(x = value, colour = feature, fill = feature), alpha = 0.3) +
  theme(legend.position = "None")
```

```{r p3}
p3 <- ggplot(transform(wages_feature_long,
                 feature = factor(feature, level = 
                                 c("min",
                                   "med", 
                                   "max")))) +
  geom_line(aes(x = feature,
                y = value, 
                group = id), alpha = 0.5)
```

```{r feature-charts}

p1+p2+p3

```

Learn:

The distribution of min, median, and max of wages is right-skewed, mean that some subjects have an extremely high hourly wages. Further, Some subjects have a minimum wages that is higher than others' maximum wages.

## Is their wage always increase?

```{r}
wages_feat_monotonic <- wages %>%
  features(mean_hourly_wage, 
           feat_monotonic
           )

wages_feat_inc <- wages_feat_monotonic %>%
  filter(increase == TRUE)
```

Learn: Only `r nrow(wages_feat_inc)` subjects out of `r nrow(wages_feat_monotonic)` or `r round(nrow(wages_feat_inc)/nrow(wages_feat_monotonic)*100, 2)` whose mean hourly wage always increase in the period of the study.

## Fit a linear model for each key using `key_slope()`

```{r}
wages_slope <- key_slope(wages, 
                         mean_hourly_wage ~ year) %>%
  left_join(wages, by = "id")

wages_slope
```

```{r decrease-slope, fig.cap= "Subject with negative wages slope"}
ggplot(wages_slope) +
  geom_line(aes(x = year,
                y = mean_hourly_wage,
                group = id)) +
  gghighlight(.slope_year < 0)
```

Learn: Some subjects had a negative slope, means that their wages tend to decrease over time.

## Find keys near some summary statistics

```{r}
wages_near <- wages %>%
  key_slope(mean_hourly_wage ~ year) %>%
  keys_near(key = id,
            var = .slope_year)

wages_near
```

### 2020-12-17 EDA_wages_hs2020.Rmd
breaking the spaghetti plot into sample

@@
 
-# Data Description
-
 ```{r data-load}
@@
 ```
+
+## The original Spaghetty Plot
+
+```{r org_spag, fig.cap = "Wages longitudinal data set"}
+fig_all <- ggplot(wages_hs2020) +
+  geom_line(aes(x = year,
+                y = mean_hourly_wage,
+                group = id))
+
+fig_all_transp <- ggplot(wages_hs2020) +
+  geom_line(aes(x = year,
+                y = mean_hourly_wage,
+                group = id),
+            alpha = 0.1)
+
+fig_all_lm <- ggplot(wages_hs2020) +
+  geom_line(aes(x = year,
+                y = mean_hourly_wage,
+                group = id),
+            alpha = 0.1) +
+  geom_smooth(aes(x = year,
+                  y = mean_hourly_wage),
+              method = lm,
+              se = FALSE)
+
+fig_all + fig_all_transp + fig_all_lm
+```
+Learn: It is really hard to see individual features since there are a lot observations over time, even if we  change the transparency of the plot.
+
+# Data Description
 
@@
 
-```{r}
+```{r first-time, fig.cap="The distribution of sample based on the year to be the first time in the study"}
 wages %>%
@@
 ```
+
+# Breaking up the spaghetti
+
+Figure \@ref(fig:org-spag) shows that when the data is presented as a whole spaghetti plot, we can only gain a high level understanding of the data.
+Hence, we have to break up the spaghetti to see a detailed figure of the individual.
+We can do this by sampling, stratifying, featuring that strata.
+
+## Sampling
+
+By using `brolgar's` `facet_sample()` function, we can randomly sample the data into a faceted plots.
+
+```{r facet-sample, fig.cap="Mean of hourly wage by sampling the individual"}
+ggplot(wages) +
+  geom_line(aes(x = year,
+                y = mean_hourly_wage,
+                group = id)) +
+  facet_sample()
+```
+Learn: By default, `facet_sample()` function created 12 facets and 3 sample in each facet.
+However, as we can see in Figure \@ref(fig:facet-sample), in panel 4 and 10, there are four lines instead of 3.
+This is due to the missing value in the individual.
+We can also learn that the subject generally have a positive trend.
+
+## Stratifying
+
+`facet_sample()` only shows part of the data, we can use `facet_strata` to show all the id in different facet.
+
+```{r facet-strata, fig.cap= "Faceted mean hourly wages"}
+
+ggplot(wages) +
+  geom_line(aes(x = year,
+                y = mean_hourly_wage,
+                group = id)) +
+  facet_strata(n_strata = 100)
+
+```
+
+As we have many ids, it is still hard to see individual data even if it is faceted in to 100 facets. However, we can see that they generally have the same pattern.
+
+## Featuring
+
+We can add the feature in Figure \@ref(fig:facet-strata) by arranging it by year. Here, the data is not randomly allocated into each facet, but they are displayed a specified range of year values.
+
+```{r}
+ggplot(wages) +
+  geom_line(aes(x = year,
+                y = mean_hourly_wage,
+                group = id),
+            alpha = 0.1) +
+  facet_strata(along = -year)
+```
+Learn: Most of the samples participated in the study in 1979, this is also inline with the finding in Figure \@ref(fig:first-time).
+We can see that generally the subjects of the study have the same starting wages but it became vary in the end of the study.

### 2020-12-22 EDA_wages_hs2020.Rmd
Add GAM model

@@
 library(stickylabeller)
+library(mgcv)
+library(modelr)
+library(tsibble)
 ```
@@
 
+set.seed(2020)
 ggplot(wages) +
@@
 ```{r}
+set.seed(2020)
 ggplot(wages) +
@@
 We can see that generally the subjects of the study have the same starting wages but it became vary in the end of the study.
+
+## Build GAM model
+
+```{r id-fct}
+wages <- wages %>%
+  mutate(id_fct = as.factor(id),
+         year0 = year - 1979)
+```
+
+```{r eval = FALSE}
+# wages_gam <- gam(
+#   mean_hourly_wage ~ s(year0, by = id_fct) + id_fct,
+#   data = wages,
+#   method = "REML"
+# )
+```
+
+Since we have a quite large data set, we have to subset it first to model the data.
+Here, we would like to build a Generalized Additive Model (GAM) model, by randomly subset the daya the data using brolgar's` `sample_frac_keys` function.
+
+```{r}
+set.seed(2020)
+wages_sample <- sample_frac_keys(wages, size = 0.01)
+
+```
+
+```{r}
+wages_gam <- gam(
+  mean_hourly_wage ~ s(year0, by = id_fct) + id_fct,
+  data = wages_sample,
+  method = "REML"
+)
+```
+
+```{r}
+wages_aug <- wages_sample %>%
+  add_predictions(wages_gam, var = "pred") %>%
+  add_residuals(wages_gam, var = "res") %>%
+  group_by_key() %>%
+  mutate(rss = sum(res^2)) %>%
+  ungroup()
+```
+
+```{r}
+set.seed(2020)
+wages_aug %>%
+  sample_n_keys(12) %>%
+  ggplot(aes(x = year,
+             y = pred,
+             group = id)) +
+  geom_line(color = "steelblue") +
+  geom_point(aes(y = mean_hourly_wage)) +
+  facet_wrap(~id)
+```

### 2021-01-05 EDA_wages_hs2020.Rmd
Add relationship with Age,Gender, and Education

@@
 
-## Build GAM model
+## Build a GAM model
 
@@
 Since we have a quite large data set, we have to subset it first to model the data.
-Here, we would like to build a Generalized Additive Model (GAM) model, by randomly subset the daya the data using brolgar's` `sample_frac_keys` function.
+Here, we would like to build a Generalized Additive Model (GAM) model, by randomly subset the data using brolgar's` `sample_frac_keys` function.
+
+Here, we sampled 1 percent portion of the data.
 
@@
 ```{r}
-set.seed(2020)
+set.seed(2021)
 wages_aug %>%
@@
 ```
+
+## Relationship with Age, Gender, and Education
+
+**Age**
+
+```{r}
+fig_wages_all <- 
+  ggplot(wages, 
+         aes(x = year,
+             y = mean_hourly_wage,
+             group = id)) +
+  geom_line(alpha = 0.1)
+```
+
+```{r facet-age}
+
+fig_wages_all +
+  facet_wrap(~age_1979)
+```
+
+Learn: There is no specific pattern of age in 1979 to the wages. Most of the subjects have the same starting wages. Moreover, the wages at the end of the study period varied between the subjects.
+
+**Gender**
+
+```{r facet-gender}
+fig_wages_all +
+  facet_wrap(~gender)
+```
+Learn: Males tend to earn more wages than females.
+
+**Education**
+
+```{r facet-edu}
+fig_wages_all +
+  facet_wrap(~hgc)
+```
+Learn: Most of subject has finished their 12th grade. Subjects with lower education tend to have lower wages.

### 2021-01-11 anomaly.Rmd
Calculate the variance of linear model residual.

---
title: "Anomaly_inspection"
author: "Dewi Lestari Amaliah"
date: "09/01/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(brolgar)
```

```{r data}
load("wages_hs2020.rda")
```

```{r wages}
wages <- as_tsibble(x = wages_hs2020,
                    key = id,
                    index = year,
                    regular = FALSE)
```

```{r lm}
wages_lm <- key_slope(wages, 
                         mean_hourly_wage ~ year) %>%
  left_join(wages, by = "id")
```

```{r}
wages_lm <- wages_lm %>%
  mutate(pred = .intercept + (.slope_year*year))
```

```{r}
wages_res <- wages_lm %>%
  mutate(resid = mean_hourly_wage - pred)
```

```{r}
wages_lm_var <- wages_res %>%
  group_by(id) %>%
  summarise(var = var(mean_hourly_wage,
                      na.rm = TRUE)) %>%
  arrange(desc(var))
```

```{r}
ggplot(filter(wages, id == 8360)) +
  geom_line(aes(x = year,
                y = mean_hourly_wage))
```

### 2021-01-13 anomaly.Rmd
Add the plot of id with high and low variance of lm's residual

@@
 
+# build a linear model
 ```{r lm}
@@
 ```{r}
+
+#calculate the prediction value
 wages_lm <- wages_lm %>%
@@
 ```{r}
+#calculate the residual
 wages_res <- wages_lm %>%
@@
 ```{r}
+
+#calculate the variance of the lm model's residual 
+#sort it from the highest to the lowest
 wages_lm_var <- wages_res %>%
@@
 ```{r}
-ggplot(filter(wages, id == 8360)) +
+wages_lm_var <- wages_lm_var %>%
+  mutate(rank = seq(1 : nrow(wages_lm_var)))
+```
+
+### Plot the wages with high variance in their lm-residuals
+
+```{r}
+top12 <- wages_lm_var %>%
+  filter(rank <=25) %>%
+  left_join(wages, by = "id")
+
+ggplot(top12) +
   geom_line(aes(x = year,
-                y = mean_hourly_wage))
+                y = mean_hourly_wage)) +
+  facet_wrap(~id)
 ```
+
+### Plot the wages with low variance in their lm-residuals
+```{r}
+bottom12 <- wages_lm_var %>%
+  filter(rank >= 5995) %>%
+  left_join(wages, by = "id")
+
+ggplot(bottom12) +
+  geom_line(aes(x = year,
+                y = mean_hourly_wage)) +
+  facet_wrap(~id)
+```
+
+# take 12 random ids 
+```{r}
+set.seed(2021)
+
+#take a random sample
+
+random <- as.data.frame(sample(wages_lm_var$id, size = 30, replace = FALSE)) %>%
+  rename("id" = "sample(wages_lm_var$id, size = 30, replace = FALSE)") %>%
+  left_join(wages, by = "id")
+
+ggplot(random) +
+  geom_line(aes(x = year,
+                y = mean_hourly_wage)) +
+  facet_wrap(~id)
+
+```
+
+Learn: The observations with high and low variability were mostly the observations with the incomplete years of observations.
+Hence, we are going to try to only look the variance on the observations with the complete years.
+
+```{r}
+wages_obs <- wages %>%
+  features(mean_hourly_wage, n_obs) %>%
+  arrange(n_obs)
+
+complete_year_obs <- wages_obs %>%
+  filter(n_obs == 27) %>%
+  left_join(wages_lm_var, by = "id") %>%
+  arrange(desc(var))
+```
+
+```{r}
+
+top20_comp <- complete_year_obs %>%
+  slice_head(n =20) %>%
+  left_join(wages, by = "id")
+
+ggplot(top20_comp) +
+  geom_line(aes(x = year,
+                y = mean_hourly_wage)) +
+  facet_wrap(~id)
+
+```
+```{r}
+bottom20_comp <- complete_year_obs %>%
+  slice_tail(n =20) %>%
+  left_join(wages, by = "id")
+
+ggplot(bottom20_comp) +
+  geom_line(aes(x = year,
+                y = mean_hourly_wage)) +
+  facet_wrap(~id)
+```
+
+```{r}
+set.seed(2021)
+
+random2 <- as.data.frame(sample(complete_year_obs$id, size = 30, replace = FALSE)) %>%
+  rename("id" = "sample(complete_year_obs$id, size = 30, replace = FALSE)") %>%
+  left_join(wages, by = "id")
+
+ggplot(random2) +
+  geom_line(aes(x = year,
+                y = mean_hourly_wage)) +
+  facet_wrap(~id)
+```
+
+The volatility in wages, could be seen in the id with high and low variance. For example, in id 951 (high variance) and 7611 (small variance).

### 2021-01-16 anomaly.Rmd
Edit some codes and add boxplot to show extreme values

@@
   group_by(id) %>%
-  summarise(var = var(mean_hourly_wage,
+  summarise(var = var(resid,
                       na.rm = TRUE)) %>%
@@
   facet_wrap(~id)
+```
+
+```{r}
+wages %>%
+  filter(id == 11974)
 ```
@@
 
-The volatility in wages, could be seen in the id with high and low variance. For example, in id 951 (high variance) and 7611 (small variance).
+```{r}
+wages %>%
+  filter(year == 1984)
+```
+
+```{r}
+ggplot(wages) +
+  geom_boxplot(aes(x = mean_hourly_wage)) +
+  facet_wrap(~year)
+```

### 2021-01-18 anomaly.Rmd
Visualize the missing data using naniar package

@@
 library(brolgar)
+library(naniar)
 ```
@@
 
-ggplot(top12) +
-  geom_line(aes(x = year,
+ggplot(top12, 
+       aes(x = year,
            y = mean_hourly_wage)) +
+  geom_miss_point() +
+  geom_line() +
   facet_wrap(~id)
@@
 
-ggplot(bottom12) +
-  geom_line(aes(x = year,
+ggplot(bottom12,
+       aes(x = year,
            y = mean_hourly_wage)) +
+  geom_miss_point() +
+  geom_line() +
   facet_wrap(~id)
@@
 
-ggplot(random) +
-  geom_line(aes(x = year,
+ggplot(random, 
+       aes(x = year,
                 y = mean_hourly_wage)) +
+  geom_miss_point() +
+  geom_line() +
   facet_wrap(~id)
@@
 
-ggplot(top20_comp) +
-  geom_line(aes(x = year,
+ggplot(top20_comp,
+       aes(x = year,
                 y = mean_hourly_wage)) +
+  geom_miss_point() +
+  geom_line()+
   facet_wrap(~id)
@@
 
-ggplot(bottom20_comp) +
-  geom_line(aes(x = year,
+ggplot(bottom20_comp,
+       aes(x = year,
                 y = mean_hourly_wage)) +
+  geom_miss_point() +
+  geom_line() +
   facet_wrap(~id)
@@
 
-ggplot(random2) +
-  geom_line(aes(x = year,
+ggplot(random2,
+       aes(x = year,
                 y = mean_hourly_wage)) +
+  geom_miss_point() +
+  geom_line() +
   facet_wrap(~id)

### 2021-01-20 School_drop_out.Rmd
Compare hs dropout with other groups.

---
title: "School_drop_out"
author: "Dewi Lestari Amaliah"
date: "13/01/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(brolgar)
library(patchwork)
library(gghighlight)
```

## filter the observations who are the high school drop out

```{r data}
load("wages_hs2020.rda")

wages <- as_tsibble(x = wages_hs2020,
                    key = id,
                    index = year,
                    regular = FALSE)

wages_do <- wages %>%
  filter(hgc %in% c("9TH GRADE", 
                    "10TH GRADE",
                    "11TH GRADE"))
```

```{r}

wages_do_fig <- ggplot(wages_do, 
                       aes(x = year,
                           y = mean_hourly_wage,
                           group = id)) +
  geom_line(alpha = 0.1)

wages_do_fig

```

```{r}
wages_do_feature<- wages_do %>%
  features(mean_hourly_wage, 
           feat_three_num
           )
```

```{r}
min <- ggplot(wages_do_feature) +
  geom_density(aes(x = min), fill = "magenta", alpha = 0.5)
```

```{r}
wages_do_feature_long <- wages_do_feature %>%
  pivot_longer(c(min, med, max), names_to = "feature", values_to = "value")

feature_fig <- ggplot(wages_do_feature_long) +
  geom_density(aes(x = value, colour = feature, fill = feature), alpha = 0.3) +
  theme(legend.position = "None")
```

```{r}
feature_spaghetty <- ggplot(transform(wages_do_feature_long,
                 feature = factor(feature, level = 
                                 c("min",
                                   "med", 
                                   "max")))) +
  geom_line(aes(x = feature,
                y = value, 
                group = id), alpha = 0.5)
```

```{r}
min+feature_fig+feature_spaghetty
```

Compared to overall cohort the range of the minimum wages of high school dropouts is smaller. The median of maximum wages in high school dropouts is smaller compared to the rest of the cohort. It is also more positively skewed. These mean that the high school dropouts generally received smaller wages then the rest of the cohort, which from the previous EDA, we found that most of the observations completed high school. Hence, the reason why the rest of the cohort has bigger wages is because most of the observation has higher education.

```{r}
ggplot(wages_do, 
       aes(x = mean_hourly_wage)) +
  geom_boxplot() +
  facet_wrap(~year)
```

```{r}
wages_do_slope <- key_slope(wages_do, 
                         mean_hourly_wage ~ year) %>%
  left_join(wages_do, by = "id")

wages_do_slope
```

```{r}
ggplot(wages_do_slope) +
  geom_line(aes(x = year,
                y = mean_hourly_wage,
                group = id)) +
  gghighlight(.slope_year < 0)
```

## filter the observations who completed high school

```{r}
wages_hs <- wages %>%
  filter(hgc == "12TH GRADE")
```

```{r}
wages_hs_fig <- ggplot(wages_hs, 
                       aes(x = year,
                           y = mean_hourly_wage,
                           group = id)) +
  geom_line(alpha = 0.1)

wages_hs_fig
```

```{r}
wages_hs_feature<- wages_hs %>%
  features(mean_hourly_wage, 
           feat_three_num
           )
```

```{r}
min_hs <- ggplot(wages_hs_feature) +
  geom_density(aes(x = min), fill = "magenta", alpha = 0.5)
```

```{r}
wages_hs_feature_long <- wages_hs_feature %>%
  pivot_longer(c(min, med, max), names_to = "feature", values_to = "value")

feature_fig_hs <- ggplot(wages_hs_feature_long) +
  geom_density(aes(x = value, colour = feature, fill = feature), alpha = 0.3) +
  theme(legend.position = "None")
```

```{r}
feature_spaghetty_hs <- ggplot(transform(wages_hs_feature_long,
                 feature = factor(feature, level = 
                                 c("min",
                                   "med", 
                                   "max")))) +
  geom_line(aes(x = feature,
                y = value, 
                group = id), alpha = 0.5)
```

```{r}
min_hs + feature_fig_hs + feature_spaghetty_hs
```

```{r}
ggplot(wages_hs, 
       aes(x = mean_hourly_wage)) +
  geom_boxplot() +
  facet_wrap(~year)
```

## Middle School

```{r}

wages_ms <- wages %>%
  filter(hgc %in% c("6TH GRADE", 
                    "7TH GRADE",
                    "8TH GRADE"))

```

```{r}
wages_ms_fig <- ggplot(wages_ms, 
                       aes(x = year,
                           y = mean_hourly_wage,
                           group = id)) +
  geom_line(alpha = 0.1)

wages_ms_fig
```

```{r}
wages_do_fig + wages_hs_fig + wages_ms_fig
```

### 2021-01-21 code_revisit2.Rmd
Tyding the code and get the raw df of rates and hours

---
title: "code_revisit_2"
author: "Dewi Lestari Amaliah"
date: "21/01/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(here)
library(readr)
library(tidyr)
library(tidyverse)
library(brolgar)
```

```{r}

# load the downloaded data 
new_data_qnames <- read_rds(here::here("data-raw/wages-high-school-demo/data-frames/wages-high-school-demo-new_data_qnames.rds"))
categories_qnames <- read_rds(here::here("data-raw/wages-high-school-demo/data-frames/wages-high-school-demo-categories_qnames.rds"))

```

## get hours of work

```{r}

year_A <- c(1979:1987, 1993)
get_hour <- function(year){
  if(year %in% year_A){
   new_data_qnames %>%
    select(CASEID_1979,
            starts_with("QES-52A") &
              ends_with(as.character(year))) %>%
    pivot_longer(!CASEID_1979,
                 names_to = "job",
                 values_to = "hours_work") %>%
    separate("job", c("job", "year"), sep = -4) %>%
    mutate(job = paste0("job_", substr(job, 9, 10))) %>%
    rename(id = CASEID_1979)}
  else{
    new_data_qnames %>%
    select(CASEID_1979,
            starts_with("QES-52D") &
              ends_with(as.character(year))) %>%
    pivot_longer(!CASEID_1979,
                 names_to = "job",
                 values_to = "hours_work") %>%
    separate("job", c("job", "year"), sep = -4) %>%
    mutate(job = paste0("job_", substr(job, 9, 10))) %>%
    rename(id = CASEID_1979)
  }
 }
```

```{r}

hours <- list()

for(ayear in c(1979:1994, 1996, 1998, 2000, 2002, 2004, 2006, 2008, 2010, 
               2012, 2014, 2016)) {
   hours[[ayear]] <- get_hour(ayear)
}

hours_all <- bind_rows(!!!hours)

```

## get rates

```{r}
get_rate <- function(year) {
  new_data_qnames %>%
     select(CASEID_1979,
            starts_with("HRP") &
              ends_with(as.character(year))) %>%
    pivot_longer(!CASEID_1979, names_to = "job", values_to = "rate_per_hour") %>%
    separate("job", c("job", "year"), sep = -4) %>%
    mutate(job = paste0("job_0", substr(job, 4, 4))) %>%
    rename(id = CASEID_1979)
}
```

```{r}
rates <- list()

for(ayear in c(1979:1994, 1996, 1998, 2000, 2002, 2004, 2006, 2008, 2010, 
               2012, 2014, 2016)) {
   rates[[ayear]] <- get_rate(ayear)
}

rates_all <- bind_rows(!!!rates)
```

## join hours and rates

```{r}

hours_wages <- left_join(rates_all, 
                         hours_all, 
                         by = c("id", "year", "job"))

```

## filter how many observations that having rates but "NA" in hours worked

```{r}
na_hours <- hours_wages %>%
  filter(!is.na(rate_per_hour) & is.na(hours_work))

percentage_na_hours <- nrow(na_hours)/ nrow(hours_wages) * 100
```

### 2021-01-21 code_revisit2.Rmd
Calculating the weighted average of wages scenario 1 and 2.

@@
 library(brolgar)
+library(patchwork)
+library(naniar)
 ```
@@
                          by = c("id", "year", "job"))
-
 ```
@@
 percentage_na_hours <- nrow(na_hours)/ nrow(hours_wages) * 100
+percentage_na_hours
 ```
+
+There are `r round(percentage_na_hours, 2)` percent of data that missing in hours worked, but not missing in wages.
+
+## filter how many observations that "having"NA" in rates, but present in hours worked
+
+```{r}
+na_rates <- hours_wages %>%
+  filter(is.na(rate_per_hour) & !is.na(hours_work))
+
+percentage_na_rates <- nrow(na_rates)/ nrow(hours_wages) * 100
+percentage_na_rates
+
+```
+
+# calculate the wages (weighted mean approach)
+
+## Scenario 1 
+- For the obs with more than one jobs, and one of the job has na in hours_work, we can not calculate the weighted average.
+- For the obs with only one job and the hours_work is na, we only will use the rates directly.
+- Calculate the weighted hours normally
+
+```{r}
+no_job <- hours_wages %>%
+  filter(!is.na(rate_per_hour)) %>%
+  group_by(id, year) %>%
+  summarise(no_jobs = n_distinct(rate_per_hour))
+
+eligible_wages <- hours_wages %>%
+  filter(!is.na(rate_per_hour)) %>%
+  left_join(no_job, by = c("id", "year"))
+```
+
+```{r}
+wages_weighted_sc1 <- 
+  eligible_wages %>%
+  group_by(id, year) %>%
+  mutate(wages = ifelse(no_jobs == 1, rate_per_hour/100,
+                        weighted.mean(rate_per_hour, hours_work, na.rm = TRUE)/100))
+
+visdat::vis_dat(wages_weighted_sc1, warn_large_data = FALSE)
+
+```
+
+Note: We lose the information in many observations with half complete of hours_work. Example:
+observation 506.
+
+```{r}
+wages_weighted %>% filter(id == 506)
+```
+```{r}
+wages_tidy_sc1 <- wages_weighted_sc1 %>%
+  filter(!is.na(wages)) %>%
+  group_by(id, year) %>%
+  summarise(wages = mean(wages)) %>%
+  mutate(year = as.numeric(year)) %>%
+  ungroup() %>%
+  mutate(wages = ifelse(wages > 300,
+                        NA, wages))
+```
+
+```{r}
+wages_sc1 <- as_tsibble(x = wages_tidy_sc1,
+                    key = id,
+                    index = year,
+                    regular = FALSE)
+
+ggplot(wages_sc1) +
+  geom_line(aes(x = year,
+                y = wages, 
+                group = id), 
+            alpha = 0.1)
+
+```
+
+```{r}
+set.seed(2020)
+ggplot(wages_sc1) +
+  geom_line(aes(x = year,
+                y = wages,
+                group = id)) +
+  facet_sample()
+```
+```{r features}
+wages_feature_sc1 <- wages_sc1 %>%
+  features(wages, 
+           feat_three_num
+           )
+```
+
+```{r p1}
+p1 <- ggplot(wages_feature_sc1) +
+  geom_density(aes(x = min), fill = "magenta", alpha = 0.5)
+```
+
+```{r p2}
+
+wages_feature_long_sc1 <- wages_feature_sc1 %>%
+  pivot_longer(c(min, med, max), names_to = "feature", values_to = "value")
+
+p2 <- ggplot(wages_feature_long_sc1) +
+  geom_density(aes(x = value, colour = feature, fill = feature), alpha = 0.3) +
+  theme(legend.position = "None")
+```
+
+```{r p3}
+p3 <- ggplot(transform(wages_feature_long_sc1,
+                 feature = factor(feature, level = 
+                                 c("min",
+                                   "med", 
+                                   "max")))) +
+  geom_line(aes(x = feature,
+                y = value, 
+                group = id), alpha = 0.5)
+```
+
+```{r feature-charts}
+
+p1+p2+p3
+
+```
+
+## SCENARIO 2
+So, let us try scenario 2, by exclude the observation with more than one job with NA in hours_work.
+
+```{r}
+wages_weighted_sc2 <- eligible_wages %>%
+  filter(no_jobs == 1 |
+           (no_jobs > 1 & !is.na(hours_work))) %>%
+  group_by(id, year) %>%
+  mutate(wages = ifelse(no_jobs == 1, rate_per_hour/100,
+                        weighted.mean(rate_per_hour, hours_work, na.rm = TRUE)/100))
+
+visdat::vis_dat(wages_weighted_sc2, warn_large_data = FALSE)
+
+```
+```{r}
+wages_tidy_sc2 <- wages_weighted_sc2 %>%
+  group_by(id, year) %>%
+  summarise(wages = mean(wages)) %>%
+  mutate(year = as.numeric(year)) %>%
+  ungroup() %>%
+  mutate(wages = ifelse(wages > 300,
+                        NA, wages))
+```
+
+```{r}
+wages <- as_tsibble(x = wages_tidy_sc2,
+                    key = id,
+                    index = year,
+                    regular = FALSE)
+```
+
+```{r}
+ggplot(wages) +
+  geom_line(aes(x = year,
+                y = wages,
+                group = id), 
+            alpha = 0.1)
+```
+
+```{r}
+set.seed(2021)
+ggplot(wages, aes(x = year,
+                y = wages,
+                group = id)) +
+  geom_line() +
+  facet_sample()
+```
+
+```{r features}
+wages_feature<- wages %>%
+  features(wages, 
+           feat_three_num
+           )
+```
+
+```{r p1}
+p4 <- ggplot(wages_feature) +
+  geom_density(aes(x = min), fill = "magenta", alpha = 0.5)
+```
+
+```{r p2}
+
+wages_feature_long <- wages_feature %>%
+  pivot_longer(c(min, med, max), names_to = "feature", values_to = "value")
+
+p5 <- ggplot(wages_feature_long) +
+  geom_density(aes(x = value, colour = feature, fill = feature), alpha = 0.3) +
+  theme(legend.position = "None")
+```
+
+```{r p3}
+p6 <- ggplot(transform(wages_feature_long,
+                 feature = factor(feature, level = 
+                                 c("min",
+                                   "med", 
+                                   "max")))) +
+  geom_line(aes(x = feature,
+                y = value, 
+                group = id), alpha = 0.5)
+```
+
+```{r feature-charts}
+
+p4+p5+p6
+
+```

### 2021-01-27 code_revisit2.Rmd
Add the summary of scenario 1 and 2

@@
 
-## Scenario 1 
+## SCENARIO 1 
 - For the obs with more than one jobs, and one of the job has na in hours_work, we can not calculate the weighted average.
@@
 ```{r}
+
+# calculate number of jobs that a person has in one year
 no_job <- hours_wages %>%
@@
 
+# filter the observations with available rate per hour
 eligible_wages <- hours_wages %>%
@@
 ```{r}
-wages_weighted %>% filter(id == 506)
+wages_weighted_sc1 %>% filter(id == 506)
 ```
@@
   group_by(id, year) %>%
-  summarise(wages = mean(wages)) %>%
+  summarise(wages = mean(wages),
+            total_hours = sum(hours_work),
+            number_of_jobs = mean(no_jobs)) %>%
   mutate(year = as.numeric(year)) %>%
   ungroup() %>%
-  mutate(wages = ifelse(wages > 300,
+  mutate(wages = ifelse(wages > 30,
                         NA, wages))
@@
 
-ggplot(wages_sc1) +
+sc1_plot <- ggplot(wages_sc1) +
   geom_line(aes(x = year,
@@
                 group = id), 
-            alpha = 0.1)
+            alpha = 0.1) +
+  ggtitle("Scenario 1")
+
+sc1_plot
 
@@
   facet_sample()
+
 ```
@@
 
-p1+p2+p3
+sc1 <- p1+p2+p3
+sc1
 
@@
 ## SCENARIO 2
+
 So, let us try scenario 2, by exclude the observation with more than one job with NA in hours_work.
@@
 ```{r}
+
+wages_weighted_sc2 %>% filter(id == 506)
+```
+
+```{r}
 wages_tidy_sc2 <- wages_weighted_sc2 %>%
   group_by(id, year) %>%
-  summarise(wages = mean(wages)) %>%
+  summarise(wages = mean(wages),
+            total_hours = sum(hours_work),
+            number_of_jobs = mean(no_jobs)) %>%
   mutate(year = as.numeric(year)) %>%
   ungroup() %>%
-  mutate(wages = ifelse(wages > 300,
+  mutate(wages = ifelse(wages > 30,
                         NA, wages))
@@
 ```{r}
-wages <- as_tsibble(x = wages_tidy_sc2,
+wages_sc2 <- as_tsibble(x = wages_tidy_sc2,
                     key = id,
@@
 ```{r}
-ggplot(wages) +
+sc2_plot <- ggplot(wages_sc2) +
   geom_line(aes(x = year,
@@
                 group = id), 
-            alpha = 0.1)
+            alpha = 0.1) +
+  ggtitle("Scenario 2")
 ```
@@
 set.seed(2021)
-ggplot(wages, aes(x = year,
+ggplot(wages_sc2, aes(x = year,
                 y = wages,
@@
 ```{r features}
-wages_feature<- wages %>%
+wages_feature_sc2<- wages_sc2 %>%
   features(wages, 
@@
 ```{r p1}
-p4 <- ggplot(wages_feature) +
+p4 <- ggplot(wages_feature_sc2) +
   geom_density(aes(x = min), fill = "magenta", alpha = 0.5)
@@
 
-wages_feature_long <- wages_feature %>%
+wages_feature_long_sc2 <- wages_feature_sc2 %>%
   pivot_longer(c(min, med, max), names_to = "feature", values_to = "value")
 
-p5 <- ggplot(wages_feature_long) +
+p5 <- ggplot(wages_feature_long_sc2) +
   geom_density(aes(x = value, colour = feature, fill = feature), alpha = 0.3) +
@@
 ```{r p3}
-p6 <- ggplot(transform(wages_feature_long,
+p6 <- ggplot(transform(wages_feature_long_sc2,
                  feature = factor(feature, level = 
@@
 
-p4+p5+p6
+sc2 <- p4+p5+p6
+sc2
 
 ```
+
+## SCENARIO 3 (modified SC1)
+
+In this scenario, we treated the observation with the wages higher than $100 as NA
+
+```{r}
+wages_tidy_sc3 <- wages_weighted_sc1 %>%
+  filter(!is.na(wages)) %>%
+  group_by(id, year) %>%
+  summarise(wages = mean(wages)) %>%
+  mutate(year = as.numeric(year)) %>%
+  ungroup() %>%
+  mutate(wages = ifelse(wages > 100,
+                        NA, wages))
+```
+
+```{r}
+wages_sc3 <- as_tsibble(x = wages_tidy_sc3,
+                    key = id,
+                    index = year,
+                    regular = FALSE)
+```
+
+```{r}
+ggplot(wages_sc3) +
+  geom_line(aes(x = year,
+                y = wages,
+                group = id), 
+            alpha = 0.1)
+```
+
+```{r}
+set.seed(2021)
+ggplot(wages_sc3, aes(x = year,
+                y = wages,
+                group = id)) +
+  geom_line() +
+  facet_sample()
+```
+
+## SCENARIO 4 (modified SC2)
+
+```{r}
+wages_tidy_sc4 <- wages_weighted_sc2 %>%
+  group_by(id, year) %>%
+  summarise(wages = mean(wages)) %>%
+  mutate(year = as.numeric(year)) %>%
+  ungroup() %>%
+  mutate(wages = ifelse(wages > 100,
+                        NA, wages))
+```
+
+```{r}
+wages_sc4 <- as_tsibble(x = wages_tidy_sc4,
+                    key = id,
+                    index = year,
+                    regular = FALSE)
+```
+
+```{r}
+ggplot(wages_sc4) +
+  geom_line(aes(x = year,
+                y = wages,
+                group = id), 
+            alpha = 0.1)
+```
+
+```{r}
+set.seed(2021)
+ggplot(wages_sc4, aes(x = year,
+                y = wages,
+                group = id)) +
+  geom_line() +
+  facet_sample()
+```
+
+## SUMMARY
+
+### The original tidy wages
+
+```{r}
+ori <- read_csv("data-raw/wages-high-school-demo/data-frames/tidy_employment.csv") %>%
+  mutate(mean_hourly_wage = ifelse( mean_hourly_wage > 30,
+                        NA, mean_hourly_wage))
+```
+
+```{r}
+fig_all_transp <- ggplot(ori) +
+  geom_line(aes(x = year,
+                y = mean_hourly_wage,
+                group = CASEID_1979),
+            alpha = 0.1) +
+  ggtitle("Original data")
+```
+
+```{r wages_tsibble}
+wages_ori <- as_tsibble(x = ori,
+                    key = CASEID_1979,
+                    index = year,
+                    regular = FALSE)
+```
+
+```{r}
+wages_obs <- wages_ori %>%
+  features(mean_hourly_wage, n_obs) %>%
+  arrange(n_obs)
+```
+
+### sc1 + sc2 + original tidy employment data
+
+```{r}
+sc1_plot + sc2_plot + fig_all_transp
+```
+
+Number of observations
+
+```{r}
+nrow(wages_ori) #regular mean
+nrow(wages_sc1) #weighted mean omit id with NA in the working hours
+nrow(wages_sc2) #weighted mean omit obs with NA in the working hours
+```
+
+Number of ids
+
+```{r}
+nrow(wages_obs)
+nrow(wages_feature_sc1)
+nrow(wages_feature_sc2)
+```

### 2021-01-27 code_di_SC1.R
Rebuild the wages_demo_hs using sc1

@@ -1,0 +1,67 @@
+library(tidyverse)
+
+wages <- read_csv("data-raw/wages-high-school-demo/data-frames/tidy_employment_sc1.csv")
+demog <- read_csv("data-raw/wages-high-school-demo/data-frames/tidy_demographics.csv")
+
+wages_demog <- left_join(wages, demog, by="CASEID_1979")
+
+wages_demog %>%
+  filter(`highest-grade` %in% c("10TH GRADE", "11TH GRADE", "12TH GRADE")) %>%
+  ggplot(aes(x=year, y=mean_hourly_wage))  +
+  geom_point() + facet_wrap(~`highest-grade`)
+# Look like some issues with wages > 10000 per hour
+# Need to organsie education
+wages_demog %>% #count(gender)
+  #filter(`highest-grade` %in% c("10TH GRADE", "11TH GRADE", "12TH GRADE")) %>%
+  ggplot(aes(x=year, y=mean_hourly_wage))  +
+  geom_point() + facet_wrap(~gender)
+
+wages_demog <- wages_demog %>%
+  mutate(year = as.numeric(year),
+         year_completed_highest_grade = as.numeric(year_completed_highest_grade)) %>%
+  mutate(years_in_workforce = year - year_completed_highest_grade) %>%
+  mutate(age_1979 = 1979 - (dob_year + 1900))
+
+wages_demog %>% filter(years_in_workforce < 0) %>% tally()
+
+wages_demog %>% filter(years_in_workforce < 0) %>%
+  count(CASEID_1979) %>% tally()
+wages_demog %>%
+  ggplot(aes(x=years_in_workforce, y=number_of_jobs)) +
+  geom_point()
+wages_demog_hs <- wages_demog  %>% filter(grepl("GRADE", `highest-grade`))
+wages_demog_hs %>% filter(years_in_workforce < 0) %>% tally()
+wages_demog_hs %>% filter(years_in_workforce < -35) %>%
+  select(CASEID_1979, year, year_completed_highest_grade, `highest-grade`, dob_year, number_of_jobs, mean_hourly_wage)
+wages_demog_hs %>% filter(CASEID_1979 == "4396") %>% print(n=50)
+
+wages_demog %>% ggplot(aes(x=age_1979)) + geom_bar()
+summary(wages_demog$mean_hourly_wage)
[…]

### 2021-01-27 code_di_SC2.R
Rebuild wages_demo_hs using SC2

@@ -1,0 +1,67 @@
+library(tidyverse)
+
+wages <- read_csv("data-raw/wages-high-school-demo/data-frames/tidy_employment_sc2.csv")
+demog <- read_csv("data-raw/wages-high-school-demo/data-frames/tidy_demographics.csv")
+
+wages_demog <- left_join(wages, demog, by="CASEID_1979")
+
+wages_demog %>%
+  filter(`highest-grade` %in% c("10TH GRADE", "11TH GRADE", "12TH GRADE")) %>%
+  ggplot(aes(x=year, y=mean_hourly_wage))  +
+  geom_point() + facet_wrap(~`highest-grade`)
+# Look like some issues with wages > 10000 per hour
+# Need to organsie education
+wages_demog %>% #count(gender)
+  #filter(`highest-grade` %in% c("10TH GRADE", "11TH GRADE", "12TH GRADE")) %>%
+  ggplot(aes(x=year, y=mean_hourly_wage))  +
+  geom_point() + facet_wrap(~gender)
+
+wages_demog <- wages_demog %>%
+  mutate(year = as.numeric(year),
+         year_completed_highest_grade = as.numeric(year_completed_highest_grade)) %>%
+  mutate(years_in_workforce = year - year_completed_highest_grade) %>%
+  mutate(age_1979 = 1979 - (dob_year + 1900))
+
+wages_demog %>% filter(years_in_workforce < 0) %>% tally()
+
+wages_demog %>% filter(years_in_workforce < 0) %>%
+  count(CASEID_1979) %>% tally()
+wages_demog %>%
+  ggplot(aes(x=years_in_workforce, y=number_of_jobs)) +
+  geom_point()
+wages_demog_hs <- wages_demog  %>% filter(grepl("GRADE", `highest-grade`))
+wages_demog_hs %>% filter(years_in_workforce < 0) %>% tally()
+wages_demog_hs %>% filter(years_in_workforce < -35) %>%
+  select(CASEID_1979, year, year_completed_highest_grade, `highest-grade`, dob_year, number_of_jobs, mean_hourly_wage)
+wages_demog_hs %>% filter(CASEID_1979 == "4396") %>% print(n=50)
+
+wages_demog %>% ggplot(aes(x=age_1979)) + geom_bar()
+summary(wages_demog$mean_hourly_wage)
[…]

### 2021-01-27 comparing_scenario.Rmd
Summarising the scenarios

---
title: "comparing_scenario"
author: "Dewi Lestari Amaliah"
date: "27/01/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(brolgar)
library(patchwork)
```

```{r}
load("wages_hs2020.rda")
load("wages_hs2020_sc1.rda")
load("wages_hs2020_sc2.rda")
```

```{r}
bp_ori<- ggplot(wages_hs2020) +
  geom_boxplot(aes(x = mean_hourly_wage)) +
  facet_wrap(~year) +
  ggtitle("Original")

bp_sc1 <- ggplot(wages_hs2020_sc1) +
  geom_boxplot(aes(x = mean_hourly_wage)) +
  facet_wrap(~year) +
  ggtitle("Scenario 1")

bp_sc2 <- ggplot(wages_hs2020_sc2) +
  geom_boxplot(aes(x = mean_hourly_wage)) +
  facet_wrap(~year) +
  ggtitle("Scenario 2")

bp_ori
bp_sc1
bp_sc2

```

```{r}
wages_ori <- as_tsibble(x = wages_hs2020,
                    key = id,
                    index = year,
                    regular = FALSE)

wages_sc1 <- as_tsibble(x = wages_hs2020_sc1,
                    key = id,
                    index = year,
                    regular = FALSE)

wages_sc2 <- as_tsibble(x = wages_hs2020_sc2,
                    key = id,
                    index = year,
                    regular = FALSE)
```

```{r}
lp_ori <- ggplot(wages_ori) +
  geom_line(aes(x = year,
                y = mean_hourly_wage,
                group = id),
            alpha = 0.1) +
  ggtitle("Original")

lp_sc1 <- ggplot(wages_sc1) +
  geom_line(aes(x = year,
                y = mean_hourly_wage,
                group = id),
            alpha = 0.1) +
  ggtitle("Scenario 1")

lp_sc2 <- ggplot(wages_sc2) +
  geom_line(aes(x = year,
                y = mean_hourly_wage,
                group = id),
            alpha = 0.1) +
  ggtitle("Scenario 2")

lp_ori + lp_sc1 + lp_sc2

```
```{r}

get_obs <- function(data){
  data %>%
  features(mean_hourly_wage, n_obs) %>%
  arrange(n_obs)
  }

wages_obs_ori <- get_obs(wages_ori)
wages_obs_sc1 <- get_obs(wages_sc1)
wages_obs_sc2 <- get_obs(wages_sc2)

nrow(wages_obs_ori)
nrow(wages_obs_sc1)
nrow(wages_obs_sc2)
```

```{r}
summary(wages_ori$mean_hourly_wage)
```

```{r}
summary(wages_sc1$mean_hourly_wage)
```

```{r}
summary(wages_sc2$mean_hourly_wage)
```

### 2021-01-29 New_Wages.Rmd
Edit the calculation of wage

---
title: "New_Wages"
author: "Dewi Lestari Amaliah"
date: "29/01/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(here)
library(readr)
library(tidyr)
library(tidyverse)
library(brolgar)
library(patchwork)
library(naniar)
```

```{r}

# load the downloaded data 
new_data_qnames <- read_rds(here::here("data-raw/wages-high-school-demo/data-frames/wages-high-school-demo-new_data_qnames.rds"))
categories_qnames <- read_rds(here::here("data-raw/wages-high-school-demo/data-frames/wages-high-school-demo-categories_qnames.rds"))

```

```{r}

year_A <- c(1979:1987, 1993)

get_hour <- function(year){
  if(year %in% year_A){
   new_data_qnames %>%
    select(CASEID_1979,
            starts_with("QES-52A") &
              ends_with(as.character(year))) %>%
    pivot_longer(!CASEID_1979,
                 names_to = "job",
                 values_to = "hours_work") %>%
    separate("job", c("job", "year"), sep = -4) %>%
    mutate(job = paste0("job_", substr(job, 9, 10))) %>%
    rename(id = CASEID_1979)}
  else{
    new_data_qnames %>%
    select(CASEID_1979,
            starts_with("QES-52D") &
              ends_with(as.character(year))) %>%
    pivot_longer(!CASEID_1979,
                 names_to = "job",
                 values_to = "hours_work") %>%
    separate("job", c("job", "year"), sep = -4) %>%
    mutate(job = paste0("job_", substr(job, 9, 10))) %>%
    rename(id = CASEID_1979)
  }
 }
```

```{r}

hours <- list()

for(ayear in c(1979:1994, 1996, 1998, 2000, 2002, 2004, 2006, 2008, 2010, 
               2012, 2014, 2016)) {
   hours[[ayear]] <- get_hour(ayear)
}

hours_all <- bind_rows(!!!hours)

```

## get rates

```{r}
get_rate <- function(year) {
  new_data_qnames %>%
     select(CASEID_1979,
            starts_with("HRP") &
              ends_with(as.character(year))) %>%
    pivot_longer(!CASEID_1979, names_to = "job", values_to = "rate_per_hour") %>%
    separate("job", c("job", "year"), sep = -4) %>%
    mutate(job = paste0("job_0", substr(job, 4, 4))) %>%
    rename(id = CASEID_1979)
}
```

```{r}
rates <- list()

for(ayear in c(1979:1994, 1996, 1998, 2000, 2002, 2004, 2006, 2008, 2010, 
               2012, 2014, 2016)) {
   rates[[ayear]] <- get_rate(ayear)
}

rates_all <- bind_rows(!!!rates)
```

## join hours and rates

```{r}

hours_wages <- left_join(rates_all, 
                         hours_all, 
                         by = c("id", "year", "job"))
```

```{r}

# calculate number of jobs that a person has in one year
no_job <- hours_wages %>%
  filter(!is.na(rate_per_hour)) %>%
  group_by(id, year) %>%
  summarise(no_jobs = length(rate_per_hour))

# filter the observations with available rate per hour
eligible_wages <- hours_wages %>%
  filter(!is.na(rate_per_hour)) %>%
  left_join(no_job, by = c("id", "year"))

# calcultae the mean_hourly_wage

# flag1 = code 1 for weighted mean
# code 0 for arithmatic mean

mean_hourly_wage <- 
  eligible_wages %>%
  group_by(id, year) %>%
  mutate(wages = ifelse(no_jobs == 1, rate_per_hour/100,
                        weighted.mean(rate_per_hour, hours_work, na.rm = TRUE)/100)) %>%
  mutate(flag1 = ifelse(!is.na(wages) & no_jobs != 1, 1,
                        0)) %>%
  mutate(wages = ifelse(is.na(wages), mean(rate_per_hour)/100,
                        wages))

mean_hourly_wage %>% filter(id == 506)

```
```{r}
visdat::vis_dat(mean_hourly_wage, warn_large_data = FALSE)
```
```{r}
mhw_to_write <- mean_hourly_wage %>%
  group_by(id, year) %>%
  summarise(wages = mean(wages),
            total_hours = sum(hours_work),
            number_of_jobs = mean(no_jobs),
            flag1 = mean(flag1)) %>%
  mutate(year = as.numeric(year)) %>%
  ungroup() %>%
  rename(mean_hourly_wage = wages,
         CASEID_1979 = id) %>%
  mutate(flag1 = ifelse(flag1 == 1, "wm",
                        "am"))
```

```{r}
summary(mhw_to_write$mean_hourly_wage)
```

```{r}
ggplot(mhw_to_write) +
  geom_boxplot(aes(x = mean_hourly_wage)) +
  facet_wrap(~year)
```
```{r}
#write.csv(mhw_to_write, here::here("data-raw/wages-high-school-demo/data-frames/tidy_employment_weighted.csv"))
```

### 2021-01-29 code_di_weighted.R
Edit the building of tidy_wage_hs data

@@ -1,0 +1,39 @@
+library(tidyverse)
+
+wages <- read_csv("data-raw/wages-high-school-demo/data-frames/tidy_employment_weighted.csv")
+demog <- read_csv("data-raw/wages-high-school-demo/data-frames/tidy_demographics.csv")
+
+d <- read_csv("data-raw/wages-high-school-demo/data-frames/tidy_employment.csv")
+
+wages_demog <- left_join(wages, demog, by="CASEID_1979")
+
+wages_demog <- wages_demog %>%
+  mutate(year_completed_highest_grade = as.numeric(year_completed_highest_grade)) %>%
+  mutate(years_in_workforce = year - year_completed_highest_grade) %>%
+  mutate(age_1979 = 1979 - (dob_year + 1900))
+
+# filter only the id with high school education
+wages_demog_hs <- wages_demog  %>% filter(grepl("GRADE", `highest-grade`))
+
+# replace the obs with mean_hourly_wage > 30 to be NA
+wages_demog_hs <- wages_demog_hs %>%
+  mutate(mean_hourly_wage = ifelse(mean_hourly_wage > 30, NA, mean_hourly_wage))
+
+# calculate the number of observation
+keep_me <- wages_demog_hs %>% count(CASEID_1979)
+keep_me %>% ggplot(aes(x=n)) + geom_bar()
+
+# filter out the id with low number of observation
+keep_me <- keep_me %>% filter(n > 4)
+
+wages_demog_hs <- wages_demog_hs %>%
+  filter(CASEID_1979 %in% keep_me$CASEID_1979)
+
+
+wages_hs2020_weighted <- wages_demog_hs %>%
+  rename(id=CASEID_1979, hgc=`highest-grade`,
+         hgc_i=`highest-grade-int`, yr_hgc=year_completed_highest_grade) %>%
+  select(id, year, mean_hourly_wage, age_1979, gender, race, hgc, hgc_i, yr_hgc, number_of_jobs, total_hours, flag1)
+
+
+save(wages_hs2020_weighted, file="wages_hs2020_weighted.rda")

### 2021-01-29 comparing_scenario.Rmd
Comparing the original vs weighted data

@@
 load("wages_hs2020.rda")
-load("wages_hs2020_sc1.rda")
-load("wages_hs2020_sc2.rda")
+load("wages_hs2020_weighted.rda")
 ```
@@
 
-bp_sc1 <- ggplot(wages_hs2020_sc1) +
+bp_wei <- ggplot(wages_hs2020_weighted) +
   geom_boxplot(aes(x = mean_hourly_wage)) +
   facet_wrap(~year) +
-  ggtitle("Scenario 1")
-
-bp_sc2 <- ggplot(wages_hs2020_sc2) +
-  geom_boxplot(aes(x = mean_hourly_wage)) +
-  facet_wrap(~year) +
-  ggtitle("Scenario 2")
+  ggtitle("Weighted")
 
 bp_ori
-bp_sc1
-bp_sc2
+bp_wei
 
@@
 wages_sc2 <- as_tsibble(x = wages_hs2020_sc2,
+                    key = id,
+                    index = year,
+                    regular = FALSE)
+
+wages_weighted <- as_tsibble(x = wages_hs2020_weighted,
                     key = id,
@@
 wages_obs_sc2 <- get_obs(wages_sc2)
+wages_obs_weig <- get_obs(wages_weighted)
 
@@
 nrow(wages_obs_sc2)
+nrow(wages_obs_weig)
 ```
@@
 ```{r}
-summary(wages_sc1$mean_hourly_wage)
+summary(wages_weighted$mean_hourly_wage)
 ```
-
-```{r}
-summary(wages_sc2$mean_hourly_wage)
-```

### 2021-01-31 code_di_weighted.R
Add flag of the extreme value

@@ -1,12 +1,16 @@
 library(tidyverse)
 
+
+# open the tidy raw data set
 wages <- read_csv("data-raw/wages-high-school-demo/data-frames/tidy_employment_weighted.csv")
 demog <- read_csv("data-raw/wages-high-school-demo/data-frames/tidy_demographics.csv")
 
-d <- read_csv("data-raw/wages-high-school-demo/data-frames/tidy_employment.csv")
 
+# join the wages information and the demographic information by case id.
 wages_demog <- left_join(wages, demog, by="CASEID_1979")
 
+
+# calculate the years in work force and the age of the subjects in 1979
 wages_demog <- wages_demog %>%
   mutate(year_completed_highest_grade = as.numeric(year_completed_highest_grade)) %>%
   mutate(years_in_workforce = year - year_completed_highest_grade) %>%
@@ -15,9 +19,15 @@ wages_demog <- wages_demog %>%
 # filter only the id with high school education
 wages_demog_hs <- wages_demog  %>% filter(grepl("GRADE", `highest-grade`))
 
-# replace the obs with mean_hourly_wage > 30 to be NA
+ggplot(wages_demog_hs, aes(x = mean_hourly_wage)) +
+  geom_boxplot()
+
+# replace the obs with mean_hourly_wage > 100 or mean_hourly_wage < 2 to be NA
 wages_demog_hs <- wages_demog_hs %>%
-  mutate(mean_hourly_wage = ifelse(mean_hourly_wage > 30, NA, mean_hourly_wage))
+  mutate(mean_hourly_wage = ifelse(mean_hourly_wage > 100 |
+                                     mean_hourly_wage < 2, NA, mean_hourly_wage))
+
+
 
 # calculate the number of observation
 keep_me <- wages_demog_hs %>% count(CASEID_1979)
@@ -30,10 +40,31 @@ wages_demog_hs <- wages_demog_hs %>%
   filter(CASEID_1979 %in% keep_me$CASEID_1979)
 
 
+# calculate the IQR to create the extreme value flag
+
+IQR_cut_off <- IQR(wages_demog_hs$mean_hourly_wage, na.rm = TRUE)*1.5
+q1 <- quantile(wages_demog_hs$mean_hourly_wage, 0.25, na.rm = TRUE)
+q3 <- quantile(wages_demog_hs$mean_hourly_wage, 0.75, na.rm = TRUE)
+
[…]

### 2021-02-01 code_di_weighted.R
Add expalantion of NA in the code documentation

@@ -23,6 +23,9 @@ ggplot(wages_demog_hs, aes(x = mean_hourly_wage)) +
   geom_boxplot()
 
 # replace the obs with mean_hourly_wage > 100 or mean_hourly_wage < 2 to be NA
+# the decision to also exclude the wages < 2 is because the minimum wages in old data is 2.030,
+# and in fact, there are 4 observations having 0 in wages.
+
 wages_demog_hs <- wages_demog_hs %>%
   mutate(mean_hourly_wage = ifelse(mean_hourly_wage > 100 |
                                      mean_hourly_wage < 2, NA, mean_hourly_wage))

### 2021-02-01 comparing_scenario.Rmd
Comparing original, weighted, and old data

@@
 load("wages_hs2020_weighted.rda")
+wages_old <- brolgar::wages %>%
+  mutate(mean_hourly_wage = exp(ln_wages))
 ```
@@
 ```{r}
-bp_ori<- ggplot(wages_hs2020) +
-  geom_boxplot(aes(x = mean_hourly_wage)) +
-  facet_wrap(~year) +
-  ggtitle("Original")
 
-bp_wei <- ggplot(wages_hs2020_weighted) +
-  geom_boxplot(aes(x = mean_hourly_wage)) +
-  facet_wrap(~year) +
-  ggtitle("Weighted")
-
-bp_ori
-bp_wei
-
-```
-
-```{r}
+# putting into tsibble
 wages_ori <- as_tsibble(x = wages_hs2020,
-                    key = id,
-                    index = year,
-                    regular = FALSE)
-
-wages_sc1 <- as_tsibble(x = wages_hs2020_sc1,
-                    key = id,
-                    index = year,
-                    regular = FALSE)
-
-wages_sc2 <- as_tsibble(x = wages_hs2020_sc2,
                     key = id,
@@
                     regular = FALSE)
+
+wages_old <- as_tsibble(x = wages_old,
+                    key = id,
+                    index = xp,
+                    regular = FALSE)
 ```
 
+## The summary statistics
+
 ```{r}
-lp_ori <- ggplot(wages_ori) +
-  geom_line(aes(x = year,
-                y = mean_hourly_wage,
-                group = id),
-            alpha = 0.1) +
-  ggtitle("Original")
+get_summary <- function(data){
+  summary(data$mean_hourly_wage)
+}
 
-lp_sc1 <- ggplot(wages_sc1) +
-  geom_line(aes(x = year,
-                y = mean_hourly_wage,
-                group = id),
-            alpha = 0.1) +
-  ggtitle("Scenario 1")
+print("Original data")
+get_summary(wages_ori)
 
-lp_sc2 <- ggplot(wages_sc2) +
-  geom_line(aes(x = year,
-                y = mean_hourly_wage,
-                group = id),
-            alpha = 0.1) +
-  ggtitle("Scenario 2")
+print("Weighted data")
+get_summary(wages_weighted)
 
-lp_ori + lp_sc1 + lp_sc2
+print("Old data")
+get_summary(wages_old)
 
+print("Weighted data, extreme value excluded")
+get_summary(filter(wages_weighted, flag2 == "non_ext_val"))
 ```
+
+## Number of ids in each data and number of observations per id
+
 ```{r}
-
 get_obs <- function(data){
@@
 wages_obs_ori <- get_obs(wages_ori)
-wages_obs_sc1 <- get_obs(wages_sc1)
-wages_obs_sc2 <- get_obs(wages_sc2)
 wages_obs_weig <- get_obs(wages_weighted)
+wages_obs_old <- get_obs(wages_old)
 
 nrow(wages_obs_ori)
-nrow(wages_obs_sc1)
-nrow(wages_obs_sc2)
 nrow(wages_obs_weig)
+nrow(wages_obs_old)
 ```
@@
 ```{r}
-summary(wages_ori$mean_hourly_wage)
+
+get_obs_plot <- function(df, plot_title){
+  ggplot(df, aes(x = n_obs)) +
+  geom_bar() +
+  xlab("Number of observations") +
+  ggtitle(plot_title)
+}
+
+get_obs_plot(wages_obs_ori, "Original data")
+get_obs_plot(wages_obs_weig, "Weighted data")
+get_obs_plot(wages_obs_old, "Old data")
+```
+
+## First time and last time participated in study (Original vs Weighted data)
+
+```{r}
+get_1st_time <- function(df, plot_title){
+  df %>%
+    features(year, list(min = min)) %>%
+    ggplot(aes(x = min)) +
+    geom_histogram(binwidth = 1) +
+    xlab("First time in study") +
+    ggtitle(plot_title)
+}
+
+get_1st_time(wages_ori, "Original data")
+get_1st_time(wages_weighted, "Weighted data")
 ```
@@
 ```{r}
-summary(wages_weighted$mean_hourly_wage)
+get_last_time <- function(df, plot_title){
+  df %>%
+    features(year, list(max = max)) %>%
+    ggplot(aes(x = max)) +
+    geom_histogram(binwidth = 1) +
+    xlab("First time in study") +
+    ggtitle(plot_title)
+}
+
+get_last_time(wages_ori, "Original data")
+get_last_time(wages_weighted, "Weighted data")
 ```
+
+## The spaghetty plot
+
+```{r}
+
+get_sp <- function(df, plot_title, x){
+  ggplot(df) +
+  geom_line(aes(x = year,
+                y = mean_hourly_wage,
+                group = id),
+            alpha = 0.1) +
+  ggtitle(plot_title)
+}
+
+get_sp(wages_ori, "Original data", year)
+get_sp(wages_weighted, "Weighted data", year)
+
+ggplot(wages_old) +
+  geom_line(aes(x = xp,
+                y = mean_hourly_wage,
+                group = id),
+            alpha = 0.1) +
+  ggtitle("Old data")
+
+```
+
+```{r}
+get_sp(filter(wages_weighted, flag2 == "non_ext_val"), "Weighted data - extreme value excluded")
+```
+
+## Take a small spoon of those spaghetty
+
+```{r}
+
+set.seed(202102)
+
+plot_sample <- function(df, plot_title){
+  ggplot(df) +
+  geom_line(aes(x = year,
+                y = mean_hourly_wage,
+                group = id)) +
+  facet_sample() +
+  ggtitle(plot_title)
+}
+
+plot_sample(wages_ori, "Original data")
+plot_sample(wages_weighted, "Weighted data")
+
+ggplot(wages_old) +
+  geom_line(aes(x = xp,
+                y = mean_hourly_wage,
+                group = id)) +
+  facet_sample() +
+  ggtitle("Old data")
+
+```

### 2021-02-02 School_drop_out.Rmd
Update the difference between the hs drop-out and the other groups

@@
 wages_do <- wages %>%
-  filter(hgc %in% c("9TH GRADE", 
+  filter((hgc %in% c("9TH GRADE", 
                     "10TH GRADE",
-                    "11TH GRADE"))
+                    "11TH GRADE")) |
+        (hgc == "12TH GRADE" & 
+        age_1979 >= 19))
 ```
 
+## filter the observations who completed highschool without dropped-out
+
 ```{r}
-
-wages_do_fig <- ggplot(wages_do, 
-                       aes(x = year,
-                           y = mean_hourly_wage,
-                           group = id)) +
-  geom_line(alpha = 0.1)
-
-wages_do_fig
-
+wages_hs_completed <- wages %>%
+  filter(hgc == "12TH GRADE" &
+           age_1979 < 19)
 ```
 
-```{r}
-wages_do_feature<- wages_do %>%
-  features(mean_hourly_wage, 
-           feat_three_num
-           )
-```
+## filter the observations who only completed middle school
 
 ```{r}
-min <- ggplot(wages_do_feature) +
-  geom_density(aes(x = min), fill = "magenta", alpha = 0.5)
-```
-
-```{r}
-wages_do_feature_long <- wages_do_feature %>%
-  pivot_longer(c(min, med, max), names_to = "feature", values_to = "value")
-
-feature_fig <- ggplot(wages_do_feature_long) +
-  geom_density(aes(x = value, colour = feature, fill = feature), alpha = 0.3) +
-  theme(legend.position = "None")
-```
-
-```{r}
-feature_spaghetty <- ggplot(transform(wages_do_feature_long,
-                 feature = factor(feature, level = 
-                                 c("min",
-                                   "med", 
-                                   "max")))) +
-  geom_line(aes(x = feature,
-                y = value, 
-                group = id), alpha = 0.5)
-```
-
-```{r}
-min+feature_fig+feature_spaghetty
-```
-
-Compared to overall cohort the range of the minimum wages of high school dropouts is smaller.
-The median of maximum wages in high school dropouts is smaller compared to the rest of the cohort.
-It is also more positively skewed.
-These mean that the high school dropouts generally received smaller wages then the rest of the cohort, which from the previous EDA, we found that most of the observations completed high school.
-Hence, the reason why the rest of the cohort has bigger wages is because most of the observation has higher education.
-
-```{r}
-ggplot(wages_do, 
-       aes(x = mean_hourly_wage)) +
-  geom_boxplot() +
-  facet_wrap(~year)
-```
-
-```{r}
-wages_do_slope <- key_slope(wages_do, 
-                         mean_hourly_wage ~ year) %>%
-  left_join(wages_do, by = "id")
-
-wages_do_slope
-```
-
-```{r}
-ggplot(wages_do_slope) +
-  geom_line(aes(x = year,
-                y = mean_hourly_wage,
-                group = id)) +
-  gghighlight(.slope_year < 0)
-```
-
-## filter the observations who completed high school
-
-```{r}
-wages_hs <- wages %>%
-  filter(hgc == "12TH GRADE")
-```
-
-```{r}
-wages_hs_fig <- ggplot(wages_hs, 
-                       aes(x = year,
-                           y = mean_hourly_wage,
-                           group = id)) +
-  geom_line(alpha = 0.1)
-
-wages_hs_fig
-```
-
-```{r}
-wages_hs_feature<- wages_hs %>%
-  features(mean_hourly_wage, 
-           feat_three_num
-           )
-```
-
-```{r}
-min_hs <- ggplot(wages_hs_feature) +
-  geom_density(aes(x = min), fill = "magenta", alpha = 0.5)
-```
-
-```{r}
-wages_hs_feature_long <- wages_hs_feature %>%
-  pivot_longer(c(min, med, max), names_to = "feature", values_to = "value")
-
-feature_fig_hs <- ggplot(wages_hs_feature_long) +
-  geom_density(aes(x = value, colour = feature, fill = feature), alpha = 0.3) +
-  theme(legend.position = "None")
-```
-
-```{r}
-feature_spaghetty_hs <- ggplot(transform(wages_hs_feature_long,
-                 feature = factor(feature, level = 
-                                 c("min",
-                                   "med", 
-                                   "max")))) +
-  geom_line(aes(x = feature,
-                y = value, 
-                group = id), alpha = 0.5)
-```
-
-```{r}
-min_hs + feature_fig_hs + feature_spaghetty_hs
-```
-
-```{r}
-ggplot(wages_hs, 
-       aes(x = mean_hourly_wage)) +
-  geom_boxplot() +
-  facet_wrap(~year)
-```
-
-## Middle School
-
-```{r}
-
-wages_ms <- wages %>%
+wages_middle <- wages %>%
   filter(hgc %in% c("6TH GRADE", 
@@
                     "8TH GRADE"))
-
 ```
 
+# calculate the summary
+
 ```{r}
-wages_ms_fig <- ggplot(wages_ms, 
-                       aes(x = year,
-                           y = mean_hourly_wage,
-                           group = id)) +
-  geom_line(alpha = 0.1)
-
-wages_ms_fig
+print("High school droup-out")
+summary(wages_do$mean_hourly_wage)
+print("Completing high school without being dropped-out")
+summary(wages_hs_completed$mean_hourly_wage)
+print("completing middle school")
+summary(wages_middle$mean_hourly_wage)
 ```
 
+# USING THE WEIGHTED DATA
+
+## filter the observations who are the high school drop out
+
+```{r data}
+load("wages_hs2020_weighted.rda")
+
+wages_w <- as_tsibble(x = wages_hs2020_weighted,
+                    key = id,
+                    index = year,
+                    regular = FALSE)
+
+wages_do_w <- wages_w %>%
+  filter((hgc %in% c("9TH GRADE", 
+                    "10TH GRADE",
+                    "11TH GRADE")) |
+        (hgc == "12TH GRADE" & 
+        age_1979 >= 19))
+```
+
+## filter the observations who completed highschool without dropped-out
+
 ```{r}
-wages_do_fig + wages_hs_fig + wages_ms_fig
+wages_hs_completed_w <- wages_w %>%
+  filter(hgc == "12TH GRADE" &
+           age_1979 < 19)
 ```
+
+## filter the observations who only completed middle school
+
+```{r}
+wages_middle_w <- wages_w %>%
+  filter(hgc %in% c("6TH GRADE", 
+                    "7TH GRADE",
+                    "8TH GRADE"))
+```
+
+# calculate the summary
+
+```{r}
+print("High school droup-out")
+summary(wages_do_w$mean_hourly_wage)
+print("Completing high school without being dropped-out")
+summary(wages_hs_completed_w$mean_hourly_wage)
+print("completing middle school")
+summary(wages_middle_w$mean_hourly_wage)
+```

### 2021-02-02 New_Wages.Rmd
Revisit the preprocessing code for the demographic information.

@@
 
+## LOAD THE RAW DATA OF DEMOGRAPHIC AND WAGES INFORMATION
+
 ```{r}
@@
 ```
+
+## DEMOGRAPHIC DATA PREPROCESSING
+
+```{r}
+
+## tidy the date of birth data
+
+dob_tidy <- new_data_qnames %>%
+  select(CASEID_1979,
+         starts_with("Q1-3_A~")) %>%
+  mutate(dob_year = case_when(
+                    # if the years recorded in both sets match, take 79 data
+                    `Q1-3_A~Y_1979` == `Q1-3_A~Y_1981` ~ `Q1-3_A~Y_1979`,
+                    # if the year in the 81 set is missing, take the 79 data
+                    is.na(`Q1-3_A~Y_1981`) ~ `Q1-3_A~Y_1979`,
+                    # if the sets don't match for dob year, take the 79 data
+                    `Q1-3_A~Y_1979` != `Q1-3_A~Y_1981` ~ `Q1-3_A~Y_1979`),
+        dob_month = case_when(
+                    # if months recorded in both sets match, take 79 data
+                    `Q1-3_A~M_1979` == `Q1-3_A~M_1981` ~ `Q1-3_A~M_1979`,
+                    # if month in 81 set is missing, take the 79 data
+                    is.na(`Q1-3_A~M_1981`) ~ `Q1-3_A~M_1979`,
+                    # if sets don't match for dob month, take the 79 data
+                    `Q1-3_A~M_1979` != `Q1-3_A~M_1981` ~ `Q1-3_A~M_1979`),
+        # flag if there is a conflict between dob recorded in 79 and 81
+        dob_conflict = case_when(     
+                      (`Q1-3_A~M_1979` != `Q1-3_A~M_1981`) & !is.na(`Q1-3_A~M_1981`) ~ TRUE,
+                      (`Q1-3_A~Y_1979` != `Q1-3_A~Y_1981`) & !is.na(`Q1-3_A~Y_1981`) ~ TRUE,
+                      (`Q1-3_A~Y_1979` == `Q1-3_A~Y_1981`) & (`Q1-3_A~M_1979` == `Q1-3_A~M_1981`) ~ FALSE,
+                      is.na(`Q1-3_A~M_1981`) | is.na(`Q1-3_A~Y_1981`) ~ FALSE)) %>%
+  select(CASEID_1979,
+         dob_month,
+         dob_year,
+         dob_conflict)
+```
+
+```{r}
+
+## tidy the sex and race data
+demog_tidy <- categories_qnames %>%
+  select(CASEID_1979,
+         SAMPLE_RACE_78SCRN,
+         SAMPLE_SEX_1979) %>%
+  rename(gender = SAMPLE_SEX_1979,
+         race = SAMPLE_RACE_78SCRN)
+```
+
+```{r}
+
+# tidy the grade 
+demog_education <- new_data_qnames %>%
+  as_tibble() %>%
+  select(CASEID_1979,
+         starts_with("HGCREV"),
+         "HGC_2012",
+         "HGC_2014",
+         "HGC_2016") %>%
+  pivot_longer(!CASEID_1979,
+               names_to = "var",
+               values_to = "grade") %>%
+  separate("var", c("var", "year"), sep = -4) %>%
+  filter(!is.na(grade)) %>%
+  select(-var)
+
+```
+
+```{r}
+
+## getting the highest year of completed education
+
+highest_year <- demog_education %>%
+  group_by(CASEID_1979) %>%
+  mutate(hgc_i = max(grade)) %>%
+  filter(hgc_i == grade) %>%
+  filter(year == first(year)) %>%
+  rename(yr_hgc = year) %>%
+  select(CASEID_1979, yr_hgc, hgc_i) %>%
+  ungroup() %>%
+  mutate('hgc' = ifelse(hgc_i == 0, "NONE",
+                                  ifelse(hgc_i == 1, "1ST GRADE",
+                                         ifelse(hgc_i == 2, "2ND GRADE",
+                                                ifelse(hgc_i == 3, "3RD GRADE",
+                                                       ifelse(hgc_i >= 4 & hgc_i <= 12, paste0(hgc_i,"TH GRADE"),
+                                                              ifelse(hgc_i == 13, "1ST YEAR COL",
+                                                                     ifelse(hgc_i == 14, "2ND YEAR COL",
+                                                                            ifelse(hgc_i == 15, "3RD YEAR COL",
+                                                                                   ifelse(hgc_i == 95, "UNGRADED", paste0((hgc_i - 12), "TH YEAR COL")))))))))))
+```
+
+## Join all of the demographic information
+
+```{r}
+full_demographics <- full_join(dob_tidy, demog_tidy, by = "CASEID_1979") %>%
+  full_join(highest_year, by = "CASEID_1979") %>%
+  rename("id" = "CASEID_1979")
+```
+
+# WAGES DATA PRE-PROCESSING
+
+## Tidying hour of work data
 
@@
 ```{r}
-
 hours <- list()
@@
 
-## get rates
+## Tidying rates per hour data
 
@@
   ungroup() %>%
-  rename(mean_hourly_wage = wages,
-         CASEID_1979 = id) %>%
+  rename(mean_hourly_wage = wages) %>%
   mutate(flag1 = ifelse(flag1 == 1, "wm",
@@
 ```{r}
-#write.csv(mhw_to_write, here::here("data-raw/wages-high-school-demo/data-frames/tidy_employment_weighted.csv"))
+write.csv(mhw_to_write, here::here("data-raw/wages-high-school-demo/data-frames/tidy_employment_weighted.csv"))
+write.csv(full_demographics, here::here("data-raw/wages-high-school-demo/data-frames/tidy_demographics_v2.csv"))
 ```

### 2021-02-02 code_di_weighted.R
Minor change in the variable name.

@@ -3,21 +3,21 @@ library(tidyverse)
 
 # open the tidy raw data set
 wages <- read_csv("data-raw/wages-high-school-demo/data-frames/tidy_employment_weighted.csv")
-demog <- read_csv("data-raw/wages-high-school-demo/data-frames/tidy_demographics.csv")
+demog <- read_csv("data-raw/wages-high-school-demo/data-frames/tidy_demographics_v2.csv")
 
 
 # join the wages information and the demographic information by case id.
-wages_demog <- left_join(wages, demog, by="CASEID_1979")
+wages_demog <- left_join(wages, demog, by="id")
 
 
 # calculate the years in work force and the age of the subjects in 1979
 wages_demog <- wages_demog %>%
-  mutate(year_completed_highest_grade = as.numeric(year_completed_highest_grade)) %>%
-  mutate(years_in_workforce = year - year_completed_highest_grade) %>%
+  mutate(yr_hgc = as.numeric(yr_hgc)) %>%
+  mutate(years_in_workforce = year - yr_hgc) %>%
   mutate(age_1979 = 1979 - (dob_year + 1900))
 
 # filter only the id with high school education
-wages_demog_hs <- wages_demog  %>% filter(grepl("GRADE", `highest-grade`))
+wages_demog_hs <- wages_demog  %>% filter(grepl("GRADE", hgc))
 
 ggplot(wages_demog_hs, aes(x = mean_hourly_wage)) +
   geom_boxplot()
@@ -33,14 +33,14 @@ wages_demog_hs <- wages_demog_hs %>%
 
 
 # calculate the number of observation
-keep_me <- wages_demog_hs %>% count(CASEID_1979)
+keep_me <- wages_demog_hs %>% count(id)
 keep_me %>% ggplot(aes(x=n)) + geom_bar()
 
 # filter out the id with low number of observation
 keep_me <- keep_me %>% filter(n > 4)
 
 wages_demog_hs <- wages_demog_hs %>%
-  filter(CASEID_1979 %in% keep_me$CASEID_1979)
+  filter(id %in% keep_me$id)
 
 
 # calculate the IQR to create the extreme value flag
@@ -55,19 +55,17 @@ upper_limit <- q3 + IQR_cut_off
 wages_demog_hs <- wages_demog_hs %>%
   mutate(flag2 = ifelse(mean_hourly_wage < lower_limit |
[…]

### 2021-02-02 School_drop_out.Rmd
Update the code for highschool drop-out.

@@
                     index = year,
-                    regular = FALSE)
+                    regular = FALSE) %>%
+  mutate(dob = 1979 - age_1979, 
+         age_hgc = yr_hgc - dob)
 
@@
         (hgc == "12TH GRADE" & 
-        age_1979 >= 19))
+        age_hgc >= 19))
 ```
@@
   filter(hgc == "12TH GRADE" &
-           age_1979 < 19)
+           age_hgc < 19)
 ```

### 2021-02-03 New_Wages.Rmd
Edit the code in flag and get_hour function.

@@
   if(year %in% year_A){
-   new_data_qnames %>%
+   temp <- new_data_qnames %>%
     select(CASEID_1979,
             starts_with("QES-52A") &
-              ends_with(as.character(year))) %>%
-    pivot_longer(!CASEID_1979,
-                 names_to = "job",
-                 values_to = "hours_work") %>%
-    separate("job", c("job", "year"), sep = -4) %>%
-    mutate(job = paste0("job_", substr(job, 9, 10))) %>%
-    rename(id = CASEID_1979)}
+              ends_with(as.character(year)))} 
   else{
-    new_data_qnames %>%
+    temp <- new_data_qnames %>%
     select(CASEID_1979,
             starts_with("QES-52D") &
-              ends_with(as.character(year))) %>%
+              ends_with(as.character(year)))} 
+    temp %>% 
       pivot_longer(!CASEID_1979,
@@
       rename(id = CASEID_1979)
-  }
+
  }
@@
   group_by(id, year) %>%
+  #calculate the weighted mean if the number of jobs > 1
   mutate(wages = ifelse(no_jobs == 1, rate_per_hour/100,
                         weighted.mean(rate_per_hour, hours_work, na.rm = TRUE)/100)) %>%
+  #give the flag if it the weighted mean
   mutate(flag1 = ifelse(!is.na(wages) & no_jobs != 1, 1,
                         0)) %>%
+  #calculate the arithmetic mean for the na
   mutate(wages = ifelse(is.na(wages), mean(rate_per_hour)/100,
@@
   rename(mean_hourly_wage = wages) %>%
-  mutate(flag1 = ifelse(flag1 == 1, "wm",
-                        "am"))
+  mutate(is_wm = ifelse(flag1 == 1, TRUE,
+                        FALSE)) %>%
+  select(-flag1)
 ```

### 2021-02-03 code_di_weighted.R
Change the label of the flag.

@@ -53,19 +53,22 @@ lower_limit <- q1 - IQR_cut_off
 upper_limit <- q3 + IQR_cut_off
 
 wages_demog_hs <- wages_demog_hs %>%
-  mutate(flag2 = ifelse(mean_hourly_wage < lower_limit |
-                          mean_hourly_wage > upper_limit, "ext_val",
-                        "non_ext_val")) %>%
-  mutate(flag = paste(flag1, flag2, sep = "-"))
+  mutate(is_extreme_val = ifelse(mean_hourly_wage < lower_limit |
+                          mean_hourly_wage > upper_limit, TRUE,
+                        FALSE)) %>%
+  mutate(flag = case_when(is_wm == TRUE & is_extreme_val == FALSE ~ 1,
+                          is_wm == TRUE & is_extreme_val == TRUE ~ 2,
+                          is_wm == FALSE & is_extreme_val == FALSE ~ 3,
+                          is_wm == FALSE & is_extreme_val == TRUE ~ 4))
 
-filtered <- filter(wages_demog_hs, flag2 == "non_ext_val")
+filtered <- filter(wages_demog_hs, is_extreme_val == FALSE)
 summary(filtered$mean_hourly_wage)
 
 
 
 # rename and select the wages in tidy
 wages_hs2020_weighted <- wages_demog_hs %>%
-  select(id, year, mean_hourly_wage, age_1979, gender, race, hgc, hgc_i, yr_hgc, number_of_jobs, total_hours, flag1, flag2, flag)
+  select(id, year, mean_hourly_wage, age_1979, gender, race, hgc, hgc_i, yr_hgc, number_of_jobs, total_hours, is_wm, is_extreme_val, flag)
 
 save(wages_hs2020_weighted, file="wages_hs2020_weighted.rda")
 

### 2021-02-08 code_di_weighted.R
Doing locf for the wages >00 per hour.

@@ -1,4 +1,5 @@
 library(tidyverse)
+library(zoo)
 
 
 # open the tidy raw data set
@@ -22,14 +23,27 @@ wages_demog_hs <- wages_demog  %>% filter(grepl("GRADE", hgc))
 ggplot(wages_demog_hs, aes(x = mean_hourly_wage)) +
   geom_boxplot()
 
-# replace the obs with mean_hourly_wage > 100 or mean_hourly_wage < 2 to be NA
-# the decision to also exclude the wages < 2 is because the minimum wages in old data is 2.030,
-# and in fact, there are 4 observations having 0 in wages.
+# replace the obs with mean_hourly_wage > 100 to be NA
+# since we will impute the missing values with the last observation carried forward (locf) for each id
+# we give the flag for that locf observation
 
 wages_demog_hs <- wages_demog_hs %>%
-  mutate(mean_hourly_wage = ifelse(mean_hourly_wage > 100 |
-                                     mean_hourly_wage < 2, NA, mean_hourly_wage))
+  mutate(mean_hourly_wage = ifelse(mean_hourly_wage > 100, NA, mean_hourly_wage),
+         is_locf = ifelse(is.na(mean_hourly_wage), TRUE, FALSE))
 
+# doing locf
+
+for_locf <- wages_demog_hs %>%
+  select(id, year, mean_hourly_wage) %>%
+  group_by(id) %>%
+  na.locf()
+
+
+# join back
+
+wages_demog_hs <- left_join(wages_demog_hs, for_locf, by = c("id", "year")) %>%
+  select(-mean_hourly_wage.x) %>%
+  rename(mean_hourly_wage = mean_hourly_wage.y)
 
 
 # calculate the number of observation
@@ -68,7 +82,10 @@ summary(filtered$mean_hourly_wage)
 
 # rename and select the wages in tidy
 wages_hs2020_weighted <- wages_demog_hs %>%
-  select(id, year, mean_hourly_wage, age_1979, gender, race, hgc, hgc_i, yr_hgc, number_of_jobs, total_hours, is_wm, is_extreme_val, flag)
+  select(id, year, mean_hourly_wage, age_1979, gender, race, hgc, hgc_i, yr_hgc, number_of_jobs, total_hours, is_wm, is_extreme_val, is_locf, flag)
 
 save(wages_hs2020_weighted, file="wages_hs2020_weighted.rda")
+
+summary(wages_hs2020_weighted$mean_hourly_wage)
+
 

### 2021-02-10 code_di_weighted.R
Change the treatment of extreme values

@@ -1,5 +1,6 @@
 library(tidyverse)
 library(zoo)
+library(rstatix)
 
 
 # open the tidy raw data set
@@ -23,13 +24,35 @@ wages_demog_hs <- wages_demog  %>% filter(grepl("GRADE", hgc))
 ggplot(wages_demog_hs, aes(x = mean_hourly_wage)) +
   geom_boxplot()
 
-# replace the obs with mean_hourly_wage > 100 to be NA
+# replace the obs with mean_hourly_wage is.extreme == TRUE to be NA
 # since we will impute the missing values with the last observation carried forward (locf) for each id
 # we give the flag for that locf observation
 
+wages_demog_ext <- wages_demog_hs %>%
+  group_by(id) %>%
+  identify_outliers("mean_hourly_wage") %>%
+  filter(is.extreme == TRUE) %>%
+  select(year, id, is.extreme)
+
+# ggplot(wages_demog_ext) +
+#   geom_histogram(aes(x = mean_hourly_wage))
+
+#summary(wages_demog_ext$mean_hourly_wage)
+
+
+# filter(wages_demog_ext, id == 3)
+# filter(wages_demog_hs, id == 6885)
+
+wages_demog_hs <- left_join(wages_demog_hs, wages_demog_ext, by = c("id", "year"))
+
 wages_demog_hs <- wages_demog_hs %>%
-  mutate(mean_hourly_wage = ifelse(mean_hourly_wage > 100, NA, mean_hourly_wage),
+  mutate(is.extreme = ifelse(is.na(is.extreme), FALSE,
+                             is.extreme))
+
+wages_demog_hs <- wages_demog_hs %>%
+  mutate(mean_hourly_wage = ifelse(is.extreme == TRUE, NA, mean_hourly_wage),
          is_locf = ifelse(is.na(mean_hourly_wage), TRUE, FALSE))
+
 
 # doing locf
 
@@ -59,33 +82,38 @@ wages_demog_hs <- wages_demog_hs %>%
 
 # calculate the IQR to create the extreme value flag
 
-IQR_cut_off <- IQR(wages_demog_hs$mean_hourly_wage, na.rm = TRUE)*1.5
-q1 <- quantile(wages_demog_hs$mean_hourly_wage, 0.25, na.rm = TRUE)
-q3 <- quantile(wages_demog_hs$mean_hourly_wage, 0.75, na.rm = TRUE)
-
-lower_limit <- q1 - IQR_cut_off
-upper_limit <- q3 + IQR_cut_off
-
[…]

### 2021-02-13 New_Wages.Rmd
treat the observations with 0 rates per hour.
Update to survey year 2018 data.

@@
 ```{r}
+#source(here::here("data-raw/NLSY79/NLSY79.R"))
+```
+
+Creates four files: new_data, has the variables using RNUM;
+new_data_qnames, has the data with qnames
+categories_vallabels, has RNUM and coding of each variable;
+categories_qnames, has qname and coding of each variable;
+(qname must be Question name?)
+
+```{r}
+
+# write_rds(as_tibble(new_data_qnames), 
+#    here::here("data-raw/NLSY79/NLSY79-new_data_qnames.rds"),       
+#       compress = "xz")
+
+# write_rds(as_tibble(categories_qnames), 
+#          here::here("data-raw/NLSY79/NLSY79-categories_qnames.rds"), 
+#        compress = "xz")
+
+```
+
+```{r}
 
 # load the downloaded data 
-new_data_qnames <- read_rds(here::here("data-raw/wages-high-school-demo/data-frames/wages-high-school-demo-new_data_qnames.rds"))
-categories_qnames <- read_rds(here::here("data-raw/wages-high-school-demo/data-frames/wages-high-school-demo-categories_qnames.rds"))
+new_data_qnames <- read_rds(here::here("data-raw/NLSY79/NLSY79-new_data_qnames.rds"))
+categories_qnames <- read_rds(here::here("data-raw/NLSY79/NLSY79-categories_qnames.rds"))
 
@@
   as_tibble() %>%
+  rename(HGC_2018 = `Q3-4_2018`) %>%
   select(CASEID_1979,
@@
          "HGC_2014",
-         "HGC_2016") %>%
+         "HGC_2016",
+         "HGC_2018") %>%
   pivot_longer(!CASEID_1979,
@@
                                                                                    ifelse(hgc_i == 95, "UNGRADED", paste0((hgc_i - 12), "TH YEAR COL")))))))))))
+
 ```
@@
 for(ayear in c(1979:1994, 1996, 1998, 2000, 2002, 2004, 2006, 2008, 2010, 
-               2012, 2014, 2016)) {
+               2012, 2014, 2016, 2018)) {
    hours[[ayear]] <- get_hour(ayear)
@@
 for(ayear in c(1979:1994, 1996, 1998, 2000, 2002, 2004, 2006, 2008, 2010, 
-               2012, 2014, 2016)) {
+               2012, 2014, 2016, 2018)) {
    rates[[ayear]] <- get_rate(ayear)
@@
                          hours_all, 
-                         by = c("id", "year", "job"))
+                         by = c("id", "year", "job")) %>%
+
+# set the 0 value in rate_per_hour as NA 
+  mutate(rate_per_hour = ifelse(rate_per_hour == 0, NA,
+                                rate_per_hour))
 ```

### 2021-02-13 School_drop_out.Rmd
Update to survey year 2018 data.

@@
         age_hgc >= 19))
+
+wages_do_comp <- wages_do %>%
+  filter(gender == "MALE" &
+           age_1979 %in% c(14, 15, 16, 17))
+summary(wages_do_comp$mean_hourly_wage)
 ```
@@
         age_hgc >= 19))
+
+wages_do_comp_w <- wages_do_w %>%
+  filter(gender == "MALE" &
+           age_1979 %in% c(14, 15, 16, 17))
+summary(wages_do_comp_w$mean_hourly_wage)
 ```
@@
                     "8TH GRADE"))
+
 ```

### 2021-02-13 anomaly_v2.Rmd
Update to survey year 2018 data.

--- 2021-01-18 anomaly.Rmd
@@
 ```{r data}
-load("wages_hs2020.rda")
+load("wages_hs2020_weighted_v2.rda")
 ```
@@
 ```{r wages}
-wages <- as_tsibble(x = wages_hs2020,
+wages <- as_tsibble(x = wages_hs2020_weighted,
                     key = id,
@@
 
-```{r}
-wages %>%
-  filter(id == 11974)
-```
-
 ### Plot the wages with low variance in their lm-residuals
@@
 bottom12 <- wages_lm_var %>%
-  filter(rank >= 5995) %>%
+  slice_tail(n =20) %>%
   left_join(wages, by = "id")
@@
 ```
-
-```{r}
-wages %>%
-  filter(year == 1984)
-```
-
-```{r}
-ggplot(wages) +
-  geom_boxplot(aes(x = mean_hourly_wage)) +
-  facet_wrap(~year)
-```

### 2021-02-13 anomaly_v3.Rmd
Update to survey year 2018 data.

--- 2021-02-13 anomaly_v2.Rmd
@@
 ```{r data}
-load("wages_hs2020_weighted_v2.rda")
+load("wages_hs2020_weighted_updated2018.rda")
 ```
@@
 complete_year_obs <- wages_obs %>%
-  filter(n_obs == 27) %>%
+  filter(n_obs == 28) %>%
   left_join(wages_lm_var, by = "id") %>%

### 2021-02-13 getting_the_data.Rmd
Update to survey year 2018 data.

---
title: "Instruction for Getting The Data"
author: "Dewi Lestari Amaliah"
date: "04/02/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, eval = FALSE)
library(tidyverse)
library(brolgar)
```

# Introduction

Data refreshing is the process to update a data set to maintain its timeliness in a consistent and accurate manner. In this case, the data refreshing is performed on the Wages Data from the National Longitudinal Survey of Youth (NLSY) 1979. It is a longitudinal project that follows the lives of a sample of American youth born between 1957-64. The cohort originally included 12,686 respondents ages 14-22 when first interviewed in 1979; after two sub-samples were dropped, 9,964 respondents remain in the eligible samples. Data are now available from Round 1 (1979 survey year) to Round 27 (2016 survey year).

Refreshing the wages data includes the process of cleaning and preprocess the data that downloaded from the NLSY's database to produce the mean hourly wages of the subject along with several demographic variables. The cleaned data set could be used for various purposes, including for teaching and research  purposes, with the demonstration of the usage of `brolgar` package, a package to analyze and visualize longitudinal data with R (Tierney, Cook, and Prvan, 2020). The tidy and cleaned data set is also made to be a package called "yowie".

# Getting The Data

## Getting The Data from The Database

The result of NLYS is stored in a [database](https://www.nlsinfo.org/content/cohorts/nlsy79/get-data) and could be downloaded by variables. In the data base, there are several variables that are ready to be downloaded, such as, education, employment, household, marriage, children, and health. However, we only downloaded these variables:

- Education, Training & Achievement Scores
    -  Education
        -  Summary measures
            - All  schools
                - By year
                    - Highest grade completed (Downloaded all of the 78 variables.)
- Employment
    - Summary measures
        - By job
            - Hours worked (Downloaded all of the 427 variables.)
            - Hourly wages (Downloaded all of the 151 variables.)

            Both hours worked and hourly wages are recorded by the job, up to five jobs for each id/subject. 
- Household, Geography & Contextual Variables
    - Context
        - Summary measures
            - Basic demographics 
              (Downloaded year and month of birth, race, and sex variables.
              There are two versions of the year and month of birth, i.e. 1979 and 1981 data. We downloaded these two versions).

The downloaded data set from the data base are in a csv (wages-high-school-demo.csv) and dat (wages-high-school-demo.dat) format. We will only use the dat format. It also came along with these files:

- wages-high-school-demo.NLSY79: This is the tagset, that can be uploaded to the web site to recreate the data set
- wages-high-school-demo.R: provided by NLSY for reading the data into R and convert the variable's name and its label into something more sensible.

## Data Preprocessing

The downloaded data came with untidy format, where all of the variables for each year being a column (the raw data contains 664 column). We will not use all of the downloaded variables, we will only use the variables as follows:

### Date of Birth

For the Date of Birth data, since there were two versions (1979 and 1981), we only used the 1979 data. Hence, we match both years and flag if it is not match.

```{r dob-tidy}
## tidy the date of birth data
dob_tidy <- new_data_qnames %>%
  select(CASEID_1979,
         starts_with("Q1-3_A~")) %>%
  mutate(dob_year = case_when(
                    # if the years recorded in both sets match, take 79 data
                    `Q1-3_A~Y_1979` == `Q1-3_A~Y_1981` ~ `Q1-3_A~Y_1979`,
                    # if the year in the 81 set is missing, take the 79 data
                    is.na(`Q1-3_A~Y_1981`) ~ `Q1-3_A~Y_1979`,
                    # if the sets don't match for dob year, take the 79 data
                    `Q1-3_A~Y_1979` != `Q1-3_A~Y_1981` ~ `Q1-3_A~Y_1979`),
        dob_month = case_when(
                    # if months recorded in both sets match, take 79 data
                    `Q1-3_A~M_1979` == `Q1-3_A~M_1981` ~ `Q1-3_A~M_1979`,
                    # if month in 81 set is missing, take the 79 data
                    is.na(`Q1-3_A~M_1981`) ~ `Q1-3_A~M_1979`,
                    # if sets don't match for dob month, take the 79 data
                    `Q1-3_A~M_1979` != `Q1-3_A~M_1981` ~ `Q1-3_A~M_1979`),
        # flag if there is a conflict between dob recorded in 79 and 81
        dob_conflict = case_when(     
                      (`Q1-3_A~M_1979` != `Q1-3_A~M_1981`) & !is.na(`Q1-3_A~M_1981`) ~ TRUE,
                      (`Q1-3_A~Y_1979` != `Q1-3_A~Y_1981`) & !is.na(`Q1-3_A~Y_1981`) ~ TRUE,
                      (`Q1-3_A~Y_1979` == `Q1-3_A~Y_1981`) & (`Q1-3_A~M_1979` == `Q1-3_A~M_1981`) ~ FALSE,
                      is.na(`Q1-3_A~M_1981`) | is.na(`Q1-3_A~Y_1981`) ~ FALSE)) %>%
  select(CASEID_1979,
         dob_month,
         dob_year,
         dob_conflict)
```

### Sex and Race

For these variables, we only rename the variable to be more clear.

```{r}
## tidy the sex and race data
demog_tidy <- categories_qnames %>%
  select(CASEID_1979,
         SAMPLE_RACE_78SCRN,
         SAMPLE_SEX_1979) %>%
  rename(gender = SAMPLE_SEX_1979,
         race = SAMPLE_RACE_78SCRN)
```

### Education i.e Highest Grade Completed

For the Education i.e. Highest Grade Completed data, we chose the revised May data because the May data seemed to have less missing and presumably the revised data has been checked. However, there was no revised May data for 2012, 2014 and 2016, so we just used the ordinary May data.

Each id will have the highest grade completed for each year. In the final data, we only used the highest grade completed ever and derived the year of when the highest grade completed.

```{r hgc-tidy}
# tidy the grade 
demog_education <- new_data_qnames %>%
  as_tibble() %>%
  select(CASEID_1979,
         starts_with("HGCREV"),
         "HGC_2012",
         "HGC_2014",
         "HGC_2016") %>%
  pivot_longer(!CASEID_1979,
               names_to = "var",
               values_to = "grade") %>%
  separate("var", c("var", "year"), sep = -4) %>%
  filter(!is.na(grade)) %>%
  select(-var)

## getting the highest year of completed education

highest_year <- demog_education %>%
  group_by(CASEID_1979) %>%
  mutate(hgc_i = max(grade)) %>%
  filter(hgc_i == grade) %>%
  filter(year == first(year)) %>%
  rename(yr_hgc = year) %>%
  select(CASEID_1979, yr_hgc, hgc_i) %>%
  ungroup() %>%
  mutate('hgc' = ifelse(hgc_i == 0, "NONE",
                                  ifelse(hgc_i == 1, "1ST GRADE",
                                         ifelse(hgc_i == 2, "2ND GRADE",
                                                ifelse(hgc_i == 3, "3RD GRADE",
                                                       ifelse(hgc_i >= 4 & hgc_i <= 12, paste0(hgc_i,"TH GRADE"),
                                                              ifelse(hgc_i == 13, "1ST YEAR COL",
                                                                     ifelse(hgc_i == 14, "2ND YEAR COL",
                                                                            ifelse(hgc_i == 15, "3RD YEAR COL",
                                                                                   ifelse(hgc_i == 95, "UNGRADED", paste0((hgc_i - 12), "TH YEAR COL")))))))))))
```

### Employment Data (Hours Work and Hourly Wage)

For the Employment i.e. Hours worked per week data, initially only one variable per job, no choice from 1979 to 1987 (QES-52A). From 1988 onward, when we had more options, we chose the variable for total hours including time spent working from home (QES-52D). However, 1993 didn't have all five D variables (the first one and the last one were missing), so we used QES-52A variable instead. Further, 2008 only had jobs 1-4 for the QES-52D variable (whereas the other years had 1-5), so we just used these.

Further, since an id might has more than one job, we calculate the wages for each id in each year using the weighted mean with the hour of work as the weight. However, there are a lot of missing value in hours work data. In this case, if the hourly wage information is available, we will only use the regular/arithmetic mean. Hence, we made a new variable to flag whether the mean hourly wage is a weighted or regular mean. Additionally, if an id only has one job, we will directly use their hourly wages information and recorded as arithmetic mean in the flag.

# Data Dictionary of The Tidy Data

- Kind of the metadata of the tidy data

### 2021-02-13 code_di_SC1.R

@@ -1,67 +1,0 @@
-library(tidyverse)
-
-wages <- read_csv("data-raw/wages-high-school-demo/data-frames/tidy_employment_sc1.csv")
-demog <- read_csv("data-raw/wages-high-school-demo/data-frames/tidy_demographics.csv")
-
-wages_demog <- left_join(wages, demog, by="CASEID_1979")
-
-wages_demog %>%
-  filter(`highest-grade` %in% c("10TH GRADE", "11TH GRADE", "12TH GRADE")) %>%
-  ggplot(aes(x=year, y=mean_hourly_wage))  +
-  geom_point() + facet_wrap(~`highest-grade`)
-# Look like some issues with wages > 10000 per hour
-# Need to organsie education
-wages_demog %>% #count(gender)
-  #filter(`highest-grade` %in% c("10TH GRADE", "11TH GRADE", "12TH GRADE")) %>%
-  ggplot(aes(x=year, y=mean_hourly_wage))  +
-  geom_point() + facet_wrap(~gender)
-
-wages_demog <- wages_demog %>%
-  mutate(year = as.numeric(year),
-         year_completed_highest_grade = as.numeric(year_completed_highest_grade)) %>%
-  mutate(years_in_workforce = year - year_completed_highest_grade) %>%
-  mutate(age_1979 = 1979 - (dob_year + 1900))
-
-wages_demog %>% filter(years_in_workforce < 0) %>% tally()
-
-wages_demog %>% filter(years_in_workforce < 0) %>%
-  count(CASEID_1979) %>% tally()
-wages_demog %>%
-  ggplot(aes(x=years_in_workforce, y=number_of_jobs)) +
-  geom_point()
-wages_demog_hs <- wages_demog  %>% filter(grepl("GRADE", `highest-grade`))
-wages_demog_hs %>% filter(years_in_workforce < 0) %>% tally()
-wages_demog_hs %>% filter(years_in_workforce < -35) %>%
-  select(CASEID_1979, year, year_completed_highest_grade, `highest-grade`, dob_year, number_of_jobs, mean_hourly_wage)
-wages_demog_hs %>% filter(CASEID_1979 == "4396") %>% print(n=50)
-
-wages_demog %>% ggplot(aes(x=age_1979)) + geom_bar()
-summary(wages_demog$mean_hourly_wage)
[…]

### 2021-02-13 code_di_SC2.R

@@ -1,67 +1,0 @@
-library(tidyverse)
-
-wages <- read_csv("data-raw/wages-high-school-demo/data-frames/tidy_employment_sc2.csv")
-demog <- read_csv("data-raw/wages-high-school-demo/data-frames/tidy_demographics.csv")
-
-wages_demog <- left_join(wages, demog, by="CASEID_1979")
-
-wages_demog %>%
-  filter(`highest-grade` %in% c("10TH GRADE", "11TH GRADE", "12TH GRADE")) %>%
-  ggplot(aes(x=year, y=mean_hourly_wage))  +
-  geom_point() + facet_wrap(~`highest-grade`)
-# Look like some issues with wages > 10000 per hour
-# Need to organsie education
-wages_demog %>% #count(gender)
-  #filter(`highest-grade` %in% c("10TH GRADE", "11TH GRADE", "12TH GRADE")) %>%
-  ggplot(aes(x=year, y=mean_hourly_wage))  +
-  geom_point() + facet_wrap(~gender)
-
-wages_demog <- wages_demog %>%
-  mutate(year = as.numeric(year),
-         year_completed_highest_grade = as.numeric(year_completed_highest_grade)) %>%
-  mutate(years_in_workforce = year - year_completed_highest_grade) %>%
-  mutate(age_1979 = 1979 - (dob_year + 1900))
-
-wages_demog %>% filter(years_in_workforce < 0) %>% tally()
-
-wages_demog %>% filter(years_in_workforce < 0) %>%
-  count(CASEID_1979) %>% tally()
-wages_demog %>%
-  ggplot(aes(x=years_in_workforce, y=number_of_jobs)) +
-  geom_point()
-wages_demog_hs <- wages_demog  %>% filter(grepl("GRADE", `highest-grade`))
-wages_demog_hs %>% filter(years_in_workforce < 0) %>% tally()
-wages_demog_hs %>% filter(years_in_workforce < -35) %>%
-  select(CASEID_1979, year, year_completed_highest_grade, `highest-grade`, dob_year, number_of_jobs, mean_hourly_wage)
-wages_demog_hs %>% filter(CASEID_1979 == "4396") %>% print(n=50)
-
-wages_demog %>% ggplot(aes(x=age_1979)) + geom_bar()
-summary(wages_demog$mean_hourly_wage)
[…]

### 2021-02-13 code_di_weighted.R

@@ -34,14 +34,8 @@ wages_demog_ext <- wages_demog_hs %>%
   filter(is.extreme == TRUE) %>%
   select(year, id, is.extreme)
 
-# ggplot(wages_demog_ext) +
-#   geom_histogram(aes(x = mean_hourly_wage))
-
-#summary(wages_demog_ext$mean_hourly_wage)
 
 
-# filter(wages_demog_ext, id == 3)
-# filter(wages_demog_hs, id == 6885)
 
 wages_demog_hs <- left_join(wages_demog_hs, wages_demog_ext, by = c("id", "year"))
 
@@ -105,9 +99,9 @@ wages_demog_hs <- wages_demog_hs %>%
 
 # rename and select the wages in tidy
 wages_hs2020_weighted <- wages_demog_hs %>%
-  select(id, year, mean_hourly_wage, age_1979, gender, race, hgc, hgc_i, yr_hgc, number_of_jobs, total_hours, is_wm, is.extreme, is_locf)
+  select(id, year, mean_hourly_wage, age_1979, gender, race, hgc, hgc_i, yr_hgc, number_of_jobs, total_hours, is_wm, is_locf)
 
-#save(wages_hs2020_weighted, file="wages_hs2020_weighted.rda")
+#save(wages_hs2020_weighted, file="wages_hs2020_weighted_updated2018.rda")
 
 summary(wages_hs2020_weighted$mean_hourly_wage)
 

### 2021-02-13 data-raw/wages-high-school-demo/tidy-wages-high-school-demo.R

@@ -114,14 +114,14 @@ full_demographics <- full_join(dob_tidy, demog_tidy, by = "CASEID_1979") %>%
   full_join(highest_year, by = "CASEID_1979")
 
 # saving data frames as csv and rds
-write.csv(full_demographics, here::here("data-raw/wages-high-school-demo/data-frames/tidy_demographics.csv"))
-write_rds(as_tibble(full_demographics),
-          here::here("data-raw/wages-high-school-demo/data-frames/tidy_demographics.rds"),
-          compress = "xz")
-write.csv(demog_education, here::here("data-raw/wages-high-school-demo/data-frames/demog-education.csv"))
-write_rds(as_tibble(demog_education),
-          here::here("data-raw/wages-high-school-demo/data-frames/demog-education.rds"),
-          compress = "xz")
+#write.csv(full_demographics, here::here("data-raw/wages-high-school-demo/data-frames/tidy_demographics.csv"))
+#write_rds(as_tibble(full_demographics),
+          #here::here("data-raw/wages-high-school-demo/data-frames/tidy_demographics.rds"),
+          #compress = "xz")
+#write.csv(demog_education, here::here("data-raw/wages-high-school-demo/data-frames/demog-education.csv"))
+#write_rds(as_tibble(demog_education),
+          #here::here("data-raw/wages-high-school-demo/data-frames/demog-education.rds"),
+          #compress = "xz")
 
 
 # tidying employment data
@@ -1104,8 +1104,12 @@ tidy_employment <- bind_rows(tidy_employment_79,
                              tidy_employment_14,
                              tidy_employment_16)
 
-write.csv(tidy_employment, here::here("data-raw/wages-high-school-demo/data-frames/tidy_employment.csv"))
-write_rds(as_tibble(tidy_employment),
-          here::here("data-raw/wages-high-school-demo/data-frames/tidy_employment.rds"),
-          compress = "xz")
+#write.csv(tidy_employment, here::here("data-raw/wages-high-school-demo/data-frames/tidy_employment.csv"))
+#write_rds(as_tibble(tidy_employment),
+          #here::here("data-raw/wages-high-school-demo/data-frames/tidy_employment.rds"),
+          #compress = "xz")
 
+
+visdat::vis_dat(tidy_employment, warn_large_data = FALSE)
+
+filter(tidy_employment, !is.na(mean_hourly_wage))

### 2021-02-13 code_di_weighted.R
add label to extreme id and extreme observation

@@ -3,6 +3,7 @@ library(zoo)
 library(rstatix)
 
 
+
 # open the tidy raw data set
 wages <- read_csv("data-raw/wages-high-school-demo/data-frames/tidy_employment_weighted.csv")
 demog <- read_csv("data-raw/wages-high-school-demo/data-frames/tidy_demographics_v2.csv")
@@ -73,27 +74,11 @@ keep_me <- keep_me %>% filter(n > 4)
 wages_demog_hs <- wages_demog_hs %>%
   filter(id %in% keep_me$id)
 
+# Add the extreme value flag
 
-# calculate the IQR to create the extreme value flag
 
-# IQR_cut_off <- IQR(wages_demog_hs$mean_hourly_wage, na.rm = TRUE)*1.5
-# q1 <- quantile(wages_demog_hs$mean_hourly_wage, 0.25, na.rm = TRUE)
-# q3 <- quantile(wages_demog_hs$mean_hourly_wage, 0.75, na.rm = TRUE)
-#
-# lower_limit <- q1 - IQR_cut_off
-# upper_limit <- q3 + IQR_cut_off
-#
-# wages_demog_hs <- wages_demog_hs %>%
-#   mutate(is_extreme_val = ifelse(mean_hourly_wage < lower_limit |
-#                           mean_hourly_wage > upper_limit, TRUE,
-#                         FALSE)) %>%
-#   mutate(flag = case_when(is_wm == TRUE & is_extreme_val == FALSE ~ 1,
-#                           is_wm == TRUE & is_extreme_val == TRUE ~ 2,
-#                           is_wm == FALSE & is_extreme_val == FALSE ~ 3,
-#                           is_wm == FALSE & is_extreme_val == TRUE ~ 4))
-#
-# filtered <- filter(wages_demog_hs, is_extreme_val == FALSE)
-# summary(filtered$mean_hourly_wage)
+
+
 
 
 

### 2021-02-17 method_comparison.R
Compare robust lmm with extreme value.

@@ -1,0 +1,124 @@
+library(tidyverse)
+library(zoo)
+library(rstatix)
+library(robustlmm)
+library(modelr)
+library(patchwork)
+
+
+
+
+# open the tidy raw data set
+wages <- read_csv("data-raw/wages-high-school-demo/data-frames/tidy_employment_weighted.csv")
+demog <- read_csv("data-raw/wages-high-school-demo/data-frames/tidy_demographics_v2.csv")
+
+# join the wages information and the demographic information by case id.
+wages_demog <- left_join(wages, demog, by="id")
+
+
+# calculate the years in work force and the age of the subjects in 1979
+wages_demog <- wages_demog %>%
+  mutate(yr_hgc = as.numeric(yr_hgc)) %>%
+  mutate(years_in_workforce = year - yr_hgc) %>%
+  mutate(age_1979 = 1979 - (dob_year + 1900))
+
+# filter only the id with high school education
+wages_demog_hs <- wages_demog  %>% filter(grepl("GRADE", hgc))
+
+ggplot(wages_demog_hs, aes(x = mean_hourly_wage)) +
+  geom_boxplot()
+
+
+# get the number of observations
+n_obs <- wages_demog_hs %>% count(id)
+
+# sample some IDs
+
+set.seed(20210217)
+
+sample_id <- sample(unique(n_obs$id), 100)
+sample <- subset(wages_demog_hs, id %in% sample_id)
+
+# check the extreme value
+
+sample_ext <- sample %>%
+  group_by(id) %>%
+  identify_outliers("mean_hourly_wage") %>%
+  select(year, id, is.outlier, is.extreme)
+
+# check the influential value using robust linear mixed model
+
+rfm <- rlmer(mean_hourly_wage ~ (1|year) + (1|id), sample)
+summary(rfm)
+
+weight_obs <- as.vector(getME(rfm, "w_e"))
+
+
+# join the weight with the data
+sample_weight <- cbind(sample, weight_obs)
+
+
+# join the extreme value with the data
+wages_hs_sample <- left_join(sample_weight, sample_ext, by = c("id", "year"))
+
+# flag the extreme value and outlier flag based on IQR methos
+wages_hs_sample <- wages_hs_sample %>%
+  mutate(is.outlier = ifelse(is.na(is.outlier), FALSE, is.outlier),
+         is.extreme = ifelse(is.na(is.extreme), FALSE, is.extreme))
+
[…]

### 2021-02-19 method_comparison.R
Compare robust lm with other method.

@@ -4,6 +4,8 @@ library(rstatix)
 library(robustlmm)
 library(modelr)
 library(patchwork)
+library(MASS)
+library(broom)
 
 
 
@@ -34,7 +36,7 @@ n_obs <- wages_demog_hs %>% count(id)
 
 # sample some IDs
 
-set.seed(20210217)
+set.seed(20210218)
 
 sample_id <- sample(unique(n_obs$id), 100)
 sample <- subset(wages_demog_hs, id %in% sample_id)
@@ -44,7 +46,7 @@ sample <- subset(wages_demog_hs, id %in% sample_id)
 sample_ext <- sample %>%
   group_by(id) %>%
   identify_outliers("mean_hourly_wage") %>%
-  select(year, id, is.outlier, is.extreme)
+  dplyr::select(year, id, is.outlier, is.extreme)
 
 # check the influential value using robust linear mixed model
 
@@ -98,14 +100,14 @@ wages_hs_sample <- wages_hs_sample %>%
 
 # doing locf
 for_locf <- wages_hs_sample %>%
-  select(id, year, wages_clean_ext) %>%
+  dplyr::select(id, year, wages_clean_ext) %>%
   group_by(id) %>%
   na.locf()
 
 
 # join back the locf value to wages_demog_hs
 wages_hs_sample <- left_join(wages_hs_sample, for_locf, by = c("id", "year")) %>%
-  select(-wages_clean_ext.x) %>%
+  dplyr::select(-wages_clean_ext.x) %>%
   rename(wages_clean_ext = wages_clean_ext.y)
 
 
@@ -120,5 +122,131 @@ p3 <- ggplot(wages_hs_sample) +
                 group = id), alpha = 0.3)
 
 p1 + p2 + p3
+
+
+wages_sample_long <- wages_hs_sample %>%
+  dplyr::select(id, year, mean_hourly_wage, wages_cleaned_rlmm, wages_clean_ext) %>%
+  pivot_longer(c(-id, -year), names_to = "type", values_to = "wages")
+
+n_obs_sample <- wages_sample_long %>% count(id) %>%
+  filter(n > 30)
+
+set.seed(20210218)
+
+sample_plot <- sample(unique(n_obs_sample$id), 20)
+sample_plot <- subset(wages_sample_long, id %in% sample_plot)
+
+ggplot(sample_plot) +
+  geom_line(aes(x = year,
+                y = wages,
+                colour = type)) +
+  facet_wrap(~id)
+
+
+##### ROBUST LINEAR MODEL
+
[…]

### 2021-02-19 method_comparison.R
Edit the comparison graph.

@@ -230,7 +230,7 @@ wages_compare <- left_join(wages_ext, wages_rlm_dat, by = c("id", "year")) %>%
 n_obs_sample_2 <- wages_compare %>% count(id) %>%
   filter(n > 30)
 
-set.seed(123456789)
+set.seed(1234)
 
 sample_plot_2 <- sample(unique(n_obs_sample_2$id), 20)
 sample_plot_2 <- subset(wages_compare, id %in% sample_plot_2)

### 2021-02-20 comparing_scenario.Rmd
Change the code as the flag changed.
Change the method to rlm.

@@
   mutate(mean_hourly_wage = exp(ln_wages))
+load("wages_hs2020_weighted_updated2018.rda")
 ```
@@
                     regular = FALSE)
+
+wages_weighted_2018 <- 
 ```
@@
 print("Weighted data, extreme value excluded")
-get_summary(filter(wages_weighted, flag2 == "non_ext_val"))
+get_summary(filter(wages_weighted, is_extreme_val == FALSE))
 ```

### 2021-02-20 code_di_weighted.R

@@ -1,6 +1,7 @@
 library(tidyverse)
 library(zoo)
-library(rstatix)
+library(MASS)
+library(broom)
 
 
 
@@ -25,44 +26,6 @@ wages_demog_hs <- wages_demog  %>% filter(grepl("GRADE", hgc))
 ggplot(wages_demog_hs, aes(x = mean_hourly_wage)) +
   geom_boxplot()
 
-# replace the obs with mean_hourly_wage is.extreme == TRUE to be NA
-# since we will impute the missing values with the last observation carried forward (locf) for each id
-# we give the flag for that locf observation
-
-wages_demog_ext <- wages_demog_hs %>%
-  group_by(id) %>%
-  identify_outliers("mean_hourly_wage") %>%
-  filter(is.extreme == TRUE) %>%
-  select(year, id, is.extreme)
-
-
-
-
-wages_demog_hs <- left_join(wages_demog_hs, wages_demog_ext, by = c("id", "year"))
-
-wages_demog_hs <- wages_demog_hs %>%
-  mutate(is.extreme = ifelse(is.na(is.extreme), FALSE,
-                             is.extreme))
-
-wages_demog_hs <- wages_demog_hs %>%
-  mutate(mean_hourly_wage = ifelse(is.extreme == TRUE, NA, mean_hourly_wage),
-         is_locf = ifelse(is.na(mean_hourly_wage), TRUE, FALSE))
-
-
-# doing locf
-
-for_locf <- wages_demog_hs %>%
-  select(id, year, mean_hourly_wage) %>%
-  group_by(id) %>%
-  na.locf()
-
-
-# join back
-
-wages_demog_hs <- left_join(wages_demog_hs, for_locf, by = c("id", "year")) %>%
-  select(-mean_hourly_wage.x) %>%
-  rename(mean_hourly_wage = mean_hourly_wage.y)
-
 
 # calculate the number of observation
 keep_me <- wages_demog_hs %>% count(id)
@@ -74,19 +37,72 @@ keep_me <- keep_me %>% filter(n > 4)
 wages_demog_hs <- wages_demog_hs %>%
   filter(id %in% keep_me$id)
 
-# Add the extreme value flag
+# nesting the data by id to build a robust linear model
+by_id <- wages_demog_hs %>%
+  dplyr::select(id, year, mean_hourly_wage) %>%
+  group_by(id) %>%
+  nest()
+
+# build a robust linear model
+id_rlm <- by_id %>%
+  mutate(model = map(.x = data,
+                     .f = function(x){
[…]

### 2021-02-20 method_comparison.R

@@ -197,13 +197,15 @@ p4 <- ggplot(wages_rlm_dat) +
   geom_line(aes(x = year,
                 y = wages_rlm,
                 group = id),
-            alpha = 0.1)
+            alpha = 0.1) +
+  ggtitle("robust linear model")
 
 p5 <- ggplot(wages_rlm_dat) +
   geom_line(aes(x = year,
                 y = wages_original,
                 group = id),
-            alpha = 0.1)
+            alpha = 0.1) +
+  ggtitle("original data")
 
 wages_ext <- yowie::wages_hs2020 %>%
   rename(wages_excl_extreme = mean_hourly_wage)
@@ -212,11 +214,12 @@ p6 <- ggplot(wages_ext) +
   geom_line(aes(x = year,
                 y = wages_excl_extreme,
                 group = id),
-            alpha = 0.1)
+            alpha = 0.1) +
+  ggtitle("Exclude the extreme value")
 
 
 
-p4+p5+p6
+p5 + p6 + p4
 
 summary(wages_rlm_dat$wages_rlm)
 summary(wages_ext$wages_excl_extreme)
@@ -230,7 +233,7 @@ wages_compare <- left_join(wages_ext, wages_rlm_dat, by = c("id", "year")) %>%
 n_obs_sample_2 <- wages_compare %>% count(id) %>%
   filter(n > 30)
 
-set.seed(1234)
+set.seed(17021992)
 
 sample_plot_2 <- sample(unique(n_obs_sample_2$id), 20)
 sample_plot_2 <- subset(wages_compare, id %in% sample_plot_2)
@@ -242,6 +245,64 @@ ggplot(sample_plot_2) +
                 linetype = type),
             alpha = 1) +
   facet_wrap(~id)
+
+library(brolgar)
+library(naniar)
+wages_rlm <- as_tsibble(x = wages_rlm_dat,
+                    key = id,
+                    index = year,
+                    regular = FALSE)
+
+wages_lm <- key_slope(wages_rlm,
+                      wages_rlm ~ year) %>%
+  left_join(wages_rlm, by = "id")
+
+#calculate the prediction value
+wages_lm <- wages_lm %>%
+  mutate(pred = .intercept + (.slope_year*year))
+
+wages_res <- wages_lm %>%
+  mutate(resid = wages_rlm - pred)
+
+wages_lm_var <- wages_res %>%
+  group_by(id) %>%
+  summarise(var = var(resid,
[…]

### 2021-02-21 report/report.Rmd
Add the report's outline.

---
title: Longitudinal Data Cleaning - A Case of the NLSY79 Data
authors:
  - name: Dewi Amaliah
    thanks: Use footnote for providing further information about author (webpage, alternative address)---*not* for acknowledging funding agencies. Optional.
    department: Department of Econometric and Business Statistics
    affiliation: Monash University
    location: Clayton, VIC 3168
    email: dama0007@student.monash.edu
  - name: Dianne Cook
    department: Department of Econometric and Business Statistics
    affiliation: Monash University
    location: Clayton, VIC 3168
    email: dicook@monash.edu
abstract: |
  Enter the text of your abstract here.
keywords:
  - longitudinal data
  - data cleaning 
  - data tidying
  - robust linear model
bibliography: references.bib
biblio-style: unsrt
output: 
  rticles::arxiv_article:
    keep_tex: true
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(brolgar)
library(patchwork)
```

# Introduction

# The NLSY79

# The NLSY79 Data Cleaning

## Getting and Tyding the Data

The NLYS data is stored in a [database](https://www.nlsinfo.org/content/cohorts/nlsy79/get-data) and could be downloaded by variables. Several variables are available to be downloaded and could be browsed by index. For the wages data set, we only extracted these variables:

- Education, Training & Achievement Scores
    -  Education -> Summary measures -> All schools -> By year -> Highest grade completed
       - Downloaded all of the 78 variables in Highest grade completed.
- Employment
    - Summary measures -> By job -> Hours worked and Hourly wages
      - Downloaded all of the 427 variables in Hours worked
      - Downloaded all of the 151 variables in Hourly wages

      Both hours worked and hourly wages are recorded by the job, up to five jobs for each id/subject. 
- Household, Geography & Contextual Variables
    - Context -> Summary measures -> Basic demographics 
      - Downloaded year and month of birth, race, and sex variables.

      There are two versions of the year and month of birth, i.e. 1979 and 1981 data. We downloaded these two versions.

The downloaded data set came in a csv (wages-high-school-demo.csv) and dat (wages-high-school-demo.dat) format. We only used the .dat format. It also came along with these files:

- wages-high-school-demo.NLSY79: This is the tagset of variables that can be uploaded to the web site to recreate the data set.
- wages-high-school-demo.R: This is an R script provided automatically by the database for reading the data into R and convert the variables' name and its label into something more sensible. We utilized this code to do an initial data tidying. It produced two data set,  `categories_qnames` (the observations are stored in categorical/interval values) and `new_data_qnames` (the observations are stored in integer form). In this case, we only used the latter.

`new_data_qnames` is still untidy, where all of the variables for each year and each job being stored in one column, hence the data contains a huge number of columns (686 columns). Thus, the data should be tidied and wrangled first to extract the demographic and employment variables that we want to put in the final data set.

### Tidying demographic variables

Date of birth, age in 1979, gender, race, highest grade completed (factor and integer), and the year when the highest grade completed are the variables that we want to use in the cleaned data set. There are two versions of date of birth variable, which are the 1979 version and the 1981 version. In this case, we only used the 1979 data. We also did a consistency check for the 1979 and 1981 data and flag the inconsistent observations.

## Initial Data Analysis

## Robust Linear Model for Influential Values Treatment

# Conclusion

### 2021-02-21 report/report.tex
Add the report's outline.

\documentclass{article}

\usepackage{arxiv}

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{lmodern}        % https://github.com/rstudio/rticles/issues/343
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{lipsum}
\usepackage{graphicx}

\title{Longitudinal Data Cleaning - A Case of the NLSY79 Data}

\author{
    Dewi Amaliah
    \thanks{Use footnote for providing further information about author (webpage,
alternative address)---\emph{not} for acknowledging funding agencies.
Optional.}
   \\
    Department of Econometric and Business Statistics \\
    Monash University \\
  Clayton, VIC 3168 \\
  \texttt{\href{mailto:dama0007@student.monash.edu}{\nolinkurl{dama0007@student.monash.edu}}} \\
   \And
    Dianne Cook
   \\
    Department of Econometric and Business Statistics \\
    Monash University \\
  Clayton, VIC 3168 \\
  \texttt{\href{mailto:dicook@monash.edu}{\nolinkurl{dicook@monash.edu}}} \\
  }

% Pandoc citation processing

\begin{document}
\maketitle

\def\tightlist{}

\begin{abstract}
Enter the text of your abstract here.
\end{abstract}

\keywords{
    longitudinal data
   \and
    data cleaning
   \and
    data tidying
   \and
    robust linear model
  }

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

\hypertarget{the-nlsy79}{%
\section{The NLSY79}\label{the-nlsy79}}

\hypertarget{the-nlsy79-data-cleaning}{%
\section{The NLSY79 Data Cleaning}\label{the-nlsy79-data-cleaning}}

\hypertarget{getting-and-tyding-the-data}{%
\subsection{Getting and Tyding the
Data}\label{getting-and-tyding-the-data}}

The NLYS data is stored in a
\href{https://www.nlsinfo.org/content/cohorts/nlsy79/get-data}{database}
and could be downloaded by variables. Several variables are available to
be downloaded and could be browsed by index. For the wages data set, we
only extracted these variables:

\begin{itemize}
\tightlist
\item
  Education, Training \& Achievement Scores

  \begin{itemize}
  \tightlist
  \item
    Education -\textgreater{} Summary measures -\textgreater{} All
    schools -\textgreater{} By year -\textgreater{} Highest grade
    completed

    \begin{itemize}
    \tightlist
    \item
      Downloaded all of the 78 variables in Highest grade completed.
    \end{itemize}
  \end{itemize}
\item
  Employment

  \begin{itemize}
  \tightlist
  \item
    Summary measures -\textgreater{} By job -\textgreater{} Hours worked
    and Hourly wages

    \begin{itemize}
    \tightlist
    \item
      Downloaded all of the 427 variables in Hours worked
    \item
      Downloaded all of the 151 variables in Hourly wages
    \end{itemize}

    Both hours worked and hourly wages are recorded by the job, up to
    five jobs for each id/subject.
  \end{itemize}
\item
  Household, Geography \& Contextual Variables

  \begin{itemize}
  \tightlist
  \item
    Context -\textgreater{} Summary measures -\textgreater{} Basic
    demographics

    \begin{itemize}
    \tightlist
    \item
      Downloaded year and month of birth, race, and sex variables.
    \end{itemize}

    There are two versions of the year and month of birth, i.e.~1979 and
    1981 data. We downloaded these two versions.
  \end{itemize}
\end{itemize}

The downloaded data set came in a csv (wages-high-school-demo.csv) and
dat (wages-high-school-demo.dat) format. We only used the .dat format.
It also came along with these files:

\begin{itemize}
\tightlist
\item
  wages-high-school-demo.NLSY79: This is the tagset of variables that
  can be uploaded to the web site to recreate the data set.
\item
  wages-high-school-demo.R: This is an R script provided automatically
  by the database for reading the data into R and convert the variables'
  name and its label into something more sensible. We utilized this code
  to do an initial data tidying. It produced two data set,
  \texttt{categories\_qnames} (the observations are stored in
  categorical/interval values) and \texttt{new\_data\_qnames} (the
  observations are stored in integer form). In this case, we only used
  the latter.
\end{itemize}

\texttt{new\_data\_qnames} is still untidy, where all of the variables
for each year and each job being stored in one column, hence the data
contains a huge number of columns (686 columns). Thus, the data should
be tidied and wrangled first to extract the demographic and employment
variables that we want to put in the final data set.

\hypertarget{tidying-demographic-variables}{%
\subsubsection{Tidying demographic
variables}\label{tidying-demographic-variables}}

Date of birth, age in 1979, gender, race, highest grade completed
(factor and integer), and the year when the highest grade completed are
the variables that we want to use in the cleaned data set. There are two
versions of date of birth variable, which are the 1979 version and the
1981 version. In this case, we only used the 1979 data. We also did a
consistency check for the 1979 and 1981 data and flag the inconsistent
observations.

\hypertarget{initial-data-analysis}{%
\subsection{Initial Data Analysis}\label{initial-data-analysis}}

\hypertarget{robust-linear-model-for-influential-values-treatment}{%
\subsection{Robust Linear Model for Influential Values
Treatment}\label{robust-linear-model-for-influential-values-treatment}}

\hypertarget{conclusion}{%
\section{Conclusion}\label{conclusion}}

\bibliographystyle{unsrt}
\bibliography{references.bib}

\end{document}

### 2021-02-22 report/report.Rmd
Add getting and tidy data section.

@@
 ```{r setup, include = FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
 library(tidyverse)
@@
 library(patchwork)
+library(kableExtra)
 ```
@@
 
-The downloaded data set came in a csv (wages-high-school-demo.csv) and dat (wages-high-school-demo.dat) format. We only used the .dat format. It also came along with these files:
+The downloaded data set came in a csv (NLSY79.csv) and dat (NLSY79.dat) format. We only used the .dat format. It also came along with these files:
 
-- wages-high-school-demo.NLSY79: This is the tagset of variables that can be uploaded to the web site to recreate the data set.
-- wages-high-school-demo.R: This is an R script provided automatically by the database for reading the data into R and convert the variables' name and its label into something more sensible.
+- NLSY79.NLSY79: This is the tagset of variables that can be uploaded to the web site to recreate the data set.
+- NLSY79.R: This is an R script provided automatically by the database for reading the data into R and convert the variables' name and its label into something more sensible.
 We utilized this code to do an initial data tidying.
@@
 
+```{r, cache = TRUE}
+source(here::here("data-raw/NLSY79/NLSY79.R"))
+```
+
 `new_data_qnames` is still untidy, where all of the variables for each year and each job being stored in one column, hence the data contains a huge number of columns (686 columns).
 Thus, the data should be tidied and wrangled first to extract the demographic and employment variables that we want to put in the final data set.
+We mainly used `tidyr` [@tidyr], `dplyr` [@dplyr], and `stringr` [@stringr] to do this job.
 
@@
 
-Date of birth, age in 1979, gender, race, highest grade completed (factor and integer), and the year when the highest grade completed are the variables that we want to use in the cleaned data set.
-There are two versions of date of birth variable, which are the 1979 version and the 1981 version.
-In this case, we only used the 1979 data.
-We also did a consistency check for the 1979 and 1981 data and flag the inconsistent observations.
+`Age in 1979`, `gender`, `race`, `highest grade completed` (factor and integer), and the `year when the highest grade completed` are the variables that we want to use in the cleaned data set.
+
+Age in 1979 are derived from year of birth and month of birth variables in the raw data set.
+The variables have two versions, which are the 1979 version and the 1981 version.
+We only used the 1979 data and did a consistency check of those years and flag the inconsistent observations.
+
+```{r dob-tidy}
+## tidy the date of birth data
+dob_tidy <- new_data_qnames %>%
+  select(CASEID_1979,
+         starts_with("Q1-3_A~")) %>%
+  mutate(dob_year = case_when(
+                    # if the years recorded in both sets match, take 79 data
+                    `Q1-3_A~Y_1979` == `Q1-3_A~Y_1981` ~ `Q1-3_A~Y_1979`,
+                    # if the year in the 81 set is missing, take the 79 data
+                    is.na(`Q1-3_A~Y_1981`) ~ `Q1-3_A~Y_1979`,
+                    # if the sets don't match for dob year, take the 79 data
+                    `Q1-3_A~Y_1979` != `Q1-3_A~Y_1981` ~ `Q1-3_A~Y_1979`),
+        dob_month = case_when(
+                    # if months recorded in both sets match, take 79 data
+                    `Q1-3_A~M_1979` == `Q1-3_A~M_1981` ~ `Q1-3_A~M_1979`,
+                    # if month in 81 set is missing, take the 79 data
+                    is.na(`Q1-3_A~M_1981`) ~ `Q1-3_A~M_1979`,
+                    # if sets don't match for dob month, take the 79 data
+                    `Q1-3_A~M_1979` != `Q1-3_A~M_1981` ~ `Q1-3_A~M_1979`),
+        # flag if there is a conflict between dob recorded in 79 and 81
+        dob_conflict = case_when(     
+                      (`Q1-3_A~M_1979` != `Q1-3_A~M_1981`) & !is.na(`Q1-3_A~M_1981`)
+                      ~ TRUE,
+                      (`Q1-3_A~Y_1979` != `Q1-3_A~Y_1981`) & !is.na(`Q1-3_A~Y_1981`)
+                      ~ TRUE,
+                      (`Q1-3_A~Y_1979` == `Q1-3_A~Y_1981`) & 
+                      (`Q1-3_A~M_1979` == `Q1-3_A~M_1981`) ~ FALSE,
+                      is.na(`Q1-3_A~M_1981`) | is.na(`Q1-3_A~Y_1981`) ~ FALSE)) %>%
+  select(CASEID_1979,
+         dob_month,
+         dob_year,
+         dob_conflict)
+```
+
+For `gender` and `race`, we only renamed these variables.
+
+```{r demog-tidy}
+## tidy the gender and race variables
+demog_tidy <- categories_qnames %>%
+  select(CASEID_1979,
+         SAMPLE_RACE_78SCRN,
+         SAMPLE_SEX_1979) %>%
+  rename(gender = SAMPLE_SEX_1979,
+         race = SAMPLE_RACE_78SCRN)
+```
+
+The `highest grade completed` came with several version in each year.
+We chose the revised May data because the May data seemed to have less missing and presumably the revised data has been checked.
+However, there was no revised May data for 2012, 2014, 2016, and 2018 so we just used the ordinary May data.
+
+```{r tidy-grade}
+# tidy the grade 
+demog_education <- new_data_qnames %>%
+  as_tibble() %>%
+  # in 2018, the variable's name is Q3-4_2018, instead of HGC_2018
+  rename(HGC_2018 = `Q3-4_2018`) %>%
+  select(CASEID_1979,
+         starts_with("HGCREV"),
+         "HGC_2012",
+         "HGC_2014",
+         "HGC_2016",
+         "HGC_2018") %>%
+  pivot_longer(!CASEID_1979,
+               names_to = "var",
+               values_to = "grade") %>%
+  separate("var", c("var", "year"), sep = -4) %>%
+  filter(!is.na(grade)) %>%
+  select(-var)
+```
+
+In the final data, we only used the highest grade completed ever and derived the year of when the highest grade completed its categorical value.
+Therefore, we wrangled the highest grade completed in each year to mutate these variables.
+
+```{r tidy-hgc}
+
+## getting the highest year of completed education ever
+highest_year <- demog_education %>%
+  group_by(CASEID_1979) %>%
+  mutate(hgc_i = max(grade)) %>%
+  filter(hgc_i == grade) %>%
+  filter(year == first(year)) %>%
+  rename(yr_hgc = year) %>%
+  select(CASEID_1979, yr_hgc, hgc_i) %>%
+  ungroup() %>%
+  mutate('hgc' = ifelse(hgc_i == 0, "NONE", ifelse(hgc_i == 1, "1ST GRADE",
+                 ifelse(hgc_i == 2, "2ND GRADE", ifelse(hgc_i == 3, "3RD GRADE",
+                 ifelse(hgc_i >= 4 & hgc_i <= 12, paste0(hgc_i,"TH GRADE"),
+                 ifelse(hgc_i == 13, "1ST YEAR COL",
+                 ifelse(hgc_i == 14, "2ND YEAR COL",
+                 ifelse(hgc_i == 15, "3RD YEAR COL",
+                 ifelse(hgc_i == 95, "UNGRADED", 
+                        paste0((hgc_i - 12), "TH YEAR COL")))))))))))
+```
+
+Finally, we join all the tidy variables in a data set called `full_demographics`.
+
+```{r full-demog}
+full_demographics <- full_join(dob_tidy, demog_tidy, by = "CASEID_1979") %>%
+  full_join(highest_year, by = "CASEID_1979") %>%
+  rename("id" = "CASEID_1979")
+
+head(full_demographics)
+```
+
+### Tidying employment variables
+
+The employment data comprises of three variables, i.e.
+`total hours of work per week`, `number of jobs that an individual has`, and `mean hourly wage`.
+For hours worked per week, initially only one version per job, no choice from 1979 to 1987 (QES-52A).
+From 1988 onward, when we had more options, we chose the variable for total hours including time spent working from home (QES-52D).
+However, 1993 did not have all the five D variables (the first one and the last one were missing), so we used QES-52A variable instead.
+In addition, 2008 only had jobs 1-4 for the QES-52D variable (whereas the other years had 1-5), so we just used these.
+
+```{r tidy-hours}
+# make a list for years where we used the "QES-52A"
+year_A <- c(1979:1987, 1993)
+#function to get the hour of work
+get_hour <- function(year){
+  if(year %in% year_A){
+   temp <- new_data_qnames %>%
+    select(CASEID_1979,
+            starts_with("QES-52A") &
+              ends_with(as.character(year)))} 
+  else{
+    temp <- new_data_qnames %>%
+    select(CASEID_1979,
+            starts_with("QES-52D") &
+              ends_with(as.character(year)))} 
+    temp %>% 
+      pivot_longer(!CASEID_1979,
+                 names_to = "job",
+                 values_to = "hours_work") %>%
+      separate("job", c("job", "year"), sep = -4) %>%
+      mutate(job = paste0("job_", substr(job, 9, 10))) %>%
+      rename(id = CASEID_1979)
+}
+
+# list to save the iteration result
+hours <- list()
+# getting the hours of work of all observations
+for(ayear in c(1979:1994, 1996, 1998, 2000, 2002, 2004, 2006, 2008, 2010, 
+               2012, 2014, 2016, 2018)) {
+   hours[[ayear]] <- get_hour(ayear)
+}
+# unlist the hours of work
+hours_all <- bind_rows(!!!hours)
+```
+
+The same algorithm was also deployed to tidy the rate of wage by year and by ID.
+The difference is that the hourly rate had only one version of each year.
+The hours of work and the hourly rate were then joined to calculate the number of jobs that an ID has and their mean hourly wage.
+Some observations had 0 in their hourly rate, which is considered as invalid value.
+Thus, their hourly rate set to be N.A.
+
+```{r tidy-rate}
+get_rate <- function(year) {
+  new_data_qnames %>%
+     select(CASEID_1979,
+            starts_with("HRP") &
+              ends_with(as.character(year))) %>%
+    pivot_longer(!CASEID_1979, names_to = "job", values_to = "rate_per_hour") %>%
+    separate("job", c("job", "year"), sep = -4) %>%
+    mutate(job = paste0("job_0", substr(job, 4, 4))) %>%
+    rename(id = CASEID_1979)
+}
+rates <- list()
+for(ayear in c(1979:1994, 1996, 1998, 2000, 2002, 2004, 2006, 2008, 2010, 
+               2012, 2014, 2016, 2018)) {
+   rates[[ayear]] <- get_rate(ayear)
+}
+rates_all <- bind_rows(!!!rates)
+# join hours and rates variable
+hours_wages <- left_join(rates_all, 
+                         hours_all, 
+                         by = c("id", "year", "job")) %>%
+  # set the 0 value in rate_per_hour as NA 
+  mutate(rate_per_hour = ifelse(rate_per_hour == 0, NA,
+                                rate_per_hour))
+head(hours_wages)
+```
+
+Since our ultimate goal is to calculate the mean hourly wage, the number of jobs is calculate based on the availability of the `rate_per_hour` information.
+For example, the number of jobs of ID 1, based on `hours_work`, is two.
+However, since the information of hourly rate of `job_02` is not available, the number of job is considered as 1.
+
+Further, we calculated the mean hourly wage for each ID in each year using a weighted mean with the hours of work as the weight.
+However, there are a lot of missing value in `hours_work` variable.
+In that case, we only calculated the mean hourly wage based on arithmetic/regular mean method.
+Hence, we created a new variable to flag whether the mean hourly wage is a weighted or a regular mean.
+Additionally, if an ID only had one job, we directly used their hourly wages information and flagged it as an arithmetic mean.
+
+```{r}
+# calculate number of jobs that a person has in one year
+no_job <- hours_wages %>%
+  filter(!is.na(rate_per_hour)) %>%
+  group_by(id, year) %>%
+  summarise(no_jobs = length(rate_per_hour))
+
+# filter the observations with available rate per hour
+eligible_wages <- hours_wages %>%
+  filter(!is.na(rate_per_hour)) %>%
+  left_join(no_job, by = c("id", "year"))
+
+# calculate the mean_hourly_wage
+# flag1 = code 1 for weighted mean
+# code 0 for arithmetic mean
+mean_hourly_wage <- 
+  eligible_wages %>%
+  group_by(id, year) %>%
+  #calculate the weighted mean if the number of jobs > 1
+  mutate(wages = ifelse(no_jobs == 1, rate_per_hour/100,
+                        weighted.mean(rate_per_hour, hours_work, na.rm = TRUE)/100)) %>%
+  #give the flag if it the weighted mean
+  mutate(flag1 = ifelse(!is.na(wages) & no_jobs != 1, 1,
+                        0)) %>%
+  #calculate the arithmetic mean for the na
+  mutate(wages = ifelse(is.na(wages), mean(rate_per_hour)/100,
+                        wages)) %>%
+  group_by(id, year) %>%
+  summarise(wages = mean(wages),
+            total_hours = sum(hours_work),
+            number_of_jobs = mean(no_jobs),
+            flag1 = mean(flag1)) %>%
+  mutate(year = as.numeric(year)) %>%
+  ungroup() %>%
+  rename(mean_hourly_wage = wages) %>%
+  mutate(is_wm = ifelse(flag1 == 1, TRUE,
+                        FALSE)) %>%
+  dplyr::select(-flag1)
+
+head(mean_hourly_wage, n = 10)
+```
+
+The `mean_hourly_wage` and `full_demographic` data are then joined.
+We also filtered the data to only have the cohort who completed the education up to 12th grade and participated at least five rounds in the survey and save it to an object called `wages_demog_hs`.
+
+```{r}
+# join the wages information and the demographic information by case id.
+wages_demog <- left_join(mean_hourly_wage, full_demographics, by="id")
+# calculate the years in work force and the age of the subjects in 1979
+wages_demog <- wages_demog %>%
+  mutate(yr_hgc = as.numeric(yr_hgc)) %>%
+  mutate(years_in_workforce = year - yr_hgc) %>%
+  mutate(age_1979 = 1979 - (dob_year + 1900))
+# filter only the id with high school education
+wages_demog_hs <- wages_demog  %>% filter(grepl("GRADE", hgc))
+# calculate the number of observation
+keep_me <- wages_demog_hs %>%
+  count(id) %>%
+  filter(n > 4)
+wages_demog_hs <- wages_demog_hs %>%
+  filter(id %in% keep_me$id)
+```
 

### 2021-02-22 report/report.tex
Add getting and tidy data section.

@@
 
+\usepackage{color}
+\usepackage{fancyvrb}
+\newcommand{\VerbBar}{|}
+\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
+\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
+% Add ',fontsize=\small' for more characters per line
+\usepackage{framed}
+\definecolor{shadecolor}{RGB}{248,248,248}
+\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
+\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
+\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
+\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
+\newcommand{\BuiltInTok}[1]{#1}
+\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
+\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
+\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
+\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
+\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
+\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
+\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
+\newcommand{\ExtensionTok}[1]{#1}
+\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
+\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
+\newcommand{\ImportTok}[1]{#1}
+\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
+\newcommand{\NormalTok}[1]{#1}
+\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
+\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
+\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
+\newcommand{\RegionMarkerTok}[1]{#1}
+\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
+\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
+\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
+\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
+\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
+\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+
 % Pandoc citation processing
+
+\usepackage{booktabs}
+\usepackage{longtable}
+\usepackage{array}
+\usepackage{multirow}
+\usepackage{wrapfig}
+\usepackage{float}
+\usepackage{colortbl}
+\usepackage{pdflscape}
+\usepackage{tabu}
+\usepackage{threeparttable}
+\usepackage{threeparttablex}
+\usepackage[normalem]{ulem}
+\usepackage{makecell}
+\usepackage{xcolor}
 
@@
 
-The downloaded data set came in a csv (wages-high-school-demo.csv) and
-dat (wages-high-school-demo.dat) format. We only used the .dat format.
-It also came along with these files:
+The downloaded data set came in a csv (NLSY79.csv) and dat (NLSY79.dat)
+format. We only used the .dat format. It also came along with these
+files:
 
@@
 \item
-  wages-high-school-demo.NLSY79: This is the tagset of variables that
-  can be uploaded to the web site to recreate the data set.
+  NLSY79.NLSY79: This is the tagset of variables that can be uploaded to
+  the web site to recreate the data set.
 \item
-  wages-high-school-demo.R: This is an R script provided automatically
-  by the database for reading the data into R and convert the variables'
-  name and its label into something more sensible. We utilized this code
-  to do an initial data tidying. It produced two data set,
+  NLSY79.R: This is an R script provided automatically by the database
+  for reading the data into R and convert the variables' name and its
+  label into something more sensible. We utilized this code to do an
+  initial data tidying. It produced two data set,
   \texttt{categories\_qnames} (the observations are stored in
@@
 
+\begin{Shaded}
+\begin{Highlighting}[]
+\KeywordTok{source}\NormalTok{(here}\OperatorTok{::}\KeywordTok{here}\NormalTok{(}\StringTok{"data-raw/NLSY79/NLSY79.R"}\NormalTok{))}
+\end{Highlighting}
+\end{Shaded}
+
 \texttt{new\_data\_qnames} is still untidy, where all of the variables
@@
 be tidied and wrangled first to extract the demographic and employment
-variables that we want to put in the final data set.
+variables that we want to put in the final data set. We mainly used
+\texttt{tidyr} (Wickham 2020), \texttt{dplyr} (Wickham et al. 2020), and
+\texttt{stringr} (Wickham 2019) to do this job.
 
@@
 
-Date of birth, age in 1979, gender, race, highest grade completed
-(factor and integer), and the year when the highest grade completed are
-the variables that we want to use in the cleaned data set. There are two
-versions of date of birth variable, which are the 1979 version and the
-1981 version. In this case, we only used the 1979 data. We also did a
-consistency check for the 1979 and 1981 data and flag the inconsistent
-observations.
+\texttt{Age\ in\ 1979}, \texttt{gender}, \texttt{race},
+\texttt{highest\ grade\ completed} (factor and integer), and the
+\texttt{year\ when\ the\ highest\ grade\ completed} are the variables
+that we want to use in the cleaned data set.
+
+Age in 1979 are derived from year of birth and month of birth variables
+in the raw data set. The variables have two versions, which are the 1979
+version and the 1981 version. We only used the 1979 data and did a
+consistency check of those years and flag the inconsistent observations.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{## tidy the date of birth data}
+\NormalTok{dob_tidy <-}\StringTok{ }\NormalTok{new_data_qnames }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{select}\NormalTok{(CASEID_}\DecValTok{1979}\NormalTok{,}
+         \KeywordTok{starts_with}\NormalTok{(}\StringTok{"Q1-3_A~"}\NormalTok{)) }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{dob_year =} \KeywordTok{case_when}\NormalTok{(}
+                    \CommentTok{# if the years recorded in both sets match, take 79 data}
+                    \StringTok{`}\DataTypeTok{Q1-3_A~Y_1979}\StringTok{`} \OperatorTok{==}\StringTok{ `}\DataTypeTok{Q1-3_A~Y_1981}\StringTok{`} \OperatorTok{~}\StringTok{ `}\DataTypeTok{Q1-3_A~Y_1979}\StringTok{`}\NormalTok{,}
+                    \CommentTok{# if the year in the 81 set is missing, take the 79 data}
+                    \KeywordTok{is.na}\NormalTok{(}\StringTok{`}\DataTypeTok{Q1-3_A~Y_1981}\StringTok{`}\NormalTok{) }\OperatorTok{~}\StringTok{ `}\DataTypeTok{Q1-3_A~Y_1979}\StringTok{`}\NormalTok{,}
+                    \CommentTok{# if the sets don't match for dob year, take the 79 data}
+                    \StringTok{`}\DataTypeTok{Q1-3_A~Y_1979}\StringTok{`} \OperatorTok{!=}\StringTok{ `}\DataTypeTok{Q1-3_A~Y_1981}\StringTok{`} \OperatorTok{~}\StringTok{ `}\DataTypeTok{Q1-3_A~Y_1979}\StringTok{`}\NormalTok{),}
+        \DataTypeTok{dob_month =} \KeywordTok{case_when}\NormalTok{(}
+                    \CommentTok{# if months recorded in both sets match, take 79 data}
+                    \StringTok{`}\DataTypeTok{Q1-3_A~M_1979}\StringTok{`} \OperatorTok{==}\StringTok{ `}\DataTypeTok{Q1-3_A~M_1981}\StringTok{`} \OperatorTok{~}\StringTok{ `}\DataTypeTok{Q1-3_A~M_1979}\StringTok{`}\NormalTok{,}
+                    \CommentTok{# if month in 81 set is missing, take the 79 data}
+                    \KeywordTok{is.na}\NormalTok{(}\StringTok{`}\DataTypeTok{Q1-3_A~M_1981}\StringTok{`}\NormalTok{) }\OperatorTok{~}\StringTok{ `}\DataTypeTok{Q1-3_A~M_1979}\StringTok{`}\NormalTok{,}
+                    \CommentTok{# if sets don't match for dob month, take the 79 data}
+                    \StringTok{`}\DataTypeTok{Q1-3_A~M_1979}\StringTok{`} \OperatorTok{!=}\StringTok{ `}\DataTypeTok{Q1-3_A~M_1981}\StringTok{`} \OperatorTok{~}\StringTok{ `}\DataTypeTok{Q1-3_A~M_1979}\StringTok{`}\NormalTok{),}
+        \CommentTok{# flag if there is a conflict between dob recorded in 79 and 81}
+        \DataTypeTok{dob_conflict =} \KeywordTok{case_when}\NormalTok{(     }
+\NormalTok{                      (}\StringTok{`}\DataTypeTok{Q1-3_A~M_1979}\StringTok{`} \OperatorTok{!=}\StringTok{ `}\DataTypeTok{Q1-3_A~M_1981}\StringTok{`}\NormalTok{) }\OperatorTok{&}\StringTok{ }\OperatorTok{!}\KeywordTok{is.na}\NormalTok{(}\StringTok{`}\DataTypeTok{Q1-3_A~M_1981}\StringTok{`}\NormalTok{)}
+                      \OperatorTok{~}\StringTok{ }\OtherTok{TRUE}\NormalTok{,}
+\NormalTok{                      (}\StringTok{`}\DataTypeTok{Q1-3_A~Y_1979}\StringTok{`} \OperatorTok{!=}\StringTok{ `}\DataTypeTok{Q1-3_A~Y_1981}\StringTok{`}\NormalTok{) }\OperatorTok{&}\StringTok{ }\OperatorTok{!}\KeywordTok{is.na}\NormalTok{(}\StringTok{`}\DataTypeTok{Q1-3_A~Y_1981}\StringTok{`}\NormalTok{)}
+                      \OperatorTok{~}\StringTok{ }\OtherTok{TRUE}\NormalTok{,}
+\NormalTok{                      (}\StringTok{`}\DataTypeTok{Q1-3_A~Y_1979}\StringTok{`} \OperatorTok{==}\StringTok{ `}\DataTypeTok{Q1-3_A~Y_1981}\StringTok{`}\NormalTok{) }\OperatorTok{&}\StringTok{ }
+\StringTok{                      }\NormalTok{(}\StringTok{`}\DataTypeTok{Q1-3_A~M_1979}\StringTok{`} \OperatorTok{==}\StringTok{ `}\DataTypeTok{Q1-3_A~M_1981}\StringTok{`}\NormalTok{) }\OperatorTok{~}\StringTok{ }\OtherTok{FALSE}\NormalTok{,}
+                      \KeywordTok{is.na}\NormalTok{(}\StringTok{`}\DataTypeTok{Q1-3_A~M_1981}\StringTok{`}\NormalTok{) }\OperatorTok{|}\StringTok{ }\KeywordTok{is.na}\NormalTok{(}\StringTok{`}\DataTypeTok{Q1-3_A~Y_1981}\StringTok{`}\NormalTok{) }\OperatorTok{~}\StringTok{ }\OtherTok{FALSE}\NormalTok{)) }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{select}\NormalTok{(CASEID_}\DecValTok{1979}\NormalTok{,}
+\NormalTok{         dob_month,}
+\NormalTok{         dob_year,}
+\NormalTok{         dob_conflict)}
+\end{Highlighting}
+\end{Shaded}
+
+For \texttt{gender} and \texttt{race}, we only renamed these variables.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{## tidy the gender and race variables}
+\NormalTok{demog_tidy <-}\StringTok{ }\NormalTok{categories_qnames }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{select}\NormalTok{(CASEID_}\DecValTok{1979}\NormalTok{,}
+\NormalTok{         SAMPLE_RACE_78SCRN,}
+\NormalTok{         SAMPLE_SEX_}\DecValTok{1979}\NormalTok{) }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{rename}\NormalTok{(}\DataTypeTok{gender =}\NormalTok{ SAMPLE_SEX_}\DecValTok{1979}\NormalTok{,}
+         \DataTypeTok{race =}\NormalTok{ SAMPLE_RACE_78SCRN)}
+\end{Highlighting}
+\end{Shaded}
+
+The \texttt{highest\ grade\ completed} came with several version in each
+year. We chose the revised May data because the May data seemed to have
+less missing and presumably the revised data has been checked. However,
+there was no revised May data for 2012, 2014, 2016, and 2018 so we just
+used the ordinary May data.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{# tidy the grade }
+\NormalTok{demog_education <-}\StringTok{ }\NormalTok{new_data_qnames }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{as_tibble}\NormalTok{() }\OperatorTok{%>%}
+\StringTok{  }\CommentTok{# in 2018, the variable's name is Q3-4_2018, instead of HGC_2018}
+\StringTok{  }\KeywordTok{rename}\NormalTok{(}\DataTypeTok{HGC_2018 =} \StringTok{`}\DataTypeTok{Q3-4_2018}\StringTok{`}\NormalTok{) }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{select}\NormalTok{(CASEID_}\DecValTok{1979}\NormalTok{,}
+         \KeywordTok{starts_with}\NormalTok{(}\StringTok{"HGCREV"}\NormalTok{),}
+         \StringTok{"HGC_2012"}\NormalTok{,}
+         \StringTok{"HGC_2014"}\NormalTok{,}
+         \StringTok{"HGC_2016"}\NormalTok{,}
+         \StringTok{"HGC_2018"}\NormalTok{) }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{pivot_longer}\NormalTok{(}\OperatorTok{!}\NormalTok{CASEID_}\DecValTok{1979}\NormalTok{,}
+               \DataTypeTok{names_to =} \StringTok{"var"}\NormalTok{,}
+               \DataTypeTok{values_to =} \StringTok{"grade"}\NormalTok{) }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{separate}\NormalTok{(}\StringTok{"var"}\NormalTok{, }\KeywordTok{c}\NormalTok{(}\StringTok{"var"}\NormalTok{, }\StringTok{"year"}\NormalTok{), }\DataTypeTok{sep =} \DecValTok{-4}\NormalTok{) }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{filter}\NormalTok{(}\OperatorTok{!}\KeywordTok{is.na}\NormalTok{(grade)) }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{select}\NormalTok{(}\OperatorTok{-}\NormalTok{var)}
+\end{Highlighting}
+\end{Shaded}
+
+In the final data, we only used the highest grade completed ever and
+derived the year of when the highest grade completed its categorical
+value. Therefore, we wrangled the highest grade completed in each year
+to mutate these variables.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{## getting the highest year of completed education ever}
+\NormalTok{highest_year <-}\StringTok{ }\NormalTok{demog_education }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{group_by}\NormalTok{(CASEID_}\DecValTok{1979}\NormalTok{) }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{hgc_i =} \KeywordTok{max}\NormalTok{(grade)) }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{filter}\NormalTok{(hgc_i }\OperatorTok{==}\StringTok{ }\NormalTok{grade) }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{filter}\NormalTok{(year }\OperatorTok{==}\StringTok{ }\KeywordTok{first}\NormalTok{(year)) }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{rename}\NormalTok{(}\DataTypeTok{yr_hgc =}\NormalTok{ year) }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{select}\NormalTok{(CASEID_}\DecValTok{1979}\NormalTok{, yr_hgc, hgc_i) }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{ungroup}\NormalTok{() }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\StringTok{'hgc'}\NormalTok{ =}\StringTok{ }\KeywordTok{ifelse}\NormalTok{(hgc_i }\OperatorTok{==}\StringTok{ }\DecValTok{0}\NormalTok{, }\StringTok{"NONE"}\NormalTok{, }\KeywordTok{ifelse}\NormalTok{(hgc_i }\OperatorTok{==}\StringTok{ }\DecValTok{1}\NormalTok{, }\StringTok{"1ST GRADE"}\NormalTok{,}
+                 \KeywordTok{ifelse}\NormalTok{(hgc_i }\OperatorTok{==}\StringTok{ }\DecValTok{2}\NormalTok{, }\StringTok{"2ND GRADE"}\NormalTok{, }\KeywordTok{ifelse}\NormalTok{(hgc_i }\OperatorTok{==}\StringTok{ }\DecValTok{3}\NormalTok{, }\StringTok{"3RD GRADE"}\NormalTok{,}
+                 \KeywordTok{ifelse}\NormalTok{(hgc_i }\OperatorTok{>=}\StringTok{ }\DecValTok{4} \OperatorTok{&}\StringTok{ }\NormalTok{hgc_i }\OperatorTok{<=}\StringTok{ }\DecValTok{12}\NormalTok{, }\KeywordTok{paste0}\NormalTok{(hgc_i,}\StringTok{"TH GRADE"}\NormalTok{),}
+                 \KeywordTok{ifelse}\NormalTok{(hgc_i }\OperatorTok{==}\StringTok{ }\DecValTok{13}\NormalTok{, }\StringTok{"1ST YEAR COL"}\NormalTok{,}
+                 \KeywordTok{ifelse}\NormalTok{(hgc_i }\OperatorTok{==}\StringTok{ }\DecValTok{14}\NormalTok{, }\StringTok{"2ND YEAR COL"}\NormalTok{,}
+                 \KeywordTok{ifelse}\NormalTok{(hgc_i }\OperatorTok{==}\StringTok{ }\DecValTok{15}\NormalTok{, }\StringTok{"3RD YEAR COL"}\NormalTok{,}
+                 \KeywordTok{ifelse}\NormalTok{(hgc_i }\OperatorTok{==}\StringTok{ }\DecValTok{95}\NormalTok{, }\StringTok{"UNGRADED"}\NormalTok{, }
+                        \KeywordTok{paste0}\NormalTok{((hgc_i }\OperatorTok{-}\StringTok{ }\DecValTok{12}\NormalTok{), }\StringTok{"TH YEAR COL"}\NormalTok{)))))))))))}
+\end{Highlighting}
+\end{Shaded}
+
+Finally, we join all the tidy variables in a data set called
+\texttt{full\_demographics}.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{full_demographics <-}\StringTok{ }\KeywordTok{full_join}\NormalTok{(dob_tidy, demog_tidy, }\DataTypeTok{by =} \StringTok{"CASEID_1979"}\NormalTok{) }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{full_join}\NormalTok{(highest_year, }\DataTypeTok{by =} \StringTok{"CASEID_1979"}\NormalTok{) }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{rename}\NormalTok{(}\StringTok{"id"}\NormalTok{ =}\StringTok{ "CASEID_1979"}\NormalTok{)}
+
+\KeywordTok{head}\NormalTok{(full_demographics)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+##   id dob_month dob_year dob_conflict                    race gender yr_hgc
+## 1  1         9       58        FALSE NON-BLACK, NON-HISPANIC FEMALE   1979
+## 2  2         1       59        FALSE NON-BLACK, NON-HISPANIC FEMALE   1985
+## 3  3         8       61        FALSE NON-BLACK, NON-HISPANIC FEMALE   1993
+## 4  4         8       62        FALSE NON-BLACK, NON-HISPANIC FEMALE   1986
+## 5  5         7       59        FALSE NON-BLACK, NON-HISPANIC   MALE   1984
+## 6  6        10       60        FALSE NON-BLACK, NON-HISPANIC   MALE   1983
+##   hgc_i          hgc
+## 1    12   12TH GRADE
+## 2    12   12TH GRADE
+## 3    12   12TH GRADE
+## 4    14 2ND YEAR COL
+## 5    18 6TH YEAR COL
+## 6    16 4TH YEAR COL
+\end{verbatim}
+
+\hypertarget{tidying-employment-variables}{%
+\subsubsection{Tidying employment
+variables}\label{tidying-employment-variables}}
+
+The employment data comprises of three variables,
+i.e.~\texttt{total\ hours\ of\ work\ per\ week},
+\texttt{number\ of\ jobs\ that\ an\ individual\ has}, and
+\texttt{mean\ hourly\ wage}. For hours worked per week, initially only
+one version per job, no choice from 1979 to 1987 (QES-52A). From 1988
+onward, when we had more options, we chose the variable for total hours
+including time spent working from home (QES-52D). However, 1993 did not
+have all the five D variables (the first one and the last one were
+missing), so we used QES-52A variable instead. In addition, 2008 only
+had jobs 1-4 for the QES-52D variable (whereas the other years had 1-5),
+so we just used these.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{# make a list for years where we used the "QES-52A"}
+\NormalTok{year_A <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\DecValTok{1979}\OperatorTok{:}\DecValTok{1987}\NormalTok{, }\DecValTok{1993}\NormalTok{)}
+\CommentTok{#function to get the hour of work}
+\NormalTok{get_hour <-}\StringTok{ }\ControlFlowTok{function}\NormalTok{(year)\{}
+  \ControlFlowTok{if}\NormalTok{(year }\OperatorTok{%in%}\StringTok{ }\NormalTok{year_A)\{}
+\NormalTok{   temp <-}\StringTok{ }\NormalTok{new_data_qnames }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{select}\NormalTok{(CASEID_}\DecValTok{1979}\NormalTok{,}
+            \KeywordTok{starts_with}\NormalTok{(}\StringTok{"QES-52A"}\NormalTok{) }\OperatorTok{&}
+\StringTok{              }\KeywordTok{ends_with}\NormalTok{(}\KeywordTok{as.character}\NormalTok{(year)))\} }
+  \ControlFlowTok{else}\NormalTok{\{}
+\NormalTok{    temp <-}\StringTok{ }\NormalTok{new_data_qnames }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{select}\NormalTok{(CASEID_}\DecValTok{1979}\NormalTok{,}
+            \KeywordTok{starts_with}\NormalTok{(}\StringTok{"QES-52D"}\NormalTok{) }\OperatorTok{&}
+\StringTok{              }\KeywordTok{ends_with}\NormalTok{(}\KeywordTok{as.character}\NormalTok{(year)))\} }
+\NormalTok{    temp }\OperatorTok{%>%}\StringTok{ }
+\StringTok{      }\KeywordTok{pivot_longer}\NormalTok{(}\OperatorTok{!}\NormalTok{CASEID_}\DecValTok{1979}\NormalTok{,}
+                 \DataTypeTok{names_to =} \StringTok{"job"}\NormalTok{,}
+                 \DataTypeTok{values_to =} \StringTok{"hours_work"}\NormalTok{) }\OperatorTok{%>%}
+\StringTok{      }\KeywordTok{separate}\NormalTok{(}\StringTok{"job"}\NormalTok{, }\KeywordTok{c}\NormalTok{(}\StringTok{"job"}\NormalTok{, }\StringTok{"year"}\NormalTok{), }\DataTypeTok{sep =} \DecValTok{-4}\NormalTok{) }\OperatorTok{%>%}
+\StringTok{      }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{job =} \KeywordTok{paste0}\NormalTok{(}\StringTok{"job_"}\NormalTok{, }\KeywordTok{substr}\NormalTok{(job, }\DecValTok{9}\NormalTok{, }\DecValTok{10}\NormalTok{))) }\OperatorTok{%>%}
+\StringTok{      }\KeywordTok{rename}\NormalTok{(}\DataTypeTok{id =}\NormalTok{ CASEID_}\DecValTok{1979}\NormalTok{)}
+\NormalTok{\}}
+
+\CommentTok{# list to save the iteration result}
+\NormalTok{hours <-}\StringTok{ }\KeywordTok{list}\NormalTok{()}
+\CommentTok{# getting the hours of work of all observations}
+\ControlFlowTok{for}\NormalTok{(ayear }\ControlFlowTok{in} \KeywordTok{c}\NormalTok{(}\DecValTok{1979}\OperatorTok{:}\DecValTok{1994}\NormalTok{, }\DecValTok{1996}\NormalTok{, }\DecValTok{1998}\NormalTok{, }\DecValTok{2000}\NormalTok{, }\DecValTok{2002}\NormalTok{, }\DecValTok{2004}\NormalTok{, }\DecValTok{2006}\NormalTok{, }\DecValTok{2008}\NormalTok{, }\DecValTok{2010}\NormalTok{, }
+               \DecValTok{2012}\NormalTok{, }\DecValTok{2014}\NormalTok{, }\DecValTok{2016}\NormalTok{, }\DecValTok{2018}\NormalTok{)) \{}
+\NormalTok{   hours[[ayear]] <-}\StringTok{ }\KeywordTok{get_hour}\NormalTok{(ayear)}
+\NormalTok{\}}
+\CommentTok{# unlist the hours of work}
+\NormalTok{hours_all <-}\StringTok{ }\KeywordTok{bind_rows}\NormalTok{(}\OperatorTok{!!!}\NormalTok{hours)}
+\end{Highlighting}
+\end{Shaded}
+
+The same algorithm was also deployed to tidy the rate of wage by year
+and by ID. The difference is that the hourly rate had only one version
+of each year. The hours of work and the hourly rate were then joined to
+calculate the number of jobs that an ID has and their mean hourly wage.
+Some observations had 0 in their hourly rate, which is considered as
+invalid value. Thus, their hourly rate set to be N.A.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{get_rate <-}\StringTok{ }\ControlFlowTok{function}\NormalTok{(year) \{}
+\NormalTok{  new_data_qnames }\OperatorTok{%>%}
+\StringTok{     }\KeywordTok{select}\NormalTok{(CASEID_}\DecValTok{1979}\NormalTok{,}
+            \KeywordTok{starts_with}\NormalTok{(}\StringTok{"HRP"}\NormalTok{) }\OperatorTok{&}
+\StringTok{              }\KeywordTok{ends_with}\NormalTok{(}\KeywordTok{as.character}\NormalTok{(year))) }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{pivot_longer}\NormalTok{(}\OperatorTok{!}\NormalTok{CASEID_}\DecValTok{1979}\NormalTok{, }\DataTypeTok{names_to =} \StringTok{"job"}\NormalTok{, }\DataTypeTok{values_to =} \StringTok{"rate_per_hour"}\NormalTok{) }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{separate}\NormalTok{(}\StringTok{"job"}\NormalTok{, }\KeywordTok{c}\NormalTok{(}\StringTok{"job"}\NormalTok{, }\StringTok{"year"}\NormalTok{), }\DataTypeTok{sep =} \DecValTok{-4}\NormalTok{) }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{job =} \KeywordTok{paste0}\NormalTok{(}\StringTok{"job_0"}\NormalTok{, }\KeywordTok{substr}\NormalTok{(job, }\DecValTok{4}\NormalTok{, }\DecValTok{4}\NormalTok{))) }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{rename}\NormalTok{(}\DataTypeTok{id =}\NormalTok{ CASEID_}\DecValTok{1979}\NormalTok{)}
+\NormalTok{\}}
+\NormalTok{rates <-}\StringTok{ }\KeywordTok{list}\NormalTok{()}
+\ControlFlowTok{for}\NormalTok{(ayear }\ControlFlowTok{in} \KeywordTok{c}\NormalTok{(}\DecValTok{1979}\OperatorTok{:}\DecValTok{1994}\NormalTok{, }\DecValTok{1996}\NormalTok{, }\DecValTok{1998}\NormalTok{, }\DecValTok{2000}\NormalTok{, }\DecValTok{2002}\NormalTok{, }\DecValTok{2004}\NormalTok{, }\DecValTok{2006}\NormalTok{, }\DecValTok{2008}\NormalTok{, }\DecValTok{2010}\NormalTok{, }
+               \DecValTok{2012}\NormalTok{, }\DecValTok{2014}\NormalTok{, }\DecValTok{2016}\NormalTok{, }\DecValTok{2018}\NormalTok{)) \{}
+\NormalTok{   rates[[ayear]] <-}\StringTok{ }\KeywordTok{get_rate}\NormalTok{(ayear)}
+\NormalTok{\}}
+\NormalTok{rates_all <-}\StringTok{ }\KeywordTok{bind_rows}\NormalTok{(}\OperatorTok{!!!}\NormalTok{rates)}
+\CommentTok{# join hours and rates variable}
+\NormalTok{hours_wages <-}\StringTok{ }\KeywordTok{left_join}\NormalTok{(rates_all, }
+\NormalTok{                         hours_all, }
+                         \DataTypeTok{by =} \KeywordTok{c}\NormalTok{(}\StringTok{"id"}\NormalTok{, }\StringTok{"year"}\NormalTok{, }\StringTok{"job"}\NormalTok{)) }\OperatorTok{%>%}
+\StringTok{  }\CommentTok{# set the 0 value in rate_per_hour as NA }
+\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{rate_per_hour =} \KeywordTok{ifelse}\NormalTok{(rate_per_hour }\OperatorTok{==}\StringTok{ }\DecValTok{0}\NormalTok{, }\OtherTok{NA}\NormalTok{,}
+\NormalTok{                                rate_per_hour))}
+\KeywordTok{head}\NormalTok{(hours_wages)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+## # A tibble: 6 x 5
+##      id job    year  rate_per_hour hours_work
+##   <int> <chr>  <chr>         <int>      <int>
+## 1     1 job_01 1979            328         38
+## 2     1 job_02 1979             NA         15
+## 3     1 job_03 1979             NA         NA
+## 4     1 job_04 1979             NA         NA
+## 5     1 job_05 1979             NA         NA
+## 6     2 job_01 1979            385         35
+\end{verbatim}
+
+Since our ultimate goal is to calculate the mean hourly wage, the number
+of jobs is calculate based on the availability of the
+\texttt{rate\_per\_hour} information. For example, the number of jobs of
+ID 1, based on \texttt{hours\_work}, is two. However, since the
+information of hourly rate of \texttt{job\_02} is not available, the
+number of job is considered as 1.
+
+Further, we calculated the mean hourly wage for each ID in each year
+using a weighted mean with the hours of work as the weight. However,
+there are a lot of missing value in \texttt{hours\_work} variable. In
+that case, we only calculated the mean hourly wage based on
+arithmetic/regular mean method. Hence, we created a new variable to flag
+whether the mean hourly wage is a weighted or a regular mean.
+Additionally, if an ID only had one job, we directly used their hourly
+wages information and flagged it as an arithmetic mean.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{# calculate number of jobs that a person has in one year}
+\NormalTok{no_job <-}\StringTok{ }\NormalTok{hours_wages }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{filter}\NormalTok{(}\OperatorTok{!}\KeywordTok{is.na}\NormalTok{(rate_per_hour)) }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{group_by}\NormalTok{(id, year) }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{summarise}\NormalTok{(}\DataTypeTok{no_jobs =} \KeywordTok{length}\NormalTok{(rate_per_hour))}
+
+\CommentTok{# filter the observations with available rate per hour}
+\NormalTok{eligible_wages <-}\StringTok{ }\NormalTok{hours_wages }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{filter}\NormalTok{(}\OperatorTok{!}\KeywordTok{is.na}\NormalTok{(rate_per_hour)) }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{left_join}\NormalTok{(no_job, }\DataTypeTok{by =} \KeywordTok{c}\NormalTok{(}\StringTok{"id"}\NormalTok{, }\StringTok{"year"}\NormalTok{)) }
+
+\CommentTok{# calculate the mean_hourly_wage}
+\CommentTok{# flag1 = code 1 for weighted mean}
+\CommentTok{# code 0 for arithmetic mean}
+\NormalTok{mean_hourly_wage <-}\StringTok{ }
+\StringTok{  }\NormalTok{eligible_wages }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{group_by}\NormalTok{(id, year) }\OperatorTok{%>%}
+\StringTok{  }\CommentTok{#calculate the weighted mean if the number of jobs > 1}
+\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{wages =} \KeywordTok{ifelse}\NormalTok{(no_jobs }\OperatorTok{==}\StringTok{ }\DecValTok{1}\NormalTok{, rate_per_hour}\OperatorTok{/}\DecValTok{100}\NormalTok{,}
+                        \KeywordTok{weighted.mean}\NormalTok{(rate_per_hour, hours_work, }\DataTypeTok{na.rm =} \OtherTok{TRUE}\NormalTok{)}\OperatorTok{/}\DecValTok{100}\NormalTok{)) }\OperatorTok{%>%}
+\StringTok{  }\CommentTok{#give the flag if it the weighted mean}
+\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{flag1 =} \KeywordTok{ifelse}\NormalTok{(}\OperatorTok{!}\KeywordTok{is.na}\NormalTok{(wages) }\OperatorTok{&}\StringTok{ }\NormalTok{no_jobs }\OperatorTok{!=}\StringTok{ }\DecValTok{1}\NormalTok{, }\DecValTok{1}\NormalTok{,}
+                        \DecValTok{0}\NormalTok{)) }\OperatorTok{%>%}
+\StringTok{  }\CommentTok{#calculate the arithmetic mean for the na}
+\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{wages =} \KeywordTok{ifelse}\NormalTok{(}\KeywordTok{is.na}\NormalTok{(wages), }\KeywordTok{mean}\NormalTok{(rate_per_hour)}\OperatorTok{/}\DecValTok{100}\NormalTok{,}
+\NormalTok{                        wages)) }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{group_by}\NormalTok{(id, year) }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{summarise}\NormalTok{(}\DataTypeTok{wages =} \KeywordTok{mean}\NormalTok{(wages),}
+            \DataTypeTok{total_hours =} \KeywordTok{sum}\NormalTok{(hours_work),}
+            \DataTypeTok{number_of_jobs =} \KeywordTok{mean}\NormalTok{(no_jobs),}
+            \DataTypeTok{flag1 =} \KeywordTok{mean}\NormalTok{(flag1)) }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{year =} \KeywordTok{as.numeric}\NormalTok{(year)) }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{ungroup}\NormalTok{() }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{rename}\NormalTok{(}\DataTypeTok{mean_hourly_wage =}\NormalTok{ wages) }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{is_wm =} \KeywordTok{ifelse}\NormalTok{(flag1 }\OperatorTok{==}\StringTok{ }\DecValTok{1}\NormalTok{, }\OtherTok{TRUE}\NormalTok{,}
+                        \OtherTok{FALSE}\NormalTok{)) }\OperatorTok{%>%}
+\StringTok{  }\NormalTok{dplyr}\OperatorTok{::}\KeywordTok{select}\NormalTok{(}\OperatorTok{-}\NormalTok{flag1)}
+
+\KeywordTok{head}\NormalTok{(mean_hourly_wage, }\DataTypeTok{n =} \DecValTok{10}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+## # A tibble: 10 x 6
+##       id  year mean_hourly_wage total_hours number_of_jobs is_wm
+##    <int> <dbl>            <dbl>       <int>          <dbl> <lgl>
+##  1     1  1979             3.28          38              1 FALSE
+##  2     1  1981             3.61          NA              1 FALSE
+##  3     2  1979             3.85          35              1 FALSE
+##  4     2  1980             4.57          NA              1 FALSE
+##  5     2  1981             5.14          NA              1 FALSE
+##  6     2  1982             5.71          35              1 FALSE
+##  7     2  1983             5.71          NA              1 FALSE
+##  8     2  1984             5.14          NA              1 FALSE
+##  9     2  1985             7.71          NA              1 FALSE
+## 10     2  1986             7.69          NA              1 FALSE
+\end{verbatim}
+
+The \texttt{mean\_hourly\_wage} and \texttt{full\_demographic} data are
+then joined. We also filtered the data to only have the cohort who
+completed the education up to 12th grade and participated at least five
+rounds in the survey.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{# join the wages information and the demographic information by case id.}
+\NormalTok{wages_demog <-}\StringTok{ }\KeywordTok{left_join}\NormalTok{(mean_hourly_wage, full_demographics, }\DataTypeTok{by=}\StringTok{"id"}\NormalTok{)}
+\CommentTok{# calculate the years in work force and the age of the subjects in 1979}
+\NormalTok{wages_demog <-}\StringTok{ }\NormalTok{wages_demog }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{yr_hgc =} \KeywordTok{as.numeric}\NormalTok{(yr_hgc)) }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{years_in_workforce =}\NormalTok{ year }\OperatorTok{-}\StringTok{ }\NormalTok{yr_hgc) }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{age_1979 =} \DecValTok{1979} \OperatorTok{-}\StringTok{ }\NormalTok{(dob_year }\OperatorTok{+}\StringTok{ }\DecValTok{1900}\NormalTok{))}
+\CommentTok{# filter only the id with high school education}
+\NormalTok{wages_demog_hs <-}\StringTok{ }\NormalTok{wages_demog  }\OperatorTok{%>%}\StringTok{ }\KeywordTok{filter}\NormalTok{(}\KeywordTok{grepl}\NormalTok{(}\StringTok{"GRADE"}\NormalTok{, hgc))}
+\CommentTok{# calculate the number of observation}
+\NormalTok{keep_me <-}\StringTok{ }\NormalTok{wages_demog_hs }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{count}\NormalTok{(id) }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{filter}\NormalTok{(n }\OperatorTok{>}\StringTok{ }\DecValTok{4}\NormalTok{)}
+\NormalTok{wages_demog_hs <-}\StringTok{ }\NormalTok{wages_demog_hs }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{filter}\NormalTok{(id }\OperatorTok{%in%}\StringTok{ }\NormalTok{keep_me}\OperatorTok{$}\NormalTok{id)}
+\end{Highlighting}
+\end{Shaded}
 
@@
 \hypertarget{conclusion}{%
-\section{Conclusion}\label{conclusion}}
+\section*{Conclusion}\label{conclusion}}
+\addcontentsline{toc}{section}{Conclusion}
+
+\hypertarget{refs}{}
+\leavevmode\hypertarget{ref-stringr}{}%
+Wickham, Hadley. 2019. \emph{Stringr: Simple, Consistent Wrappers for
+Common String Operations}.
+\url{https://CRAN.R-project.org/package=stringr}.
+
+\leavevmode\hypertarget{ref-tidyr}{}%
+---------. 2020. \emph{Tidyr: Tidy Messy Data}.
+\url{https://CRAN.R-project.org/package=tidyr}.
+
+\leavevmode\hypertarget{ref-dplyr}{}%
+Wickham, Hadley, Romain François, Lionel Henry, and Kirill Müller. 2020.
+\emph{Dplyr: A Grammar of Data Manipulation}.
+\url{https://CRAN.R-project.org/package=dplyr}.
 

### 2021-02-22 report/report.Rmd
Add part of IDA.

@@
 library(kableExtra)
+library(MASS)
 ```
@@
 dob_tidy <- new_data_qnames %>%
-  select(CASEID_1979,
+  dplyr::select(CASEID_1979,
          starts_with("Q1-3_A~")) %>%
@@
                       is.na(`Q1-3_A~M_1981`) | is.na(`Q1-3_A~Y_1981`) ~ FALSE)) %>%
-  select(CASEID_1979,
+  dplyr::select(CASEID_1979,
          dob_month,
@@
 demog_tidy <- categories_qnames %>%
-  select(CASEID_1979,
+  dplyr::select(CASEID_1979,
          SAMPLE_RACE_78SCRN,
@@
   rename(HGC_2018 = `Q3-4_2018`) %>%
-  select(CASEID_1979,
+  dplyr::select(CASEID_1979,
          starts_with("HGCREV"),
@@
   filter(!is.na(grade)) %>%
-  select(-var)
+  dplyr::select(-var)
 ```
@@
   rename(yr_hgc = year) %>%
-  select(CASEID_1979, yr_hgc, hgc_i) %>%
+  dplyr::select(CASEID_1979, yr_hgc, hgc_i) %>%
   ungroup() %>%
@@
    temp <- new_data_qnames %>%
-    select(CASEID_1979,
+    dplyr::select(CASEID_1979,
             starts_with("QES-52A") &
@@
     temp <- new_data_qnames %>%
-    select(CASEID_1979,
+    dplyr::select(CASEID_1979,
             starts_with("QES-52D") &
@@
   new_data_qnames %>%
-     select(CASEID_1979,
+    dplyr::select(CASEID_1979,
             starts_with("HRP") &
@@
 
-## Robust Linear Model for Influential Values Treatment
+According to @HuebnerMariannePhD2016Asat, Initial Data Analysis (IDA) is the step of inspecting and screening the data after being collected to ensure that the data is clean, valid, and ready to be deployed in the later formal statistical analysis.
+Moreover, @Chatfield1985TIEo argued that the two main objectives of IDA is data description, which is to assess the structure and the quality of the data; and model formulation without any formal statistical inference.
 
-# Conclusion
+In this paper, we conducted an IDA or a preliminary data analysis to assess the consistency of the data with the cohort information that is provided by the NLSY.
+In addition, we also aimed to find the anomaly in the wages values using this approach.
+We mainly used graphical summary to do the IDA using `ggplot2`[@ggplot2] and `brolgar` [@brolgar].
+
+As stated previously, the respondents' ages ranged from 12 to 22 when first interviewed in 1979.
+Hence, we would like to validate whether all of the respondents were in this range.
+For the demographic data validation, we used the `full_demographic` i.e.
+the data with the survey years 1979 sample.
+
+```{r valid-age, echo = FALSE, fig.cap="Subjects' Age Distribution"}
+age <- yowie::demographic_nlsy79 %>%
+  ggplot(aes(x = age_1979)) +
+  geom_bar(fill = "#45B39D")
+
+gender <- yowie::demographic_nlsy79 %>%
+  ggplot(aes(x = gender,
+             fill = gender)) +
+  geom_bar() +
+  geom_text(aes(label = ..count..), stat = "count", vjust = 1.5, colour = "white")+
+  scale_fill_manual("gender", values = c("MALE" = "#2471A3", "FEMALE" = "#9B59B6"))+
+  theme(legend.position = "none")
+
+race <- yowie::demographic_nlsy79 %>%
+  ggplot(aes(x = race,
+             fill = race)) +
+  geom_bar() +
+  geom_text(aes(label = ..count..), stat = "count", vjust = 1.5, colour = "white") +
+  theme(legend.position = "none",
+        axis.text.x = element_text(angle = 10))
+
+age / (gender|race)
+```
+
+### Robust Linear Model for Influential Values Treatment
+
+# Summary

### 2021-02-22 report/report.tex
Add part of IDA.

@@
 \NormalTok{dob_tidy <-}\StringTok{ }\NormalTok{new_data_qnames }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{select}\NormalTok{(CASEID_}\DecValTok{1979}\NormalTok{,}
+\StringTok{  }\NormalTok{dplyr}\OperatorTok{::}\KeywordTok{select}\NormalTok{(CASEID_}\DecValTok{1979}\NormalTok{,}
          \KeywordTok{starts_with}\NormalTok{(}\StringTok{"Q1-3_A~"}\NormalTok{)) }\OperatorTok{%>%}
@@
                       \KeywordTok{is.na}\NormalTok{(}\StringTok{`}\DataTypeTok{Q1-3_A~M_1981}\StringTok{`}\NormalTok{) }\OperatorTok{|}\StringTok{ }\KeywordTok{is.na}\NormalTok{(}\StringTok{`}\DataTypeTok{Q1-3_A~Y_1981}\StringTok{`}\NormalTok{) }\OperatorTok{~}\StringTok{ }\OtherTok{FALSE}\NormalTok{)) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{select}\NormalTok{(CASEID_}\DecValTok{1979}\NormalTok{,}
+\StringTok{  }\NormalTok{dplyr}\OperatorTok{::}\KeywordTok{select}\NormalTok{(CASEID_}\DecValTok{1979}\NormalTok{,}
 \NormalTok{         dob_month,}
@@
 \NormalTok{demog_tidy <-}\StringTok{ }\NormalTok{categories_qnames }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{select}\NormalTok{(CASEID_}\DecValTok{1979}\NormalTok{,}
+\StringTok{  }\NormalTok{dplyr}\OperatorTok{::}\KeywordTok{select}\NormalTok{(CASEID_}\DecValTok{1979}\NormalTok{,}
 \NormalTok{         SAMPLE_RACE_78SCRN,}
@@
 \StringTok{  }\KeywordTok{rename}\NormalTok{(}\DataTypeTok{HGC_2018 =} \StringTok{`}\DataTypeTok{Q3-4_2018}\StringTok{`}\NormalTok{) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{select}\NormalTok{(CASEID_}\DecValTok{1979}\NormalTok{,}
+\StringTok{  }\NormalTok{dplyr}\OperatorTok{::}\KeywordTok{select}\NormalTok{(CASEID_}\DecValTok{1979}\NormalTok{,}
          \KeywordTok{starts_with}\NormalTok{(}\StringTok{"HGCREV"}\NormalTok{),}
@@
 \StringTok{  }\KeywordTok{filter}\NormalTok{(}\OperatorTok{!}\KeywordTok{is.na}\NormalTok{(grade)) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{select}\NormalTok{(}\OperatorTok{-}\NormalTok{var)}
+\StringTok{  }\NormalTok{dplyr}\OperatorTok{::}\KeywordTok{select}\NormalTok{(}\OperatorTok{-}\NormalTok{var)}
 \end{Highlighting}
@@
 \StringTok{  }\KeywordTok{rename}\NormalTok{(}\DataTypeTok{yr_hgc =}\NormalTok{ year) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{select}\NormalTok{(CASEID_}\DecValTok{1979}\NormalTok{, yr_hgc, hgc_i) }\OperatorTok{%>%}
+\StringTok{  }\NormalTok{dplyr}\OperatorTok{::}\KeywordTok{select}\NormalTok{(CASEID_}\DecValTok{1979}\NormalTok{, yr_hgc, hgc_i) }\OperatorTok{%>%}
 \StringTok{  }\KeywordTok{ungroup}\NormalTok{() }\OperatorTok{%>%}
@@
 \NormalTok{   temp <-}\StringTok{ }\NormalTok{new_data_qnames }\OperatorTok{%>%}
-\StringTok{    }\KeywordTok{select}\NormalTok{(CASEID_}\DecValTok{1979}\NormalTok{,}
+\StringTok{    }\NormalTok{dplyr}\OperatorTok{::}\KeywordTok{select}\NormalTok{(CASEID_}\DecValTok{1979}\NormalTok{,}
             \KeywordTok{starts_with}\NormalTok{(}\StringTok{"QES-52A"}\NormalTok{) }\OperatorTok{&}
@@
 \NormalTok{    temp <-}\StringTok{ }\NormalTok{new_data_qnames }\OperatorTok{%>%}
-\StringTok{    }\KeywordTok{select}\NormalTok{(CASEID_}\DecValTok{1979}\NormalTok{,}
+\StringTok{    }\NormalTok{dplyr}\OperatorTok{::}\KeywordTok{select}\NormalTok{(CASEID_}\DecValTok{1979}\NormalTok{,}
             \KeywordTok{starts_with}\NormalTok{(}\StringTok{"QES-52D"}\NormalTok{) }\OperatorTok{&}
@@
 \NormalTok{  new_data_qnames }\OperatorTok{%>%}
-\StringTok{     }\KeywordTok{select}\NormalTok{(CASEID_}\DecValTok{1979}\NormalTok{,}
+\StringTok{    }\NormalTok{dplyr}\OperatorTok{::}\KeywordTok{select}\NormalTok{(CASEID_}\DecValTok{1979}\NormalTok{,}
             \KeywordTok{starts_with}\NormalTok{(}\StringTok{"HRP"}\NormalTok{) }\OperatorTok{&}
@@
 completed the education up to 12th grade and participated at least five
-rounds in the survey.
+rounds in the survey and save it to an object called
+\texttt{wages\_demog\_hs}.
 
@@
 
+According to Huebner, Vach, and Cessie (2016), Initial Data Analysis
+(IDA) is the step of inspecting and screening the data after being
+collected to ensure that the data is clean, valid, and ready to be
+deployed in the later formal statistical analysis. Moreover, Chatfield
+(1985) argued that the two main objectives of IDA is data description,
+which is to assess the structure and the quality of the data; and model
+formulation without any formal statistical inference.
+
+In this paper, we conducted an IDA or a preliminary data analysis to
+assess the consistency of the data with the cohort information that is
+provided by the NLSY. In addition, we also aimed to find the anomaly in
+the wages values using this approach. We mainly used graphical summary
+to do the IDA using \texttt{ggplot2}(Wickham 2016) and \texttt{brolgar}
+(Tierney, Cook, and Prvan 2020).
+
+As stated previously, the respondents' ages ranged from 12 to 22 when
+first interviewed in 1979. Hence, we would like to validate whether all
+of the respondents were in this range. For the demographic data
+validation, we used the \texttt{full\_demographic} i.e.~the data with
+the survey years 1979 sample.
+
+\begin{figure}
+\centering
+\includegraphics{report_files/figure-latex/valid-age-1.pdf}
+\caption{Subjects' Age Distribution}
+\end{figure}
+
 \hypertarget{robust-linear-model-for-influential-values-treatment}{%
-\subsection{Robust Linear Model for Influential Values
+\subsubsection{Robust Linear Model for Influential Values
 Treatment}\label{robust-linear-model-for-influential-values-treatment}}
 
-\hypertarget{conclusion}{%
-\section*{Conclusion}\label{conclusion}}
-\addcontentsline{toc}{section}{Conclusion}
+\hypertarget{summary}{%
+\section*{Summary}\label{summary}}
+\addcontentsline{toc}{section}{Summary}
 
 \hypertarget{refs}{}
+\leavevmode\hypertarget{ref-Chatfield1985TIEo}{}%
+Chatfield, C. 1985. ``The Initial Examination of Data.'' \emph{Journal
+of the Royal Statistical Society. Series A. General} 148 (3): 214--53.
+
+\leavevmode\hypertarget{ref-HuebnerMariannePhD2016Asat}{}%
+Huebner, PhD, Marianne, Dr rer. nat Vach Werner, and PhD le Cessie
+Saskia. 2016. ``A Systematic Approach to Initial Data Analysis Is Good
+Research Practice.'' \emph{The Journal of Thoracic and Cardiovascular
+Surgery} 151 (1): 25--27.
+
+\leavevmode\hypertarget{ref-brolgar}{}%
+Tierney, Nicholas, Di Cook, and Tania Prvan. 2020. \emph{Brolgar: BRowse
+over Longitudinal Data Graphically and Analytically in R}.
+\url{https://github.com/njtierney/brolgar}.
+
+\leavevmode\hypertarget{ref-ggplot2}{}%
+Wickham, Hadley. 2016. \emph{Ggplot2: Elegant Graphics for Data
+Analysis}. Springer-Verlag New York.
+\url{https://ggplot2.tidyverse.org}.
+
 \leavevmode\hypertarget{ref-stringr}{}%
-Wickham, Hadley. 2019. \emph{Stringr: Simple, Consistent Wrappers for
-Common String Operations}.
-\url{https://CRAN.R-project.org/package=stringr}.
+---------. 2019. \emph{Stringr: Simple, Consistent Wrappers for Common
+String Operations}. \url{https://CRAN.R-project.org/package=stringr}.
 

### 2021-02-23 report/report.Rmd
Add plots to IDA section.

@@
 Hence, we would like to validate whether all of the respondents were in this range.
-For the demographic data validation, we used the `full_demographic` i.e.
+Additionally, the [NLSY](https://www.nlsinfo.org/content/cohorts/nlsy79/intro-to-the-sample/nlsy79-sample-introduction) also provided the number of the survey cohort by their gender (6,403 males and 6,283 females) and race (7,510 Non-Black/Non-Hispanic; 3,174 Black; 2,002 Hispanic).
+To validate this, we used the `full_demographic` i.e.
 the data with the survey years 1979 sample.
+Figure \ref{fig:demoprofile} suggests that the demographic data we had is consistent with the sample information in the database.
 
-```{r valid-age, echo = FALSE, fig.cap="Subjects' Age Distribution"}
+```{r demoprofile, echo = FALSE, fig.cap="Demography Profile of the NLSY79 Cohort", fig.width=6}
 age <- yowie::demographic_nlsy79 %>%
   ggplot(aes(x = age_1979)) +
-  geom_bar(fill = "#45B39D")
+  geom_bar(fill = "#45B39D") +
+  ggtitle("age distribution") +
+  theme(plot.title = element_text(size = 10))
 
@@
   scale_fill_manual("gender", values = c("MALE" = "#2471A3", "FEMALE" = "#9B59B6"))+
-  theme(legend.position = "none")
+  theme(legend.position = "none")  +
+  ggtitle("gender_distribution") +
+  theme(plot.title = element_text(size = 10))
 
@@
   theme(legend.position = "none",
-        axis.text.x = element_text(angle = 10))
+        axis.text.x = element_text(angle = 10)) +
+  ggtitle("race_distribution") +
+  theme(plot.title = element_text(size = 10),
+        axis.text.x = element_text(size = 6))
 
 age / (gender|race)
+```
+
+The next step is that we explored the mean hourly wage data, in this case, we only explored the wages data in `wages_demog_hs`.
+Figure \ref{fig:featureplot} conveys that there is clearly a problem in the mean hourly wage values.
+Figure \ref{fig:featureplot} a.) shows that some observations had an exceptionally high figure of wages, even more than US$10,000 per hour.
+In Figure \ref{fig:featureplot} b.), we barely see any difference in the minimum, median, an maximum value of the wages since the distribution is heavily skewed to the right.
+Additionally, Table \ref{tab:summarytable} shows that the overall wages median of the cohort is only 7.2, while the mean is 11.87.
+It indicates that the data might contains a lot of extreme values.
+
+```{r spaghetti, echo = FALSE}
+spag <- wages_demog_hs %>%
+  ggplot(aes(x = year,
+             y = mean_hourly_wage,
+             group = id)) +
+  geom_line(alpha = 0.1) +
+  ggtitle("a)") +
+  theme(plot.title = element_text(size = 10))
+```
+
+```{r featureplot, echo = FALSE, fig.cap = "a. The mean hourly wage of cohort whose highest grade completed is up to 12th grade, b. Three features of the cohort", fig.width=6, fig.height=4}
+wages_hs2020 <- as_tsibble(x = wages_demog_hs,
+                    key = id,
+                    index = year,
+                    regular = FALSE)
+wages_three_feat <- wages_hs2020 %>%
+  features(mean_hourly_wage, 
+           feat_three_num
+           )
+wages_feat_long <- wages_three_feat %>%
+  pivot_longer(c(min, med, max), names_to = "feature", values_to = "value")
+ feature <- ggplot(wages_feat_long) +
+  geom_density(aes(x = value, colour = feature, fill = feature), alpha = 0.3) +
+   ggtitle("b)") +
+   theme(plot.title = element_text(size = 10))
+
+spag + feature
+```
+```{r summarytable, echo = FALSE}
+kable(as.array(summary(wages_demog_hs$mean_hourly_wage)), 
+      caption = "Summary Statistics of Wages of High School Data",
+      col.names = c("Statistics", "Value")) %>%
+  kable_styling(latex_options = "striped")
+```
+
+Further, we took 36 samples randomly from the data and plotted it as is seen in Figure \ref{fig:sampleplot}.
+It implies that not only that some observations earned an extremely high figures of wages, but some also had a very fluctuative wages, fo example in panel number 5, 7, and 11.
+
+```{r sampleplot, echo = FALSE, fig.cap="Some samples of observations"}
+set.seed(20210225)
+
+ggplot(wages_hs2020, 
+       aes(x = year,
+                y = mean_hourly_wage,
+                group = id)) +
+  geom_line(alpha = 0.7) +
+  facet_sample() +
+  ylab("mean hourly wage") 
 ```

### 2021-02-23 report/report.tex
Add plots to IDA section.

@@
 first interviewed in 1979. Hence, we would like to validate whether all
-of the respondents were in this range. For the demographic data
-validation, we used the \texttt{full\_demographic} i.e.~the data with
-the survey years 1979 sample.
+of the respondents were in this range. Additionally, the
+\href{https://www.nlsinfo.org/content/cohorts/nlsy79/intro-to-the-sample/nlsy79-sample-introduction}{NLSY}
+also provided the number of the survey cohort by their gender (6,403
+males and 6,283 females) and race (7,510 Non-Black/Non-Hispanic; 3,174
+Black; 2,002 Hispanic). To validate this, we used the
+\texttt{full\_demographic} i.e.~the data with the survey years 1979
+sample. Figure \ref{fig:demoprofile} suggests that the demographic data
+we had is consistent with the sample information in the database.
 
@@
 \centering
-\includegraphics{report_files/figure-latex/valid-age-1.pdf}
-\caption{Subjects' Age Distribution}
+\includegraphics{report_files/figure-latex/demoprofile-1.pdf}
+\caption{Demography Profile of the NLSY79 Cohort}
+\end{figure}
+
+The next step is that we explored the mean hourly wage data, in this
+case, we only explored the wages data in \texttt{wages\_demog\_hs}.
+Figure \ref{fig:featureplot} conveys that there is clearly a problem in
+the mean hourly wage values. Figure \ref{fig:featureplot} a.) shows that
+some observations had an exceptionally high figure of wages, even more
+than US\$10,000 per hour. In Figure \ref{fig:featureplot} b.), we barely
+see any difference in the minimum, median, an maximum value of the wages
+since the distribution is heavily skewed to the right. Additionally,
+Table \ref{tab:summarytable} shows that the overall wages median of the
+cohort is only 7.2, while the mean is 11.87. It indicates that the data
+might contains a lot of extreme values.
+
+\includegraphics{report_files/figure-latex/featureplot-1.pdf}
+
+\begin{table}
+
+\caption{\label{tab:summarytable}Summary Statistics of Wages of High School Data}
+\centering
+\begin{tabular}[t]{l|r}
+\hline
+Statistics & Value\\
+\hline
+\rowcolor{gray!6}  Min. & 0.01000\\
+\hline
+1st Qu. & 4.50000\\
+\hline
+\rowcolor{gray!6}  Median & 7.20000\\
+\hline
+Mean & 11.86578\\
+\hline
+\rowcolor{gray!6}  3rd Qu. & 11.74000\\
+\hline
+Max. & 33400.00000\\
+\hline
+\end{tabular}
+\end{table}
+
+Further, we took 36 samples randomly from the data and plotted it as is
+seen in Figure \ref{fig:sampleplot}. It implies that not only that some
+observations earned an extremely high figures of wages, but some also
+had a very fluctuative wages, fo example in panel number 5, 7, and 11.
+
+\begin{figure}
+\centering
+\includegraphics{report_files/figure-latex/sampleplot-1.pdf}
+\caption{Some samples of observations}
 \end{figure}

### 2021-02-24 report_alt.Rmd
Edit the report template and add rlm section.

--- 2021-02-23 report/report.Rmd
@@
 ---
-title: Longitudinal Data Cleaning - A Case of the NLSY79 Data
-authors:
-  - name: Dewi Amaliah
-    thanks: Use footnote for providing further information about author (webpage, alternative address)---*not* for acknowledging funding agencies. Optional.
-    department: Department of Econometric and Business Statistics
-    affiliation: Monash University
-    location: Clayton, VIC 3168
-    email: dama0007@student.monash.edu
-  - name: Dianne Cook
-    department: Department of Econometric and Business Statistics
-    affiliation: Monash University
-    location: Clayton, VIC 3168
-    email: dicook@monash.edu
-abstract: |
-  Enter the text of your abstract here.
-keywords:
-  - longitudinal data
-  - data cleaning 
-  - data tidying
-  - robust linear model
+title: "Longitudinal Data Cleaning"
+subtitle: "A Case of The NLSY79 Data"
+author: "Dewi Lestari Amaliah"
+date: "24/02/2021"
+output: 
+  bookdown::pdf_document2:
+    toc : false
 bibliography: references.bib
-biblio-style: unsrt
-output: 
-  rticles::arxiv_article:
-    keep_tex: true
 ---
@@
 # Introduction
+
+Real world or wild caugh data
 
@@
 
-In this paper, we conducted an IDA or a preliminary data analysis to assess the consistency of the data with the cohort information that is provided by the NLSY.
-In addition, we also aimed to find the anomaly in the wages values using this approach.
-We mainly used graphical summary to do the IDA using `ggplot2`[@ggplot2] and `brolgar` [@brolgar].
-
-As stated previously, the respondents' ages ranged from 12 to 22 when first interviewed in 1979.
-Hence, we would like to validate whether all of the respondents were in this range.
-Additionally, the [NLSY](https://www.nlsinfo.org/content/cohorts/nlsy79/intro-to-the-sample/nlsy79-sample-introduction) also provided the number of the survey cohort by their gender (6,403 males and 6,283 females) and race (7,510 Non-Black/Non-Hispanic; 3,174 Black; 2,002 Hispanic).
-To validate this, we used the `full_demographic` i.e.
-the data with the survey years 1979 sample.
-Figure \ref{fig:demoprofile} suggests that the demographic data we had is consistent with the sample information in the database.
-
-```{r demoprofile, echo = FALSE, fig.cap="Demography Profile of the NLSY79 Cohort", fig.width=6}
+```{r demoprofile, echo = FALSE, fig.cap="Demography Profile of the NLSY79 Cohort", fig.width=5}
 age <- yowie::demographic_nlsy79 %>%
@@
   theme(plot.title = element_text(size = 10),
-        axis.text.x = element_text(size = 6))
+        axis.text.x = element_text(size = 4))
 
@@
 
-The next step is that we explored the mean hourly wage data, in this case, we only explored the wages data in `wages_demog_hs`.
-Figure \ref{fig:featureplot} conveys that there is clearly a problem in the mean hourly wage values.
-Figure \ref{fig:featureplot} a.) shows that some observations had an exceptionally high figure of wages, even more than US$10,000 per hour.
-In Figure \ref{fig:featureplot} b.), we barely see any difference in the minimum, median, an maximum value of the wages since the distribution is heavily skewed to the right.
-Additionally, Table \ref{tab:summarytable} shows that the overall wages median of the cohort is only 7.2, while the mean is 11.87.
-It indicates that the data might contains a lot of extreme values.
+In this paper, we conducted an IDA or a preliminary data analysis to assess the consistency of the data with the cohort information that is provided by the NLSY.
+In addition, we also aimed to find the anomaly in the wages values using this approach.
+We mainly used graphical summary to do the IDA using `ggplot2`[@ggplot2] and `brolgar` [@brolgar].
 
-```{r spaghetti, echo = FALSE}
+As stated previously, the respondents' ages ranged from 12 to 22 when first interviewed in 1979.
+Hence, we would like to validate whether all of the respondents were in this range.
+Additionally, the [NLSY](https://www.nlsinfo.org/content/cohorts/nlsy79/intro-to-the-sample/nlsy79-sample-introduction) also provided the number of the survey cohort by their gender (6,403 males and 6,283 females) and race (7,510 Non-Black/Non-Hispanic; 3,174 Black; 2,002 Hispanic).
+To validate this, we used the `full_demographic` i.e.
+the data with the survey years 1979 sample.
+Figure \ref{fig:demoprofile} suggests that the demographic data we had is consistent with the sample information in the database.
+
+```{r featureplot, echo = FALSE, fig.cap = "a) Pattern of Mean Hourly Wage of High School Cohort; b) The Three Features of the Cohort", fig.width=6, fig.height=4}
+
 spag <- wages_demog_hs %>%
@@
   theme(plot.title = element_text(size = 10))
-```
 
-```{r featureplot, echo = FALSE, fig.cap = "a. The mean hourly wage of cohort whose highest grade completed is up to 12th grade, b. Three features of the cohort", fig.width=6, fig.height=4}
-wages_hs2020 <- as_tsibble(x = wages_demog_hs,
+wages_demog_hs_tsibble <- as_tsibble(x = wages_demog_hs,
                     key = id,
@@
                     regular = FALSE)
-wages_three_feat <- wages_hs2020 %>%
+wages_three_feat <- wages_demog_hs_tsibble %>%
   features(mean_hourly_wage, 
@@
 ```
+
+The next step is that we explored the mean hourly wage data, in this case, we only explored the wages data in `wages_demog_hs`.
+Figure \ref{fig:featureplot} conveys that there is clearly a problem in the mean hourly wage values.
+Figure \ref{fig:featureplot} a.) shows that some observations had an exceptionally high figure of wages, even more than US$10,000 per hour.
+In Figure \ref{fig:featureplot} b.), we barely see any difference in the minimum, median, an maximum value of the wages since the distribution is heavily skewed to the right.
+Additionally, Table \ref{tab:summarytable} shows that the overall wages median of the cohort is only 7.2, while the mean is 11.87.
+It indicates that the data might contains a lot of extreme values.
+
 ```{r summarytable, echo = FALSE}
@@
 Further, we took 36 samples randomly from the data and plotted it as is seen in Figure \ref{fig:sampleplot}.
-It implies that not only that some observations earned an extremely high figures of wages, but some also had a very fluctuative wages, fo example in panel number 5, 7, and 11.
+It implies that not only that some observations earned an extremely high figures of wages, but some also had a very fluctuate wages, fo example in panel number 5, 7, and 11.
 
-```{r sampleplot, echo = FALSE, fig.cap="Some samples of observations"}
+```{r sampleplot, echo = FALSE, fig.cap="Some samples of observations", fig.height=4, fig.width=5}
 set.seed(20210225)
 
-ggplot(wages_hs2020, 
+ggplot(wages_demog_hs_tsibble, 
        aes(x = year,
@@
   facet_sample() +
+  theme(axis.text.x = element_text(angle = 10, size = 6)) +
   ylab("mean hourly wage") 
@@
 
-### Robust Linear Model for Influential Values Treatment
+### Robust Linear Model for Noises Treatment
+
+As it is seen from figure \ref{fig:sampleplot}, there are many spikes in the mean hourly wage data.
+As part of the IDA, which is the model formulation, we built a robust linear regression model to address this issue.
+The notion of robust linear regression is to yield an estimation that is robust to the influence of noise or contamination [@KollerManuel2016rARP].
+It also aims to detect the contamination by weighting each observation based on how "well-behave" they are, known as robustness weight.
+Observations with lower robustness weight are suggested as an outliers by this method [@KollerManuel2016rARP].
+
+In this paper, we built the model using the `rlm` function from `MASS` package [@mass].
+We set the `mean_hourly_wage` and `year` as the dependent and predictor respectively.
+Furthermore, we used M-Estimation with Huber weighting where the observation with small residual get a weight of 1, while the larger the residual, the smaller the weight (less than 1) [@rlm].
+
+Since we worked with longitudinal data, we should built the model for each ID, instead of the overall data.
+The robust mixed model is actually the best model to be employed in this case.
+However, this method is too computationally and memory expensive, especially for a large data set, like the NLSY79 data.
+Thus, the model for each ID is built utilizing the `nest` and `map` function from `tidyr` [@tidyr] and `purrr` [@purrr] respectively.
+After getting the weight values, we flagged the observations with the weight less than 1, and imputed their mean hourly wage with the models' predicted value.
+
+Figure \ref{fig:comppict} a) shows that after imputing the outliers with the models' predicted value, the highest wages value is not exceed US$200.
+The spikes were still observed, but are not as extreme as the original data.
+In Figure \ref{fig:comppict} b), although the distributions of the features are still positively skewed, we can still examine it clearly.
+We also learned that some IDs' have a minimum wages that is higher than others' maximum wages.
+
+```{r rlm, cache=TRUE, warning = FALSE, message = FALSE}
+
+# nest the data by id to build a robust linear model
+by_id <- wages_demog_hs %>%
+  dplyr::select(id, year, mean_hourly_wage) %>%
+  group_by(id) %>%
+  nest()
+
+# build a robust linear model
+id_rlm <- by_id %>%
+  mutate(model = map(.x = data,
+                     .f = function(x){
+                       rlm(mean_hourly_wage ~ year, data = x)
+                     }))
+# extract the property of the regression model
+id_aug <- id_rlm %>%
+  mutate(augmented = map(model, broom::augment)) %>%
+  unnest(augmented)
+
+# extract the weight of each observation
+id_w <- id_rlm %>%
+  mutate(w = map(.x = model,
+                 .f = function(x){
+                   x$w
+                 })) %>%
+  unnest(w) %>%
+  dplyr::select(w)
+
+# bind the property of each observation with their weight
+id_aug_w <- cbind(id_aug, id_w) %>%
+  dplyr::select(`id...1`,
+                year,
+                mean_hourly_wage,
+                .fitted,
+                .resid,
+                .hat,
+                .sigma,
+                w) %>%
+  rename(id = `id...1`)
+
+# if the weight < 1, the mean_hourly_wage is replaced by the model's fitted/predicted value.
+# and add the flag whether the observation is predicted value or not.
+# since the fitted value is sometimes <0, and wages value could never be negative,
+# we keep the mean hourly wage value even its weight < 1.
+
+wages_rlm_dat <- id_aug_w %>%
+  mutate(wages_rlm = ifelse(w != 1 & .fitted >= 0, .fitted,
+                            mean_hourly_wage)) %>%
+  mutate(is_pred = ifelse(w != 1 & .fitted >= 0, TRUE,
+                          FALSE)) %>%
+  dplyr::select(id, year, wages_rlm, is_pred)
+
+# join back the `wages_rlm_dat` to `wages_demog_hs`
+
+wages_demog_hs <- left_join(wages_demog_hs, wages_rlm_dat, by = c("id", "year"))
+
+```
+
+```{r comppict, echo = FALSE, fig.cap="a) The Wages and b) The Feature Distribution after the Noises Treatment", fig.width=6, fig.height=4}
+
+spag2 <- wages_demog_hs %>%
+  ggplot(aes(x = year,
+             y = wages_rlm,
+             group = id)) +
+  geom_line(alpha = 0.1) +
+  ggtitle("a)") +
+  theme(plot.title = element_text(size = 10))
+
+wages_hs2020_rlm <- as_tsibble(x = wages_demog_hs,
+                    key = id,
+                    index = year,
+                    regular = FALSE)
+wages_three_feat_rlm <- wages_hs2020_rlm %>%
+  features(wages_rlm, 
+           feat_three_num
+           )
+wages_feat_long_rlm <- wages_three_feat_rlm %>%
+  pivot_longer(c(min, med, max), names_to = "feature", values_to = "value")
+feature2 <- ggplot(wages_feat_long_rlm) +
+  geom_density(aes(x = value, colour = feature, fill = feature), alpha = 0.3) +
+  ggtitle("b)") +
+  theme(plot.title = element_text(size = 10))
+
+spag2 + feature2
+```
+
+Furthermore, the robust linear regression has reduce the level of noise in the data set as is seen in Figure \ref{fig:compare}.
+The result is that we got a data set with the reasonable degree of fluctuation.
+
+```{r compare, echo = FALSE, fig.cap="Comparison between the Original and Treated Mean Hourly Wage", fig.height=4, fig.width=5}
+set.seed(20210228)
+
+sample_id <- sample(unique(wages_demog_hs$id), 20)
+sample <- subset(wages_demog_hs, id %in% sample_id)
+
+wages_compare <- sample %>%
+  dplyr::select(id, year, mean_hourly_wage, wages_rlm) %>%
+  rename(mean_hourly_wage_rlm = wages_rlm) %>%
+  pivot_longer(c(-id, -year), names_to = "type", values_to = "wages")
+
+ggplot(wages_compare) +
+  geom_line(aes(x = year,
+                y = wages,
+                colour = type,
+                linetype = type),
+            alpha = 1) +
+  theme(axis.text.x = element_text(angle = 10, size = 5),
+        legend.position = "bottom") +
+  facet_wrap(~id)
+
+```
+
+Finally, we saved the imputed data and set the appropriate data type for the variables.
+We also saved the NLSY79 cohort's demographic information and the high school dropout cohort in a separate data sets.
+
+```{r}
+# select out the old value of mean hourly wage and change it with the wages_rlm value
+wages_demog_hs <- wages_demog_hs %>%
+  dplyr::select(-mean_hourly_wage) %>%
+  rename(mean_hourly_wage = wages_rlm)
+
+# rename and select the wages in tidy
+wages_hs2020 <- wages_demog_hs %>%
+  dplyr::select(id, year, mean_hourly_wage, age_1979, gender, race, hgc, hgc_i, yr_hgc,
+                number_of_jobs, total_hours, is_wm, is_pred) %>%
+  mutate(hgc = as.factor(hgc),
+         year = as.integer(year),
+         age_1979 = as.integer(age_1979),
+         yr_hgc = as.integer(yr_hgc),
+         number_of_jobs = as.integer(number_of_jobs))
+
+# Create a data set for demographic variables
+demographic_nlsy79 <- full_demographics %>%
+  mutate(age_1979 = 1979 - (dob_year + 1900)) %>%
+  dplyr::select(id,
+         age_1979,
+         gender,
+         race,
+         hgc,
+         hgc_i,
+         yr_hgc) %>%
+  mutate(age_1979 = as.integer(age_1979),
+         hgc = as.factor(hgc),
+         yr_hgc = as.integer(yr_hgc))
+
+# Create a data set for the high school dropouts cohort
+wages_hs_dropout <- wages_hs2020 %>%
+  mutate(dob = 1979 - age_1979,
+         age_hgc = yr_hgc - dob) %>%
+  filter((hgc %in% c("9TH GRADE",
+                     "10TH GRADE",
+                     "11TH GRADE")) |
+           (hgc == "12TH GRADE" &
+              age_hgc >= 19)) %>%
+  dplyr::select(-dob,
+         -age_hgc)
+
+```
 
 # Summary
+
+# Reference

### 2021-02-24 report_alt.Rmd
Add introduction to the report.

@@
 
-Real world or wild caugh data
+Wild caught data are often untidy and are not free from anomaly.
+Hence it is should be tidied and cleaned before a further analysis conducted to prevent inappropriate usage of statistical method that may leads to a misleading result [@HuebnerMariannePhD2016Asat].
+@HuebnerMariannePhD2016Asat also argued that data tidying and cleaning is the first part of Initial Data Analysis (IDA).
+The second and third steps are to explore the data properties and to document and report the process for the later formal analysis.
+This second step is actually the same with Chatfield's perspective on the goal of Initial Data Analysis [@Chatfield1985TIEo], which is to get the "feel" of the data.
+
+@DasuTamraparni2003Edma mentioned that data cleaning and exploration, which is the IDA notion, is a difficult task and determine 80 percent of data mining result.
+Despite its importance, this stage is often undervalued and neglected [@Chatfield1985TIEo].
+@WickhamHadley2014TD also stated that there has been few research of the good practice of data cleaning.
+Furthermore, the decisions made in this stage is often unreported [@HuebnerMariannePhD2016Asat].
+
+This paper aims to demonstrate the steps of IDA, which is tidying, cleaning, and documenting the data.
+We used the longitudinal wages data of Americans from the National Longitudinal Survey of Youth 1979 or is known as the NLSY79.
 

### 2021-02-26 report_alt.Rmd
Compare the threshold of weight.

@@
 subtitle: "A Case of The NLSY79 Data"
-author: "Dewi Lestari Amaliah"
+author: "Dewi Lestari Amaliah, Dianne Cook, Emi Tanaka, Nicholas Tierney, Kate Hyde"
 date: "24/02/2021"
@@
 This paper aims to demonstrate the steps of IDA, which is tidying, cleaning, and documenting the data.
-We used the longitudinal wages data of Americans from the National Longitudinal Survey of Youth 1979 or is known as the NLSY79.
+We used the longitudinal data of Americans from the National Longitudinal Survey of Youth 1979 or is known as the NLSY79.
+Particularly, we are interested in doing the IDA to the wages and several demographic variables of the cohort whose highest grade are up to 12th grade and the high school dropouts.
 
@@
 
+The NLSY79  is a longitudinal survey held by the U.S Bureau of Labor Statistics that follows the lives of American Youth a sample of American youth born between 1957-64.
+The cohort originally included 12,686 respondents ages 14-22 when first interviewed in 1979.
+There were three race interviewed in the survey, which are Black, Hispanic, and non-Black/non-Hispanic.
+In 1984 interview, 1,079 members of the military sample were no longer eligible for interview; 201 respondents randomly selected from the military sample remained in the survey.
+Following the 1990 interview, none of the 1,643 members of the economically disadvantaged, non-Black/non-Hispanic sample were eligible for interview.
+Hence, 9,964 respondents remain in the eligible samples.
+The surveys were conducted annually from 1979 to 1994 and biennially thereafter.
+Data are now available from Round 1 (1979 survey year) to Round 28 (2018 survey year).
+
+The NLSY79 covers several topics including Education, Training & Achievement Scores; Employment; Household, Geography & Contextual Variables; Family Background; Dating, Marriage & Cohabitation, Sexual Activity, Pregnancy & Fertility; Children; Income, Assets & Program Participation; Health; Attitudes & Expectations; and Crime and Substance Use.
+
 # The NLSY79 Data Cleaning
 
-## Getting and Tyding the Data
+## Getting and Tidyng the Data
 

### 2021-02-26 method_comparison.R

@@ -187,15 +187,29 @@ id_aug_w <- cbind(id_aug, id_w) %>%
   rename(id = `id...1`)
 
 wages_rlm_dat <- id_aug_w %>%
-  mutate(wages_rlm = ifelse(w != 1 & .fitted >= 0, .fitted,
-                            mean_hourly_wage)) %>%
-  mutate(is_pred = ifelse(w != 1 & .fitted >= 0, TRUE,
-                          FALSE)) %>%
+  mutate(wages_rlm_1 = ifelse(w != 1 & .fitted >= 0, .fitted,
+                            mean_hourly_wage),
+         wages_rlm_2 = ifelse(w < 0.7 & .fitted >= 0, .fitted,
+                              mean_hourly_wage),
+         wages_rlm_3 = ifelse(w < 0.5 & .fitted >= 0, .fitted,
+                              mean_hourly_wage),
+         wages_rlm_4 = ifelse(w < 0.3 & .fitted >= 0, .fitted,
+                              mean_hourly_wage)) %>%
+  mutate(is_pred1 = ifelse(w != 1 & .fitted >= 0, TRUE,
+                          FALSE),
+         is_pred2 = ifelse(w < 0.7 & .fitted >= 0, TRUE,
+                           FALSE),
+         is_pred3 = ifelse(w < 0.5 & .fitted >= 0, TRUE,
+                           FALSE),
+         is_pred4 = ifelse(w < 0.3 & .fitted >= 0, TRUE,
+                           FALSE)) %>%
   rename(wages_original = mean_hourly_wage)
+
+
 
 p4 <- ggplot(wages_rlm_dat) +
   geom_line(aes(x = year,
-                y = wages_rlm,
+                y = wages_rlm_1,
                 group = id),
             alpha = 0.1) +
   ggtitle("robust linear model")
@@ -207,99 +221,68 @@ p5 <- ggplot(wages_rlm_dat) +
             alpha = 0.1) +
   ggtitle("original data")
 
-wages_ext <- yowie::wages_hs2020 %>%
-  rename(wages_excl_extreme = mean_hourly_wage)
-
-p6 <- ggplot(wages_ext) +
+p6 <- ggplot(wages_rlm_dat) +
   geom_line(aes(x = year,
-                y = wages_excl_extreme,
+                y = wages_rlm_2,
                 group = id),
             alpha = 0.1) +
-  ggtitle("Exclude the extreme value")
+  ggtitle("original data")
+
[…]

### 2021-03-02 report_alt.Rmd
Adding plot of comparison.

@@
 ---
-title: "Longitudinal Data Cleaning"
+title: "Caring and Feeding the Wild Caught Data"
 subtitle: "A Case of The NLSY79 Data"
@@
 Wild caught data are often untidy and are not free from anomaly.
-Hence it is should be tidied and cleaned before a further analysis conducted to prevent inappropriate usage of statistical method that may leads to a misleading result [@HuebnerMariannePhD2016Asat].
+Hence, it is should be tidied and cleaned before a further analysis conducted to prevent inappropriate usage of statistical method that may leads to a misleading result [@HuebnerMariannePhD2016Asat].
 @HuebnerMariannePhD2016Asat also argued that data tidying and cleaning is the first part of Initial Data Analysis (IDA).
@@
 
-@DasuTamraparni2003Edma mentioned that data cleaning and exploration, which is the IDA notion, is a difficult task and determine 80 percent of data mining result.
+@DasuTamraparni2003Edma mentioned that data cleaning and exploration, which is the IDA notion, is a difficult task and determine 80 percent of the data mining result.
 Despite its importance, this stage is often undervalued and neglected [@Chatfield1985TIEo].
@@
 This paper aims to demonstrate the steps of IDA, which is tidying, cleaning, and documenting the data.
-We used the longitudinal data of Americans from the National Longitudinal Survey of Youth 1979 or is known as the NLSY79.
+We used the longitudinal data from the National Longitudinal Survey of Youth 1979 or is known as the NLSY79.
 Particularly, we are interested in doing the IDA to the wages and several demographic variables of the cohort whose highest grade are up to 12th grade and the high school dropouts.
@@
 
-The NLSY79  is a longitudinal survey held by the U.S Bureau of Labor Statistics that follows the lives of American Youth a sample of American youth born between 1957-64.
+The NLSY79  is a longitudinal survey held by the U.S Bureau of Labor Statistics that follows the lives of a sample of American youth born between 1957-64.
 The cohort originally included 12,686 respondents ages 14-22 when first interviewed in 1979.
@@
 
-## Getting and Tidyng the Data
+## Getting and Tidying the Data
 

### 2021-03-02 method_comparison.R

@@ -274,15 +274,79 @@ ggplot(sample) +
   facet_wrap(~id)
 
 
-ggplot(filter(sample, type %in% c("wages_original", "wages_rlm_3"))) +
+w05 <- ggplot(filter(sample, type %in% c("wages_original", "wages_rlm_3"))) +
   geom_line(aes(x = year,
                 y = wages,
                 colour = type,
                 linetype = type),
             alpha = 1) +
+  geom_point(aes(x = year,
+                y = wages,
+                colour = type),
+            alpha = 0.5,
+            size = 0.5) +
   theme(axis.text.x = element_text(angle = 10, size = 5),
         legend.position = "bottom") +
   facet_wrap(~id)
+
+
+w07 <- ggplot(filter(sample, type %in% c("wages_original", "wages_rlm_2"))) +
+  geom_line(aes(x = year,
+                y = wages,
+                colour = type,
+                linetype = type),
+            alpha = 1) +
+  geom_point(aes(x = year,
+                 y = wages,
+                 colour = type),
+             alpha = 0.5,
+             size = 0.5) +
+  theme(axis.text.x = element_text(angle = 10, size = 5),
+        legend.position = "bottom") +
+  facet_wrap(~id)
+
+w07
+
+
+w03 <- ggplot(filter(sample, type %in% c("wages_original", "wages_rlm_4"))) +
+  geom_line(aes(x = year,
+                y = wages,
+                colour = type,
+                linetype = type),
+            alpha = 1) +
+  geom_point(aes(x = year,
+                 y = wages,
+                 colour = type),
+             alpha = 0.5,
+             size = 0.5) +
+  theme(axis.text.x = element_text(angle = 10, size = 5),
+        legend.position = "bottom") +
+  facet_wrap(~id)
+
+w03
+
+
+w01 <- ggplot(filter(sample, type %in% c("wages_original", "wages_rlm_1"))) +
+  geom_line(aes(x = year,
+                y = wages,
+                colour = type,
+                linetype = type),
+            alpha = 1) +
+  geom_point(aes(x = year,
+                 y = wages,
+                 colour = type),
+             alpha = 0.5,
+             size = 0.5) +
+  theme(axis.text.x = element_text(angle = 10, size = 5),
+        legend.position = "bottom") +
+  facet_wrap(~id)
+
+w01
+
+
+w01 + w07 + w05 + w03
+
+
 
 
 

### 2021-03-02 report_alt.Rmd
Edit demography figure to be tables.

@@
 subtitle: "A Case of The NLSY79 Data"
-author: "Dewi Lestari Amaliah, Dianne Cook, Emi Tanaka, Nicholas Tierney, Kate Hyde"
+author: "Dewi Amaliah, Dianne Cook, Emi Tanaka, Nicholas Tierney, Kate Hyde"
 date: "24/02/2021"
@@
 library(MASS)
+library(janitor)
 ```
@@
 # Introduction
+
+*Explain a little bit about wild caught data*
 
@@
 The second and third steps are to explore the data properties and to document and report the process for the later formal analysis.
-This second step is actually the same with Chatfield's perspective on the goal of Initial Data Analysis [@Chatfield1985TIEo], which is to get the "feel" of the data.
+This second step is actually the same with Chatfield's perspective on the goal of IDA [@Chatfield1985TIEo], which is to get the "feel" of the data.
 
@@
 We used the longitudinal data from the National Longitudinal Survey of Youth 1979 or is known as the NLSY79.
+*Why using NLSY79 data*.
 Particularly, we are interested in doing the IDA to the wages and several demographic variables of the cohort whose highest grade are up to 12th grade and the high school dropouts.
@@
 
-```{r demoprofile, echo = FALSE, fig.cap="Demography Profile of the NLSY79 Cohort", fig.width=5}
-age <- yowie::demographic_nlsy79 %>%
-  ggplot(aes(x = age_1979)) +
-  geom_bar(fill = "#45B39D") +
-  ggtitle("age distribution") +
-  theme(plot.title = element_text(size = 10))
-
-gender <- yowie::demographic_nlsy79 %>%
-  ggplot(aes(x = gender,
-             fill = gender)) +
-  geom_bar() +
-  geom_text(aes(label = ..count..), stat = "count", vjust = 1.5, colour = "white")+
-  scale_fill_manual("gender", values = c("MALE" = "#2471A3", "FEMALE" = "#9B59B6"))+
-  theme(legend.position = "none")  +
-  ggtitle("gender_distribution") +
-  theme(plot.title = element_text(size = 10))
-
-race <- yowie::demographic_nlsy79 %>%
-  ggplot(aes(x = race,
-             fill = race)) +
-  geom_bar() +
-  geom_text(aes(label = ..count..), stat = "count", vjust = 1.5, colour = "white") +
-  theme(legend.position = "none",
-        axis.text.x = element_text(angle = 10)) +
-  ggtitle("race_distribution") +
-  theme(plot.title = element_text(size = 10),
-        axis.text.x = element_text(size = 4))
-
-age / (gender|race)
-```
-
 In this paper, we conducted an IDA or a preliminary data analysis to assess the consistency of the data with the cohort information that is provided by the NLSY.
@@
 the data with the survey years 1979 sample.
-Figure \ref{fig:demoprofile} suggests that the demographic data we had is consistent with the sample information in the database.
+Table \ref{tab:age-table} and Table \ref{tab:gender-race-table} suggest that the demographic data we had is consistent with the sample information in the database.
+
+```{r age-table}
+age_table <- yowie::demographic_nlsy79 %>%
+  group_by(age_1979) %>%
+  count(age_1979)
+
+kable(age_table,
+      caption = "Age Distribution of the NLSY79 samples",
+      col.names = c("Age", "Number of Sample")) %>%
+  kable_styling(latex_options = "striped")
+```
+
+```{r gender-race-table}
+gender_race_table <- yowie::demographic_nlsy79 %>%
+  tabyl(gender, race) %>%
+  adorn_totals(c("row", "col")) %>%
+  adorn_percentages("row") %>%
+  adorn_pct_formatting(digits = 2) %>%
+  adorn_ns(position = "front") %>%
+  mutate(gender = ifelse(gender == "MALE", "Male",
+                         ifelse(gender == "FEMALE", "Female", "Total")))
+
+kable(gender_race_table,
+      caption = "Gender and Race Distribution of the NLSY79 Samples",
+      col.names = c("Gender", "Hispanic", "Black", "Non-Black, Non-Hispanic", "Total")) %>%
+  kable_styling(latex_options = "striped") %>%
+  add_header_above(c(" " = 1, "Race" = 3, " " = 1))
+
+```
 

### 2021-03-03 report_alt.Rmd
Edit the introduction of the report.

@@
 
-*Explain a little bit about wild caught data*
+Just like the living things, the usage of the term "wild" could also be applied in the data domain.
+Wild data is the form of data that could be acquired from various source and might be generated from experiments, surveys, or may be from sensors or social media.
+Wild data also often associated with "Open Data", means that the data is freely accessible, modified, and shared by anyone for any purpose [@opendata].
 
-Wild caught data are often untidy and are not free from anomaly.
+Despite from its traits, open data is often have an issue with its quality in the sense of is often not tidy and clean.
 Hence, it is should be tidied and cleaned before a further analysis conducted to prevent inappropriate usage of statistical method that may leads to a misleading result [@HuebnerMariannePhD2016Asat].
@@
 This second step is actually the same with Chatfield's perspective on the goal of IDA [@Chatfield1985TIEo], which is to get the "feel" of the data.
+In other word, IDA could be regarded as a way to assess the data quality.
 
@@
 @WickhamHadley2014TD also stated that there has been few research of the good practice of data cleaning.
-Furthermore, the decisions made in this stage is often unreported [@HuebnerMariannePhD2016Asat].
+Furthermore, the decisions made in this stage is often unreported  in term that the IDA are often performed in an unplanned and unstructured way and is only shared among restricted party [@HuebnerMarianne2020Haar].
+Further, In fact, documenting the result of IDA is utterly essential as the information preservation for the later statistical analyses and model building [@HuebnerMariannePhD2016Asat] and to avoid publication bias [@HuebnerMarianne2020Haar].
 
@@
 We used the longitudinal data from the National Longitudinal Survey of Youth 1979 or is known as the NLSY79.
-*Why using NLSY79 data*.
-Particularly, we are interested in doing the IDA to the wages and several demographic variables of the cohort whose highest grade are up to 12th grade and the high school dropouts.
+Particularly, we are interested in doing the IDA to the wages and several demographic variables of the NLSY79 cohort whose highest grade are up to 12th grade and the high school dropouts.
+
+We chose the NLSY data because this data has been played an important role in research in various discipline including economics, sociology, education, public policy, and public health for more than a quarter of century  [@MichaelRPergamit2001DWTN].
+In addition, The NLS is considered as survey with high retention rates and carefully designed making it suitable for a life course research [@MichaelRPergamit2001DWTN] and [@eliznlsy].
+According to @eliznlsy as of 2015, thousand of articles, and hundreds of book chapters and monographs has incorporated the NLSY data.
+Moreover, the NLSY79 is considered as the most widely used and most important cohort in the NLS data sets [@MichaelRPergamit2001DWTN].
 
@@
 
-The NLSY79  is a longitudinal survey held by the U.S Bureau of Labor Statistics that follows the lives of a sample of American youth born between 1957-64.
+The NLSY79  is a longitudinal survey held by the U.S Bureau of Labor Statistics that follows the lives of a sample of American youth and born between 1957-1964.
 The cohort originally included 12,686 respondents ages 14-22 when first interviewed in 1979.
-There were three race interviewed in the survey, which are Black, Hispanic, and non-Black/non-Hispanic.
-In 1984 interview, 1,079 members of the military sample were no longer eligible for interview; 201 respondents randomly selected from the military sample remained in the survey.
-Following the 1990 interview, none of the 1,643 members of the economically disadvantaged, non-Black/non-Hispanic sample were eligible for interview.
+It was comprised of Blacks, Hispanics, economically disadvantaged non-Black non-Hispanics, and youth in the military.
+In 1984 and 1990, two sub-samples were dropped from the interview.
+They are  1,079 members of the military sample and 1,643 members of the economically disadvantaged non-Black non-Hispanics respectively.
 Hence, 9,964 respondents remain in the eligible samples.
@@
 
-The NLSY79 covers several topics including Education, Training & Achievement Scores; Employment; Household, Geography & Contextual Variables; Family Background; Dating, Marriage & Cohabitation, Sexual Activity, Pregnancy & Fertility; Children; Income, Assets & Program Participation; Health; Attitudes & Expectations; and Crime and Substance Use.
+Although the main focus area of the NLSY is labor and employment, the NLSY also cover several other topics including education; training and achievement; household, geography and contextual variables; dating, marriage, cohabitation; sexual activity, pregnancy, and fertility; children; income, assets and program participation; health; attitudes and expectations; and crime and substance use.
+
+There are two ways to conduct the interview of the NLSY, which are face to face interview or by telephone.
+In recent survey years, more than 90 percent of respondents were interviewed by telephone [@eliznlsy].
 
@@
 
-```{r age-table}
+```{r age-table, echo = FALSE}
 age_table <- yowie::demographic_nlsy79 %>%
@@
 
-```{r gender-race-table}
+```{r gender-race-table, echo = FALSE}
 gender_race_table <- yowie::demographic_nlsy79 %>%

### 2021-03-03 report_alt.Rmd
report intro + NLSY79

@@
 Wild data is the form of data that could be acquired from various source and might be generated from experiments, surveys, or may be from sensors or social media.
-Wild data also often associated with "Open Data", means that the data is freely accessible, modified, and shared by anyone for any purpose [@opendata].
+Wild data also often associated with "Open Data", means that the data are freely accessible, modified, and shared by anyone for any purpose [@opendata].
 
-Despite from its traits, open data is often have an issue with its quality in the sense of is often not tidy and clean.
-Hence, it is should be tidied and cleaned before a further analysis conducted to prevent inappropriate usage of statistical method that may leads to a misleading result [@HuebnerMariannePhD2016Asat].
+Despite its traits, wild data often have an issue with its quality in the sense that it is often not tidy and clean.
+Hence, it should be tidied and cleaned before further analysis to prevent inappropriate usage of statistical method that may lead to a misleading result [@HuebnerMariannePhD2016Asat].
 @HuebnerMariannePhD2016Asat also argued that data tidying and cleaning is the first part of Initial Data Analysis (IDA).
 The second and third steps are to explore the data properties and to document and report the process for the later formal analysis.
-This second step is actually the same with Chatfield's perspective on the goal of IDA [@Chatfield1985TIEo], which is to get the "feel" of the data.
+This second step is the same with Chatfield's perspective on the goal of IDA [@Chatfield1985TIEo], which is to get the "feel" of the data.
 In other word, IDA could be regarded as a way to assess the data quality.
@@
 Furthermore, the decisions made in this stage is often unreported  in term that the IDA are often performed in an unplanned and unstructured way and is only shared among restricted party [@HuebnerMarianne2020Haar].
-Further, In fact, documenting the result of IDA is utterly essential as the information preservation for the later statistical analyses and model building [@HuebnerMariannePhD2016Asat] and to avoid publication bias [@HuebnerMarianne2020Haar].
+Furthermore, documenting the result of IDA is utterly essential as the information preservation for the downstream statistical analyses and model building [@HuebnerMariannePhD2016Asat] and to avoid publication bias [@HuebnerMarianne2020Haar].
 
 This paper aims to demonstrate the steps of IDA, which is tidying, cleaning, and documenting the data.
-We used the longitudinal data from the National Longitudinal Survey of Youth 1979 or is known as the NLSY79.
+We used the longitudinal data from the National Longitudinal Survey of Youth 1979 (NLSY79) or is known as the NLSY79.
 Particularly, we are interested in doing the IDA to the wages and several demographic variables of the NLSY79 cohort whose highest grade are up to 12th grade and the high school dropouts.
 
-We chose the NLSY data because this data has been played an important role in research in various discipline including economics, sociology, education, public policy, and public health for more than a quarter of century  [@MichaelRPergamit2001DWTN].
-In addition, The NLS is considered as survey with high retention rates and carefully designed making it suitable for a life course research [@MichaelRPergamit2001DWTN] and [@eliznlsy].
-According to @eliznlsy as of 2015, thousand of articles, and hundreds of book chapters and monographs has incorporated the NLSY data.
-Moreover, the NLSY79 is considered as the most widely used and most important cohort in the NLS data sets [@MichaelRPergamit2001DWTN].
+We chose the NLSY79 data because this data play an important role in research in various discipline including economics, sociology, education, public policy, and public health for more than a quarter of the century [@MichaelRPergamit2001DWTN].
+In addition, The National Longitudinal Survey is considered as survey with high retention rates and carefully designed making it suitable for a life course research [@MichaelRPergamit2001DWTN] and [@eliznlsy].
+According to @eliznlsy as of 2015, thousand of articles, and hundreds of book chapters and monographs has incorporated the NLSY79 data.
+Moreover, the NLSY79 is considered as the most widely used and most important cohort in the NLSY79 data sets [@MichaelRPergamit2001DWTN].
 
@@
 It was comprised of Blacks, Hispanics, economically disadvantaged non-Black non-Hispanics, and youth in the military.
-In 1984 and 1990, two sub-samples were dropped from the interview.
-They are  1,079 members of the military sample and 1,643 members of the economically disadvantaged non-Black non-Hispanics respectively.
+In 1984 and 1990, two sub-samples were dropped from the interview; the dropped subjects are the 1,079 members of the military sample and 1,643 members of the economically disadvantaged non-Black non-Hispanics respectively.
 Hence, 9,964 respondents remain in the eligible samples.
@@
 
-There are two ways to conduct the interview of the NLSY, which are face to face interview or by telephone.
+There are two ways to conduct the interview of the NLSY79, which are face to face interview or by telephone.
 In recent survey years, more than 90 percent of respondents were interviewed by telephone [@eliznlsy].

### 2021-03-03 report_alt.Rmd
cache

@@
 ```{r setup, include = FALSE}
-knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
+knitr::opts_chunk$set(echo = TRUE, 
+                      warning = FALSE, 
+                      message = FALSE,
+                      cache = TRUE, 
+                      cache.path = "cache/")
 library(tidyverse)
@@
 
-The NLYS data is stored in a [database](https://www.nlsinfo.org/content/cohorts/nlsy79/get-data) and could be downloaded by variables.
-Several variables are available to be downloaded and could be browsed by index.
+The NLSY79 data are stored in a [database](https://www.nlsinfo.org/content/cohorts/nlsy79/get-data) and could be downloaded by variables.
+Several variables are available for download and could be browsed by index.
 For the wages data set, we only extracted these variables:

### 2021-03-03 report_alt.Rmd
author order changed. Di last. Middle authors alphabetical

@@
 subtitle: "A Case of The NLSY79 Data"
-author: "Dewi Amaliah, Dianne Cook, Emi Tanaka, Nicholas Tierney, Kate Hyde"
+author: 
+ - Dewi Amaliah
+ - Kate Hyde
+ - Emi Tanaka
+ - Nicholas Tierney
+ - Dianne Cook
 date: "24/02/2021"

### 2021-03-03 report_alt.Rmd
arxiv version

--- 2021-03-03 report_alt.Rmd
@@
 subtitle: "A Case of The NLSY79 Data"
-author: "Dewi Amaliah, Dianne Cook, Emi Tanaka, Nicholas Tierney, Kate Hyde"
+authors:
+  - name: Dewi Amaliah
+    thanks: ""
+    department: Department of Econometrics and Business Statistics
+    affiliation: Monash University
+    location: Clayton, VIC 3800
+    email: ""
+  - name: Kate Hyde
+    thanks: ""
+    department: ""
+    affiliation: ""
+    location: ""
+    email: ""
+  - name: Emi Tanaka
+    thanks: emitanaka.org
+    department: Department of Econometrics and Business Statistics
+    affiliation: Monash University
+    location: Clayton, VIC 3800
+    email: emi.tanaka@monash.edu
+  - name: Nicholas Tierney
+    thanks: ""
+    department: ""
+    affiliation: ""
+    location: ""
+    email: ""
+  - name: Dianne Cook
+    thanks: dicook.org
+    department: Department of Econometrics and Business Statistics
+    affiliation: Monash University
+    location: Clayton, VIC 3800
+    email: dicook@monash.edu
+bibliography: references.bib
+biblio-style: unsrt
+#preamble: >
+output: 
+  rticles::arxiv_article:
+    keep_tex: false
 date: "24/02/2021"
-output: 
-  bookdown::pdf_document2:
-    toc : false
-bibliography: references.bib
 ---
@@
                       cache = TRUE, 
-                      cache.path = "cache/")
+                      cache.path = "cache/",
+                      fig.path = "figures/")
 library(tidyverse)
@@
 
-```{r featureplot, echo = FALSE, fig.cap = "a) Pattern of Mean Hourly Wage of High School Cohort; b) The Three Features of the Cohort", fig.width=6, fig.height=4}
+```{r featureplot, echo = FALSE, fig.cap = "a) Pattern of Mean Hourly Wage of High School Cohort; b) The Three Features of the Cohort", fig.width=6, fig.height=4, cache = FALSE}
 

### 2021-03-08 report_alt.Rmd
Edit caption, and wages threshold.

@@
 
-`new_data_qnames` is still untidy, where all of the variables for each year and each job being stored in one column, hence the data contains a huge number of columns (686 columns).
-Thus, the data should be tidied and wrangled first to extract the demographic and employment variables that we want to put in the final data set.
+According to @WickhamHadley2014TD, a tidy data sets comply with three rules, the first is that each variable forms a column, the second is that each observation forms a row, and the last is that each type of observational unit forms a table.
+Unfortunately, the `new_data_qnames` did not meet these requirements in the way that the value for particular year and job are stored in different columns, hence the data contains a huge number of columns (686 columns).
+The example of the untidy of the data set is displayed in Table \ref{tab:untidy-data}, which actually intended to display the hourly rate of each respondent by job (HRP1 to HRP5) and by year (1979 and 1980).
+The table implies that the column headers are values of the year and job, not variable names.
+
+```{r untidy-data, echo = FALSE}
+
+untidy_demo <- new_data_qnames %>%
+  as_tibble() %>%
+  # in 2018, the variable's name is Q3-4_2018, instead of HGC_2018
+  rename(HGC_2018 = `Q3-4_2018`) %>%
+  dplyr::select(CASEID_1979,
+            starts_with("HRP") &
+              ends_with(c("1979", "1980", "1981", "1982", "1983"))) %>%
+  dplyr::select(1:7) %>%
+  head()
+
+kable(untidy_demo,
+      caption = "The Untidy Form of the NLSY79 Raw Data") %>%
+  kable_styling(latex_options = "striped")
+```
+
+Consequently, the data should be tidied and wrangled first to extract the demographic and employment variables that we want to put in the final data set.
 We mainly used `tidyr` [@tidyr], `dplyr` [@dplyr], and `stringr` [@stringr] to do this job.
@@
 
-```{r featureplot, echo = FALSE, fig.cap = "a) Pattern of Mean Hourly Wage of High School Cohort; b) The Three Features of the Cohort", fig.width=6, fig.height=4, cache = FALSE}
+```{r featureplot, echo = FALSE, fig.cap = "Two plots showing the distribution of the mean hourly wage.
+Plot A portrays the pattern of mean hourly wage of high school cohort from 1979 to 2018 of each ID in US Dollar; Plot B shows the distribution of their minimum, median, and maximum value.
+We can see that some IDs had an extremely high of wages and it made the distribution of the three features is extremely skewed.", fig.width=6, fig.height=4}
 
@@
 Figure \ref{fig:featureplot} conveys that there is clearly a problem in the mean hourly wage values.
-Figure \ref{fig:featureplot} a.) shows that some observations had an exceptionally high figure of wages, even more than US$10,000 per hour.
-In Figure \ref{fig:featureplot} b.), we barely see any difference in the minimum, median, an maximum value of the wages since the distribution is heavily skewed to the right.
+Figure \ref{fig:featureplot} A shows that some observations had an exceptionally high figure of wages, even more than US$10,000 per hour.
+In Figure \ref{fig:featureplot} B, we barely see any difference in the minimum, median, an maximum value of the wages since the distribution is heavily skewed to the right.
 Additionally, Table \ref{tab:summarytable} shows that the overall wages median of the cohort is only 7.2, while the mean is 11.87.
@@
 Further, we took 36 samples randomly from the data and plotted it as is seen in Figure \ref{fig:sampleplot}.
-It implies that not only that some observations earned an extremely high figures of wages, but some also had a very fluctuate wages, fo example in panel number 5, 7, and 11.
+It implies that not only that some observations earned an extremely high figures of wages, but some also had a reasonably fluctuate wages, for example the IDs in panel number 5, 7, and 11.
+The plot also implies that the samples had a different pattern of mean hourly wages.
+Some had a flat wages for years but had a sudden increase in a year than it went down again, while the other experienced a upsurge in their wage, for instance the IDs in panel 9.
 
-```{r sampleplot, echo = FALSE, fig.cap="Some samples of observations", fig.height=4, fig.width=5}
+According to @MichaelRPergamit2001DWTN, one of the flaws of the NLSY79 employment data is that since the NLSY79 collect the information of the working hours since the last interview, it might be challenging for the respondents to track the within-job hours changes that happens between survey year, especially for the respondents with fluctuate working hours or whose job is seasonal.
+It even has been more challenging since 1994, where the respondents had to recall two years period.
+This shortcoming might also contribute on the fluctuation of one's wages data.
+
+```{r sampleplot, echo = FALSE, fig.cap="The mean hourly wages of some random samples are shown in twelve facets, three IDs per facet.
+It suggests that some IDs had a reasonably fluctuate wages.", fig.height=4, fig.width=5}
+
 set.seed(20210225)
@@
 Thus, the model for each ID is built utilizing the `nest` and `map` function from `tidyr` [@tidyr] and `purrr` [@purrr] respectively.
-After getting the weight values, we flagged the observations with the weight less than 1, and imputed their mean hourly wage with the models' predicted value.
 
-Figure \ref{fig:comppict} a) shows that after imputing the outliers with the models' predicted value, the highest wages value is not exceed US$200.
-The spikes were still observed, but are not as extreme as the original data.
-In Figure \ref{fig:comppict} b), although the distributions of the features are still positively skewed, we can still examine it clearly.
-We also learned that some IDs' have a minimum wages that is higher than others' maximum wages.
+The challenging part of detecting the anomaly using the robustness weight is to determine the threshold of the weight where the observations considered as outliers.
+Moreover, it should be noted that not all the outliers is due to an error, instead it might be that one had a reasonably increasing or decreasing wages.
+To, minimize the risk of being mistakenly regard an outlier as an error outlier, we have simulated some threshold and study the behavior of the spikes in each threshold.
+We found that 0.3 is the most reasonable value to be the threshold to minimize the risk of that drawback because it still capture the sensible spikes in the data.
+After deciding the threshold, we flagged the observations with the weight less than 0.3, and imputed their mean hourly wage with the models' predicted value.
 
@@
 wages_rlm_dat <- id_aug_w %>%
-  mutate(wages_rlm = ifelse(w != 1 & .fitted >= 0, .fitted,
+  mutate(wages_rlm = ifelse(w < 0.3  & .fitted >= 0, .fitted,
                             mean_hourly_wage)) %>%
-  mutate(is_pred = ifelse(w != 1 & .fitted >= 0, TRUE,
+  mutate(is_pred = ifelse(w < 0.3 & .fitted >= 0, TRUE,
                           FALSE)) %>%
@@
 
-```{r comppict, echo = FALSE, fig.cap="a) The Wages and b) The Feature Distribution after the Noises Treatment", fig.width=6, fig.height=4}
+Figure \ref{fig:comppict} A shows that after imputing the "error outliers" with the models' predicted value, the highest wages value has decreased to be around US$250.
+The spikes were still observed, but are not as extreme as the original data set.
+In Figure \ref{fig:comppict} B) although the distributions of the features are still positively skewed, we can still examine it clearly the difference shape of those features.
+The minimum value is heavily skewed, means that most the subjects have a small minimum value of wages, but there are still extreme cases where their minimum wages was extremely higher than others.
+Moreover, these cases had a minimum wages that is higher than others' maximum wages.
+
+Furthermore, the robust linear regression has reduce the level of noise in the data set as is seen in Figure \ref{fig:compare}.
+The figure also implies that after the treatment, the fluctuation can still be observed in the data and only the large spikes, which are considered as "error outliers", are eliminated from the data.
+Hence, the model results a data set with the reasonable degree of fluctuation.
+
+```{r comppict, echo = FALSE, fig.cap = "The distribution of the mean hourly wage after treating the extreme values.
+Plot A portrays the pattern of mean hourly wage of high school cohort from 1979 to 2018 of each ID in US Dollar; Plot B shows the distribution of their minimum, median, and maximum value.
+We can see that some observations still had reasonbaly higher wages than the others.
+The minimum, median, and maximum distribution is positively skewed, where some IDs' have a minimum wages that is higher than others' maximum wages.", fig.width=6, fig.height=4}
 
@@
   pivot_longer(c(min, med, max), names_to = "feature", values_to = "value")
+
 feature2 <- ggplot(wages_feat_long_rlm) +
@@
 
-Furthermore, the robust linear regression has reduce the level of noise in the data set as is seen in Figure \ref{fig:compare}.
-The result is that we got a data set with the reasonable degree of fluctuation.
-
-```{r compare, echo = FALSE, fig.cap="Comparison between the Original and Treated Mean Hourly Wage", fig.height=4, fig.width=5}
+```{r compare-data, echo = FALSE}
 set.seed(20210228)
@@
   pivot_longer(c(-id, -year), names_to = "type", values_to = "wages")
+```
 
+```{r compare, echo = FALSE, fig.cap="Comparison between the original and the treated mean hourly wage.
+The orange line portray the original value of mean hourly wage, while the turquoise line display the mean hourly wages value after the extreme values imputed with the robust linear model's prediction value.
+We can see that some extreme spikes has been reduced by the model.", fig.height=4, fig.width=5}
 ggplot(wages_compare) +
@@
             alpha = 1) +
+  geom_point(aes(x = year,
+                y = wages,
+                colour = type),
+            alpha = 0.5,
+            size = 0.5) +
   theme(axis.text.x = element_text(angle = 10, size = 5),
@@
 
-# Reference
+Wild caught data is often not tidy and clean.
+IDA, as a way to assess data quality should be performed and most importantly documented to avoid bias publication.
+In this paper, we performed an IDA on the NLSY79 data, specifically the wages data.
+We chose this data set because it play an important role in various discipline of research, especially the life course research.
+Getting the data from the database is pretty straightforward.
+However, it cannot be directly deployed to the analysis since the form is untidy.
+Data exploration also found some anomalies in the wages data where some IDs had an extremely high wages and large spikes.
+Hence, we performed data wrangling to tidy the data and robust linear regression model for each ID to find and treat the anomalies.
+We used the robustness weight of  less than 0.5 as threshold of the wages considered as outliers.
+
+- From this IDA, we learned that the open data is not only the matter of opennes, but also how that data could be analyzed flexibly with a tidy dataset.
+Hence, our recommendation that the database for every open data should be built to produce a tidy data.
+
+- For the future study, the room for improvement: linear mixed model and more sophisticated method to determine the threshold?

### 2021-03-08 report_old.Rmd
Edit caption, and wages threshold.

--- 2021-03-08 report_alt.Rmd
@@
 ---
-title: "Caring and Feeding the Wild Caught Data"
+title: "Initial Data Analysis to Care and Feed the Wild Caught Data"
 subtitle: "A Case of The NLSY79 Data"
-authors:
-  - name: Dewi Amaliah
-    thanks: ""
-    department: Department of Econometrics and Business Statistics
-    affiliation: Monash University
-    location: Clayton, VIC 3800
-    email: ""
-  - name: Kate Hyde
-    thanks: ""
-    department: ""
-    affiliation: ""
-    location: ""
-    email: ""
-  - name: Emi Tanaka
-    thanks: emitanaka.org
-    department: Department of Econometrics and Business Statistics
-    affiliation: Monash University
-    location: Clayton, VIC 3800
-    email: emi.tanaka@monash.edu
-  - name: Nicholas Tierney
-    thanks: ""
-    department: ""
-    affiliation: ""
-    location: ""
-    email: ""
-  - name: Dianne Cook
-    thanks: dicook.org
-    department: Department of Econometrics and Business Statistics
-    affiliation: Monash University
-    location: Clayton, VIC 3800
-    email: dicook@monash.edu
+author: "Dewi Amaliah, Dianne Cook, Emi Tanaka, Nicholas Tierney, Kate Hyde"
+date: "24/02/2021"
+output: 
+  bookdown::pdf_document2:
+    toc : false
 bibliography: references.bib
-biblio-style: unsrt
-#preamble: >
-output: 
-  rticles::arxiv_article:
-    keep_tex: false
-date: "24/02/2021"
 ---
@@
 ```{r setup, include = FALSE}
-knitr::opts_chunk$set(echo = TRUE, 
-                      warning = FALSE, 
-                      message = FALSE,
-                      cache = TRUE, 
-                      cache.path = "cache/",
-                      fig.path = "figures/")
+knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
 library(tidyverse)
@@
 Just like the living things, the usage of the term "wild" could also be applied in the data domain.
-Wild data is the form of data that could be acquired from various source and might be generated from experiments, surveys, or may be from sensors or social media.
-Wild data also often associated with "Open Data", means that the data are freely accessible, modified, and shared by anyone for any purpose [@opendata].
+Wild data is the form of data that could be acquired from various source and generated from either experiments, surveys, sensors or social media.
+Wild data also often associated with "Open Data", means that the data is freely accessible, modified, and shared by anyone for any purpose [@opendata].
 
-Despite its traits, wild data often have an issue with its quality in the sense that it is often not tidy and clean.
-Hence, it should be tidied and cleaned before further analysis to prevent inappropriate usage of statistical method that may lead to a misleading result [@HuebnerMariannePhD2016Asat].
+Despite from its traits, open data is often have an issue with its quality in the sense of is often not tidy and clean.
+Hence, it is should be tidied and cleaned before a further analysis conducted to prevent inappropriate usage of statistical method that may leads to a misleading result [@HuebnerMariannePhD2016Asat].
 @HuebnerMariannePhD2016Asat also argued that data tidying and cleaning is the first part of Initial Data Analysis (IDA).
 The second and third steps are to explore the data properties and to document and report the process for the later formal analysis.
-This second step is the same with Chatfield's perspective on the goal of IDA [@Chatfield1985TIEo], which is to get the "feel" of the data.
+This second step is actually the same with Chatfield's perspective on the goal of IDA [@Chatfield1985TIEo], which is to get the "feel" of the data.
 In other word, IDA could be regarded as a way to assess the data quality.
@@
 Furthermore, the decisions made in this stage is often unreported  in term that the IDA are often performed in an unplanned and unstructured way and is only shared among restricted party [@HuebnerMarianne2020Haar].
-Furthermore, documenting the result of IDA is utterly essential as the information preservation for the downstream statistical analyses and model building [@HuebnerMariannePhD2016Asat] and to avoid publication bias [@HuebnerMarianne2020Haar].
+Further, In fact, documenting the result of IDA is utterly essential as the information preservation for the later statistical analyses and model building [@HuebnerMariannePhD2016Asat] and to avoid publication bias [@HuebnerMarianne2020Haar].
 
-This paper aims to demonstrate the steps of IDA, which is tidying, cleaning, and documenting the data.
-We used the longitudinal data from the National Longitudinal Survey of Youth 1979 (NLSY79) or is known as the NLSY79.
+This paper aims to demonstrate the steps of IDA, which is tidying, cleaning, and documenting those process.
+We used the longitudinal data from the National Longitudinal Survey of Youth 1979 or is known as the NLSY79.
 Particularly, we are interested in doing the IDA to the wages and several demographic variables of the NLSY79 cohort whose highest grade are up to 12th grade and the high school dropouts.
 
-We chose the NLSY79 data because this data play an important role in research in various discipline including economics, sociology, education, public policy, and public health for more than a quarter of the century [@MichaelRPergamit2001DWTN].
-In addition, The National Longitudinal Survey is considered as survey with high retention rates and carefully designed making it suitable for a life course research [@MichaelRPergamit2001DWTN] and [@eliznlsy].
-According to @eliznlsy as of 2015, thousand of articles, and hundreds of book chapters and monographs has incorporated the NLSY79 data.
-Moreover, the NLSY79 is considered as the most widely used and most important cohort in the NLSY79 data sets [@MichaelRPergamit2001DWTN].
+We chose the NLSY data because this data has been played an important role in research in various discipline including economics, sociology, education, public policy, and public health for more than a quarter of century  [@MichaelRPergamit2001DWTN].
+In addition, The NLS is considered as survey with high retention rates and carefully designed making it suitable for a life course research [@MichaelRPergamit2001DWTN] and [@eliznlsy].
+According to @eliznlsy as of 2015, thousand of articles, and hundreds of book chapters and monographs has incorporated the NLSY data.
+Moreover, the NLSY79 is considered as the most widely used and most important cohort in the NLS data sets [@MichaelRPergamit2001DWTN].
 
@@
 It was comprised of Blacks, Hispanics, economically disadvantaged non-Black non-Hispanics, and youth in the military.
-In 1984 and 1990, two sub-samples were dropped from the interview; the dropped subjects are the 1,079 members of the military sample and 1,643 members of the economically disadvantaged non-Black non-Hispanics respectively.
+In 1984 and 1990, two sub-samples were dropped from the interview.
+They are  1,079 members of the military sample and 1,643 members of the economically disadvantaged non-Black non-Hispanics respectively.
 Hence, 9,964 respondents remain in the eligible samples.
@@
 
-There are two ways to conduct the interview of the NLSY79, which are face to face interview or by telephone.
+There are two ways to conduct the interview of the NLSY, which are face to face interview or by telephone.
 In recent survey years, more than 90 percent of respondents were interviewed by telephone [@eliznlsy].
 
-# The NLSY79 Data Cleaning
+# Initial Data Analysis of the NLSY79
 
@@
 
-The NLSY79 data are stored in a [database](https://www.nlsinfo.org/content/cohorts/nlsy79/get-data) and could be downloaded by variables.
-Several variables are available for download and could be browsed by index.
+The NLYS data is stored in a [database](https://www.nlsinfo.org/content/cohorts/nlsy79/get-data) and could be downloaded by variables.
+Several variables are available to be downloaded and could be browsed by index.
 For the wages data set, we only extracted these variables:
@@
 
-Since our ultimate goal is to calculate the mean hourly wage, the number of jobs is calculate based on the availability of the `rate_per_hour` information.
+Since our ultimate goal is to calculate the mean hourly wage, the number of jobs is calculate based on the availability of the `rate_per_hour` information, instead of the `hours of work`.
 For example, the number of jobs of ID 1, based on `hours_work`, is two.
@@
 
-## Initial Data Analysis
+## Exploring and Finding the Anomalies
 
@@
 In addition, we also aimed to find the anomaly in the wages values using this approach.
-We mainly used graphical summary to do the IDA using `ggplot2`[@ggplot2] and `brolgar` [@brolgar].
+We mainly used graphical summary to do the IDA using `ggplot2` [@ggplot2] and `brolgar` [@brolgar].
 
 As stated previously, the respondents' ages ranged from 12 to 22 when first interviewed in 1979.
-Hence, we would like to validate whether all of the respondents were in this range.
+Hence, we would like to validate whether all of the respondents' age were in this range.
 Additionally, the [NLSY](https://www.nlsinfo.org/content/cohorts/nlsy79/intro-to-the-sample/nlsy79-sample-introduction) also provided the number of the survey cohort by their gender (6,403 males and 6,283 females) and race (7,510 Non-Black/Non-Hispanic; 3,174 Black; 2,002 Hispanic).
@@
   geom_line(alpha = 0.1) +
-  ggtitle("a)") +
+  ggtitle("A") +
+  xlab("Year") +
+  ylab("Mean hourly wage") +
   theme(plot.title = element_text(size = 10))
@@
   geom_density(aes(x = value, colour = feature, fill = feature), alpha = 0.3) +
-   ggtitle("b)") +
+   ggtitle("B") +
+   xlab("Value") +
+   ylab("Density") +
    theme(plot.title = element_text(size = 10))
@@
 It suggests that some IDs had a reasonably fluctuate wages.", fig.height=4, fig.width=5}
-
 set.seed(20210225)
@@
 We set the `mean_hourly_wage` and `year` as the dependent and predictor respectively.
-Furthermore, we used M-Estimation with Huber weighting where the observation with small residual get a weight of 1, while the larger the residual, the smaller the weight (less than 1) [@rlm].
+Furthermore, we used M-Estimation with Huber weighting where the observation with small residual get a weight of 1.
+The larger the residual, the smaller the weight (less than 1) [@rlm].
 
@@
 To, minimize the risk of being mistakenly regard an outlier as an error outlier, we have simulated some threshold and study the behavior of the spikes in each threshold.
-We found that 0.3 is the most reasonable value to be the threshold to minimize the risk of that drawback because it still capture the sensible spikes in the data.
-After deciding the threshold, we flagged the observations with the weight less than 0.3, and imputed their mean hourly wage with the models' predicted value.
+We found that 0.5 is the most reasonable value to be the threshold to minimize the risk of that drawback because it still capture the sensible spikes in the data.
+After deciding the threshold, we flagged the observations with the weight less than 0.5, and imputed their mean hourly wage with the models' predicted value.
 
@@
 wages_rlm_dat <- id_aug_w %>%
-  mutate(wages_rlm = ifelse(w < 0.3  & .fitted >= 0, .fitted,
+  mutate(wages_rlm = ifelse(w < 0.5 & .fitted >= 0, .fitted,
                             mean_hourly_wage)) %>%
-  mutate(is_pred = ifelse(w < 0.3 & .fitted >= 0, TRUE,
+  mutate(is_pred = ifelse(w < 0.5 & .fitted >= 0, TRUE,
                           FALSE)) %>%
@@
   geom_line(alpha = 0.1) +
-  ggtitle("a)") +
+  ggtitle("A") +
   theme(plot.title = element_text(size = 10))
@@
   pivot_longer(c(min, med, max), names_to = "feature", values_to = "value")
-
 feature2 <- ggplot(wages_feat_long_rlm) +
   geom_density(aes(x = value, colour = feature, fill = feature), alpha = 0.3) +
-  ggtitle("b)") +
+  ggtitle("B") +
   theme(plot.title = element_text(size = 10))
@@
 - For the future study, the room for improvement: linear mixed model and more sophisticated method to determine the threshold?
+
+# Reference

### 2021-03-10 report_alt.Rmd
Add overall flowchart.

@@
 date: "24/02/2021"
+always_allow_html: true
 ---
@@
 library(janitor)
+library(DiagrammeR)
 ```
@@
 It produced two data set,  `categories_qnames` (the observations are stored in categorical/interval values) and `new_data_qnames` (the observations are stored in integer form).
-In this case, we only used the latter.
 
@@
 We also saved the NLSY79 cohort's demographic information and the high school dropout cohort in a separate data sets.
+Here, we defined high school dropouts as respondents who only completed either $9^{th}$, $10^{th}$, $11^{th}$ grade, or who completed $12^{th}$ when their age are at least 19 years old (means that these IDs being dropped out in certain year, but they returned back to high school to complete it).
+
+The complete flow from the raw data to these data set is displayed in Figure \ref(fig:flowchart).
 
@@
 
-# Summary
+```{r nrow, echo FALSE}
+a = nrow(categories_qnames)
+b = nrow(full_demographics)
+d = eligible_wages %>%
+  group_by(id) %>%
+  count(id) %>%
+  nrow()
+n_hgc = wages_hs2020 %>%
+  group_by(id) %>%
+  count(id) %>%
+  nrow()
+n_obs_hgc = nrow(wages_hs2020)
+n_obs_hgc_pred = filter(wages_hs2020, is_pred == TRUE) %>%
+  nrow()
+
+n_do = wages_hs_dropout %>%
+  group_by(id) %>%
+  count(id) %>%
+  nrow()
+n_obs_do = nrow(wages_hs_dropout)
+n_obs_do_pred = filter(wages_hs_dropout, is_pred == TRUE) %>%
+  nrow()
+```
+
+```{r flowchart, fig.cap="The stages of data filtering from the raw data to get three datasets that are contained in an R package called 'yowie', 'n' means the number of ID, while 'n_obs' means the number of observations."}
+
+grViz("digraph flowchart {
+      node [fontname = Helvetica, shape = rectangle]        
+      tab1 [label = '@@1']
+      tab2 [label = '@@2']
+      tab3 [label = '@@3']
+      tab4 [label = '@@4']
+      tab5 [label = '@@5']
+      tab6 [label = '@@6']
+      tab7 [label = '@@7']
+
+      tab1 -> tab2;
+      tab1 -> tab3;
+      tab1 -> tab4 -> tab5 -> tab6;
+      tab5 -> tab7;
+      tab6 -> tab7;
+      tab2 -> tab7
+      }
+
+      [1]: paste0('NLSY79, n = ', a)
+      [2]: paste0('Extract demographic variable, n = ', b)
+      [3]: paste0('Exclude ', a - d, ' IDs whose hourly rate is missing')
+      [4]: paste0('Eligible ID, n = ', d)
+      [5]: paste0('Cohort whose hgc is up to 12th grade and \\n participated at least 5 years in the survey, \\n n = ', n_hgc, ', n_obs = ', n_obs_hgc, ' \\n (', n_obs_hgc_pred, ' observations are predicted value)')
+      [6]: paste0('High school dropouts cohort, \\n n = ', n_do, ', n_obs = ', n_obs_do, ' \\n (', n_obs_do_pred, ' observations are predicted value)')
+      [7]: 'yowie Package'
+      ")
+```
+
+# Exploratory Data Analysis
+
+# Summary and Discussion
 

### 2021-03-14 New_Wages.Rmd
Add getting and tidy data section.
Change the threshold to be 0.1

@@
 
-## getting the highest year of completed education
+## getting the highest year of completed education ever
 
@@
 ```{r}
-
 hours_wages <- left_join(rates_all, 
@@
                                 rate_per_hour))
+
+hours_wages <- hours_wages %>%
+  mutate(rate_per_hour_adj = ifelse(as.numeric(year) < 1994 & rate_per_hour > 1000, 1000,
+                                    ifelse(as.numeric(year) >= 1994 & rate_per_hour > 5000, 5000, rate_per_hour)))
 ```
@@
 no_job <- hours_wages %>%
-  filter(!is.na(rate_per_hour)) %>%
+  filter(!is.na(rate_per_hour_adj)) %>%
   group_by(id, year) %>%
-  summarise(no_jobs = length(rate_per_hour))
+  summarise(no_jobs = length(rate_per_hour_adj))
 
@@
 eligible_wages <- hours_wages %>%
-  filter(!is.na(rate_per_hour)) %>%
+  filter(!is.na(rate_per_hour_adj)) %>%
   left_join(no_job, by = c("id", "year"))
@@
   #calculate the weighted mean if the number of jobs > 1
-  mutate(wages = ifelse(no_jobs == 1, rate_per_hour/100,
-                        weighted.mean(rate_per_hour, hours_work, na.rm = TRUE)/100)) %>%
+  mutate(wages = ifelse(no_jobs == 1, rate_per_hour_adj/100,
+                        weighted.mean(rate_per_hour_adj, hours_work, na.rm = TRUE)/100)) %>%
   #give the flag if it the weighted mean
@@
   #calculate the arithmetic mean for the na
-  mutate(wages = ifelse(is.na(wages), mean(rate_per_hour)/100,
+  mutate(wages = ifelse(is.na(wages), mean(rate_per_hour_adj)/100,
                         wages))

### 2021-03-14 report_alt.Rmd
Change the threshold to be 0.1

@@
 library(DiagrammeR)
+library(DiagrammeRsvg)
+library(rsvg)
+library(knitr)
+library(webshot)
 ```
@@
 
-Just like the living things, the usage of the term "wild" could also be applied in the data domain.
-Wild data is the form of data that could be acquired from various source and might be generated from experiments, surveys, or may be from sensors or social media.
-Wild data also often associated with "Open Data", means that the data are freely accessible, modified, and shared by anyone for any purpose [@opendata].
+The idea of "Open Data" is that the data are freely accessible, modified, and shared by anyone for any purpose [@opendata].
+Due too its openness, Open Data are sometimes used and modified to be incorporated as the data in statistical text book or in research.
+Despite its traits, open data often have an issue with its quality in the sense that it is often not tidy and clean.
+Hence, it should be tidied and cleaned before further analysis to prevent inappropriate usage of statistical method that may lead to a misleading result [@HuebnerMariannePhD2016Asat].
 
-Despite its traits, wild data often have an issue with its quality in the sense that it is often not tidy and clean.
-Hence, it should be tidied and cleaned before further analysis to prevent inappropriate usage of statistical method that may lead to a misleading result [@HuebnerMariannePhD2016Asat].
 @HuebnerMariannePhD2016Asat also argued that data tidying and cleaning is the first part of Initial Data Analysis (IDA).
@@
           (hgc == "12TH GRADE" &
-              age_hgc >= 19)) %>%
+              age_hgc > 19)) %>%
+  filter(age_1979 <= 17,
+         gender == "MALE") %>%
   dplyr::select(-dob,
@@
 
-```{r nrow, echo FALSE}
+```{r nrow, echo=FALSE}
 a = nrow(categories_qnames)
@@
 
-```{r flowchart, fig.cap="The stages of data filtering from the raw data to get three datasets that are contained in an R package called 'yowie', 'n' means the number of ID, while 'n_obs' means the number of observations."}
+```{r flowchart, echo=FALSE, fig.cap="The stages of data filtering from the raw data to get three datasets that are contained in an R package called yowie, n means the number of ID, while n_obs means the number of observations.", fig.height=3, fig.width=3}
 
-grViz("digraph flowchart {
+fc <- grViz("digraph flowchart {
       node [fontname = Helvetica, shape = rectangle]        
@@
       [7]: 'yowie Package'
-      ")
+      ",
+      height = 200)
+
+fc
 ```
@@
 - For the future study, the room for improvement: linear mixed model and more sophisticated method to determine the threshold?
+
+- Data provider should provide validation rules for data user

### 2021-03-14 method_comparison.R

@@ -131,7 +131,7 @@ wages_sample_long <- wages_hs_sample %>%
 n_obs_sample <- wages_sample_long %>% count(id) %>%
   filter(n > 30)
 
-set.seed(20210218)
+set.seed(31251587)
 
 sample_plot <- sample(unique(n_obs_sample$id), 20)
 sample_plot <- subset(wages_sample_long, id %in% sample_plot)
@@ -194,6 +194,10 @@ wages_rlm_dat <- id_aug_w %>%
          wages_rlm_3 = ifelse(w < 0.5 & .fitted >= 0, .fitted,
                               mean_hourly_wage),
          wages_rlm_4 = ifelse(w < 0.3 & .fitted >= 0, .fitted,
+                              mean_hourly_wage),
+         wages_rlm_5 = ifelse(w < 0.1 & .fitted >= 0, .fitted,
+                              mean_hourly_wage),
+         wages_rlm_6 = ifelse(w < 0.2 & .fitted >= 0, .fitted,
                               mean_hourly_wage)) %>%
   mutate(is_pred1 = ifelse(w != 1 & .fitted >= 0, TRUE,
                           FALSE),
@@ -202,47 +206,13 @@ wages_rlm_dat <- id_aug_w %>%
          is_pred3 = ifelse(w < 0.5 & .fitted >= 0, TRUE,
                            FALSE),
          is_pred4 = ifelse(w < 0.3 & .fitted >= 0, TRUE,
+                           FALSE),
+         is_pred5 = ifelse(w < 0.1 & .fitted >= 0, TRUE,
+                           FALSE),
+         is_pred6 = ifelse(w < 0.2 & .fitted >= 0, TRUE,
                            FALSE)) %>%
   rename(wages_original = mean_hourly_wage)
 
-
-
-p4 <- ggplot(wages_rlm_dat) +
-  geom_line(aes(x = year,
-                y = wages_rlm_1,
-                group = id),
-            alpha = 0.1) +
-  ggtitle("robust linear model")
-
-p5 <- ggplot(wages_rlm_dat) +
-  geom_line(aes(x = year,
-                y = wages_original,
-                group = id),
-            alpha = 0.1) +
-  ggtitle("original data")
-
-p6 <- ggplot(wages_rlm_dat) +
-  geom_line(aes(x = year,
-                y = wages_rlm_2,
-                group = id),
-            alpha = 0.1) +
-  ggtitle("original data")
-
-p7 <- ggplot(wages_rlm_dat) +
[…]

### 2021-03-14 method_comparison.R
Examine observations with >150 US$ wages.

@@ -227,8 +227,17 @@ wages_compare <- wages_rlm_dat %>%
   pivot_longer(c(-id, -year), names_to = "type", values_to = "wages")
 
 
+ggplot(filter(wages_compare, type == "wages_rlm_5")) +
+  geom_line(aes(x = year,
+                y = wages,
+                group = id),
+            alpha = 0.1)
 
-set.seed(333222123)
+
+
+
+
+set.seed(156370113)
 
 sample_id <- sample(unique(wages_compare$id), 20)
 sample <- subset(wages_compare, id %in% sample_id)
@@ -265,6 +274,8 @@ w05 <- ggplot(filter(sample, type %in% c("wages_original", "wages_rlm_5"))) +
 
 w05
 
+
+## INSPECT OBS WITH PREDICTED VALUE (THRESHOL < 0.1)
 wages_rlm_zero_pone <- wages_rlm_dat %>%
   dplyr::select(id, year, wages_original, w, wages_rlm_5, is_pred5)
 
@@ -309,11 +320,31 @@ ggplot(filter(sum_pred_filtered2, type %in% c("wages_original", "wages_rlm_5")))
 
 
 
+## CHECK WAGES MORE THAN US$ 150
+
+id_more_than_150 <- wages_rlm_zero_pone %>%
+  filter(wages_rlm_5 > 150)
 
 
 
+wages_more_than_150 <- wages_compare %>%
+  filter(id %in% id_more_than_150$id)
 
 
+ggplot(filter(wages_more_than_150, type %in% c("wages_original", "wages_rlm_5"))) +
+  geom_line(aes(x = year,
+                y = wages,
+                colour = type,
+                linetype = type),
+            alpha = 1) +
+  geom_point(aes(x = year,
+                 y = wages,
+                 colour = type),
+             alpha = 0.5,
+             size = 1) +
+  theme(axis.text.x = element_text(angle = 10, size = 5),
+        legend.position = "bottom") +
+  facet_wrap(~id, scales = "free_y")
 
 
 

### 2021-03-17 New_Wages.Rmd
Add discussion, edit the intro and the title.

--- 2021-02-13 New_Wages.Rmd
@@
   filter(!is.na(grade)) %>%
-  select(-var)
+  dplyr::select(-var)
 
@@
   new_data_qnames %>%
-     select(CASEID_1979,
+     dplyr::select(CASEID_1979,
             starts_with("HRP") &
@@
                                 rate_per_hour))
+
 ```
@@
                         FALSE)) %>%
-  select(-flag1)
+  dplyr::select(-flag1)
 ```
@@
   geom_boxplot(aes(x = mean_hourly_wage)) +
-  facet_wrap(~year)
+  facet_wrap(~year, scales = "free_x")
 ```

### 2021-03-17 report_alt.Rmd
Add discussion, edit the intro and the title.

@@
 ---
-title: "Caring and Feeding the Wild Caught Data"
+title: "From Open to Text Book Data: The Role of Initial Data Analysis in Longitudinal Data Cleaning"
 subtitle: "A Case of The NLSY79 Data"
@@
 The idea of "Open Data" is that the data are freely accessible, modified, and shared by anyone for any purpose [@opendata].
-Due too its openness, Open Data are sometimes used and modified to be incorporated as the data in statistical text book or in research.
+Due too its openness, Open Data are sometimes utilized and incorporated as the example data in statistical text book or in research.
 Despite its traits, open data often have an issue with its quality in the sense that it is often not tidy and clean.
@@
 Furthermore, the decisions made in this stage is often unreported  in term that the IDA are often performed in an unplanned and unstructured way and is only shared among restricted party [@HuebnerMarianne2020Haar].
-Furthermore, documenting the result of IDA is utterly essential as the information preservation for the downstream statistical analyses and model building [@HuebnerMariannePhD2016Asat] and to avoid publication bias [@HuebnerMarianne2020Haar].
+In fact, documenting the result of IDA is utterly essential as the information preservation for the downstream statistical analyses and model building [@HuebnerMariannePhD2016Asat] as well as to avoid publication bias [@HuebnerMarianne2020Haar].
+It is also important to keep the data cleaning process accountable and transparent.
 
@@
 
-We chose the NLSY79 data because this data play an important role in research in various discipline including economics, sociology, education, public policy, and public health for more than a quarter of the century [@MichaelRPergamit2001DWTN].
+We chose the NLSY79 data because this data is one of the instances of Open Data that has been playing an important role in research in various disciplines including but are not limited to economics, sociology, education, public policy, and public health for more than a quarter of the century [@MichaelRPergamit2001DWTN].
 In addition, The National Longitudinal Survey is considered as survey with high retention rates and carefully designed making it suitable for a life course research [@MichaelRPergamit2001DWTN] and [@eliznlsy].
@@
 Moreover, the NLSY79 is considered as the most widely used and most important cohort in the NLSY79 data sets [@MichaelRPergamit2001DWTN].
+@SingerJudithD2003Alda used the wages and other variables of the NLSY79 data in their book in longitudinal data analysis.
+This is also being our motivation in incorporating the wages of high school cohort, specifically the high school dropouts, is that we want to make an open data to be suitable as an example data in text book or in research.
 
@@
 
+In Figure \ref{fig:high-wages}, we plotted some respondents with a high value of mean hourly wages.
+We filtered all of the IDs who earned more than US$ 500 per hour averagely.
+We found that most of these respondents only experienced one point of extremely high wages.
+Thus, we suspected that these high values are erroneous values resulted from a data entry error.
+
+```{r high-wages, echo = FALSE, fig.cap= "Some observations with extremely high mean hourly wage. Most of the IDs only have one point of high wage.", fig.height=8, fig.width=6}
+
+wages_high <- filter(wages_demog_hs, mean_hourly_wage > 500) %>%
+  as_tibble()
+
+wages_high2 <- wages_demog_hs %>%
+  filter(id %in% wages_high$id)
+
+ggplot(wages_high2) +
+  geom_line(aes(x = year,
+                y = mean_hourly_wage)) +
+  geom_point(aes(x = year,
+                y = mean_hourly_wage),
+             size = 0.5,
+             alpha = 0.5) +
+  facet_wrap(~id, scales = "free_y", ncol = 5) +
+  theme(axis.text.x = element_text(angle = 10, size = 6)) +
+  ylab("mean hourly wage")
+
+```
+
 Further, we took 36 samples randomly from the data and plotted it as is seen in Figure \ref{fig:sampleplot}.
@@
 To, minimize the risk of being mistakenly regard an outlier as an error outlier, we have simulated some threshold and study the behavior of the spikes in each threshold.
-We found that 0.3 is the most reasonable value to be the threshold to minimize the risk of that drawback because it still capture the sensible spikes in the data.
-After deciding the threshold, we flagged the observations with the weight less than 0.3, and imputed their mean hourly wage with the models' predicted value.
+We found that 0.1 is the most reasonable value to be the threshold to minimize the risk of that drawback because it still capture the sensible spikes in the data.
+After deciding the threshold, we flagged the observations with the weight less than 0.1, and imputed their mean hourly wage with the models' predicted value.
 
@@
 wages_rlm_dat <- id_aug_w %>%
-  mutate(wages_rlm = ifelse(w < 0.3  & .fitted >= 0, .fitted,
+  mutate(wages_rlm = ifelse(w < 0.1  & .fitted >= 0, .fitted,
                             mean_hourly_wage)) %>%
-  mutate(is_pred = ifelse(w < 0.3 & .fitted >= 0, TRUE,
+  mutate(is_pred = ifelse(w < 0.1 & .fitted >= 0, TRUE,
                           FALSE)) %>%
@@
 ```{r compare-data, echo = FALSE}
-set.seed(20210228)
+set.seed(31251587)
 
@@
 The orange line portray the original value of mean hourly wage, while the turquoise line display the mean hourly wages value after the extreme values imputed with the robust linear model's prediction value.
-We can see that some extreme spikes has been reduced by the model.", fig.height=4, fig.width=5}
+We can see that some extreme spikes has been reduced by the model.", fig.height=5, fig.width=6}
 ggplot(wages_compare) +
@@
             alpha = 0.5,
-            size = 0.5) +
+            size = 1) +
   theme(axis.text.x = element_text(angle = 10, size = 5),
         legend.position = "bottom") +
-  facet_wrap(~id)
+  facet_wrap(~id, scales = "free_y")
 
@@
 Here, we defined high school dropouts as respondents who only completed either $9^{th}$, $10^{th}$, $11^{th}$ grade, or who completed $12^{th}$ when their age are at least 19 years old (means that these IDs being dropped out in certain year, but they returned back to high school to complete it).
+We also filtered the data to be only male and aged between 14 and 17 years old to refresh the high school dropouts data in the `brolgar` package.
 
+We then make these three data sets and its processing documentation to be publicly available through an R data container package called `yowie`.
 The complete flow from the raw data to these data set is displayed in Figure \ref(fig:flowchart).
@@
 
-Wild caught data is often not tidy and clean.
-IDA, as a way to assess data quality should be performed and most importantly documented to avoid bias publication.
-In this paper, we performed an IDA on the NLSY79 data, specifically the wages data.
-We chose this data set because it play an important role in various discipline of research, especially the life course research.
-Getting the data from the database is pretty straightforward.
-However, it cannot be directly deployed to the analysis since the form is untidy.
-Data exploration also found some anomalies in the wages data where some IDs had an extremely high wages and large spikes.
-Hence, we performed data wrangling to tidy the data and robust linear regression model for each ID to find and treat the anomalies.
-We used the robustness weight of  less than 0.5 as threshold of the wages considered as outliers.
+One take away from this paper is to make an open data set to be suitable for a text book data or to make it ready to be incorporated in a research, an IDA should be performed and documented.
+The IDA shows that the NLSY79 data has two problems.
+Firstly, the downloaded data format is untidy, in term of its column name contains the year values, not just the variable names.
+Secondly, some anomalies i.e.
+an extreme high hourly wages are found.
+We addressed the first problem by tidying the data according to three principles of a tidy data.
+For the latter problem, we fixed the anomalies using the predicted value of the robust linear regression model.
+Further, we kept maintaining the natural variability of the wages while minimizing the presence of anomalies because of the error in the data recording.
+As an effort to provide the good practice IDA documentation, we also provided two additional variables in the final data sets, which is `is_wm` and `is_pred`.
+These variables could be a precaution for the data users in analyzing the data.
 
-- From this IDA, we learned that the open data is not only the matter of opennes, but also how that data could be analyzed flexibly with a tidy dataset.
-Hence, our recommendation that the database for every open data should be built to produce a tidy data.
+In addition to demonstrating the importance of IDA to transform the open data to be a analysis-ready data, the finding in this paper also implies that data providers should also design the database that is able to produce tidy data sets.
+A data provider should also check for data anomalies prior the data publishing or at least provides a set of rules or threshold value, for example, in this case is the threshold of reasonable wages.
+This will greatly support the data users to carry out further validation and set the same line of which data are considered as outliers by all of the data users.
+Moreover, some data cleaning tools also require validation rules.
+For example, `validates`, an R package that is powerful enough to perform a data validation, cannot be used in this case.
 
-- For the future study, the room for improvement: linear mixed model and more sophisticated method to determine the threshold?
+For the future study, we suggested the usage of linear mixed model to find and fixed the anomalies in the data set.
 
-- Data provider should provide validation rules for data user
+# References

### 2021-03-17 method_comparison.R

@@ -27,121 +27,6 @@ wages_demog <- wages_demog %>%
 # filter only the id with high school education
 wages_demog_hs <- wages_demog  %>% filter(grepl("GRADE", hgc))
 
-ggplot(wages_demog_hs, aes(x = mean_hourly_wage)) +
-  geom_boxplot()
-
-
-# get the number of observations
-n_obs <- wages_demog_hs %>% count(id)
-
-# sample some IDs
-
-set.seed(20210218)
-
-sample_id <- sample(unique(n_obs$id), 100)
-sample <- subset(wages_demog_hs, id %in% sample_id)
-
-# check the extreme value
-
-sample_ext <- sample %>%
-  group_by(id) %>%
-  identify_outliers("mean_hourly_wage") %>%
-  dplyr::select(year, id, is.outlier, is.extreme)
-
-# check the influential value using robust linear mixed model
-
-rfm <- rlmer(mean_hourly_wage ~ (1|year) + (1|id), sample)
-summary(rfm)
-
-weight_obs <- as.vector(getME(rfm, "w_e"))
-
-
-# join the weight with the data
-sample_weight <- cbind(sample, weight_obs)
-
-
-# join the extreme value with the data
-wages_hs_sample <- left_join(sample_weight, sample_ext, by = c("id", "year"))
-
-# flag the extreme value and outlier flag based on IQR methos
-wages_hs_sample <- wages_hs_sample %>%
-  mutate(is.outlier = ifelse(is.na(is.outlier), FALSE, is.outlier),
-         is.extreme = ifelse(is.na(is.extreme), FALSE, is.extreme))
-
-# add the prediction and residual of robust lmm model
-wages_hs_sample <- wages_hs_sample %>%
-  add_predictions(rfm, var = "pred") %>%
-  add_residuals(rfm, var = "res")
-
-# create a clean wages data based on robust lmm model
-wages_hs_sample <- wages_hs_sample %>%
-  mutate(wages_cleaned_rlmm = ifelse(weight_obs != 1, pred,
-                                     mean_hourly_wage),
-         is_pred = ifelse(weight_obs != 1, TRUE, FALSE),
-         wages_cleaned_rlmm = ifelse(wages_cleaned_rlmm < 0, mean_hourly_wage, wages_cleaned_rlmm))
-
-summary(wages_hs_sample$wages_cleaned_rlmm)
-
[…]

### 2021-03-17 report_alt.Rmd
Revise the introduction section.

@@
 ---
-title: "From Open to Text Book Data: The Role of Initial Data Analysis in Longitudinal Data Cleaning"
-subtitle: "A Case of The NLSY79 Data"
+title: "The Journey from Wild to Text Book Data: A Case Study of the National Longitudinal Survey of Youth"
 authors:
@@
 
-The idea of "Open Data" is that the data are freely accessible, modified, and shared by anyone for any purpose [@opendata].
-Due too its openness, Open Data are sometimes utilized and incorporated as the example data in statistical text book or in research.
-Despite its traits, open data often have an issue with its quality in the sense that it is often not tidy and clean.
+"Open data" is data that are freely accessible, modifiable, and shareable by anyone for any purpose [@opendata].
+This type of data can be useful as example data in statistical text books and for research purposes.
+However, open data are often, what we might call "wild data", because it requires substantial cleaning and tidying to tame it into text book shape.
 Hence, it should be tidied and cleaned before further analysis to prevent inappropriate usage of statistical method that may lead to a misleading result [@HuebnerMariannePhD2016Asat].
+In fact, documenting the data cleaning is utterly essential as the information preservation for the downstream statistical analyses and model building [@HuebnerMariannePhD2016Asat] as well as to avoid publication bias [@HuebnerMarianne2020Haar].
+Making the data cleaning process accountable and transparent is imperative.
 
-@HuebnerMariannePhD2016Asat also argued that data tidying and cleaning is the first part of Initial Data Analysis (IDA).
-The second and third steps are to explore the data properties and to document and report the process for the later formal analysis.
-This second step is the same with Chatfield's perspective on the goal of IDA [@Chatfield1985TIEo], which is to get the "feel" of the data.
-In other word, IDA could be regarded as a way to assess the data quality.
+@HuebnerMariannePhD2016Asat considers data cleaning to be part of "initial data analysis (IDA)".
+The other IDA steps are to explore the data properties and to document and report the process for the later formal analysis.
+This is similar to Chatfield's perspective on IDA [@Chatfield1985TIEo], which is to get the "feel" of the data.
+@DasuTamraparni2003Edma say that data cleaning and exploration, without naming it as IDA, is a difficult task and consumes 80% of the data mining task.
 
-@DasuTamraparni2003Edma mentioned that data cleaning and exploration, which is the IDA notion, is a difficult task and determine 80 percent of the data mining result.
 Despite its importance, this stage is often undervalued and neglected [@Chatfield1985TIEo].
-@WickhamHadley2014TD also stated that there has been few research of the good practice of data cleaning.
-Furthermore, the decisions made in this stage is often unreported  in term that the IDA are often performed in an unplanned and unstructured way and is only shared among restricted party [@HuebnerMarianne2020Haar].
-In fact, documenting the result of IDA is utterly essential as the information preservation for the downstream statistical analyses and model building [@HuebnerMariannePhD2016Asat] as well as to avoid publication bias [@HuebnerMarianne2020Haar].
-It is also important to keep the data cleaning process accountable and transparent.
+There are few research papers that document the data cleaning [@WickhamHadley2014TD].
+Furthermore, the decisions made in this stage often go unreported  in term that the IDA are often performed in an unplanned and unstructured way and is only shared among restricted party [@HuebnerMarianne2020Haar].
 
@@
 @SingerJudithD2003Alda used the wages and other variables of the NLSY79 data in their book in longitudinal data analysis.
-This is also being our motivation in incorporating the wages of high school cohort, specifically the high school dropouts, is that we want to make an open data to be suitable as an example data in text book or in research.
+This is also being our motivation in incorporating the wages of high school cohort, specifically the high school dropouts, is that we want to refresh this data set as well as demonstrating the process taken to transform and open data to be suitable as an example data in text book or in research.
 

### 2021-03-17 method_comparison.R

@@ -142,7 +142,9 @@ ggplot(sample) +
 
 # plot by threshold
 
-w05 <- ggplot(filter(sample, type %in% c("wages_original", "wages_rlm_5"))) +
+for_plot <- dplyr::filter(sample, type %in% c("wages_original", "wages_rlm_5"))
+
+w05 <- ggplot(for_plot) +
   geom_line(aes(x = year,
                 y = wages,
                 colour = type,

### 2021-03-18 method_comparison.R
Simulated wages imputation with w < 0.12 threshold.

@@ -82,7 +82,7 @@ wages_rlm_dat <- id_aug_w %>%
                               mean_hourly_wage),
          wages_rlm_5 = ifelse(w < 0.1 & .fitted >= 0, .fitted,
                               mean_hourly_wage),
-         wages_rlm_6 = ifelse(w < 0.15 & .fitted >= 0, .fitted,
+         wages_rlm_6 = ifelse(w < 0.12 & .fitted >= 0, .fitted,
                               mean_hourly_wage)) %>%
   mutate(is_pred1 = ifelse(w != 1 & .fitted >= 0, TRUE,
                           FALSE),
@@ -94,7 +94,7 @@ wages_rlm_dat <- id_aug_w %>%
                            FALSE),
          is_pred5 = ifelse(w < 0.1 & .fitted >= 0, TRUE,
                            FALSE),
-         is_pred6 = ifelse(w < 0.15 & .fitted >= 0, TRUE,
+         is_pred6 = ifelse(w < 0.12 & .fitted >= 0, TRUE,
                            FALSE)) %>%
   rename(wages_original = mean_hourly_wage)
 
@@ -142,7 +142,7 @@ ggplot(sample) +
 
 # plot by threshold
 
-for_plot <- dplyr::filter(sample, type %in% c("wages_original", "wages_rlm_5"))
+for_plot <- dplyr::filter(sample, type %in% c("wages_original", "wages_rlm_6"))
 
 w05 <- ggplot(for_plot) +
   geom_line(aes(x = year,
@@ -164,14 +164,14 @@ w05
 
 ## INSPECT OBS WITH PREDICTED VALUE (THRESHOLD < 0.1)
 wages_rlm_zero_pone <- wages_rlm_dat %>%
-  dplyr::select(id, year, wages_original, w, wages_rlm_5, is_pred5)
+  dplyr::select(id, year, wages_original, w, wages_rlm_6, is_pred6)
 
 sum_id <- wages_rlm_zero_pone %>%
   group_by(id) %>%
   summarise(n_obs = length(id))
 
 sum_pred <- wages_rlm_zero_pone %>%
-  filter(is_pred5 == TRUE) %>%
+  filter(is_pred6 == TRUE) %>%
   group_by(id) %>%
   summarise(n_pred = length(id))
 
@@ -190,7 +190,7 @@ sum_pred_filtered2 <- wages_compare %>%
 
 
 
[…]

### 2021-03-23 report_alt.Rmd
Change the threshold weight and add EDA.

@@
 wages_rlm_dat <- id_aug_w %>%
-  mutate(wages_rlm = ifelse(w < 0.1  & .fitted >= 0, .fitted,
+  mutate(wages_rlm = ifelse(w < 0.12  & .fitted >= 0, .fitted,
                             mean_hourly_wage)) %>%
-  mutate(is_pred = ifelse(w < 0.1 & .fitted >= 0, TRUE,
+  mutate(is_pred = ifelse(w < 0.12 & .fitted >= 0, TRUE,
                           FALSE)) %>%
@@
 
-Figure \ref{fig:comppict} A shows that after imputing the "error outliers" with the models' predicted value, the highest wages value has decreased to be around US$250.
+Figure \ref{fig:comppict} A shows that after imputing the anomalies with the models' predicted value, the highest wages value has decreased to be around $350.
 The spikes were still observed, but are not as extreme as the original data set.
@@
 ```{r comppict, echo = FALSE, fig.cap = "The distribution of the mean hourly wage after treating the extreme values.
-Plot A portrays the pattern of mean hourly wage of high school cohort from 1979 to 2018 of each ID in US Dollar; Plot B shows the distribution of their minimum, median, and maximum value.
+Plot A portrays the pattern of mean hourly wage of high school cohort from 1979 to 2018 of each ID in US Dollar; Plot B shows the distribution of their minimum, median, and maximum value transformed to log10 scale.
 We can see that some observations still had reasonbaly higher wages than the others.
@@
   ggtitle("b)") +
-  theme(plot.title = element_text(size = 10))
+  scale_x_log10() + 
+  theme(plot.title = element_text(size = 10)) +
+  xlab("value (log10)")
 
@@
 
-# Summary and Discussion
+In this part, we gave some examples of how the data might be used for data analysis in text book.
+The first example is that we observed the relationship between wage and the respondents' demographic variables.
+Using `features` function in `brolgar`, we extracted the median of the mean hourly age of each ID and examine its relationship with respondents' gender, race, and education.
+
+```{r}
+wages_hs2020_tsibble <- wages_hs2020 %>%
+  as_tsibble(key = id,
+             index = year,
+             regular = FALSE)
+
+wages_hs2020_feature <- wages_hs2020_tsibble %>%
+  features(mean_hourly_wage,
+           feat_three_num)
+
+wages_median_demog <- left_join(wages_hs2020_feature, demographic_nlsy79, by = "id") %>%
+  mutate(hgc = factor(hgc, levels = c("1ST GRADE", "2ND GRADE", "3RD GRADE", 
+                                      "4TH GRADE", "5TH GRADE", "6TH GRADE", 
+                                      "7TH GRADE", "8TH GRADE", "9TH GRADE",
+                                      "10TH GRADE", "11TH GRADE", "12TH GRADE", 
+                                      "UNGRADED")))
+```
+
+```{r}
+ggplot(filter(wages_median_demog, hgc != "UNGRADED"), aes(x = gender, y = med, color = gender)) +
+  geom_boxplot() +
+  facet_wrap(~hgc, scales = "free_y") +
+  theme_bw()
+```
+```{r}
+wages_hs2020_grouped <- wages_hs2020 %>%
+  group_by(race, year) %>%
+  summarise(median = median(mean_hourly_wage))
+
+ggplot(wages_hs2020_grouped) +
+  geom_line(aes(x = year,
+                y = median,
+                group = race,
+                colour = race)) +
+  theme_classic()
+```
+
+# Summary
 
@@
 
-In addition to demonstrating the importance of IDA to transform the open data to be a analysis-ready data, the finding in this paper also implies that data providers should also design the database that is able to produce tidy data sets.
+In addition to demonstrating the importance of IDA to transform the open data to be a analysis-ready data, the finding in this paper also implies that data providers should design the database that is able to produce tidy data sets.
 A data provider should also check for data anomalies prior the data publishing or at least provides a set of rules or threshold value, for example, in this case is the threshold of reasonable wages.
-This will greatly support the data users to carry out further validation and set the same line of which data are considered as outliers by all of the data users.
+This will greatly support the data users to carry out further validation and set the same understanding of which data are considered as outliers by all of the data users.
 Moreover, some data cleaning tools also require validation rules.

### 2021-03-23 method_comparison.R

@@ -122,7 +122,7 @@ ggplot(filter(wages_compare, type == "wages_rlm_6")) +
 
 
 
-set.seed(156370113)
+set.seed(31251587)
 
 sample_id <- sample(unique(wages_compare$id), 20)
 sample <- subset(wages_compare, id %in% sample_id)

### 2021-03-24 report_alt.Rmd
Add explanation of eda plots.

@@
   geom_line(alpha = 0.1) +
-  ggtitle("a)") +
+  ggtitle("A)") +
   theme(plot.title = element_text(size = 10))
@@
   geom_density(aes(x = value, colour = feature, fill = feature), alpha = 0.3) +
-  ggtitle("b)") +
+  ggtitle("B)") +
   theme(plot.title = element_text(size = 10))
@@
   geom_density(aes(x = value, colour = feature, fill = feature), alpha = 0.3) +
-  ggtitle("b)") +
+  ggtitle("B)") +
   scale_x_log10() + 
@@
 In this part, we gave some examples of how the data might be used for data analysis in text book.
-The first example is that we observed the relationship between wage and the respondents' demographic variables.
-Using `features` function in `brolgar`, we extracted the median of the mean hourly age of each ID and examine its relationship with respondents' gender, race, and education.
+The first example is that we observed relationship between wages and education.
+Using `features` function in `brolgar`, we extracted the median of the mean hourly age of each ID and plotted according to highest grade completed.
+We found in Figure \ref{fig:eda-plot} A that the higher the education completed does not necessarily means the higher wages earned.
+However, people who completed 12th grade tend to have higher wage compared to the other groups.
+Further, respondents who only completed 4th grade happen to have median about as high as the respondents who completed 10th grade.
 
-```{r}
+Secondly, we examined the pattern of mean hourly wage corresponding to different groups of race and gender.
+Figure \ref{fig:eda-plot} B and C suggest that in general, the mean hourly wage median increased over the years.
+Regarding the race, we found that Non-Black Non-Hispanic group had a higher wages than the other groups.
+We can also see a reasonable gap between the wages median of Black people and the Hispanic and the Non-Black Non-Hispanic.
+Regarding the gender, male earn more than female, and unfortunately, the gap keep increasing over time.
+
+```{r, echo = FALSE}
 wages_hs2020_tsibble <- wages_hs2020 %>%
@@
 
-```{r}
-ggplot(filter(wages_median_demog, hgc != "UNGRADED"), aes(x = gender, y = med, color = gender)) +
+```{r eda-plot, echo = FALSE, fig.height=6, fig.width=6, fig.cap="The example of EDA plot.
+Plot A shows the pattern of the mean hourly wage median against the highest grade completed.
+It shows that the higher eductaion completed does not necessarily means the more wages earned.
+Plot B and C show the mean hourly wage median over the years based on the race and gender respectively.
+Generally, the respondents experienced a positive trend over time despite of what their race and gender."}
+hgc_plot <- ggplot(filter(wages_median_demog, hgc != "UNGRADED"), aes(y = hgc, x = med)) +
   geom_boxplot() +
-  facet_wrap(~hgc, scales = "free_y") +
-  theme_bw()
-```
-```{r}
+  theme(axis.text.y = element_text(size = 6),
+        axis.title = element_text(size = 9)) +
+  ggtitle("A)") +
+  xlab("median of mean hourly wage (US$)") +
+  ylab("highest grade completed")
+
 wages_hs2020_grouped <- wages_hs2020 %>%
@@
 
-ggplot(wages_hs2020_grouped) +
+race_plot <- ggplot(wages_hs2020_grouped) +
   geom_line(aes(x = year,
@@
                 colour = race)) +
-  theme_classic()
+  theme(legend.position = "bottom",
+        legend.text = element_text(size = 5),
+        axis.title = element_text(size = 9)) +
+  scale_color_brewer(palette = "Accent") +
+  ggtitle("B)") +
+  ylab("median of mean hourly wage (US$)")
+
+wages_hs2020_grouped_gender <- wages_hs2020 %>%
+  group_by(gender, year) %>%
+  summarise(median = median(mean_hourly_wage))
+
+gender_plot <- ggplot(wages_hs2020_grouped_gender) +
+  geom_line(aes(x = year,
+                y = median,
+                group = gender,
+                colour = gender)) +
+  theme(legend.position = "bottom",
+        legend.text = element_text(size = 5),
+        axis.title = element_text(size = 9)) +
+  scale_color_brewer(palette = "Dark2") +
+  ggtitle("C)") +
+  ylab("median of mean hourly wage (US$)")
+
+hgc_plot / (race_plot + gender_plot)
 ```
@@
 
-In addition to demonstrating the importance of IDA to transform the open data to be a analysis-ready data, the finding in this paper also implies that data providers should design the database that is able to produce tidy data sets.
+In addition, the finding in this paper implies that data providers should design the database that is able to produce tidy data sets.
 A data provider should also check for data anomalies prior the data publishing or at least provides a set of rules or threshold value, for example, in this case is the threshold of reasonable wages.
-This will greatly support the data users to carry out further validation and set the same understanding of which data are considered as outliers by all of the data users.
-Moreover, some data cleaning tools also require validation rules.
-For example, `validates`, an R package that is powerful enough to perform a data validation, cannot be used in this case.
+This will greatly support the data users to carry out further validation and set the same understanding of which data are considered as outliers.
+Moreover, providing validation rules would facilitate the usage of any establish data validation tool, such as `validates` package.
+In this case, we cannot use this handy validation package due to the absence of validation rules.
 

### 2021-03-24 app/app.R
Add a shiny app to do threshold simulation.

@@ -1,0 +1,132 @@
+#
+# This is a Shiny web application. You can run the application by clicking
+# the 'Run App' button above.
+#
+# Find out more about building applications with Shiny here:
+#
+#    http://shiny.rstudio.com/
+#
+
+library(shiny)
+library(tidyverse)
+
+
+# LOAD THE DATA
+
+wages_hs <- readRDS(here::here("app/wages_hs.rds")) %>%
+    dplyr::select(-`X1.x`, -`X1.y`)
+
+# Define UI for application that draws a histogram
+ui <- fluidPage(
+
+    # Application title
+    titlePanel("Customize Your Threshold"),
+
+    # Sidebar with a slider input for number of bins
+    fluidRow(
+        column(width = 4,
+            sliderInput("th",
+                        "Set the weight threshold:",
+                        min = 0.01,
+                        max = 1,
+                        value = 5),
+            br(),
+            br(),
+            textInput("seed",
+                      "Set the seed:",
+                      value = "123")
+
+        ),
+
+        # Show a plot of the generated distribution
+        column(width = 8,
+           plotOutput("compar_plot",
+                      height = 500)
+        )
+    ),
+    fluidRow(
+        column(
+            width = 12,
+            h2("Potential Outliers"),
+            plotOutput("outlier_plot",
+                       height = 800)
+
+        )
+    )
+)
+
+# Define server logic required to draw a histogram
+server <- function(input, output) {
+
+    thres_data <- reactive({
+        temp_data <- wages_hs %>%
+            mutate(wages_rlm = ifelse(w < input$th  & .fitted >= 0, .fitted,
+                                      mean_hourly_wage)) %>%
+            mutate(is_pred = ifelse(w < input$th & .fitted >= 0, TRUE,
+                                    FALSE))
+    })
+
+    compare_data <- reactive({
+        set.seed(input$seed)
+
+        sample_id <- sample(unique(thres_data()$id), 30)
+        sample <- subset(thres_data(), id %in% sample_id)
+
+        wages_compare <- sample %>%
+            dplyr::select(id, year, mean_hourly_wage, wages_rlm) %>%
+            rename(mean_hourly_wage_rlm = wages_rlm) %>%
+            pivot_longer(c(-id, -year), names_to = "type", values_to = "wages")
+    })
+
+    pot_out_data <- reactive({
+        wages_high <- filter(wages_hs, mean_hourly_wage > 500) %>%
+            as_tibble()
[…]

### 2021-03-24 method_comparison.R

@@ -30,7 +30,7 @@ wages_demog_hs <- wages_demog  %>% filter(grepl("GRADE", hgc))
 
 ##### ROBUST LINEAR MODEL
 
-keep_me <- wages_demog_hs %>% count(id) %>% filter(n >4)
+keep_me <- wages_demog_hs %>% count(id) %>% filter(n > 4)
 
 wages_demog_hs <- wages_demog_hs %>%
   filter(id %in% keep_me$id)
@@ -70,6 +70,8 @@ id_aug_w <- cbind(id_aug, id_w) %>%
                 .sigma,
                 w) %>%
   rename(id = `id...1`)
+
+
 
 wages_rlm_dat <- id_aug_w %>%
   mutate(wages_rlm_1 = ifelse(w != 1 & .fitted >= 0, .fitted,

### 2021-03-24 app/app.R
Edit the shiny app.

@@ -50,7 +50,28 @@ ui <- fluidPage(
             h2("Potential Outliers"),
             plotOutput("outlier_plot",
                        height = 800)
-
+            )
+    ),
+    fluidRow(
+        column(
+            width = 6,
+            h2("The Spaghetti Plots"),
+            plotOutput("spag1"),
+            br(),
+            textOutput("summary_before"),
+            textOutput("n_obs"),
+            textOutput("n_id")
+        ),
+        column(
+            width = 6,
+            br(),
+            br(),
+            br(),
+            plotOutput("spag2"),
+            br(),
+            textOutput("summary_after"),
+            textOutput("n_obs_pred"),
+            textOutput("n_id_pred")
         )
     )
 )
@@ -122,9 +143,63 @@ server <- function(input, output) {
             facet_wrap(~id, scales = "free_y", ncol = 5) +
             theme(axis.text.x = element_text(angle = 10, size = 6)) +
             ylab("mean hourly wage")
+    })
 
+    output$spag1 <- renderPlot({
 
+        ggplot(wages_hs) +
+            geom_line(aes(x = year,
+                          y = mean_hourly_wage,
+                          group = id),
+                      alpha = 0.1) +
+            ggtitle("Before")
+    })
 
+    output$spag2 <- renderPlot({
+
+        ggplot(thres_data()) +
+            geom_line(aes(x = year,
+                          y = wages_rlm,
+                          group = id),
+                      alpha = 0.1) +
+            ggtitle("After")
+    })
+
+    output$summary_before <- renderText({
+
+        summary(wages_hs$mean_hourly_wage)
+    })
+
+    output$summary_after <- renderText({
+        summary(thres_data()$wages_rlm)
+    })
+
+    output$n_obs <- renderText({
+        paste0("Number of observations: ", nrow(wages_hs))
+    })
+
+    output$n_id <- renderText({
+        n_id <- wages_hs %>%
+            group_by(id) %>%
+            count(id) %>%
+            nrow()
+
+        paste0("Number of IDs: ", n_id)
+    })
+
+    output$n_obs_pred <- renderText({
+        n_obs_pred <- filter(thres_data(), is_pred == TRUE) %>%
+            nrow()
+        paste0("Number of observations predicted: ", n_obs_pred)
+    })
+
+    output$n_id_pred <- renderText({
+        n_id_pred <- filter(thres_data(), is_pred == TRUE) %>%
[…]

### 2021-04-07 report_alt.Rmd
Edit the introduction and summary.

@@
 However, open data are often, what we might call "wild data", because it requires substantial cleaning and tidying to tame it into text book shape.
-Hence, it should be tidied and cleaned before further analysis to prevent inappropriate usage of statistical method that may lead to a misleading result [@HuebnerMariannePhD2016Asat].
-In fact, documenting the data cleaning is utterly essential as the information preservation for the downstream statistical analyses and model building [@HuebnerMariannePhD2016Asat] as well as to avoid publication bias [@HuebnerMarianne2020Haar].
-Making the data cleaning process accountable and transparent is imperative.
+@HuebnerMariannePhD2016Asat emphasize that making the data cleaning process accountable and transparent is imperative.
+Documenting the data cleaning is essential for the integrity of downstream statistical analyses and model building [@HuebnerMarianne2020Haar].
 
-@HuebnerMariannePhD2016Asat considers data cleaning to be part of "initial data analysis (IDA)".
-The other IDA steps are to explore the data properties and to document and report the process for the later formal analysis.
-This is similar to Chatfield's perspective on IDA [@Chatfield1985TIEo], which is to get the "feel" of the data.
+Data cleaning is part of what is called "initial data analysis (IDA)" [@Chatfield1985TIEo].
+The other IDA steps are to explore the data properties and check assumption of modeling to document and report the process for the later formal analysis.
+Hence, exploratory data analysis, which is defined by @tukey as a detective work to get the clue from data, either numerically or graphically, before confirmatory data analysis is performed, encompasses IDA.
 @DasuTamraparni2003Edma say that data cleaning and exploration, without naming it as IDA, is a difficult task and consumes 80% of the data mining task.
 
-Despite its importance, this stage is often undervalued and neglected [@Chatfield1985TIEo].
+Despite its importance, this IDA stage is often undervalued and neglected [@Chatfield1985TIEo].
 There are few research papers that document the data cleaning [@WickhamHadley2014TD].
-Furthermore, the decisions made in this stage often go unreported  in term that the IDA are often performed in an unplanned and unstructured way and is only shared among restricted party [@HuebnerMarianne2020Haar].
+Furthermore, the decisions made in this stage often go unreported  in the sense that IDA is often performed in an unplanned and unstructured way, and is only shared among restricted parties [@HuebnerMarianne2020Haar].
 
-This paper aims to demonstrate the steps of IDA, which is tidying, cleaning, and documenting the data.
-We used the longitudinal data from the National Longitudinal Survey of Youth 1979 (NLSY79) or is known as the NLSY79.
-Particularly, we are interested in doing the IDA to the wages and several demographic variables of the NLSY79 cohort whose highest grade are up to 12th grade and the high school dropouts.
+This paper aims to demonstrate the steps of IDA,  tidying, cleaning, and documenting the process, for a prominent open data source, National Longitudinal Survey of Youth (NLSY) 1979, henceforth referred to NSLY79.
+This data has been playing an important role in research in various disciplines including but are not limited to economics, sociology, education, public policy, and public health for more than a quarter of the century [@MichaelRPergamit2001DWTN].
+In addition, The National Longitudinal Survey is considered as survey with high retention rates and carefully designed making it suitable for a life course research [@MichaelRPergamit2001DWTN] and [@eliznlsy].
+According to @eliznlsy, thousand of articles, and hundreds of book chapters and monographs has incorporated the NLSY data.
+Moreover, the NLSY79 is considered as the most widely used and most important cohort in the NLSY79 data sets [@MichaelRPergamit2001DWTN].
 
-We chose the NLSY79 data because this data is one of the instances of Open Data that has been playing an important role in research in various disciplines including but are not limited to economics, sociology, education, public policy, and public health for more than a quarter of the century [@MichaelRPergamit2001DWTN].
-In addition, The National Longitudinal Survey is considered as survey with high retention rates and carefully designed making it suitable for a life course research [@MichaelRPergamit2001DWTN] and [@eliznlsy].
-According to @eliznlsy as of 2015, thousand of articles, and hundreds of book chapters and monographs has incorporated the NLSY79 data.
-Moreover, the NLSY79 is considered as the most widely used and most important cohort in the NLSY79 data sets [@MichaelRPergamit2001DWTN].
-@SingerJudithD2003Alda used the wages and other variables of the NLSY79 data in their book in longitudinal data analysis.
-This is also being our motivation in incorporating the wages of high school cohort, specifically the high school dropouts, is that we want to refresh this data set as well as demonstrating the process taken to transform and open data to be suitable as an example data in text book or in research.
+@SingerJudithD2003Alda used the  wages and other variables of high school dropouts from the NLSY79 data as an example data set to illustrate longitudinal data modeling.
+Our aim is to refresh this text book data to be more up to date.
+Here, we investigate the process of getting from the raw NLSY79 to this text book data set.
+However, we are not able to create the exact same data set as published in their book since we do not have the information of what age threshold did they use to determine the high school dropouts.
+
+This paper is structured to have 5 sections.
+Section 2 describes the original data source.
+Section 3 presents the steps of cleaning the data, including getting and tidying the data from the NLSY79 and initial data analysis to find and treat the anomalies in the data.
+We also gave examples of exploratory data analysis using the clean data in Section 4.
+Finally, Section 6 summarizes the paper.
 
@@
 Moreover, it should be noted that not all the outliers is due to an error, instead it might be that one had a reasonably increasing or decreasing wages.
-To, minimize the risk of being mistakenly regard an outlier as an error outlier, we have simulated some threshold and study the behavior of the spikes in each threshold.
-We found that 0.1 is the most reasonable value to be the threshold to minimize the risk of that drawback because it still capture the sensible spikes in the data.
-After deciding the threshold, we flagged the observations with the weight less than 0.1, and imputed their mean hourly wage with the models' predicted value.
+To, minimize the risk of being mistakenly regard an outlier as an "erronous outlier", we have simulated some threshold and study the behavior of the spikes in each threshold.
+We found that 0.12 is the most reasonable value to be the threshold to minimize the risk of that drawback because it still capture the sensible spikes in the data.
+In other words, we kept maintaining the natural variability of the wages while minimizing the presence of anomalies because of the error in the data recording.
+After deciding the threshold, we imputed the observations with weight less than 0.2 wage with the models' predicted value.
+We then flagged those observations in a new variable called `is_pred`,
 
@@
 The spikes were still observed, but are not as extreme as the original data set.
-In Figure \ref{fig:comppict} B) although the distributions of the features are still positively skewed, we can still examine it clearly the difference shape of those features.
-The minimum value is heavily skewed, means that most the subjects have a small minimum value of wages, but there are still extreme cases where their minimum wages was extremely higher than others.
-Moreover, these cases had a minimum wages that is higher than others' maximum wages.
+In Figure \ref{fig:comppict} B), we plotted the three features of mean hourly wages, namely the minimum, median, an maximum value, transformed to log scale.
+The plot implies that the skewness are slightly negative.
+We also see that the three features are overlapped each other.
+It indicates that some ID's minimum wage are higher than some ID's maximum wage.
 
-Furthermore, the robust linear regression has reduce the level of noise in the data set as is seen in Figure \ref{fig:compare}.
-The figure also implies that after the treatment, the fluctuation can still be observed in the data and only the large spikes, which are considered as "error outliers", are eliminated from the data.
+Furthermore, Figure \ref{fig:compare} implies that after the treatment, the fluctuation can still be observed in the data and only the large spikes, which are considered as "erronous outliers", are eliminated from the data.
 Hence, the model results a data set with the reasonable degree of fluctuation.
@@
 We can see that some observations still had reasonbaly higher wages than the others.
-The minimum, median, and maximum distribution is positively skewed, where some IDs' have a minimum wages that is higher than others' maximum wages.", fig.width=6, fig.height=4}
+Also, some IDs' have a minimum wages that is higher than others' maximum wages.", fig.width=6, fig.height=4}
 
@@
 hgc_plot / (race_plot + gender_plot)
+
+#### using brolgar to explore the individual experiences <- change the EDA
 ```
@@
 
-One take away from this paper is to make an open data set to be suitable for a text book data or to make it ready to be incorporated in a research, an IDA should be performed and documented.
-The IDA shows that the NLSY79 data has two problems.
-Firstly, the downloaded data format is untidy, in term of its column name contains the year values, not just the variable names.
-Secondly, some anomalies i.e.
-an extreme high hourly wages are found.
-We addressed the first problem by tidying the data according to three principles of a tidy data.
-For the latter problem, we fixed the anomalies using the predicted value of the robust linear regression model.
-Further, we kept maintaining the natural variability of the wages while minimizing the presence of anomalies because of the error in the data recording.
-As an effort to provide the good practice IDA documentation, we also provided two additional variables in the final data sets, which is `is_wm` and `is_pred`.
-These variables could be a precaution for the data users in analyzing the data.
+This paper has performed a set of stages to make an open data suitable for a text book data or to make it ready for research.
+In the first stage, we showed the steps that are performed to get the data from the NLSY79 database.
+Since the data format is untidy, we showed how the data has been tidied.
+After that, we conducted an initial data analysis to investigate and screen the quality of the data.
+Using robust linear regression model, we found and fixed the anomalous observations in the data set with its predicted values.
+We also performed an example of exploratory data analysis using the cleaned data set.
 
-In addition, the finding in this paper implies that data providers should design the database that is able to produce tidy data sets.
+This paper also has demonstrated the documentation of data cleaning.
+We also provided all of the codes in performing data tidying and initial data analysis.
+Thus, this paper provided an opportunity to continuously refresh the text book data whenever the updated data is published in the NLSY79 database.
+This could be done by following the documentation of the code that is provided in this paper.
+
+Moreover, the documentation also includes the process of how did we generate the robustness weight and how did we decide the threshold of anomalous observations.
+It is also documented along with the flag of whether an observation is imputed value or not.
+Accordingly, if somebody wish to make another decision, it can be done by making a small changes in the code provided.
+Further, this paper is also supplemented by a `shiny`  [@shiny] app as a simulation tool to customize the weight threshold.
+
+Finally, this paper implies that data providers should design the database that is able to produce tidy data sets.
 A data provider should also check for data anomalies prior the data publishing or at least provides a set of rules or threshold value, for example, in this case is the threshold of reasonable wages.
 This will greatly support the data users to carry out further validation and set the same understanding of which data are considered as outliers.
-Moreover, providing validation rules would facilitate the usage of any establish data validation tool, such as `validates` package.
+Moreover, providing validation rules would facilitate the usage of any establish data validation tool, such as `validate` [@validate] package.
 In this case, we cannot use this handy validation package due to the absence of validation rules.
 
-For the future study, we suggested the usage of linear mixed model to find and fixed the anomalies in the data set.
+## Acknowledgements (Aarathy, software recources (R packages used to write the paper and packages that are used for entire analysis) )
+
+## Supplementary materials (github repo for, package, and shiny app)
 

### 2021-04-07 app/app.R

@@ -8,6 +8,7 @@
 #
 
 library(shiny)
+library(shinythemes)
 library(tidyverse)
 
 
@@ -16,68 +17,85 @@ library(tidyverse)
 wages_hs <- readRDS(here::here("app/wages_hs.rds")) %>%
     dplyr::select(-`X1.x`, -`X1.y`)
 
-# Define UI for application that draws a histogram
-ui <- fluidPage(
 
-    # Application title
-    titlePanel("Customize Your Threshold"),
+## Make it separate tab for these three plots
 
-    # Sidebar with a slider input for number of bins
-    fluidRow(
-        column(width = 4,
-            sliderInput("th",
-                        "Set the weight threshold:",
-                        min = 0.01,
-                        max = 1,
-                        value = 5),
-            br(),
-            br(),
-            textInput("seed",
-                      "Set the seed:",
-                      value = "123")
+## Add progress bar for how long will it take to recalculate threshold and resample the ids
 
-        ),
+ui <- bootstrapPage(
+    navbarPage(theme = shinytheme("cerulean"),
+               collapsible = TRUE,
+               # Application title
+               "Customize Your Threshold",
+               tabPanel("Set the weight threshold",
+                        sidebarLayout(
+                            sidebarPanel(width = 4,
+                                   sliderInput("th",
+                                               "Set the weight threshold:",
+                                               min = 0.01,
+                                               max = 1,
+                                               value = 5),
+                                   br(),
+                                   br(),
+                                   actionButton("resample",
+                                             "Compare individuals")
 
-        # Show a plot of the generated distribution
-        column(width = 8,
-           plotOutput("compar_plot",
-                      height = 500)
-        )
-    ),
-    fluidRow(
-        column(
-            width = 12,
-            h2("Potential Outliers"),
-            plotOutput("outlier_plot",
-                       height = 800)
-            )
-    ),
-    fluidRow(
-        column(
-            width = 6,
-            h2("The Spaghetti Plots"),
-            plotOutput("spag1"),
-            br(),
-            textOutput("summary_before"),
-            textOutput("n_obs"),
-            textOutput("n_id")
-        ),
-        column(
-            width = 6,
-            br(),
-            br(),
-            br(),
-            plotOutput("spag2"),
-            br(),
-            textOutput("summary_after"),
-            textOutput("n_obs_pred"),
-            textOutput("n_id_pred")
+                            ),
+
+                            # Show a plot of the samples of indivudial before and after
+                            mainPanel(width = 8,
[…]

### 2021-04-07 report_alt.Rmd
Add acknowledgment and supplementary materials.

@@
 library(DiagrammeR)
-library(DiagrammeRsvg)
 library(rsvg)
-library(knitr)
 library(webshot)
@@
 
-The NLSY79 is a longitudinal survey held by the U.S Bureau of Labor Statistics that follows the lives of a sample of American youth and born between 1957-1964.
+The NLSY79 is a longitudinal survey held by the U.S Bureau of Labor Statistics that follows the lives of a sample of American youth and born between 1957-1964 [@nlsy79].
 The cohort originally included 12,686 respondents ages 14-22 when first interviewed in 1979.
@@
 
-## Acknowledgements (Aarathy, software recources (R packages used to write the paper and packages that are used for entire analysis) )
+# Acknowledgements
 
-## Supplementary materials (github repo for, package, and shiny app)
+We would like to thank Aarathy Babu for the insight and discussion during the writing of this paper.
+
+The entire analysis is conducted using `R` [@R] in `rstudio` using these packages: `tidyverse` [@tidyverse], `ggplot2` [@ggplot2], `dplyr` [@dplyr], `readr` [@readr], `tidyr` [@tidyr], `stringr` [@stringr], `purrr` [@purrr], `broom` [@broom], `blorgar` [@brolgar], `patchwork` [@patchwork], `kableExtra` [@kableExtra], `MASS` [@mass], `janitor` [@janitor], `DiagrammeR` [@DiagrammeR], `rsvg` [@rsvg], and `webshot` [@webshot].
+The paper are generated using `knitr` [@knitr], `rmarkdown` [@rmarkdown], and `rticles` [@rticles].
+
+# Supplementary Materials
+
+**Codes** : R script to reproduce data tidying and cleaning are available in this [Github Repository](https://github.com/numbats/yowie/blob/master/data-raw/data_preprocessing.R)
+
+**R Package `yowie`**:`yowie` is a data container R package that contains 3 datasets, namely the high school mean hourly wage data, high school dropouts mean hourly wage data, and demographic data of the NLSY79 cohort.
+This package could be accessed [here](https://github.com/numbats/yowie).
+
+**shiny app**: An web interactive `shiny` app to run a simulation to customize the weight threshold. This app could be accessed [here](https://github.com/numbats/summer-wages-refresh/tree/main/app).
 

### 2021-04-07 report_alt.Rmd
Add abstract.

@@
 ```
+
+The NSLY79 is a prominent open data source that has been playing an important role in multidisciplinary research.
+This article aims to demonstrate the steps of initial data analysis (IDA), namely tidying, cleaning, and documenting the process to make the data available for text book or research data.
+Besides, this article provides the opportunity to continuously refresh the text book data by updating the codes that are provided in this article.
+The IDA shows that the NLSY79 data has two problems.
+Firstly, the downloaded data format is untidy.
+Secondly, some anomalies in terms of extreme high hourly wages are found.
+We addressed the first problem by tidying the data according to three principles of a tidy data.
+For the latter problem, we found the anomalies using the robustness weight threshold of robust linear regression.
+An observation with weight < 1.12 is considered as an outlier.
+Further, we fixed the anomalies using the predicted value of the robust linear regression model.
+Using the cleaned data, we provided three data sets, the high school mean hourly wage data, the high school dropouts mean hourly wage data, and demographic variables of the NLSY79 cohort.
+These data sets are contained in an open source R package `yowie`.
+Supplementary materials for this article are available online.
+
+**Keywords**: Data cleaning; Data tidying; Longitudinal data; NLSY79; Open data; Initial data analysis; Outlier detection; Robust linear regression
 

### 2021-04-07 report_alt.Rmd
Change the exploratory data analysis

@@
 
-In this part, we gave some examples of how the data might be used for data analysis in text book.
+In this part, we gave some examples of how the data might be used for data analysis in text book. We used `brolgar` [@brolgar] to perform exploratory data analysis. In first example,
+
+```{r wages-tsibble, echo = FALSE}
+# create the wages data as tsibble
+wages_hs2020_tsibble <- wages_hs2020 %>%
+  as_tsibble(key = id,
+             index = year,
+             regular = FALSE)
+```
+
+```{r gender-plot, echo = FALSE}
+# sample the data by gender
+
+set.seed(31251587)
+wages_hs2020_male <- wages_hs2020_tsibble %>%
+  filter(gender == "MALE") %>%
+  sample_n_keys(size = 120)
+
+set.seed(31251587)
+wages_hs2020_female <- wages_hs2020_tsibble %>%
+  filter(gender == "FEMALE") %>%
+  sample_n_keys(size = 120)
+
+wages_hs2020_sample_gender <- bind_rows(wages_hs2020_male, wages_hs2020_female)
+
+ggplot(wages_hs2020_sample_gender, 
+       aes(x = year,
+           y = mean_hourly_wage,
+           group = id,
+           color = gender)) + 
+  geom_line(alpha = 0.4) +
+  facet_strata(along = gender)
+```
+
+```{r race-plot, echo = FALSE}
+
+set.seed(31251587)
+wages_hs2020_white <- wages_hs2020_tsibble %>%
+  filter(race == "NON-BLACK, NON-HISPANIC") %>%
+  sample_n_keys(size = 60)
+
+set.seed(31251587)
+wages_hs2020_black <- wages_hs2020_tsibble %>%
+  filter(race == "BLACK") %>%
+  sample_n_keys(size = 60)
+
+set.seed(31251587)
+wages_hs2020_hisp <- wages_hs2020_tsibble %>%
+  filter(race == "HISPANIC") %>%
+  sample_n_keys(size = 60)
+
+wages_hs2020_sample_race <- bind_rows(wages_hs2020_white, wages_hs2020_black, wages_hs2020_hisp)
+
+ggplot(wages_hs2020_sample_race, 
+       aes(x = year,
+           y = mean_hourly_wage,
+           group = id,
+           color = race)) + 
+  geom_line(alpha = 0.4) +
+  facet_strata(along = race) +
+  theme(legend.position = "bottom")
+```
+
+```{r wages-slope, echo = FALSE}
+wages_slope <- key_slope(wages_hs2020_tsibble, 
+                         mean_hourly_wage ~ year) %>%
+  left_join(demographic_nlsy79, by = "id")
+
+ggplot(wages_slope, aes(y = hgc,
+                        x = .slope_year)) +
+  geom_boxplot()
+```
+
 The first example is that we observed relationship between wages and education.
@@
 Regarding the gender, male earn more than female, and unfortunately, the gap keep increasing over time.
-
-```{r, echo = FALSE}
-wages_hs2020_tsibble <- wages_hs2020 %>%
-  as_tsibble(key = id,
-             index = year,
-             regular = FALSE)
-
-wages_hs2020_feature <- wages_hs2020_tsibble %>%
-  features(mean_hourly_wage,
-           feat_three_num)
-
-wages_median_demog <- left_join(wages_hs2020_feature, demographic_nlsy79, by = "id") %>%
-  mutate(hgc = factor(hgc, levels = c("1ST GRADE", "2ND GRADE", "3RD GRADE", 
-                                      "4TH GRADE", "5TH GRADE", "6TH GRADE", 
-                                      "7TH GRADE", "8TH GRADE", "9TH GRADE",
-                                      "10TH GRADE", "11TH GRADE", "12TH GRADE", 
-                                      "UNGRADED")))
-```
-
-```{r eda-plot, echo = FALSE, fig.height=6, fig.width=6, fig.cap="The example of EDA plot.
-Plot A shows the pattern of the mean hourly wage median against the highest grade completed.
-It shows that the higher eductaion completed does not necessarily means the more wages earned.
-Plot B and C show the mean hourly wage median over the years based on the race and gender respectively.
-Generally, the respondents experienced a positive trend over time despite of what their race and gender."}
-hgc_plot <- ggplot(filter(wages_median_demog, hgc != "UNGRADED"), aes(y = hgc, x = med)) +
-  geom_boxplot() +
-  theme(axis.text.y = element_text(size = 6),
-        axis.title = element_text(size = 9)) +
-  ggtitle("A)") +
-  xlab("median of mean hourly wage (US$)") +
-  ylab("highest grade completed")
-
-wages_hs2020_grouped <- wages_hs2020 %>%
-  group_by(race, year) %>%
-  summarise(median = median(mean_hourly_wage))
-
-race_plot <- ggplot(wages_hs2020_grouped) +
-  geom_line(aes(x = year,
-                y = median,
-                group = race,
-                colour = race)) +
-  theme(legend.position = "bottom",
-        legend.text = element_text(size = 5),
-        axis.title = element_text(size = 9)) +
-  scale_color_brewer(palette = "Accent") +
-  ggtitle("B)") +
-  ylab("median of mean hourly wage (US$)")
-
-wages_hs2020_grouped_gender <- wages_hs2020 %>%
-  group_by(gender, year) %>%
-  summarise(median = median(mean_hourly_wage))
-
-gender_plot <- ggplot(wages_hs2020_grouped_gender) +
-  geom_line(aes(x = year,
-                y = median,
-                group = gender,
-                colour = gender)) +
-  theme(legend.position = "bottom",
-        legend.text = element_text(size = 5),
-        axis.title = element_text(size = 9)) +
-  scale_color_brewer(palette = "Dark2") +
-  ggtitle("C)") +
-  ylab("median of mean hourly wage (US$)")
-
-hgc_plot / (race_plot + gender_plot)
-
-#### using brolgar to explore the individual experiences <- change the EDA
-```
 

### 2021-04-08 report_alt.Rmd
Add explanation to wages slope plot.

@@
 
-In this part, we gave some examples of how the data might be used for data analysis in text book. We used `brolgar` [@brolgar] to perform exploratory data analysis. In first example,
+In this part, we gave some examples of how the data might be used for data analysis in text book.
+We used `brolgar` [@brolgar] to perform exploratory data analysis.
+The first example is that we observed relationship between wages and education in terms of how the wages increase year by year.
+Using `key_slope` function in `brolgar`, we performed a liner regression with  mean hourly wage as response variable and year as predictor and extracted the regression slope of each ID and plotted it according to highest grade completed.
+
+We found in Figure \ref{fig:wages-slope} that the higher the education completed does not necessarily increase the wages more from year to year.
+We can see that people who completed 4th grade happen to have relatively the same median as people who completed 11th grade, even though it is probably because some respondents in this group had an extremely high mean hourly wages.
+Moreover, the annual increasing wages of people who completed 6th grade to 11th grade are relatively same.
+People who completed 12th grade tend to have higher slope of year compared to the other groups.
+We can also see that some respondents have a negative slope, meaning that their wages tend to decrease over the year.
 
@@
              regular = FALSE)
+```
+
+```{r wages-slope, echo = FALSE, fig.cap="Regression slope of mean hourly wage regression by highest grade completed.
+The regression model has mean hourly wage as response variable and year as predictor.
+Respondents who completed 12th grade has the highest median of slope."}
+
+wages_slope <- key_slope(wages_hs2020_tsibble, 
+            mean_hourly_wage ~ year) %>%
+  left_join(demographic_nlsy79, by = "id") %>%
+  mutate(hgc = factor(hgc, levels = c("1ST GRADE", "2ND GRADE", "3RD GRADE", 
+                                      "4TH GRADE", "5TH GRADE", "6TH GRADE", 
+                                      "7TH GRADE", "8TH GRADE", "9TH GRADE",
+                                      "10TH GRADE", "11TH GRADE", "12TH GRADE", 
+                                      "UNGRADED"))) %>%
+  filter(hgc != "UNGRADED")
+
+ggplot(wages_slope, aes(y = hgc,
+                      x = .slope_year)) +
+  geom_boxplot()
 ```
@@
 
-```{r wages-slope, echo = FALSE}
-wages_slope <- key_slope(wages_hs2020_tsibble, 
-                         mean_hourly_wage ~ year) %>%
-  left_join(demographic_nlsy79, by = "id")
-
-ggplot(wages_slope, aes(y = hgc,
-                        x = .slope_year)) +
-  geom_boxplot()
-```
-
-The first example is that we observed relationship between wages and education.
-Using `features` function in `brolgar`, we extracted the median of the mean hourly age of each ID and plotted according to highest grade completed.
-We found in Figure \ref{fig:eda-plot} A that the higher the education completed does not necessarily means the higher wages earned.
-However, people who completed 12th grade tend to have higher wage compared to the other groups.
-Further, respondents who only completed 4th grade happen to have median about as high as the respondents who completed 10th grade.
-
-Secondly, we examined the pattern of mean hourly wage corresponding to different groups of race and gender.
-Figure \ref{fig:eda-plot} B and C suggest that in general, the mean hourly wage median increased over the years.
-Regarding the race, we found that Non-Black Non-Hispanic group had a higher wages than the other groups.
-We can also see a reasonable gap between the wages median of Black people and the Hispanic and the Non-Black Non-Hispanic.
-Regarding the gender, male earn more than female, and unfortunately, the gap keep increasing over time.
-
 # Summary
@@
 
-This paper also has demonstrated the documentation of data cleaning.
-We also provided all of the codes in performing data tidying and initial data analysis.
+This paper also has demonstrated the documentation of data cleaning by providing all of the codes in performing data tidying and initial data analysis.
 Thus, this paper provided an opportunity to continuously refresh the text book data whenever the updated data is published in the NLSY79 database.

### 2021-04-13 report_alt.Rmd
Add GAM model to EDA section.

@@
 library(webshot)
+library(mgcv)
+library(modelr)
+library(tsibble)
 ```
@@
 We found in Figure \ref{fig:wages-slope} that the higher the education completed does not necessarily increase the wages more from year to year.
-We can see that people who completed 4th grade happen to have relatively the same median as people who completed 11th grade, even though it is probably because some respondents in this group had an extremely high mean hourly wages.
+We can see that people who completed 4th grade happen to have relatively the same median of slope as people who completed 11th grade, even though it is probably because some respondents in this group had an extremely high mean hourly wages.
 Moreover, the annual increasing wages of people who completed 6th grade to 11th grade are relatively same.
 People who completed 12th grade tend to have higher slope of year compared to the other groups.
-We can also see that some respondents have a negative slope, meaning that their wages tend to decrease over the year.
+Some outliers are also spotted in this group.
+Additionally, we can see that some respondents have a negative slope, meaning that their wages tend to decrease over the year.
 
@@
 
-```{r gender-plot, echo = FALSE}
+In the second example, we examined the pattern of mean hourly wage corresponding to different groups of gender and race.
+We took 240 samples of female and male equally and spread it into twelve facets as is seen in Figure \ref{fig:gender-plot} using `facet_strata` function from `brolgar` [@brolgar].
+We learn that the mean hourly wage of female and male fluctuated over the years.
+In some facets, for example in facet 4, 5, and 10, males tend to earn more wages than females.
+This finding is also shown in Figure \ref{fig:gender-plot2} where we plotted the median wage of males and females against the year.
+We learn that males had a higher mean of mean hourly wage over the period of survey than females.
+Unfortunately, the gap between them tend to be wider from time to time.
+
+```{r gender-plot, echo = FALSE, warning = FALSE, fig.cap= "The pattern of mean hourly wages overtime displayed in 12 facets coloured by gender.
+The respondents belong to high school cohort are randomly sampled.
+We learn that both male and female respondents had a fluctuate mean hourly wage over the year." }
 # sample the data by gender
@@
            color = gender)) + 
-  geom_line(alpha = 0.4) +
-  facet_strata(along = gender) 
-```
-
-```{r race-plot, echo = FALSE}
-
-set.seed(31251587)
-wages_hs2020_white <- wages_hs2020_tsibble %>%
-  filter(race == "NON-BLACK, NON-HISPANIC") %>%
-  sample_n_keys(size = 60)
-
-set.seed(31251587)
-wages_hs2020_black <- wages_hs2020_tsibble %>%
-  filter(race == "BLACK") %>%
-  sample_n_keys(size = 60)
-
-set.seed(31251587)
-wages_hs2020_hisp <- wages_hs2020_tsibble %>%
-  filter(race == "HISPANIC") %>%
-  sample_n_keys(size = 60)
-
-wages_hs2020_sample_race <- bind_rows(wages_hs2020_white, wages_hs2020_black, wages_hs2020_hisp)
-
-ggplot(wages_hs2020_sample_race, 
+  geom_line(alpha = 0.7) +
+  scale_color_brewer(palette = "Dark2") +
+  facet_strata(along = ~gender) 
+```
+
+```{r gender-plot2, echo=FALSE, fig.cap="The yearly mean of mean hourly wage by gender along with its standard error.
+Males generally earn more wages than females.
+The error interval tend to be wider over the years."}
+wages_hs2020_grouped_gender <- wages_hs2020 %>%
+  group_by(gender, year) %>%
+  summarise(mean = mean(mean_hourly_wage),
+             sd = sd(mean_hourly_wage),
+             samplesize = n()) %>%
+  ungroup() %>%
+  mutate(upper = mean + sd/sqrt(samplesize),
+         lower = mean - sd/sqrt(samplesize))
+
+gender_plot <- ggplot(wages_hs2020_grouped_gender,
                       aes(x = year,
-           y = mean_hourly_wage,
-           group = id,
-           color = race)) + 
-  geom_line(alpha = 0.4) +
-  facet_strata(along = race) +
-  theme(legend.position = "bottom")
+                          y = mean,
+                          group = gender,
+                          colour = gender)) +
+  geom_point(size = 1) +
+  geom_errorbar(aes(ymin=lower, ymax=upper), width=1,
+                 position=position_dodge(0.1)) +
+  theme(legend.position = "bottom",
+         legend.text = element_text(size = 5),
+         axis.title = element_text(size = 9)) +
+  scale_color_brewer(palette = "Dark2") +
+  ylab("median of mean hourly wage (US$)")
+
+gender_plot
+
+```
+
+For the final example, we show the example of modeling the longitudinal data using Generelized Additive Model (GAM) for each ID.
+We fit a more flexible model to fit in to the data since the relationship between mean hourly wage and year is not linear due to the spikes.
+Since this approach is computationally expensive, we only show the model in a little fraction of the data.
+We sample 1 percent of the total respondents using `brolgar's` `sample_frac_keys` function.
+
+```{r gam, cache=TRUE, echo=FALSE, fig.cap="Exploration of wages data by fitting a GAM.
+The fitted model displayed by blue line.
+The fitted line shows that the model flexible enough to follow the pattern of the data."}
+
+# take a small fraction of key
+set.seed(2020)
+wages_sample_gam <- sample_frac_keys(wages_hs2020_tsibble, size = 0.01) %>%
+  mutate(id_fct = as.factor(id),
+         year0 = year - 1979)
+
+# build GAM model
+wages_gam <- gam(
+  mean_hourly_wage ~ s(year0, by = id_fct) + id_fct,
+  data = wages_sample_gam,
+  method = "REML"
+)
+
+wages_aug_gam <- wages_sample_gam %>%
+  add_predictions(wages_gam, var = "pred") %>%
+  add_residuals(wages_gam, var = "res") %>%
+  group_by_key() %>%
+  mutate(rss = sum(res^2)) %>%
+  ungroup()
+
+set.seed(2021)
+wages_aug_gam %>%
+  sample_n_keys(12) %>%
+  ggplot(aes(x = year,
+             y = pred,
+             group = id)) +
+  geom_line(color = "steelblue") +
+  geom_point(aes(y = mean_hourly_wage)) +
+  facet_wrap(~id)
+
 ```
@@
 
-The entire analysis is conducted using `R` [@R] in `rstudio` using these packages: `tidyverse` [@tidyverse], `ggplot2` [@ggplot2], `dplyr` [@dplyr], `readr` [@readr], `tidyr` [@tidyr], `stringr` [@stringr], `purrr` [@purrr], `broom` [@broom], `blorgar` [@brolgar], `patchwork` [@patchwork], `kableExtra` [@kableExtra], `MASS` [@mass], `janitor` [@janitor], `DiagrammeR` [@DiagrammeR], `rsvg` [@rsvg], and `webshot` [@webshot].
+The entire analysis is conducted using `R` [@R] in `rstudio` using these packages: `tidyverse` [@tidyverse], `ggplot2` [@ggplot2], `dplyr` [@dplyr], `readr` [@readr], `tidyr` [@tidyr], `stringr` [@stringr], `purrr` [@purrr], `broom` [@broom], `blorgar` [@brolgar], `patchwork` [@patchwork], `kableExtra` [@kableExtra], `MASS` [@mass], `janitor` [@janitor], `DiagrammeR` [@DiagrammeR], `rsvg` [@rsvg], `webshot` [@webshot], `mgcv` [@mgcv], `tsibble` [@tsibble], and `modelr` [@modelr].
 The paper are generated using `knitr` [@knitr], `rmarkdown` [@rmarkdown], and `rticles` [@rticles].

### 2021-04-13 report_alt.Rmd
Add explanation to the GAM model plot.

@@
 
-For the final example, we show the example of modeling the longitudinal data using Generelized Additive Model (GAM) for each ID.
+For the final example, we show the example of modeling the longitudinal data using Generelized Additive Model (GAM) with `mgcv` package [@mgcv] for each ID.
 We fit a more flexible model to fit in to the data since the relationship between mean hourly wage and year is not linear due to the spikes.
@@
 We sample 1 percent of the total respondents using `brolgar's` `sample_frac_keys` function.
+Further, we plotted the some sample of the fitted model in Figure \@ref{fig:gam}.
+We learn that the model flexible enough to deal with the fluctuation in the data.
+Moreover, model with small spikes tend to be linearly fitted.
 

### 2021-04-21 report_alt.Rmd
abstract

@@
   - name: Emi Tanaka
-    thanks: emitanaka.org
     department: Department of Econometrics and Business Statistics
@@
   - name: Dianne Cook
-    thanks: dicook.org
     department: Department of Econometrics and Business Statistics
@@
 
-The NSLY79 is a prominent open data source that has been playing an important role in multidisciplinary research.
-This article aims to demonstrate the steps of initial data analysis (IDA), namely tidying, cleaning, and documenting the process to make the data available for text book or research data.
-Besides, this article provides the opportunity to continuously refresh the text book data by updating the codes that are provided in this article.
-The IDA shows that the NLSY79 data has two problems.
-Firstly, the downloaded data format is untidy.
-Secondly, some anomalies in terms of extreme high hourly wages are found.
-We addressed the first problem by tidying the data according to three principles of a tidy data.
-For the latter problem, we found the anomalies using the robustness weight threshold of robust linear regression.
-An observation with weight < 1.12 is considered as an outlier.
-Further, we fixed the anomalies using the predicted value of the robust linear regression model.
-Using the cleaned data, we provided three data sets, the high school mean hourly wage data, the high school dropouts mean hourly wage data, and demographic variables of the NLSY79 cohort.
-These data sets are contained in an open source R package `yowie`.
-Supplementary materials for this article are available online.
+NSLY79 is a prominent open data source that has been an important support in multidisciplinary research on longitudinal data.
+Subsets of this data can be found in numerous textbooks and research articles, however the steps and decisions taken to get from the original data to these subsets is never clearly articulated.
+This article describes our journey when trying to re-create a textbook example data set from the original data, with a goal to refresh it for use in the classroom.
+It thus demonstrates the process of initial data analysis (IDA) -- tidying, cleaning, and documenting the process -- to make data available for text book examples or research.
+Three new data sets, and the code to produce them are provided in an accompanying open source R package, called `yowie`.
+The code might be useful for future refreshing of this particular text book data.
+As a result of this process, some recommendations are made for the NLSY79 curators for incorporating data quality checks, and providing more convenient samples of the data to potential users.
 
@@
 
-**Codes** : R script to reproduce data tidying and cleaning are available in this [Github Repository](https://github.com/numbats/yowie/blob/master/data-raw/data_preprocessing.R)
-
-**R Package `yowie`**:`yowie` is a data container R package that contains 3 datasets, namely the high school mean hourly wage data, high school dropouts mean hourly wage data, and demographic data of the NLSY79 cohort.
+- **Codes** : R script to reproduce data tidying and cleaning are available in this [Github Repository](https://github.com/numbats/yowie/blob/master/data-raw/data_preprocessing.R).
+
+- **R Package `yowie`**:`yowie` is a data container R package that contains 3 datasets, namely the high school mean hourly wage data, high school dropouts mean hourly wage data, and demographic data of the NLSY79 cohort.
 This package could be accessed [here](https://github.com/numbats/yowie).
 
-**shiny app**: An web interactive `shiny` app to run a simulation to customize the weight threshold. This app could be accessed [here](https://github.com/numbats/summer-wages-refresh/tree/main/app).
+- **shiny app**: An web interactive `shiny` app to run a simulation to customize the weight threshold. This app could be accessed [here](https://github.com/numbats/summer-wages-refresh/tree/main/app).
 

### 2021-05-05 report_alt.Rmd
section titles and some reference fixes
Add GAM and gender with SE plots.

@@
 
-### Robust Linear Model for Noises Treatment
+### Replacing extreme values
 
@@
 In some facets, for example in facet 4, 5, and 10, males tend to earn more wages than females.
-This finding is also shown in Figure \ref{fig:gender-plot2} where we plotted the median wage of males and females against the year.
+This finding is also shown in Figure \ref{fig:gender-se} where we plotted the median wage of males and females against the year.
 We learn that males had a higher mean of mean hourly wage over the period of survey than females.
@@
 
-```{r gender-plot2, echo=FALSE, fig.cap="The yearly mean of mean hourly wage by gender along with its standard error.
+```{r gender-se, echo=FALSE, fig.cap="The yearly mean of mean hourly wage by gender along with its standard error.
 Males generally earn more wages than females.
@@
 We sample 1 percent of the total respondents using `brolgar's` `sample_frac_keys` function.
-Further, we plotted the some sample of the fitted model in Figure \@ref{fig:gam}.
+Further, we plotted the some sample of the fitted model in Figure \@ref{fig:gam-plot}.
 We learn that the model flexible enough to deal with the fluctuation in the data.
@@
 
-```{r gam, cache=TRUE, echo=FALSE, fig.cap="Exploration of wages data by fitting a GAM.
+```{r gam-plot, cache=TRUE, echo=FALSE, fig.cap="Exploration of wages data by fitting a GAM.
 The fitted model displayed by blue line.

### 2021-05-05 report_alt.Rmd
create two subsections in The NLSY79 section i.e Database and Target Data.

@@
 
+## Database
+
 The NLSY79 is a longitudinal survey held by the U.S Bureau of Labor Statistics that follows the lives of a sample of American youth and born between 1957-1964 [@nlsy79].
@@
 In recent survey years, more than 90 percent of respondents were interviewed by telephone [@eliznlsy].
+
+## Target Data
+
+As mentioned in Section 1, this paper aims to refresh the NLSY79 data as used in Singer and Willet's text book (@SingerJudithD2003Alda).
+This data set contains the information of yearly mean hourly wages from the NLSY79 cohort, with education and race as covariates.
+This data set represents the measurement of male high-school dropouts, aged from 14 to 17 years old when first time measured.
+Hence, in this paper, after getting the data set of high school graduates, we subset it by only having the high school dropouts.
+
+Since there is no information of high school drop-outs in Singer and Willet's book, we define high school dropouts as respondents who only completed either $9^{th}$, $10^{th}$, $11^{th}$ grade, or who completed $12^{th}$ when their age are at least 19 years old (means that these IDs being dropped out in certain year, but they returned back to high school to complete it).
 
@@
 Finally, we saved the imputed data and set the appropriate data type for the variables.
-We also saved the NLSY79 cohort's demographic information and the high school dropout cohort in a separate data sets.
-Here, we defined high school dropouts as respondents who only completed either $9^{th}$, $10^{th}$, $11^{th}$ grade, or who completed $12^{th}$ when their age are at least 19 years old (means that these IDs being dropped out in certain year, but they returned back to high school to complete it).
-We also filtered the data to be only male and aged between 14 and 17 years old to refresh the high school dropouts data in the `brolgar` package.
-
+As our target data is the mean hourly wage of the high-school dropouts, we then subset the high-school graduate data set to be only having the male-high school dropout data.
+We also saved the NLSY79 cohort's demographic information in a separate dataset.
 We then make these three data sets and its processing documentation to be publicly available through an R data container package called `yowie`.

### 2021-05-05 report_alt.Rmd
Set some code chunks to echo = FALSE

@@
 output: 
-  rticles::arxiv_article:
-    keep_tex: false
+  #rticles::arxiv_article:
+    #keep_tex: false
+  bookdown::pdf_book:
+    base_format: rticles::arxiv_article
 date: "24/02/2021"
@@
 
-```{r, cache = TRUE}
+```{r, cache = TRUE, echo = FALSE}
 source(here::here("data-raw/NLSY79/NLSY79.R"))
@@
 
-The `highest grade completed` came with several version in each year.
-We chose the revised May data because the May data seemed to have less missing and presumably the revised data has been checked.
-However, there was no revised May data for 2012, 2014, 2016, and 2018 so we just used the ordinary May data.
-
-```{r tidy-grade}
+The next step is tidying the `highest grade completed` variable.
+This variable came with several version in each year.
+We chose the revised May data because it seemed to have less missing and presumably has been checked.
+However, there was no revised May data for 2012, 2014, 2016, and 2018 so we just used the ordinary May data for these years.
+The code for tidying this variable is provided in the supplementary code whose details could be seen in the Supplementary Materials section of this paper.
+
+```{r tidy-grade, echo = FALSE}
 # tidy the grade 
@@
 
-In the final data, we only used the highest grade completed ever and derived the year of when the highest grade completed its categorical value.
-Therefore, we wrangled the highest grade completed in each year to mutate these variables.
-
-```{r tidy-hgc}
+Further, `highest grade completed` is measured and could be updated in each period of the survey.
+Hence, we only used the highest grade completed ever and derived the year when it is completed.
+The code for tidying this variable is also available in the supplementary code of this paper.
+
+```{r tidy-hgc, echo = FALSE}
 
@@
 
-Finally, we join all the tidy variables in a data set called `full_demographics`.
-
-```{r full-demog}
+Finally, we join all of the tidy variables in a data set called `full_demographics` as displayed as follows:
+
+```{r full-demog, echo = FALSE}
 full_demographics <- full_join(dob_tidy, demog_tidy, by = "CASEID_1979") %>%
@@
 
-The employment data comprises of three variables, i.e.
-`total hours of work per week`, `number of jobs that an individual has`, and `mean hourly wage`.
-For hours worked per week, initially only one version per job, no choice from 1979 to 1987 (QES-52A).
+The employment data comprises of three variables, i.e. `total hours of work per week`, `number of jobs that an individual has`, and `mean hourly wage`.
+
+`hours worked per week` initially has only one version per job, no choice from 1979 to 1987 (QES-52A).
 From 1988 onward, when we had more options, we chose the variable for total hours including time spent working from home (QES-52D).
-However, 1993 did not have all the five D variables (the first one and the last one were missing), so we used QES-52A variable instead.
+However, in 1993, this variable did not have all the five D variables (the first one and the last one were missing), so we used QES-52A variable instead.
 In addition, 2008 only had jobs 1-4 for the QES-52D variable (whereas the other years had 1-5), so we just used these.
@@
 
-The same algorithm was also deployed to tidy the rate of wage by year and by ID.
+The same way was also deployed to tidy the rate of wage by year and by ID.
 The difference is that the hourly rate had only one version of each year.
-The hours of work and the hourly rate were then joined to calculate the number of jobs that an ID has and their mean hourly wage.
+The hours of work and the hourly rate were then joined to calculate the number of jobs that a respondent has and their mean hourly wage.
 Some observations had 0 in their hourly rate, which is considered as invalid value.
 Thus, their hourly rate set to be N.A.
-
-```{r tidy-rate}
+the result of the joining variables could be seen in the following tibble.
+
+```{r tidy-rate, echo = FALSE}
 get_rate <- function(year) {
@@
 Since our ultimate goal is to calculate the mean hourly wage, the number of jobs is calculate based on the availability of the `rate_per_hour` information.
-For example, the number of jobs of ID 1, based on `hours_work`, is two.
+For example, the number of jobs of ID 1, based on `hours_work`, is 2.
 However, since the information of hourly rate of `job_02` is not available, the number of job is considered as 1.
@@
 Additionally, if an ID only had one job, we directly used their hourly wages information and flagged it as an arithmetic mean.
-
-```{r}
+The following tibble shows the result of this tidying process.
+
+```{r tidy-nojob, echo = FALSE}
 # calculate number of jobs that a person has in one year
@@
 
-```{r}
+```{r wages-demog-hs, echo = FALSE}
 # join the wages information and the demographic information by case id.
@@
 The next step is that we explored the mean hourly wage data, in this case, we only explored the wages data in `wages_demog_hs`.
-Figure \ref{fig:featureplot} conveys that there is clearly a problem in the mean hourly wage values.
+Figure \@ref(fig:featureplot) conveys that there is clearly a problem in the mean hourly wage values.
 Figure \ref{fig:featureplot} A shows that some observations had an exceptionally high figure of wages, even more than US$10,000 per hour.
@@
 
-```{r high-wages, echo = FALSE, fig.cap= "Some observations with extremely high mean hourly wage. Most of the IDs only have one point of high wage.", fig.height=8, fig.width=6}
+```{r high-wages, echo = FALSE, fig.cap= "6 out of 45 IDs with extremely high mean hourly wage. Most of the IDs only have one point of high wage."}
 
 wages_high <- filter(wages_demog_hs, mean_hourly_wage > 500) %>%
-  as_tibble()
+  as_tibble() %>%
+  head(n = 6)
 
@@
              alpha = 0.5) +
-  facet_wrap(~id, scales = "free_y", ncol = 5) +
+  facet_wrap(~id, scales = "free_y") +
   theme(axis.text.x = element_text(angle = 10, size = 6)) +
@@
 
-```{r rlm, cache=TRUE, warning = FALSE, message = FALSE}
+```{r rlm, cache=TRUE, warning = FALSE, message = FALSE, echo = FALSE}
 
@@
 
-```{r}
+```{r save-data, echo = FALSE}
 # select out the old value of mean hourly wage and change it with the wages_rlm value

### 2021-05-05 report_alt.tex
Set some code chunks to echo = FALSE

\documentclass{article}

\usepackage{arxiv}

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{lmodern}        % https://github.com/rstudio/rticles/issues/343
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{lipsum}
\usepackage{graphicx}

\title{The Journey from Wild to Text Book Data: A Case Study of the National Longitudinal Survey of Youth}

\author{
    Dewi Amaliah
   \\
    Department of Econometrics and Business Statistics \\
    Monash University \\
  Clayton, VIC 3800 \\
  \texttt{} \\
   \And
    Kate Hyde
   \\
     \\
   \\
  \texttt{} \\
   \And
    Emi Tanaka
   \\
    Department of Econometrics and Business Statistics \\
    Monash University \\
  Clayton, VIC 3800 \\
  \texttt{\href{mailto:emi.tanaka@monash.edu}{\nolinkurl{emi.tanaka@monash.edu}}} \\
   \And
    Nicholas Tierney
   \\
     \\
   \\
  \texttt{} \\
   \And
    Dianne Cook
   \\
    Department of Econometrics and Business Statistics \\
    Monash University \\
  Clayton, VIC 3800 \\
  \texttt{\href{mailto:dicook@monash.edu}{\nolinkurl{dicook@monash.edu}}} \\
  }

\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}

% Pandoc citation processing

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}

\begin{document}
\maketitle

\def\tightlist{}

\begin{abstract}

\end{abstract}

NSLY79 is a prominent open data source that has been an important support in multidisciplinary research on longitudinal data. Subsets of this data can be found in numerous textbooks and research articles, however the steps and decisions taken to get from the original data to these subsets is never clearly articulated. This article describes our journey when trying to re-create a textbook example data set from the original data, with a goal to refresh it for use in the classroom. It thus demonstrates the process of initial data analysis (IDA) -- tidying, cleaning, and documenting the process -- to make data available for text book examples or research. Three new data sets, and the code to produce them are provided in an accompanying open source R package, called \texttt{yowie}. The code might be useful for future refreshing of this particular text book data. As a result of this process, some recommendations are made for the NLSY79 curators for incorporating data quality checks, and providing more convenient samples of the data to potential users.

\textbf{Keywords}: Data cleaning; Data tidying; Longitudinal data; NLSY79; Open data; Initial data analysis; Outlier detection; Robust linear regression

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

``Open data'' is data that are freely accessible, modifiable, and shareable by anyone for any purpose (Open Knowledge Foundation 2021). This type of data can be useful as example data in statistical text books and for research purposes. However, open data are often, what we might call ``wild data'', because it requires substantial cleaning and tidying to tame it into text book shape. Huebner, Vach, and Cessie (2016) emphasize that making the data cleaning process accountable and transparent is imperative. Documenting the data cleaning is essential for the integrity of downstream statistical analyses and model building (Huebner et al. 2020).

Data cleaning is part of what is called ``initial data analysis (IDA)'' (Chatfield 1985). The other IDA steps are to explore the data properties and check assumption of modeling to document and report the process for the later formal analysis. Hence, exploratory data analysis, which is defined by Tukey (1977) as a detective work to get the clue from data, either numerically or graphically, before confirmatory data analysis is performed, encompasses IDA. Dasu and Johnson (2003) say that data cleaning and exploration, without naming it as IDA, is a difficult task and consumes 80\% of the data mining task.

Despite its importance, this IDA stage is often undervalued and neglected (Chatfield 1985). There are few research papers that document the data cleaning (Wickham 2014). Furthermore, the decisions made in this stage often go unreported in the sense that IDA is often performed in an unplanned and unstructured way, and is only shared among restricted parties (Huebner et al. 2020).

This paper aims to demonstrate the steps of IDA, tidying, cleaning, and documenting the process, for a prominent open data source, National Longitudinal Survey of Youth (NLSY) 1979, henceforth referred to NSLY79. This data has been playing an important role in research in various disciplines including but are not limited to economics, sociology, education, public policy, and public health for more than a quarter of the century (Pergamit et al. 2001). In addition, The National Longitudinal Survey is considered as survey with high retention rates and carefully designed making it suitable for a life course research (Pergamit et al. 2001) and (Cooksey 2017). According to Cooksey (2017), thousand of articles, and hundreds of book chapters and monographs has incorporated the NLSY data. Moreover, the NLSY79 is considered as the most widely used and most important cohort in the NLSY79 data sets (Pergamit et al. 2001).

Singer and Willett (2003) used the wages and other variables of high school dropouts from the NLSY79 data as an example data set to illustrate longitudinal data modeling. Our aim is to refresh this text book data to be more up to date. Here, we investigate the process of getting from the raw NLSY79 to this text book data set. However, we are not able to create the exact same data set as published in their book since we do not have the information of what age threshold did they use to determine the high school dropouts.

This paper is structured to have 5 sections. Section 2 describes the original data source. Section 3 presents the steps of cleaning the data, including getting and tidying the data from the NLSY79 and initial data analysis to find and treat the anomalies in the data. We also gave examples of exploratory data analysis using the clean data in Section 4. Finally, Section 6 summarizes the paper.

\hypertarget{the-nlsy79}{%
\section{The NLSY79}\label{the-nlsy79}}

\hypertarget{database}{%
\subsection{Database}\label{database}}

The NLSY79 is a longitudinal survey held by the U.S Bureau of Labor Statistics that follows the lives of a sample of American youth and born between 1957-1964 (The U.S. Bureau of Labor Statistics, n.d.). The cohort originally included 12,686 respondents ages 14-22 when first interviewed in 1979. It was comprised of Blacks, Hispanics, economically disadvantaged non-Black non-Hispanics, and youth in the military. In 1984 and 1990, two sub-samples were dropped from the interview; the dropped subjects are the 1,079 members of the military sample and 1,643 members of the economically disadvantaged non-Black non-Hispanics respectively. Hence, 9,964 respondents remain in the eligible samples. The surveys were conducted annually from 1979 to 1994 and biennially thereafter. Data are now available from Round 1 (1979 survey year) to Round 28 (2018 survey year).

Although the main focus area of the NLSY is labor and employment, the NLSY also cover several other topics including education; training and achievement; household, geography and contextual variables; dating, marriage, cohabitation; sexual activity, pregnancy, and fertility; children; income, assets and program participation; health; attitudes and expectations; and crime and substance use.

There are two ways to conduct the interview of the NLSY79, which are face to face interview or by telephone. In recent survey years, more than 90 percent of respondents were interviewed by telephone (Cooksey 2017).

\hypertarget{target-data}{%
\subsection{Target Data}\label{target-data}}

As mentioned in Section 1, this paper aims to refresh the NLSY79 data as used in Singer and Willet's text book (Singer and Willett (2003)). This data set contains the information of yearly mean hourly wages from the NLSY79 cohort, with education and race as covariates. This data set represents the measurement of male high-school dropouts, aged from 14 to 17 years old when first time measured. Hence, in this paper, after getting the data set of high school graduates, we subset it by only having the high school dropouts.

Since there is no information of high school drop-outs in Singer and Willet's book, we define high school dropouts as respondents who only completed either \(9^{th}\), \(10^{th}\), \(11^{th}\) grade, or who completed \(12^{th}\) when their age are at least 19 years old (means that these IDs being dropped out in certain year, but they returned back to high school to complete it).

\hypertarget{the-nlsy79-data-cleaning}{%
\section{The NLSY79 Data Cleaning}\label{the-nlsy79-data-cleaning}}

\hypertarget{getting-and-tidying-the-data}{%
\subsection{Getting and Tidying the Data}\label{getting-and-tidying-the-data}}

The NLSY79 data are stored in a \href{https://www.nlsinfo.org/content/cohorts/nlsy79/get-data}{database} and could be downloaded by variables. Several variables are available for download and could be browsed by index. For the wages data set, we only extracted these variables:

\begin{itemize}
\tightlist
\item
  Education, Training \& Achievement Scores

  \begin{itemize}
  \tightlist
  \item
    Education -\textgreater{} Summary measures -\textgreater{} All schools -\textgreater{} By year -\textgreater{} Highest grade completed

    \begin{itemize}
    \tightlist
    \item
      Downloaded all of the 78 variables in Highest grade completed.
    \end{itemize}
  \end{itemize}
\item
  Employment

  \begin{itemize}
  \tightlist
  \item
    Summary measures -\textgreater{} By job -\textgreater{} Hours worked and Hourly wages

    \begin{itemize}
    \tightlist
    \item
      Downloaded all of the 427 variables in Hours worked
    \item
      Downloaded all of the 151 variables in Hourly wages
    \end{itemize}

    Both hours worked and hourly wages are recorded by the job, up to five jobs for each id/subject.
  \end{itemize}
\item
  Household, Geography \& Contextual Variables

  \begin{itemize}
  \tightlist
  \item
    Context -\textgreater{} Summary measures -\textgreater{} Basic demographics

    \begin{itemize}
    \tightlist
    \item
      Downloaded year and month of birth, race, and sex variables.
    \end{itemize}

    There are two versions of the year and month of birth, i.e.~1979 and 1981 data. We downloaded these two versions.
  \end{itemize}
\end{itemize}

The downloaded data set came in a csv (NLSY79.csv) and dat (NLSY79.dat) format. We only used the .dat format. It also came along with these files:

\begin{itemize}
\tightlist
\item
  NLSY79.NLSY79: This is the tagset of variables that can be uploaded to the web site to recreate the data set.
\item
  NLSY79.R: This is an R script provided automatically by the database for reading the data into R and convert the variables' name and its label into something more sensible. We utilized this code to do an initial data tidying. It produced two data set, \texttt{categories\_qnames} (the observations are stored in categorical/interval values) and \texttt{new\_data\_qnames} (the observations are stored in integer form).
\end{itemize}

According to Wickham (2014), a tidy data sets comply with three rules, the first is that each variable forms a column, the second is that each observation forms a row, and the last is that each type of observational unit forms a table. Unfortunately, the \texttt{new\_data\_qnames} did not meet these requirements in the way that the value for particular year and job are stored in different columns, hence the data contains a huge number of columns (686 columns). The example of the untidy of the data set is displayed in Table \ref{tab:untidy-data}, which actually intended to display the hourly rate of each respondent by job (HRP1 to HRP5) and by year (1979 and 1980). The table implies that the column headers are values of the year and job, not variable names.

\begin{table}

\caption{\label{tab:untidy-data}The Untidy Form of the NLSY79 Raw Data}
\centering
\begin{tabular}[t]{r|r|r|r|r|r|r}
\hline
CASEID\_1979 & HRP1\_1979 & HRP2\_1979 & HRP3\_1979 & HRP4\_1979 & HRP5\_1979 & HRP1\_1980\\
\hline
\rowcolor{gray!6}  1 & 328 & NA & NA & NA & NA & NA\\
\hline
2 & 385 & NA & NA & NA & NA & 457\\
\hline
\rowcolor{gray!6}  3 & 365 & NA & 275 & NA & NA & 397\\
\hline
4 & NA & NA & NA & NA & NA & NA\\
\hline
\rowcolor{gray!6}  5 & 310 & 375 & NA & NA & NA & 333\\
\hline
6 & NA & NA & NA & 250 & NA & 275\\
\hline
\end{tabular}
\end{table}

Consequently, the data should be tidied and wrangled first to extract the demographic and employment variables that we want to put in the final data set. We mainly used \texttt{tidyr} (H. Wickham 2020b), \texttt{dplyr} (H. Wickham, François, et al. 2020), and \texttt{stringr} (H. Wickham 2019) to do this job.

\hypertarget{tidying-demographic-variables}{%
\subsubsection{Tidying demographic variables}\label{tidying-demographic-variables}}

\texttt{Age\ in\ 1979}, \texttt{gender}, \texttt{race}, \texttt{highest\ grade\ completed} (factor and integer), and the \texttt{year\ when\ the\ highest\ grade\ completed} are the variables that we want to use in the cleaned data set.

Age in 1979 are derived from year of birth and month of birth variables in the raw data set. The variables have two versions, which are the 1979 version and the 1981 version. We only used the 1979 data and did a consistency check of those years and flag the inconsistent observations.

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{## tidy the date of birth data}
\NormalTok{dob_tidy <-}\StringTok{ }\NormalTok{new_data_qnames }\OperatorTok{%>%}
\StringTok{  }\NormalTok{dplyr}\OperatorTok{::}\KeywordTok{select}\NormalTok{(CASEID_}\DecValTok{1979}\NormalTok{,}
         \KeywordTok{starts_with}\NormalTok{(}\StringTok{"Q1-3_A~"}\NormalTok{)) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{dob_year =} \KeywordTok{case_when}\NormalTok{(}
                    \CommentTok{# if the years recorded in both sets match, take 79 data}
                    \StringTok{`}\DataTypeTok{Q1-3_A~Y_1979}\StringTok{`} \OperatorTok{==}\StringTok{ `}\DataTypeTok{Q1-3_A~Y_1981}\StringTok{`} \OperatorTok{~}\StringTok{ `}\DataTypeTok{Q1-3_A~Y_1979}\StringTok{`}\NormalTok{,}
                    \CommentTok{# if the year in the 81 set is missing, take the 79 data}
                    \KeywordTok{is.na}\NormalTok{(}\StringTok{`}\DataTypeTok{Q1-3_A~Y_1981}\StringTok{`}\NormalTok{) }\OperatorTok{~}\StringTok{ `}\DataTypeTok{Q1-3_A~Y_1979}\StringTok{`}\NormalTok{,}
                    \CommentTok{# if the sets don't match for dob year, take the 79 data}
                    \StringTok{`}\DataTypeTok{Q1-3_A~Y_1979}\StringTok{`} \OperatorTok{!=}\StringTok{ `}\DataTypeTok{Q1-3_A~Y_1981}\StringTok{`} \OperatorTok{~}\StringTok{ `}\DataTypeTok{Q1-3_A~Y_1979}\StringTok{`}\NormalTok{),}
        \DataTypeTok{dob_month =} \KeywordTok{case_when}\NormalTok{(}
                    \CommentTok{# if months recorded in both sets match, take 79 data}
                    \StringTok{`}\DataTypeTok{Q1-3_A~M_1979}\StringTok{`} \OperatorTok{==}\StringTok{ `}\DataTypeTok{Q1-3_A~M_1981}\StringTok{`} \OperatorTok{~}\StringTok{ `}\DataTypeTok{Q1-3_A~M_1979}\StringTok{`}\NormalTok{,}
                    \CommentTok{# if month in 81 set is missing, take the 79 data}
                    \KeywordTok{is.na}\NormalTok{(}\StringTok{`}\DataTypeTok{Q1-3_A~M_1981}\StringTok{`}\NormalTok{) }\OperatorTok{~}\StringTok{ `}\DataTypeTok{Q1-3_A~M_1979}\StringTok{`}\NormalTok{,}
                    \CommentTok{# if sets don't match for dob month, take the 79 data}
                    \StringTok{`}\DataTypeTok{Q1-3_A~M_1979}\StringTok{`} \OperatorTok{!=}\StringTok{ `}\DataTypeTok{Q1-3_A~M_1981}\StringTok{`} \OperatorTok{~}\StringTok{ `}\DataTypeTok{Q1-3_A~M_1979}\StringTok{`}\NormalTok{),}
        \CommentTok{# flag if there is a conflict between dob recorded in 79 and 81}
        \DataTypeTok{dob_conflict =} \KeywordTok{case_when}\NormalTok{(     }
\NormalTok{                      (}\StringTok{`}\DataTypeTok{Q1-3_A~M_1979}\StringTok{`} \OperatorTok{!=}\StringTok{ `}\DataTypeTok{Q1-3_A~M_1981}\StringTok{`}\NormalTok{) }\OperatorTok{&}\StringTok{ }\OperatorTok{!}\KeywordTok{is.na}\NormalTok{(}\StringTok{`}\DataTypeTok{Q1-3_A~M_1981}\StringTok{`}\NormalTok{)}
                      \OperatorTok{~}\StringTok{ }\OtherTok{TRUE}\NormalTok{,}
\NormalTok{                      (}\StringTok{`}\DataTypeTok{Q1-3_A~Y_1979}\StringTok{`} \OperatorTok{!=}\StringTok{ `}\DataTypeTok{Q1-3_A~Y_1981}\StringTok{`}\NormalTok{) }\OperatorTok{&}\StringTok{ }\OperatorTok{!}\KeywordTok{is.na}\NormalTok{(}\StringTok{`}\DataTypeTok{Q1-3_A~Y_1981}\StringTok{`}\NormalTok{)}
                      \OperatorTok{~}\StringTok{ }\OtherTok{TRUE}\NormalTok{,}
\NormalTok{                      (}\StringTok{`}\DataTypeTok{Q1-3_A~Y_1979}\StringTok{`} \OperatorTok{==}\StringTok{ `}\DataTypeTok{Q1-3_A~Y_1981}\StringTok{`}\NormalTok{) }\OperatorTok{&}\StringTok{ }
\StringTok{                      }\NormalTok{(}\StringTok{`}\DataTypeTok{Q1-3_A~M_1979}\StringTok{`} \OperatorTok{==}\StringTok{ `}\DataTypeTok{Q1-3_A~M_1981}\StringTok{`}\NormalTok{) }\OperatorTok{~}\StringTok{ }\OtherTok{FALSE}\NormalTok{,}
                      \KeywordTok{is.na}\NormalTok{(}\StringTok{`}\DataTypeTok{Q1-3_A~M_1981}\StringTok{`}\NormalTok{) }\OperatorTok{|}\StringTok{ }\KeywordTok{is.na}\NormalTok{(}\StringTok{`}\DataTypeTok{Q1-3_A~Y_1981}\StringTok{`}\NormalTok{) }\OperatorTok{~}\StringTok{ }\OtherTok{FALSE}\NormalTok{)) }\OperatorTok{%>%}
\StringTok{  }\NormalTok{dplyr}\OperatorTok{::}\KeywordTok{select}\NormalTok{(CASEID_}\DecValTok{1979}\NormalTok{,}
\NormalTok{         dob_month,}
\NormalTok{         dob_year,}
\NormalTok{         dob_conflict)}
\end{Highlighting}
\end{Shaded}

For \texttt{gender} and \texttt{race}, we only renamed these variables.

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{## tidy the gender and race variables}
\NormalTok{demog_tidy <-}\StringTok{ }\NormalTok{categories_qnames }\OperatorTok{%>%}
\StringTok{  }\NormalTok{dplyr}\OperatorTok{::}\KeywordTok{select}\NormalTok{(CASEID_}\DecValTok{1979}\NormalTok{,}
\NormalTok{         SAMPLE_RACE_78SCRN,}
\NormalTok{         SAMPLE_SEX_}\DecValTok{1979}\NormalTok{) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{rename}\NormalTok{(}\DataTypeTok{gender =}\NormalTok{ SAMPLE_SEX_}\DecValTok{1979}\NormalTok{,}
         \DataTypeTok{race =}\NormalTok{ SAMPLE_RACE_78SCRN)}
\end{Highlighting}
\end{Shaded}

The next step is tidying the \texttt{highest\ grade\ completed} variable. This variable came with several version in each year. We chose the revised May data because it seemed to have less missing and presumably has been checked. However, there was no revised May data for 2012, 2014, 2016, and 2018 so we just used the ordinary May data for these years. The code for tidying this variable is provided in the supplementary code whose details could be seen in the Supplementary Materials section of this paper.

Further, \texttt{highest\ grade\ completed} is measured and could be updated in each period of the survey. Hence, we only used the highest grade completed ever and derived the year when it is completed. The code for tidying this variable is also available in the supplementary code of this paper.

Finally, we join all of the tidy variables in a data set called \texttt{full\_demographics} as displayed as follows:

\begin{verbatim}
##   id dob_month dob_year dob_conflict                    race gender yr_hgc
## 1  1         9       58        FALSE NON-BLACK, NON-HISPANIC FEMALE   1979
## 2  2         1       59        FALSE NON-BLACK, NON-HISPANIC FEMALE   1985
## 3  3         8       61        FALSE NON-BLACK, NON-HISPANIC FEMALE   1993
## 4  4         8       62        FALSE NON-BLACK, NON-HISPANIC FEMALE   1986
## 5  5         7       59        FALSE NON-BLACK, NON-HISPANIC   MALE   1984
## 6  6        10       60        FALSE NON-BLACK, NON-HISPANIC   MALE   1983
##   hgc_i          hgc
## 1    12   12TH GRADE
## 2    12   12TH GRADE
## 3    12   12TH GRADE
## 4    14 2ND YEAR COL
## 5    18 6TH YEAR COL
## 6    16 4TH YEAR COL
\end{verbatim}

\hypertarget{tidying-employment-variables}{%
\subsubsection{Tidying employment variables}\label{tidying-employment-variables}}

The employment data comprises of three variables, i.e.~\texttt{total\ hours\ of\ work\ per\ week}, \texttt{number\ of\ jobs\ that\ an\ individual\ has}, and \texttt{mean\ hourly\ wage}.

\texttt{hours\ worked\ per\ week} initially has only one version per job, no choice from 1979 to 1987 (QES-52A). From 1988 onward, when we had more options, we chose the variable for total hours including time spent working from home (QES-52D). However, in 1993, this variable did not have all the five D variables (the first one and the last one were missing), so we used QES-52A variable instead. In addition, 2008 only had jobs 1-4 for the QES-52D variable (whereas the other years had 1-5), so we just used these.

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# make a list for years where we used the "QES-52A"}
\NormalTok{year_A <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\DecValTok{1979}\OperatorTok{:}\DecValTok{1987}\NormalTok{, }\DecValTok{1993}\NormalTok{)}
\CommentTok{#function to get the hour of work}
\NormalTok{get_hour <-}\StringTok{ }\ControlFlowTok{function}\NormalTok{(year)\{}
  \ControlFlowTok{if}\NormalTok{(year }\OperatorTok{%in%}\StringTok{ }\NormalTok{year_A)\{}
\NormalTok{   temp <-}\StringTok{ }\NormalTok{new_data_qnames }\OperatorTok{%>%}
\StringTok{    }\NormalTok{dplyr}\OperatorTok{::}\KeywordTok{select}\NormalTok{(CASEID_}\DecValTok{1979}\NormalTok{,}
            \KeywordTok{starts_with}\NormalTok{(}\StringTok{"QES-52A"}\NormalTok{) }\OperatorTok{&}
\StringTok{              }\KeywordTok{ends_with}\NormalTok{(}\KeywordTok{as.character}\NormalTok{(year)))\} }
  \ControlFlowTok{else}\NormalTok{\{}
\NormalTok{    temp <-}\StringTok{ }\NormalTok{new_data_qnames }\OperatorTok{%>%}
\StringTok{    }\NormalTok{dplyr}\OperatorTok{::}\KeywordTok{select}\NormalTok{(CASEID_}\DecValTok{1979}\NormalTok{,}
            \KeywordTok{starts_with}\NormalTok{(}\StringTok{"QES-52D"}\NormalTok{) }\OperatorTok{&}
\StringTok{              }\KeywordTok{ends_with}\NormalTok{(}\KeywordTok{as.character}\NormalTok{(year)))\} }
\NormalTok{    temp }\OperatorTok{%>%}\StringTok{ }
\StringTok{      }\KeywordTok{pivot_longer}\NormalTok{(}\OperatorTok{!}\NormalTok{CASEID_}\DecValTok{1979}\NormalTok{,}
                 \DataTypeTok{names_to =} \StringTok{"job"}\NormalTok{,}
                 \DataTypeTok{values_to =} \StringTok{"hours_work"}\NormalTok{) }\OperatorTok{%>%}
\StringTok{      }\KeywordTok{separate}\NormalTok{(}\StringTok{"job"}\NormalTok{, }\KeywordTok{c}\NormalTok{(}\StringTok{"job"}\NormalTok{, }\StringTok{"year"}\NormalTok{), }\DataTypeTok{sep =} \DecValTok{-4}\NormalTok{) }\OperatorTok{%>%}
\StringTok{      }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{job =} \KeywordTok{paste0}\NormalTok{(}\StringTok{"job_"}\NormalTok{, }\KeywordTok{substr}\NormalTok{(job, }\DecValTok{9}\NormalTok{, }\DecValTok{10}\NormalTok{))) }\OperatorTok{%>%}
\StringTok{      }\KeywordTok{rename}\NormalTok{(}\DataTypeTok{id =}\NormalTok{ CASEID_}\DecValTok{1979}\NormalTok{)}
\NormalTok{\}}

\CommentTok{# list to save the iteration result}
\NormalTok{hours <-}\StringTok{ }\KeywordTok{list}\NormalTok{()}
\CommentTok{# getting the hours of work of all observations}
\ControlFlowTok{for}\NormalTok{(ayear }\ControlFlowTok{in} \KeywordTok{c}\NormalTok{(}\DecValTok{1979}\OperatorTok{:}\DecValTok{1994}\NormalTok{, }\DecValTok{1996}\NormalTok{, }\DecValTok{1998}\NormalTok{, }\DecValTok{2000}\NormalTok{, }\DecValTok{2002}\NormalTok{, }\DecValTok{2004}\NormalTok{, }\DecValTok{2006}\NormalTok{, }\DecValTok{2008}\NormalTok{, }\DecValTok{2010}\NormalTok{, }
               \DecValTok{2012}\NormalTok{, }\DecValTok{2014}\NormalTok{, }\DecValTok{2016}\NormalTok{, }\DecValTok{2018}\NormalTok{)) \{}
\NormalTok{   hours[[ayear]] <-}\StringTok{ }\KeywordTok{get_hour}\NormalTok{(ayear)}
\NormalTok{\}}
\CommentTok{# unlist the hours of work}
\NormalTok{hours_all <-}\StringTok{ }\KeywordTok{bind_rows}\NormalTok{(}\OperatorTok{!!!}\NormalTok{hours)}
\end{Highlighting}
\end{Shaded}

The same way was also deployed to tidy the rate of wage by year and by ID. The difference is that the hourly rate had only one version of each year. The hours of work and the hourly rate were then joined to calculate the number of jobs that a respondent has and their mean hourly wage. Some observations had 0 in their hourly rate, which is considered as invalid value. Thus, their hourly rate set to be N.A. the result of the joining variables could be seen in the following tibble.

\begin{verbatim}
## # A tibble: 6 x 5
##      id job    year  rate_per_hour hours_work
##   <int> <chr>  <chr>         <int>      <int>
## 1     1 job_01 1979            328         38
## 2     1 job_02 1979             NA         15
## 3     1 job_03 1979             NA         NA
## 4     1 job_04 1979             NA         NA
## 5     1 job_05 1979             NA         NA
## 6     2 job_01 1979            385         35
\end{verbatim}

Since our ultimate goal is to calculate the mean hourly wage, the number of jobs is calculate based on the availability of the \texttt{rate\_per\_hour} information. For example, the number of jobs of ID 1, based on \texttt{hours\_work}, is 2. However, since the information of hourly rate of \texttt{job\_02} is not available, the number of job is considered as 1.

Further, we calculated the mean hourly wage for each ID in each year using a weighted mean with the hours of work as the weight. However, there are a lot of missing value in \texttt{hours\_work} variable. In that case, we only calculated the mean hourly wage based on arithmetic/regular mean method. Hence, we created a new variable to flag whether the mean hourly wage is a weighted or a regular mean. Additionally, if an ID only had one job, we directly used their hourly wages information and flagged it as an arithmetic mean. The following tibble shows the result of this tidying process.

\begin{verbatim}
## # A tibble: 10 x 6
##       id  year mean_hourly_wage total_hours number_of_jobs is_wm
##    <int> <dbl>            <dbl>       <int>          <dbl> <lgl>
##  1     1  1979             3.28          38              1 FALSE
##  2     1  1981             3.61          NA              1 FALSE
##  3     2  1979             3.85          35              1 FALSE
##  4     2  1980             4.57          NA              1 FALSE
##  5     2  1981             5.14          NA              1 FALSE
##  6     2  1982             5.71          35              1 FALSE
##  7     2  1983             5.71          NA              1 FALSE
##  8     2  1984             5.14          NA              1 FALSE
##  9     2  1985             7.71          NA              1 FALSE
## 10     2  1986             7.69          NA              1 FALSE
\end{verbatim}

The \texttt{mean\_hourly\_wage} and \texttt{full\_demographic} data are then joined. We also filtered the data to only have the cohort who completed the education up to 12th grade and participated at least five rounds in the survey and save it to an object called \texttt{wages\_demog\_hs}.

\hypertarget{initial-data-analysis}{%
\subsection{Initial Data Analysis}\label{initial-data-analysis}}

According to Huebner, Vach, and Cessie (2016), Initial Data Analysis (IDA) is the step of inspecting and screening the data after being collected to ensure that the data is clean, valid, and ready to be deployed in the later formal statistical analysis. Moreover, Chatfield (1985) argued that the two main objectives of IDA is data description, which is to assess the structure and the quality of the data; and model formulation without any formal statistical inference.

In this paper, we conducted an IDA or a preliminary data analysis to assess the consistency of the data with the cohort information that is provided by the NLSY. In addition, we also aimed to find the anomaly in the wages values using this approach. We mainly used graphical summary to do the IDA using \texttt{ggplot2}(Wickham 2016) and \texttt{brolgar} (Tierney, Cook, and Prvan 2020).

As stated previously, the respondents' ages ranged from 12 to 22 when first interviewed in 1979. Hence, we would like to validate whether all of the respondents were in this range. Additionally, the \href{https://www.nlsinfo.org/content/cohorts/nlsy79/intro-to-the-sample/nlsy79-sample-introduction}{NLSY} also provided the number of the survey cohort by their gender (6,403 males and 6,283 females) and race (7,510 Non-Black/Non-Hispanic; 3,174 Black; 2,002 Hispanic). To validate this, we used the \texttt{full\_demographic} i.e.~the data with the survey years 1979 sample. Table \ref{tab:age-table} and Table \ref{tab:gender-race-table} suggest that the demographic data we had is consistent with the sample information in the database.

\begin{table}

\caption{\label{tab:age-table}Age Distribution of the NLSY79 samples}
\centering
\begin{tabular}[t]{r|r}
\hline
Age & Number of Sample\\
\hline
\rowcolor{gray!6}  15 & 1265\\
\hline
16 & 1550\\
\hline
\rowcolor{gray!6}  17 & 1600\\
\hline
18 & 1530\\
\hline
\rowcolor{gray!6}  19 & 1662\\
\hline
20 & 1722\\
\hline
\rowcolor{gray!6}  21 & 1677\\
\hline
22 & 1680\\
\hline
\end{tabular}
\end{table}

\begin{table}

\caption{\label{tab:gender-race-table}Gender and Race Distribution of the NLSY79 Samples}
\centering
\begin{tabular}[t]{l|l|l|l|l}
\hline
\multicolumn{1}{c|}{ } & \multicolumn{3}{c|}{Race} & \multicolumn{1}{c}{ } \\
\cline{2-4}
Gender & Hispanic & Black & Non-Black, Non-Hispanic & Total\\
\hline
\rowcolor{gray!6}  Male & 1000 (15.62\%) & 1613 (25.19\%) & 3790 (59.19\%) & 6403 (100.00\%)\\
\hline
Female & 1002 (15.95\%) & 1561 (24.84\%) & 3720 (59.21\%) & 6283 (100.00\%)\\
\hline
\rowcolor{gray!6}  Total & 2002 (15.78\%) & 3174 (25.02\%) & 7510 (59.20\%) & 12686 (100.00\%)\\
\hline
\end{tabular}
\end{table}

\begin{figure}
\centering
\includegraphics{figures/featureplot-1.pdf}
\caption{Two plots showing the distribution of the mean hourly wage. Plot A portrays the pattern of mean hourly wage of high school cohort from 1979 to 2018 of each ID in US Dollar; Plot B shows the distribution of their minimum, median, and maximum value. We can see that some IDs had an extremely high of wages and it made the distribution of the three features is extremely skewed.}
\end{figure}

The next step is that we explored the mean hourly wage data, in this case, we only explored the wages data in \texttt{wages\_demog\_hs}. Figure \ref{fig:featureplot} conveys that there is clearly a problem in the mean hourly wage values. Figure \ref{fig:featureplot} A shows that some observations had an exceptionally high figure of wages, even more than US\$10,000 per hour. In Figure \ref{fig:featureplot} B, we barely see any difference in the minimum, median, an maximum value of the wages since the distribution is heavily skewed to the right. Additionally, Table \ref{tab:summarytable} shows that the overall wages median of the cohort is only 7.2, while the mean is 11.87. It indicates that the data might contains a lot of extreme values.

\begin{table}

\caption{\label{tab:summarytable}Summary Statistics of Wages of High School Data}
\centering
\begin{tabular}[t]{l|r}
\hline
Statistics & Value\\
\hline
\rowcolor{gray!6}  Min. & 0.01000\\
\hline
1st Qu. & 4.50000\\
\hline
\rowcolor{gray!6}  Median & 7.20000\\
\hline
Mean & 11.86578\\
\hline
\rowcolor{gray!6}  3rd Qu. & 11.74000\\
\hline
Max. & 33400.00000\\
\hline
\end{tabular}
\end{table}

In Figure \ref{fig:high-wages}, we plotted some respondents with a high value of mean hourly wages. We filtered all of the IDs who earned more than US\$ 500 per hour averagely. We found that most of these respondents only experienced one point of extremely high wages. Thus, we suspected that these high values are erroneous values resulted from a data entry error.

\begin{figure}
\centering
\includegraphics{figures/high-wages-1.pdf}
\caption{\label{fig:high-wages}6 out of 45 IDs with extremely high mean hourly wage. Most of the IDs only have one point of high wage.}
\end{figure}

Further, we took 36 samples randomly from the data and plotted it as is seen in Figure \ref{fig:sampleplot}. It implies that not only that some observations earned an extremely high figures of wages, but some also had a reasonably fluctuate wages, for example the IDs in panel number 5, 7, and 11. The plot also implies that the samples had a different pattern of mean hourly wages. Some had a flat wages for years but had a sudden increase in a year than it went down again, while the other experienced a upsurge in their wage, for instance the IDs in panel 9.

According to Pergamit et al. (2001), one of the flaws of the NLSY79 employment data is that since the NLSY79 collect the information of the working hours since the last interview, it might be challenging for the respondents to track the within-job hours changes that happens between survey year, especially for the respondents with fluctuate working hours or whose job is seasonal. It even has been more challenging since 1994, where the respondents had to recall two years period. This shortcoming might also contribute on the fluctuation of one's wages data.

\begin{figure}
\centering
\includegraphics{figures/sampleplot-1.pdf}
\caption{The mean hourly wages of some random samples are shown in twelve facets, three IDs per facet. It suggests that some IDs had a reasonably fluctuate wages.}
\end{figure}

\hypertarget{replacing-extreme-values}{%
\subsubsection{Replacing extreme values}\label{replacing-extreme-values}}

As it is seen from figure \ref{fig:sampleplot}, there are many spikes in the mean hourly wage data. As part of the IDA, which is the model formulation, we built a robust linear regression model to address this issue. The notion of robust linear regression is to yield an estimation that is robust to the influence of noise or contamination (Koller 2016). It also aims to detect the contamination by weighting each observation based on how ``well-behave'' they are, known as robustness weight. Observations with lower robustness weight are suggested as an outliers by this method (Koller 2016).

In this paper, we built the model using the \texttt{rlm} function from \texttt{MASS} package (Venables and Ripley 2002). We set the \texttt{mean\_hourly\_wage} and \texttt{year} as the dependent and predictor respectively. Furthermore, we used M-Estimation with Huber weighting where the observation with small residual get a weight of 1, while the larger the residual, the smaller the weight (less than 1) (UCLA: Statistical Consulting Group 2021).

Since we worked with longitudinal data, we should built the model for each ID, instead of the overall data. The robust mixed model is actually the best model to be employed in this case. However, this method is too computationally and memory expensive, especially for a large data set, like the NLSY79 data. Thus, the model for each ID is built utilizing the \texttt{nest} and \texttt{map} function from \texttt{tidyr} (H. Wickham 2020b) and \texttt{purrr} (Henry and Wickham 2020) respectively.

The challenging part of detecting the anomaly using the robustness weight is to determine the threshold of the weight where the observations considered as outliers. Moreover, it should be noted that not all the outliers is due to an error, instead it might be that one had a reasonably increasing or decreasing wages. To, minimize the risk of being mistakenly regard an outlier as an ``erronous outlier'', we have simulated some threshold and study the behavior of the spikes in each threshold. We found that 0.12 is the most reasonable value to be the threshold to minimize the risk of that drawback because it still capture the sensible spikes in the data. In other words, we kept maintaining the natural variability of the wages while minimizing the presence of anomalies because of the error in the data recording. After deciding the threshold, we imputed the observations with weight less than 0.2 wage with the models' predicted value. We then flagged those observations in a new variable called \texttt{is\_pred},

Figure \ref{fig:comppict} A shows that after imputing the anomalies with the models' predicted value, the highest wages value has decreased to be around \$350. The spikes were still observed, but are not as extreme as the original data set. In Figure \ref{fig:comppict} B), we plotted the three features of mean hourly wages, namely the minimum, median, an maximum value, transformed to log scale. The plot implies that the skewness are slightly negative. We also see that the three features are overlapped each other. It indicates that some ID's minimum wage are higher than some ID's maximum wage.

Furthermore, Figure \ref{fig:compare} implies that after the treatment, the fluctuation can still be observed in the data and only the large spikes, which are considered as ``erronous outliers'', are eliminated from the data. Hence, the model results a data set with the reasonable degree of fluctuation.

\begin{figure}
\centering
\includegraphics{figures/comppict-1.pdf}
\caption{The distribution of the mean hourly wage after treating the extreme values. Plot A portrays the pattern of mean hourly wage of high school cohort from 1979 to 2018 of each ID in US Dollar; Plot B shows the distribution of their minimum, median, and maximum value transformed to log10 scale. We can see that some observations still had reasonbaly higher wages than the others. Also, some IDs' have a minimum wages that is higher than others' maximum wages.}
\end{figure}

\begin{figure}
\centering
\includegraphics{figures/compare-1.pdf}
\caption{Comparison between the original and the treated mean hourly wage. The orange line portray the original value of mean hourly wage, while the turquoise line display the mean hourly wages value after the extreme values imputed with the robust linear model's prediction value. We can see that some extreme spikes has been reduced by the model.}
\end{figure}

Finally, we saved the imputed data and set the appropriate data type for the variables. As our target data is the mean hourly wage of the high-school dropouts, we then subset the high-school graduate data set to be only having the male-high school dropout data. We also saved the NLSY79 cohort's demographic information in a separate dataset. We then make these three data sets and its processing documentation to be publicly available through an R data container package called \texttt{yowie}. The complete flow from the raw data to these data set is displayed in Figure \ref(fig:flowchart).

\begin{figure}
\centering
\includegraphics{figures/flowchart-1.pdf}
\caption{The stages of data filtering from the raw data to get three datasets that are contained in an R package called yowie, n means the number of ID, while n\_obs means the number of observations.}
\end{figure}

\hypertarget{exploratory-data-analysis}{%
\section{Exploratory Data Analysis}\label{exploratory-data-analysis}}

In this part, we gave some examples of how the data might be used for data analysis in text book. We used \texttt{brolgar} (Tierney, Cook, and Prvan 2020) to perform exploratory data analysis. The first example is that we observed relationship between wages and education in terms of how the wages increase year by year. Using \texttt{key\_slope} function in \texttt{brolgar}, we performed a liner regression with mean hourly wage as response variable and year as predictor and extracted the regression slope of each ID and plotted it according to highest grade completed.

We found in Figure \ref{fig:wages-slope} that the higher the education completed does not necessarily increase the wages more from year to year. We can see that people who completed 4th grade happen to have relatively the same median of slope as people who completed 11th grade, even though it is probably because some respondents in this group had an extremely high mean hourly wages. Moreover, the annual increasing wages of people who completed 6th grade to 11th grade are relatively same. People who completed 12th grade tend to have higher slope of year compared to the other groups. Some outliers are also spotted in this group. Additionally, we can see that some respondents have a negative slope, meaning that their wages tend to decrease over the year.

\begin{figure}
\centering
\includegraphics{figures/wages-slope-1.pdf}
\caption{Regression slope of mean hourly wage regression by highest grade completed. The regression model has mean hourly wage as response variable and year as predictor. Respondents who completed 12th grade has the highest median of slope.}
\end{figure}

In the second example, we examined the pattern of mean hourly wage corresponding to different groups of gender and race. We took 240 samples of female and male equally and spread it into twelve facets as is seen in Figure \ref{fig:gender-plot} using \texttt{facet\_strata} function from \texttt{brolgar} (Tierney, Cook, and Prvan 2020). We learn that the mean hourly wage of female and male fluctuated over the years. In some facets, for example in facet 4, 5, and 10, males tend to earn more wages than females. This finding is also shown in Figure \ref{fig:gender-se} where we plotted the median wage of males and females against the year. We learn that males had a higher mean of mean hourly wage over the period of survey than females. Unfortunately, the gap between them tend to be wider from time to time.

\begin{figure}
\centering
\includegraphics{figures/gender-plot-1.pdf}
\caption{The pattern of mean hourly wages overtime displayed in 12 facets coloured by gender. The respondents belong to high school cohort are randomly sampled. We learn that both male and female respondents had a fluctuate mean hourly wage over the year.}
\end{figure}

\begin{figure}
\centering
\includegraphics{figures/gender-se-1.pdf}
\caption{The yearly mean of mean hourly wage by gender along with its standard error. Males generally earn more wages than females. The error interval tend to be wider over the years.}
\end{figure}

For the final example, we show the example of modeling the longitudinal data using Generelized Additive Model (GAM) with \texttt{mgcv} package (Wood 2003) for each ID. We fit a more flexible model to fit in to the data since the relationship between mean hourly wage and year is not linear due to the spikes. Since this approach is computationally expensive, we only show the model in a little fraction of the data. We sample 1 percent of the total respondents using \texttt{brolgar\textquotesingle{}s} \texttt{sample\_frac\_keys} function. Further, we plotted the some sample of the fitted model in Figure @ref\{fig:gam-plot\}. We learn that the model flexible enough to deal with the fluctuation in the data. Moreover, model with small spikes tend to be linearly fitted.

\begin{figure}
\centering
\includegraphics{figures/gam-plot-1.pdf}
\caption{Exploration of wages data by fitting a GAM. The fitted model displayed by blue line. The fitted line shows that the model flexible enough to follow the pattern of the data.}
\end{figure}

\hypertarget{summary}{%
\section{Summary}\label{summary}}

This paper has performed a set of stages to make an open data suitable for a text book data or to make it ready for research. In the first stage, we showed the steps that are performed to get the data from the NLSY79 database. Since the data format is untidy, we showed how the data has been tidied. After that, we conducted an initial data analysis to investigate and screen the quality of the data. Using robust linear regression model, we found and fixed the anomalous observations in the data set with its predicted values. We also performed an example of exploratory data analysis using the cleaned data set.

This paper also has demonstrated the documentation of data cleaning by providing all of the codes in performing data tidying and initial data analysis. Thus, this paper provided an opportunity to continuously refresh the text book data whenever the updated data is published in the NLSY79 database. This could be done by following the documentation of the code that is provided in this paper.

Moreover, the documentation also includes the process of how did we generate the robustness weight and how did we decide the threshold of anomalous observations. It is also documented along with the flag of whether an observation is imputed value or not. Accordingly, if somebody wish to make another decision, it can be done by making a small changes in the code provided. Further, this paper is also supplemented by a \texttt{shiny} (Chang et al. 2020) app as a simulation tool to customize the weight threshold.

Finally, this paper implies that data providers should design the database that is able to produce tidy data sets. A data provider should also check for data anomalies prior the data publishing or at least provides a set of rules or threshold value, for example, in this case is the threshold of reasonable wages. This will greatly support the data users to carry out further validation and set the same understanding of which data are considered as outliers. Moreover, providing validation rules would facilitate the usage of any establish data validation tool, such as \texttt{validate} (van der Loo and de Jonge 2021) package. In this case, we cannot use this handy validation package due to the absence of validation rules.

\hypertarget{acknowledgements}{%
\section{Acknowledgements}\label{acknowledgements}}

We would like to thank Aarathy Babu for the insight and discussion during the writing of this paper.

The entire analysis is conducted using \texttt{R} (R Core Team 2020) in \texttt{rstudio} using these packages: \texttt{tidyverse} (H. Wickham, Averick, et al. 2019), \texttt{ggplot2} (Wickham 2016), \texttt{dplyr} (H. Wickham, François, et al. 2020), \texttt{readr} (H. Wickham and Hester 2020), \texttt{tidyr} (H. Wickham 2020b), \texttt{stringr} (H. Wickham 2019), \texttt{purrr} (Henry and Wickham 2020), \texttt{broom} (Robinson, Hayes, and Couch 2021), \texttt{blorgar} (Tierney, Cook, and Prvan 2020), \texttt{patchwork} (Pedersen 2020), \texttt{kableExtra} (Zhu 2019), \texttt{MASS} (Venables and Ripley 2002), \texttt{janitor} (Firke 2020), \texttt{DiagrammeR} (Iannone 2020), \texttt{rsvg} (Ooms 2020), \texttt{webshot} (Chang 2019), \texttt{mgcv} (Wood 2003), \texttt{tsibble} (Wang, Cook, and Hyndman 2020), and \texttt{modelr} (H. Wickham 2020a). The paper are generated using \texttt{knitr} (Xie 2014), \texttt{rmarkdown} (Xie, Dervieux, and Riederer 2020), and \texttt{rticles} (Allaire et al. 2021).

\hypertarget{supplementary-materials}{%
\section{Supplementary Materials}\label{supplementary-materials}}

\begin{itemize}
\item
  \textbf{Codes} : R script to reproduce data tidying and cleaning are available in this \href{https://github.com/numbats/yowie/blob/master/data-raw/data_preprocessing.R}{Github Repository}.
\item
  \textbf{R Package \texttt{yowie}}:\texttt{yowie} is a data container R package that contains 3 datasets, namely the high school mean hourly wage data, high school dropouts mean hourly wage data, and demographic data of the NLSY79 cohort. This package could be accessed \href{https://github.com/numbats/yowie}{here}.
\item
  \textbf{shiny app}: An web interactive \texttt{shiny} app to run a simulation to customize the weight threshold. This app could be accessed \href{https://github.com/numbats/summer-wages-refresh/tree/main/app}{here}.
\end{itemize}

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-rticles}{}%
Allaire, JJ, Yihui Xie, R Foundation, Hadley Wickham, Journal of Statistical Software, Ramnath Vaidyanathan, Association for Computing Machinery, et al. 2021. \emph{Rticles: Article Formats for R Markdown}. \url{https://CRAN.R-project.org/package=rticles}.

\leavevmode\hypertarget{ref-webshot}{}%
Chang, Winston. 2019. \emph{Webshot: Take Screenshots of Web Pages}. \url{https://CRAN.R-project.org/package=webshot}.

\leavevmode\hypertarget{ref-shiny}{}%
Chang, Winston, Joe Cheng, JJ Allaire, Yihui Xie, and Jonathan McPherson. 2020. \emph{Shiny: Web Application Framework for R}. \url{https://CRAN.R-project.org/package=shiny}.

\leavevmode\hypertarget{ref-Chatfield1985TIEo}{}%
Chatfield, C. 1985. ``The Initial Examination of Data.'' \emph{Journal of the Royal Statistical Society. Series A. General} 148 (3): 214--53.

\leavevmode\hypertarget{ref-eliznlsy}{}%
Cooksey, Elizabeth C. 2017. ``Using the National Longitudinal Surveys of Youth (Nlsy) to Conduct Life Course Analyses.'' In \emph{Handbook of Life Course Health Development}, edited by Richard M. Lerner Neal Halfon Christoper B. Forrest, 561--77. Cham: Springer. \url{https://doi.org/https://doi.org/10.1007/978-3-319-47143-3_23}.

\leavevmode\hypertarget{ref-DasuTamraparni2003Edma}{}%
Dasu, Tamraparni, and Theodore Johnson. 2003. \emph{Exploratory Data Mining and Data Cleaning}. Wiley Series in Probability and Statistics. Hoboken: WILEY.

\leavevmode\hypertarget{ref-janitor}{}%
Firke, Sam. 2020. \emph{Janitor: Simple Tools for Examining and Cleaning Dirty Data}. \url{https://CRAN.R-project.org/package=janitor}.

\leavevmode\hypertarget{ref-purrr}{}%
Henry, Lionel, and Hadley Wickham. 2020. \emph{Purrr: Functional Programming Tools}. \url{https://CRAN.R-project.org/package=purrr}.

\leavevmode\hypertarget{ref-HuebnerMariannePhD2016Asat}{}%
Huebner, Marianne, Werner Vach, and Saskia le Cessie. 2016. ``A Systematic Approach to Initial Data Analysis Is Good Research Practice.'' \emph{The Journal of Thoracic and Cardiovascular Surgery} 151 (1): 25--27.

\leavevmode\hypertarget{ref-HuebnerMarianne2020Haar}{}%
Huebner, Marianne, Werner Vach, Saskia le Cessie, Carsten Oliver Schmidt, and Lara Lusa. 2020. ``Hidden Analyses: A Review of Reporting Practice and Recommendations for More Transparent Reporting of Initial Data Analyses.'' \emph{BMC Medical Research Methodology} 20 (1): 61--61.

\leavevmode\hypertarget{ref-DiagrammeR}{}%
Iannone, Richard. 2020. \emph{DiagrammeR: Graph/Network Visualization}. \url{https://CRAN.R-project.org/package=DiagrammeR}.

\leavevmode\hypertarget{ref-KollerManuel2016rARP}{}%
Koller, Manuel. 2016. ``Robustlmm: An R Package for Robust Estimation of Linear Mixed-Effects Models.'' \emph{Journal of Statistical Software} 75 (6): 1--24.

\leavevmode\hypertarget{ref-rsvg}{}%
Ooms, Jeroen. 2020. \emph{Rsvg: Render Svg Images into Pdf, Png, Postscript, or Bitmap Arrays}. \url{https://CRAN.R-project.org/package=rsvg}.

\leavevmode\hypertarget{ref-opendata}{}%
Open Knowledge Foundation. 2021. ``Open Definition. Defining Open in Open Data, Open Content, and Open Knowledge.'' 2021. \url{http://opendefinition.org/od/2.1/en/}.

\leavevmode\hypertarget{ref-patchwork}{}%
Pedersen, Thomas Lin. 2020. \emph{Patchwork: The Composer of Plots}. \url{https://CRAN.R-project.org/package=patchwork}.

\leavevmode\hypertarget{ref-MichaelRPergamit2001DWTN}{}%
Pergamit, Michael R., Charles R. Pierret, Donna S. Rothstein, and Jonathan R. Veum. 2001. ``Data Watch: The National Longitudinal Surveys.'' \emph{The Journal of Economic Perspectives} 15 (2): 239--53.

\leavevmode\hypertarget{ref-R}{}%
R Core Team. 2020. \emph{R: A Language and Environment for Statistical Computing}. Vienna, Austria: R Foundation for Statistical Computing. \url{https://www.R-project.org/}.

\leavevmode\hypertarget{ref-broom}{}%
Robinson, David, Alex Hayes, and Simon Couch. 2021. \emph{Broom: Convert Statistical Objects into Tidy Tibbles}. \url{https://CRAN.R-project.org/package=broom}.

\leavevmode\hypertarget{ref-SingerJudithD2003Alda}{}%
Singer, Judith D, and John B Willett. 2003. \emph{Applied Longitudinal Data Analysis: Modeling Change and Event Occurrence}. Oxford u.a: Oxford Univ. Pr.

\leavevmode\hypertarget{ref-nlsy79}{}%
The U.S. Bureau of Labor Statistics. n.d. ``National Longitudinal Survey of Youth 1979.'' Available at \url{https://www.nlsinfo.org/content/cohorts/nlsy79} (2021/25/02).

\leavevmode\hypertarget{ref-brolgar}{}%
Tierney, Nicholas, Di Cook, and Tania Prvan. 2020. \emph{Brolgar: BRowse over Longitudinal Data Graphically and Analytically in R}. \url{https://github.com/njtierney/brolgar}.

\leavevmode\hypertarget{ref-tukey}{}%
Tukey, John W. (John Wilder). 1977. \emph{Exploratory Data Analysis}. Addison-Wesley Series in Behavioral Science. Reading, Mass.: Addison-Wesley Pub. Co.

\leavevmode\hypertarget{ref-rlm}{}%
UCLA: Statistical Consulting Group. 2021. ``Robust Regression \textbar{} R Data Analysis Examples.'' February 2021. \url{https://stats.idre.ucla.edu/r/dae/robust-regression/}.

\leavevmode\hypertarget{ref-validate}{}%
van der Loo, Mark P. J., and Edwin de Jonge. 2021. ``Data Validation Infrastructure for R.'' \emph{Journal of Statistical Software} 97 (10): 1--31. \url{https://doi.org/10.18637/jss.v097.i10}.

\leavevmode\hypertarget{ref-mass}{}%
Venables, W. N., and B. D. Ripley. 2002. \emph{Modern Applied Statistics with S}. Fourth. New York: Springer. \url{http://www.stats.ox.ac.uk/pub/MASS4}.

\leavevmode\hypertarget{ref-tsibble}{}%
Wang, Earo, Dianne Cook, and Rob J Hyndman. 2020. ``A New Tidy Data Structure to Support Exploration and Modeling of Temporal Data.'' \emph{Journal of Computational and Graphical Statistics} 29 (3): 466--78. \url{https://doi.org/10.1080/10618600.2019.1695624}.

\leavevmode\hypertarget{ref-WickhamHadley2014TD}{}%
Wickham, Hadley. 2014. ``Tidy Data.'' \emph{Journal of Statistical Software} 59 (10): 1--23.

\leavevmode\hypertarget{ref-ggplot2}{}%
---------. 2016. \emph{Ggplot2: Elegant Graphics for Data Analysis}. Springer-Verlag New York. \url{https://ggplot2.tidyverse.org}.

\leavevmode\hypertarget{ref-stringr}{}%
---------. 2019. \emph{Stringr: Simple, Consistent Wrappers for Common String Operations}. \url{https://CRAN.R-project.org/package=stringr}.

\leavevmode\hypertarget{ref-modelr}{}%
---------. 2020a. \emph{Modelr: Modelling Functions That Work with the Pipe}. \url{https://CRAN.R-project.org/package=modelr}.

\leavevmode\hypertarget{ref-tidyr}{}%
---------. 2020b. \emph{Tidyr: Tidy Messy Data}. \url{https://CRAN.R-project.org/package=tidyr}.

\leavevmode\hypertarget{ref-tidyverse}{}%
Wickham, Hadley, Mara Averick, Jennifer Bryan, Winston Chang, Lucy D'Agostino McGowan, Romain François, Garrett Grolemund, et al. 2019. ``Welcome to the tidyverse.'' \emph{Journal of Open Source Software} 4 (43): 1686. \url{https://doi.org/10.21105/joss.01686}.

\leavevmode\hypertarget{ref-dplyr}{}%
Wickham, Hadley, Romain François, Lionel Henry, and Kirill Müller. 2020. \emph{Dplyr: A Grammar of Data Manipulation}. \url{https://CRAN.R-project.org/package=dplyr}.

\leavevmode\hypertarget{ref-readr}{}%
Wickham, Hadley, and Jim Hester. 2020. \emph{Readr: Read Rectangular Text Data}. \url{https://CRAN.R-project.org/package=readr}.

\leavevmode\hypertarget{ref-mgcv}{}%
Wood, S. N. 2003. ``Thin-Plate Regression Splines.'' \emph{Journal of the Royal Statistical Society (B)} 65 (1): 95--114.

\leavevmode\hypertarget{ref-knitr}{}%
Xie, Yihui. 2014. ``Knitr: A Comprehensive Tool for Reproducible Research in R.'' In \emph{Implementing Reproducible Computational Research}, edited by Victoria Stodden, Friedrich Leisch, and Roger D. Peng. Chapman; Hall/CRC. \url{http://www.crcpress.com/product/isbn/9781466561595}.

\leavevmode\hypertarget{ref-rmarkdown}{}%
Xie, Yihui, Christophe Dervieux, and Emily Riederer. 2020. \emph{R Markdown Cookbook}. Boca Raton, Florida: Chapman; Hall/CRC. \url{https://bookdown.org/yihui/rmarkdown-cookbook}.

\leavevmode\hypertarget{ref-kableExtra}{}%
Zhu, Hao. 2019. \emph{KableExtra: Construct Complex Table with 'Kable' and Pipe Syntax}. \url{https://CRAN.R-project.org/package=kableExtra}.

\bibliographystyle{unsrt}
\bibliography{references.bib}

\end{document}

### 2021-05-05 report_alt.Rmd
fixed cross-referencing

@@
 
-```{r featureplot, echo = FALSE, fig.cap = "Two plots showing the distribution of the mean hourly wage.
+The next step is that we explored the mean hourly wage data, in this case, we only explored the wages data in `wages_demog_hs`.
+Table \ref{tab:summarytable} shows that the overall wages median of the cohort is only 7.2, while the mean is 11.87.
+It indicates that the data might contains a lot of extreme values.
+
+```{r summarytable, echo = FALSE}
+kable(as.array(summary(wages_demog_hs$mean_hourly_wage)), 
+      caption = "Summary Statistics of Wages of High School Data",
+      col.names = c("Statistics", "Value")) %>%
+  kable_styling(latex_options = "striped")
+```
+
+However, we can not be sure only by observing the summary statistics.
+Hence, we used visualisation techniques to investigate this matter.
+Figure \ref{fig:feature-plot} conveys a problem in the mean hourly wage values.
+Figure \ref{fig:feature-plot} A shows that some observations had an exceptionally high figure of wages, even more than US$10,000 per hour.
+In Figure \ref{fig:feature-plot} B, we barely see any difference in the minimum, median, and maximum value of the wages since the distribution is heavily skewed to the right.
+
+```{r feature-plot, echo = FALSE, fig.cap = "Two plots showing the distribution of the mean hourly wage.
 Plot A portrays the pattern of mean hourly wage of high school cohort from 1979 to 2018 of each ID in US Dollar; Plot B shows the distribution of their minimum, median, and maximum value.
@@
 
-The next step is that we explored the mean hourly wage data, in this case, we only explored the wages data in `wages_demog_hs`.
-Figure \@ref(fig:featureplot) conveys that there is clearly a problem in the mean hourly wage values.
-Figure \ref{fig:featureplot} A shows that some observations had an exceptionally high figure of wages, even more than US$10,000 per hour.
-In Figure \ref{fig:featureplot} B, we barely see any difference in the minimum, median, an maximum value of the wages since the distribution is heavily skewed to the right.
-Additionally, Table \ref{tab:summarytable} shows that the overall wages median of the cohort is only 7.2, while the mean is 11.87.
-It indicates that the data might contains a lot of extreme values.
-
-```{r summarytable, echo = FALSE}
-kable(as.array(summary(wages_demog_hs$mean_hourly_wage)), 
-      caption = "Summary Statistics of Wages of High School Data",
-      col.names = c("Statistics", "Value")) %>%
-  kable_styling(latex_options = "striped")
-```
-
 In Figure \ref{fig:high-wages}, we plotted some respondents with a high value of mean hourly wages.
-We filtered all of the IDs who earned more than US$ 500 per hour averagely.
-We found that most of these respondents only experienced one point of extremely high wages.
-Thus, we suspected that these high values are erroneous values resulted from a data entry error.
+We filtered all of the IDs who earned more than US$ 500 per hour on average.
+We found that these respondents only experienced one point of extremely high wages.
+Thus, we suspected that these high values are erroneous resulted from a data entry error.
 
@@
 
-Further, we took 36 samples randomly from the data and plotted it as is seen in Figure \ref{fig:sampleplot}.
-It implies that not only that some observations earned an extremely high figures of wages, but some also had a reasonably fluctuate wages, for example the IDs in panel number 5, 7, and 11.
+Further, we took 36 samples randomly from the data and plotted them, as shown in Figure \ref{fig:sample-plot}.
+It implies that not only that some observations earned extremely high figures of wages, but some also had reasonably fluctuated wages, for example, the IDs in panel numbers 5, 7, and 11.
 The plot also implies that the samples had a different pattern of mean hourly wages.
-Some had a flat wages for years but had a sudden increase in a year than it went down again, while the other experienced a upsurge in their wage, for instance the IDs in panel 9.
-
-According to @MichaelRPergamit2001DWTN, one of the flaws of the NLSY79 employment data is that since the NLSY79 collect the information of the working hours since the last interview, it might be challenging for the respondents to track the within-job hours changes that happens between survey year, especially for the respondents with fluctuate working hours or whose job is seasonal.
-It even has been more challenging since 1994, where the respondents had to recall two years period.
-This shortcoming might also contribute on the fluctuation of one's wages data.
-
-```{r sampleplot, echo = FALSE, fig.cap="The mean hourly wages of some random samples are shown in twelve facets, three IDs per facet.
+Some had flat wages for years but had a sudden increase in one particular year, then it went down again, while the others experienced an upsurge in their wage, for instance, the IDs in panel 9.
+
+```{r sample-plot, echo = FALSE, fig.cap="The mean hourly wages of some random samples are shown in twelve facets, three IDs per facet.
 It suggests that some IDs had a reasonably fluctuate wages.", fig.height=4, fig.width=5}
@@
 
+According to @MichaelRPergamit2001DWTN, one of the flaws of the NLSY79 employment data is that since the NLSY79 collect the information of the working hours since the last interview, it might be challenging for the respondents to track the within-job hours' changes that happens between survey year, especially for the respondents with fluctuate working hours or whose job is seasonal.
+It even has been more challenging since 1994, where the respondents had to recall two years period.
+This shortcoming might also contribute to the fluctuation of one's wages data.
+
 ### Replacing extreme values
 
-As it is seen from figure \ref{fig:sampleplot}, there are many spikes in the mean hourly wage data.
-As part of the IDA, which is the model formulation, we built a robust linear regression model to address this issue.
-The notion of robust linear regression is to yield an estimation that is robust to the influence of noise or contamination [@KollerManuel2016rARP].
-It also aims to detect the contamination by weighting each observation based on how "well-behave" they are, known as robustness weight.
-Observations with lower robustness weight are suggested as an outliers by this method [@KollerManuel2016rARP].
-
-In this paper, we built the model using the `rlm` function from `MASS` package [@mass].
-We set the `mean_hourly_wage` and `year` as the dependent and predictor respectively.
-Furthermore, we used M-Estimation with Huber weighting where the observation with small residual get a weight of 1, while the larger the residual, the smaller the weight (less than 1) [@rlm].
-
-Since we worked with longitudinal data, we should built the model for each ID, instead of the overall data.
-The robust mixed model is actually the best model to be employed in this case.
+As part of the IDA, which is the model formulation, we built a robust linear regression model to handle the extreme values in the data.
+Robust linear regression yields an estimation that is robust to the influence of noise or contamination [@KollerManuel2016rARP].
+It also aims to detect the contamination by weighting each observation based on how "well-behaved" they are, known as robustness weight.
+Observations with lower robustness weight are suggested as an outlier by this method [@KollerManuel2016rARP].
+
+Since we worked with longitudinal data, we built the model for each ID instead of the overall data.
+The robust mixed model is might be the best model to be employed in this case.
 However, this method is too computationally and memory expensive, especially for a large data set, like the NLSY79 data.
@@
 
-The challenging part of detecting the anomaly using the robustness weight is to determine the threshold of the weight where the observations considered as outliers.
-Moreover, it should be noted that not all the outliers is due to an error, instead it might be that one had a reasonably increasing or decreasing wages.
-To, minimize the risk of being mistakenly regard an outlier as an "erronous outlier", we have simulated some threshold and study the behavior of the spikes in each threshold.
-We found that 0.12 is the most reasonable value to be the threshold to minimize the risk of that drawback because it still capture the sensible spikes in the data.
-In other words, we kept maintaining the natural variability of the wages while minimizing the presence of anomalies because of the error in the data recording.
-After deciding the threshold, we imputed the observations with weight less than 0.2 wage with the models' predicted value.
-We then flagged those observations in a new variable called `is_pred`,
+We built the model using the `rlm` function from `MASS` package [@mass].
+We set the `mean_hourly_wage` and `year` as the dependent and predictor, respectively.
+Furthermore, we used M-Estimation with Huber weighting where the observation with a slight residual gets a weight of 1, while the larger the residual, the smaller the weight (less than 1) [@rlm].
+However, the challenging part of detecting the anomaly using the robustness weight is determining the threshold of the weight in which the observations are considered outliers.
+Moreover, it should be noted that not all the outliers are due to an error.
+Instead, it might be that one had reasonably increasing or decreasing wages in a particular period.
+
+To minimize the risk of mistakenly regarding an outlier as an "erroneous outlier", we have simulated some thresholds and study how they affected the data.
+We found that 0.12 is the most reasonable value to be the threshold to minimize that drawback's risk because it still captures the sensible spikes in the data.
+In other words, we kept maintaining the natural variability of the wages while minimizing anomalies because of the error in the data recording.
+After deciding the threshold, we imputed the observations whose weight less than 0.12 with the models' predicted value.
+We then flagged those observations in a new variable called `is_pred`.
 
@@
 
-Figure \ref{fig:comppict} A shows that after imputing the anomalies with the models' predicted value, the highest wages value has decreased to be around $350.
-The spikes were still observed, but are not as extreme as the original data set.
-In Figure \ref{fig:comppict} B), we plotted the three features of mean hourly wages, namely the minimum, median, an maximum value, transformed to log scale.
-The plot implies that the skewness are slightly negative.
+Figure \ref{fig:comppict} A shows the mean hourly wage feature after the extreme values are replaced.
+It shows that the highest value has decreased to be around $350.
+The spikes were still observed but are not as extreme as the original data set.
+In Figure \ref{fig:comppict} B), we plotted the three features of mean hourly wages, namely the minimum, median, and maximum value, transformed to log scale.
+The plot implies that the skewness is slightly negative.
 We also see that the three features are overlapped each other.
-It indicates that some ID's minimum wage are higher than some ID's maximum wage.
-
-Furthermore, Figure \ref{fig:compare} implies that after the treatment, the fluctuation can still be observed in the data and only the large spikes, which are considered as "erronous outliers", are eliminated from the data.
-Hence, the model results a data set with the reasonable degree of fluctuation.
-
-```{r comppict, echo = FALSE, fig.cap = "The distribution of the mean hourly wage after treating the extreme values.
+It indicates that some ID's minimum wages are higher than some ID's maximum wages.
+
+```{r comppict, echo = FALSE, fig.cap = "The distribution of the mean hourly wage after the extreme values are replaced.
 Plot A portrays the pattern of mean hourly wage of high school cohort from 1979 to 2018 of each ID in US Dollar; Plot B shows the distribution of their minimum, median, and maximum value transformed to log10 scale.
@@
 
-```{r compare, echo = FALSE, fig.cap="Comparison between the original and the treated mean hourly wage.
+Furthermore, Figure \ref{fig:compare-plot} implies that the fluctuation can still be observed in the data after the treatment.
+However, the large spikes, which are considered "erroneous outliers", are already eliminated from the data.
+Hence, the model produces a data set with a more reasonable degree of fluctuation.
+
+```{r compare-plot, echo = FALSE, fig.cap="Comparison between the original and the treated mean hourly wage.
 The orange line portray the original value of mean hourly wage, while the turquoise line display the mean hourly wages value after the extreme values imputed with the robust linear model's prediction value.
@@
 Finally, we saved the imputed data and set the appropriate data type for the variables.
-As our target data is the mean hourly wage of the high-school dropouts, we then subset the high-school graduate data set to be only having the male-high school dropout data.
+As our target data is the mean hourly wage of the high-school dropouts, we then subset the high-school graduate data set to have only the male-high school dropout data.
 We also saved the NLSY79 cohort's demographic information in a separate dataset.
-We then make these three data sets and its processing documentation to be publicly available through an R data container package called `yowie`.
-The complete flow from the raw data to these data set is displayed in Figure \ref(fig:flowchart).
+We then make these three data sets and their processing documentation publicly available through an R data container package called `yowie`.
+The complete flow from the raw data to these data set is displayed in Figure \ref{fig:flow-chart}.
 
@@
 
-```{r flowchart, echo=FALSE, fig.cap="The stages of data filtering from the raw data to get three datasets that are contained in an R package called yowie, n means the number of ID, while n_obs means the number of observations.", fig.height=3, fig.width=3}
+```{r flow-chart, echo=FALSE, fig.cap="The stages of data filtering from the raw data to get three datasets that are contained in an R package called yowie, n means the number of ID, while n_obs means the number of observations.", fig.height=3, fig.width=3}
 
@@
 
-In this part, we gave some examples of how the data might be used for data analysis in text book.
+This part gives some examples of how the data might be used for data analysis in a textbook.
 We used `brolgar` [@brolgar] to perform exploratory data analysis.
-The first example is that we observed relationship between wages and education in terms of how the wages increase year by year.
-Using `key_slope` function in `brolgar`, we performed a liner regression with  mean hourly wage as response variable and year as predictor and extracted the regression slope of each ID and plotted it according to highest grade completed.
-
-We found in Figure \ref{fig:wages-slope} that the higher the education completed does not necessarily increase the wages more from year to year.
-We can see that people who completed 4th grade happen to have relatively the same median of slope as people who completed 11th grade, even though it is probably because some respondents in this group had an extremely high mean hourly wages.
-Moreover, the annual increasing wages of people who completed 6th grade to 11th grade are relatively same.
-People who completed 12th grade tend to have higher slope of year compared to the other groups.
+In the first example, we observed the relationship between wages and education regarding how the wages increase year by year.
+Using `key_slope` function in `brolgar`, we performed linear regression with the mean hourly wage as response variable and year as the predictor, extracted the regression slope of each ID, and plotted it according to the highest grade completed.
+
+Figure \ref{fig:slope-fig} shows that the higher the education completed does not necessarily increase the wages more from year to year.
+We can see that people who completed 4th grade happen to have relatively the same median of slope as people who completed 11th grade, even though it is probably because some respondents in this group had extremely high mean hourly wages.
+Moreover, the annual increasing wages of people who completed 6th grade to 11th grade are relatively the same.
+People who completed 12th grade tend to have a higher slope of the year than the other groups.
 Some outliers are also spotted in this group.
@@
 
-```{r wages-slope, echo = FALSE, fig.cap="Regression slope of mean hourly wage regression by highest grade completed.
+```{r slope-fig, echo = FALSE, fig.cap="Regression slope of mean hourly wage regression by highest grade completed.
 The regression model has mean hourly wage as response variable and year as predictor.
@@
 In the second example, we examined the pattern of mean hourly wage corresponding to different groups of gender and race.
-We took 240 samples of female and male equally and spread it into twelve facets as is seen in Figure \ref{fig:gender-plot} using `facet_strata` function from `brolgar` [@brolgar].
-We learn that the mean hourly wage of female and male fluctuated over the years.
-In some facets, for example in facet 4, 5, and 10, males tend to earn more wages than females.
-This finding is also shown in Figure \ref{fig:gender-se} where we plotted the median wage of males and females against the year.
-We learn that males had a higher mean of mean hourly wage over the period of survey than females.
-Unfortunately, the gap between them tend to be wider from time to time.
-
-```{r gender-plot, echo = FALSE, warning = FALSE, fig.cap= "The pattern of mean hourly wages overtime displayed in 12 facets coloured by gender.
+We took 240 samples of females and males equally and spread them into twelve facets, as shown in Figure \ref{fig:plot-gender} using `facet_strata` function from `brolgar` [@brolgar].
+We learn that the mean hourly wage of females and males fluctuated over the years.
+For example,  in facets 4, 5, and 10, males tend to earn more wages than females.
+This finding is also shown in Figure \ref{fig:se-gender} where we plotted the median wage of males and females against the year.
+We learn that males had a higher mean of mean hourly wage throughout the survey than females.
+Unfortunately, the gap between them tends to be wider from time to time.
+
+```{r plot-gender, echo = FALSE, warning = FALSE, fig.cap= "The pattern of mean hourly wages overtime displayed in 12 facets coloured by gender.
 The respondents belong to high school cohort are randomly sampled.
@@
 
-```{r gender-se, echo=FALSE, fig.cap="The yearly mean of mean hourly wage by gender along with its standard error.
+```{r se-gender, echo=FALSE, fig.cap="The yearly mean of mean hourly wage by gender along with its standard error.
 Males generally earn more wages than females.
@@
 
-For the final example, we show the example of modeling the longitudinal data using Generelized Additive Model (GAM) with `mgcv` package [@mgcv] for each ID.
-We fit a more flexible model to fit in to the data since the relationship between mean hourly wage and year is not linear due to the spikes.
-Since this approach is computationally expensive, we only show the model in a little fraction of the data.
-We sample 1 percent of the total respondents using `brolgar's` `sample_frac_keys` function.
-Further, we plotted the some sample of the fitted model in Figure \@ref{fig:gam-plot}.
+For the final example, we showed modeling the longitudinal data using the Generalized Additive Model (GAM) with `mgcv` package [@mgcv] for each ID.
+We fitted a more flexible model to fit into the data since the relationship between mean hourly wage and year is not linear due to the spikes.
+Since this approach is computationally expensive, we only showed the model in a small fraction of the data.
+We sampled 1 percent of the total respondents using `brolgar's` `sample_frac_keys` function.
+Further, we plotted some IDs overlayed by the fitted GAM model in Figure \ref{fig:plot-gam}.
 We learn that the model flexible enough to deal with the fluctuation in the data.
-Moreover, model with small spikes tend to be linearly fitted.
-
-```{r gam-plot, cache=TRUE, echo=FALSE, fig.cap="Exploration of wages data by fitting a GAM.
+Moreover, the data with small spikes tend to be linearly fitted.
+
+```{r plot-gam, cache=TRUE, echo=FALSE, fig.cap="Exploration of wages data by fitting a GAM.
 The fitted model displayed by blue line.
@@
 
-This paper has performed a set of stages to make an open data suitable for a text book data or to make it ready for research.
-In the first stage, we showed the steps that are performed to get the data from the NLSY79 database.
-Since the data format is untidy, we showed how the data has been tidied.
-After that, we conducted an initial data analysis to investigate and screen the quality of the data.
-Using robust linear regression model, we found and fixed the anomalous observations in the data set with its predicted values.
+This paper has performed a set of stages to make open data suitable for textbook data or make it ready for research.
+In the first stage, we showed the steps performed to get the data from the NLSY79 database.
+Since the data format is untidy, we showed how the data had been tidied.
+After that, we conducted initial data analysis to investigate and screen the quality of the data.
+Using the robust linear regression model, we found and fixed the anomalous observations in the data set with its predicted values.
 We also performed an example of exploratory data analysis using the cleaned data set.
 
-This paper also has demonstrated the documentation of data cleaning by providing all of the codes in performing data tidying and initial data analysis.
-Thus, this paper provided an opportunity to continuously refresh the text book data whenever the updated data is published in the NLSY79 database.
-This could be done by following the documentation of the code that is provided in this paper.
-
-Moreover, the documentation also includes the process of how did we generate the robustness weight and how did we decide the threshold of anomalous observations.
+This paper has also demonstrated data cleaning documentation by providing all of the codes in performing data tidying and initial data analysis.
+Thus, this paper provided an opportunity to continuously refresh the textbook data whenever the updated data is published in the NLSY79 database.
+It could be done by following the documentation of the code that is provided in this paper.
+
+Moreover, the documentation also includes how we generated the robustness weight and how we decided the threshold of anomalous observations.
 It is also documented along with the flag of whether an observation is imputed value or not.
-Accordingly, if somebody wish to make another decision, it can be done by making a small changes in the code provided.
+Accordingly, if somebody wishes to make another decision, it can be done by making a small change in the code provided.
 Further, this paper is also supplemented by a `shiny`  [@shiny] app as a simulation tool to customize the weight threshold.
 
-Finally, this paper implies that data providers should design the database that is able to produce tidy data sets.
-A data provider should also check for data anomalies prior the data publishing or at least provides a set of rules or threshold value, for example, in this case is the threshold of reasonable wages.
-This will greatly support the data users to carry out further validation and set the same understanding of which data are considered as outliers.
-Moreover, providing validation rules would facilitate the usage of any establish data validation tool, such as `validate` [@validate] package.
+Finally, this paper implies that data providers should design a database that is able to produce tidy data sets.
+A data provider should also check for data anomalies before the data publishing or at least provides a set of rules or threshold values.
+For example, in this case, is the threshold of reasonable wages.
+This will greatly support the data users to validate and set the same understanding of which data are considered outliers.
+Moreover, providing validation rules would facilitate any established data validation tool, such as `validate` [@validate] package.
 In this case, we cannot use this handy validation package due to the absence of validation rules.

### 2021-05-05 report_alt.tex
fixed cross-referencing

@@
 
-\begin{figure}
-\centering
-\includegraphics{figures/featureplot-1.pdf}
-\caption{Two plots showing the distribution of the mean hourly wage.
-Plot A portrays the pattern of mean hourly wage of high school cohort from 1979 to 2018 of each ID in US Dollar; Plot B shows the distribution of their minimum, median, and maximum value.
-We can see that some IDs had an extremely high of wages and it made the distribution of the three features is extremely skewed.}
-\end{figure}
-
 The next step is that we explored the mean hourly wage data, in this case, we only explored the wages data in \texttt{wages\_demog\_hs}.
-Figure \ref{fig:featureplot} conveys that there is clearly a problem in the mean hourly wage values.
-Figure \ref{fig:featureplot} A shows that some observations had an exceptionally high figure of wages, even more than US\$10,000 per hour.
-In Figure \ref{fig:featureplot} B, we barely see any difference in the minimum, median, an maximum value of the wages since the distribution is heavily skewed to the right.
-Additionally, Table \ref{tab:summarytable} shows that the overall wages median of the cohort is only 7.2, while the mean is 11.87.
+Table \ref{tab:summarytable} shows that the overall wages median of the cohort is only 7.2, while the mean is 11.87.
 It indicates that the data might contains a lot of extreme values.
@@
 
+However, we can not be sure only by observing the summary statistics.
+Hence, we used visualisation techniques to investigate this matter.
+Figure \ref{fig:feature-plot} conveys a problem in the mean hourly wage values.
+Figure \ref{fig:feature-plot} A shows that some observations had an exceptionally high figure of wages, even more than US\$10,000 per hour.
+In Figure \ref{fig:feature-plot} B, we barely see any difference in the minimum, median, and maximum value of the wages since the distribution is heavily skewed to the right.
+
+\begin{figure}
+\centering
+\includegraphics{figures/feature-plot-1.pdf}
+\caption{\label{fig:feature-plot}Two plots showing the distribution of the mean hourly wage.
+Plot A portrays the pattern of mean hourly wage of high school cohort from 1979 to 2018 of each ID in US Dollar; Plot B shows the distribution of their minimum, median, and maximum value.
+We can see that some IDs had an extremely high of wages and it made the distribution of the three features is extremely skewed.}
+\end{figure}
+
 In Figure \ref{fig:high-wages}, we plotted some respondents with a high value of mean hourly wages.
-We filtered all of the IDs who earned more than US\$ 500 per hour averagely.
-We found that most of these respondents only experienced one point of extremely high wages.
-Thus, we suspected that these high values are erroneous values resulted from a data entry error.
+We filtered all of the IDs who earned more than US\$ 500 per hour on average.
+We found that these respondents only experienced one point of extremely high wages.
+Thus, we suspected that these high values are erroneous resulted from a data entry error.
 
@@
 
-Further, we took 36 samples randomly from the data and plotted it as is seen in Figure \ref{fig:sampleplot}.
-It implies that not only that some observations earned an extremely high figures of wages, but some also had a reasonably fluctuate wages, for example the IDs in panel number 5, 7, and 11.
+Further, we took 36 samples randomly from the data and plotted them, as shown in Figure \ref{fig:sample-plot}.
+It implies that not only that some observations earned extremely high figures of wages, but some also had reasonably fluctuated wages, for example, the IDs in panel numbers 5, 7, and 11.
 The plot also implies that the samples had a different pattern of mean hourly wages.
-Some had a flat wages for years but had a sudden increase in a year than it went down again, while the other experienced a upsurge in their wage, for instance the IDs in panel 9.
-
-According to Pergamit et al.
-(2001), one of the flaws of the NLSY79 employment data is that since the NLSY79 collect the information of the working hours since the last interview, it might be challenging for the respondents to track the within-job hours changes that happens between survey year, especially for the respondents with fluctuate working hours or whose job is seasonal.
-It even has been more challenging since 1994, where the respondents had to recall two years period.
-This shortcoming might also contribute on the fluctuation of one's wages data.
+Some had flat wages for years but had a sudden increase in one particular year, then it went down again, while the others experienced an upsurge in their wage, for instance, the IDs in panel 9.
 
@@
 \centering
-\includegraphics{figures/sampleplot-1.pdf}
-\caption{The mean hourly wages of some random samples are shown in twelve facets, three IDs per facet. It suggests that some IDs had a reasonably fluctuate wages.}
+\includegraphics{figures/sample-plot-1.pdf}
+\caption{\label{fig:sample-plot}The mean hourly wages of some random samples are shown in twelve facets, three IDs per facet. It suggests that some IDs had a reasonably fluctuate wages.}
 \end{figure}
+
+According to Pergamit et al.
+(2001), one of the flaws of the NLSY79 employment data is that since the NLSY79 collect the information of the working hours since the last interview, it might be challenging for the respondents to track the within-job hours' changes that happens between survey year, especially for the respondents with fluctuate working hours or whose job is seasonal.
+It even has been more challenging since 1994, where the respondents had to recall two years period.
+This shortcoming might also contribute to the fluctuation of one's wages data.
 
@@
 
-As it is seen from figure \ref{fig:sampleplot}, there are many spikes in the mean hourly wage data.
-As part of the IDA, which is the model formulation, we built a robust linear regression model to address this issue.
-The notion of robust linear regression is to yield an estimation that is robust to the influence of noise or contamination (Koller 2016).
-It also aims to detect the contamination by weighting each observation based on how ``well-behave'' they are, known as robustness weight.
-Observations with lower robustness weight are suggested as an outliers by this method (Koller 2016).
+As part of the IDA, which is the model formulation, we built a robust linear regression model to handle the extreme values in the data.
+Robust linear regression yields an estimation that is robust to the influence of noise or contamination (Koller 2016).
+It also aims to detect the contamination by weighting each observation based on how ``well-behaved'' they are, known as robustness weight.
+Observations with lower robustness weight are suggested as an outlier by this method (Koller 2016).
 
-In this paper, we built the model using the \texttt{rlm} function from \texttt{MASS} package (Venables and Ripley 2002).
-We set the \texttt{mean\_hourly\_wage} and \texttt{year} as the dependent and predictor respectively.
-Furthermore, we used M-Estimation with Huber weighting where the observation with small residual get a weight of 1, while the larger the residual, the smaller the weight (less than 1) (UCLA: Statistical Consulting Group 2021).
-
-Since we worked with longitudinal data, we should built the model for each ID, instead of the overall data.
-The robust mixed model is actually the best model to be employed in this case.
+Since we worked with longitudinal data, we built the model for each ID instead of the overall data.
+The robust mixed model is might be the best model to be employed in this case.
 However, this method is too computationally and memory expensive, especially for a large data set, like the NLSY79 data.
@@
 
-The challenging part of detecting the anomaly using the robustness weight is to determine the threshold of the weight where the observations considered as outliers.
-Moreover, it should be noted that not all the outliers is due to an error, instead it might be that one had a reasonably increasing or decreasing wages.
-To, minimize the risk of being mistakenly regard an outlier as an ``erronous outlier'', we have simulated some threshold and study the behavior of the spikes in each threshold.
-We found that 0.12 is the most reasonable value to be the threshold to minimize the risk of that drawback because it still capture the sensible spikes in the data.
-In other words, we kept maintaining the natural variability of the wages while minimizing the presence of anomalies because of the error in the data recording.
-After deciding the threshold, we imputed the observations with weight less than 0.2 wage with the models' predicted value.
-We then flagged those observations in a new variable called \texttt{is\_pred},
+We built the model using the \texttt{rlm} function from \texttt{MASS} package (Venables and Ripley 2002).
+We set the \texttt{mean\_hourly\_wage} and \texttt{year} as the dependent and predictor, respectively.
+Furthermore, we used M-Estimation with Huber weighting where the observation with a slight residual gets a weight of 1, while the larger the residual, the smaller the weight (less than 1) (UCLA: Statistical Consulting Group 2021).
+However, the challenging part of detecting the anomaly using the robustness weight is determining the threshold of the weight in which the observations are considered outliers.
+Moreover, it should be noted that not all the outliers are due to an error.
+Instead, it might be that one had reasonably increasing or decreasing wages in a particular period.
 
-Figure \ref{fig:comppict} A shows that after imputing the anomalies with the models' predicted value, the highest wages value has decreased to be around \$350.
-The spikes were still observed, but are not as extreme as the original data set.
-In Figure \ref{fig:comppict} B), we plotted the three features of mean hourly wages, namely the minimum, median, an maximum value, transformed to log scale.
-The plot implies that the skewness are slightly negative.
+To minimize the risk of mistakenly regarding an outlier as an ``erroneous outlier'', we have simulated some thresholds and study how they affected the data.
+We found that 0.12 is the most reasonable value to be the threshold to minimize that drawback's risk because it still captures the sensible spikes in the data.
+In other words, we kept maintaining the natural variability of the wages while minimizing anomalies because of the error in the data recording.
+After deciding the threshold, we imputed the observations whose weight less than 0.12 with the models' predicted value.
+We then flagged those observations in a new variable called \texttt{is\_pred}.
+
+Figure \ref{fig:comppict} A shows the mean hourly wage feature after the extreme values are replaced.
+It shows that the highest value has decreased to be around \$350.
+The spikes were still observed but are not as extreme as the original data set.
+In Figure \ref{fig:comppict} B), we plotted the three features of mean hourly wages, namely the minimum, median, and maximum value, transformed to log scale.
+The plot implies that the skewness is slightly negative.
 We also see that the three features are overlapped each other.
-It indicates that some ID's minimum wage are higher than some ID's maximum wage.
-
-Furthermore, Figure \ref{fig:compare} implies that after the treatment, the fluctuation can still be observed in the data and only the large spikes, which are considered as ``erronous outliers'', are eliminated from the data.
-Hence, the model results a data set with the reasonable degree of fluctuation.
+It indicates that some ID's minimum wages are higher than some ID's maximum wages.
 
@@
 \includegraphics{figures/comppict-1.pdf}
-\caption{The distribution of the mean hourly wage after treating the extreme values.
+\caption{\label{fig:comppict}The distribution of the mean hourly wage after the extreme values are replaced.
 Plot A portrays the pattern of mean hourly wage of high school cohort from 1979 to 2018 of each ID in US Dollar; Plot B shows the distribution of their minimum, median, and maximum value transformed to log10 scale.
@@
 
+Furthermore, Figure \ref{fig:compare-plot} implies that the fluctuation can still be observed in the data after the treatment.
+However, the large spikes, which are considered ``erroneous outliers'', are already eliminated from the data.
+Hence, the model produces a data set with a more reasonable degree of fluctuation.
+
 \begin{figure}
 \centering
-\includegraphics{figures/compare-1.pdf}
-\caption{Comparison between the original and the treated mean hourly wage.
+\includegraphics{figures/compare-plot-1.pdf}
+\caption{\label{fig:compare-plot}Comparison between the original and the treated mean hourly wage.
 The orange line portray the original value of mean hourly wage, while the turquoise line display the mean hourly wages value after the extreme values imputed with the robust linear model's prediction value.
@@
 Finally, we saved the imputed data and set the appropriate data type for the variables.
-As our target data is the mean hourly wage of the high-school dropouts, we then subset the high-school graduate data set to be only having the male-high school dropout data.
+As our target data is the mean hourly wage of the high-school dropouts, we then subset the high-school graduate data set to have only the male-high school dropout data.
 We also saved the NLSY79 cohort's demographic information in a separate dataset.
-We then make these three data sets and its processing documentation to be publicly available through an R data container package called \texttt{yowie}.
-The complete flow from the raw data to these data set is displayed in Figure \ref(fig:flowchart).
+We then make these three data sets and their processing documentation publicly available through an R data container package called \texttt{yowie}.
+The complete flow from the raw data to these data set is displayed in Figure \ref{fig:flow-chart}.
 
@@
 \centering
-\includegraphics{figures/flowchart-1.pdf}
-\caption{The stages of data filtering from the raw data to get three datasets that are contained in an R package called yowie, n means the number of ID, while n\_obs means the number of observations.}
+\includegraphics{figures/flow-chart-1.pdf}
+\caption{\label{fig:flow-chart}The stages of data filtering from the raw data to get three datasets that are contained in an R package called yowie, n means the number of ID, while n\_obs means the number of observations.}
 \end{figure}
@@
 
-In this part, we gave some examples of how the data might be used for data analysis in text book.
+This part gives some examples of how the data might be used for data analysis in a textbook.
 We used \texttt{brolgar} (Tierney, Cook, and Prvan 2020) to perform exploratory data analysis.
-The first example is that we observed relationship between wages and education in terms of how the wages increase year by year.
-Using \texttt{key\_slope} function in \texttt{brolgar}, we performed a liner regression with mean hourly wage as response variable and year as predictor and extracted the regression slope of each ID and plotted it according to highest grade completed.
+In the first example, we observed the relationship between wages and education regarding how the wages increase year by year.
+Using \texttt{key\_slope} function in \texttt{brolgar}, we performed linear regression with the mean hourly wage as response variable and year as the predictor, extracted the regression slope of each ID, and plotted it according to the highest grade completed.
 
-We found in Figure \ref{fig:wages-slope} that the higher the education completed does not necessarily increase the wages more from year to year.
-We can see that people who completed 4th grade happen to have relatively the same median of slope as people who completed 11th grade, even though it is probably because some respondents in this group had an extremely high mean hourly wages.
-Moreover, the annual increasing wages of people who completed 6th grade to 11th grade are relatively same.
-People who completed 12th grade tend to have higher slope of year compared to the other groups.
+Figure \ref{fig:slope-fig} shows that the higher the education completed does not necessarily increase the wages more from year to year.
+We can see that people who completed 4th grade happen to have relatively the same median of slope as people who completed 11th grade, even though it is probably because some respondents in this group had extremely high mean hourly wages.
+Moreover, the annual increasing wages of people who completed 6th grade to 11th grade are relatively the same.
+People who completed 12th grade tend to have a higher slope of the year than the other groups.
 Some outliers are also spotted in this group.
@@
 \centering
-\includegraphics{figures/wages-slope-1.pdf}
-\caption{Regression slope of mean hourly wage regression by highest grade completed.
+\includegraphics{figures/slope-fig-1.pdf}
+\caption{\label{fig:slope-fig}Regression slope of mean hourly wage regression by highest grade completed.
 The regression model has mean hourly wage as response variable and year as predictor.
@@
 In the second example, we examined the pattern of mean hourly wage corresponding to different groups of gender and race.
-We took 240 samples of female and male equally and spread it into twelve facets as is seen in Figure \ref{fig:gender-plot} using \texttt{facet\_strata} function from \texttt{brolgar} (Tierney, Cook, and Prvan 2020).
-We learn that the mean hourly wage of female and male fluctuated over the years.
-In some facets, for example in facet 4, 5, and 10, males tend to earn more wages than females.
-This finding is also shown in Figure \ref{fig:gender-se} where we plotted the median wage of males and females against the year.
-We learn that males had a higher mean of mean hourly wage over the period of survey than females.
-Unfortunately, the gap between them tend to be wider from time to time.
+We took 240 samples of females and males equally and spread them into twelve facets, as shown in Figure \ref{fig:plot-gender} using \texttt{facet\_strata} function from \texttt{brolgar} (Tierney, Cook, and Prvan 2020).
+We learn that the mean hourly wage of females and males fluctuated over the years.
+For example, in facets 4, 5, and 10, males tend to earn more wages than females.
+This finding is also shown in Figure \ref{fig:se-gender} where we plotted the median wage of males and females against the year.
+We learn that males had a higher mean of mean hourly wage throughout the survey than females.
+Unfortunately, the gap between them tends to be wider from time to time.
 
@@
 \centering
-\includegraphics{figures/gender-plot-1.pdf}
-\caption{The pattern of mean hourly wages overtime displayed in 12 facets coloured by gender.
+\includegraphics{figures/plot-gender-1.pdf}
+\caption{\label{fig:plot-gender}The pattern of mean hourly wages overtime displayed in 12 facets coloured by gender.
 The respondents belong to high school cohort are randomly sampled.
@@
 \centering
-\includegraphics{figures/gender-se-1.pdf}
-\caption{The yearly mean of mean hourly wage by gender along with its standard error. Males generally earn more wages than females. The error interval tend to be wider over the years.}
+\includegraphics{figures/se-gender-1.pdf}
+\caption{\label{fig:se-gender}The yearly mean of mean hourly wage by gender along with its standard error.
+Males generally earn more wages than females.
+The error interval tend to be wider over the years.}
 \end{figure}
 
-For the final example, we show the example of modeling the longitudinal data using Generelized Additive Model (GAM) with \texttt{mgcv} package (Wood 2003) for each ID.
-We fit a more flexible model to fit in to the data since the relationship between mean hourly wage and year is not linear due to the spikes.
-Since this approach is computationally expensive, we only show the model in a little fraction of the data.
-We sample 1 percent of the total respondents using \texttt{brolgar\textquotesingle{}s} \texttt{sample\_frac\_keys} function.
-Further, we plotted the some sample of the fitted model in Figure @ref\{fig:gam-plot\}.
+For the final example, we showed modeling the longitudinal data using the Generalized Additive Model (GAM) with \texttt{mgcv} package (Wood 2003) for each ID.
+We fitted a more flexible model to fit into the data since the relationship between mean hourly wage and year is not linear due to the spikes.
+Since this approach is computationally expensive, we only showed the model in a small fraction of the data.
+We sampled 1 percent of the total respondents using \texttt{brolgar\textquotesingle{}s} \texttt{sample\_frac\_keys} function.
+Further, we plotted some IDs overlayed by the fitted GAM model in Figure \ref{fig:plot-gam}.
 We learn that the model flexible enough to deal with the fluctuation in the data.
-Moreover, model with small spikes tend to be linearly fitted.
+Moreover, the data with small spikes tend to be linearly fitted.
 
@@
 \centering
-\includegraphics{figures/gam-plot-1.pdf}
-\caption{Exploration of wages data by fitting a GAM. The fitted model displayed by blue line. The fitted line shows that the model flexible enough to follow the pattern of the data.}
+\includegraphics{figures/plot-gam-1.pdf}
+\caption{\label{fig:plot-gam}Exploration of wages data by fitting a GAM.
+The fitted model displayed by blue line.
+The fitted line shows that the model flexible enough to follow the pattern of the data.}
 \end{figure}
@@
 
-This paper has performed a set of stages to make an open data suitable for a text book data or to make it ready for research.
-In the first stage, we showed the steps that are performed to get the data from the NLSY79 database.
-Since the data format is untidy, we showed how the data has been tidied.
-After that, we conducted an initial data analysis to investigate and screen the quality of the data.
-Using robust linear regression model, we found and fixed the anomalous observations in the data set with its predicted values.
+This paper has performed a set of stages to make open data suitable for textbook data or make it ready for research.
+In the first stage, we showed the steps performed to get the data from the NLSY79 database.
+Since the data format is untidy, we showed how the data had been tidied.
+After that, we conducted initial data analysis to investigate and screen the quality of the data.
+Using the robust linear regression model, we found and fixed the anomalous observations in the data set with its predicted values.
 We also performed an example of exploratory data analysis using the cleaned data set.
 
-This paper also has demonstrated the documentation of data cleaning by providing all of the codes in performing data tidying and initial data analysis.
-Thus, this paper provided an opportunity to continuously refresh the text book data whenever the updated data is published in the NLSY79 database.
-This could be done by following the documentation of the code that is provided in this paper.
+This paper has also demonstrated data cleaning documentation by providing all of the codes in performing data tidying and initial data analysis.
+Thus, this paper provided an opportunity to continuously refresh the textbook data whenever the updated data is published in the NLSY79 database.
+It could be done by following the documentation of the code that is provided in this paper.
 
-Moreover, the documentation also includes the process of how did we generate the robustness weight and how did we decide the threshold of anomalous observations.
+Moreover, the documentation also includes how we generated the robustness weight and how we decided the threshold of anomalous observations.
 It is also documented along with the flag of whether an observation is imputed value or not.
-Accordingly, if somebody wish to make another decision, it can be done by making a small changes in the code provided.
+Accordingly, if somebody wishes to make another decision, it can be done by making a small change in the code provided.
 Further, this paper is also supplemented by a \texttt{shiny} (Chang et al.
@@
 
-Finally, this paper implies that data providers should design the database that is able to produce tidy data sets.
-A data provider should also check for data anomalies prior the data publishing or at least provides a set of rules or threshold value, for example, in this case is the threshold of reasonable wages.
-This will greatly support the data users to carry out further validation and set the same understanding of which data are considered as outliers.
-Moreover, providing validation rules would facilitate the usage of any establish data validation tool, such as \texttt{validate} (van der Loo and de Jonge 2021) package.
+Finally, this paper implies that data providers should design a database that is able to produce tidy data sets.
+A data provider should also check for data anomalies before the data publishing or at least provides a set of rules or threshold values.
+For example, in this case, is the threshold of reasonable wages.
+This will greatly support the data users to validate and set the same understanding of which data are considered outliers.
+Moreover, providing validation rules would facilitate any established data validation tool, such as \texttt{validate} (van der Loo and de Jonge 2021) package.
 In this case, we cannot use this handy validation package due to the absence of validation rules.

### 2021-05-05 report_alt.Rmd
Change the sequence of the plot in 3.2.1

@@
 
-Figure \ref{fig:comppict} A shows the mean hourly wage feature after the extreme values are replaced.
-It shows that the highest value has decreased to be around $350.
-The spikes were still observed but are not as extreme as the original data set.
-In Figure \ref{fig:comppict} B), we plotted the three features of mean hourly wages, namely the minimum, median, and maximum value, transformed to log scale.
-The plot implies that the skewness is slightly negative.
-We also see that the three features are overlapped each other.
-It indicates that some ID's minimum wages are higher than some ID's maximum wages.
-
-```{r comppict, echo = FALSE, fig.cap = "The distribution of the mean hourly wage after the extreme values are replaced.
-Plot A portrays the pattern of mean hourly wage of high school cohort from 1979 to 2018 of each ID in US Dollar; Plot B shows the distribution of their minimum, median, and maximum value transformed to log10 scale.
-We can see that some observations still had reasonbaly higher wages than the others.
-Also, some IDs' have a minimum wages that is higher than others' maximum wages.", fig.width=6, fig.height=4}
-
-spag2 <- wages_demog_hs %>%
-  ggplot(aes(x = year,
-             y = wages_rlm,
-             group = id)) +
-  geom_line(alpha = 0.1) +
-  ggtitle("A)") +
-  theme(plot.title = element_text(size = 10))
-
-wages_hs2020_rlm <- as_tsibble(x = wages_demog_hs,
-                    key = id,
-                    index = year,
-                    regular = FALSE)
-wages_three_feat_rlm <- wages_hs2020_rlm %>%
-  features(wages_rlm, 
-           feat_three_num
-           )
-wages_feat_long_rlm <- wages_three_feat_rlm %>%
-  pivot_longer(c(min, med, max), names_to = "feature", values_to = "value")
-
-feature2 <- ggplot(wages_feat_long_rlm) +
-  geom_density(aes(x = value, colour = feature, fill = feature), alpha = 0.3) +
-  ggtitle("B)") +
-  scale_x_log10() + 
-  theme(plot.title = element_text(size = 10)) +
-  xlab("value (log10)")
-
-spag2 + feature2
-```
+Figure \ref{fig:compare-plot} shows the mean hourly wage before and after the extreme values are replaced.
+It implies that the fluctuation can still be observed in the data after the treatment.
+However, the large spikes, which are considered "erroneous outliers", are already eliminated from the data.
+Hence, the model produces a data set with a more reasonable degree of fluctuation.
 
@@
 ```
-
-Furthermore, Figure \ref{fig:compare-plot} implies that the fluctuation can still be observed in the data after the treatment.
-However, the large spikes, which are considered "erroneous outliers", are already eliminated from the data.
-Hence, the model produces a data set with a more reasonable degree of fluctuation.
 
@@
 
+Further, \ref{fig:comppict} A shows that after eliminating the extreme values, the highest value has decreased to be around $350.
+The spikes were still observed but are not as extreme as the original data set.
+In Figure \ref{fig:comppict} B), we plotted the three features of mean hourly wages, namely the minimum, median, and maximum value, transformed to log scale.
+The plot implies that the skewness is slightly negative.
+We also see that the three features are overlapped each other.
+It indicates that some ID's minimum wages are higher than some ID's maximum wages.
+
+```{r comppict, echo = FALSE, fig.cap = "The distribution of the mean hourly wage after the extreme values are replaced.
+Plot A portrays the pattern of mean hourly wage of high school cohort from 1979 to 2018 of each ID in US Dollar; Plot B shows the distribution of their minimum, median, and maximum value transformed to log10 scale.
+We can see that some observations still had reasonbaly higher wages than the others.
+Also, some IDs' have a minimum wages that is higher than others' maximum wages.", fig.width=6, fig.height=4}
+
+spag2 <- wages_demog_hs %>%
+  ggplot(aes(x = year,
+             y = wages_rlm,
+             group = id)) +
+  geom_line(alpha = 0.1) +
+  ggtitle("A)") +
+  theme(plot.title = element_text(size = 10))
+
+wages_hs2020_rlm <- as_tsibble(x = wages_demog_hs,
+                    key = id,
+                    index = year,
+                    regular = FALSE)
+wages_three_feat_rlm <- wages_hs2020_rlm %>%
+  features(wages_rlm, 
+           feat_three_num
+           )
+wages_feat_long_rlm <- wages_three_feat_rlm %>%
+  pivot_longer(c(min, med, max), names_to = "feature", values_to = "value")
+
+feature2 <- ggplot(wages_feat_long_rlm) +
+  geom_density(aes(x = value, colour = feature, fill = feature), alpha = 0.3) +
+  ggtitle("B)") +
+  scale_x_log10() + 
+  theme(plot.title = element_text(size = 10)) +
+  xlab("value (log10)")
+
+spag2 + feature2
+```
+
 Finally, we saved the imputed data and set the appropriate data type for the variables.

### 2021-05-05 report_alt.tex
Change the sequence of the plot in 3.2.1

@@
 
-Figure \ref{fig:comppict} A shows the mean hourly wage feature after the extreme values are replaced.
-It shows that the highest value has decreased to be around \$350.
+Figure \ref{fig:compare-plot} shows the mean hourly wage before and after the extreme values are replaced.
+It implies that the fluctuation can still be observed in the data after the treatment.
+However, the large spikes, which are considered ``erroneous outliers'', are already eliminated from the data.
+Hence, the model produces a data set with a more reasonable degree of fluctuation.
+
+\begin{figure}
+\centering
+\includegraphics{figures/compare-plot-1.pdf}
+\caption{\label{fig:compare-plot}Comparison between the original and the treated mean hourly wage.
+The orange line portray the original value of mean hourly wage, while the turquoise line display the mean hourly wages value after the extreme values imputed with the robust linear model's prediction value.
+We can see that some extreme spikes has been reduced by the model.}
+\end{figure}
+
+Further, \ref{fig:comppict} A shows that after eliminating the extreme values, the highest value has decreased to be around \$350.
 The spikes were still observed but are not as extreme as the original data set.
@@
 Also, some IDs' have a minimum wages that is higher than others' maximum wages.}
-\end{figure}
-
-Furthermore, Figure \ref{fig:compare-plot} implies that the fluctuation can still be observed in the data after the treatment.
-However, the large spikes, which are considered ``erroneous outliers'', are already eliminated from the data.
-Hence, the model produces a data set with a more reasonable degree of fluctuation.
-
-\begin{figure}
-\centering
-\includegraphics{figures/compare-plot-1.pdf}
-\caption{\label{fig:compare-plot}Comparison between the original and the treated mean hourly wage.
-The orange line portray the original value of mean hourly wage, while the turquoise line display the mean hourly wages value after the extreme values imputed with the robust linear model's prediction value.
-We can see that some extreme spikes has been reduced by the model.}
 \end{figure}

### 2021-08-12 Report/report_alt.Rmd
Calculate gender wage gap and fit linear mixed model.

--- 2021-05-05 report_alt.Rmd
@@
                       message = FALSE,
-                      cache = TRUE, 
-                      cache.path = "cache/",
+                      #cache = TRUE, 
+                      #cache.path = "cache/",
                       fig.path = "figures/")
@@
 library(tsibble)
+library(lme4)
+library(modelr)
 ```
@@
 
-```{r rlm, cache=TRUE, warning = FALSE, message = FALSE, echo = FALSE}
+```{r rlm, eval=FALSE, warning = FALSE, message = FALSE, echo = FALSE}
 
@@
 
-# if the weight < 1, the mean_hourly_wage is replaced by the model's fitted/predicted value.
+# if the weight < 0.12, the mean_hourly_wage is replaced by the model's fitted/predicted value.
 # and add the flag whether the observation is predicted value or not.
 # since the fitted value is sometimes <0, and wages value could never be negative,
-# we keep the mean hourly wage value even its weight < 1.
+# we keep the mean hourly wage value even its weight < 0.12.
 
@@
 
+# save it to rds file so it would faster the knitting process
+saveRDS(wages_demog_hs, here::here("Report/Result/wages_demog_hs_tmp.rds"))
 ```
@@
 
+wages_demog_hs <- readRDS(here::here("Report/Result/wages_demog_hs_tmp.rds"))
 sample_id <- sample(unique(wages_demog_hs$id), 20)
@@
 For example,  in facets 4, 5, and 10, males tend to earn more wages than females.
-This finding is also shown in Figure \ref{fig:se-gender} where we plotted the median wage of males and females against the year.
-We learn that males had a higher mean of mean hourly wage throughout the survey than females.
-Unfortunately, the gap between them tends to be wider from time to time.
 
@@
 
-```{r se-gender, echo=FALSE, fig.cap="The yearly mean of mean hourly wage by gender along with its standard error.
-Males generally earn more wages than females.
-The error interval tend to be wider over the years."}
+To further investigate this, we calculated the median wage of females relative to males, which is called gender wage gap.
+According to @OECD, gender pay gap is the difference between median wage of males and females relative to median wage of males.
+We find that, as shown in \ref{fig:se-gender}, over the years, females tend to earn fewer than males.
+For example, in 2018 females earn about 78 cents for every $1 earned by males.
+The widest gender gaps were observed in 1990s.
+
+```{r se-gender, echo=FALSE, fig.cap="Gender wage gap or the median wages of females relatives to median wages of males.
+Females tend to earn fewer than males.
+For example, in 2018 females earn about 78 cents for every $1 earned by males."}
+
 wages_hs2020_grouped_gender <- wages_hs2020 %>%
@@
   summarise(mean = mean(mean_hourly_wage),
+            median = median(mean_hourly_wage),
              sd = sd(mean_hourly_wage),
@@
 
-gender_plot <- ggplot(wages_hs2020_grouped_gender,
-                      aes(x = year,
-                          y = mean,
-                          group = gender,
-                          colour = gender)) +
-  geom_point(size = 1) +
-  geom_errorbar(aes(ymin=lower, ymax=upper), width=1,
-                 position=position_dodge(0.1)) +
-  theme(legend.position = "bottom",
-         legend.text = element_text(size = 5),
-         axis.title = element_text(size = 9)) +
-  scale_color_brewer(palette = "Dark2") +
-  ylab("median of mean hourly wage (US$)")
-
-gender_plot
-
-```
-
-For the final example, we showed modeling the longitudinal data using the Generalized Additive Model (GAM) with `mgcv` package [@mgcv] for each ID.
+wage_female <- wages_hs2020_grouped_gender %>%
+  filter(gender == "FEMALE") %>%
+  dplyr::select(year, median) %>%
+  rename(female = median)
+
+wage_male <- wages_hs2020_grouped_gender %>%
+  filter(gender == "MALE") %>%
+  dplyr::select(median) %>%
+  rename(male = median)
+
+gender_wage_gap <- cbind(wage_female, wage_male) %>%
+  mutate(gap = (male-female)/male) %>%
+  mutate(relative = 1-gap)
+
+ggplot(gender_wage_gap) +
+  geom_line(aes(x = year,
+                y = relative),
+            color = "#DF7401", size = 0.8) +
+  geom_point(aes(x = year,
+                y = relative),
+             color = "#DF7401", size = 2) +
+  geom_hline(yintercept = 1, colour = "#088A68", size = 1) +
+  scale_x_continuous(limits = c(1978.7,2018.3), expand = c(0, 0)) +
+  ylab("females' median wages relative to males (US$)") +
+  annotate("text", x = 2014.5, y = 1, label = "$1 of male wages", vjust = 1.4) +
+  annotate("text", x = 2012, y = 0.8, label = "female wages relative to male", vjust = 1)
+```
+
+For the final example, we showed modeling the longitudinal data using linear mixed model with `lme4` package [@lme4].
+
 We fitted a more flexible model to fit into the data since the relationship between mean hourly wage and year is not linear due to the spikes.
@@
 Further, we plotted some IDs overlayed by the fitted GAM model in Figure \ref{fig:plot-gam}.
-We learn that the model flexible enough to deal with the fluctuation in the data.
+We learn that the model is flexible enough to deal with the fluctuation in the data.
 Moreover, the data with small spikes tend to be linearly fitted.
@@
 
-# take a small fraction of key
-set.seed(2020)
-wages_sample_gam <- sample_frac_keys(wages_hs2020_tsibble, size = 0.01) %>%
+# transform the data type and year for modeling
+wages_model <- wages_hs2020_tsibble %>%
   mutate(id_fct = as.factor(id),
-         year0 = year - 1979)
+         year0 = year - 1979,
+         ln_wages = log(mean_hourly_wage))
 
 # build GAM model
-wages_gam <- gam(
-  mean_hourly_wage ~ s(year0, by = id_fct) + id_fct,
-  data = wages_sample_gam,
-  method = "REML"
+wages_fit <- lmer(
+  ln_wages ~ year0 + hgc_i + gender + (1|id),
+  data = wages_model
 )
 
-wages_aug_gam <- wages_sample_gam %>%
-  add_predictions(wages_gam, var = "pred") %>%
-  add_residuals(wages_gam, var = "res") %>%
-  group_by_key() %>%
-  mutate(rss = sum(res^2)) %>%
-  ungroup()
-
-set.seed(2021)
-wages_aug_gam %>%
-  sample_n_keys(12) %>%
-  ggplot(aes(x = year,
+wages_fit_aug<- wages_model %>%
+  add_predictions(wages_fit, var = "pred") %>%
+  add_residuals(wages_fit, var = "resid")
+
+slope_only <- wages_slope %>%
+  dplyr::select(id, .slope_year) %>%
+  arrange(desc(.slope_year)) %>%
+  slice_head(n = 12)
+
+wages_fit_aug2 <- wages_fit_aug %>%
+  filter(id %in% slope_only$id)
+
+ggplot(wages_fit_aug2, aes(x = year,
              y = pred,
-             group = id)) +
-  geom_line(color = "steelblue") +
-  geom_point(aes(y = mean_hourly_wage)) +
-  facet_wrap(~id)
+             group = id,
+             colour = factor(id))) + 
+  geom_line() + 
+  geom_point(aes(x = year,
+                 y = ln_wages,
+                 colour = factor(id))) + 
+  facet_wrap(~id) + 
+  theme(legend.position = "none")
 

### 2021-08-12 Report/report-alt.R

@@ -1,0 +1,693 @@
+## ----setup, include = FALSE---------------------------------------------------
+knitr::opts_chunk$set(echo = TRUE, 
+                      warning = FALSE, 
+                      message = FALSE,
+                      cache = TRUE, 
+                      cache.path = "cache/",
+                      fig.path = "figures/")
+library(tidyverse)
+library(brolgar)
+library(patchwork)
+library(kableExtra)
+library(MASS)
+library(janitor)
+library(DiagrammeR)
+library(rsvg)
+library(webshot)
+library(mgcv)
+library(modelr)
+library(tsibble)
+
+
+## ---- cache = TRUE, echo = FALSE----------------------------------------------
+source(here::here("data-raw/NLSY79/NLSY79.R"))
+
+
+## ----untidy-data, echo = FALSE------------------------------------------------
+
+untidy_demo <- new_data_qnames %>%
+  as_tibble() %>%
+  # in 2018, the variable's name is Q3-4_2018, instead of HGC_2018
+  rename(HGC_2018 = `Q3-4_2018`) %>%
+  dplyr::select(CASEID_1979,
+            starts_with("HRP") &
+              ends_with(c("1979", "1980", "1981", "1982", "1983"))) %>%
+  dplyr::select(1:7) %>%
+  head()
+
+kable(untidy_demo,
+      caption = "The Untidy Form of the NLSY79 Raw Data") %>%
+  kable_styling(latex_options = "striped")
+
+
+## ----dob-tidy-----------------------------------------------------------------
+## tidy the date of birth data
+dob_tidy <- new_data_qnames %>%
+  dplyr::select(CASEID_1979,
+         starts_with("Q1-3_A~")) %>%
+  mutate(dob_year = case_when(
+                    # if the years recorded in both sets match, take 79 data
+                    `Q1-3_A~Y_1979` == `Q1-3_A~Y_1981` ~ `Q1-3_A~Y_1979`,
+                    # if the year in the 81 set is missing, take the 79 data
+                    is.na(`Q1-3_A~Y_1981`) ~ `Q1-3_A~Y_1979`,
+                    # if the sets don't match for dob year, take the 79 data
+                    `Q1-3_A~Y_1979` != `Q1-3_A~Y_1981` ~ `Q1-3_A~Y_1979`),
+        dob_month = case_when(
[…]

### 2021-08-12 code_di_weighted.R

@@ -1,113 +1,0 @@
-library(tidyverse)
-library(zoo)
-library(MASS)
-library(broom)
-
-
-
-# open the tidy raw data set
-wages <- read_csv("data-raw/wages-high-school-demo/data-frames/tidy_employment_weighted.csv")
-demog <- read_csv("data-raw/wages-high-school-demo/data-frames/tidy_demographics_v2.csv")
-
-
-# join the wages information and the demographic information by case id.
-wages_demog <- left_join(wages, demog, by="id")
-
-
-# calculate the years in work force and the age of the subjects in 1979
-wages_demog <- wages_demog %>%
-  mutate(yr_hgc = as.numeric(yr_hgc)) %>%
-  mutate(years_in_workforce = year - yr_hgc) %>%
-  mutate(age_1979 = 1979 - (dob_year + 1900))
-
-# filter only the id with high school education
-wages_demog_hs <- wages_demog  %>% filter(grepl("GRADE", hgc))
-
-ggplot(wages_demog_hs, aes(x = mean_hourly_wage)) +
-  geom_boxplot()
-
-
-# calculate the number of observation
-keep_me <- wages_demog_hs %>% count(id)
-keep_me %>% ggplot(aes(x=n)) + geom_bar()
-
-# filter out the id with low number of observation
-keep_me <- keep_me %>% filter(n > 4)
-
-wages_demog_hs <- wages_demog_hs %>%
-  filter(id %in% keep_me$id)
-
-# nesting the data by id to build a robust linear model
-by_id <- wages_demog_hs %>%
-  dplyr::select(id, year, mean_hourly_wage) %>%
-  group_by(id) %>%
-  nest()
-
-# build a robust linear model
-id_rlm <- by_id %>%
-  mutate(model = map(.x = data,
-                     .f = function(x){
-                       rlm(mean_hourly_wage ~ year, data = x)
-                     }))
-
-# extract the property of the regression model
-id_aug <- id_rlm %>%
-  mutate(augmented = map(model, broom::augment)) %>%
-  unnest(augmented)
-
-# extract the weight of each observation
-id_w <- id_rlm %>%
-  mutate(w = map(.x = model,
-                 .f = function(x){
-                   x$w
-                 })) %>%
-  unnest(w) %>%
-  dplyr::select(w)
-
-# bind the property of each observation with their weight
-id_aug_w <- cbind(id_aug, id_w) %>%
-  dplyr::select(`id...1`,
-                year,
[…]

### 2021-08-18 Report/report_alt.Rmd
Modify the EDA.

@@
 library(modelr)
+library(forcats)
 ```
@@
 
-wages_slope <- key_slope(wages_hs2020_tsibble, 
-            mean_hourly_wage ~ year) %>%
+# transform the data for modeling purpose
+wages_model_data <- wages_hs2020_tsibble %>%
+  mutate(id_fct = as.factor(id),
+         year0 = year - 1979,
+         ln_wages = log(mean_hourly_wage)) %>%
+  mutate(hgc_regroup = fct_collapse(hgc, `8TH grade or less` = c("UNGRADED","1ST GRADE", "3RD GRADE",
+                                                                 "4TH GRADE", "5TH GRADE",
+                                                                 "6TH GRADE", "7TH GRADE",
+                                                                 "8TH GRADE"),
+                                    `9TH to 11TH grade` = c("9TH GRADE",
+                                                            "10TH GRADE",
+                                                            "11TH GRADE"),
+                                    `12TH grade` = "12TH GRADE"))
+
+# extract the slope for every individual 
+wages_slope <- key_slope(wages_model_data, 
+            ln_wages ~ year0) %>%
   left_join(demographic_nlsy79, by = "id") %>%
-  mutate(hgc = factor(hgc, levels = c("1ST GRADE", "2ND GRADE", "3RD GRADE", 
-                                      "4TH GRADE", "5TH GRADE", "6TH GRADE", 
-                                      "7TH GRADE", "8TH GRADE", "9TH GRADE",
-                                      "10TH GRADE", "11TH GRADE", "12TH GRADE", 
-                                      "UNGRADED"))) %>%
-  filter(hgc != "UNGRADED")
-
-ggplot(wages_slope, aes(y = hgc,
-                      x = .slope_year)) +
-  geom_boxplot()
+  mutate(hgc_regroup = fct_collapse(hgc, `8TH grade or less` = c("UNGRADED","1ST GRADE", "3RD GRADE",
+                                                                 "4TH GRADE", "5TH GRADE",
+                                                                 "6TH GRADE", "7TH GRADE",
+                                                                 "8TH GRADE"),
+                                    `9TH to 11TH grade` = c("9TH GRADE",
+                                                            "10TH GRADE",
+                                                            "11TH GRADE"),
+                                    `12TH grade` = "12TH GRADE"))
+
+ggplot(wages_slope) +
+  geom_boxplot(aes(x = `.slope_year0`, y = hgc_regroup))
+
+```
+```{r}
+ggplot(wages_slope, aes(x = .slope_year0)) +
+  geom_histogram()
+```
+
+```{r}
+slope_bottom_5<- wages_slope %>%
+  arrange(.slope_year0) %>%
+  slice_head(n = 5)
+
+slope_top_5<- wages_slope %>%
+  arrange(desc(.slope_year0)) %>%
+  slice_head(n = 5)
+
+keep_contrast <- rbind(slope_bottom_5, slope_top_5) %>%
+  dplyr::select(.slope_year0, id)
+
+contrast_slope <- left_join(keep_contrast, wages_model_data, by = "id") %>%
+  arrange(.slope_year0) %>%
+  mutate(id = factor(id, levels = c("7918", "11695", "9176", "10119", "6464", "1271", 
+                                    "12371", "9039", "10217", "7782")))
+
+ggplot(contrast_slope, aes(x = year, y = ln_wages, color = gender)) +
+  geom_line() +
+  geom_point() +
+  facet_wrap(~id)
+
+```
+
+```{r}
+ggplot(contrast_slope, aes(x = year, y = ln_wages, color = hgc_regroup)) +
+  geom_line() +
+  geom_point() +
+  facet_wrap(~id)
+```
+
+```{r}
+quartile <- wages_slope %>%
+  mutate(quartile = as.factor(ntile(.slope_year0, 4)))
+
+ggplot(quartile) +
+  geom_bar(aes(x = quartile, fill = gender, position = "fill"))
+```
+
+```{r}
+ggplot(quartile) +
+  geom_bar(aes(x = quartile, fill = hgc_regroup, position = "fill"))
 ```
@@
 
-# transform the data type and year for modeling
-wages_model <- wages_hs2020_tsibble %>%
-  mutate(id_fct = as.factor(id),
-         year0 = year - 1979,
-         ln_wages = log(mean_hourly_wage))
-
-# build GAM model
+# build linear-mixed model
 wages_fit <- lmer(
-  ln_wages ~ year0 + hgc_i + gender + (1|id),
-  data = wages_model
+  ln_wages ~ year0*hgc_regroup*gender + (1|id),
+  data = wages_model_data
 )
 
-wages_fit_aug<- wages_model %>%
+wages_fit_aug<- wages_model_data %>%
   add_predictions(wages_fit, var = "pred") %>%
@@
 
-slope_only <- wages_slope %>%
-  dplyr::select(id, .slope_year) %>%
-  arrange(desc(.slope_year)) %>%
+slope_bottom_12 <- wages_slope %>%
+  dplyr::select(id, .slope_year0) %>%
+  arrange(.slope_year0) %>%
   slice_head(n = 12)
@@
 wages_fit_aug2 <- wages_fit_aug %>%
-  filter(id %in% slope_only$id)
+  filter(id %in% slope_bottom_12$id)
 

### 2021-08-18 Report/report_alt.tex
Calculate gender wage gap and fit linear mixed model.
Modify the EDA.

--- 2021-05-05 report_alt.tex
@@
 For example, in facets 4, 5, and 10, males tend to earn more wages than females.
-This finding is also shown in Figure \ref{fig:se-gender} where we plotted the median wage of males and females against the year.
-We learn that males had a higher mean of mean hourly wage throughout the survey than females.
-Unfortunately, the gap between them tends to be wider from time to time.
 
@@
 
+To further investigate this, we calculated the median wage of females relative to males, which is called gender wage gap.
+According to OECD (n.d.), gender pay gap is the difference between median wage of males and females relative to median wage of males.
+We find that, as shown in \ref{fig:se-gender}, over the years, females tend to earn fewer than males.
+For example, in 2018 females earn about 78 cents for every \$1 earned by males.
+The widest gender gaps were observed in 1990s.
+
 \begin{figure}
@@
 \includegraphics{figures/se-gender-1.pdf}
-\caption{\label{fig:se-gender}The yearly mean of mean hourly wage by gender along with its standard error.
-Males generally earn more wages than females.
-The error interval tend to be wider over the years.}
+\caption{\label{fig:se-gender}Gender wage gap or the median wages of females relatives to median wages of males.
+Females tend to earn fewer than males.
+For example, in 2018 females earn about 78 cents for every \$1 earned by males.}
 \end{figure}
 
-For the final example, we showed modeling the longitudinal data using the Generalized Additive Model (GAM) with \texttt{mgcv} package (Wood 2003) for each ID.
+For the final example, we showed modeling the longitudinal data using linear mixed model with \texttt{lme4} package (Bates et al. 2015).
+
 We fitted a more flexible model to fit into the data since the relationship between mean hourly wage and year is not linear due to the spikes.
@@
 
+\leavevmode\hypertarget{ref-lme4}{}%
+Bates, Douglas, Martin Mächler, Ben Bolker, and Steve Walker.
+2015.
+``Fitting Linear Mixed-Effects Models Using lme4.'' \emph{Journal of Statistical Software} 67 (1): 1--48.
+\url{https://doi.org/10.18637/jss.v067.i01}.
+
 \leavevmode\hypertarget{ref-webshot}{}%
@@
 Koller, Manuel. 2016. ``Robustlmm: An R Package for Robust Estimation of Linear Mixed-Effects Models.'' \emph{Journal of Statistical Software} 75 (6): 1--24.
+
+\leavevmode\hypertarget{ref-OECD}{}%
+OECD. n.d. ``Gender Wage Gap.'' https://data.oecd.org/earnwage/gender-wage-gap.htm.
 

### 2021-08-26 Report/report_alt.Rmd
dis code

@@
 ---
-title: "The Journey from Wild to Text Book Data: A Case Study of the National Longitudinal Survey of Youth"
+title: "The Journey from Wild to Text Book Data: A Case Study from the National Longitudinal Survey of Youth"
 authors:

### 2021-08-26 code_di.R

@@ -111,3 +111,52 @@ wages_demog %>% filter(CASEID_1979 == "6803") %>% print(n=50)
 wages_demog %>% filter(CASEID_1979 == "7980") %>% print(n=50)
 wages_demog %>% filter(CASEID_1979 == "5140") %>% print(n=50)
 
+# RE-DO
+library(tidyverse)
+library(yowie)
+data("wages_hs_dropout")
+glimpse(wages_hs_dropout)
+
+# Number of observations per person
+wages_hs_dropout %>% count(id, sort=TRUE) %>% summary(n)
+wages_hs_dropout %>% count(id, sort=TRUE) %>%
+  ggplot(aes(x=n)) +
+  geom_histogram(breaks = seq(4.5, 28.5, 1), colour="white") +
+  xlab("Number of observations per individual")
+
+# Check values for long individuals
+long <- wages_hs_dropout %>%
+  count(id, sort=TRUE) %>%
+  filter(n == 28)
+wages_hs_dropout %>%
+  filter(id %in% long$id) %>%
+  ggplot(aes(x=year, y=mean_hourly_wage)) +
+    geom_point() +
+    geom_smooth(method="lm", se=FALSE) +
+    facet_wrap(~id) +
+  scale_x_continuous("", breaks = seq(1980, 2020, 10),
+                     labels = c("80", "90", "00", "10", "20")) +
+  ylab("Hourly wage")
+
+# Check values for short individuals
+short <- wages_hs_dropout %>%
+  count(id, sort=TRUE) %>%
+  filter(n == 5)
+wages_hs_dropout %>%
+  filter(id %in% short$id) %>%
+  ggplot(aes(x=year, y=mean_hourly_wage)) +
+  geom_point() +
+  geom_smooth(method="lm", se=FALSE) +
+  facet_wrap(~id) +
+  scale_x_continuous("", breaks = seq(1980, 2020, 10),
+                     labels = c("80", "90", "00", "10", "20")) +
+  ylab("Hourly wage")
+
+# Random collection of individuals
+
+# By demographics
+# hgc
+# race
+# gender
+# number of jobs or hours worked, aggregated over time
+

### 2021-09-01 Report/report_alt.Rmd
Modify the EDA.

--- 2021-08-12 Report/report_alt.Rmd
@@
 library(modelr)
+library(forcats)
+library(yowie)
 ```
@@
 Some observations had 0 in their hourly rate, which is considered as invalid value.
-Thus, their hourly rate set to be N.A.
-the result of the joining variables could be seen in the following tibble.
+Thus, their hourly rate set to be NA.
+The result of the joining variables could be seen in the following tibble.
 
@@
   mutate(rate_per_hour = ifelse(rate_per_hour == 0, NA,
-                                rate_per_hour))
+                                rate_per_hour),
+         hours_work = ifelse(hours_work == 0 | hours_work > 84, NA, hours_work))
+
 head(hours_wages)
@@
 # code 0 for arithmetic mean
-mean_hourly_wage <- 
-  eligible_wages %>%
+mean_hourly_wage <- eligible_wages %>%
   group_by(id, year) %>%
@@
 ```{r age-table, echo = FALSE}
-age_table <- yowie::demographic_nlsy79 %>%
+age_table <- yowie::demog_nlsy79 %>%
   group_by(age_1979) %>%
@@
 ```{r gender-race-table, echo = FALSE}
-gender_race_table <- yowie::demographic_nlsy79 %>%
+gender_race_table <- yowie::demog_nlsy79 %>%
   tabyl(gender, race) %>%
@@
 
-```{r wages-tsibble, echo = FALSE}
-# create the wages data as tsibble
-wages_hs2020_tsibble <- wages_hs2020 %>%
-  as_tsibble(key = id,
-             index = year,
-             regular = FALSE)
-```
-
-```{r slope-fig, echo = FALSE, fig.cap="Regression slope of mean hourly wage regression by highest grade completed.
-The regression model has mean hourly wage as response variable and year as predictor.
-Respondents who completed 12th grade has the highest median of slope."}
-
-wages_slope <- key_slope(wages_hs2020_tsibble, 
-            mean_hourly_wage ~ year) %>%
-  left_join(demographic_nlsy79, by = "id") %>%
-  mutate(hgc = factor(hgc, levels = c("1ST GRADE", "2ND GRADE", "3RD GRADE", 
-                                      "4TH GRADE", "5TH GRADE", "6TH GRADE", 
-                                      "7TH GRADE", "8TH GRADE", "9TH GRADE",
-                                      "10TH GRADE", "11TH GRADE", "12TH GRADE", 
-                                      "UNGRADED"))) %>%
-  filter(hgc != "UNGRADED")
-
-ggplot(wages_slope, aes(y = hgc,
-                      x = .slope_year)) +
-  geom_boxplot()
-```
-
-In the second example, we examined the pattern of mean hourly wage corresponding to different groups of gender and race.
-We took 240 samples of females and males equally and spread them into twelve facets, as shown in Figure \ref{fig:plot-gender} using `facet_strata` function from `brolgar` [@brolgar].
-We learn that the mean hourly wage of females and males fluctuated over the years.
-For example,  in facets 4, 5, and 10, males tend to earn more wages than females.
-
-```{r plot-gender, echo = FALSE, warning = FALSE, fig.cap= "The pattern of mean hourly wages overtime displayed in 12 facets coloured by gender.
-The respondents belong to high school cohort are randomly sampled.
-We learn that both male and female respondents had a fluctuate mean hourly wage over the year." }
-# sample the data by gender
-
-set.seed(31251587)
-wages_hs2020_male <- wages_hs2020_tsibble %>%
-  filter(gender == "MALE") %>%
-  sample_n_keys(size = 120)
-
-set.seed(31251587)
-wages_hs2020_female <- wages_hs2020_tsibble %>%
-  filter(gender == "FEMALE") %>%
-  sample_n_keys(size = 120)
-
-wages_hs2020_sample_gender <- bind_rows(wages_hs2020_male, wages_hs2020_female)
-
-ggplot(wages_hs2020_sample_gender, 
-       aes(x = year,
-           y = mean_hourly_wage,
-           group = id,
-           color = gender)) + 
-  geom_line(alpha = 0.7) +
-  scale_color_brewer(palette = "Dark2") +
-  facet_strata(along = ~gender) 
-```
-
-To further investigate this, we calculated the median wage of females relative to males, which is called gender wage gap.
+```{r}
+wages <- yowie::wages %>%
+  mutate(year0 = year - 1979,
+         ln_wages = log(wage))
+
+wages_slope <- key_slope(wages, ln_wages ~ year0) %>%
+  left_join(yowie::demog_nlsy79, by = "id") %>%
+  mutate(hgc_regroup = fct_collapse(hgc, `ungraded` = "UNGRADED",
+                                    `8TH grade or less` = c("1ST GRADE", "3RD GRADE",
+                                                                 "4TH GRADE", "5TH GRADE",
+                                                                 "6TH GRADE", "7TH GRADE",
+                                                                 "8TH GRADE"),
+                                    `9TH to 11TH grade` = c("9TH GRADE",
+                                                            "10TH GRADE",
+                                                            "11TH GRADE"),
+                                    `12TH grade` = "12TH GRADE")) %>%
+  mutate(hgc_regroup = factor(hgc_regroup, levels = c("8TH grade or less", 
+                                                      "9TH to 11TH grade", 
+                                                      "12TH grade")))
+
+```
+
+EDA:
+
+Exploring the wages slope of each individual:
+  - By sex
+  - By education 
+  - Exploring the wages slope pattern based on the number of observation
+
+- Slope by sex
+
+```{r}
+slope_bottom_5<- wages_slope %>%
+  arrange(.slope_year0) %>%
+  slice_head(n = 5) %>%
+  arrange(desc(.slope_year0)) %>%
+  mutate(note = "bottom 5") %>%
+  mutate(rank = seq_along(1:5))
+
+slope_top_5<- wages_slope %>%
+  arrange(desc(.slope_year0)) %>%
+  slice_head(n = 5) %>%
+  mutate(note = "top 5") %>%
+  mutate(rank = seq_along(1:5))
+
+keep_contrast <- rbind(slope_bottom_5, slope_top_5) %>%
+  dplyr::select(.slope_year0, id, note, rank, hgc_regroup)
+
+contrast_slope <- left_join(keep_contrast, wages, by = "id") %>%
+  arrange(.slope_year0) %>%
+  mutate(id = factor(id, levels = c("7918", "11695", "9176", "10119", "6464", "1271", 
+                                    "12371", "9039", "10217", "7782")),
+         note = factor(note, levels = c("top 5", "bottom 5")))
+
+ggplot(contrast_slope, aes(x = year, y = ln_wages, color = gender)) +
+  geom_point() +
+  geom_smooth(method = "lm") +
+  facet_grid(note ~ rank) +
+  scale_color_manual(values = c("#04B4AE", "#BF00FF"))
+
+```
+
+```{r}
+ggplot(contrast_slope, aes(x = year, y = ln_wages, color = hgc_regroup)) +
+  geom_point() +
+  geom_smooth(method = "lm") +
+  facet_grid(note ~ rank)
+```
+
+```{r}
+wages_obs <- wages %>%
+  features(wage, n_obs) %>%
+  left_join(wages_slope, by = "id")
+
+ggplot(wages_obs, aes(x = n_obs, y = .slope_year0)) +
+  geom_point() +
+  geom_jitter()
+```
+
 According to @OECD, gender pay gap is the difference between median wage of males and females relative to median wage of males.
@@
 
-wages_hs2020_grouped_gender <- wages_hs2020 %>%
+wages_gender <- as.tibble(wages) %>%
   group_by(gender, year) %>%
-  summarise(mean = mean(mean_hourly_wage),
-            median = median(mean_hourly_wage),
-             sd = sd(mean_hourly_wage),
+  summarise(mean = mean(wage),
+            median = median(wage),
+            sd = sd(wage),
             samplesize = n()) %>%
@@
 
-wage_female <- wages_hs2020_grouped_gender %>%
+wage_female <- wages_gender %>%
   filter(gender == "FEMALE") %>%
@@
 
-wage_male <- wages_hs2020_grouped_gender %>%
+wage_male <- wages_gender %>%
   filter(gender == "MALE") %>%

### 2021-09-01 Report/report_alt.tex

@@
 
-\title{The Journey from Wild to Text Book Data: A Case Study of the National Longitudinal Survey of Youth}
+\title{The Journey from Wild to Text Book Data: A Case Study from the National Longitudinal Survey of Youth}
 
@@
 % Pandoc citation processing
+\newlength{\csllabelwidth}
+\setlength{\csllabelwidth}{3em}
+\newlength{\cslhangindent}
+\setlength{\cslhangindent}{1.5em}
+% for Pandoc 2.8 to 2.10.1
+\newenvironment{cslreferences}%
+  {}%
+  {\par}
+% For Pandoc 2.11+
+\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
+ {% don't indent paragraphs
+  \setlength{\parindent}{0pt}
+  % turn on hanging indent if param 1 is 1
+  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
+  % set entry spacing
+  \ifnum #2 > 0
+  \setlength{\parskip}{#2\baselineskip}
+  \fi
+ }%
+ {}
+\usepackage{calc} % for calculating minipage widths
+\newcommand{\CSLBlock}[1]{#1\hfill\break}
+\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
+\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
+\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
 
@@
 This type of data can be useful as example data in statistical text books and for research purposes.
-However, open data are often, what we might call ``wild data'', because it requires substantial cleaning and tidying to tame it into text book shape.
-Huebner, Vach, and Cessie (2016) emphasize that making the data cleaning process accountable and transparent is imperative.
-Documenting the data cleaning is essential for the integrity of downstream statistical analyses and model building (Huebner et al.
+However, open data are often, what we might call ``wild data,'' because it requires substantial cleaning and tidying to tame it into text book shape.
+Huebner Marianne, Vach, and Cessie (2016) emphasize that making the data cleaning process accountable and transparent is imperative.
+Documenting the data cleaning is essential for the integrity of downstream statistical analyses and model building (M.
+Huebner et al.
 2020).
@@
 There are few research papers that document the data cleaning (Wickham 2014).
-Furthermore, the decisions made in this stage often go unreported in the sense that IDA is often performed in an unplanned and unstructured way, and is only shared among restricted parties (Huebner et al.
+Furthermore, the decisions made in this stage often go unreported in the sense that IDA is often performed in an unplanned and unstructured way, and is only shared among restricted parties (M.
+Huebner et al.
 2020).
@@
 \hline
-\rowcolor{gray!6}  1 & 328 & NA & NA & NA & NA & NA\\
+\cellcolor{gray!6}{1} & \cellcolor{gray!6}{328} & \cellcolor{gray!6}{NA} & \cellcolor{gray!6}{NA} & \cellcolor{gray!6}{NA} & \cellcolor{gray!6}{NA} & \cellcolor{gray!6}{NA}\\
 \hline
@@
 \hline
-\rowcolor{gray!6}  3 & 365 & NA & 275 & NA & NA & 397\\
+\cellcolor{gray!6}{3} & \cellcolor{gray!6}{365} & \cellcolor{gray!6}{NA} & \cellcolor{gray!6}{275} & \cellcolor{gray!6}{NA} & \cellcolor{gray!6}{NA} & \cellcolor{gray!6}{397}\\
 \hline
@@
 \hline
-\rowcolor{gray!6}  5 & 310 & 375 & NA & NA & NA & 333\\
+\cellcolor{gray!6}{5} & \cellcolor{gray!6}{310} & \cellcolor{gray!6}{375} & \cellcolor{gray!6}{NA} & \cellcolor{gray!6}{NA} & \cellcolor{gray!6}{NA} & \cellcolor{gray!6}{333}\\
 \hline
@@
 Consequently, the data should be tidied and wrangled first to extract the demographic and employment variables that we want to put in the final data set.
-We mainly used \texttt{tidyr} (H.
-Wickham 2020b), \texttt{dplyr} (H.
-Wickham, François, et al.
-2020), and \texttt{stringr} (H.
-Wickham 2019) to do this job.
+We mainly used \texttt{tidyr} (Wickham 2020b), \texttt{dplyr} (Wickham et al.
+2020), and \texttt{stringr} (Wickham 2019) to do this job.
 
@@
 \begin{Highlighting}[]
-\CommentTok{## tidy the date of birth data}
-\NormalTok{dob_tidy <-}\StringTok{ }\NormalTok{new_data_qnames }\OperatorTok{%>%}
-\StringTok{  }\NormalTok{dplyr}\OperatorTok{::}\KeywordTok{select}\NormalTok{(CASEID_}\DecValTok{1979}\NormalTok{,}
-         \KeywordTok{starts_with}\NormalTok{(}\StringTok{"Q1-3_A~"}\NormalTok{)) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{dob_year =} \KeywordTok{case_when}\NormalTok{(}
-                    \CommentTok{# if the years recorded in both sets match, take 79 data}
-                    \StringTok{`}\DataTypeTok{Q1-3_A~Y_1979}\StringTok{`} \OperatorTok{==}\StringTok{ `}\DataTypeTok{Q1-3_A~Y_1981}\StringTok{`} \OperatorTok{~}\StringTok{ `}\DataTypeTok{Q1-3_A~Y_1979}\StringTok{`}\NormalTok{,}
-                    \CommentTok{# if the year in the 81 set is missing, take the 79 data}
-                    \KeywordTok{is.na}\NormalTok{(}\StringTok{`}\DataTypeTok{Q1-3_A~Y_1981}\StringTok{`}\NormalTok{) }\OperatorTok{~}\StringTok{ `}\DataTypeTok{Q1-3_A~Y_1979}\StringTok{`}\NormalTok{,}
-                    \CommentTok{# if the sets don't match for dob year, take the 79 data}
-                    \StringTok{`}\DataTypeTok{Q1-3_A~Y_1979}\StringTok{`} \OperatorTok{!=}\StringTok{ `}\DataTypeTok{Q1-3_A~Y_1981}\StringTok{`} \OperatorTok{~}\StringTok{ `}\DataTypeTok{Q1-3_A~Y_1979}\StringTok{`}\NormalTok{),}
-        \DataTypeTok{dob_month =} \KeywordTok{case_when}\NormalTok{(}
-                    \CommentTok{# if months recorded in both sets match, take 79 data}
-                    \StringTok{`}\DataTypeTok{Q1-3_A~M_1979}\StringTok{`} \OperatorTok{==}\StringTok{ `}\DataTypeTok{Q1-3_A~M_1981}\StringTok{`} \OperatorTok{~}\StringTok{ `}\DataTypeTok{Q1-3_A~M_1979}\StringTok{`}\NormalTok{,}
-                    \CommentTok{# if month in 81 set is missing, take the 79 data}
-                    \KeywordTok{is.na}\NormalTok{(}\StringTok{`}\DataTypeTok{Q1-3_A~M_1981}\StringTok{`}\NormalTok{) }\OperatorTok{~}\StringTok{ `}\DataTypeTok{Q1-3_A~M_1979}\StringTok{`}\NormalTok{,}
-                    \CommentTok{# if sets don't match for dob month, take the 79 data}
-                    \StringTok{`}\DataTypeTok{Q1-3_A~M_1979}\StringTok{`} \OperatorTok{!=}\StringTok{ `}\DataTypeTok{Q1-3_A~M_1981}\StringTok{`} \OperatorTok{~}\StringTok{ `}\DataTypeTok{Q1-3_A~M_1979}\StringTok{`}\NormalTok{),}
-        \CommentTok{# flag if there is a conflict between dob recorded in 79 and 81}
-        \DataTypeTok{dob_conflict =} \KeywordTok{case_when}\NormalTok{(     }
-\NormalTok{                      (}\StringTok{`}\DataTypeTok{Q1-3_A~M_1979}\StringTok{`} \OperatorTok{!=}\StringTok{ `}\DataTypeTok{Q1-3_A~M_1981}\StringTok{`}\NormalTok{) }\OperatorTok{&}\StringTok{ }\OperatorTok{!}\KeywordTok{is.na}\NormalTok{(}\StringTok{`}\DataTypeTok{Q1-3_A~M_1981}\StringTok{`}\NormalTok{)}
-                      \OperatorTok{~}\StringTok{ }\OtherTok{TRUE}\NormalTok{,}
-\NormalTok{                      (}\StringTok{`}\DataTypeTok{Q1-3_A~Y_1979}\StringTok{`} \OperatorTok{!=}\StringTok{ `}\DataTypeTok{Q1-3_A~Y_1981}\StringTok{`}\NormalTok{) }\OperatorTok{&}\StringTok{ }\OperatorTok{!}\KeywordTok{is.na}\NormalTok{(}\StringTok{`}\DataTypeTok{Q1-3_A~Y_1981}\StringTok{`}\NormalTok{)}
-                      \OperatorTok{~}\StringTok{ }\OtherTok{TRUE}\NormalTok{,}
-\NormalTok{                      (}\StringTok{`}\DataTypeTok{Q1-3_A~Y_1979}\StringTok{`} \OperatorTok{==}\StringTok{ `}\DataTypeTok{Q1-3_A~Y_1981}\StringTok{`}\NormalTok{) }\OperatorTok{&}\StringTok{ }
-\StringTok{                      }\NormalTok{(}\StringTok{`}\DataTypeTok{Q1-3_A~M_1979}\StringTok{`} \OperatorTok{==}\StringTok{ `}\DataTypeTok{Q1-3_A~M_1981}\StringTok{`}\NormalTok{) }\OperatorTok{~}\StringTok{ }\OtherTok{FALSE}\NormalTok{,}
-                      \KeywordTok{is.na}\NormalTok{(}\StringTok{`}\DataTypeTok{Q1-3_A~M_1981}\StringTok{`}\NormalTok{) }\OperatorTok{|}\StringTok{ }\KeywordTok{is.na}\NormalTok{(}\StringTok{`}\DataTypeTok{Q1-3_A~Y_1981}\StringTok{`}\NormalTok{) }\OperatorTok{~}\StringTok{ }\OtherTok{FALSE}\NormalTok{)) }\OperatorTok{%>%}
-\StringTok{  }\NormalTok{dplyr}\OperatorTok{::}\KeywordTok{select}\NormalTok{(CASEID_}\DecValTok{1979}\NormalTok{,}
-\NormalTok{         dob_month,}
-\NormalTok{         dob_year,}
-\NormalTok{         dob_conflict)}
+\DocumentationTok{\#\# tidy the date of birth data}
+\NormalTok{dob\_tidy }\OtherTok{\textless{}{-}}\NormalTok{ new\_data\_qnames }\SpecialCharTok{\%\textgreater{}\%}
+\NormalTok{  dplyr}\SpecialCharTok{::}\FunctionTok{select}\NormalTok{(CASEID\_1979,}
+         \FunctionTok{starts\_with}\NormalTok{(}\StringTok{"Q1{-}3\_A\textasciitilde{}"}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{dob\_year =} \FunctionTok{case\_when}\NormalTok{(}
+                    \CommentTok{\# if the years recorded in both sets match, take 79 data}
+                    \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}Y\_1979}\StringTok{\textasciigrave{}} \SpecialCharTok{==} \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}Y\_1981}\StringTok{\textasciigrave{}} \SpecialCharTok{\textasciitilde{}} \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}Y\_1979}\StringTok{\textasciigrave{}}\NormalTok{,}
+                    \CommentTok{\# if the year in the 81 set is missing, take the 79 data}
+                    \FunctionTok{is.na}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}Y\_1981}\StringTok{\textasciigrave{}}\NormalTok{) }\SpecialCharTok{\textasciitilde{}} \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}Y\_1979}\StringTok{\textasciigrave{}}\NormalTok{,}
+                    \CommentTok{\# if the sets don\textquotesingle{}t match for dob year, take the 79 data}
+                    \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}Y\_1979}\StringTok{\textasciigrave{}} \SpecialCharTok{!=} \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}Y\_1981}\StringTok{\textasciigrave{}} \SpecialCharTok{\textasciitilde{}} \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}Y\_1979}\StringTok{\textasciigrave{}}\NormalTok{),}
+        \AttributeTok{dob\_month =} \FunctionTok{case\_when}\NormalTok{(}
+                    \CommentTok{\# if months recorded in both sets match, take 79 data}
+                    \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}M\_1979}\StringTok{\textasciigrave{}} \SpecialCharTok{==} \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}M\_1981}\StringTok{\textasciigrave{}} \SpecialCharTok{\textasciitilde{}} \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}M\_1979}\StringTok{\textasciigrave{}}\NormalTok{,}
+                    \CommentTok{\# if month in 81 set is missing, take the 79 data}
+                    \FunctionTok{is.na}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}M\_1981}\StringTok{\textasciigrave{}}\NormalTok{) }\SpecialCharTok{\textasciitilde{}} \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}M\_1979}\StringTok{\textasciigrave{}}\NormalTok{,}
+                    \CommentTok{\# if sets don\textquotesingle{}t match for dob month, take the 79 data}
+                    \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}M\_1979}\StringTok{\textasciigrave{}} \SpecialCharTok{!=} \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}M\_1981}\StringTok{\textasciigrave{}} \SpecialCharTok{\textasciitilde{}} \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}M\_1979}\StringTok{\textasciigrave{}}\NormalTok{),}
+        \CommentTok{\# flag if there is a conflict between dob recorded in 79 and 81}
+        \AttributeTok{dob\_conflict =} \FunctionTok{case\_when}\NormalTok{(     }
+\NormalTok{                      (}\StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}M\_1979}\StringTok{\textasciigrave{}} \SpecialCharTok{!=} \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}M\_1981}\StringTok{\textasciigrave{}}\NormalTok{) }\SpecialCharTok{\&} \SpecialCharTok{!}\FunctionTok{is.na}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}M\_1981}\StringTok{\textasciigrave{}}\NormalTok{)}
+                      \SpecialCharTok{\textasciitilde{}} \ConstantTok{TRUE}\NormalTok{,}
+\NormalTok{                      (}\StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}Y\_1979}\StringTok{\textasciigrave{}} \SpecialCharTok{!=} \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}Y\_1981}\StringTok{\textasciigrave{}}\NormalTok{) }\SpecialCharTok{\&} \SpecialCharTok{!}\FunctionTok{is.na}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}Y\_1981}\StringTok{\textasciigrave{}}\NormalTok{)}
+                      \SpecialCharTok{\textasciitilde{}} \ConstantTok{TRUE}\NormalTok{,}
+\NormalTok{                      (}\StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}Y\_1979}\StringTok{\textasciigrave{}} \SpecialCharTok{==} \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}Y\_1981}\StringTok{\textasciigrave{}}\NormalTok{) }\SpecialCharTok{\&}
+\NormalTok{                      (}\StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}M\_1979}\StringTok{\textasciigrave{}} \SpecialCharTok{==} \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}M\_1981}\StringTok{\textasciigrave{}}\NormalTok{) }\SpecialCharTok{\textasciitilde{}} \ConstantTok{FALSE}\NormalTok{,}
+                      \FunctionTok{is.na}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}M\_1981}\StringTok{\textasciigrave{}}\NormalTok{) }\SpecialCharTok{|} \FunctionTok{is.na}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}Y\_1981}\StringTok{\textasciigrave{}}\NormalTok{) }\SpecialCharTok{\textasciitilde{}} \ConstantTok{FALSE}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%}
+\NormalTok{  dplyr}\SpecialCharTok{::}\FunctionTok{select}\NormalTok{(CASEID\_1979,}
+\NormalTok{         dob\_month,}
+\NormalTok{         dob\_year,}
+\NormalTok{         dob\_conflict)}
 \end{Highlighting}
@@
 \begin{Highlighting}[]
-\CommentTok{## tidy the gender and race variables}
-\NormalTok{demog_tidy <-}\StringTok{ }\NormalTok{categories_qnames }\OperatorTok{%>%}
-\StringTok{  }\NormalTok{dplyr}\OperatorTok{::}\KeywordTok{select}\NormalTok{(CASEID_}\DecValTok{1979}\NormalTok{,}
-\NormalTok{         SAMPLE_RACE_78SCRN,}
-\NormalTok{         SAMPLE_SEX_}\DecValTok{1979}\NormalTok{) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{rename}\NormalTok{(}\DataTypeTok{gender =}\NormalTok{ SAMPLE_SEX_}\DecValTok{1979}\NormalTok{,}
-         \DataTypeTok{race =}\NormalTok{ SAMPLE_RACE_78SCRN)}
+\DocumentationTok{\#\# tidy the gender and race variables}
+\NormalTok{demog\_tidy }\OtherTok{\textless{}{-}}\NormalTok{ categories\_qnames }\SpecialCharTok{\%\textgreater{}\%}
+\NormalTok{  dplyr}\SpecialCharTok{::}\FunctionTok{select}\NormalTok{(CASEID\_1979,}
+\NormalTok{         SAMPLE\_RACE\_78SCRN,}
+\NormalTok{         SAMPLE\_SEX\_1979) }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{rename}\NormalTok{(}\AttributeTok{gender =}\NormalTok{ SAMPLE\_SEX\_1979,}
+         \AttributeTok{race =}\NormalTok{ SAMPLE\_RACE\_78SCRN)}
 \end{Highlighting}
@@
 \begin{Highlighting}[]
-\CommentTok{# make a list for years where we used the "QES-52A"}
-\NormalTok{year_A <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\DecValTok{1979}\OperatorTok{:}\DecValTok{1987}\NormalTok{, }\DecValTok{1993}\NormalTok{)}
-\CommentTok{#function to get the hour of work}
-\NormalTok{get_hour <-}\StringTok{ }\ControlFlowTok{function}\NormalTok{(year)\{}
-  \ControlFlowTok{if}\NormalTok{(year }\OperatorTok{%in%}\StringTok{ }\NormalTok{year_A)\{}
-\NormalTok{   temp <-}\StringTok{ }\NormalTok{new_data_qnames }\OperatorTok{%>%}
-\StringTok{    }\NormalTok{dplyr}\OperatorTok{::}\KeywordTok{select}\NormalTok{(CASEID_}\DecValTok{1979}\NormalTok{,}
-            \KeywordTok{starts_with}\NormalTok{(}\StringTok{"QES-52A"}\NormalTok{) }\OperatorTok{&}
-\StringTok{              }\KeywordTok{ends_with}\NormalTok{(}\KeywordTok{as.character}\NormalTok{(year)))\} }
+\CommentTok{\# make a list for years where we used the "QES{-}52A"}
+\NormalTok{year\_A }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(}\DecValTok{1979}\SpecialCharTok{:}\DecValTok{1987}\NormalTok{, }\DecValTok{1993}\NormalTok{)}
+\CommentTok{\#function to get the hour of work}
+\NormalTok{get\_hour }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{(year)\{}
+  \ControlFlowTok{if}\NormalTok{(year }\SpecialCharTok{\%in\%}\NormalTok{ year\_A)\{}
+\NormalTok{   temp }\OtherTok{\textless{}{-}}\NormalTok{ new\_data\_qnames }\SpecialCharTok{\%\textgreater{}\%}
+\NormalTok{    dplyr}\SpecialCharTok{::}\FunctionTok{select}\NormalTok{(CASEID\_1979,}
+            \FunctionTok{starts\_with}\NormalTok{(}\StringTok{"QES{-}52A"}\NormalTok{) }\SpecialCharTok{\&}
+              \FunctionTok{ends\_with}\NormalTok{(}\FunctionTok{as.character}\NormalTok{(year)))\} }
   \ControlFlowTok{else}\NormalTok{\{}
-\NormalTok{    temp <-}\StringTok{ }\NormalTok{new_data_qnames }\OperatorTok{%>%}
-\StringTok{    }\NormalTok{dplyr}\OperatorTok{::}\KeywordTok{select}\NormalTok{(CASEID_}\DecValTok{1979}\NormalTok{,}
-            \KeywordTok{starts_with}\NormalTok{(}\StringTok{"QES-52D"}\NormalTok{) }\OperatorTok{&}
-\StringTok{              }\KeywordTok{ends_with}\NormalTok{(}\KeywordTok{as.character}\NormalTok{(year)))\} }
-\NormalTok{    temp }\OperatorTok{%>%}\StringTok{ }
-\StringTok{      }\KeywordTok{pivot_longer}\NormalTok{(}\OperatorTok{!}\NormalTok{CASEID_}\DecValTok{1979}\NormalTok{,}
-                 \DataTypeTok{names_to =} \StringTok{"job"}\NormalTok{,}
-                 \DataTypeTok{values_to =} \StringTok{"hours_work"}\NormalTok{) }\OperatorTok{%>%}
-\StringTok{      }\KeywordTok{separate}\NormalTok{(}\StringTok{"job"}\NormalTok{, }\KeywordTok{c}\NormalTok{(}\StringTok{"job"}\NormalTok{, }\StringTok{"year"}\NormalTok{), }\DataTypeTok{sep =} \DecValTok{-4}\NormalTok{) }\OperatorTok{%>%}
-\StringTok{      }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{job =} \KeywordTok{paste0}\NormalTok{(}\StringTok{"job_"}\NormalTok{, }\KeywordTok{substr}\NormalTok{(job, }\DecValTok{9}\NormalTok{, }\DecValTok{10}\NormalTok{))) }\OperatorTok{%>%}
-\StringTok{      }\KeywordTok{rename}\NormalTok{(}\DataTypeTok{id =}\NormalTok{ CASEID_}\DecValTok{1979}\NormalTok{)}
+\NormalTok{    temp }\OtherTok{\textless{}{-}}\NormalTok{ new\_data\_qnames }\SpecialCharTok{\%\textgreater{}\%}
+\NormalTok{    dplyr}\SpecialCharTok{::}\FunctionTok{select}\NormalTok{(CASEID\_1979,}
+            \FunctionTok{starts\_with}\NormalTok{(}\StringTok{"QES{-}52D"}\NormalTok{) }\SpecialCharTok{\&}
+              \FunctionTok{ends\_with}\NormalTok{(}\FunctionTok{as.character}\NormalTok{(year)))\} }
+\NormalTok{    temp }\SpecialCharTok{\%\textgreater{}\%} 
+      \FunctionTok{pivot\_longer}\NormalTok{(}\SpecialCharTok{!}\NormalTok{CASEID\_1979,}
+                 \AttributeTok{names\_to =} \StringTok{"job"}\NormalTok{,}
+                 \AttributeTok{values\_to =} \StringTok{"hours\_work"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
+      \FunctionTok{separate}\NormalTok{(}\StringTok{"job"}\NormalTok{, }\FunctionTok{c}\NormalTok{(}\StringTok{"job"}\NormalTok{, }\StringTok{"year"}\NormalTok{), }\AttributeTok{sep =} \SpecialCharTok{{-}}\DecValTok{4}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
+      \FunctionTok{mutate}\NormalTok{(}\AttributeTok{job =} \FunctionTok{paste0}\NormalTok{(}\StringTok{"job\_"}\NormalTok{, }\FunctionTok{substr}\NormalTok{(job, }\DecValTok{9}\NormalTok{, }\DecValTok{10}\NormalTok{))) }\SpecialCharTok{\%\textgreater{}\%}
+      \FunctionTok{rename}\NormalTok{(}\AttributeTok{id =}\NormalTok{ CASEID\_1979)}
 \NormalTok{\}}
 
-\CommentTok{# list to save the iteration result}
-\NormalTok{hours <-}\StringTok{ }\KeywordTok{list}\NormalTok{()}
-\CommentTok{# getting the hours of work of all observations}
-\ControlFlowTok{for}\NormalTok{(ayear }\ControlFlowTok{in} \KeywordTok{c}\NormalTok{(}\DecValTok{1979}\OperatorTok{:}\DecValTok{1994}\NormalTok{, }\DecValTok{1996}\NormalTok{, }\DecValTok{1998}\NormalTok{, }\DecValTok{2000}\NormalTok{, }\DecValTok{2002}\NormalTok{, }\DecValTok{2004}\NormalTok{, }\DecValTok{2006}\NormalTok{, }\DecValTok{2008}\NormalTok{, }\DecValTok{2010}\NormalTok{, }
+\CommentTok{\# list to save the iteration result}
+\NormalTok{hours }\OtherTok{\textless{}{-}} \FunctionTok{list}\NormalTok{()}
+\CommentTok{\# getting the hours of work of all observations}
+\ControlFlowTok{for}\NormalTok{(ayear }\ControlFlowTok{in} \FunctionTok{c}\NormalTok{(}\DecValTok{1979}\SpecialCharTok{:}\DecValTok{1994}\NormalTok{, }\DecValTok{1996}\NormalTok{, }\DecValTok{1998}\NormalTok{, }\DecValTok{2000}\NormalTok{, }\DecValTok{2002}\NormalTok{, }\DecValTok{2004}\NormalTok{, }\DecValTok{2006}\NormalTok{, }\DecValTok{2008}\NormalTok{, }\DecValTok{2010}\NormalTok{, }
                \DecValTok{2012}\NormalTok{, }\DecValTok{2014}\NormalTok{, }\DecValTok{2016}\NormalTok{, }\DecValTok{2018}\NormalTok{)) \{}
-\NormalTok{   hours[[ayear]] <-}\StringTok{ }\KeywordTok{get_hour}\NormalTok{(ayear)}
+\NormalTok{   hours[[ayear]] }\OtherTok{\textless{}{-}} \FunctionTok{get\_hour}\NormalTok{(ayear)}
 \NormalTok{\}}
-\CommentTok{# unlist the hours of work}
-\NormalTok{hours_all <-}\StringTok{ }\KeywordTok{bind_rows}\NormalTok{(}\OperatorTok{!!!}\NormalTok{hours)}
+\CommentTok{\# unlist the hours of work}
+\NormalTok{hours\_all }\OtherTok{\textless{}{-}} \FunctionTok{bind\_rows}\NormalTok{(}\SpecialCharTok{!!!}\NormalTok{hours)}
 \end{Highlighting}
@@
 
-According to Huebner, Vach, and Cessie (2016), Initial Data Analysis (IDA) is the step of inspecting and screening the data after being collected to ensure that the data is clean, valid, and ready to be deployed in the later formal statistical analysis.
+According to Huebner Marianne, Vach, and Cessie (2016), Initial Data Analysis (IDA) is the step of inspecting and screening the data after being collected to ensure that the data is clean, valid, and ready to be deployed in the later formal statistical analysis.
 Moreover, Chatfield (1985) argued that the two main objectives of IDA is data description, which is to assess the structure and the quality of the data; and model formulation without any formal statistical inference.
@@
 \hline
-\rowcolor{gray!6}  15 & 1265\\
+\cellcolor{gray!6}{15} & \cellcolor{gray!6}{1265}\\
 \hline
@@
 \hline
-\rowcolor{gray!6}  17 & 1600\\
+\cellcolor{gray!6}{17} & \cellcolor{gray!6}{1600}\\
 \hline
@@
 \hline
-\rowcolor{gray!6}  19 & 1662\\
+\cellcolor{gray!6}{19} & \cellcolor{gray!6}{1662}\\
 \hline
@@
 \hline
-\rowcolor{gray!6}  21 & 1677\\
+\cellcolor{gray!6}{21} & \cellcolor{gray!6}{1677}\\
 \hline
@@
 \hline
-\rowcolor{gray!6}  Male & 1000 (15.62\%) & 1613 (25.19\%) & 3790 (59.19\%) & 6403 (100.00\%)\\
+\cellcolor{gray!6}{Male} & \cellcolor{gray!6}{1000 (15.62\%)} & \cellcolor{gray!6}{1613 (25.19\%)} & \cellcolor{gray!6}{3790 (59.19\%)} & \cellcolor{gray!6}{6403 (100.00\%)}\\
 \hline
@@
 \hline
-\rowcolor{gray!6}  Total & 2002 (15.78\%) & 3174 (25.02\%) & 7510 (59.20\%) & 12686 (100.00\%)\\
+\cellcolor{gray!6}{Total} & \cellcolor{gray!6}{2002 (15.78\%)} & \cellcolor{gray!6}{3174 (25.02\%)} & \cellcolor{gray!6}{7510 (59.20\%)} & \cellcolor{gray!6}{12686 (100.00\%)}\\
 \hline
@@
 \hline
-\rowcolor{gray!6}  Min. & 0.01000\\
+\cellcolor{gray!6}{Min.} & \cellcolor{gray!6}{0.01000}\\
 \hline
@@
 \hline
-\rowcolor{gray!6}  Median & 7.20000\\
+\cellcolor{gray!6}{Median} & \cellcolor{gray!6}{7.20000}\\
 \hline
@@
 \hline
-\rowcolor{gray!6}  3rd Qu. & 11.74000\\
+\cellcolor{gray!6}{3rd Qu.} & \cellcolor{gray!6}{11.74000}\\
 \hline
@@
 However, this method is too computationally and memory expensive, especially for a large data set, like the NLSY79 data.
-Thus, the model for each ID is built utilizing the \texttt{nest} and \texttt{map} function from \texttt{tidyr} (H.
-Wickham 2020b) and \texttt{purrr} (Henry and Wickham 2020) respectively.
+Thus, the model for each ID is built utilizing the \texttt{nest} and \texttt{map} function from \texttt{tidyr} (Wickham 2020b) and \texttt{purrr} (Henry and Wickham 2020) respectively.
 
@@
 
-To minimize the risk of mistakenly regarding an outlier as an ``erroneous outlier'', we have simulated some thresholds and study how they affected the data.
+To minimize the risk of mistakenly regarding an outlier as an ``erroneous outlier,'' we have simulated some thresholds and study how they affected the data.
 We found that 0.12 is the most reasonable value to be the threshold to minimize that drawback's risk because it still captures the sensible spikes in the data.
@@
 It implies that the fluctuation can still be observed in the data after the treatment.
-However, the large spikes, which are considered ``erroneous outliers'', are already eliminated from the data.
+However, the large spikes, which are considered ``erroneous outliers,'' are already eliminated from the data.
 Hence, the model produces a data set with a more reasonable degree of fluctuation.
@@
 
+\begin{Shaded}
+\begin{Highlighting}[]
+\FunctionTok{ggplot}\NormalTok{(wages\_slope, }\FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ .slope\_year0)) }\SpecialCharTok{+}
+  \FunctionTok{geom\_histogram}\NormalTok{()}
+\end{Highlighting}
+\end{Shaded}
+
+\includegraphics{figures/unnamed-chunk-2-1.pdf}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{slope\_bottom\_5}\OtherTok{\textless{}{-}}\NormalTok{ wages\_slope }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{arrange}\NormalTok{(.slope\_year0) }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{slice\_head}\NormalTok{(}\AttributeTok{n =} \DecValTok{5}\NormalTok{)}
+
+\NormalTok{slope\_top\_5}\OtherTok{\textless{}{-}}\NormalTok{ wages\_slope }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{arrange}\NormalTok{(}\FunctionTok{desc}\NormalTok{(.slope\_year0)) }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{slice\_head}\NormalTok{(}\AttributeTok{n =} \DecValTok{5}\NormalTok{)}
+
+\NormalTok{keep\_contrast }\OtherTok{\textless{}{-}} \FunctionTok{rbind}\NormalTok{(slope\_bottom\_5, slope\_top\_5) }\SpecialCharTok{\%\textgreater{}\%}
+\NormalTok{  dplyr}\SpecialCharTok{::}\FunctionTok{select}\NormalTok{(.slope\_year0, id)}
+
+\NormalTok{contrast\_slope }\OtherTok{\textless{}{-}} \FunctionTok{left\_join}\NormalTok{(keep\_contrast, wages\_model\_data, }\AttributeTok{by =} \StringTok{"id"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{arrange}\NormalTok{(.slope\_year0) }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{id =} \FunctionTok{factor}\NormalTok{(id, }\AttributeTok{levels =} \FunctionTok{c}\NormalTok{(}\StringTok{"7918"}\NormalTok{, }\StringTok{"11695"}\NormalTok{, }\StringTok{"9176"}\NormalTok{, }\StringTok{"10119"}\NormalTok{, }\StringTok{"6464"}\NormalTok{, }\StringTok{"1271"}\NormalTok{, }
+                                    \StringTok{"12371"}\NormalTok{, }\StringTok{"9039"}\NormalTok{, }\StringTok{"10217"}\NormalTok{, }\StringTok{"7782"}\NormalTok{)))}
+
+\FunctionTok{ggplot}\NormalTok{(contrast\_slope, }\FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ year, }\AttributeTok{y =}\NormalTok{ ln\_wages, }\AttributeTok{color =}\NormalTok{ gender)) }\SpecialCharTok{+}
+  \FunctionTok{geom\_line}\NormalTok{() }\SpecialCharTok{+}
+  \FunctionTok{geom\_point}\NormalTok{() }\SpecialCharTok{+}
+  \FunctionTok{facet\_wrap}\NormalTok{(}\SpecialCharTok{\textasciitilde{}}\NormalTok{id)}
+\end{Highlighting}
+\end{Shaded}
+
+\includegraphics{figures/unnamed-chunk-3-1.pdf}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\FunctionTok{ggplot}\NormalTok{(contrast\_slope, }\FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ year, }\AttributeTok{y =}\NormalTok{ ln\_wages, }\AttributeTok{color =}\NormalTok{ hgc\_regroup)) }\SpecialCharTok{+}
+  \FunctionTok{geom\_line}\NormalTok{() }\SpecialCharTok{+}
+  \FunctionTok{geom\_point}\NormalTok{() }\SpecialCharTok{+}
+  \FunctionTok{facet\_wrap}\NormalTok{(}\SpecialCharTok{\textasciitilde{}}\NormalTok{id)}
+\end{Highlighting}
+\end{Shaded}
+
+\includegraphics{figures/unnamed-chunk-4-1.pdf}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{quartile }\OtherTok{\textless{}{-}}\NormalTok{ wages\_slope }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{quartile =} \FunctionTok{as.factor}\NormalTok{(}\FunctionTok{ntile}\NormalTok{(.slope\_year0, }\DecValTok{4}\NormalTok{)))}
+
+\FunctionTok{ggplot}\NormalTok{(quartile) }\SpecialCharTok{+}
+  \FunctionTok{geom\_bar}\NormalTok{(}\FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ quartile, }\AttributeTok{fill =}\NormalTok{ gender, }\AttributeTok{position =} \StringTok{"fill"}\NormalTok{))}
+\end{Highlighting}
+\end{Shaded}
+
+\includegraphics{figures/unnamed-chunk-5-1.pdf}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\FunctionTok{ggplot}\NormalTok{(quartile) }\SpecialCharTok{+}
+  \FunctionTok{geom\_bar}\NormalTok{(}\FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ quartile, }\AttributeTok{fill =}\NormalTok{ hgc\_regroup, }\AttributeTok{position =} \StringTok{"fill"}\NormalTok{))}
+\end{Highlighting}
+\end{Shaded}
+
+\includegraphics{figures/unnamed-chunk-6-1.pdf}
+
 In the second example, we examined the pattern of mean hourly wage corresponding to different groups of gender and race.
@@
 
-The entire analysis is conducted using \texttt{R} (R Core Team 2020) in \texttt{rstudio} using these packages: \texttt{tidyverse} (H.
-Wickham, Averick, et al.
-2019), \texttt{ggplot2} (Wickham 2016), \texttt{dplyr} (H.
-Wickham, François, et al.
-2020), \texttt{readr} (H.
-Wickham and Hester 2020), \texttt{tidyr} (H.
-Wickham 2020b), \texttt{stringr} (H.
-Wickham 2019), \texttt{purrr} (Henry and Wickham 2020), \texttt{broom} (Robinson, Hayes, and Couch 2021), \texttt{blorgar} (Tierney, Cook, and Prvan 2020), \texttt{patchwork} (Pedersen 2020), \texttt{kableExtra} (Zhu 2019), \texttt{MASS} (Venables and Ripley 2002), \texttt{janitor} (Firke 2020), \texttt{DiagrammeR} (Iannone 2020), \texttt{rsvg} (Ooms 2020), \texttt{webshot} (Chang 2019), \texttt{mgcv} (Wood 2003), \texttt{tsibble} (Wang, Cook, and Hyndman 2020), and \texttt{modelr} (H.
-Wickham 2020a).
+The entire analysis is conducted using \texttt{R} (R Core Team 2020) in \texttt{rstudio} using these packages: \texttt{tidyverse} (Wickham et al.
+2019), \texttt{ggplot2} (Wickham 2016), \texttt{dplyr} (Wickham et al.
+2020), \texttt{readr} (Wickham and Hester 2020), \texttt{tidyr} (Wickham 2020b), \texttt{stringr} (Wickham 2019), \texttt{purrr} (Henry and Wickham 2020), \texttt{broom} (Robinson, Hayes, and Couch 2021), \texttt{blorgar} (Tierney, Cook, and Prvan 2020), \texttt{patchwork} (Pedersen 2020), \texttt{kableExtra} (Zhu 2019), \texttt{MASS} (Venables and Ripley 2002), \texttt{janitor} (Firke 2020), \texttt{DiagrammeR} (Iannone 2020), \texttt{rsvg} (Ooms 2020), \texttt{webshot} (Chang 2019), \texttt{mgcv} (Wood 2003), \texttt{tsibble} (Wang, Cook, and Hyndman 2020), and \texttt{modelr} (Wickham 2020a).
 The paper are generated using \texttt{knitr} (Xie 2014), \texttt{rmarkdown} (Xie, Dervieux, and Riederer 2020), and \texttt{rticles} (Allaire et al.
@@
 \hypertarget{refs}{}
+\begin{CSLReferences}{1}{0}
 \leavevmode\hypertarget{ref-rticles}{}%
@@
 2021.
-\emph{Rticles: Article Formats for R Markdown}.
+\emph{Rticles: Article Formats for r Markdown}.
 \url{https://CRAN.R-project.org/package=rticles}.
@@
 2015.
-``Fitting Linear Mixed-Effects Models Using lme4.'' \emph{Journal of Statistical Software} 67 (1): 1--48.
+{``Fitting Linear Mixed-Effects Models Using {lme4}.''} \emph{Journal of Statistical Software} 67 (1): 1--48.
 \url{https://doi.org/10.18637/jss.v067.i01}.
@@
 \leavevmode\hypertarget{ref-shiny}{}%
-Chang, Winston, Joe Cheng, JJ Allaire, Yihui Xie, and Jonathan McPherson. 2020. \emph{Shiny: Web Application Framework for R}. \url{https://CRAN.R-project.org/package=shiny}.
+Chang, Winston, Joe Cheng, JJ Allaire, Yihui Xie, and Jonathan McPherson. 2020. \emph{Shiny: Web Application Framework for r}. \url{https://CRAN.R-project.org/package=shiny}.
 
 \leavevmode\hypertarget{ref-Chatfield1985TIEo}{}%
-Chatfield, C. 1985. ``The Initial Examination of Data.'' \emph{Journal of the Royal Statistical Society. Series A. General} 148 (3): 214--53.
+Chatfield, C. 1985. {``The Initial Examination of Data.''} \emph{Journal of the Royal Statistical Society. Series A. General} 148 (3): 214--53.
 
@@
 2017.
-``Using the National Longitudinal Surveys of Youth (Nlsy) to Conduct Life Course Analyses.'' In \emph{Handbook of Life Course Health Development}, edited by Richard M.
+{``Using the National Longitudinal Surveys of Youth (NLSY) to Conduct Life Course Analyses.''} In \emph{Handbook of Life Course Health Development}, edited by Richard M.
 Lerner Neal Halfon Christoper B.
@@
 Cham: Springer.
-\url{https://doi.org/https://doi.org/10.1007/978-3-319-47143-3_23}.
+https://doi.org/\url{https://doi.org/10.1007/978-3-319-47143-3_23}.
 
@@
 2016.
-``A Systematic Approach to Initial Data Analysis Is Good Research Practice.'' \emph{The Journal of Thoracic and Cardiovascular Surgery} 151 (1): 25--27.
+{``A Systematic Approach to Initial Data Analysis Is Good Research Practice.''} \emph{The Journal of Thoracic and Cardiovascular Surgery} 151 (1): 25--27.
 
@@
 2020.
-``Hidden Analyses: A Review of Reporting Practice and Recommendations for More Transparent Reporting of Initial Data Analyses.'' \emph{BMC Medical Research Methodology} 20 (1): 61--61.
+{``Hidden Analyses: A Review of Reporting Practice and Recommendations for More Transparent Reporting of Initial Data Analyses.''} \emph{BMC Medical Research Methodology} 20 (1): 61--61.
 
@@
 \leavevmode\hypertarget{ref-KollerManuel2016rARP}{}%
-Koller, Manuel. 2016. ``Robustlmm: An R Package for Robust Estimation of Linear Mixed-Effects Models.'' \emph{Journal of Statistical Software} 75 (6): 1--24.
+Koller, Manuel. 2016. {``Robustlmm: An r Package for Robust Estimation of Linear Mixed-Effects Models.''} \emph{Journal of Statistical Software} 75 (6): 1--24.
 
 \leavevmode\hypertarget{ref-OECD}{}%
-OECD. n.d. ``Gender Wage Gap.'' https://data.oecd.org/earnwage/gender-wage-gap.htm.
+OECD. n.d. {``Gender Wage Gap.''} https://data.oecd.org/earnwage/gender-wage-gap.htm.
 
 \leavevmode\hypertarget{ref-rsvg}{}%
-Ooms, Jeroen. 2020. \emph{Rsvg: Render Svg Images into Pdf, Png, Postscript, or Bitmap Arrays}. \url{https://CRAN.R-project.org/package=rsvg}.
+Ooms, Jeroen. 2020. \emph{Rsvg: Render SVG Images into PDF, PNG, PostScript, or Bitmap Arrays}. \url{https://CRAN.R-project.org/package=rsvg}.
 
 \leavevmode\hypertarget{ref-opendata}{}%
-Open Knowledge Foundation. 2021. ``Open Definition. Defining Open in Open Data, Open Content, and Open Knowledge.'' 2021. \url{http://opendefinition.org/od/2.1/en/}.
+Open Knowledge Foundation. 2021. {``Open Definition. Defining Open in Open Data, Open Content, and Open Knowledge.''} 2021. \url{http://opendefinition.org/od/2.1/en/}.
 
@@
 2001.
-``Data Watch: The National Longitudinal Surveys.'' \emph{The Journal of Economic Perspectives} 15 (2): 239--53.
+{``Data Watch: The National Longitudinal Surveys.''} \emph{The Journal of Economic Perspectives} 15 (2): 239--53.
 
@@
 \leavevmode\hypertarget{ref-nlsy79}{}%
-The U.S. Bureau of Labor Statistics. n.d. ``National Longitudinal Survey of Youth 1979.'' Available at \url{https://www.nlsinfo.org/content/cohorts/nlsy79} (2021/25/02).
+The U.S. Bureau of Labor Statistics. n.d. {``National Longitudinal Survey of Youth 1979.''} Available at \url{https://www.nlsinfo.org/content/cohorts/nlsy79} (2021/25/02).
 
 \leavevmode\hypertarget{ref-brolgar}{}%
-Tierney, Nicholas, Di Cook, and Tania Prvan. 2020. \emph{Brolgar: BRowse over Longitudinal Data Graphically and Analytically in R}. \url{https://github.com/njtierney/brolgar}.
+Tierney, Nicholas, Di Cook, and Tania Prvan. 2020. \emph{Brolgar: BRowse over Longitudinal Data Graphically and Analytically in r}. \url{https://github.com/njtierney/brolgar}.
 
@@
 \leavevmode\hypertarget{ref-rlm}{}%
-UCLA: Statistical Consulting Group. 2021. ``Robust Regression \textbar{} R Data Analysis Examples.'' February 2021. \url{https://stats.idre.ucla.edu/r/dae/robust-regression/}.
+UCLA: Statistical Consulting Group. 2021. {``Robust Regression \textbar{} r Data Analysis Examples.''} February 2021. \url{https://stats.idre.ucla.edu/r/dae/robust-regression/}.
 
 \leavevmode\hypertarget{ref-validate}{}%
-van der Loo, Mark P. J., and Edwin de Jonge. 2021. ``Data Validation Infrastructure for R.'' \emph{Journal of Statistical Software} 97 (10): 1--31. \url{https://doi.org/10.18637/jss.v097.i10}.
+van der Loo, Mark P. J., and Edwin de Jonge. 2021. {``Data Validation Infrastructure for {R}.''} \emph{Journal of Statistical Software} 97 (10): 1--31. \url{https://doi.org/10.18637/jss.v097.i10}.
 
 \leavevmode\hypertarget{ref-mass}{}%
-Venables, W. N., and B. D. Ripley. 2002. \emph{Modern Applied Statistics with S}. Fourth. New York: Springer. \url{http://www.stats.ox.ac.uk/pub/MASS4}.
+Venables, W. N., and B. D. Ripley. 2002. \emph{Modern Applied Statistics with s}. Fourth. New York: Springer. \url{http://www.stats.ox.ac.uk/pub/MASS4}.
 
@@
 2020.
-``A New Tidy Data Structure to Support Exploration and Modeling of Temporal Data.'' \emph{Journal of Computational and Graphical Statistics} 29 (3): 466--78.
+{``A New Tidy Data Structure to Support Exploration and Modeling of Temporal Data.''} \emph{Journal of Computational and Graphical Statistics} 29 (3): 466--78.
 \url{https://doi.org/10.1080/10618600.2019.1695624}.
@@
 \leavevmode\hypertarget{ref-WickhamHadley2014TD}{}%
-Wickham, Hadley. 2014. ``Tidy Data.'' \emph{Journal of Statistical Software} 59 (10): 1--23.
+Wickham, Hadley. 2014. {``Tidy Data.''} \emph{Journal of Statistical Software} 59 (10): 1--23.
 
@@
 2019.
-``Welcome to the tidyverse.'' \emph{Journal of Open Source Software} 4 (43): 1686.
+{``Welcome to the {tidyverse}.''} \emph{Journal of Open Source Software} 4 (43): 1686.
 \url{https://doi.org/10.21105/joss.01686}.
@@
 \leavevmode\hypertarget{ref-mgcv}{}%
-Wood, S. N. 2003. ``Thin-Plate Regression Splines.'' \emph{Journal of the Royal Statistical Society (B)} 65 (1): 95--114.
+Wood, S. N. 2003. {``Thin-Plate Regression Splines.''} \emph{Journal of the Royal Statistical Society (B)} 65 (1): 95--114.
 
@@
 2014.
-``Knitr: A Comprehensive Tool for Reproducible Research in R.'' In \emph{Implementing Reproducible Computational Research}, edited by Victoria Stodden, Friedrich Leisch, and Roger D.
+{``Knitr: A Comprehensive Tool for Reproducible Research in {R}.''} In \emph{Implementing Reproducible Computational Research}, edited by Victoria Stodden, Friedrich Leisch, and Roger D.
 Peng.
@@
 \leavevmode\hypertarget{ref-kableExtra}{}%
-Zhu, Hao. 2019. \emph{KableExtra: Construct Complex Table with 'Kable' and Pipe Syntax}. \url{https://CRAN.R-project.org/package=kableExtra}.
+Zhu, Hao. 2019. \emph{kableExtra: Construct Complex Table with 'Kable' and Pipe Syntax}. \url{https://CRAN.R-project.org/package=kableExtra}.
+
+\end{CSLReferences}
 

### 2021-09-06 paper/paper.Rmd
Change filing structure.

--- 2021-09-01 Report/report_alt.Rmd
@@
 
-This part gives some examples of how the data might be used for data analysis in a textbook.
-We used `brolgar` [@brolgar] to perform exploratory data analysis.
-In the first example, we observed the relationship between wages and education regarding how the wages increase year by year.
-Using `key_slope` function in `brolgar`, we performed linear regression with the mean hourly wage as response variable and year as the predictor, extracted the regression slope of each ID, and plotted it according to the highest grade completed.
-
-Figure \ref{fig:slope-fig} shows that the higher the education completed does not necessarily increase the wages more from year to year.
-We can see that people who completed 4th grade happen to have relatively the same median of slope as people who completed 11th grade, even though it is probably because some respondents in this group had extremely high mean hourly wages.
-Moreover, the annual increasing wages of people who completed 6th grade to 11th grade are relatively the same.
-People who completed 12th grade tend to have a higher slope of the year than the other groups.
-Some outliers are also spotted in this group.
-Additionally, we can see that some respondents have a negative slope, meaning that their wages tend to decrease over the year.
-
 ```{r}
@@
                                     `12TH grade` = "12TH GRADE")) %>%
-  mutate(hgc_regroup = factor(hgc_regroup, levels = c("8TH grade or less", 
+  mutate(hgc_regroup = factor(hgc_regroup, levels = c("UNGRADED",
+                                                      "8TH grade or less", 
                                                       "9TH to 11TH grade", 
@@
 
-EDA:
+This part gives some examples of exploratory data analysis using the data.
+In this example, we are interested to explore how people experience changes in their wages over the years.
+Hence, we focus the analysis in their slope of wages.
+Using `key_slope` function in `brolgar` [@brolgar], we perform linear regression with the mean hourly wage as response variable and year as the predictor for each individual.
 
 Exploring the wages slope of each individual:
-  - By sex
-  - By education 
   - Exploring the wages slope pattern based on the number of observation
 
-- Slope by sex
-
-```{r}
+```{r slope-gender, fig.cap=""}
 slope_bottom_5<- wages_slope %>%
@@
   geom_point() +
-  geom_smooth(method = "lm") +
+  geom_smooth(method = "lm", se = FALSE) +
   facet_grid(note ~ rank) +
@@
   geom_point() +
-  geom_smooth(method = "lm") +
+  geom_smooth(method = "lm", se = FALSE) +
   facet_grid(note ~ rank)
@@
 ```
+
+Notes:
+
+Trimming the slope by 95% percentile and 5% percentile. 
+After that explore the slope after trimmed:
+- By sex (including the wage gender gap)
+- By education
 
@@
 
-For the final example, we showed modeling the longitudinal data using linear mixed model with `lme4` package [@lme4].
-
-We fitted a more flexible model to fit into the data since the relationship between mean hourly wage and year is not linear due to the spikes.
-Since this approach is computationally expensive, we only showed the model in a small fraction of the data.
-We sampled 1 percent of the total respondents using `brolgar's` `sample_frac_keys` function.
-Further, we plotted some IDs overlayed by the fitted GAM model in Figure \ref{fig:plot-gam}.
-We learn that the model is flexible enough to deal with the fluctuation in the data.
-Moreover, the data with small spikes tend to be linearly fitted.
-
-```{r plot-gam, cache=TRUE, echo=FALSE, fig.cap="Exploration of wages data by fitting a GAM.
-The fitted model displayed by blue line.
-The fitted line shows that the model flexible enough to follow the pattern of the data."}
-
-# build linear-mixed model
-wages_fit <- lmer(
-  ln_wages ~ year0*hgc_regroup*gender + (1|id),
-  data = wages_model_data
-)
-
-wages_fit_aug<- wages_model_data %>%
-  add_predictions(wages_fit, var = "pred") %>%
-  add_residuals(wages_fit, var = "resid")
-
-slope_bottom_12 <- wages_slope %>%
-  dplyr::select(id, .slope_year0) %>%
-  arrange(.slope_year0) %>%
-  slice_head(n = 12)
-
-wages_fit_aug2 <- wages_fit_aug %>%
-  filter(id %in% slope_bottom_12$id)
-
-ggplot(wages_fit_aug2, aes(x = year,
-             y = pred,
-             group = id,
-             colour = factor(id))) + 
-  geom_line() + 
-  geom_point(aes(x = year,
-                 y = ln_wages,
-                 colour = factor(id))) + 
-  facet_wrap(~id) + 
-  theme(legend.position = "none")
-
-```
-
 # Summary

### 2021-09-06 paper/report-alt.tex
Change filing structure.

% !TeX root = RJwrapper.tex
\title{The Journey from Wild to Text Book Data: A Case Study of the National
Longitudinal Survey of Youth}
\author{by }

\maketitle

NSLY79 is a prominent open data source that has been an important
support in multidisciplinary research on longitudinal data. Subsets of
this data can be found in numerous textbooks and research articles,
however the steps and decisions taken to get from the original data to
these subsets is never clearly articulated. This article describes our
journey when trying to re-create a textbook example data set from the
original data, with a goal to refresh it for use in the classroom. It
thus demonstrates the process of initial data analysis (IDA) -- tidying,
cleaning, and documenting the process -- to make data available for text
book examples or research. Three new data sets, and the code to produce
them are provided in an accompanying open source R package, called
\texttt{yowie}. The code might be useful for future refreshing of this
particular text book data. As a result of this process, some
recommendations are made for the NLSY79 curators for incorporating data
quality checks, and providing more convenient samples of the data to
potential users.

\textbf{Keywords}: Data cleaning; Data tidying; Longitudinal data;
NLSY79; Open data; Initial data analysis; Outlier detection; Robust
linear regression

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

``Open data'' is data that are freely accessible, modifiable, and
shareable by anyone for any purpose \citep{opendata}. This type of data
can be useful as example data in statistical text books and for research
purposes. However, open data are often, what we might call ``wild
data'', because it requires substantial cleaning and tidying to tame it
into text book shape. \citet{HuebnerMariannePhD2016Asat} emphasize that
making the data cleaning process accountable and transparent is
imperative. Documenting the data cleaning is essential for the integrity
of downstream statistical analyses and model building
\citep{HuebnerMarianne2020Haar}.

Data cleaning is part of what is called ``initial data analysis (IDA)''
\citep{Chatfield1985TIEo}. The other IDA steps are to explore the data
properties and check assumption of modeling to document and report the
process for the later formal analysis. Hence, exploratory data analysis,
which is defined by \citet{tukey} as a detective work to get the clue
from data, either numerically or graphically, before confirmatory data
analysis is performed, encompasses IDA. \citet{DasuTamraparni2003Edma}
say that data cleaning and exploration, without naming it as IDA, is a
difficult task and consumes 80\% of the data mining task.

Despite its importance, this IDA stage is often undervalued and
neglected \citep{Chatfield1985TIEo}. There are few research papers that
document the data cleaning \citep{WickhamHadley2014TD}. Furthermore, the
decisions made in this stage often go unreported in the sense that IDA
is often performed in an unplanned and unstructured way, and is only
shared among restricted parties \citep{HuebnerMarianne2020Haar}.

This paper aims to demonstrate the steps of IDA, tidying, cleaning, and
documenting the process, for a prominent open data source, National
Longitudinal Survey of Youth (NLSY) 1979, henceforth referred to NSLY79.
This data has been playing an important role in research in various
disciplines including but are not limited to economics, sociology,
education, public policy, and public health for more than a quarter of
the century \citep{MichaelRPergamit2001DWTN}. In addition, The National
Longitudinal Survey is considered as survey with high retention rates
and carefully designed making it suitable for a life course research
\citep{MichaelRPergamit2001DWTN} and \citep{eliznlsy}. According to
\citet{eliznlsy}, thousand of articles, and hundreds of book chapters
and monographs has incorporated the NLSY data. Moreover, the NLSY79 is
considered as the most widely used and most important cohort in the
NLSY79 data sets \citep{MichaelRPergamit2001DWTN}.

\citet{SingerJudithD2003Alda} used the wages and other variables of high
school dropouts from the NLSY79 data as an example data set to
illustrate longitudinal data modeling. Our aim is to refresh this text
book data to be more up to date. Here, we investigate the process of
getting from the raw NLSY79 to this text book data set. However, we are
not able to create the exact same data set as published in their book
since we do not have the information of what age threshold did they use
to determine the high school dropouts.

This paper is structured to have 5 sections. Section 2 describes the
original data source. Section 3 presents the steps of cleaning the data,
including getting and tidying the data from the NLSY79 and initial data
analysis to find and treat the anomalies in the data. We also gave
examples of exploratory data analysis using the clean data in Section 4.
Finally, Section 6 summarizes the paper.

\hypertarget{the-nlsy79}{%
\section{The NLSY79}\label{the-nlsy79}}

\hypertarget{database}{%
\subsection{Database}\label{database}}

The NLSY79 is a longitudinal survey held by the U.S Bureau of Labor
Statistics that follows the lives of a sample of American youth and born
between 1957-1964 \citep{nlsy79}. The cohort originally included 12,686
respondents ages 14-22 when first interviewed in 1979. It was comprised
of Blacks, Hispanics, economically disadvantaged non-Black
non-Hispanics, and youth in the military. In 1984 and 1990, two
sub-samples were dropped from the interview; the dropped subjects are
the 1,079 members of the military sample and 1,643 members of the
economically disadvantaged non-Black non-Hispanics respectively. Hence,
9,964 respondents remain in the eligible samples. The surveys were
conducted annually from 1979 to 1994 and biennially thereafter. Data are
now available from Round 1 (1979 survey year) to Round 28 (2018 survey
year).

Although the main focus area of the NLSY is labor and employment, the
NLSY also cover several other topics including education; training and
achievement; household, geography and contextual variables; dating,
marriage, cohabitation; sexual activity, pregnancy, and fertility;
children; income, assets and program participation; health; attitudes
and expectations; and crime and substance use.

There are two ways to conduct the interview of the NLSY79, which are
face to face interview or by telephone. In recent survey years, more
than 90 percent of respondents were interviewed by telephone
\citep{eliznlsy}.

\hypertarget{target-data}{%
\subsection{Target Data}\label{target-data}}

As mentioned in Section 1, this paper aims to refresh the NLSY79 data as
used in Singer and Willet's text book (\citet{SingerJudithD2003Alda}).
This data set contains the information of yearly mean hourly wages from
the NLSY79 cohort, with education and race as covariates. This data set
represents the measurement of male high-school dropouts, aged from 14 to
17 years old when first time measured. Hence, in this paper, after
getting the data set of high school graduates, we subset it by only
having the high school dropouts.

Since there is no information of high school drop-outs in Singer and
Willet's book, we define high school dropouts as respondents who only
completed either \(9^{th}\), \(10^{th}\), \(11^{th}\) grade, or who
completed \(12^{th}\) when their age are at least 19 years old (means
that these IDs being dropped out in certain year, but they returned back
to high school to complete it).

\hypertarget{the-nlsy79-data-cleaning}{%
\section{The NLSY79 Data Cleaning}\label{the-nlsy79-data-cleaning}}

\hypertarget{getting-and-tidying-the-data}{%
\subsection{Getting and Tidying the
Data}\label{getting-and-tidying-the-data}}

The NLSY79 data are stored in a
\href{https://www.nlsinfo.org/content/cohorts/nlsy79/get-data}{database}
and could be downloaded by variables. Several variables are available
for download and could be browsed by index. For the wages data set, we
only extracted these variables:

\begin{itemize}
\tightlist
\item
  Education, Training \& Achievement Scores

  \begin{itemize}
  \tightlist
  \item
    Education -\textgreater{} Summary measures -\textgreater{} All
    schools -\textgreater{} By year -\textgreater{} Highest grade
    completed

    \begin{itemize}
    \tightlist
    \item
      Downloaded all of the 78 variables in Highest grade completed.
    \end{itemize}
  \end{itemize}
\item
  Employment

  \begin{itemize}
  \tightlist
  \item
    Summary measures -\textgreater{} By job -\textgreater{} Hours worked
    and Hourly wages

    \begin{itemize}
    \tightlist
    \item
      Downloaded all of the 427 variables in Hours worked
    \item
      Downloaded all of the 151 variables in Hourly wages
    \end{itemize}

    Both hours worked and hourly wages are recorded by the job, up to
    five jobs for each id/subject.
  \end{itemize}
\item
  Household, Geography \& Contextual Variables

  \begin{itemize}
  \tightlist
  \item
    Context -\textgreater{} Summary measures -\textgreater{} Basic
    demographics

    \begin{itemize}
    \tightlist
    \item
      Downloaded year and month of birth, race, and sex variables.
    \end{itemize}

    There are two versions of the year and month of birth, i.e.~1979 and
    1981 data. We downloaded these two versions.
  \end{itemize}
\end{itemize}

The downloaded data set came in a csv (NLSY79.csv) and dat (NLSY79.dat)
format. We only used the .dat format. It also came along with these
files:

\begin{itemize}
\tightlist
\item
  NLSY79.NLSY79: This is the tagset of variables that can be uploaded to
  the web site to recreate the data set.
\item
  NLSY79.R: This is an R script provided automatically by the database
  for reading the data into R and convert the variables' name and its
  label into something more sensible. We utilized this code to do an
  initial data tidying. It produced two data set,
  \texttt{categories\_qnames} (the observations are stored in
  categorical/interval values) and \texttt{new\_data\_qnames} (the
  observations are stored in integer form).
\end{itemize}

According to \citet{WickhamHadley2014TD}, a tidy data sets comply with
three rules, the first is that each variable forms a column, the second
is that each observation forms a row, and the last is that each type of
observational unit forms a table. Unfortunately, the
\texttt{new\_data\_qnames} did not meet these requirements in the way
that the value for particular year and job are stored in different
columns, hence the data contains a huge number of columns (686 columns).
The example of the untidy of the data set is displayed in Table
\ref{tab:untidy-data}, which actually intended to display the hourly
rate of each respondent by job (HRP1 to HRP5) and by year (1979 and
1980). The table implies that the column headers are values of the year
and job, not variable names.

\begin{Schunk}
\begin{table}

\caption{\label{tab:untidy-data}The Untidy Form of the NLSY79 Raw Data}
\centering
\begin{tabular}[t]{r|r|r|r|r|r|r}
\hline
CASEID\_1979 & HRP1\_1979 & HRP2\_1979 & HRP3\_1979 & HRP4\_1979 & HRP5\_1979 & HRP1\_1980\\
\hline
\rowcolor{gray!6}  1 & 328 & NA & NA & NA & NA & NA\\
\hline
2 & 385 & NA & NA & NA & NA & 457\\
\hline
\rowcolor{gray!6}  3 & 365 & NA & 275 & NA & NA & 397\\
\hline
4 & NA & NA & NA & NA & NA & NA\\
\hline
\rowcolor{gray!6}  5 & 310 & 375 & NA & NA & NA & 333\\
\hline
6 & NA & NA & NA & 250 & NA & 275\\
\hline
\end{tabular}
\end{table}

\end{Schunk}

Consequently, the data should be tidied and wrangled first to extract
the demographic and employment variables that we want to put in the
final data set. We mainly used \texttt{tidyr} \citep{tidyr},
\texttt{dplyr} \citep{dplyr}, and \texttt{stringr} \citep{stringr} to do
this job.

\hypertarget{tidying-demographic-variables}{%
\subsubsection{Tidying demographic
variables}\label{tidying-demographic-variables}}

\texttt{Age\ in\ 1979}, \texttt{gender}, \texttt{race},
\texttt{highest\ grade\ completed} (factor and integer), and the
\texttt{year\ when\ the\ highest\ grade\ completed} are the variables
that we want to use in the cleaned data set.

Age in 1979 are derived from year of birth and month of birth variables
in the raw data set. The variables have two versions, which are the 1979
version and the 1981 version. We only used the 1979 data and did a
consistency check of those years and flag the inconsistent observations.

\begin{Schunk}
\begin{Sinput}
## tidy the date of birth data
dob_tidy <- new_data_qnames %>%
  dplyr::select(CASEID_1979,
         starts_with("Q1-3_A~")) %>%
  mutate(dob_year = case_when(
                    # if the years recorded in both sets match, take 79 data
                    `Q1-3_A~Y_1979` == `Q1-3_A~Y_1981` ~ `Q1-3_A~Y_1979`,
                    # if the year in the 81 set is missing, take the 79 data
                    is.na(`Q1-3_A~Y_1981`) ~ `Q1-3_A~Y_1979`,
                    # if the sets don't match for dob year, take the 79 data
                    `Q1-3_A~Y_1979` != `Q1-3_A~Y_1981` ~ `Q1-3_A~Y_1979`),
        dob_month = case_when(
                    # if months recorded in both sets match, take 79 data
                    `Q1-3_A~M_1979` == `Q1-3_A~M_1981` ~ `Q1-3_A~M_1979`,
                    # if month in 81 set is missing, take the 79 data
                    is.na(`Q1-3_A~M_1981`) ~ `Q1-3_A~M_1979`,
                    # if sets don't match for dob month, take the 79 data
                    `Q1-3_A~M_1979` != `Q1-3_A~M_1981` ~ `Q1-3_A~M_1979`),
        # flag if there is a conflict between dob recorded in 79 and 81
        dob_conflict = case_when(     
                      (`Q1-3_A~M_1979` != `Q1-3_A~M_1981`) & !is.na(`Q1-3_A~M_1981`)
                      ~ TRUE,
                      (`Q1-3_A~Y_1979` != `Q1-3_A~Y_1981`) & !is.na(`Q1-3_A~Y_1981`)
                      ~ TRUE,
                      (`Q1-3_A~Y_1979` == `Q1-3_A~Y_1981`) & 
                      (`Q1-3_A~M_1979` == `Q1-3_A~M_1981`) ~ FALSE,
                      is.na(`Q1-3_A~M_1981`) | is.na(`Q1-3_A~Y_1981`) ~ FALSE)) %>%
  dplyr::select(CASEID_1979,
         dob_month,
         dob_year,
         dob_conflict)
\end{Sinput}
\end{Schunk}

For \texttt{gender} and \texttt{race}, we only renamed these variables.

\begin{Schunk}
\begin{Sinput}
## tidy the gender and race variables
demog_tidy <- categories_qnames %>%
  dplyr::select(CASEID_1979,
         SAMPLE_RACE_78SCRN,
         SAMPLE_SEX_1979) %>%
  rename(gender = SAMPLE_SEX_1979,
         race = SAMPLE_RACE_78SCRN)
\end{Sinput}
\end{Schunk}

The next step is tidying the \texttt{highest\ grade\ completed}
variable. This variable came with several version in each year. We chose
the revised May data because it seemed to have less missing and
presumably has been checked. However, there was no revised May data for
2012, 2014, 2016, and 2018 so we just used the ordinary May data for
these years. The code for tidying this variable is provided in the
supplementary code whose details could be seen in the Supplementary
Materials section of this paper.

Further, \texttt{highest\ grade\ completed} is measured and could be
updated in each period of the survey. Hence, we only used the highest
grade completed ever and derived the year when it is completed. The code
for tidying this variable is also available in the supplementary code of
this paper.

Finally, we join all of the tidy variables in a data set called
\texttt{full\_demographics} as displayed as follows:

\begin{Schunk}
\begin{Soutput}
#>   id dob_month dob_year dob_conflict                    race gender yr_hgc
#> 1  1         9       58        FALSE NON-BLACK, NON-HISPANIC FEMALE   1979
#> 2  2         1       59        FALSE NON-BLACK, NON-HISPANIC FEMALE   1985
#> 3  3         8       61        FALSE NON-BLACK, NON-HISPANIC FEMALE   1993
#> 4  4         8       62        FALSE NON-BLACK, NON-HISPANIC FEMALE   1986
#> 5  5         7       59        FALSE NON-BLACK, NON-HISPANIC   MALE   1984
#> 6  6        10       60        FALSE NON-BLACK, NON-HISPANIC   MALE   1983
#>   hgc_i          hgc
#> 1    12   12TH GRADE
#> 2    12   12TH GRADE
#> 3    12   12TH GRADE
#> 4    14 2ND YEAR COL
#> 5    18 6TH YEAR COL
#> 6    16 4TH YEAR COL
\end{Soutput}
\end{Schunk}

\hypertarget{tidying-employment-variables}{%
\subsubsection{Tidying employment
variables}\label{tidying-employment-variables}}

The employment data comprises of three variables,
i.e.~\texttt{total\ hours\ of\ work\ per\ week},
\texttt{number\ of\ jobs\ that\ an\ individual\ has}, and
\texttt{mean\ hourly\ wage}.

\texttt{hours\ worked\ per\ week} initially has only one version per
job, no choice from 1979 to 1987 (QES-52A). From 1988 onward, when we
had more options, we chose the variable for total hours including time
spent working from home (QES-52D). However, in 1993, this variable did
not have all the five D variables (the first one and the last one were
missing), so we used QES-52A variable instead. In addition, 2008 only
had jobs 1-4 for the QES-52D variable (whereas the other years had 1-5),
so we just used these.

\begin{Schunk}
\begin{Sinput}
# make a list for years where we used the "QES-52A"
year_A <- c(1979:1987, 1993)
#function to get the hour of work
get_hour <- function(year){
  if(year %in% year_A){
   temp <- new_data_qnames %>%
    dplyr::select(CASEID_1979,
            starts_with("QES-52A") &
              ends_with(as.character(year)))} 
  else{
    temp <- new_data_qnames %>%
    dplyr::select(CASEID_1979,
            starts_with("QES-52D") &
              ends_with(as.character(year)))} 
    temp %>% 
      pivot_longer(!CASEID_1979,
                 names_to = "job",
                 values_to = "hours_work") %>%
      separate("job", c("job", "year"), sep = -4) %>%
      mutate(job = paste0("job_", substr(job, 9, 10))) %>%
      rename(id = CASEID_1979)
}

# list to save the iteration result
hours <- list()
# getting the hours of work of all observations
for(ayear in c(1979:1994, 1996, 1998, 2000, 2002, 2004, 2006, 2008, 2010, 
               2012, 2014, 2016, 2018)) {
   hours[[ayear]] <- get_hour(ayear)
}
# unlist the hours of work
hours_all <- bind_rows(!!!hours)
\end{Sinput}
\end{Schunk}

The same way was also deployed to tidy the rate of wage by year and by
ID. The difference is that the hourly rate had only one version of each
year. The hours of work and the hourly rate were then joined to
calculate the number of jobs that a respondent has and their mean hourly
wage. Some observations had 0 in their hourly rate, which is considered
as invalid value. Thus, their hourly rate set to be N.A. the result of
the joining variables could be seen in the following tibble.

\begin{Schunk}
\begin{Soutput}
#> # A tibble: 6 x 5
#>      id job    year  rate_per_hour hours_work
#>   <int> <chr>  <chr>         <int>      <int>
#> 1     1 job_01 1979            328         38
#> 2     1 job_02 1979             NA         15
#> 3     1 job_03 1979             NA         NA
#> 4     1 job_04 1979             NA         NA
#> 5     1 job_05 1979             NA         NA
#> 6     2 job_01 1979            385         35
\end{Soutput}
\end{Schunk}

Since our ultimate goal is to calculate the mean hourly wage, the number
of jobs is calculate based on the availability of the
\texttt{rate\_per\_hour} information. For example, the number of jobs of
ID 1, based on \texttt{hours\_work}, is 2. However, since the
information of hourly rate of \texttt{job\_02} is not available, the
number of job is considered as 1.

Further, we calculated the mean hourly wage for each ID in each year
using a weighted mean with the hours of work as the weight. However,
there are a lot of missing value in \texttt{hours\_work} variable. In
that case, we only calculated the mean hourly wage based on
arithmetic/regular mean method. Hence, we created a new variable to flag
whether the mean hourly wage is a weighted or a regular mean.
Additionally, if an ID only had one job, we directly used their hourly
wages information and flagged it as an arithmetic mean. The following
tibble shows the result of this tidying process.

\begin{Schunk}
\begin{Soutput}
#> # A tibble: 10 x 6
#>       id  year mean_hourly_wage total_hours number_of_jobs is_wm
#>    <int> <dbl>            <dbl>       <int>          <dbl> <lgl>
#>  1     1  1979             3.28          38              1 FALSE
#>  2     1  1981             3.61          NA              1 FALSE
#>  3     2  1979             3.85          35              1 FALSE
#>  4     2  1980             4.57          NA              1 FALSE
#>  5     2  1981             5.14          NA              1 FALSE
#>  6     2  1982             5.71          35              1 FALSE
#>  7     2  1983             5.71          NA              1 FALSE
#>  8     2  1984             5.14          NA              1 FALSE
#>  9     2  1985             7.71          NA              1 FALSE
#> 10     2  1986             7.69          NA              1 FALSE
\end{Soutput}
\end{Schunk}

The \texttt{mean\_hourly\_wage} and \texttt{full\_demographic} data are
then joined. We also filtered the data to only have the cohort who
completed the education up to 12th grade and participated at least five
rounds in the survey and save it to an object called
\texttt{wages\_demog\_hs}.

\hypertarget{initial-data-analysis}{%
\subsection{Initial Data Analysis}\label{initial-data-analysis}}

According to \citet{HuebnerMariannePhD2016Asat}, Initial Data Analysis
(IDA) is the step of inspecting and screening the data after being
collected to ensure that the data is clean, valid, and ready to be
deployed in the later formal statistical analysis. Moreover,
\citet{Chatfield1985TIEo} argued that the two main objectives of IDA is
data description, which is to assess the structure and the quality of
the data; and model formulation without any formal statistical
inference.

In this paper, we conducted an IDA or a preliminary data analysis to
assess the consistency of the data with the cohort information that is
provided by the NLSY. In addition, we also aimed to find the anomaly in
the wages values using this approach. We mainly used graphical summary
to do the IDA using \texttt{ggplot2}\citep{ggplot2} and \texttt{brolgar}
\citep{brolgar}.

As stated previously, the respondents' ages ranged from 12 to 22 when
first interviewed in 1979. Hence, we would like to validate whether all
of the respondents were in this range. Additionally, the
\href{https://www.nlsinfo.org/content/cohorts/nlsy79/intro-to-the-sample/nlsy79-sample-introduction}{NLSY}
also provided the number of the survey cohort by their gender (6,403
males and 6,283 females) and race (7,510 Non-Black/Non-Hispanic; 3,174
Black; 2,002 Hispanic). To validate this, we used the
\texttt{full\_demographic} i.e.~the data with the survey years 1979
sample. Table \ref{tab:age-table} and Table \ref{tab:gender-race-table}
suggest that the demographic data we had is consistent with the sample
information in the database.

\begin{Schunk}
\begin{table}

\caption{\label{tab:age-table}Age Distribution of the NLSY79 samples}
\centering
\begin{tabular}[t]{r|r}
\hline
Age & Number of Sample\\
\hline
\rowcolor{gray!6}  15 & 1265\\
\hline
16 & 1550\\
\hline
\rowcolor{gray!6}  17 & 1600\\
\hline
18 & 1530\\
\hline
\rowcolor{gray!6}  19 & 1662\\
\hline
20 & 1722\\
\hline
\rowcolor{gray!6}  21 & 1677\\
\hline
22 & 1680\\
\hline
\end{tabular}
\end{table}

\end{Schunk}

\begin{Schunk}
\begin{table}

\caption{\label{tab:gender-race-table}Gender and Race Distribution of the NLSY79 Samples}
\centering
\begin{tabular}[t]{l|l|l|l|l}
\hline
\multicolumn{1}{c|}{ } & \multicolumn{3}{c|}{Race} & \multicolumn{1}{c}{ } \\
\cline{2-4}
Gender & Hispanic & Black & Non-Black, Non-Hispanic & Total\\
\hline
\rowcolor{gray!6}  Male & 1000 (15.62\%) & 1613 (25.19\%) & 3790 (59.19\%) & 6403 (100.00\%)\\
\hline
Female & 1002 (15.95\%) & 1561 (24.84\%) & 3720 (59.21\%) & 6283 (100.00\%)\\
\hline
\rowcolor{gray!6}  Total & 2002 (15.78\%) & 3174 (25.02\%) & 7510 (59.20\%) & 12686 (100.00\%)\\
\hline
\end{tabular}
\end{table}

\end{Schunk}

\begin{Schunk}
\begin{figure}
\includegraphics{figures/featureplot-1} \caption[Two plots showing the distribution of the mean hourly wage]{Two plots showing the distribution of the mean hourly wage. Plot A portrays the pattern of mean hourly wage of high school cohort from 1979 to 2018 of each ID in US Dollar; Plot B shows the distribution of their minimum, median, and maximum value. We can see that some IDs had an extremely high of wages and it made the distribution of the three features is extremely skewed.}\label{fig:featureplot}
\end{figure}
\end{Schunk}

The next step is that we explored the mean hourly wage data, in this
case, we only explored the wages data in \texttt{wages\_demog\_hs}.
Figure @ref(fig:featureplot) conveys that there is clearly a problem in
the mean hourly wage values. Figure \ref{fig:featureplot} A shows that
some observations had an exceptionally high figure of wages, even more
than US\$10,000 per hour. In Figure \ref{fig:featureplot} B, we barely
see any difference in the minimum, median, an maximum value of the wages
since the distribution is heavily skewed to the right. Additionally,
Table \ref{tab:summarytable} shows that the overall wages median of the
cohort is only 7.2, while the mean is 11.87. It indicates that the data
might contains a lot of extreme values.

\begin{Schunk}
\begin{table}

\caption{\label{tab:summarytable}Summary Statistics of Wages of High School Data}
\centering
\begin{tabular}[t]{l|r}
\hline
Statistics & Value\\
\hline
\rowcolor{gray!6}  Min. & 0.01000\\
\hline
1st Qu. & 4.50000\\
\hline
\rowcolor{gray!6}  Median & 7.20000\\
\hline
Mean & 11.86578\\
\hline
\rowcolor{gray!6}  3rd Qu. & 11.74000\\
\hline
Max. & 33400.00000\\
\hline
\end{tabular}
\end{table}

\end{Schunk}

In Figure \ref{fig:high-wages}, we plotted some respondents with a high
value of mean hourly wages. We filtered all of the IDs who earned more
than US\$ 500 per hour averagely. We found that most of these
respondents only experienced one point of extremely high wages. Thus, we
suspected that these high values are erroneous values resulted from a
data entry error.

\begin{Schunk}
\begin{figure}
\includegraphics{figures/high-wages-1} \caption[Some observations with extremely high mean hourly wage]{Some observations with extremely high mean hourly wage. Most of the IDs only have one point of high wage.}\label{fig:high-wages}
\end{figure}
\end{Schunk}

Further, we took 36 samples randomly from the data and plotted it as is
seen in Figure \ref{fig:sampleplot}. It implies that not only that some
observations earned an extremely high figures of wages, but some also
had a reasonably fluctuate wages, for example the IDs in panel number 5,
7, and 11. The plot also implies that the samples had a different
pattern of mean hourly wages. Some had a flat wages for years but had a
sudden increase in a year than it went down again, while the other
experienced a upsurge in their wage, for instance the IDs in panel 9.

According to \citet{MichaelRPergamit2001DWTN}, one of the flaws of the
NLSY79 employment data is that since the NLSY79 collect the information
of the working hours since the last interview, it might be challenging
for the respondents to track the within-job hours changes that happens
between survey year, especially for the respondents with fluctuate
working hours or whose job is seasonal. It even has been more
challenging since 1994, where the respondents had to recall two years
period. This shortcoming might also contribute on the fluctuation of
one's wages data.

\begin{Schunk}
\begin{figure}
\includegraphics{figures/sampleplot-1} \caption[The mean hourly wages of some random samples are shown in twelve facets, three IDs per facet]{The mean hourly wages of some random samples are shown in twelve facets, three IDs per facet. It suggests that some IDs had a reasonably fluctuate wages.}\label{fig:sampleplot}
\end{figure}
\end{Schunk}

\hypertarget{replacing-extreme-values}{%
\subsubsection{Replacing extreme
values}\label{replacing-extreme-values}}

As it is seen from figure \ref{fig:sampleplot}, there are many spikes in
the mean hourly wage data. As part of the IDA, which is the model
formulation, we built a robust linear regression model to address this
issue. The notion of robust linear regression is to yield an estimation
that is robust to the influence of noise or contamination
\citep{KollerManuel2016rARP}. It also aims to detect the contamination
by weighting each observation based on how ``well-behave'' they are,
known as robustness weight. Observations with lower robustness weight
are suggested as an outliers by this method
\citep{KollerManuel2016rARP}.

In this paper, we built the model using the \texttt{rlm} function from
\texttt{MASS} package \citep{mass}. We set the
\texttt{mean\_hourly\_wage} and \texttt{year} as the dependent and
predictor respectively. Furthermore, we used M-Estimation with Huber
weighting where the observation with small residual get a weight of 1,
while the larger the residual, the smaller the weight (less than 1)
\citep{rlm}.

Since we worked with longitudinal data, we should built the model for
each ID, instead of the overall data. The robust mixed model is actually
the best model to be employed in this case. However, this method is too
computationally and memory expensive, especially for a large data set,
like the NLSY79 data. Thus, the model for each ID is built utilizing the
\texttt{nest} and \texttt{map} function from \texttt{tidyr}
\citep{tidyr} and \texttt{purrr} \citep{purrr} respectively.

The challenging part of detecting the anomaly using the robustness
weight is to determine the threshold of the weight where the
observations considered as outliers. Moreover, it should be noted that
not all the outliers is due to an error, instead it might be that one
had a reasonably increasing or decreasing wages. To, minimize the risk
of being mistakenly regard an outlier as an ``erronous outlier'', we
have simulated some threshold and study the behavior of the spikes in
each threshold. We found that 0.12 is the most reasonable value to be
the threshold to minimize the risk of that drawback because it still
capture the sensible spikes in the data. In other words, we kept
maintaining the natural variability of the wages while minimizing the
presence of anomalies because of the error in the data recording. After
deciding the threshold, we imputed the observations with weight less
than 0.2 wage with the models' predicted value. We then flagged those
observations in a new variable called \texttt{is\_pred},

\begin{Schunk}
\begin{Sinput}
# nest the data by id to build a robust linear model
by_id <- wages_demog_hs %>%
  dplyr::select(id, year, mean_hourly_wage) %>%
  group_by(id) %>%
  nest()

# build a robust linear model
id_rlm <- by_id %>%
  mutate(model = map(.x = data,
                     .f = function(x){
                       rlm(mean_hourly_wage ~ year, data = x)
                     }))
# extract the property of the regression model
id_aug <- id_rlm %>%
  mutate(augmented = map(model, broom::augment)) %>%
  unnest(augmented)

# extract the weight of each observation
id_w <- id_rlm %>%
  mutate(w = map(.x = model,
                 .f = function(x){
                   x$w
                 })) %>%
  unnest(w) %>%
  dplyr::select(w)

# bind the property of each observation with their weight
id_aug_w <- cbind(id_aug, id_w) %>%
  dplyr::select(`id...1`,
                year,
                mean_hourly_wage,
                .fitted,
                .resid,
                .hat,
                .sigma,
                w) %>%
  rename(id = `id...1`)

# if the weight < 1, the mean_hourly_wage is replaced by the model's fitted/predicted value.
# and add the flag whether the observation is predicted value or not.
# since the fitted value is sometimes <0, and wages value could never be negative,
# we keep the mean hourly wage value even its weight < 1.

wages_rlm_dat <- id_aug_w %>%
  mutate(wages_rlm = ifelse(w < 0.12  & .fitted >= 0, .fitted,
                            mean_hourly_wage)) %>%
  mutate(is_pred = ifelse(w < 0.12 & .fitted >= 0, TRUE,
                          FALSE)) %>%
  dplyr::select(id, year, wages_rlm, is_pred)

# join back the `wages_rlm_dat` to `wages_demog_hs`

wages_demog_hs <- left_join(wages_demog_hs, wages_rlm_dat, by = c("id", "year"))
\end{Sinput}
\end{Schunk}

Figure \ref{fig:comppict} A shows that after imputing the anomalies with
the models' predicted value, the highest wages value has decreased to be
around \$350. The spikes were still observed, but are not as extreme as
the original data set. In Figure \ref{fig:comppict} B), we plotted the
three features of mean hourly wages, namely the minimum, median, an
maximum value, transformed to log scale. The plot implies that the
skewness are slightly negative. We also see that the three features are
overlapped each other. It indicates that some ID's minimum wage are
higher than some ID's maximum wage.

Furthermore, Figure \ref{fig:compare} implies that after the treatment,
the fluctuation can still be observed in the data and only the large
spikes, which are considered as ``erronous outliers'', are eliminated
from the data. Hence, the model results a data set with the reasonable
degree of fluctuation.

\begin{Schunk}
\begin{figure}
\includegraphics{figures/comppict-1} \caption[The distribution of the mean hourly wage after treating the extreme values]{The distribution of the mean hourly wage after treating the extreme values. Plot A portrays the pattern of mean hourly wage of high school cohort from 1979 to 2018 of each ID in US Dollar; Plot B shows the distribution of their minimum, median, and maximum value transformed to log10 scale. We can see that some observations still had reasonbaly higher wages than the others. Also, some IDs' have a minimum wages that is higher than others' maximum wages.}\label{fig:comppict}
\end{figure}
\end{Schunk}

\begin{Schunk}
\begin{figure}
\includegraphics{figures/compare-1} \caption[Comparison between the original and the treated mean hourly wage]{Comparison between the original and the treated mean hourly wage. The orange line portray the original value of mean hourly wage, while the turquoise line display the mean hourly wages value after the extreme values imputed with the robust linear model's prediction value. We can see that some extreme spikes has been reduced by the model.}\label{fig:compare}
\end{figure}
\end{Schunk}

Finally, we saved the imputed data and set the appropriate data type for
the variables. As our target data is the mean hourly wage of the
high-school dropouts, we then subset the high-school graduate data set
to be only having the male-high school dropout data. We also saved the
NLSY79 cohort's demographic information in a separate dataset. We then
make these three data sets and its processing documentation to be
publicly available through an R data container package called
\texttt{yowie}. The complete flow from the raw data to these data set is
displayed in Figure \ref(fig:flowchart).

\begin{Schunk}
\begin{Sinput}
# select out the old value of mean hourly wage and change it with the wages_rlm value
wages_demog_hs <- wages_demog_hs %>%
  dplyr::select(-mean_hourly_wage) %>%
  rename(mean_hourly_wage = wages_rlm)

# rename and select the wages in tidy
wages_hs2020 <- wages_demog_hs %>%
  dplyr::select(id, year, mean_hourly_wage, age_1979, gender, race, hgc, hgc_i, yr_hgc,
                number_of_jobs, total_hours, is_wm, is_pred) %>%
  mutate(hgc = as.factor(hgc),
         year = as.integer(year),
         age_1979 = as.integer(age_1979),
         yr_hgc = as.integer(yr_hgc),
         number_of_jobs = as.integer(number_of_jobs))

# Create a data set for demographic variables
demographic_nlsy79 <- full_demographics %>%
  mutate(age_1979 = 1979 - (dob_year + 1900)) %>%
  dplyr::select(id,
         age_1979,
         gender,
         race,
         hgc,
         hgc_i,
         yr_hgc) %>%
  mutate(age_1979 = as.integer(age_1979),
         hgc = as.factor(hgc),
         yr_hgc = as.integer(yr_hgc))

# Create a data set for the high school dropouts cohort
wages_hs_dropout <- wages_hs2020 %>%
  mutate(dob = 1979 - age_1979,
         age_hgc = yr_hgc - dob) %>%
  filter((hgc %in% c("9TH GRADE",
                     "10TH GRADE",
                     "11TH GRADE")) |
          (hgc == "12TH GRADE" &
              age_hgc > 19)) %>%
  filter(age_1979 <= 17,
         gender == "MALE") %>%
  dplyr::select(-dob,
         -age_hgc)
\end{Sinput}
\end{Schunk}

\begin{Schunk}
\begin{figure}
\includegraphics{figures/flowchart-1} \caption[The stages of data filtering from the raw data to get three datasets that are contained in an R package called yowie, n means the number of ID, while n_obs means the number of observations]{The stages of data filtering from the raw data to get three datasets that are contained in an R package called yowie, n means the number of ID, while n_obs means the number of observations.}\label{fig:flowchart}
\end{figure}
\end{Schunk}

\hypertarget{exploratory-data-analysis}{%
\section{Exploratory Data Analysis}\label{exploratory-data-analysis}}

In this part, we gave some examples of how the data might be used for
data analysis in text book. We used \texttt{brolgar} \citep{brolgar} to
perform exploratory data analysis. The first example is that we observed
relationship between wages and education in terms of how the wages
increase year by year. Using \texttt{key\_slope} function in
\texttt{brolgar}, we performed a liner regression with mean hourly wage
as response variable and year as predictor and extracted the regression
slope of each ID and plotted it according to highest grade completed.

We found in Figure \ref{fig:wages-slope} that the higher the education
completed does not necessarily increase the wages more from year to
year. We can see that people who completed 4th grade happen to have
relatively the same median of slope as people who completed 11th grade,
even though it is probably because some respondents in this group had an
extremely high mean hourly wages. Moreover, the annual increasing wages
of people who completed 6th grade to 11th grade are relatively same.
People who completed 12th grade tend to have higher slope of year
compared to the other groups. Some outliers are also spotted in this
group. Additionally, we can see that some respondents have a negative
slope, meaning that their wages tend to decrease over the year.

\begin{Schunk}
\begin{figure}
\includegraphics{figures/wages-slope-1} \caption[Regression slope of mean hourly wage regression by highest grade completed]{Regression slope of mean hourly wage regression by highest grade completed. The regression model has mean hourly wage as response variable and year as predictor. Respondents who completed 12th grade has the highest median of slope.}\label{fig:wages-slope}
\end{figure}
\end{Schunk}

In the second example, we examined the pattern of mean hourly wage
corresponding to different groups of gender and race. We took 240
samples of female and male equally and spread it into twelve facets as
is seen in Figure \ref{fig:gender-plot} using \texttt{facet\_strata}
function from \texttt{brolgar} \citep{brolgar}. We learn that the mean
hourly wage of female and male fluctuated over the years. In some
facets, for example in facet 4, 5, and 10, males tend to earn more wages
than females. This finding is also shown in Figure \ref{fig:gender-se}
where we plotted the median wage of males and females against the year.
We learn that males had a higher mean of mean hourly wage over the
period of survey than females. Unfortunately, the gap between them tend
to be wider from time to time.

\begin{Schunk}
\begin{figure}
\includegraphics{figures/gender-plot-1} \caption[The pattern of mean hourly wages overtime displayed in 12 facets coloured by gender]{The pattern of mean hourly wages overtime displayed in 12 facets coloured by gender. The respondents belong to high school cohort are randomly sampled. We learn that both male and female respondents had a fluctuate mean hourly wage over the year.}\label{fig:gender-plot}
\end{figure}
\end{Schunk}

\begin{Schunk}
\begin{figure}
\includegraphics{figures/gender-se-1} \caption[The yearly mean of mean hourly wage by gender along with its standard error]{The yearly mean of mean hourly wage by gender along with its standard error. Males generally earn more wages than females. The error interval tend to be wider over the years.}\label{fig:gender-se}
\end{figure}
\end{Schunk}

For the final example, we show the example of modeling the longitudinal
data using Generelized Additive Model (GAM) with \texttt{mgcv} package
\citep{mgcv} for each ID. We fit a more flexible model to fit in to the
data since the relationship between mean hourly wage and year is not
linear due to the spikes. Since this approach is computationally
expensive, we only show the model in a little fraction of the data. We
sample 1 percent of the total respondents using
\texttt{brolgar\textquotesingle{}s} \texttt{sample\_frac\_keys}
function. Further, we plotted the some sample of the fitted model in
Figure @ref\{fig:gam-plot\}. We learn that the model flexible enough to
deal with the fluctuation in the data. Moreover, model with small spikes
tend to be linearly fitted.

\begin{Schunk}
\begin{figure}
\includegraphics{figures/gam-plot-1} \caption[Exploration of wages data by fitting a GAM]{Exploration of wages data by fitting a GAM. The fitted model displayed by blue line. The fitted line shows that the model flexible enough to follow the pattern of the data.}\label{fig:gam-plot}
\end{figure}
\end{Schunk}

\hypertarget{summary}{%
\section{Summary}\label{summary}}

This paper has performed a set of stages to make an open data suitable
for a text book data or to make it ready for research. In the first
stage, we showed the steps that are performed to get the data from the
NLSY79 database. Since the data format is untidy, we showed how the data
has been tidied. After that, we conducted an initial data analysis to
investigate and screen the quality of the data. Using robust linear
regression model, we found and fixed the anomalous observations in the
data set with its predicted values. We also performed an example of
exploratory data analysis using the cleaned data set.

This paper also has demonstrated the documentation of data cleaning by
providing all of the codes in performing data tidying and initial data
analysis. Thus, this paper provided an opportunity to continuously
refresh the text book data whenever the updated data is published in the
NLSY79 database. This could be done by following the documentation of
the code that is provided in this paper.

Moreover, the documentation also includes the process of how did we
generate the robustness weight and how did we decide the threshold of
anomalous observations. It is also documented along with the flag of
whether an observation is imputed value or not. Accordingly, if somebody
wish to make another decision, it can be done by making a small changes
in the code provided. Further, this paper is also supplemented by a
\texttt{shiny} \citep{shiny} app as a simulation tool to customize the
weight threshold.

Finally, this paper implies that data providers should design the
database that is able to produce tidy data sets. A data provider should
also check for data anomalies prior the data publishing or at least
provides a set of rules or threshold value, for example, in this case is
the threshold of reasonable wages. This will greatly support the data
users to carry out further validation and set the same understanding of
which data are considered as outliers. Moreover, providing validation
rules would facilitate the usage of any establish data validation tool,
such as \texttt{validate} \citep{validate} package. In this case, we
cannot use this handy validation package due to the absence of
validation rules.

\hypertarget{acknowledgements}{%
\section{Acknowledgements}\label{acknowledgements}}

We would like to thank Aarathy Babu for the insight and discussion
during the writing of this paper.

The entire analysis is conducted using \texttt{R} \citep{R} in
\texttt{rstudio} using these packages: \texttt{tidyverse}
\citep{tidyverse}, \texttt{ggplot2} \citep{ggplot2}, \texttt{dplyr}
\citep{dplyr}, \texttt{readr} \citep{readr}, \texttt{tidyr}
\citep{tidyr}, \texttt{stringr} \citep{stringr}, \texttt{purrr}
\citep{purrr}, \texttt{broom} \citep{broom}, \texttt{blorgar}
\citep{brolgar}, \texttt{patchwork} \citep{patchwork},
\texttt{kableExtra} \citep{kableExtra}, \texttt{MASS} \citep{mass},
\texttt{janitor} \citep{janitor}, \texttt{DiagrammeR}
\citep{DiagrammeR}, \texttt{rsvg} \citep{rsvg}, \texttt{webshot}
\citep{webshot}, \texttt{mgcv} \citep{mgcv}, \texttt{tsibble}
\citep{tsibble}, and \texttt{modelr} \citep{modelr}. The paper are
generated using \texttt{knitr} \citep{knitr}, \texttt{rmarkdown}
\citep{rmarkdown}, and \texttt{rticles} \citep{rticles}.

\hypertarget{supplementary-materials}{%
\section{Supplementary Materials}\label{supplementary-materials}}

\begin{itemize}
\item
  \textbf{Codes} : R script to reproduce data tidying and cleaning are
  available in this
  \href{https://github.com/numbats/yowie/blob/master/data-raw/data_preprocessing.R}{Github
  Repository}.
\item
  \textbf{R Package \texttt{yowie}}:\texttt{yowie} is a data container R
  package that contains 3 datasets, namely the high school mean hourly
  wage data, high school dropouts mean hourly wage data, and demographic
  data of the NLSY79 cohort. This package could be accessed
  \href{https://github.com/numbats/yowie}{here}.
\item
  \textbf{shiny app}: An web interactive \texttt{shiny} app to run a
  simulation to customize the weight threshold. This app could be
  accessed
  \href{https://github.com/numbats/summer-wages-refresh/tree/main/app}{here}.
\end{itemize}

\bibliography{references.bib}

### 2021-09-06 R/linear_mixed_model.R

@@ -1,0 +1,51 @@
+library(tidyverse)
+library(brolgar)
+library(modelr)
+
+# read data
+wages <- yowie::wages %>%
+  mutate(year0 = year - 1979,
+         ln_wages = log(wage)) %>%
+  mutate(hgc_regroup = fct_collapse(hgc, `ungraded` = "UNGRADED",
+                                    `8TH grade or less` = c("1ST GRADE", "3RD GRADE",
+                                                            "4TH GRADE", "5TH GRADE",
+                                                            "6TH GRADE", "7TH GRADE",
+                                                            "8TH GRADE"),
+                                    `9TH to 11TH grade` = c("9TH GRADE",
+                                                            "10TH GRADE",
+                                                            "11TH GRADE"),
+                                    `12TH grade` = "12TH GRADE"))
+
+# calculate slope
+wages_slope <- key_slope(wages, ln_wages ~ year0) %>%
+  left_join(yowie::demog_nlsy79, by = "id")
+
+
+# build linear-mixed model
+wages_fit <- lmer(
+  ln_wages ~ year0*hgc_regroup*gender + (1|id),
+  data = wages
+)
+
+wages_fit_aug<- wages %>%
+  add_predictions(wages_fit, var = "pred") %>%
+  add_residuals(wages_fit, var = "resid")
+
+slope_bottom_12 <- wages_slope %>%
+  dplyr::select(id, .slope_year0) %>%
+  arrange(.slope_year0) %>%
+  slice_head(n = 12)
+
+wages_fit_aug2 <- wages_fit_aug %>%
+  filter(id %in% slope_bottom_12$id)
+
+ggplot(wages_fit_aug2, aes(x = year,
+                           y = pred,
+                           group = id,
+                           colour = factor(id))) +
+  geom_line() +
+  geom_point(aes(x = year,
+                 y = ln_wages,
+                 colour = factor(id))) +
+  facet_wrap(~id) +
+  theme(legend.position = "none")

### 2021-09-06 method_comparison.R

@@ -1,285 +1,0 @@
-library(tidyverse)
-library(zoo)
-library(rstatix)
-library(robustlmm)
-library(modelr)
-library(patchwork)
-library(MASS)
-library(broom)
-
-
-
-
-# open the tidy raw data set
-wages <- read_csv("data-raw/wages-high-school-demo/data-frames/tidy_employment_weighted.csv")
-demog <- read_csv("data-raw/wages-high-school-demo/data-frames/tidy_demographics_v2.csv")
-
-# join the wages information and the demographic information by case id.
-wages_demog <- left_join(wages, demog, by="id")
-
-
-# calculate the years in work force and the age of the subjects in 1979
-wages_demog <- wages_demog %>%
-  mutate(yr_hgc = as.numeric(yr_hgc)) %>%
-  mutate(years_in_workforce = year - yr_hgc) %>%
-  mutate(age_1979 = 1979 - (dob_year + 1900))
-
-# filter only the id with high school education
-wages_demog_hs <- wages_demog  %>% filter(grepl("GRADE", hgc))
-
-
-##### ROBUST LINEAR MODEL
-
-keep_me <- wages_demog_hs %>% count(id) %>% filter(n > 4)
-
-wages_demog_hs <- wages_demog_hs %>%
-  filter(id %in% keep_me$id)
-
-by_id <- wages_demog_hs %>%
-  dplyr::select(id, year, mean_hourly_wage) %>%
-  group_by(id) %>%
-  nest()
-
-
-id_rlm <- by_id %>%
-  mutate(model = map(.x = data,
-                     .f = function(x){
-                   rlm(mean_hourly_wage ~ year, data = x)
-                     }))
-
-id_aug <- id_rlm %>%
-  mutate(augmented = map(model, broom::augment)) %>%
-  unnest(augmented)
-
-id_w <- id_rlm %>%
-  mutate(w = map(.x = model,
-                 .f = function(x){
-                   x$w
-                 })) %>%
-  unnest(w) %>%
-  dplyr::select(w)
-
-
-id_aug_w <- cbind(id_aug, id_w) %>%
-  dplyr::select(`id...1`,
-                year,
-                mean_hourly_wage,
-                .fitted,
-                .resid,
-                .hat,
-                .sigma,
-                w) %>%
-  rename(id = `id...1`)
-
-
-
-wages_rlm_dat <- id_aug_w %>%
[…]

### 2021-09-07 paper/paper.Rmd
Change paper's file structure.

@@
     location: Clayton, VIC 3800
-    email: ""
-  - name: Kate Hyde
-    thanks: ""
-    department: ""
-    affiliation: ""
-    location: ""
-    email: ""
+    email: dama0007@student.monash.edu
+  - name: Dianne Cook
+    department: Department of Econometrics and Business Statistics
+    affiliation: Monash University
+    location: Clayton, VIC 3800
+    email: dicook@monash.edu
   - name: Emi Tanaka
@@
     email: emi.tanaka@monash.edu
+  - name: Kate Hyde
+    thanks: ""
+    department: Department of Econometrics and Business Statistics
+    affiliation: Monash University
+    location: Clayton, VIC 3800
+    email: ""
   - name: Nicholas Tierney
@@
     email: ""
-  - name: Dianne Cook
-    department: Department of Econometrics and Business Statistics
-    affiliation: Monash University
-    location: Clayton, VIC 3800
-    email: dicook@monash.edu
 bibliography: references.bib
@@
 
-```{r setup, include = FALSE}
-knitr::opts_chunk$set(echo = TRUE, 
-                      warning = FALSE, 
+```{r setup, echo = FALSE, cache = FALSE, include = FALSE}
+library(knitr)
+opts_chunk$set(echo = FALSE, warning = FALSE, 
                message = FALSE,
-                      #cache = TRUE, 
-                      #cache.path = "cache/",
-                      fig.path = "figures/")
-library(tidyverse)
-library(brolgar)
-library(patchwork)
-library(kableExtra)
-library(MASS)
-library(janitor)
-library(DiagrammeR)
-library(rsvg)
-library(webshot)
-library(mgcv)
-library(modelr)
-library(tsibble)
-library(lme4)
-library(modelr)
-library(forcats)
-library(yowie)
+               cache = TRUE, 
+               cache.path = "paper_cache/",
+               fig.path = "figures/",
+               comment = "#>",
+               fig.align = 'center', fig.show = 'hold',
+               out.width = ifelse(is_html_output(), "100%","\\textwidth"))
+
+read_chunk(here::here("paper/scripts/data-preprocessing.R"))
+read_chunk(here::here("paper/scripts/ida.R"))
 ```
@@
 **Keywords**: Data cleaning; Data tidying; Longitudinal data; NLSY79; Open data; Initial data analysis; Outlier detection; Robust linear regression
+
+```{r, load-pkgs}
+```
 
@@
 
-```{r, cache = TRUE, echo = FALSE}
-source(here::here("data-raw/NLSY79/NLSY79.R"))
+```{r, raw-data}
 ```
@@
 
-```{r untidy-data, echo = FALSE}
-
-untidy_demo <- new_data_qnames %>%
-  as_tibble() %>%
-  # in 2018, the variable's name is Q3-4_2018, instead of HGC_2018
-  rename(HGC_2018 = `Q3-4_2018`) %>%
-  dplyr::select(CASEID_1979,
-            starts_with("HRP") &
-              ends_with(c("1979", "1980", "1981", "1982", "1983"))) %>%
-  dplyr::select(1:7) %>%
-  head()
-
-kable(untidy_demo,
-      caption = "The Untidy Form of the NLSY79 Raw Data") %>%
-  kable_styling(latex_options = "striped")
+```{r untidy-data}
 ```
@@
 
-```{r dob-tidy}
-## tidy the date of birth data
-dob_tidy <- new_data_qnames %>%
-  dplyr::select(CASEID_1979,
-         starts_with("Q1-3_A~")) %>%
-  mutate(dob_year = case_when(
-                    # if the years recorded in both sets match, take 79 data
-                    `Q1-3_A~Y_1979` == `Q1-3_A~Y_1981` ~ `Q1-3_A~Y_1979`,
-                    # if the year in the 81 set is missing, take the 79 data
-                    is.na(`Q1-3_A~Y_1981`) ~ `Q1-3_A~Y_1979`,
-                    # if the sets don't match for dob year, take the 79 data
-                    `Q1-3_A~Y_1979` != `Q1-3_A~Y_1981` ~ `Q1-3_A~Y_1979`),
-        dob_month = case_when(
-                    # if months recorded in both sets match, take 79 data
-                    `Q1-3_A~M_1979` == `Q1-3_A~M_1981` ~ `Q1-3_A~M_1979`,
-                    # if month in 81 set is missing, take the 79 data
-                    is.na(`Q1-3_A~M_1981`) ~ `Q1-3_A~M_1979`,
-                    # if sets don't match for dob month, take the 79 data
-                    `Q1-3_A~M_1979` != `Q1-3_A~M_1981` ~ `Q1-3_A~M_1979`),
-        # flag if there is a conflict between dob recorded in 79 and 81
-        dob_conflict = case_when(     
-                      (`Q1-3_A~M_1979` != `Q1-3_A~M_1981`) & !is.na(`Q1-3_A~M_1981`)
-                      ~ TRUE,
-                      (`Q1-3_A~Y_1979` != `Q1-3_A~Y_1981`) & !is.na(`Q1-3_A~Y_1981`)
-                      ~ TRUE,
-                      (`Q1-3_A~Y_1979` == `Q1-3_A~Y_1981`) & 
-                      (`Q1-3_A~M_1979` == `Q1-3_A~M_1981`) ~ FALSE,
-                      is.na(`Q1-3_A~M_1981`) | is.na(`Q1-3_A~Y_1981`) ~ FALSE)) %>%
-  dplyr::select(CASEID_1979,
-         dob_month,
-         dob_year,
-         dob_conflict)
+```{r dob-tidy, echo=TRUE}
 ```
@@
 
-```{r demog-tidy}
-## tidy the gender and race variables
-demog_tidy <- categories_qnames %>%
-  dplyr::select(CASEID_1979,
-         SAMPLE_RACE_78SCRN,
-         SAMPLE_SEX_1979) %>%
-  rename(gender = SAMPLE_SEX_1979,
-         race = SAMPLE_RACE_78SCRN)
+```{r demog-tidy, echo = TRUE}
 ```
@@
 
-```{r tidy-grade, echo = FALSE}
-# tidy the grade 
-demog_education <- new_data_qnames %>%
-  as_tibble() %>%
-  # in 2018, the variable's name is Q3-4_2018, instead of HGC_2018
-  rename(HGC_2018 = `Q3-4_2018`) %>%
-  dplyr::select(CASEID_1979,
-         starts_with("HGCREV"),
-         "HGC_2012",
-         "HGC_2014",
-         "HGC_2016",
-         "HGC_2018") %>%
-  pivot_longer(!CASEID_1979,
-               names_to = "var",
-               values_to = "grade") %>%
-  separate("var", c("var", "year"), sep = -4) %>%
-  filter(!is.na(grade)) %>%
-  dplyr::select(-var)
+```{r tidy-grade}
 ```
@@
 
-```{r tidy-hgc, echo = FALSE}
-
-## getting the highest year of completed education ever
-highest_year <- demog_education %>%
-  group_by(CASEID_1979) %>%
-  mutate(hgc_i = max(grade)) %>%
-  filter(hgc_i == grade) %>%
-  filter(year == first(year)) %>%
-  rename(yr_hgc = year) %>%
-  dplyr::select(CASEID_1979, yr_hgc, hgc_i) %>%
-  ungroup() %>%
-  mutate('hgc' = ifelse(hgc_i == 0, "NONE", ifelse(hgc_i == 1, "1ST GRADE",
-                 ifelse(hgc_i == 2, "2ND GRADE", ifelse(hgc_i == 3, "3RD GRADE",
-                 ifelse(hgc_i >= 4 & hgc_i <= 12, paste0(hgc_i,"TH GRADE"),
-                 ifelse(hgc_i == 13, "1ST YEAR COL",
-                 ifelse(hgc_i == 14, "2ND YEAR COL",
-                 ifelse(hgc_i == 15, "3RD YEAR COL",
-                 ifelse(hgc_i == 95, "UNGRADED", 
-                        paste0((hgc_i - 12), "TH YEAR COL")))))))))))
+```{r tidy-hgc}
 ```
@@
 
-```{r full-demog, echo = FALSE}
-full_demographics <- full_join(dob_tidy, demog_tidy, by = "CASEID_1979") %>%
-  full_join(highest_year, by = "CASEID_1979") %>%
-  rename("id" = "CASEID_1979")
-
-head(full_demographics)
+```{r full-demog}
 ```
@@
 
-```{r tidy-hours}
-# make a list for years where we used the "QES-52A"
-year_A <- c(1979:1987, 1993)
-#function to get the hour of work
-get_hour <- function(year){
-  if(year %in% year_A){
-   temp <- new_data_qnames %>%
-    dplyr::select(CASEID_1979,
-            starts_with("QES-52A") &
-              ends_with(as.character(year)))} 
-  else{
-    temp <- new_data_qnames %>%
-    dplyr::select(CASEID_1979,
-            starts_with("QES-52D") &
-              ends_with(as.character(year)))} 
-    temp %>% 
-      pivot_longer(!CASEID_1979,
-                 names_to = "job",
-                 values_to = "hours_work") %>%
-      separate("job", c("job", "year"), sep = -4) %>%
-      mutate(job = paste0("job_", substr(job, 9, 10))) %>%
-      rename(id = CASEID_1979)
-}
-
-# list to save the iteration result
-hours <- list()
-# getting the hours of work of all observations
-for(ayear in c(1979:1994, 1996, 1998, 2000, 2002, 2004, 2006, 2008, 2010, 
-               2012, 2014, 2016, 2018)) {
-   hours[[ayear]] <- get_hour(ayear)
-}
-# unlist the hours of work
-hours_all <- bind_rows(!!!hours)
+```{r tidy-hours, echo=TRUE}
 ```
@@
 
-```{r tidy-rate, echo = FALSE}
-get_rate <- function(year) {
-  new_data_qnames %>%
-    dplyr::select(CASEID_1979,
-            starts_with("HRP") &
-              ends_with(as.character(year))) %>%
-    pivot_longer(!CASEID_1979, names_to = "job", values_to = "rate_per_hour") %>%
-    separate("job", c("job", "year"), sep = -4) %>%
-    mutate(job = paste0("job_0", substr(job, 4, 4))) %>%
-    rename(id = CASEID_1979)
-}
-rates <- list()
-for(ayear in c(1979:1994, 1996, 1998, 2000, 2002, 2004, 2006, 2008, 2010, 
-               2012, 2014, 2016, 2018)) {
-   rates[[ayear]] <- get_rate(ayear)
-}
-rates_all <- bind_rows(!!!rates)
-# join hours and rates variable
-hours_wages <- left_join(rates_all, 
-                         hours_all, 
-                         by = c("id", "year", "job")) %>%
-  # set the 0 value in rate_per_hour as NA 
-  mutate(rate_per_hour = ifelse(rate_per_hour == 0, NA,
-                                rate_per_hour),
-         hours_work = ifelse(hours_work == 0 | hours_work > 84, NA, hours_work))
-
-head(hours_wages)
+```{r tidy-rate, echo = TRUE}
 ```
@@
 
-```{r tidy-nojob, echo = FALSE}
-# calculate number of jobs that a person has in one year
-no_job <- hours_wages %>%
-  filter(!is.na(rate_per_hour)) %>%
-  group_by(id, year) %>%
-  summarise(no_jobs = length(rate_per_hour))
-
-# filter the observations with available rate per hour
-eligible_wages <- hours_wages %>%
-  filter(!is.na(rate_per_hour)) %>%
-  left_join(no_job, by = c("id", "year"))
-
-# calculate the mean_hourly_wage
-# flag1 = code 1 for weighted mean
-# code 0 for arithmetic mean
-mean_hourly_wage <- eligible_wages %>%
-  group_by(id, year) %>%
-  #calculate the weighted mean if the number of jobs > 1
-  mutate(wages = ifelse(no_jobs == 1, rate_per_hour/100,
-                        weighted.mean(rate_per_hour, hours_work, na.rm = TRUE)/100)) %>%
-  #give the flag if it the weighted mean
-  mutate(flag1 = ifelse(!is.na(wages) & no_jobs != 1, 1,
-                        0)) %>%
-  #calculate the arithmetic mean for the na
-  mutate(wages = ifelse(is.na(wages), mean(rate_per_hour)/100,
-                        wages)) %>%
-  group_by(id, year) %>%
-  summarise(wages = mean(wages),
-            total_hours = sum(hours_work),
-            number_of_jobs = mean(no_jobs),
-            flag1 = mean(flag1)) %>%
-  mutate(year = as.numeric(year)) %>%
-  ungroup() %>%
-  rename(mean_hourly_wage = wages) %>%
-  mutate(is_wm = ifelse(flag1 == 1, TRUE,
-                        FALSE)) %>%
-  dplyr::select(-flag1)
-
-head(mean_hourly_wage, n = 10)
+```{r tidy-nojob}
 ```
@@
 
-```{r wages-demog-hs, echo = FALSE}
-# join the wages information and the demographic information by case id.
-wages_demog <- left_join(mean_hourly_wage, full_demographics, by="id")
-# calculate the years in work force and the age of the subjects in 1979
-wages_demog <- wages_demog %>%
-  mutate(yr_hgc = as.numeric(yr_hgc)) %>%
-  mutate(years_in_workforce = year - yr_hgc) %>%
-  mutate(age_1979 = 1979 - (dob_year + 1900))
-# filter only the id with high school education
-wages_demog_hs <- wages_demog  %>% filter(grepl("GRADE", hgc))
-# calculate the number of observation
-keep_me <- wages_demog_hs %>%
-  count(id) %>%
-  filter(n > 4)
-wages_demog_hs <- wages_demog_hs %>%
-  filter(id %in% keep_me$id)
+```{r wages-demog-hs}
 ```
@@
 
-```{r age-table, echo = FALSE}
-age_table <- yowie::demog_nlsy79 %>%
-  group_by(age_1979) %>%
-  count(age_1979)
-
-kable(age_table,
-      caption = "Age Distribution of the NLSY79 samples",
-      col.names = c("Age", "Number of Sample")) %>%
-  kable_styling(latex_options = "striped")
+```{r age-table}
 ```
 
-```{r gender-race-table, echo = FALSE}
-gender_race_table <- yowie::demog_nlsy79 %>%
-  tabyl(gender, race) %>%
-  adorn_totals(c("row", "col")) %>%
-  adorn_percentages("row") %>%
-  adorn_pct_formatting(digits = 2) %>%
-  adorn_ns(position = "front") %>%
-  mutate(gender = ifelse(gender == "MALE", "Male",
-                         ifelse(gender == "FEMALE", "Female", "Total")))
-
-kable(gender_race_table,
-      caption = "Gender and Race Distribution of the NLSY79 Samples",
-      col.names = c("Gender", "Hispanic", "Black", "Non-Black, Non-Hispanic", "Total")) %>%
-  kable_styling(latex_options = "striped") %>%
-  add_header_above(c(" " = 1, "Race" = 3, " " = 1))
-
+```{r gender-race-table}
 ```
@@
 
-```{r summarytable, echo = FALSE}
-kable(as.array(summary(wages_demog_hs$mean_hourly_wage)), 
-      caption = "Summary Statistics of Wages of High School Data",
-      col.names = c("Statistics", "Value")) %>%
-  kable_styling(latex_options = "striped")
+```{r summarytable}
 ```
@@
 
-```{r feature-plot, echo = FALSE, fig.cap = "Two plots showing the distribution of the mean hourly wage.
+```{r feature-plot, fig.cap = "Two plots showing the distribution of the mean hourly wage.
 Plot A portrays the pattern of mean hourly wage of high school cohort from 1979 to 2018 of each ID in US Dollar; Plot B shows the distribution of their minimum, median, and maximum value.
 We can see that some IDs had an extremely high of wages and it made the distribution of the three features is extremely skewed.", fig.width=6, fig.height=4}
-
-spag <- wages_demog_hs %>%
-  ggplot(aes(x = year,
-             y = mean_hourly_wage,
-             group = id)) +
-  geom_line(alpha = 0.1) +
-  ggtitle("A)") +
-  theme(plot.title = element_text(size = 10))
-
-wages_demog_hs_tsibble <- as_tsibble(x = wages_demog_hs,
-                    key = id,
-                    index = year,
-                    regular = FALSE)
-wages_three_feat <- wages_demog_hs_tsibble %>%
-  features(mean_hourly_wage, 
-           feat_three_num
-           )
-wages_feat_long <- wages_three_feat %>%
-  pivot_longer(c(min, med, max), names_to = "feature", values_to = "value")
-feature <- ggplot(wages_feat_long) +
-  geom_density(aes(x = value, colour = feature, fill = feature), alpha = 0.3) +
-  ggtitle("B)") +
-  theme(plot.title = element_text(size = 10))
-
-spag + feature
 ```
@@
 
-```{r high-wages, echo = FALSE, fig.cap= "6 out of 45 IDs with extremely high mean hourly wage. Most of the IDs only have one point of high wage."}
-
-wages_high <- filter(wages_demog_hs, mean_hourly_wage > 500) %>%
-  as_tibble() %>%
-  head(n = 6)
-
-wages_high2 <- wages_demog_hs %>%
-  filter(id %in% wages_high$id)
-
-ggplot(wages_high2) +
-  geom_line(aes(x = year,
-                y = mean_hourly_wage)) +
-  geom_point(aes(x = year,
-                y = mean_hourly_wage),
-             size = 0.5,
-             alpha = 0.5) +
-  facet_wrap(~id, scales = "free_y") +
-  theme(axis.text.x = element_text(angle = 10, size = 6)) +
-  ylab("mean hourly wage")
-
+```{r high-wages, fig.cap= "6 out of 45 IDs with extremely high mean hourly wage. Most of the IDs only have one point of high wage."}
 ```
@@
 It suggests that some IDs had a reasonably fluctuate wages.", fig.height=4, fig.width=5}
-
-set.seed(20210225)
-
-ggplot(wages_demog_hs_tsibble, 
-       aes(x = year,
-                y = mean_hourly_wage,
-                group = id)) +
-  geom_line(alpha = 0.7) +
-  facet_sample() +
-  theme(axis.text.x = element_text(angle = 10, size = 6)) +
-  ylab("mean hourly wage") 
 ```
@@
 
-```{r rlm, eval=FALSE, warning = FALSE, message = FALSE, echo = FALSE}
-
-# nest the data by id to build a robust linear model
-by_id <- wages_demog_hs %>%
-  dplyr::select(id, year, mean_hourly_wage) %>%
-  group_by(id) %>%
-  nest()
-
-# build a robust linear model
-id_rlm <- by_id %>%
-  mutate(model = map(.x = data,
-                     .f = function(x){
-                       rlm(mean_hourly_wage ~ year, data = x)
-                     }))
-# extract the property of the regression model
-id_aug <- id_rlm %>%
-  mutate(augmented = map(model, broom::augment)) %>%
-  unnest(augmented)
-
-# extract the weight of each observation
-id_w <- id_rlm %>%
-  mutate(w = map(.x = model,
-                 .f = function(x){
-                   x$w
-                 })) %>%
-  unnest(w) %>%
-  dplyr::select(w)
-
-# bind the property of each observation with their weight
-id_aug_w <- cbind(id_aug, id_w) %>%
-  dplyr::select(`id...1`,
-                year,
-                mean_hourly_wage,
-                .fitted,
-                .resid,
-                .hat,
-                .sigma,
-                w) %>%
-  rename(id = `id...1`)
-
-# if the weight < 0.12, the mean_hourly_wage is replaced by the model's fitted/predicted value.
-# and add the flag whether the observation is predicted value or not.
-# since the fitted value is sometimes <0, and wages value could never be negative,
-# we keep the mean hourly wage value even its weight < 0.12.
-
-wages_rlm_dat <- id_aug_w %>%
-  mutate(wages_rlm = ifelse(w < 0.12  & .fitted >= 0, .fitted,
-                            mean_hourly_wage)) %>%
-  mutate(is_pred = ifelse(w < 0.12 & .fitted >= 0, TRUE,
-                          FALSE)) %>%
-  dplyr::select(id, year, wages_rlm, is_pred)
-
-# join back the `wages_rlm_dat` to `wages_demog_hs`
-
-wages_demog_hs <- left_join(wages_demog_hs, wages_rlm_dat, by = c("id", "year"))
-
-# save it to rds file so it would faster the knitting process
-saveRDS(wages_demog_hs, here::here("Report/Result/wages_demog_hs_tmp.rds"))
+```{r rlm, eval=FALSE}
 ```
@@
 
-```{r compare-data, echo = FALSE}
-set.seed(31251587)
-
-wages_demog_hs <- readRDS(here::here("Report/Result/wages_demog_hs_tmp.rds"))
-sample_id <- sample(unique(wages_demog_hs$id), 20)
-sample <- subset(wages_demog_hs, id %in% sample_id)
-
-wages_compare <- sample %>%
-  dplyr::select(id, year, mean_hourly_wage, wages_rlm) %>%
-  rename(mean_hourly_wage_rlm = wages_rlm) %>%
-  pivot_longer(c(-id, -year), names_to = "type", values_to = "wages")
+```{r compare-data}
 ```
 
-```{r compare-plot, echo = FALSE, fig.cap="Comparison between the original and the treated mean hourly wage.
+```{r compare-plot, fig.cap="Comparison between the original and the treated mean hourly wage.
 The orange line portray the original value of mean hourly wage, while the turquoise line display the mean hourly wages value after the extreme values imputed with the robust linear model's prediction value.
 We can see that some extreme spikes has been reduced by the model.", fig.height=5, fig.width=6}
-ggplot(wages_compare) +
-  geom_line(aes(x = year,
-                y = wages,
-                colour = type,
-                linetype = type),
-            alpha = 1) +
-  geom_point(aes(x = year,
-                y = wages,
-                colour = type),
-            alpha = 0.5,
-            size = 1) +
-  theme(axis.text.x = element_text(angle = 10, size = 5),
-        legend.position = "bottom") +
-  facet_wrap(~id, scales = "free_y")
-
 ```
@@
 
-```{r comppict, echo = FALSE, fig.cap = "The distribution of the mean hourly wage after the extreme values are replaced.
+```{r comppict, fig.cap = "The distribution of the mean hourly wage after the extreme values are replaced.
 Plot A portrays the pattern of mean hourly wage of high school cohort from 1979 to 2018 of each ID in US Dollar; Plot B shows the distribution of their minimum, median, and maximum value transformed to log10 scale.
@@
 Also, some IDs' have a minimum wages that is higher than others' maximum wages.", fig.width=6, fig.height=4}
-
-spag2 <- wages_demog_hs %>%
-  ggplot(aes(x = year,
-             y = wages_rlm,
-             group = id)) +
-  geom_line(alpha = 0.1) +
-  ggtitle("A)") +
-  theme(plot.title = element_text(size = 10))
-
-wages_hs2020_rlm <- as_tsibble(x = wages_demog_hs,
-                    key = id,
-                    index = year,
-                    regular = FALSE)
-wages_three_feat_rlm <- wages_hs2020_rlm %>%
-  features(wages_rlm, 
-           feat_three_num
-           )
-wages_feat_long_rlm <- wages_three_feat_rlm %>%
-  pivot_longer(c(min, med, max), names_to = "feature", values_to = "value")
-
-feature2 <- ggplot(wages_feat_long_rlm) +
-  geom_density(aes(x = value, colour = feature, fill = feature), alpha = 0.3) +
-  ggtitle("B)") +
-  scale_x_log10() + 
-  theme(plot.title = element_text(size = 10)) +
-  xlab("value (log10)")
-
-spag2 + feature2
 ```
@@
 
-```{r save-data, echo = FALSE}
-# select out the old value of mean hourly wage and change it with the wages_rlm value
-wages_demog_hs <- wages_demog_hs %>%
-  dplyr::select(-mean_hourly_wage) %>%
-  rename(mean_hourly_wage = wages_rlm)
-
-# rename and select the wages in tidy
-wages_hs2020 <- wages_demog_hs %>%
-  dplyr::select(id, year, mean_hourly_wage, age_1979, gender, race, hgc, hgc_i, yr_hgc,
-                number_of_jobs, total_hours, is_wm, is_pred) %>%
-  mutate(hgc = as.factor(hgc),
-         year = as.integer(year),
-         age_1979 = as.integer(age_1979),
-         yr_hgc = as.integer(yr_hgc),
-         number_of_jobs = as.integer(number_of_jobs))
-
-# Create a data set for demographic variables
-demographic_nlsy79 <- full_demographics %>%
-  mutate(age_1979 = 1979 - (dob_year + 1900)) %>%
-  dplyr::select(id,
-         age_1979,
-         gender,
-         race,
-         hgc,
-         hgc_i,
-         yr_hgc) %>%
-  mutate(age_1979 = as.integer(age_1979),
-         hgc = as.factor(hgc),
-         yr_hgc = as.integer(yr_hgc))
-
-# Create a data set for the high school dropouts cohort
-wages_hs_dropout <- wages_hs2020 %>%
-  mutate(dob = 1979 - age_1979,
-         age_hgc = yr_hgc - dob) %>%
-  filter((hgc %in% c("9TH GRADE",
-                     "10TH GRADE",
-                     "11TH GRADE")) |
-          (hgc == "12TH GRADE" &
-              age_hgc > 19)) %>%
-  filter(age_1979 <= 17,
-         gender == "MALE") %>%
-  dplyr::select(-dob,
-         -age_hgc)
-
+```{r save-data}
 ```
 
-```{r nrow, echo=FALSE}
-a = nrow(categories_qnames)
-b = nrow(full_demographics)
-d = eligible_wages %>%
-  group_by(id) %>%
-  count(id) %>%
-  nrow()
-n_hgc = wages_hs2020 %>%
-  group_by(id) %>%
-  count(id) %>%
-  nrow()
-n_obs_hgc = nrow(wages_hs2020)
-n_obs_hgc_pred = filter(wages_hs2020, is_pred == TRUE) %>%
-  nrow()
-
-n_do = wages_hs_dropout %>%
-  group_by(id) %>%
-  count(id) %>%
-  nrow()
-n_obs_do = nrow(wages_hs_dropout)
-n_obs_do_pred = filter(wages_hs_dropout, is_pred == TRUE) %>%
-  nrow()
+```{r nrow}
 ```
 
-```{r flow-chart, echo=FALSE, fig.cap="The stages of data filtering from the raw data to get three datasets that are contained in an R package called yowie, n means the number of ID, while n_obs means the number of observations.", fig.height=3, fig.width=3}
-
-fc <- grViz("digraph flowchart {
-      node [fontname = Helvetica, shape = rectangle]        
-      tab1 [label = '@@1']
-      tab2 [label = '@@2']
-      tab3 [label = '@@3']
-      tab4 [label = '@@4']
-      tab5 [label = '@@5']
-      tab6 [label = '@@6']
-      tab7 [label = '@@7']
-
-      tab1 -> tab2;
-      tab1 -> tab3;
-      tab1 -> tab4 -> tab5 -> tab6;
-      tab5 -> tab7;
-      tab6 -> tab7;
-      tab2 -> tab7
-      }
-
-      [1]: paste0('NLSY79, n = ', a)
-      [2]: paste0('Extract demographic variable, n = ', b)
-      [3]: paste0('Exclude ', a - d, ' IDs whose hourly rate is missing')
-      [4]: paste0('Eligible ID, n = ', d)
-      [5]: paste0('Cohort whose hgc is up to 12th grade and \\n participated at least 5 years in the survey, \\n n = ', n_hgc, ', n_obs = ', n_obs_hgc, ' \\n (', n_obs_hgc_pred, ' observations are predicted value)')
-      [6]: paste0('High school dropouts cohort, \\n n = ', n_do, ', n_obs = ', n_obs_do, ' \\n (', n_obs_do_pred, ' observations are predicted value)')
-      [7]: 'yowie Package'
-      ",
-      height = 200)
-
-fc
+```{r flow-chart, echo=FALSE, fig.cap="The stages of data filtering from the raw data to get three datasets that are contained in an R package called yowie, n means the number of ID, while n_obs means the number of observations.", fig.height=5, fig.width=5}
 ```
@@
 
-```{r slope-gender, fig.cap=""}
+```{r slope-gender, fig.cap="slope-gender"}
 slope_bottom_5<- wages_slope %>%
@@
 
-```{r}
+```{r slope-educ}
 ggplot(contrast_slope, aes(x = year, y = ln_wages, color = hgc_regroup)) +
@@
 
-```{r}
+```{r slope-nobs}
 wages_obs <- wages %>%
@@
 
-Notes:
-
-Trimming the slope by 95% percentile and 5% percentile. 
-After that explore the slope after trimmed:
-- By sex (including the wage gender gap)
-- By education
-
 According to @OECD, gender pay gap is the difference between median wage of males and females relative to median wage of males.
@@
 
-```{r se-gender, echo=FALSE, fig.cap="Gender wage gap or the median wages of females relatives to median wages of males.
+```{r wages-gender-gap, echo=FALSE, fig.cap="Gender wage gap or the median wages of females relatives to median wages of males.
 Females tend to earn fewer than males.

### 2021-09-07 paper/paper.tex
Change paper's file structure.

--- 2021-05-05 report_alt.tex
@@
   Clayton, VIC 3800 \\
-  \texttt{} \\
+  \texttt{\href{mailto:dama0007@student.monash.edu}{\nolinkurl{dama0007@student.monash.edu}}} \\
    \And
-    Kate Hyde
+    Dianne Cook
    \\
-     \\
-   \\
-  \texttt{} \\
+    Department of Econometrics and Business Statistics \\
+    Monash University \\
+  Clayton, VIC 3800 \\
+  \texttt{\href{mailto:dicook@monash.edu}{\nolinkurl{dicook@monash.edu}}} \\
    \And
@@
    \And
+    Kate Hyde
+   \\
+    Department of Econometrics and Business Statistics \\
+    Monash University \\
+  Clayton, VIC 3800 \\
+  \texttt{} \\
+   \And
     Nicholas Tierney
@@
   \texttt{} \\
-   \And
-    Dianne Cook
-   \\
-    Department of Econometrics and Business Statistics \\
-    Monash University \\
-  Clayton, VIC 3800 \\
-  \texttt{\href{mailto:dicook@monash.edu}{\nolinkurl{dicook@monash.edu}}} \\
   }
@@
 \begin{Highlighting}[]
-\CommentTok{## tidy the date of birth data}
+\CommentTok{#tidy the date of birth data}
 \NormalTok{dob_tidy <-}\StringTok{ }\NormalTok{new_data_qnames }\OperatorTok{%>%}
@@
     \CommentTok{# flag if there is a conflict between dob recorded in 79 and 81}
-        \DataTypeTok{dob_conflict =} \KeywordTok{case_when}\NormalTok{(     }
+    \DataTypeTok{dob_conflict =} \KeywordTok{case_when}\NormalTok{(}
 \NormalTok{      (}\StringTok{`}\DataTypeTok{Q1-3_A~M_1979}\StringTok{`} \OperatorTok{!=}\StringTok{ `}\DataTypeTok{Q1-3_A~M_1981}\StringTok{`}\NormalTok{) }\OperatorTok{&}\StringTok{ }\OperatorTok{!}\KeywordTok{is.na}\NormalTok{(}\StringTok{`}\DataTypeTok{Q1-3_A~M_1981}\StringTok{`}\NormalTok{)}
@@
       \OperatorTok{~}\StringTok{ }\OtherTok{TRUE}\NormalTok{,}
-\NormalTok{                      (}\StringTok{`}\DataTypeTok{Q1-3_A~Y_1979}\StringTok{`} \OperatorTok{==}\StringTok{ `}\DataTypeTok{Q1-3_A~Y_1981}\StringTok{`}\NormalTok{) }\OperatorTok{&}\StringTok{ }
+\NormalTok{      (}\StringTok{`}\DataTypeTok{Q1-3_A~Y_1979}\StringTok{`} \OperatorTok{==}\StringTok{ `}\DataTypeTok{Q1-3_A~Y_1981}\StringTok{`}\NormalTok{) }\OperatorTok{&}
 \StringTok{        }\NormalTok{(}\StringTok{`}\DataTypeTok{Q1-3_A~M_1979}\StringTok{`} \OperatorTok{==}\StringTok{ `}\DataTypeTok{Q1-3_A~M_1981}\StringTok{`}\NormalTok{) }\OperatorTok{~}\StringTok{ }\OtherTok{FALSE}\NormalTok{,}
@@
 \begin{Highlighting}[]
-\CommentTok{## tidy the gender and race variables}
+\CommentTok{# tidy the gender and race variables}
 \NormalTok{demog_tidy <-}\StringTok{ }\NormalTok{categories_qnames }\OperatorTok{%>%}
@@
 \begin{verbatim}
-##   id dob_month dob_year dob_conflict                    race gender yr_hgc
-## 1  1         9       58        FALSE NON-BLACK, NON-HISPANIC FEMALE   1979
-## 2  2         1       59        FALSE NON-BLACK, NON-HISPANIC FEMALE   1985
-## 3  3         8       61        FALSE NON-BLACK, NON-HISPANIC FEMALE   1993
-## 4  4         8       62        FALSE NON-BLACK, NON-HISPANIC FEMALE   1986
-## 5  5         7       59        FALSE NON-BLACK, NON-HISPANIC   MALE   1984
-## 6  6        10       60        FALSE NON-BLACK, NON-HISPANIC   MALE   1983
-##   hgc_i          hgc
-## 1    12   12TH GRADE
-## 2    12   12TH GRADE
-## 3    12   12TH GRADE
-## 4    14 2ND YEAR COL
-## 5    18 6TH YEAR COL
-## 6    16 4TH YEAR COL
+#>   id dob_month dob_year dob_conflict                    race gender yr_hgc
+#> 1  1         9       58        FALSE NON-BLACK, NON-HISPANIC FEMALE   1979
+#> 2  2         1       59        FALSE NON-BLACK, NON-HISPANIC FEMALE   1985
+#> 3  3         8       61        FALSE NON-BLACK, NON-HISPANIC FEMALE   1993
+#> 4  4         8       62        FALSE NON-BLACK, NON-HISPANIC FEMALE   1986
+#> 5  5         7       59        FALSE NON-BLACK, NON-HISPANIC   MALE   1984
+#> 6  6        10       60        FALSE NON-BLACK, NON-HISPANIC   MALE   1983
+#>   hgc_i          hgc
+#> 1    12   12TH GRADE
+#> 2    12   12TH GRADE
+#> 3    12   12TH GRADE
+#> 4    14 2ND YEAR COL
+#> 5    18 6TH YEAR COL
+#> 6    16 4TH YEAR COL
 \end{verbatim}
@@
                     \KeywordTok{starts_with}\NormalTok{(}\StringTok{"QES-52A"}\NormalTok{) }\OperatorTok{&}
-\StringTok{              }\KeywordTok{ends_with}\NormalTok{(}\KeywordTok{as.character}\NormalTok{(year)))\} }
+\StringTok{                      }\KeywordTok{ends_with}\NormalTok{(}\KeywordTok{as.character}\NormalTok{(year)))\}}
   \ControlFlowTok{else}\NormalTok{\{}
@@
                     \KeywordTok{starts_with}\NormalTok{(}\StringTok{"QES-52D"}\NormalTok{) }\OperatorTok{&}
-\StringTok{              }\KeywordTok{ends_with}\NormalTok{(}\KeywordTok{as.character}\NormalTok{(year)))\} }
-\NormalTok{    temp }\OperatorTok{%>%}\StringTok{ }
+\StringTok{                      }\KeywordTok{ends_with}\NormalTok{(}\KeywordTok{as.character}\NormalTok{(year)))\}}
+\NormalTok{  temp }\OperatorTok{%>%}
 \StringTok{    }\KeywordTok{pivot_longer}\NormalTok{(}\OperatorTok{!}\NormalTok{CASEID_}\DecValTok{1979}\NormalTok{,}
@@
 \NormalTok{\}}
-
 \CommentTok{# list to save the iteration result}
@@
 \CommentTok{# getting the hours of work of all observations}
-\ControlFlowTok{for}\NormalTok{(ayear }\ControlFlowTok{in} \KeywordTok{c}\NormalTok{(}\DecValTok{1979}\OperatorTok{:}\DecValTok{1994}\NormalTok{, }\DecValTok{1996}\NormalTok{, }\DecValTok{1998}\NormalTok{, }\DecValTok{2000}\NormalTok{, }\DecValTok{2002}\NormalTok{, }\DecValTok{2004}\NormalTok{, }\DecValTok{2006}\NormalTok{, }\DecValTok{2008}\NormalTok{, }\DecValTok{2010}\NormalTok{, }
+\ControlFlowTok{for}\NormalTok{(ayear }\ControlFlowTok{in} \KeywordTok{c}\NormalTok{(}\DecValTok{1979}\OperatorTok{:}\DecValTok{1994}\NormalTok{, }\DecValTok{1996}\NormalTok{, }\DecValTok{1998}\NormalTok{, }\DecValTok{2000}\NormalTok{, }\DecValTok{2002}\NormalTok{, }\DecValTok{2004}\NormalTok{, }\DecValTok{2006}\NormalTok{, }\DecValTok{2008}\NormalTok{, }\DecValTok{2010}\NormalTok{,}
                \DecValTok{2012}\NormalTok{, }\DecValTok{2014}\NormalTok{, }\DecValTok{2016}\NormalTok{, }\DecValTok{2018}\NormalTok{)) \{}
@@
 Some observations had 0 in their hourly rate, which is considered as invalid value.
-Thus, their hourly rate set to be N.A.
-the result of the joining variables could be seen in the following tibble.
+Thus, their hourly rate set to be NA.
+The result of the joining variables could be seen in the following tibble.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{get_rate <-}\StringTok{ }\ControlFlowTok{function}\NormalTok{(year) \{}
+\NormalTok{  new_data_qnames }\OperatorTok{%>%}
+\StringTok{    }\NormalTok{dplyr}\OperatorTok{::}\KeywordTok{select}\NormalTok{(CASEID_}\DecValTok{1979}\NormalTok{,}
+                  \KeywordTok{starts_with}\NormalTok{(}\StringTok{"HRP"}\NormalTok{) }\OperatorTok{&}
+\StringTok{                    }\KeywordTok{ends_with}\NormalTok{(}\KeywordTok{as.character}\NormalTok{(year))) }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{pivot_longer}\NormalTok{(}\OperatorTok{!}\NormalTok{CASEID_}\DecValTok{1979}\NormalTok{, }\DataTypeTok{names_to =} \StringTok{"job"}\NormalTok{, }\DataTypeTok{values_to =} \StringTok{"rate_per_hour"}\NormalTok{) }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{separate}\NormalTok{(}\StringTok{"job"}\NormalTok{, }\KeywordTok{c}\NormalTok{(}\StringTok{"job"}\NormalTok{, }\StringTok{"year"}\NormalTok{), }\DataTypeTok{sep =} \DecValTok{-4}\NormalTok{) }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{job =} \KeywordTok{paste0}\NormalTok{(}\StringTok{"job_0"}\NormalTok{, }\KeywordTok{substr}\NormalTok{(job, }\DecValTok{4}\NormalTok{, }\DecValTok{4}\NormalTok{))) }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{rename}\NormalTok{(}\DataTypeTok{id =}\NormalTok{ CASEID_}\DecValTok{1979}\NormalTok{)}
+\NormalTok{\}}
+\NormalTok{rates <-}\StringTok{ }\KeywordTok{list}\NormalTok{()}
+\ControlFlowTok{for}\NormalTok{(ayear }\ControlFlowTok{in} \KeywordTok{c}\NormalTok{(}\DecValTok{1979}\OperatorTok{:}\DecValTok{1994}\NormalTok{, }\DecValTok{1996}\NormalTok{, }\DecValTok{1998}\NormalTok{, }\DecValTok{2000}\NormalTok{, }\DecValTok{2002}\NormalTok{, }\DecValTok{2004}\NormalTok{, }\DecValTok{2006}\NormalTok{, }\DecValTok{2008}\NormalTok{, }\DecValTok{2010}\NormalTok{,}
+               \DecValTok{2012}\NormalTok{, }\DecValTok{2014}\NormalTok{, }\DecValTok{2016}\NormalTok{, }\DecValTok{2018}\NormalTok{)) \{}
+\NormalTok{  rates[[ayear]] <-}\StringTok{ }\KeywordTok{get_rate}\NormalTok{(ayear)}
+\NormalTok{\}}
+\NormalTok{rates_all <-}\StringTok{ }\KeywordTok{bind_rows}\NormalTok{(}\OperatorTok{!!!}\NormalTok{rates)}
+\CommentTok{# join hours and rates variable}
+\NormalTok{hours_wages <-}\StringTok{ }\KeywordTok{left_join}\NormalTok{(rates_all,}
+\NormalTok{                         hours_all,}
+                         \DataTypeTok{by =} \KeywordTok{c}\NormalTok{(}\StringTok{"id"}\NormalTok{, }\StringTok{"year"}\NormalTok{, }\StringTok{"job"}\NormalTok{)) }\OperatorTok{%>%}
+\StringTok{  }\CommentTok{# set the 0 value in rate_per_hour as NA}
+\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{rate_per_hour =} \KeywordTok{ifelse}\NormalTok{(rate_per_hour }\OperatorTok{==}\StringTok{ }\DecValTok{0}\NormalTok{, }\OtherTok{NA}\NormalTok{,}
+\NormalTok{                                rate_per_hour),}
+         \DataTypeTok{hours_work =} \KeywordTok{ifelse}\NormalTok{(hours_work }\OperatorTok{==}\StringTok{ }\DecValTok{0} \OperatorTok{|}\StringTok{ }\NormalTok{hours_work }\OperatorTok{>}\StringTok{ }\DecValTok{84}\NormalTok{, }\OtherTok{NA}\NormalTok{, hours_work))}
+\KeywordTok{head}\NormalTok{(hours_wages)}
+\end{Highlighting}
+\end{Shaded}
 
 \begin{verbatim}
-## # A tibble: 6 x 5
-##      id job    year  rate_per_hour hours_work
-##   <int> <chr>  <chr>         <int>      <int>
-## 1     1 job_01 1979            328         38
-## 2     1 job_02 1979             NA         15
-## 3     1 job_03 1979             NA         NA
-## 4     1 job_04 1979             NA         NA
-## 5     1 job_05 1979             NA         NA
-## 6     2 job_01 1979            385         35
+#> # A tibble: 6 x 5
+#>      id job    year  rate_per_hour hours_work
+#>   <int> <chr>  <chr>         <int>      <int>
+#> 1     1 job_01 1979            328         38
+#> 2     1 job_02 1979             NA         15
+#> 3     1 job_03 1979             NA         NA
+#> 4     1 job_04 1979             NA         NA
+#> 5     1 job_05 1979             NA         NA
+#> 6     2 job_01 1979            385         35
 \end{verbatim}
@@
 \begin{verbatim}
-## # A tibble: 10 x 6
-##       id  year mean_hourly_wage total_hours number_of_jobs is_wm
-##    <int> <dbl>            <dbl>       <int>          <dbl> <lgl>
-##  1     1  1979             3.28          38              1 FALSE
-##  2     1  1981             3.61          NA              1 FALSE
-##  3     2  1979             3.85          35              1 FALSE
-##  4     2  1980             4.57          NA              1 FALSE
-##  5     2  1981             5.14          NA              1 FALSE
-##  6     2  1982             5.71          35              1 FALSE
-##  7     2  1983             5.71          NA              1 FALSE
-##  8     2  1984             5.14          NA              1 FALSE
-##  9     2  1985             7.71          NA              1 FALSE
-## 10     2  1986             7.69          NA              1 FALSE
+#> # A tibble: 10 x 6
+#>       id  year mean_hourly_wage total_hours number_of_jobs is_wm
+#>    <int> <dbl>            <dbl>       <int>          <dbl> <lgl>
+#>  1     1  1979             3.28          38              1 FALSE
+#>  2     1  1981             3.61          NA              1 FALSE
+#>  3     2  1979             3.85          35              1 FALSE
+#>  4     2  1980             4.57          NA              1 FALSE
+#>  5     2  1981             5.14          NA              1 FALSE
+#>  6     2  1982             5.71          35              1 FALSE
+#>  7     2  1983             5.71          NA              1 FALSE
+#>  8     2  1984             5.14          NA              1 FALSE
+#>  9     2  1985             7.71          NA              1 FALSE
+#> 10     2  1986             7.69          NA              1 FALSE
 \end{verbatim}
@@
 \hline
-Mean & 11.86578\\
+Mean & 11.86593\\
 \hline
@@
 \begin{figure}
-\centering
-\includegraphics{figures/feature-plot-1.pdf}
-\caption{\label{fig:feature-plot}Two plots showing the distribution of the mean hourly wage.
+
+{\centering \includegraphics[width=\textwidth]{figures/feature-plot-1}
+
+}
+
+\caption{Two plots showing the distribution of the mean hourly wage.
 Plot A portrays the pattern of mean hourly wage of high school cohort from 1979 to 2018 of each ID in US Dollar; Plot B shows the distribution of their minimum, median, and maximum value.
-We can see that some IDs had an extremely high of wages and it made the distribution of the three features is extremely skewed.}
+We can see that some IDs had an extremely high of wages and it made the distribution of the three features is extremely skewed.}\label{fig:feature-plot}
 \end{figure}
@@
 \begin{figure}
-\centering
-\includegraphics{figures/high-wages-1.pdf}
-\caption{\label{fig:high-wages}6 out of 45 IDs with extremely high mean hourly wage. Most of the IDs only have one point of high wage.}
+
+{\centering \includegraphics[width=\textwidth]{figures/high-wages-1}
+
+}
+
+\caption{6 out of 45 IDs with extremely high mean hourly wage. Most of the IDs only have one point of high wage.}\label{fig:high-wages}
 \end{figure}
@@
 \begin{figure}
-\centering
-\includegraphics{figures/sample-plot-1.pdf}
-\caption{\label{fig:sample-plot}The mean hourly wages of some random samples are shown in twelve facets, three IDs per facet. It suggests that some IDs had a reasonably fluctuate wages.}
+
+{\centering \includegraphics[width=\textwidth]{figures/sample-plot-1}
+
+}
+
+\caption{The mean hourly wages of some random samples are shown in twelve facets, three IDs per facet. It suggests that some IDs had a reasonably fluctuate wages.}\label{fig:sample-plot}
 \end{figure}
@@
 \begin{figure}
-\centering
-\includegraphics{figures/compare-plot-1.pdf}
-\caption{\label{fig:compare-plot}Comparison between the original and the treated mean hourly wage.
+
+{\centering \includegraphics[width=\textwidth]{figures/compare-plot-1}
+
+}
+
+\caption{Comparison between the original and the treated mean hourly wage.
 The orange line portray the original value of mean hourly wage, while the turquoise line display the mean hourly wages value after the extreme values imputed with the robust linear model's prediction value.
-We can see that some extreme spikes has been reduced by the model.}
+We can see that some extreme spikes has been reduced by the model.}\label{fig:compare-plot}
 \end{figure}
@@
 \begin{figure}
-\centering
-\includegraphics{figures/comppict-1.pdf}
-\caption{\label{fig:comppict}The distribution of the mean hourly wage after the extreme values are replaced.
+
+{\centering \includegraphics[width=\textwidth]{figures/comppict-1}
+
+}
+
+\caption{The distribution of the mean hourly wage after the extreme values are replaced.
 Plot A portrays the pattern of mean hourly wage of high school cohort from 1979 to 2018 of each ID in US Dollar; Plot B shows the distribution of their minimum, median, and maximum value transformed to log10 scale.
 We can see that some observations still had reasonbaly higher wages than the others.
-Also, some IDs' have a minimum wages that is higher than others' maximum wages.}
+Also, some IDs' have a minimum wages that is higher than others' maximum wages.}\label{fig:comppict}
 \end{figure}
@@
 
-\begin{figure}
-\centering
-\includegraphics{figures/flow-chart-1.pdf}
-\caption{\label{fig:flow-chart}The stages of data filtering from the raw data to get three datasets that are contained in an R package called yowie, n means the number of ID, while n\_obs means the number of observations.}
-\end{figure}
+\textbackslash begin\{figure\}
+
+\{\centering \includegraphics[width=\textwidth]{figures/flow-chart-1}
 
@@
 
-This part gives some examples of how the data might be used for data analysis in a textbook.
-We used \texttt{brolgar} (Tierney, Cook, and Prvan 2020) to perform exploratory data analysis.
-In the first example, we observed the relationship between wages and education regarding how the wages increase year by year.
-Using \texttt{key\_slope} function in \texttt{brolgar}, we performed linear regression with the mean hourly wage as response variable and year as the predictor, extracted the regression slope of each ID, and plotted it according to the highest grade completed.
+This part gives some examples of exploratory data analysis using the data.
+In this example, we are interested to explore how people experience changes in their wages over the years.
+Hence, we focus the analysis in their slope of wages.
+Using \texttt{key\_slope} function in \texttt{brolgar} (Tierney, Cook, and Prvan 2020), we perform linear regression with the mean hourly wage as response variable and year as the predictor for each individual.
 
-Figure \ref{fig:slope-fig} shows that the higher the education completed does not necessarily increase the wages more from year to year.
-We can see that people who completed 4th grade happen to have relatively the same median of slope as people who completed 11th grade, even though it is probably because some respondents in this group had extremely high mean hourly wages.
-Moreover, the annual increasing wages of people who completed 6th grade to 11th grade are relatively the same.
-People who completed 12th grade tend to have a higher slope of the year than the other groups.
-Some outliers are also spotted in this group.
-Additionally, we can see that some respondents have a negative slope, meaning that their wages tend to decrease over the year.
+Exploring the wages slope of each individual:
+- Exploring the wages slope pattern based on the number of observation
 
 \begin{figure}
-\centering
-\includegraphics{figures/slope-fig-1.pdf}
-\caption{\label{fig:slope-fig}Regression slope of mean hourly wage regression by highest grade completed.
-The regression model has mean hourly wage as response variable and year as predictor.
-Respondents who completed 12th grade has the highest median of slope.}
+
+{\centering \includegraphics[width=\textwidth]{figures/slope-gender-1}
+
+}
+
+\caption{slope-gender}\label{fig:slope-gender}
 \end{figure}
 
-In the second example, we examined the pattern of mean hourly wage corresponding to different groups of gender and race.
-We took 240 samples of females and males equally and spread them into twelve facets, as shown in Figure \ref{fig:plot-gender} using \texttt{facet\_strata} function from \texttt{brolgar} (Tierney, Cook, and Prvan 2020).
-We learn that the mean hourly wage of females and males fluctuated over the years.
-For example, in facets 4, 5, and 10, males tend to earn more wages than females.
-This finding is also shown in Figure \ref{fig:se-gender} where we plotted the median wage of males and females against the year.
-We learn that males had a higher mean of mean hourly wage throughout the survey than females.
-Unfortunately, the gap between them tends to be wider from time to time.
+\begin{center}\includegraphics[width=\textwidth]{figures/slope-educ-1} \end{center}
 
-\begin{figure}
-\centering
-\includegraphics{figures/plot-gender-1.pdf}
-\caption{\label{fig:plot-gender}The pattern of mean hourly wages overtime displayed in 12 facets coloured by gender.
-The respondents belong to high school cohort are randomly sampled.
-We learn that both male and female respondents had a fluctuate mean hourly wage over the year.}
-\end{figure}
+\begin{center}\includegraphics[width=\textwidth]{figures/slope-nobs-1} \end{center}
 
-\begin{figure}
-\centering
-\includegraphics{figures/se-gender-1.pdf}
-\caption{\label{fig:se-gender}The yearly mean of mean hourly wage by gender along with its standard error.
-Males generally earn more wages than females.
-The error interval tend to be wider over the years.}
-\end{figure}
+According to OECD (n.d.), gender pay gap is the difference between median wage of males and females relative to median wage of males.
+We find that, as shown in \ref{fig:se-gender}, over the years, females tend to earn fewer than males.
+For example, in 2018 females earn about 78 cents for every \$1 earned by males.
+The widest gender gaps were observed in 1990s.
 
-For the final example, we showed modeling the longitudinal data using the Generalized Additive Model (GAM) with \texttt{mgcv} package (Wood 2003) for each ID.
-We fitted a more flexible model to fit into the data since the relationship between mean hourly wage and year is not linear due to the spikes.
-Since this approach is computationally expensive, we only showed the model in a small fraction of the data.
-We sampled 1 percent of the total respondents using \texttt{brolgar\textquotesingle{}s} \texttt{sample\_frac\_keys} function.
-Further, we plotted some IDs overlayed by the fitted GAM model in Figure \ref{fig:plot-gam}.
-We learn that the model flexible enough to deal with the fluctuation in the data.
-Moreover, the data with small spikes tend to be linearly fitted.
+\textbackslash begin\{figure\}
 
-\begin{figure}
-\centering
-\includegraphics{figures/plot-gam-1.pdf}
-\caption{\label{fig:plot-gam}Exploration of wages data by fitting a GAM.
-The fitted model displayed by blue line.
-The fitted line shows that the model flexible enough to follow the pattern of the data.}
-\end{figure}
+\{\centering \includegraphics[width=\textwidth]{figures/wages-gender-gap-1}
+
+\}
+
+\textbackslash caption\{Gender wage gap or the median wages of females relatives to median wages of males.
+Females tend to earn fewer than males.
+For example, in 2018 females earn about 78 cents for every \$1 earned by males.\}\label{fig:wages-gender-gap}
+\textbackslash end\{figure\}
 
@@
 
+\leavevmode\hypertarget{ref-OECD}{}%
+OECD. n.d. ``Gender Wage Gap.'' https://data.oecd.org/earnwage/gender-wage-gap.htm.
+
 \leavevmode\hypertarget{ref-rsvg}{}%

### 2021-09-07 paper/report-alt.R

@@ -1,693 +1,0 @@
-## ----setup, include = FALSE---------------------------------------------------
-knitr::opts_chunk$set(echo = TRUE, 
-                      warning = FALSE, 
-                      message = FALSE,
-                      cache = TRUE, 
-                      cache.path = "cache/",
-                      fig.path = "figures/")
-library(tidyverse)
-library(brolgar)
-library(patchwork)
-library(kableExtra)
-library(MASS)
-library(janitor)
-library(DiagrammeR)
-library(rsvg)
-library(webshot)
-library(mgcv)
-library(modelr)
-library(tsibble)
-
-
-## ---- cache = TRUE, echo = FALSE----------------------------------------------
-source(here::here("data-raw/NLSY79/NLSY79.R"))
-
-
-## ----untidy-data, echo = FALSE------------------------------------------------
-
-untidy_demo <- new_data_qnames %>%
-  as_tibble() %>%
-  # in 2018, the variable's name is Q3-4_2018, instead of HGC_2018
-  rename(HGC_2018 = `Q3-4_2018`) %>%
-  dplyr::select(CASEID_1979,
-            starts_with("HRP") &
-              ends_with(c("1979", "1980", "1981", "1982", "1983"))) %>%
-  dplyr::select(1:7) %>%
-  head()
-
-kable(untidy_demo,
-      caption = "The Untidy Form of the NLSY79 Raw Data") %>%
-  kable_styling(latex_options = "striped")
-
-
-## ----dob-tidy-----------------------------------------------------------------
-## tidy the date of birth data
-dob_tidy <- new_data_qnames %>%
-  dplyr::select(CASEID_1979,
-         starts_with("Q1-3_A~")) %>%
-  mutate(dob_year = case_when(
-                    # if the years recorded in both sets match, take 79 data
-                    `Q1-3_A~Y_1979` == `Q1-3_A~Y_1981` ~ `Q1-3_A~Y_1979`,
-                    # if the year in the 81 set is missing, take the 79 data
-                    is.na(`Q1-3_A~Y_1981`) ~ `Q1-3_A~Y_1979`,
-                    # if the sets don't match for dob year, take the 79 data
-                    `Q1-3_A~Y_1979` != `Q1-3_A~Y_1981` ~ `Q1-3_A~Y_1979`),
-        dob_month = case_when(
[…]

### 2021-09-07 paper/scripts/data-preprocessing.R

@@ -1,0 +1,237 @@
+## ---- load-pkgs
+
+library(tidyverse)
+library(brolgar)
+library(patchwork)
+library(kableExtra)
+library(MASS)
+library(janitor)
+library(DiagrammeR)
+library(rsvg)
+library(webshot)
+library(mgcv)
+library(modelr)
+library(tsibble)
+library(lme4)
+library(modelr)
+library(forcats)
+library(yowie)
+
+## ---- raw-data
+source(here::here("data-raw/NLSY79/NLSY79.R"))
+
+## ---- untidy-data
+untidy_demo <- new_data_qnames %>%
+  as_tibble() %>%
+  # in 2018, the variable's name is Q3-4_2018, instead of HGC_2018
+  rename(HGC_2018 = `Q3-4_2018`) %>%
+  dplyr::select(CASEID_1979,
+                starts_with("HRP") &
+                  ends_with(c("1979", "1980", "1981", "1982", "1983"))) %>%
+  dplyr::select(1:7) %>%
+  head()
+
+kable(untidy_demo,
+      caption = "The Untidy Form of the NLSY79 Raw Data") %>%
+  kable_styling(latex_options = "striped")
+
+## ---- dob-tidy
+#tidy the date of birth data
+dob_tidy <- new_data_qnames %>%
+  dplyr::select(CASEID_1979,
+                starts_with("Q1-3_A~")) %>%
+  mutate(dob_year = case_when(
+    # if the years recorded in both sets match, take 79 data
+    `Q1-3_A~Y_1979` == `Q1-3_A~Y_1981` ~ `Q1-3_A~Y_1979`,
+    # if the year in the 81 set is missing, take the 79 data
+    is.na(`Q1-3_A~Y_1981`) ~ `Q1-3_A~Y_1979`,
+    # if the sets don't match for dob year, take the 79 data
+    `Q1-3_A~Y_1979` != `Q1-3_A~Y_1981` ~ `Q1-3_A~Y_1979`),
+    dob_month = case_when(
+      # if months recorded in both sets match, take 79 data
+      `Q1-3_A~M_1979` == `Q1-3_A~M_1981` ~ `Q1-3_A~M_1979`,
+      # if month in 81 set is missing, take the 79 data
+      is.na(`Q1-3_A~M_1981`) ~ `Q1-3_A~M_1979`,
+      # if sets don't match for dob month, take the 79 data
[…]

### 2021-09-07 paper/scripts/ida.R

@@ -1,0 +1,313 @@
+## ---- age-table
+age_table <- yowie::demog_nlsy79 %>%
+group_by(age_1979) %>%
+  count(age_1979)
+
+kable(age_table,
+      caption = "Age Distribution of the NLSY79 samples",
+      col.names = c("Age", "Number of Sample")) %>%
+  kable_styling(latex_options = "striped")
+
+## ---- gender-race-table
+gender_race_table <- yowie::demog_nlsy79 %>%
+  tabyl(gender, race) %>%
+  adorn_totals(c("row", "col")) %>%
+  adorn_percentages("row") %>%
+  adorn_pct_formatting(digits = 2) %>%
+  adorn_ns(position = "front") %>%
+  mutate(gender = ifelse(gender == "MALE", "Male",
+                         ifelse(gender == "FEMALE", "Female", "Total")))
+
+kable(gender_race_table,
+      caption = "Gender and Race Distribution of the NLSY79 Samples",
+      col.names = c("Gender", "Hispanic", "Black", "Non-Black, Non-Hispanic", "Total")) %>%
+  kable_styling(latex_options = "striped") %>%
+  add_header_above(c(" " = 1, "Race" = 3, " " = 1))
+
+## ---- summarytable
+kable(as.array(summary(wages_demog_hs$mean_hourly_wage)),
+      caption = "Summary Statistics of Wages of High School Data",
+      col.names = c("Statistics", "Value")) %>%
+  kable_styling(latex_options = "striped")
+
+## ---- feature-plot
+spag <- wages_demog_hs %>%
+  ggplot(aes(x = year,
+             y = mean_hourly_wage,
+             group = id)) +
+  geom_line(alpha = 0.1) +
+  ggtitle("A)") +
+  theme(plot.title = element_text(size = 10))
+
+wages_demog_hs_tsibble <- as_tsibble(x = wages_demog_hs,
+                                     key = id,
+                                     index = year,
+                                     regular = FALSE)
+wages_three_feat <- wages_demog_hs_tsibble %>%
+  features(mean_hourly_wage,
+           feat_three_num
+  )
+wages_feat_long <- wages_three_feat %>%
+  pivot_longer(c(min, med, max), names_to = "feature", values_to = "value")
+feature <- ggplot(wages_feat_long) +
+  geom_density(aes(x = value, colour = feature, fill = feature), alpha = 0.3) +
[…]

### 2021-09-07 paper/paper.Rmd
Edit EDA part and add it's script.

@@
 read_chunk(here::here("paper/scripts/ida.R"))
+read_chunk(here::here("paper/scripts/eda.R"))
 ```
@@
 
-```{r}
-wages <- yowie::wages %>%
-  mutate(year0 = year - 1979,
-         ln_wages = log(wage))
-
-wages_slope <- key_slope(wages, ln_wages ~ year0) %>%
-  left_join(yowie::demog_nlsy79, by = "id") %>%
-  mutate(hgc_regroup = fct_collapse(hgc, `ungraded` = "UNGRADED",
-                                    `8TH grade or less` = c("1ST GRADE", "3RD GRADE",
-                                                                 "4TH GRADE", "5TH GRADE",
-                                                                 "6TH GRADE", "7TH GRADE",
-                                                                 "8TH GRADE"),
-                                    `9TH to 11TH grade` = c("9TH GRADE",
-                                                            "10TH GRADE",
-                                                            "11TH GRADE"),
-                                    `12TH grade` = "12TH GRADE")) %>%
-  mutate(hgc_regroup = factor(hgc_regroup, levels = c("UNGRADED",
-                                                      "8TH grade or less", 
-                                                      "9TH to 11TH grade", 
-                                                      "12TH grade")))
-
-```
-
 This part gives some examples of exploratory data analysis using the data.
-In this example, we are interested to explore how people experience changes in their wages over the years.
+We are interested to explore how people experience changes in their wages over the years.
 Hence, we focus the analysis in their slope of wages.
@@
 
-Exploring the wages slope of each individual:
-  - Exploring the wages slope pattern based on the number of observation
-
-```{r slope-gender, fig.cap="slope-gender"}
-slope_bottom_5<- wages_slope %>%
-  arrange(.slope_year0) %>%
-  slice_head(n = 5) %>%
-  arrange(desc(.slope_year0)) %>%
-  mutate(note = "bottom 5") %>%
-  mutate(rank = seq_along(1:5))
-
-slope_top_5<- wages_slope %>%
-  arrange(desc(.slope_year0)) %>%
-  slice_head(n = 5) %>%
-  mutate(note = "top 5") %>%
-  mutate(rank = seq_along(1:5))
-
-keep_contrast <- rbind(slope_bottom_5, slope_top_5) %>%
-  dplyr::select(.slope_year0, id, note, rank, hgc_regroup)
-
-contrast_slope <- left_join(keep_contrast, wages, by = "id") %>%
-  arrange(.slope_year0) %>%
-  mutate(id = factor(id, levels = c("7918", "11695", "9176", "10119", "6464", "1271", 
-                                    "12371", "9039", "10217", "7782")),
-         note = factor(note, levels = c("top 5", "bottom 5")))
-
-ggplot(contrast_slope, aes(x = year, y = ln_wages, color = gender)) +
-  geom_point() +
-  geom_smooth(method = "lm", se = FALSE) +
-  facet_grid(note ~ rank) +
-  scale_color_manual(values = c("#04B4AE", "#BF00FF"))
-
+```{r nobs-plot, fig.cap="Distribution of observations. The numbers of rounds varies among respondents. Most of respondents participated in 28 rounds of survey."}
 ```
 
-```{r slope-educ}
-ggplot(contrast_slope, aes(x = year, y = ln_wages, color = hgc_regroup)) +
-  geom_point() +
-  geom_smooth(method = "lm", se = FALSE) +
-  facet_grid(note ~ rank)
+```{r slope}
 ```
 
-```{r slope-nobs}
-wages_obs <- wages %>%
-  features(wage, n_obs) %>%
-  left_join(wages_slope, by = "id")
+```{r slope-nobs, fig.cap="Distribution of wages slope by number of observations. The slopes' range are more stable with more observations."}
+```
 
-ggplot(wages_obs, aes(x = n_obs, y = .slope_year0)) +
-  geom_point() +
-  geom_jitter()
+```{r trim-slope}
+```
+
+```{r slope-gender, fig.cap = "Wages slope over the years based on gender.
+Most of respondents who experienced high wages slope are males.
+In contrast, respondents who experienced lower wages slope are mostly females."}
 ```
@@
 Females tend to earn fewer than males.
-For example, in 2018 females earn about 78 cents for every $1 earned by males."}
+For example, in 2018 females earn about 78 cents for every $1 earned by males.
+Note that this is a crude measure as gap is calculated not based on other variable, such as experience and job class."}
+```
 
-wages_gender <- as.tibble(wages) %>%
-  group_by(gender, year) %>%
-  summarise(mean = mean(wage),
-            median = median(wage),
-            sd = sd(wage),
-            samplesize = n()) %>%
-  ungroup() %>%
-  mutate(upper = mean + sd/sqrt(samplesize),
-         lower = mean - sd/sqrt(samplesize))
-
-wage_female <- wages_gender %>%
-  filter(gender == "FEMALE") %>%
-  dplyr::select(year, median) %>%
-  rename(female = median)
-
-wage_male <- wages_gender %>%
-  filter(gender == "MALE") %>%
-  dplyr::select(median) %>%
-  rename(male = median)
-
-gender_wage_gap <- cbind(wage_female, wage_male) %>%
-  mutate(gap = (male-female)/male) %>%
-  mutate(relative = 1-gap)
-
-ggplot(gender_wage_gap) +
-  geom_line(aes(x = year,
-                y = relative),
-            color = "#DF7401", size = 0.8) +
-  geom_point(aes(x = year,
-                y = relative),
-             color = "#DF7401", size = 2) +
-  geom_hline(yintercept = 1, colour = "#088A68", size = 1) +
-  scale_x_continuous(limits = c(1978.7,2018.3), expand = c(0, 0)) +
-  ylab("females' median wages relative to males (US$)") +
-  annotate("text", x = 2014.5, y = 1, label = "$1 of male wages", vjust = 1.4) +
-  annotate("text", x = 2012, y = 0.8, label = "female wages relative to male", vjust = 1)
+```{r slope-educ, fig.cap = "Wages slope over the years based on highest education completed. People who are in top-5 and bottom-5 wages are dominated by people who completed high school."}
 ```

### 2021-09-07 paper/paper.tex
Edit EDA part and add it's script.

@@
 This part gives some examples of exploratory data analysis using the data.
-In this example, we are interested to explore how people experience changes in their wages over the years.
+We are interested to explore how people experience changes in their wages over the years.
 Hence, we focus the analysis in their slope of wages.
@@
 
-Exploring the wages slope of each individual:
-- Exploring the wages slope pattern based on the number of observation
+\begin{figure}
+
+{\centering \includegraphics[width=\textwidth]{figures/nobs-plot-1}
+
+}
+
+\caption{Distribution of observations. The numbers of rounds varies among respondents. Most of respondents participated in 28 rounds of survey.}\label{fig:nobs-plot}
+\end{figure}
+
+\begin{figure}
+
+{\centering \includegraphics[width=\textwidth]{figures/slope-nobs-1}
+
+}
+
+\caption{Distribution of wages slope by number of observations. The slopes' range are more stable with more observations.}\label{fig:slope-nobs}
+\end{figure}
 
@@
 
-\caption{slope-gender}\label{fig:slope-gender}
+\caption{Wages slope over the years based on gender.
+Most of respondents who experienced high wages slope are males.
+In contrast, respondents who experienced lower wages slope are mostly females.}\label{fig:slope-gender}
 \end{figure}
-
-\begin{center}\includegraphics[width=\textwidth]{figures/slope-educ-1} \end{center}
-
-\begin{center}\includegraphics[width=\textwidth]{figures/slope-nobs-1} \end{center}
 
@@
 Females tend to earn fewer than males.
-For example, in 2018 females earn about 78 cents for every \$1 earned by males.\}\label{fig:wages-gender-gap}
+For example, in 2018 females earn about 78 cents for every \$1 earned by males.
+Note that this is a crude measure as gap is calculated not based on other variable, such as experience and job class.\}\label{fig:wages-gender-gap}
 \textbackslash end\{figure\}
+
+\begin{figure}
+
+{\centering \includegraphics[width=\textwidth]{figures/slope-educ-1}
+
+}
+
+\caption{Wages slope over the years based on highest education completed. People who are in top-5 and bottom-5 wages are dominated by people who completed high school.}\label{fig:slope-educ}
+\end{figure}
 

### 2021-09-07 paper/scripts/eda.R

@@ -1,0 +1,120 @@
+## ---- nobs-plot
+wages_obs <- yowie::wages %>%
+  features(wage, n_obs) %>%
+  arrange(n_obs)
+ggplot(wages_obs, aes(x = n_obs)) +
+  geom_bar() +
+  xlab("number of observations") +
+  theme_bw()
+
+## ---- slope
+wages <- yowie::wages %>%
+  mutate(year0 = year - 1979,
+         ln_wages = log(wage))
+
+wages_slope <- key_slope(wages, ln_wages ~ year0) %>%
+  left_join(yowie::demog_nlsy79, by = "id") %>%
+  mutate(hgc_regroup = fct_collapse(hgc, `ungraded` = "UNGRADED",
+                                    `8TH grade or less` = c("1ST GRADE", "3RD GRADE",
+                                                            "4TH GRADE", "5TH GRADE",
+                                                            "6TH GRADE", "7TH GRADE",
+                                                            "8TH GRADE"),
+                                    `9TH to 11TH grade` = c("9TH GRADE",
+                                                            "10TH GRADE",
+                                                            "11TH GRADE"),
+                                    `12TH grade` = "12TH GRADE")) %>%
+  mutate(hgc_regroup = factor(hgc_regroup, levels = c("UNGRADED",
+                                                      "8TH grade or less",
+                                                      "9TH to 11TH grade",
+                                                      "12TH grade")))
+## ---- slope-nobs
+wages_obs_slope <- left_join(wages_obs, wages_slope, by = "id")
+ggplot(wages_obs_slope, aes(x = n_obs, y = .slope_year0)) +
+  geom_point() +
+  geom_jitter() +
+  xlab("number of observations") +
+  ylab("slope") +
+  theme_bw()
+
+## ---- trim-slope
+slope_trimmed <- wages_slope %>%
+  filter(.slope_year0 < quantile(.slope_year0, 0.95) &
+           .slope_year0 > quantile(.slope_year0, 0.05))
+
+## ---- slope-gender
+slope_bottom_5<- slope_trimmed %>%
+  arrange(.slope_year0) %>%
+  slice_head(n = 5) %>%
+  arrange(desc(.slope_year0)) %>%
+  mutate(note = "bottom 5") %>%
[…]

### 2021-09-07 paper/scripts/ida.R

@@ -37,7 +37,8 @@ spag <- wages_demog_hs %>%
              group = id)) +
   geom_line(alpha = 0.1) +
   ggtitle("A)") +
-  theme(plot.title = element_text(size = 10))
+  theme(plot.title = element_text(size = 10)) +
+  theme_bw()
 
 wages_demog_hs_tsibble <- as_tsibble(x = wages_demog_hs,
                                      key = id,
@@ -52,7 +53,8 @@ wages_feat_long <- wages_three_feat %>%
 feature <- ggplot(wages_feat_long) +
   geom_density(aes(x = value, colour = feature, fill = feature), alpha = 0.3) +
   ggtitle("B)") +
-  theme(plot.title = element_text(size = 10))
+  theme(plot.title = element_text(size = 10))+
+  theme_bw()
 
 spag + feature
 
@@ -73,7 +75,8 @@ ggplot(wages_high2) +
              alpha = 0.5) +
   facet_wrap(~id, scales = "free_y") +
   theme(axis.text.x = element_text(angle = 10, size = 6)) +
-  ylab("mean hourly wage")
+  ylab("mean hourly wage") +
+  theme_bw()
 
 ## ---- sample-plot
 set.seed(20210225)
@@ -84,7 +87,8 @@ ggplot(wages_demog_hs_tsibble,
   geom_line(alpha = 0.7) +
   facet_sample() +
   theme(axis.text.x = element_text(angle = 10, size = 6)) +
-  ylab("mean hourly wage")
+  ylab("mean hourly wage") +
+  theme_bw()
 
 ## ---- rlm
 # nest the data by id to build a robust linear model
@@ -170,7 +174,8 @@ ggplot(wages_compare) +
              size = 1) +
   theme(axis.text.x = element_text(angle = 10, size = 5),
         legend.position = "bottom") +
-  facet_wrap(~id, scales = "free_y")
+  facet_wrap(~id, scales = "free_y") +
+  theme_bw()
 
 ## ---- comppict
 spag2 <- wages_demog_hs %>%
@@ -179,7 +184,8 @@ spag2 <- wages_demog_hs %>%
              group = id)) +
   geom_line(alpha = 0.1) +
   ggtitle("A)") +
-  theme(plot.title = element_text(size = 10))
+  theme(plot.title = element_text(size = 10)) +
+  theme_bw()
 
 wages_hs2020_rlm <- as_tsibble(x = wages_demog_hs,
[…]

### 2021-09-07 paper/scripts/data-preprocessing.R
Removed striped styling in kable.

@@ -33,7 +33,7 @@ untidy_demo <- new_data_qnames %>%
 
 kable(untidy_demo,
       caption = "The Untidy Form of the NLSY79 Raw Data") %>%
-  kable_styling(latex_options = "striped")
+  kable_styling()
 
 ## ---- dob-tidy
 #tidy the date of birth data

### 2021-09-07 paper/scripts/ida.R

@@ -6,7 +6,7 @@ group_by(age_1979) %>%
 kable(age_table,
       caption = "Age Distribution of the NLSY79 samples",
       col.names = c("Age", "Number of Sample")) %>%
-  kable_styling(latex_options = "striped")
+  kable_styling()
 
 ## ---- gender-race-table
 gender_race_table <- yowie::demog_nlsy79 %>%
@@ -21,14 +21,14 @@ gender_race_table <- yowie::demog_nlsy79 %>%
 kable(gender_race_table,
       caption = "Gender and Race Distribution of the NLSY79 Samples",
       col.names = c("Gender", "Hispanic", "Black", "Non-Black, Non-Hispanic", "Total")) %>%
-  kable_styling(latex_options = "striped") %>%
+  kable_styling() %>%
   add_header_above(c(" " = 1, "Race" = 3, " " = 1))
 
 ## ---- summarytable
 kable(as.array(summary(wages_demog_hs$mean_hourly_wage)),
       caption = "Summary Statistics of Wages of High School Data",
       col.names = c("Statistics", "Value")) %>%
-  kable_styling(latex_options = "striped")
+  kable_styling()
 
 ## ---- feature-plot
 spag <- wages_demog_hs %>%

### 2021-09-08 paper/paper.Rmd
Check spelling and grammar.

@@
 ---
-title: "The Journey from Wild to Text Book Data: A Case Study from the National Longitudinal Survey of Youth"
+title: "The Journey from Wild to Textbook Data: A Case Study from the National Longitudinal Survey of Youth"
 authors:
@@
                comment = "#>",
-               fig.align = 'center', fig.show = 'hold',
-               out.width = ifelse(is_html_output(), "100%","\\textwidth"))
+               fig.align = 'center', #fig.show = 'hold',
+               out.width = "70%")
 
@@
 NSLY79 is a prominent open data source that has been an important support in multidisciplinary research on longitudinal data.
-Subsets of this data can be found in numerous textbooks and research articles, however the steps and decisions taken to get from the original data to these subsets is never clearly articulated.
+Subsets of this data can be found in numerous textbooks and research articles.
+However, the steps and decisions taken to get from the original data to these subsets is never clearly articulated.
 This article describes our journey when trying to re-create a textbook example data set from the original data, with a goal to refresh it for use in the classroom.
@@
 
-# Introduction
+# Introduction {#intro}
 
-"Open data" is data that are freely accessible, modifiable, and shareable by anyone for any purpose [@opendata].
-This type of data can be useful as example data in statistical text books and for research purposes.
-However, open data are often, what we might call "wild data", because it requires substantial cleaning and tidying to tame it into text book shape.
+"Open data" is data that is freely accessible, modifiable, and shareable by anyone for any purpose [@opendata].
+This type of data can be useful as example data in statistical textbooks and for research purposes.
+However, open data is often referred to as what we might call "wild data" because it requires substantial cleaning and tidying to tame it into textbook shape.
 @HuebnerMariannePhD2016Asat emphasize that making the data cleaning process accountable and transparent is imperative.
@@
 Data cleaning is part of what is called "initial data analysis (IDA)" [@Chatfield1985TIEo].
-The other IDA steps are to explore the data properties and check assumption of modeling to document and report the process for the later formal analysis.
+The other IDA steps are to explore the data properties and check the assumption of modeling to document and report the process for the later formal analysis.
 Hence, exploratory data analysis, which is defined by @tukey as a detective work to get the clue from data, either numerically or graphically, before confirmatory data analysis is performed, encompasses IDA.
@@
 There are few research papers that document the data cleaning [@WickhamHadley2014TD].
-Furthermore, the decisions made in this stage often go unreported  in the sense that IDA is often performed in an unplanned and unstructured way, and is only shared among restricted parties [@HuebnerMarianne2020Haar].
+Furthermore, the decisions made in this stage often go unreported in the sense that IDA is often performed in an unplanned and unstructured way and is only shared among restricted parties [@HuebnerMarianne2020Haar].
 
 This paper aims to demonstrate the steps of IDA,  tidying, cleaning, and documenting the process, for a prominent open data source, National Longitudinal Survey of Youth (NLSY) 1979, henceforth referred to NSLY79.
-This data has been playing an important role in research in various disciplines including but are not limited to economics, sociology, education, public policy, and public health for more than a quarter of the century [@MichaelRPergamit2001DWTN].
-In addition, The National Longitudinal Survey is considered as survey with high retention rates and carefully designed making it suitable for a life course research [@MichaelRPergamit2001DWTN] and [@eliznlsy].
+This data has been playing an important role in research in various disciplines, including but are not limited to economics, sociology, education, public policy, and public health for more than a quarter of the century [@MichaelRPergamit2001DWTN].
+In addition, The National Longitudinal Survey is considered a survey with high retention rates and carefully designed, making it suitable for a life course research [@MichaelRPergamit2001DWTN] and [@eliznlsy].
 According to @eliznlsy, thousand of articles, and hundreds of book chapters and monographs has incorporated the NLSY data.
-Moreover, the NLSY79 is considered as the most widely used and most important cohort in the NLSY79 data sets [@MichaelRPergamit2001DWTN].
+Moreover, the NLSY79 is considered the most widely used and most important cohort in the NLSY79 data sets [@MichaelRPergamit2001DWTN].
 
 @SingerJudithD2003Alda used the wages and other variables of high school dropouts from the NLSY79 data as an example data set to illustrate longitudinal data modeling.
-Our aim is to refresh this text book data to be more up to date.
-Here, we investigate the process of getting from the raw NLSY79 to this text book data set.
-However, we are not able to create the exact same data set as published in their book since we do not have the information of what age threshold did they use to determine the high school dropouts.
+Our aim is to refresh this textbook data to be more up to date.
+Here, we investigate the process of getting from the raw NLSY79 to this textbook data set.
+However, we cannot create the exact same data set as published in their book since we do not have the information of what age threshold they used to determine the high school dropouts.
 
 This paper is structured to have 5 sections.
-Section 2 describes the original data source.
-Section 3 presents the steps of cleaning the data, including getting and tidying the data from the NLSY79 and initial data analysis to find and treat the anomalies in the data.
-We also gave examples of exploratory data analysis using the clean data in Section 4.
-Finally, Section 6 summarizes the paper.
+Section \@ref(database) describes the original data source.
+Section \@ref(cleaning) presents the steps of cleaning the data, including getting and tidying the data from the NLSY79 and initial data analysis to find and treat the anomalies in the data.
+We also gave examples of exploratory data analysis using the clean data in Section \@ref(eda).
+Finally, Section \@ref(summary) summarises the paper.
 
-# The NLSY79
+# The NLSY79 {#database}
 
@@
 It was comprised of Blacks, Hispanics, economically disadvantaged non-Black non-Hispanics, and youth in the military.
-In 1984 and 1990, two sub-samples were dropped from the interview; the dropped subjects are the 1,079 members of the military sample and 1,643 members of the economically disadvantaged non-Black non-Hispanics respectively.
+In 1984 and 1990, two sub-samples were dropped from the interview; the dropped subjects are the 1,079 members of the military sample and 1,643 members of the economically disadvantaged non-Black non-Hispanics, respectively.
 Hence, 9,964 respondents remain in the eligible samples.
@@
 
-Although the main focus area of the NLSY is labor and employment, the NLSY also cover several other topics including education; training and achievement; household, geography and contextual variables; dating, marriage, cohabitation; sexual activity, pregnancy, and fertility; children; income, assets and program participation; health; attitudes and expectations; and crime and substance use.
+Although the main focus area of the NLSY is labor and employment, the NLSY also covers several other topics, including education; training and achievement; household, geography, and contextual variables; dating, marriage, cohabitation; sexual activity, pregnancy, and fertility; children; income, assets and program participation; health; attitudes and expectations; and crime and substance use.
 
-There are two ways to conduct the interview of the NLSY79, which are face to face interview or by telephone.
+There are two ways to conduct the interview of the NLSY79, which are face-to-face interviews or by telephone.
 In recent survey years, more than 90 percent of respondents were interviewed by telephone [@eliznlsy].
@@
 
-As mentioned in Section 1, this paper aims to refresh the NLSY79 data as used in Singer and Willet's text book (@SingerJudithD2003Alda).
+As mentioned in Section \@ref(intro), this paper aims to refresh the NLSY79 data as used in Singer and Willet's textbook (@SingerJudithD2003Alda).
 This data set contains the information of yearly mean hourly wages from the NLSY79 cohort, with education and race as covariates.
 This data set represents the measurement of male high-school dropouts, aged from 14 to 17 years old when first time measured.
-Hence, in this paper, after getting the data set of high school graduates, we subset it by only having the high school dropouts.
+Accordingly, to get this data, we have two additional target data, NSLY79 cohort's demographic profiles and wages data of people who completed education up to high school.
 
-Since there is no information of high school drop-outs in Singer and Willet's book, we define high school dropouts as respondents who only completed either $9^{th}$, $10^{th}$, $11^{th}$ grade, or who completed $12^{th}$ when their age are at least 19 years old (means that these IDs being dropped out in certain year, but they returned back to high school to complete it).
+Since there is no information on high school dropouts in Singer and Willet's book, we define high school dropouts as respondents who only completed either $9^{th}$, $10^{th}$, $11^{th}$ grade, or who completed $12^{th}$ when their age is at least 19 years old (means that these IDs being dropped out in a certain year, but they returned back to high school to complete it).
 
-# The NLSY79 Data Cleaning
+# The NLSY79 Data Cleaning {#cleaning}
 
@@
 
-- Education, Training & Achievement Scores
+**Education, Training & Achievement Scores**
+
 - Education -> Summary measures -> All schools -> By year -> Highest grade completed
-       - Downloaded all of the 78 variables in Highest grade completed.
-- Employment
+  - Downloaded all of variables in Highest grade completed.
+
+**Employment**
+
 - Summary measures -> By job -> Hours worked and Hourly wages
-      - Downloaded all of the 427 variables in Hours worked
-      - Downloaded all of the 151 variables in Hourly wages
+  - Downloaded all of the variables in Hours worked.
+  - Downloaded all of the variables in Hourly wages
+  Both hours worked and hourly wages are recorded by the job, up to five jobs for each id/subject.
 
-      Both hours worked and hourly wages are recorded by the job, up to five jobs for each id/subject. 
-- Household, Geography & Contextual Variables
+**Household, Geography & Contextual Variables**
+
 - Context -> Summary measures -> Basic demographics 
   - Downloaded year and month of birth, race, and sex variables.
-
-      There are two versions of the year and month of birth, i.e. 1979 and 1981 data. We downloaded these two versions.
+  There are two versions of the year and month of birth, i.e., 1979 and 1981 data. We downloaded these two versions.
 
@@
 
-- NLSY79.NLSY79: This is the tagset of variables that can be uploaded to the web site to recreate the data set.
-- NLSY79.R: This is an R script provided automatically by the database for reading the data into R and convert the variables' name and its label into something more sensible.
+- NLSY79.NLSY79: Tagset of variables that can be uploaded to the website to recreate the data set.
+- NLSY79.R: R script provided automatically by the database for reading the data into R and converting the variables' names and label into something more sensible.
 We utilized this code to do an initial data tidying.
@@
 
-According to @WickhamHadley2014TD, a tidy data sets comply with three rules, the first is that each variable forms a column, the second is that each observation forms a row, and the last is that each type of observational unit forms a table.
-Unfortunately, the `new_data_qnames` did not meet these requirements in the way that the value for particular year and job are stored in different columns, hence the data contains a huge number of columns (686 columns).
-The example of the untidy of the data set is displayed in Table \ref{tab:untidy-data}, which actually intended to display the hourly rate of each respondent by job (HRP1 to HRP5) and by year (1979 and 1980).
+According to @WickhamHadley2014TD, tidy data sets comply with three rules.
+Firstly, each variable forms a column.
+Secondly, each observation forms a row.
+Lastly, each type of observational unit forms a table.
+Unfortunately, the `new_data_qnames` did not meet these requirements in the way that data for each year and job for the same respondent is stored as one variable.
+Hence, the data contains a huge number of columns (686 columns).
+The example of the untidy data set is displayed in Table \ref{tab:untidy-data}.
+The table intended to display the hourly rate of each respondent by job (HRP1 to HRP5) and by year (1979 and 1980).
 The table implies that the column headers are values of the year and job, not variable names.
+Consequently, the data should be tidied and wrangled first to extract the demographic and employment variables we want to put in the final data set.
+We mainly used `tidyr` [@tidyr], `dplyr` [@dplyr], and `stringr` [@stringr] to do this job.
 
@@
 
-Consequently, the data should be tidied and wrangled first to extract the demographic and employment variables that we want to put in the final data set.
-We mainly used `tidyr` [@tidyr], `dplyr` [@dplyr], and `stringr` [@stringr] to do this job.
-
 ### Tidying demographic variables
 
-`Age in 1979`, `gender`, `race`, `highest grade completed` (factor and integer), and the `year when the highest grade completed` are the variables that we want to use in the cleaned data set.
+`Age in 1979`, `gender`, `race`, `highest grade completed`, and the `year when the highest grade completed` are the variables that we want to use in the cleaned data set.
 
-Age in 1979 are derived from year of birth and month of birth variables in the raw data set.
+Age in 1979 are derived from year and month of birth variables in the raw data.
 The variables have two versions, which are the 1979 version and the 1981 version.
@@
 
-For `gender` and `race`, we only renamed these variables.
+For `gender` and `race`, we only rename the column names.
+It is worth noting that using these two variable needs special attention.
+Gender, as reported in the data, only has two categories, which is recognised today as inadequate.
+Gender is not binary.
+Further, race is as reported in the database.
+When doing analysis with this variable, one should keep in mind that the purpose is to study racism rather than race.
 
-```{r demog-tidy, echo = TRUE}
+```{r demog-tidy}
 ```
@@
 This variable came with several version in each year.
-We chose the revised May data because it seemed to have less missing and presumably has been checked.
-However, there was no revised May data for 2012, 2014, 2016, and 2018 so we just used the ordinary May data for these years.
-The code for tidying this variable is provided in the supplementary code whose details could be seen in the Supplementary Materials section of this paper.
+We choose the revised May data because it seemed to have less missing and presumably has been checked.
+However, there is no revised May data for 2012, 2014, 2016, and 2018.
+Thus, we just use the ordinary May data for these years.
+The code for tidying this variable is provided in the supplementary materials.
 
@@
 Hence, we only used the highest grade completed ever and derived the year when it is completed.
-The code for tidying this variable is also available in the supplementary code of this paper.
+The code for tidying this variable is also available in the supplementary materials.
 
@@
 
-Finally, we join all of the tidy variables in a data set called `full_demographics` as displayed as follows:
+Finally, we get all of the demographic profiles of the NLSY79 cohort. We then save this data as `demog_nlsy79`.
 
@@
 
-The employment data comprises of three variables, i.e. `total hours of work per week`, `number of jobs that an individual has`, and `mean hourly wage`.
+The employment data comprises of three variables, i.e., `total hours of work per week`, `number of jobs that an individual has`, and `mean hourly wage`.
 
 `hours worked per week` initially has only one version per job, no choice from 1979 to 1987 (QES-52A).
-From 1988 onward, when we had more options, we chose the variable for total hours including time spent working from home (QES-52D).
+From 1988 onward, when we had more options, we chose the variable for total hours, including time spent working from home (QES-52D).
 However, in 1993, this variable did not have all the five D variables (the first one and the last one were missing), so we used QES-52A variable instead.
-In addition, 2008 only had jobs 1-4 for the QES-52D variable (whereas the other years had 1-5), so we just used these.
+In addition, 2008 only had jobs 1-4 for the QES-52D variable (whereas the other years had 1-5), so we use these.
 
@@
 The same way was also deployed to tidy the rate of wage by year and by ID.
-The difference is that the hourly rate had only one version of each year.
-The hours of work and the hourly rate were then joined to calculate the number of jobs that a respondent has and their mean hourly wage.
-Some observations had 0 in their hourly rate, which is considered as invalid value.
-Thus, their hourly rate set to be NA.
-The result of the joining variables could be seen in the following tibble.
+The difference is that the hourly rate has only one version of each year.
+The hours of work and the hourly rate are then joined to calculate the number of jobs that a respondent has and their mean hourly wage.
+Some observations have 0 in their hourly rate and work hours, which is considered an invalid value.
+Thus, these observations are set to be NA.
+Besides, there are some observations with unusual hours of work.
+Here, we also censored the observations with > 84 hours of work a week to be NA.
 
@@
 
-Since our ultimate goal is to calculate the mean hourly wage, the number of jobs is calculate based on the availability of the `rate_per_hour` information.
+Since our ultimate goal is to calculate the mean hourly wage, the number of jobs is calculated based on the availability of the `rate_per_hour` information.
 For example, the number of jobs of ID 1, based on `hours_work`, is 2.
-However, since the information of hourly rate of `job_02` is not available, the number of job is considered as 1.
+However, since the information of hourly rate of `job_02` is not available, the number of jobs is considered as 1.
 
 Further, we calculated the mean hourly wage for each ID in each year using a weighted mean with the hours of work as the weight.
-However, there are a lot of missing value in `hours_work` variable.
-In that case, we only calculated the mean hourly wage based on arithmetic/regular mean method.
-Hence, we created a new variable to flag whether the mean hourly wage is a weighted or a regular mean.
-Additionally, if an ID only had one job, we directly used their hourly wages information and flagged it as an arithmetic mean.
-The following tibble shows the result of this tidying process.
+However, there is a lot of missing value in `hours_work` variable.
+In that case, we only calculated the mean hourly wage based on the arithmetic/regular mean method.
+Hence, we created a new variable to flag whether the mean hourly wage is weighted or regular.
+Additionally, if an ID only had one job, we directly used their hourly wages information and flagged it as arithmetic mean.
 
@@
 
-The `mean_hourly_wage` and `full_demographic` data are then joined.
-We also filtered the data to only have the cohort who completed the education up to 12th grade and participated at least five rounds in the survey and save it to an object called `wages_demog_hs`.
+The employment and demographic variables are then joined.
+We also filtered the data to only have the cohort who completed the education up to 12th grade and participated at least five rounds in the survey and save it to an object called `wages`.
 
@@
 According to @HuebnerMariannePhD2016Asat, Initial Data Analysis (IDA) is the step of inspecting and screening the data after being collected to ensure that the data is clean, valid, and ready to be deployed in the later formal statistical analysis.
-Moreover, @Chatfield1985TIEo argued that the two main objectives of IDA is data description, which is to assess the structure and the quality of the data; and model formulation without any formal statistical inference.
+Moreover, @Chatfield1985TIEo argues that the two main objectives of IDA are data description, which is to assess the structure and the quality of the data, and model formulation without any formal statistical inference.
 
-In this paper, we conducted an IDA or a preliminary data analysis to assess the consistency of the data with the cohort information that is provided by the NLSY.
-In addition, we also aimed to find the anomaly in the wages values using this approach.
-We mainly used graphical summary to do the IDA using `ggplot2`[@ggplot2] and `brolgar` [@brolgar].
+In this paper, we conduct an IDA or a preliminary data analysis to assess the consistency of the data with the cohort information that the NLSY provides.
+In addition, we also aim to find the anomaly in the wages values using this approach.
+We mainly use graphical summaries to do the IDA using `ggplot2`[@ggplot2] and `brolgar` [@brolgar].
 
 As stated previously, the respondents' ages ranged from 12 to 22 when first interviewed in 1979.
-Hence, we would like to validate whether all of the respondents were in this range.
-Additionally, the [NLSY](https://www.nlsinfo.org/content/cohorts/nlsy79/intro-to-the-sample/nlsy79-sample-introduction) also provided the number of the survey cohort by their gender (6,403 males and 6,283 females) and race (7,510 Non-Black/Non-Hispanic; 3,174 Black; 2,002 Hispanic).
-To validate this, we used the `full_demographic` i.e.
-the data with the survey years 1979 sample.
+Hence, we validate whether all of the respondents were in this range.
+Additionally, the [NLSY](https://www.nlsinfo.org/content/cohorts/nlsy79/intro-to-the-sample/nlsy79-sample-introduction) also provides the number of the survey cohort by their gender (6,403 males and 6,283 females) and race (7,510 Non-Black/Non-Hispanic; 3,174 Black; 2,002 Hispanic).
+To validate this, we used the `demog_nlsy79`, i.e., the data with the survey years 1979 sample.
 Table \ref{tab:age-table} and Table \ref{tab:gender-race-table} suggest that the demographic data we had is consistent with the sample information in the database.
@@
 
-The next step is that we explored the mean hourly wage data, in this case, we only explored the wages data in `wages_demog_hs`.
-Table \ref{tab:summarytable} shows that the overall wages median of the cohort is only 7.2, while the mean is 11.87.
+The next step is that explore the mean hourly wage data.
+In this case, we only explore the wages data of up to $12^{th}$ grade.
+Table \ref{tab:summarytable} shows that the median of the overall wage of the cohort is only 7.2, while the mean is 11.87.
 It indicates that the data might contains a lot of extreme values.
@@
 However, we can not be sure only by observing the summary statistics.
-Hence, we used visualisation techniques to investigate this matter.
+Hence, we use visualisation techniques to investigate this matter.
 Figure \ref{fig:feature-plot} conveys a problem in the mean hourly wage values.
-Figure \ref{fig:feature-plot} A shows that some observations had an exceptionally high figure of wages, even more than US$10,000 per hour.
+Figure \ref{fig:feature-plot} A shows that some observations have an exceptionally high figure of wages, even more than US$10,000 per hour.
 In Figure \ref{fig:feature-plot} B, we barely see any difference in the minimum, median, and maximum value of the wages since the distribution is heavily skewed to the right.
@@
 
-In Figure \ref{fig:high-wages}, we plotted some respondents with a high value of mean hourly wages.
-We filtered all of the IDs who earned more than US$ 500 per hour on average.
-We found that these respondents only experienced one point of extremely high wages.
-Thus, we suspected that these high values are erroneous resulted from a data entry error.
+In Figure \ref{fig:high-wages}, we plot some respondents with a high value of mean hourly wages.
+We filter all of the IDs who earned more than US$ 500 per hour on average.
+We find that these respondents only experienced one point of extremely high wages.
+Thus, we suspect that these high values are erroneous resulted from a data entry error.
 
@@
 
-Further, we took 36 samples randomly from the data and plotted them, as shown in Figure \ref{fig:sample-plot}.
-It implies that not only that some observations earned extremely high figures of wages, but some also had reasonably fluctuated wages, for example, the IDs in panel numbers 5, 7, and 11.
-The plot also implies that the samples had a different pattern of mean hourly wages.
-Some had flat wages for years but had a sudden increase in one particular year, then it went down again, while the others experienced an upsurge in their wage, for instance, the IDs in panel 9.
+Further, we take 36 samples randomly from the data and plot them, as shown in Figure \ref{fig:sample-plot}.
+It implies that not only that some observations earn extremely high figures of wages, but some also have reasonably fluctuated wages, for example, the IDs in panel numbers 5, 7, and 11.
+The plot also implies that the samples have a different pattern of mean hourly wages.
+Some have had flat wages for years but had a sudden increase in one particular year, then it gone down again, while the others experienced an upsurge in their wage, for instance, the IDs in panel 9.
 
@@
 
-According to @MichaelRPergamit2001DWTN, one of the flaws of the NLSY79 employment data is that since the NLSY79 collect the information of the working hours since the last interview, it might be challenging for the respondents to track the within-job hours' changes that happens between survey year, especially for the respondents with fluctuate working hours or whose job is seasonal.
-It even has been more challenging since 1994, where the respondents had to recall two years period.
+The anomalies are also found in the total hours of work, where some observations work for 420 hours a week in total.
+According to @MichaelRPergamit2001DWTN, one of the flaws of the NLSY79 employment data is that the NLSY79 collects the information of the working hours since the last interview.
+Thus, it might be challenging for the respondents to track the within-job hours' changes between survey years, especially for the respondents with fluctuated working hours or seasonal jobs.
+It even has been more challenging since 1994, where the respondents had to recall two years periods.
 This shortcoming might also contribute to the fluctuation of one's wages data.
@@
 
-As part of the IDA, which is the model formulation, we built a robust linear regression model to handle the extreme values in the data.
-Robust linear regression yields an estimation that is robust to the influence of noise or contamination [@KollerManuel2016rARP].
+As part of the IDA, which is the model formulation, we build a robust linear regression model to treat the extreme values in the data.
+Robust linear regression yields an estimation robust to the influence of noise or contamination [@KollerManuel2016rARP].
 It also aims to detect the contamination by weighting each observation based on how "well-behaved" they are, known as robustness weight.
@@
 
-Since we worked with longitudinal data, we built the model for each ID instead of the overall data.
-The robust mixed model is might be the best model to be employed in this case.
+Since we work with longitudinal data, we build the model for each ID instead of the overall data.
+The robust mixed model could be the best model to be employed in this case.
 However, this method is too computationally and memory expensive, especially for a large data set, like the NLSY79 data.
-Thus, the model for each ID is built utilizing the `nest` and `map` function from `tidyr` [@tidyr] and `purrr` [@purrr] respectively.
+Thus, the model for each ID is built utilizing the `nest` and `map` function from `tidyr` [@tidyr] and `purrr` [@purrr], respectively.
 
-We built the model using the `rlm` function from `MASS` package [@mass].
+We build the model using the `rlm` function from `MASS` package [@mass].
 We set the `mean_hourly_wage` and `year` as the dependent and predictor, respectively.
-Furthermore, we used M-Estimation with Huber weighting where the observation with a slight residual gets a weight of 1, while the larger the residual, the smaller the weight (less than 1) [@rlm].
-However, the challenging part of detecting the anomaly using the robustness weight is determining the threshold of the weight in which the observations are considered outliers.
+Furthermore, we use M-Estimation with Huber weighting, where the observation with a slight residual gets a weight of 1, while the larger the residual, the smaller the weight (less than 1) [@rlm].
+However, the challenging part of detecting the anomaly using the robustness weight is determining the weight threshold in which the observations are considered outliers.
 Moreover, it should be noted that not all the outliers are due to an error.
@@
 
-To minimize the risk of mistakenly regarding an outlier as an "erroneous outlier", we have simulated some thresholds and study how they affected the data.
-We found that 0.12 is the most reasonable value to be the threshold to minimize that drawback's risk because it still captures the sensible spikes in the data.
-In other words, we kept maintaining the natural variability of the wages while minimizing anomalies because of the error in the data recording.
-After deciding the threshold, we imputed the observations whose weight less than 0.12 with the models' predicted value.
-We then flagged those observations in a new variable called `is_pred`.
+To minimize the risk of mistakenly identifying an outlier as an "erroneous outlier", we simulate some thresholds and study how they affect the data.
+We find that 0.12 is the most reasonable value to be the threshold to minimize that drawback's risk because it still captures the sensible spikes in the data.
+In other words, we keep maintaining the natural variability of the wages while minimizing anomalies because of the error in the data recording.
+After deciding the threshold, we impute the observations whose weights are less than 0.12 with the models' predicted value.
+We then flag those observations in a new variable called `is_pred`.
 
@@
 Further, \ref{fig:comppict} A shows that after eliminating the extreme values, the highest value has decreased to be around $350.
-The spikes were still observed but are not as extreme as the original data set.
-In Figure \ref{fig:comppict} B), we plotted the three features of mean hourly wages, namely the minimum, median, and maximum value, transformed to log scale.
+The spikes are still observed but not as extreme as the original data set.
+In Figure \ref{fig:comppict} B), we plot the three features of mean hourly wages, namely the minimum, median, and maximum value, transformed to log scale.
 The plot implies that the skewness is slightly negative.
@@
 
-Finally, we saved the imputed data and set the appropriate data type for the variables.
+Finally, we save the imputed data and set the appropriate data type for the variables.
 As our target data is the mean hourly wage of the high-school dropouts, we then subset the high-school graduate data set to have only the male-high school dropout data.
-We also saved the NLSY79 cohort's demographic information in a separate dataset.
 We then make these three data sets and their processing documentation publicly available through an R data container package called `yowie`.
@@
 
-```{r flow-chart, echo=FALSE, fig.cap="The stages of data filtering from the raw data to get three datasets that are contained in an R package called yowie, n means the number of ID, while n_obs means the number of observations.", fig.height=5, fig.width=5}
+```{r flow-chart, fig.cap="The stages of data filtering from the raw data to get three datasets contained yowie, n means the number of ID, while n\\_obs means the number of observations."}
 ```
 
-# Exploratory Data Analysis
+# Exploratory Data Analysis {#eda}
 
@@
 
-```{r nobs-plot, fig.cap="Distribution of observations. The numbers of rounds varies among respondents. Most of respondents participated in 28 rounds of survey."}
+```{r nobs-plot}
 ```
@@
 
-```{r slope-nobs, fig.cap="Distribution of wages slope by number of observations. The slopes' range are more stable with more observations."}
+```{r slope-nobs}
+```
+
+```{r slope-nobs-plot, fig.cap = "A) Distribution of observations.
+The numbers of rounds varies among respondents.
+Most of respondents participated in 28 rounds of survey.
+B) Distribution of wages slope by number of observations.
+The slopes' range are more stable with more observations."}
 ```
@@
 According to @OECD, gender pay gap is the difference between median wage of males and females relative to median wage of males.
-We find that, as shown in \ref{fig:se-gender}, over the years, females tend to earn fewer than males.
+We find that, as shown in \ref{fig:gap-gender}, over the years, females tend to earn fewer than males.
 For example, in 2018 females earn about 78 cents for every $1 earned by males.
@@
 
-```{r wages-gender-gap, echo=FALSE, fig.cap="Gender wage gap or the median wages of females relatives to median wages of males.
+```{r gap-gender, fig.cap="Gender wage gap or the median wages of females relatives to median wages of males.
 Females tend to earn fewer than males.
-For example, in 2018 females earn about 78 cents for every $1 earned by males.
+For example, in 2018 females earn about 78 cents for every \\$1 earned by males.
 Note that this is a crude measure as gap is calculated not based on other variable, such as experience and job class."}
 ```
+
+a
+b
+c
+d
+e
+f
+gh
 
@@
 
-# Summary
+# Summary {#summary}
 

### 2021-09-08 paper/paper.tex
Removed striped styling in kable.
Check spelling and grammar.

@@
 
-\title{The Journey from Wild to Text Book Data: A Case Study from the National Longitudinal Survey of Youth}
+\title{The Journey from Wild to Textbook Data: A Case Study from the National Longitudinal Survey of Youth}
 
@@
 
-\hypertarget{introduction}{%
-\section{Introduction}\label{introduction}}
+\hypertarget{intro}{%
+\section{Introduction}\label{intro}}
 
-``Open data'' is data that are freely accessible, modifiable, and shareable by anyone for any purpose (Open Knowledge Foundation 2021).
-This type of data can be useful as example data in statistical text books and for research purposes.
-However, open data are often, what we might call ``wild data'', because it requires substantial cleaning and tidying to tame it into text book shape.
+``Open data'' is data that is freely accessible, modifiable, and shareable by anyone for any purpose (Open Knowledge Foundation 2021).
+This type of data can be useful as example data in statistical textbooks and for research purposes.
+However, open data is often referred to as what we might call ``wild data'' because it requires substantial cleaning and tidying to tame it into textbook shape.
 Huebner, Vach, and Cessie (2016) emphasize that making the data cleaning process accountable and transparent is imperative.
@@
 Data cleaning is part of what is called ``initial data analysis (IDA)'' (Chatfield 1985).
-The other IDA steps are to explore the data properties and check assumption of modeling to document and report the process for the later formal analysis.
+The other IDA steps are to explore the data properties and check the assumption of modeling to document and report the process for the later formal analysis.
 Hence, exploratory data analysis, which is defined by Tukey (1977) as a detective work to get the clue from data, either numerically or graphically, before confirmatory data analysis is performed, encompasses IDA.
@@
 There are few research papers that document the data cleaning (Wickham 2014).
-Furthermore, the decisions made in this stage often go unreported in the sense that IDA is often performed in an unplanned and unstructured way, and is only shared among restricted parties (Huebner et al.
+Furthermore, the decisions made in this stage often go unreported in the sense that IDA is often performed in an unplanned and unstructured way and is only shared among restricted parties (Huebner et al.
 2020).
@@
 This paper aims to demonstrate the steps of IDA, tidying, cleaning, and documenting the process, for a prominent open data source, National Longitudinal Survey of Youth (NLSY) 1979, henceforth referred to NSLY79.
-This data has been playing an important role in research in various disciplines including but are not limited to economics, sociology, education, public policy, and public health for more than a quarter of the century (Pergamit et al.
+This data has been playing an important role in research in various disciplines, including but are not limited to economics, sociology, education, public policy, and public health for more than a quarter of the century (Pergamit et al.
 2001).
-In addition, The National Longitudinal Survey is considered as survey with high retention rates and carefully designed making it suitable for a life course research (Pergamit et al.
+In addition, The National Longitudinal Survey is considered a survey with high retention rates and carefully designed, making it suitable for a life course research (Pergamit et al.
 2001) and (Cooksey 2017).
 According to Cooksey (2017), thousand of articles, and hundreds of book chapters and monographs has incorporated the NLSY data.
-Moreover, the NLSY79 is considered as the most widely used and most important cohort in the NLSY79 data sets (Pergamit et al.
+Moreover, the NLSY79 is considered the most widely used and most important cohort in the NLSY79 data sets (Pergamit et al.
 2001).
@@
 Singer and Willett (2003) used the wages and other variables of high school dropouts from the NLSY79 data as an example data set to illustrate longitudinal data modeling.
-Our aim is to refresh this text book data to be more up to date.
-Here, we investigate the process of getting from the raw NLSY79 to this text book data set.
-However, we are not able to create the exact same data set as published in their book since we do not have the information of what age threshold did they use to determine the high school dropouts.
+Our aim is to refresh this textbook data to be more up to date.
+Here, we investigate the process of getting from the raw NLSY79 to this textbook data set.
+However, we cannot create the exact same data set as published in their book since we do not have the information of what age threshold they used to determine the high school dropouts.
 
 This paper is structured to have 5 sections.
-Section 2 describes the original data source.
-Section 3 presents the steps of cleaning the data, including getting and tidying the data from the NLSY79 and initial data analysis to find and treat the anomalies in the data.
-We also gave examples of exploratory data analysis using the clean data in Section 4.
-Finally, Section 6 summarizes the paper.
-
-\hypertarget{the-nlsy79}{%
-\section{The NLSY79}\label{the-nlsy79}}
+Section \ref{database} describes the original data source.
+Section \ref{cleaning} presents the steps of cleaning the data, including getting and tidying the data from the NLSY79 and initial data analysis to find and treat the anomalies in the data.
+We also gave examples of exploratory data analysis using the clean data in Section \ref{eda}.
+Finally, Section \ref{summary} summarises the paper.
 
 \hypertarget{database}{%
-\subsection{Database}\label{database}}
+\section{The NLSY79}\label{database}}
+
+\hypertarget{database-1}{%
+\subsection{Database}\label{database-1}}
 
@@
 
-Although the main focus area of the NLSY is labor and employment, the NLSY also cover several other topics including education; training and achievement; household, geography and contextual variables; dating, marriage, cohabitation; sexual activity, pregnancy, and fertility; children; income, assets and program participation; health; attitudes and expectations; and crime and substance use.
+Although the main focus area of the NLSY is labor and employment, the NLSY also covers several other topics, including education; training and achievement; household, geography, and contextual variables; dating, marriage, cohabitation; sexual activity, pregnancy, and fertility; children; income, assets and program participation; health; attitudes and expectations; and crime and substance use.
 
-There are two ways to conduct the interview of the NLSY79, which are face to face interview or by telephone.
+There are two ways to conduct the interview of the NLSY79, which are face-to-face interviews or by telephone.
 In recent survey years, more than 90 percent of respondents were interviewed by telephone (Cooksey 2017).
@@
 
-As mentioned in Section 1, this paper aims to refresh the NLSY79 data as used in Singer and Willet's text book (Singer and Willett (2003)).
+As mentioned in Section \ref{intro}, this paper aims to refresh the NLSY79 data as used in Singer and Willet's textbook (Singer and Willett (2003)).
 This data set contains the information of yearly mean hourly wages from the NLSY79 cohort, with education and race as covariates.
 This data set represents the measurement of male high-school dropouts, aged from 14 to 17 years old when first time measured.
-Hence, in this paper, after getting the data set of high school graduates, we subset it by only having the high school dropouts.
+Accordingly, to get this data, we have two additional target data, NSLY79 cohort's demographic profiles and wages data of people who completed education up to high school.
 
-Since there is no information of high school drop-outs in Singer and Willet's book, we define high school dropouts as respondents who only completed either \(9^{th}\), \(10^{th}\), \(11^{th}\) grade, or who completed \(12^{th}\) when their age are at least 19 years old (means that these IDs being dropped out in certain year, but they returned back to high school to complete it).
+Since there is no information on high school dropouts in Singer and Willet's book, we define high school dropouts as respondents who only completed either \(9^{th}\), \(10^{th}\), \(11^{th}\) grade, or who completed \(12^{th}\) when their age is at least 19 years old (means that these IDs being dropped out in a certain year, but they returned back to high school to complete it).
 
-\hypertarget{the-nlsy79-data-cleaning}{%
-\section{The NLSY79 Data Cleaning}\label{the-nlsy79-data-cleaning}}
+\hypertarget{cleaning}{%
+\section{The NLSY79 Data Cleaning}\label{cleaning}}
 
@@
 
-\begin{itemize}
-\tightlist
-\item
-  Education, Training \& Achievement Scores
+\textbf{Education, Training \& Achievement Scores}
 
@@
   \item
-      Downloaded all of the 78 variables in Highest grade completed.
+    Downloaded all of variables in Highest grade completed.
   \end{itemize}
 \end{itemize}
-\item
-  Employment
+
+\textbf{Employment}
 
@@
   \item
-      Downloaded all of the 427 variables in Hours worked
+    Downloaded all of the variables in Hours worked.
   \item
-      Downloaded all of the 151 variables in Hourly wages
+    Downloaded all of the variables in Hourly wages
+    Both hours worked and hourly wages are recorded by the job, up to five jobs for each id/subject.
+  \end{itemize}
 \end{itemize}
 
-    Both hours worked and hourly wages are recorded by the job, up to five jobs for each id/subject.
-  \end{itemize}
-\item
-  Household, Geography \& Contextual Variables
+\textbf{Household, Geography \& Contextual Variables}
 
@@
     Downloaded year and month of birth, race, and sex variables.
-    \end{itemize}
-
-    There are two versions of the year and month of birth, i.e.~1979 and 1981 data. We downloaded these two versions.
+    There are two versions of the year and month of birth, i.e., 1979 and 1981 data. We downloaded these two versions.
   \end{itemize}
@@
 \item
-  NLSY79.NLSY79: This is the tagset of variables that can be uploaded to the web site to recreate the data set.
+  NLSY79.NLSY79: Tagset of variables that can be uploaded to the website to recreate the data set.
 \item
-  NLSY79.R: This is an R script provided automatically by the database for reading the data into R and convert the variables' name and its label into something more sensible.
+  NLSY79.R: R script provided automatically by the database for reading the data into R and converting the variables' names and label into something more sensible.
 We utilized this code to do an initial data tidying.
@@
 
-According to Wickham (2014), a tidy data sets comply with three rules, the first is that each variable forms a column, the second is that each observation forms a row, and the last is that each type of observational unit forms a table.
-Unfortunately, the \texttt{new\_data\_qnames} did not meet these requirements in the way that the value for particular year and job are stored in different columns, hence the data contains a huge number of columns (686 columns).
-The example of the untidy of the data set is displayed in Table \ref{tab:untidy-data}, which actually intended to display the hourly rate of each respondent by job (HRP1 to HRP5) and by year (1979 and 1980).
+According to Wickham (2014), tidy data sets comply with three rules.
+Firstly, each variable forms a column.
+Secondly, each observation forms a row.
+Lastly, each type of observational unit forms a table.
+Unfortunately, the \texttt{new\_data\_qnames} did not meet these requirements in the way that data for each year and job for the same respondent is stored as one variable.
+Hence, the data contains a huge number of columns (686 columns).
+The example of the untidy data set is displayed in Table \ref{tab:untidy-data}.
+The table intended to display the hourly rate of each respondent by job (HRP1 to HRP5) and by year (1979 and 1980).
 The table implies that the column headers are values of the year and job, not variable names.
+Consequently, the data should be tidied and wrangled first to extract the demographic and employment variables we want to put in the final data set.
+We mainly used \texttt{tidyr} (H.
+Wickham 2020b), \texttt{dplyr} (H.
+Wickham, François, et al.
+2020), and \texttt{stringr} (H.
+Wickham 2019) to do this job.
 
@@
 \hline
-\rowcolor{gray!6}  1 & 328 & NA & NA & NA & NA & NA\\
+1 & 328 & NA & NA & NA & NA & NA\\
 \hline
@@
 \hline
-\rowcolor{gray!6}  3 & 365 & NA & 275 & NA & NA & 397\\
+3 & 365 & NA & 275 & NA & NA & 397\\
 \hline
@@
 \hline
-\rowcolor{gray!6}  5 & 310 & 375 & NA & NA & NA & 333\\
+5 & 310 & 375 & NA & NA & NA & 333\\
 \hline
@@
 
-Consequently, the data should be tidied and wrangled first to extract the demographic and employment variables that we want to put in the final data set.
-We mainly used \texttt{tidyr} (H.
-Wickham 2020b), \texttt{dplyr} (H.
-Wickham, François, et al.
-2020), and \texttt{stringr} (H.
-Wickham 2019) to do this job.
-
 \hypertarget{tidying-demographic-variables}{%
@@
 
-\texttt{Age\ in\ 1979}, \texttt{gender}, \texttt{race}, \texttt{highest\ grade\ completed} (factor and integer), and the \texttt{year\ when\ the\ highest\ grade\ completed} are the variables that we want to use in the cleaned data set.
+\texttt{Age\ in\ 1979}, \texttt{gender}, \texttt{race}, \texttt{highest\ grade\ completed}, and the \texttt{year\ when\ the\ highest\ grade\ completed} are the variables that we want to use in the cleaned data set.
 
-Age in 1979 are derived from year of birth and month of birth variables in the raw data set.
+Age in 1979 are derived from year and month of birth variables in the raw data.
 The variables have two versions, which are the 1979 version and the 1981 version.
@@
 
-For \texttt{gender} and \texttt{race}, we only renamed these variables.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\CommentTok{# tidy the gender and race variables}
-\NormalTok{demog_tidy <-}\StringTok{ }\NormalTok{categories_qnames }\OperatorTok{%>%}
-\StringTok{  }\NormalTok{dplyr}\OperatorTok{::}\KeywordTok{select}\NormalTok{(CASEID_}\DecValTok{1979}\NormalTok{,}
-\NormalTok{                SAMPLE_RACE_78SCRN,}
-\NormalTok{                SAMPLE_SEX_}\DecValTok{1979}\NormalTok{) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{rename}\NormalTok{(}\DataTypeTok{gender =}\NormalTok{ SAMPLE_SEX_}\DecValTok{1979}\NormalTok{,}
-         \DataTypeTok{race =}\NormalTok{ SAMPLE_RACE_78SCRN)}
-\end{Highlighting}
-\end{Shaded}
+For \texttt{gender} and \texttt{race}, we only rename the column names.
+It is worth noting that using these two variable needs special attention.
+Gender, as reported in the data, only has two categories, which is recognised today as inadequate.
+Gender is not binary.
+Further, race is as reported in the database.
+When doing analysis with this variable, one should keep in mind that the purpose is to study racism rather than race.
 
@@
 Hence, we only used the highest grade completed ever and derived the year when it is completed.
-The code for tidying this variable is also available in the supplementary code of this paper.
+The code for tidying this variable is also available in the supplementary materials.
 
-Finally, we join all of the tidy variables in a data set called \texttt{full\_demographics} as displayed as follows:
-
-\begin{verbatim}
-#>   id dob_month dob_year dob_conflict                    race gender yr_hgc
-#> 1  1         9       58        FALSE NON-BLACK, NON-HISPANIC FEMALE   1979
-#> 2  2         1       59        FALSE NON-BLACK, NON-HISPANIC FEMALE   1985
-#> 3  3         8       61        FALSE NON-BLACK, NON-HISPANIC FEMALE   1993
-#> 4  4         8       62        FALSE NON-BLACK, NON-HISPANIC FEMALE   1986
-#> 5  5         7       59        FALSE NON-BLACK, NON-HISPANIC   MALE   1984
-#> 6  6        10       60        FALSE NON-BLACK, NON-HISPANIC   MALE   1983
-#>   hgc_i          hgc
-#> 1    12   12TH GRADE
-#> 2    12   12TH GRADE
-#> 3    12   12TH GRADE
-#> 4    14 2ND YEAR COL
-#> 5    18 6TH YEAR COL
-#> 6    16 4TH YEAR COL
-\end{verbatim}
+Finally, we get all of the demographic profiles of the NLSY79 cohort. We then save this data as \texttt{demog\_nlsy79}.
 
@@
 
-The employment data comprises of three variables, i.e.~\texttt{total\ hours\ of\ work\ per\ week}, \texttt{number\ of\ jobs\ that\ an\ individual\ has}, and \texttt{mean\ hourly\ wage}.
+The employment data comprises of three variables, i.e., \texttt{total\ hours\ of\ work\ per\ week}, \texttt{number\ of\ jobs\ that\ an\ individual\ has}, and \texttt{mean\ hourly\ wage}.
 
 \texttt{hours\ worked\ per\ week} initially has only one version per job, no choice from 1979 to 1987 (QES-52A).
-From 1988 onward, when we had more options, we chose the variable for total hours including time spent working from home (QES-52D).
+From 1988 onward, when we had more options, we chose the variable for total hours, including time spent working from home (QES-52D).
 However, in 1993, this variable did not have all the five D variables (the first one and the last one were missing), so we used QES-52A variable instead.
-In addition, 2008 only had jobs 1-4 for the QES-52D variable (whereas the other years had 1-5), so we just used these.
+In addition, 2008 only had jobs 1-4 for the QES-52D variable (whereas the other years had 1-5), so we use these.
 
@@
 The same way was also deployed to tidy the rate of wage by year and by ID.
-The difference is that the hourly rate had only one version of each year.
-The hours of work and the hourly rate were then joined to calculate the number of jobs that a respondent has and their mean hourly wage.
-Some observations had 0 in their hourly rate, which is considered as invalid value.
-Thus, their hourly rate set to be NA.
-The result of the joining variables could be seen in the following tibble.
+The difference is that the hourly rate has only one version of each year.
+The hours of work and the hourly rate are then joined to calculate the number of jobs that a respondent has and their mean hourly wage.
+Some observations have 0 in their hourly rate and work hours, which is considered an invalid value.
+Thus, these observations are set to be NA.
+Besides, there are some observations with unusual hours of work.
+Here, we also censored the observations with \textgreater{} 84 hours of work a week to be NA.
 
@@
          \DataTypeTok{hours_work =} \KeywordTok{ifelse}\NormalTok{(hours_work }\OperatorTok{==}\StringTok{ }\DecValTok{0} \OperatorTok{|}\StringTok{ }\NormalTok{hours_work }\OperatorTok{>}\StringTok{ }\DecValTok{84}\NormalTok{, }\OtherTok{NA}\NormalTok{, hours_work))}
-\KeywordTok{head}\NormalTok{(hours_wages)}
 \end{Highlighting}
@@
 
-\begin{verbatim}
-#> # A tibble: 6 x 5
-#>      id job    year  rate_per_hour hours_work
-#>   <int> <chr>  <chr>         <int>      <int>
-#> 1     1 job_01 1979            328         38
-#> 2     1 job_02 1979             NA         15
-#> 3     1 job_03 1979             NA         NA
-#> 4     1 job_04 1979             NA         NA
-#> 5     1 job_05 1979             NA         NA
-#> 6     2 job_01 1979            385         35
-\end{verbatim}
-
-Since our ultimate goal is to calculate the mean hourly wage, the number of jobs is calculate based on the availability of the \texttt{rate\_per\_hour} information.
+Since our ultimate goal is to calculate the mean hourly wage, the number of jobs is calculated based on the availability of the \texttt{rate\_per\_hour} information.
 For example, the number of jobs of ID 1, based on \texttt{hours\_work}, is 2.
-However, since the information of hourly rate of \texttt{job\_02} is not available, the number of job is considered as 1.
+However, since the information of hourly rate of \texttt{job\_02} is not available, the number of jobs is considered as 1.
 
 Further, we calculated the mean hourly wage for each ID in each year using a weighted mean with the hours of work as the weight.
-However, there are a lot of missing value in \texttt{hours\_work} variable.
-In that case, we only calculated the mean hourly wage based on arithmetic/regular mean method.
-Hence, we created a new variable to flag whether the mean hourly wage is a weighted or a regular mean.
-Additionally, if an ID only had one job, we directly used their hourly wages information and flagged it as an arithmetic mean.
-The following tibble shows the result of this tidying process.
+However, there is a lot of missing value in \texttt{hours\_work} variable.
+In that case, we only calculated the mean hourly wage based on the arithmetic/regular mean method.
+Hence, we created a new variable to flag whether the mean hourly wage is weighted or regular.
+Additionally, if an ID only had one job, we directly used their hourly wages information and flagged it as arithmetic mean.
 
@@
 
-The \texttt{mean\_hourly\_wage} and \texttt{full\_demographic} data are then joined.
-We also filtered the data to only have the cohort who completed the education up to 12th grade and participated at least five rounds in the survey and save it to an object called \texttt{wages\_demog\_hs}.
+The employment and demographic variables are then joined.
+We also filtered the data to only have the cohort who completed the education up to 12th grade and participated at least five rounds in the survey and save it to an object called \texttt{wages}.
 
@@
 According to Huebner, Vach, and Cessie (2016), Initial Data Analysis (IDA) is the step of inspecting and screening the data after being collected to ensure that the data is clean, valid, and ready to be deployed in the later formal statistical analysis.
-Moreover, Chatfield (1985) argued that the two main objectives of IDA is data description, which is to assess the structure and the quality of the data; and model formulation without any formal statistical inference.
+Moreover, Chatfield (1985) argues that the two main objectives of IDA are data description, which is to assess the structure and the quality of the data, and model formulation without any formal statistical inference.
 
-In this paper, we conducted an IDA or a preliminary data analysis to assess the consistency of the data with the cohort information that is provided by the NLSY.
-In addition, we also aimed to find the anomaly in the wages values using this approach.
-We mainly used graphical summary to do the IDA using \texttt{ggplot2}(Wickham 2016) and \texttt{brolgar} (Tierney, Cook, and Prvan 2020).
+In this paper, we conduct an IDA or a preliminary data analysis to assess the consistency of the data with the cohort information that the NLSY provides.
+In addition, we also aim to find the anomaly in the wages values using this approach.
+We mainly use graphical summaries to do the IDA using \texttt{ggplot2}(Wickham 2016) and \texttt{brolgar} (Tierney, Cook, and Prvan 2020).
 
 As stated previously, the respondents' ages ranged from 12 to 22 when first interviewed in 1979.
-Hence, we would like to validate whether all of the respondents were in this range.
-Additionally, the \href{https://www.nlsinfo.org/content/cohorts/nlsy79/intro-to-the-sample/nlsy79-sample-introduction}{NLSY} also provided the number of the survey cohort by their gender (6,403 males and 6,283 females) and race (7,510 Non-Black/Non-Hispanic; 3,174 Black; 2,002 Hispanic).
-To validate this, we used the \texttt{full\_demographic} i.e.~the data with the survey years 1979 sample.
+Hence, we validate whether all of the respondents were in this range.
+Additionally, the \href{https://www.nlsinfo.org/content/cohorts/nlsy79/intro-to-the-sample/nlsy79-sample-introduction}{NLSY} also provides the number of the survey cohort by their gender (6,403 males and 6,283 females) and race (7,510 Non-Black/Non-Hispanic; 3,174 Black; 2,002 Hispanic).
+To validate this, we used the \texttt{demog\_nlsy79}, i.e., the data with the survey years 1979 sample.
 Table \ref{tab:age-table} and Table \ref{tab:gender-race-table} suggest that the demographic data we had is consistent with the sample information in the database.
@@
 \hline
-\rowcolor{gray!6}  15 & 1265\\
+15 & 1265\\
 \hline
@@
 \hline
-\rowcolor{gray!6}  17 & 1600\\
+17 & 1600\\
 \hline
@@
 \hline
-\rowcolor{gray!6}  19 & 1662\\
+19 & 1662\\
 \hline
@@
 \hline
-\rowcolor{gray!6}  21 & 1677\\
+21 & 1677\\
 \hline
@@
 \hline
-\rowcolor{gray!6}  Male & 1000 (15.62\%) & 1613 (25.19\%) & 3790 (59.19\%) & 6403 (100.00\%)\\
+Male & 1000 (15.62\%) & 1613 (25.19\%) & 3790 (59.19\%) & 6403 (100.00\%)\\
 \hline
@@
 \hline
-\rowcolor{gray!6}  Total & 2002 (15.78\%) & 3174 (25.02\%) & 7510 (59.20\%) & 12686 (100.00\%)\\
+Total & 2002 (15.78\%) & 3174 (25.02\%) & 7510 (59.20\%) & 12686 (100.00\%)\\
 \hline
@@
 
-The next step is that we explored the mean hourly wage data, in this case, we only explored the wages data in \texttt{wages\_demog\_hs}.
-Table \ref{tab:summarytable} shows that the overall wages median of the cohort is only 7.2, while the mean is 11.87.
+The next step is that explore the mean hourly wage data.
+In this case, we only explore the wages data of up to \(12^{th}\) grade.
+Table \ref{tab:summarytable} shows that the median of the overall wage of the cohort is only 7.2, while the mean is 11.87.
 It indicates that the data might contains a lot of extreme values.
@@
 \hline
-\rowcolor{gray!6}  Min. & 0.01000\\
+Min. & 0.01000\\
 \hline
@@
 \hline
-\rowcolor{gray!6}  Median & 7.20000\\
+Median & 7.20000\\
 \hline
@@
 \hline
-\rowcolor{gray!6}  3rd Qu. & 11.74000\\
+3rd Qu. & 11.74000\\
 \hline
@@
 However, we can not be sure only by observing the summary statistics.
-Hence, we used visualisation techniques to investigate this matter.
+Hence, we use visualisation techniques to investigate this matter.
 Figure \ref{fig:feature-plot} conveys a problem in the mean hourly wage values.
-Figure \ref{fig:feature-plot} A shows that some observations had an exceptionally high figure of wages, even more than US\$10,000 per hour.
+Figure \ref{fig:feature-plot} A shows that some observations have an exceptionally high figure of wages, even more than US\$10,000 per hour.
 In Figure \ref{fig:feature-plot} B, we barely see any difference in the minimum, median, and maximum value of the wages since the distribution is heavily skewed to the right.
@@
 
-{\centering \includegraphics[width=\textwidth]{figures/feature-plot-1}
+{\centering \includegraphics[width=0.7\linewidth]{figures/feature-plot-1}
 
@@
 
-In Figure \ref{fig:high-wages}, we plotted some respondents with a high value of mean hourly wages.
-We filtered all of the IDs who earned more than US\$ 500 per hour on average.
-We found that these respondents only experienced one point of extremely high wages.
-Thus, we suspected that these high values are erroneous resulted from a data entry error.
+In Figure \ref{fig:high-wages}, we plot some respondents with a high value of mean hourly wages.
+We filter all of the IDs who earned more than US\$ 500 per hour on average.
+We find that these respondents only experienced one point of extremely high wages.
+Thus, we suspect that these high values are erroneous resulted from a data entry error.
 
@@
 
-{\centering \includegraphics[width=\textwidth]{figures/high-wages-1}
+{\centering \includegraphics[width=0.7\linewidth]{figures/high-wages-1}
 
@@
 
-{\centering \includegraphics[width=\textwidth]{figures/sample-plot-1}
+{\centering \includegraphics[width=0.7\linewidth]{figures/sample-plot-1}
 
@@
 
+The anomalies are also found in the total hours of work, where some observations work for 420 hours a week in total.
 According to Pergamit et al.
-(2001), one of the flaws of the NLSY79 employment data is that since the NLSY79 collect the information of the working hours since the last interview, it might be challenging for the respondents to track the within-job hours' changes that happens between survey year, especially for the respondents with fluctuate working hours or whose job is seasonal.
-It even has been more challenging since 1994, where the respondents had to recall two years period.
+(2001), one of the flaws of the NLSY79 employment data is that the NLSY79 collects the information of the working hours since the last interview.
+Thus, it might be challenging for the respondents to track the within-job hours' changes between survey years, especially for the respondents with fluctuated working hours or seasonal jobs.
+It even has been more challenging since 1994, where the respondents had to recall two years periods.
 This shortcoming might also contribute to the fluctuation of one's wages data.
@@
 
-As part of the IDA, which is the model formulation, we built a robust linear regression model to handle the extreme values in the data.
-Robust linear regression yields an estimation that is robust to the influence of noise or contamination (Koller 2016).
+As part of the IDA, which is the model formulation, we build a robust linear regression model to treat the extreme values in the data.
+Robust linear regression yields an estimation robust to the influence of noise or contamination (Koller 2016).
 It also aims to detect the contamination by weighting each observation based on how ``well-behaved'' they are, known as robustness weight.
@@
 
-Since we worked with longitudinal data, we built the model for each ID instead of the overall data.
-The robust mixed model is might be the best model to be employed in this case.
+Since we work with longitudinal data, we build the model for each ID instead of the overall data.
+The robust mixed model could be the best model to be employed in this case.
 However, this method is too computationally and memory expensive, especially for a large data set, like the NLSY79 data.
 Thus, the model for each ID is built utilizing the \texttt{nest} and \texttt{map} function from \texttt{tidyr} (H.
-Wickham 2020b) and \texttt{purrr} (Henry and Wickham 2020) respectively.
+Wickham 2020b) and \texttt{purrr} (Henry and Wickham 2020), respectively.
 
-We built the model using the \texttt{rlm} function from \texttt{MASS} package (Venables and Ripley 2002).
+We build the model using the \texttt{rlm} function from \texttt{MASS} package (Venables and Ripley 2002).
 We set the \texttt{mean\_hourly\_wage} and \texttt{year} as the dependent and predictor, respectively.
-Furthermore, we used M-Estimation with Huber weighting where the observation with a slight residual gets a weight of 1, while the larger the residual, the smaller the weight (less than 1) (UCLA: Statistical Consulting Group 2021).
-However, the challenging part of detecting the anomaly using the robustness weight is determining the threshold of the weight in which the observations are considered outliers.
+Furthermore, we use M-Estimation with Huber weighting, where the observation with a slight residual gets a weight of 1, while the larger the residual, the smaller the weight (less than 1) (UCLA: Statistical Consulting Group 2021).
+However, the challenging part of detecting the anomaly using the robustness weight is determining the weight threshold in which the observations are considered outliers.
 Moreover, it should be noted that not all the outliers are due to an error.
@@
 
-To minimize the risk of mistakenly regarding an outlier as an ``erroneous outlier'', we have simulated some thresholds and study how they affected the data.
-We found that 0.12 is the most reasonable value to be the threshold to minimize that drawback's risk because it still captures the sensible spikes in the data.
-In other words, we kept maintaining the natural variability of the wages while minimizing anomalies because of the error in the data recording.
-After deciding the threshold, we imputed the observations whose weight less than 0.12 with the models' predicted value.
-We then flagged those observations in a new variable called \texttt{is\_pred}.
+To minimize the risk of mistakenly identifying an outlier as an ``erroneous outlier'', we simulate some thresholds and study how they affect the data.
+We find that 0.12 is the most reasonable value to be the threshold to minimize that drawback's risk because it still captures the sensible spikes in the data.
+In other words, we keep maintaining the natural variability of the wages while minimizing anomalies because of the error in the data recording.
+After deciding the threshold, we impute the observations whose weights are less than 0.12 with the models' predicted value.
+We then flag those observations in a new variable called \texttt{is\_pred}.
 
@@
 
-{\centering \includegraphics[width=\textwidth]{figures/compare-plot-1}
+{\centering \includegraphics[width=0.7\linewidth]{figures/compare-plot-1}
 
@@
 Further, \ref{fig:comppict} A shows that after eliminating the extreme values, the highest value has decreased to be around \$350.
-The spikes were still observed but are not as extreme as the original data set.
-In Figure \ref{fig:comppict} B), we plotted the three features of mean hourly wages, namely the minimum, median, and maximum value, transformed to log scale.
+The spikes are still observed but not as extreme as the original data set.
+In Figure \ref{fig:comppict} B), we plot the three features of mean hourly wages, namely the minimum, median, and maximum value, transformed to log scale.
 The plot implies that the skewness is slightly negative.
@@
 
-{\centering \includegraphics[width=\textwidth]{figures/comppict-1}
+{\centering \includegraphics[width=0.7\linewidth]{figures/comppict-1}
 
@@
 
-Finally, we saved the imputed data and set the appropriate data type for the variables.
+Finally, we save the imputed data and set the appropriate data type for the variables.
 As our target data is the mean hourly wage of the high-school dropouts, we then subset the high-school graduate data set to have only the male-high school dropout data.
-We also saved the NLSY79 cohort's demographic information in a separate dataset.
 We then make these three data sets and their processing documentation publicly available through an R data container package called \texttt{yowie}.
@@
 
-\textbackslash begin\{figure\}
+\begin{figure}
 
-\{\centering \includegraphics[width=\textwidth]{figures/flow-chart-1}
+{\centering \includegraphics[width=0.7\linewidth]{figures/flow-chart-1}
 
-\hypertarget{exploratory-data-analysis}{%
-\section{Exploratory Data Analysis}\label{exploratory-data-analysis}}
+}
+
+\caption{The stages of data filtering from the raw data to get three datasets contained yowie, n means the number of ID, while n\_obs means the number of observations.}\label{fig:flow-chart}
+\end{figure}
+
+\hypertarget{eda}{%
+\section{Exploratory Data Analysis}\label{eda}}
 
@@
 
-{\centering \includegraphics[width=\textwidth]{figures/nobs-plot-1}
+{\centering \includegraphics[width=0.7\linewidth]{figures/slope-nobs-plot-1}
 
@@
 
-\caption{Distribution of observations. The numbers of rounds varies among respondents. Most of respondents participated in 28 rounds of survey.}\label{fig:nobs-plot}
+\caption{A) Distribution of observations.
+The numbers of rounds varies among respondents.
+Most of respondents participated in 28 rounds of survey.
+B) Distribution of wages slope by number of observations.
+The slopes' range are more stable with more observations.}\label{fig:slope-nobs-plot}
 \end{figure}
@@
 
-{\centering \includegraphics[width=\textwidth]{figures/slope-nobs-1}
-
-}
-
-\caption{Distribution of wages slope by number of observations. The slopes' range are more stable with more observations.}\label{fig:slope-nobs}
-\end{figure}
-
-\begin{figure}
-
-{\centering \includegraphics[width=\textwidth]{figures/slope-gender-1}
+{\centering \includegraphics[width=0.7\linewidth]{figures/slope-gender-1}
 
@@
 According to OECD (n.d.), gender pay gap is the difference between median wage of males and females relative to median wage of males.
-We find that, as shown in \ref{fig:se-gender}, over the years, females tend to earn fewer than males.
+We find that, as shown in \ref{fig:gap-gender}, over the years, females tend to earn fewer than males.
 For example, in 2018 females earn about 78 cents for every \$1 earned by males.
@@
 
-\textbackslash begin\{figure\}
+\begin{figure}
 
-\{\centering \includegraphics[width=\textwidth]{figures/wages-gender-gap-1}
+{\centering \includegraphics[width=0.7\linewidth]{figures/gap-gender-1}
 
-\}
+}
 
-\textbackslash caption\{Gender wage gap or the median wages of females relatives to median wages of males.
+\caption{Gender wage gap or the median wages of females relatives to median wages of males.
 Females tend to earn fewer than males.
 For example, in 2018 females earn about 78 cents for every \$1 earned by males.
-Note that this is a crude measure as gap is calculated not based on other variable, such as experience and job class.\}\label{fig:wages-gender-gap}
-\textbackslash end\{figure\}
+Note that this is a crude measure as gap is calculated not based on other variable, such as experience and job class.}\label{fig:gap-gender}
+\end{figure}
+
+a
+b
+c
+d
+e
+f
+gh
 
@@
 
-{\centering \includegraphics[width=\textwidth]{figures/slope-educ-1}
+{\centering \includegraphics[width=0.7\linewidth]{figures/slope-educ-1}
 

### 2021-09-08 paper/scripts/data-preprocessing.R

@@ -17,6 +17,7 @@ library(modelr)
 library(forcats)
 library(yowie)
 
+
 ## ---- raw-data
 source(here::here("data-raw/NLSY79/NLSY79.R"))
 
@@ -118,7 +119,6 @@ highest_year <- demog_education %>%
 full_demographics <- full_join(dob_tidy, demog_tidy, by = "CASEID_1979") %>%
   full_join(highest_year, by = "CASEID_1979") %>%
   rename("id" = "CASEID_1979")
-head(full_demographics)
 
 ## ---- tidy-hours
 # make a list for years where we used the "QES-52A"
@@ -178,7 +178,7 @@ hours_wages <- left_join(rates_all,
   mutate(rate_per_hour = ifelse(rate_per_hour == 0, NA,
                                 rate_per_hour),
          hours_work = ifelse(hours_work == 0 | hours_work > 84, NA, hours_work))
-head(hours_wages)
+
 
 ## ---- tidy-nojob
 # calculate number of jobs that a person has in one year

### 2021-09-08 paper/scripts/eda.R

@@ -2,10 +2,12 @@
 wages_obs <- yowie::wages %>%
   features(wage, n_obs) %>%
   arrange(n_obs)
-ggplot(wages_obs, aes(x = n_obs)) +
+
+wages_obs_p <- ggplot(wages_obs, aes(x = n_obs)) +
   geom_bar() +
   xlab("number of observations") +
-  theme_bw()
+  theme_bw() +
+  ggtitle("A)")
 
 ## ---- slope
 wages <- yowie::wages %>%
@@ -29,12 +31,16 @@ wages_slope <- key_slope(wages, ln_wages ~ year0) %>%
                                                       "12TH grade")))
 ## ---- slope-nobs
 wages_obs_slope <- left_join(wages_obs, wages_slope, by = "id")
-ggplot(wages_obs_slope, aes(x = n_obs, y = .slope_year0)) +
+wages_obs_slop_p <- ggplot(wages_obs_slope, aes(x = n_obs, y = .slope_year0)) +
   geom_point() +
   geom_jitter() +
   xlab("number of observations") +
   ylab("slope") +
-  theme_bw()
+  theme_bw() +
+  ggtitle("B)")
+
+## ---- slope-nobs-plot
+wages_obs_p + wages_obs_slop_p
 
 ## ---- trim-slope
 slope_trimmed <- wages_slope %>%
@@ -69,9 +75,10 @@ ggplot(contrast_slope, aes(x = year, y = ln_wages, color = gender)) +
   scale_color_manual(values = c("#04B4AE", "#BF00FF")) +
   ylab("ln(wages)") +
   theme_bw() +
-  theme(axis.text.x = element_text(angle = 25))
+  theme(axis.text.x = element_text(angle = 25),
+        legend.position = "bottom")
 
-## ---- wages-gender-gap
+## ---- gap-gender
 wages_gender <- as.tibble(wages) %>%
   group_by(gender, year) %>%
   summarise(mean = mean(wage),
@@ -104,7 +111,7 @@ ggplot(gender_wage_gap) +
                  y = relative),
              color = "#DF7401", size = 2) +
   geom_hline(yintercept = 1, colour = "#088A68", size = 1) +
-  scale_x_continuous(limits = c(1978.7,2018.3), expand = c(0, 0)) +
+  #scale_x_continuous(limits = c(1978.7,2018.3), expand = c(0, 0)) +
   ylab("females' median wages relative to males (US$)") +
[…]

### 2021-09-08 paper/scripts/ida.R

@@ -172,10 +172,10 @@ ggplot(wages_compare) +
                  colour = type),
              alpha = 0.5,
              size = 1) +
-  theme(axis.text.x = element_text(angle = 10, size = 5),
-        legend.position = "bottom") +
   facet_wrap(~id, scales = "free_y") +
-  theme_bw()
+  theme_bw() +
+  theme(axis.text.x = element_text(angle = 10, size = 5),
+        legend.position = "bottom")
 
 ## ---- comppict
 spag2 <- wages_demog_hs %>%
@@ -290,7 +290,7 @@ n_obs_do_pred = filter(wages_hs_do, is_pred == TRUE) %>%
   nrow()
 
 ### ---- flow-chart
-fc <- grViz("digraph flowchart {
+grViz("digraph flowchart {
       node [fontname = Helvetica, shape = rectangle]
       tab1 [label = '@@1']
       tab2 [label = '@@2']
@@ -316,5 +316,4 @@ fc <- grViz("digraph flowchart {
       [6]: paste0('High school dropouts cohort, \\n n = ', n_do, ', n_obs = ', n_obs_do, ' \\n (', n_obs_do_pred, ' observations are predicted value)')
       [7]: 'yowie Package'
       ",
-            height = 200)
-fc
+            height = 700)

### 2021-09-09 paper/paper.Rmd
Add explanation to EDA section.

@@
 
-This part gives some examples of exploratory data analysis using the data.
-We are interested to explore how people experience changes in their wages over the years.
-Hence, we focus the analysis in their slope of wages.
-Using `key_slope` function in `brolgar` [@brolgar], we perform linear regression with the mean hourly wage as response variable and year as the predictor for each individual.
+This section will provide an example of how data sets could be used to do exploratory data analysis as in textbook data.
+We are interested in examining how the respondents experience changes in their wages based on their demographic characteristics.
+Thus, we use `wages` data instead of the refreshed `wages_hs_do` as it only has male respondents.
+
+The statistics we use in measuring the change in wages experience is the slope of the wages of each respondent. 
+Using `key_slope` function in `brolgar` [@brolgar], we perform linear regression with the log of mean hourly wage as the response variable and year as the predictor.
+Since we do modeling, it is important to look at the number of observations from each respondent.
+Figure \ref{fig:slope-nobs-plot} shows that the number of observations for each individual varies.
+Moreover, each number of observations seems to have an impact on the slope.
+Extreme values in slope are observed in a few observations, while the range of slope is more stable with more observations.
+Accordingly, we trim the data to only include $5^{th} - 12^{th}$ percentile.
 
@@
 
+We then create some visualizations of five respondents with the highest slope and five respondents with the lowest slope, henceforth referred to as top-5 slopes and bottom-5 slopes, respectively.
+We plot them based on their gender and education.
+Figure \ref{fig:slope-gender} shows top-5 slopes are dominated by males, while females dominate bottom-5 slopes.
+However, as this figure only shows a small portion of the data, we further inspect the gender wage gap.
+
 ```{r slope-gender, fig.cap = "Wages slope over the years based on gender.
@@
 
-According to @OECD, gender pay gap is the difference between median wage of males and females relative to median wage of males.
+According to @OECD, the gender pay gap is the difference between the median wage of males and females relative to the median wage of males.
 We find that, as shown in \ref{fig:gap-gender}, over the years, females tend to earn fewer than males.
-For example, in 2018 females earn about 78 cents for every $1 earned by males.
-The widest gender gaps were observed in 1990s.
+For example, in 2018, females earned about 78 cents for every $1 earned by males.
+The widest gender gaps were observed in the 1990s.
 
@@
 
-a
-b
-c
-d
-e
-f
-gh
+According to their highest education completed, both top-5 and bottom-5 slopes are dominated by people who completed high school.
+This is understandable since the respondents are mostly completed high school.
 
@@
 
-This paper has performed a set of stages to make open data suitable for textbook data or make it ready for research.
+This paper has performed stages to make open data suitable for textbook data or make it ready for research.
 In the first stage, we showed the steps performed to get the data from the NLSY79 database.
 Since the data format is untidy, we showed how the data had been tidied.
-After that, we conducted initial data analysis to investigate and screen the quality of the data.
-Using the robust linear regression model, we found and fixed the anomalous observations in the data set with its predicted values.
+After that, we conducted an initial data analysis to investigate and screen the quality of the data.
+We found and fixed the anomalous observations in the data set using the robust linear regression model with its predicted values.
 We also performed an example of exploratory data analysis using the cleaned data set.
@@
 Moreover, the documentation also includes how we generated the robustness weight and how we decided the threshold of anomalous observations.
-It is also documented along with the flag of whether an observation is imputed value or not.
+It is also documented along with the flag of whether an observation is an imputed value or not.
 Accordingly, if somebody wishes to make another decision, it can be done by making a small change in the code provided.

### 2021-09-09 paper/paper.tex
Add explanation to EDA section.

@@
 % Pandoc citation processing
-
-\usepackage{booktabs}
-\usepackage{longtable}
-\usepackage{array}
-\usepackage{multirow}
-\usepackage{wrapfig}
-\usepackage{float}
-\usepackage{colortbl}
-\usepackage{pdflscape}
-\usepackage{tabu}
-\usepackage{threeparttable}
-\usepackage{threeparttablex}
-\usepackage[normalem]{ulem}
-\usepackage{makecell}
-\usepackage{xcolor}
 
@@
 
-This part gives some examples of exploratory data analysis using the data.
-We are interested to explore how people experience changes in their wages over the years.
-Hence, we focus the analysis in their slope of wages.
-Using \texttt{key\_slope} function in \texttt{brolgar} (Tierney, Cook, and Prvan 2020), we perform linear regression with the mean hourly wage as response variable and year as the predictor for each individual.
+This section will provide an example of how data sets could be used to do exploratory data analysis as in textbook data.
+We are interested in examining how the respondents experience changes in their wages based on their demographic characteristics.
+Thus, we use \texttt{wages} data instead of the refreshed \texttt{wages\_hs\_do} as it only has male respondents.
+
+The statistics we use in measuring the change in wages experience is the slope of the wages of each respondent.
+Using \texttt{key\_slope} function in \texttt{brolgar} (Tierney, Cook, and Prvan 2020), we perform linear regression with the log of mean hourly wage as the response variable and year as the predictor.
+Since we do modeling, it is important to look at the number of observations from each respondent.
+Figure \ref{fig:slope-nobs-plot} shows that the number of observations for each individual varies.
+Moreover, each number of observations seems to have an impact on the slope.
+Extreme values in slope are observed in a few observations, while the range of slope is more stable with more observations.
+Accordingly, we trim the data to only include \(5^{th} - 12^{th}\) percentile.
 
@@
 
+We then create some visualizations of five respondents with the highest slope and five respondents with the lowest slope, henceforth referred to as top-5 slopes and bottom-5 slopes, respectively.
+We plot them based on their gender and education.
+Figure \ref{fig:slope-gender} shows top-5 slopes are dominated by males, while females dominate bottom-5 slopes.
+However, as this figure only shows a small portion of the data, we further inspect the gender wage gap.
+
 \begin{figure}
@@
 
-According to OECD (n.d.), gender pay gap is the difference between median wage of males and females relative to median wage of males.
+According to OECD (n.d.), the gender pay gap is the difference between the median wage of males and females relative to the median wage of males.
 We find that, as shown in \ref{fig:gap-gender}, over the years, females tend to earn fewer than males.
-For example, in 2018 females earn about 78 cents for every \$1 earned by males.
-The widest gender gaps were observed in 1990s.
+For example, in 2018, females earned about 78 cents for every \$1 earned by males.
+The widest gender gaps were observed in the 1990s.
 

### 2021-09-13 paper/paper.Rmd
Change data trimming for eda.

@@
                comment = "#>",
-               fig.align = 'center', #fig.show = 'hold',
+               fig.align = 'center', 
                out.width = "70%")
@@
 Extreme values in slope are observed in a few observations, while the range of slope is more stable with more observations.
-Accordingly, we trim the data to only include $5^{th} - 12^{th}$ percentile.
+Accordingly, we trim the data to only include $5^{th} - 95^{th}$ percentile.
 
@@
 
-```{r gap-gender, fig.cap="Gender wage gap or the median wages of females relatives to median wages of males.
+```{r gap-gender, fig.cap="A) Number of observations by gender and by year.
+There are more males participated in survey than females.
+The number of observations decreased over the years.
+B) Gender wage gap or the median wages of females relatives to median wages of males.
 Females tend to earn fewer than males.
 For example, in 2018 females earn about 78 cents for every \\$1 earned by males.
-Note that this is a crude measure as gap is calculated not based on other variable, such as experience and job class."}
+Note that this is a crude measure as gap is calculated not based on other variable, such as experience and job class.", out.width="90%"}
 ```

### 2021-09-13 paper/paper.tex
Change data trimming for eda.

@@
 % Pandoc citation processing
+
+\usepackage{booktabs}
+\usepackage{longtable}
+\usepackage{array}
+\usepackage{multirow}
+\usepackage{wrapfig}
+\usepackage{float}
+\usepackage{colortbl}
+\usepackage{pdflscape}
+\usepackage{tabu}
+\usepackage{threeparttable}
+\usepackage{threeparttablex}
+\usepackage[normalem]{ulem}
+\usepackage{makecell}
+\usepackage{xcolor}
 
@@
 Extreme values in slope are observed in a few observations, while the range of slope is more stable with more observations.
-Accordingly, we trim the data to only include \(5^{th} - 12^{th}\) percentile.
+Accordingly, we trim the data to only include \(5^{th} - 95^{th}\) percentile.
 
@@
 
-{\centering \includegraphics[width=0.7\linewidth]{figures/gap-gender-1}
+{\centering \includegraphics[width=0.9\linewidth]{figures/gap-gender-1}
 
@@
 
-\caption{Gender wage gap or the median wages of females relatives to median wages of males.
+\caption{A) Number of observations by gender and by year.
+There are more males participated in survey than females.
+The number of observations decreased over the years.
+B) Gender wage gap or the median wages of females relatives to median wages of males.
 Females tend to earn fewer than males.

### 2021-09-13 paper/scripts/eda.R

@@ -43,9 +43,8 @@ wages_obs_slop_p <- ggplot(wages_obs_slope, aes(x = n_obs, y = .slope_year0)) +
 wages_obs_p + wages_obs_slop_p
 
 ## ---- trim-slope
-slope_trimmed <- wages_slope %>%
-  filter(.slope_year0 < quantile(.slope_year0, 0.95) &
-           .slope_year0 > quantile(.slope_year0, 0.05))
+slope_trimmed <- wages_obs_slope %>%
+  filter(n_obs >= 10)
 
 ## ---- slope-gender
 slope_bottom_5<- slope_trimmed %>%
@@ -80,6 +79,8 @@ ggplot(contrast_slope, aes(x = year, y = ln_wages, color = gender)) +
 
 ## ---- gap-gender
 wages_gender <- as.tibble(wages) %>%
+  left_join(wages_obs, by = "id") %>%
+  filter(n_obs >= 10) %>%
   group_by(gender, year) %>%
   summarise(mean = mean(wage),
             median = median(wage),
@@ -89,6 +90,15 @@ wages_gender <- as.tibble(wages) %>%
   mutate(upper = mean + sd/sqrt(samplesize),
          lower = mean - sd/sqrt(samplesize))
 
+# number of observations by gender
+nobs_gender <- ggplot(wages_gender, aes(fill=gender, y=samplesize, x=year)) +
+  geom_bar(position="dodge", stat="identity") +
+  theme(legend.position = "bottom") +
+  ylab("number of observations") +
+  scale_fill_manual(values = c("#088A68", "#DF7401")) +
+  ggtitle("A)")
+
+# wages by gender
 wage_female <- wages_gender %>%
   filter(gender == "FEMALE") %>%
   dplyr::select(year, median) %>%
@@ -103,7 +113,7 @@ gender_wage_gap <- cbind(wage_female, wage_male) %>%
   mutate(gap = (male-female)/male) %>%
   mutate(relative = 1-gap)
 
-ggplot(gender_wage_gap) +
+gender_gap_plot <- ggplot(gender_wage_gap) +
   geom_line(aes(x = year,
                 y = relative),
             color = "#DF7401", size = 0.8) +
@@ -113,8 +123,11 @@ ggplot(gender_wage_gap) +
   geom_hline(yintercept = 1, colour = "#088A68", size = 1) +
   #scale_x_continuous(limits = c(1978.7,2018.3), expand = c(0, 0)) +
[…]

### 2021-09-14 paper/paper.Rmd
abstract

@@
 
-NSLY79 is a prominent open data source that has been an important support in multidisciplinary research on longitudinal data.
+NSLY79 is a prominent open data source that has been an important support for educational purposes and multidisciplinary research on longitudinal data.
 Subsets of this data can be found in numerous textbooks and research articles.
 However, the steps and decisions taken to get from the original data to these subsets is never clearly articulated.
-This article describes our journey when trying to re-create a textbook example data set from the original data, with a goal to refresh it for use in the classroom.
-It thus demonstrates the process of initial data analysis (IDA) -- tidying, cleaning, and documenting the process -- to make data available for text book examples or research.
+This article describes our journey when trying to re-create a textbook example data set from the original database.
+Ultimately, the goal is to provide the resources so that refreshing data subsets can be done on a regular basis, especially when updates are made to the database.
+Thus, this paper demonstrates the process of initial data analysis (IDA) -- tidying, cleaning, and documenting the process -- to make data available for text book examples or research.
+A brief exploratory analysis of the resulting data is also conducted.
 Three new data sets, and the code to produce them are provided in an accompanying open source R package, called `yowie`.
-The code might be useful for future refreshing of this particular text book data.
-As a result of this process, some recommendations are made for the NLSY79 curators for incorporating data quality checks, and providing more convenient samples of the data to potential users.
+As a result of this process, some recommendations are also made for the NLSY79 curators for incorporating data quality checks, and providing more convenient samples of the data to potential users.
 

### 2021-09-14 paper/paper.Rmd
pull the latest change.

--- 2021-09-13 paper/paper.Rmd
@@
 Extreme values in slope are observed in a few observations, while the range of slope is more stable with more observations.
-Accordingly, we trim the data to only include $5^{th} - 95^{th}$ percentile.
+Accordingly, we trim the data to only include respondents with at least 10 observations.
 
@@
 The widest gender gaps were observed in the 1990s.
+However, as more males participated in the survey than females from year to year, these gaps could be affected by imbalance number of observations.
 
@@
 ```
+
+Further, we perform a linear regression to see the effect of education on wages experience.
 

### 2021-09-14 paper/scripts/eda.R

@@ -139,3 +139,8 @@ ggplot(contrast_slope, aes(x = year, y = ln_wages, color = hgc_regroup)) +
   theme_bw() +
   theme(axis.text.x = element_text(angle = 25),
         legend.position = "bottom")
+
+## ---- regression
+
+fit <- lm(.slope_year0 ~ age_1979 + hgc_regroup*gender, data = slope_trimmed)
+summary(fit)

### 2021-09-14 paper/paper.Rmd
twaking plot captions and order of plots

--- 2021-09-14 paper/paper.Rmd
@@
 
-# The NLSY79 Data Cleaning {#cleaning}
+# Data Cleaning {#cleaning}
 
@@
 
+```{r sample-plot, echo = FALSE, fig.cap="Getting a glimpse of the data: random sample of 36 individuals, divided into  twelve subplots.
+Wages are shown against year.
+There is a lot of variability in wages.", fig.height=4, fig.width=5}
+```
+
 However, we can not be sure only by observing the summary statistics.
@@
 
-```{r feature-plot, fig.cap = "Two plots showing the distribution of the mean hourly wage.
+```{r feature-plot, fig.cap = "Several plots showing anomalies observed the input data.
+XX Add a couple more plots that reveal the initial problems in the data.
 Plot A portrays the pattern of mean hourly wage of high school cohort from 1979 to 2018 of each ID in US Dollar; Plot B shows the distribution of their minimum, median, and maximum value.
@@
 Some have had flat wages for years but had a sudden increase in one particular year, then it gone down again, while the others experienced an upsurge in their wage, for instance, the IDs in panel 9.
-
-```{r sample-plot, echo = FALSE, fig.cap="The mean hourly wages of some random samples are shown in twelve facets, three IDs per facet.
-It suggests that some IDs had a reasonably fluctuate wages.", fig.height=4, fig.width=5}
-```
 

### 2021-09-14 paper/paper.tex
twaking plot captions and order of plots

@@
 
-NSLY79 is a prominent open data source that has been an important support in multidisciplinary research on longitudinal data.
+NSLY79 is a prominent open data source that has been an important support for educational purposes and multidisciplinary research on longitudinal data.
 Subsets of this data can be found in numerous textbooks and research articles.
 However, the steps and decisions taken to get from the original data to these subsets is never clearly articulated.
-This article describes our journey when trying to re-create a textbook example data set from the original data, with a goal to refresh it for use in the classroom.
-It thus demonstrates the process of initial data analysis (IDA) -- tidying, cleaning, and documenting the process -- to make data available for text book examples or research.
+This article describes our journey when trying to re-create a textbook example data set from the original database.
+Ultimately, the goal is to provide the resources so that refreshing data subsets can be done on a regular basis, especially when updates are made to the database.
+Thus, this paper demonstrates the process of initial data analysis (IDA) -- tidying, cleaning, and documenting the process -- to make data available for text book examples or research.
+A brief exploratory analysis of the resulting data is also conducted.
 Three new data sets, and the code to produce them are provided in an accompanying open source R package, called \texttt{yowie}.
-The code might be useful for future refreshing of this particular text book data.
-As a result of this process, some recommendations are made for the NLSY79 curators for incorporating data quality checks, and providing more convenient samples of the data to potential users.
+As a result of this process, some recommendations are also made for the NLSY79 curators for incorporating data quality checks, and providing more convenient samples of the data to potential users.
 
@@
 However, open data is often referred to as what we might call ``wild data'' because it requires substantial cleaning and tidying to tame it into textbook shape.
-Huebner, Vach, and Cessie (2016) emphasize that making the data cleaning process accountable and transparent is imperative.
-Documenting the data cleaning is essential for the integrity of downstream statistical analyses and model building (Huebner et al.
+Huebner Marianne, Vach, and Cessie (2016) emphasize that making the data cleaning process accountable and transparent is imperative.
+Documenting the data cleaning is essential for the integrity of downstream statistical analyses and model building (M.
+Huebner et al.
 2020).
@@
 There are few research papers that document the data cleaning (Wickham 2014).
-Furthermore, the decisions made in this stage often go unreported in the sense that IDA is often performed in an unplanned and unstructured way and is only shared among restricted parties (Huebner et al.
+Furthermore, the decisions made in this stage often go unreported in the sense that IDA is often performed in an unplanned and unstructured way and is only shared among restricted parties (M.
+Huebner et al.
 2020).
@@
 \hypertarget{cleaning}{%
-\section{The NLSY79 Data Cleaning}\label{cleaning}}
+\section{Data Cleaning}\label{cleaning}}
 
@@
 Consequently, the data should be tidied and wrangled first to extract the demographic and employment variables we want to put in the final data set.
-We mainly used \texttt{tidyr} (H.
-Wickham 2020b), \texttt{dplyr} (H.
-Wickham, François, et al.
-2020), and \texttt{stringr} (H.
-Wickham 2019) to do this job.
+We mainly used \texttt{tidyr} (Wickham 2020b), \texttt{dplyr} (Wickham et al.
+2020), and \texttt{stringr} (Wickham 2019) to do this job.
 
@@
 \begin{Highlighting}[]
-\CommentTok{#tidy the date of birth data}
-\NormalTok{dob_tidy <-}\StringTok{ }\NormalTok{new_data_qnames }\OperatorTok{%>%}
-\StringTok{  }\NormalTok{dplyr}\OperatorTok{::}\KeywordTok{select}\NormalTok{(CASEID_}\DecValTok{1979}\NormalTok{,}
-                \KeywordTok{starts_with}\NormalTok{(}\StringTok{"Q1-3_A~"}\NormalTok{)) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{dob_year =} \KeywordTok{case_when}\NormalTok{(}
-    \CommentTok{# if the years recorded in both sets match, take 79 data}
-    \StringTok{`}\DataTypeTok{Q1-3_A~Y_1979}\StringTok{`} \OperatorTok{==}\StringTok{ `}\DataTypeTok{Q1-3_A~Y_1981}\StringTok{`} \OperatorTok{~}\StringTok{ `}\DataTypeTok{Q1-3_A~Y_1979}\StringTok{`}\NormalTok{,}
-    \CommentTok{# if the year in the 81 set is missing, take the 79 data}
-    \KeywordTok{is.na}\NormalTok{(}\StringTok{`}\DataTypeTok{Q1-3_A~Y_1981}\StringTok{`}\NormalTok{) }\OperatorTok{~}\StringTok{ `}\DataTypeTok{Q1-3_A~Y_1979}\StringTok{`}\NormalTok{,}
-    \CommentTok{# if the sets don't match for dob year, take the 79 data}
-    \StringTok{`}\DataTypeTok{Q1-3_A~Y_1979}\StringTok{`} \OperatorTok{!=}\StringTok{ `}\DataTypeTok{Q1-3_A~Y_1981}\StringTok{`} \OperatorTok{~}\StringTok{ `}\DataTypeTok{Q1-3_A~Y_1979}\StringTok{`}\NormalTok{),}
-    \DataTypeTok{dob_month =} \KeywordTok{case_when}\NormalTok{(}
-      \CommentTok{# if months recorded in both sets match, take 79 data}
-      \StringTok{`}\DataTypeTok{Q1-3_A~M_1979}\StringTok{`} \OperatorTok{==}\StringTok{ `}\DataTypeTok{Q1-3_A~M_1981}\StringTok{`} \OperatorTok{~}\StringTok{ `}\DataTypeTok{Q1-3_A~M_1979}\StringTok{`}\NormalTok{,}
-      \CommentTok{# if month in 81 set is missing, take the 79 data}
-      \KeywordTok{is.na}\NormalTok{(}\StringTok{`}\DataTypeTok{Q1-3_A~M_1981}\StringTok{`}\NormalTok{) }\OperatorTok{~}\StringTok{ `}\DataTypeTok{Q1-3_A~M_1979}\StringTok{`}\NormalTok{,}
-      \CommentTok{# if sets don't match for dob month, take the 79 data}
-      \StringTok{`}\DataTypeTok{Q1-3_A~M_1979}\StringTok{`} \OperatorTok{!=}\StringTok{ `}\DataTypeTok{Q1-3_A~M_1981}\StringTok{`} \OperatorTok{~}\StringTok{ `}\DataTypeTok{Q1-3_A~M_1979}\StringTok{`}\NormalTok{),}
-    \CommentTok{# flag if there is a conflict between dob recorded in 79 and 81}
-    \DataTypeTok{dob_conflict =} \KeywordTok{case_when}\NormalTok{(}
-\NormalTok{      (}\StringTok{`}\DataTypeTok{Q1-3_A~M_1979}\StringTok{`} \OperatorTok{!=}\StringTok{ `}\DataTypeTok{Q1-3_A~M_1981}\StringTok{`}\NormalTok{) }\OperatorTok{&}\StringTok{ }\OperatorTok{!}\KeywordTok{is.na}\NormalTok{(}\StringTok{`}\DataTypeTok{Q1-3_A~M_1981}\StringTok{`}\NormalTok{)}
-      \OperatorTok{~}\StringTok{ }\OtherTok{TRUE}\NormalTok{,}
-\NormalTok{      (}\StringTok{`}\DataTypeTok{Q1-3_A~Y_1979}\StringTok{`} \OperatorTok{!=}\StringTok{ `}\DataTypeTok{Q1-3_A~Y_1981}\StringTok{`}\NormalTok{) }\OperatorTok{&}\StringTok{ }\OperatorTok{!}\KeywordTok{is.na}\NormalTok{(}\StringTok{`}\DataTypeTok{Q1-3_A~Y_1981}\StringTok{`}\NormalTok{)}
-      \OperatorTok{~}\StringTok{ }\OtherTok{TRUE}\NormalTok{,}
-\NormalTok{      (}\StringTok{`}\DataTypeTok{Q1-3_A~Y_1979}\StringTok{`} \OperatorTok{==}\StringTok{ `}\DataTypeTok{Q1-3_A~Y_1981}\StringTok{`}\NormalTok{) }\OperatorTok{&}
-\StringTok{        }\NormalTok{(}\StringTok{`}\DataTypeTok{Q1-3_A~M_1979}\StringTok{`} \OperatorTok{==}\StringTok{ `}\DataTypeTok{Q1-3_A~M_1981}\StringTok{`}\NormalTok{) }\OperatorTok{~}\StringTok{ }\OtherTok{FALSE}\NormalTok{,}
-      \KeywordTok{is.na}\NormalTok{(}\StringTok{`}\DataTypeTok{Q1-3_A~M_1981}\StringTok{`}\NormalTok{) }\OperatorTok{|}\StringTok{ }\KeywordTok{is.na}\NormalTok{(}\StringTok{`}\DataTypeTok{Q1-3_A~Y_1981}\StringTok{`}\NormalTok{) }\OperatorTok{~}\StringTok{ }\OtherTok{FALSE}\NormalTok{)) }\OperatorTok{%>%}
-\StringTok{  }\NormalTok{dplyr}\OperatorTok{::}\KeywordTok{select}\NormalTok{(CASEID_}\DecValTok{1979}\NormalTok{,}
-\NormalTok{                dob_month,}
-\NormalTok{                dob_year,}
-\NormalTok{                dob_conflict)}
+\CommentTok{\#tidy the date of birth data}
+\NormalTok{dob\_tidy }\OtherTok{\textless{}{-}}\NormalTok{ new\_data\_qnames }\SpecialCharTok{\%\textgreater{}\%}
+\NormalTok{  dplyr}\SpecialCharTok{::}\FunctionTok{select}\NormalTok{(CASEID\_1979,}
+                \FunctionTok{starts\_with}\NormalTok{(}\StringTok{"Q1{-}3\_A\textasciitilde{}"}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{dob\_year =} \FunctionTok{case\_when}\NormalTok{(}
+    \CommentTok{\# if the years recorded in both sets match, take 79 data}
+    \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}Y\_1979}\StringTok{\textasciigrave{}} \SpecialCharTok{==} \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}Y\_1981}\StringTok{\textasciigrave{}} \SpecialCharTok{\textasciitilde{}} \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}Y\_1979}\StringTok{\textasciigrave{}}\NormalTok{,}
+    \CommentTok{\# if the year in the 81 set is missing, take the 79 data}
+    \FunctionTok{is.na}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}Y\_1981}\StringTok{\textasciigrave{}}\NormalTok{) }\SpecialCharTok{\textasciitilde{}} \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}Y\_1979}\StringTok{\textasciigrave{}}\NormalTok{,}
+    \CommentTok{\# if the sets don\textquotesingle{}t match for dob year, take the 79 data}
+    \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}Y\_1979}\StringTok{\textasciigrave{}} \SpecialCharTok{!=} \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}Y\_1981}\StringTok{\textasciigrave{}} \SpecialCharTok{\textasciitilde{}} \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}Y\_1979}\StringTok{\textasciigrave{}}\NormalTok{),}
+    \AttributeTok{dob\_month =} \FunctionTok{case\_when}\NormalTok{(}
+      \CommentTok{\# if months recorded in both sets match, take 79 data}
+      \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}M\_1979}\StringTok{\textasciigrave{}} \SpecialCharTok{==} \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}M\_1981}\StringTok{\textasciigrave{}} \SpecialCharTok{\textasciitilde{}} \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}M\_1979}\StringTok{\textasciigrave{}}\NormalTok{,}
+      \CommentTok{\# if month in 81 set is missing, take the 79 data}
+      \FunctionTok{is.na}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}M\_1981}\StringTok{\textasciigrave{}}\NormalTok{) }\SpecialCharTok{\textasciitilde{}} \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}M\_1979}\StringTok{\textasciigrave{}}\NormalTok{,}
+      \CommentTok{\# if sets don\textquotesingle{}t match for dob month, take the 79 data}
+      \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}M\_1979}\StringTok{\textasciigrave{}} \SpecialCharTok{!=} \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}M\_1981}\StringTok{\textasciigrave{}} \SpecialCharTok{\textasciitilde{}} \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}M\_1979}\StringTok{\textasciigrave{}}\NormalTok{),}
+    \CommentTok{\# flag if there is a conflict between dob recorded in 79 and 81}
+    \AttributeTok{dob\_conflict =} \FunctionTok{case\_when}\NormalTok{(}
+\NormalTok{      (}\StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}M\_1979}\StringTok{\textasciigrave{}} \SpecialCharTok{!=} \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}M\_1981}\StringTok{\textasciigrave{}}\NormalTok{) }\SpecialCharTok{\&} \SpecialCharTok{!}\FunctionTok{is.na}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}M\_1981}\StringTok{\textasciigrave{}}\NormalTok{)}
+      \SpecialCharTok{\textasciitilde{}} \ConstantTok{TRUE}\NormalTok{,}
+\NormalTok{      (}\StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}Y\_1979}\StringTok{\textasciigrave{}} \SpecialCharTok{!=} \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}Y\_1981}\StringTok{\textasciigrave{}}\NormalTok{) }\SpecialCharTok{\&} \SpecialCharTok{!}\FunctionTok{is.na}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}Y\_1981}\StringTok{\textasciigrave{}}\NormalTok{)}
+      \SpecialCharTok{\textasciitilde{}} \ConstantTok{TRUE}\NormalTok{,}
+\NormalTok{      (}\StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}Y\_1979}\StringTok{\textasciigrave{}} \SpecialCharTok{==} \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}Y\_1981}\StringTok{\textasciigrave{}}\NormalTok{) }\SpecialCharTok{\&}
+\NormalTok{        (}\StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}M\_1979}\StringTok{\textasciigrave{}} \SpecialCharTok{==} \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}M\_1981}\StringTok{\textasciigrave{}}\NormalTok{) }\SpecialCharTok{\textasciitilde{}} \ConstantTok{FALSE}\NormalTok{,}
+      \FunctionTok{is.na}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}M\_1981}\StringTok{\textasciigrave{}}\NormalTok{) }\SpecialCharTok{|} \FunctionTok{is.na}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}Y\_1981}\StringTok{\textasciigrave{}}\NormalTok{) }\SpecialCharTok{\textasciitilde{}} \ConstantTok{FALSE}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%}
+\NormalTok{  dplyr}\SpecialCharTok{::}\FunctionTok{select}\NormalTok{(CASEID\_1979,}
+\NormalTok{                dob\_month,}
+\NormalTok{                dob\_year,}
+\NormalTok{                dob\_conflict)}
 \end{Highlighting}
@@
 \begin{Highlighting}[]
-\CommentTok{# make a list for years where we used the "QES-52A"}
-\NormalTok{year_A <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\DecValTok{1979}\OperatorTok{:}\DecValTok{1987}\NormalTok{, }\DecValTok{1993}\NormalTok{)}
-\CommentTok{#function to get the hour of work}
-\NormalTok{get_hour <-}\StringTok{ }\ControlFlowTok{function}\NormalTok{(year)\{}
-  \ControlFlowTok{if}\NormalTok{(year }\OperatorTok{%in%}\StringTok{ }\NormalTok{year_A)\{}
-\NormalTok{    temp <-}\StringTok{ }\NormalTok{new_data_qnames }\OperatorTok{%>%}
-\StringTok{      }\NormalTok{dplyr}\OperatorTok{::}\KeywordTok{select}\NormalTok{(CASEID_}\DecValTok{1979}\NormalTok{,}
-                    \KeywordTok{starts_with}\NormalTok{(}\StringTok{"QES-52A"}\NormalTok{) }\OperatorTok{&}
-\StringTok{                      }\KeywordTok{ends_with}\NormalTok{(}\KeywordTok{as.character}\NormalTok{(year)))\}}
+\CommentTok{\# make a list for years where we used the "QES{-}52A"}
+\NormalTok{year\_A }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(}\DecValTok{1979}\SpecialCharTok{:}\DecValTok{1987}\NormalTok{, }\DecValTok{1993}\NormalTok{)}
+\CommentTok{\#function to get the hour of work}
+\NormalTok{get\_hour }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{(year)\{}
+  \ControlFlowTok{if}\NormalTok{(year }\SpecialCharTok{\%in\%}\NormalTok{ year\_A)\{}
+\NormalTok{    temp }\OtherTok{\textless{}{-}}\NormalTok{ new\_data\_qnames }\SpecialCharTok{\%\textgreater{}\%}
+\NormalTok{      dplyr}\SpecialCharTok{::}\FunctionTok{select}\NormalTok{(CASEID\_1979,}
+                    \FunctionTok{starts\_with}\NormalTok{(}\StringTok{"QES{-}52A"}\NormalTok{) }\SpecialCharTok{\&}
+                      \FunctionTok{ends\_with}\NormalTok{(}\FunctionTok{as.character}\NormalTok{(year)))\}}
   \ControlFlowTok{else}\NormalTok{\{}
-\NormalTok{    temp <-}\StringTok{ }\NormalTok{new_data_qnames }\OperatorTok{%>%}
-\StringTok{      }\NormalTok{dplyr}\OperatorTok{::}\KeywordTok{select}\NormalTok{(CASEID_}\DecValTok{1979}\NormalTok{,}
-                    \KeywordTok{starts_with}\NormalTok{(}\StringTok{"QES-52D"}\NormalTok{) }\OperatorTok{&}
-\StringTok{                      }\KeywordTok{ends_with}\NormalTok{(}\KeywordTok{as.character}\NormalTok{(year)))\}}
-\NormalTok{  temp }\OperatorTok{%>%}
-\StringTok{    }\KeywordTok{pivot_longer}\NormalTok{(}\OperatorTok{!}\NormalTok{CASEID_}\DecValTok{1979}\NormalTok{,}
-                 \DataTypeTok{names_to =} \StringTok{"job"}\NormalTok{,}
-                 \DataTypeTok{values_to =} \StringTok{"hours_work"}\NormalTok{) }\OperatorTok{%>%}
-\StringTok{    }\KeywordTok{separate}\NormalTok{(}\StringTok{"job"}\NormalTok{, }\KeywordTok{c}\NormalTok{(}\StringTok{"job"}\NormalTok{, }\StringTok{"year"}\NormalTok{), }\DataTypeTok{sep =} \DecValTok{-4}\NormalTok{) }\OperatorTok{%>%}
-\StringTok{    }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{job =} \KeywordTok{paste0}\NormalTok{(}\StringTok{"job_"}\NormalTok{, }\KeywordTok{substr}\NormalTok{(job, }\DecValTok{9}\NormalTok{, }\DecValTok{10}\NormalTok{))) }\OperatorTok{%>%}
-\StringTok{    }\KeywordTok{rename}\NormalTok{(}\DataTypeTok{id =}\NormalTok{ CASEID_}\DecValTok{1979}\NormalTok{)}
+\NormalTok{    temp }\OtherTok{\textless{}{-}}\NormalTok{ new\_data\_qnames }\SpecialCharTok{\%\textgreater{}\%}
+\NormalTok{      dplyr}\SpecialCharTok{::}\FunctionTok{select}\NormalTok{(CASEID\_1979,}
+                    \FunctionTok{starts\_with}\NormalTok{(}\StringTok{"QES{-}52D"}\NormalTok{) }\SpecialCharTok{\&}
+                      \FunctionTok{ends\_with}\NormalTok{(}\FunctionTok{as.character}\NormalTok{(year)))\}}
+\NormalTok{  temp }\SpecialCharTok{\%\textgreater{}\%}
+    \FunctionTok{pivot\_longer}\NormalTok{(}\SpecialCharTok{!}\NormalTok{CASEID\_1979,}
+                 \AttributeTok{names\_to =} \StringTok{"job"}\NormalTok{,}
+                 \AttributeTok{values\_to =} \StringTok{"hours\_work"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
+    \FunctionTok{separate}\NormalTok{(}\StringTok{"job"}\NormalTok{, }\FunctionTok{c}\NormalTok{(}\StringTok{"job"}\NormalTok{, }\StringTok{"year"}\NormalTok{), }\AttributeTok{sep =} \SpecialCharTok{{-}}\DecValTok{4}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
+    \FunctionTok{mutate}\NormalTok{(}\AttributeTok{job =} \FunctionTok{paste0}\NormalTok{(}\StringTok{"job\_"}\NormalTok{, }\FunctionTok{substr}\NormalTok{(job, }\DecValTok{9}\NormalTok{, }\DecValTok{10}\NormalTok{))) }\SpecialCharTok{\%\textgreater{}\%}
+    \FunctionTok{rename}\NormalTok{(}\AttributeTok{id =}\NormalTok{ CASEID\_1979)}
 \NormalTok{\}}
-\CommentTok{# list to save the iteration result}
-\NormalTok{hours <-}\StringTok{ }\KeywordTok{list}\NormalTok{()}
-\CommentTok{# getting the hours of work of all observations}
-\ControlFlowTok{for}\NormalTok{(ayear }\ControlFlowTok{in} \KeywordTok{c}\NormalTok{(}\DecValTok{1979}\OperatorTok{:}\DecValTok{1994}\NormalTok{, }\DecValTok{1996}\NormalTok{, }\DecValTok{1998}\NormalTok{, }\DecValTok{2000}\NormalTok{, }\DecValTok{2002}\NormalTok{, }\DecValTok{2004}\NormalTok{, }\DecValTok{2006}\NormalTok{, }\DecValTok{2008}\NormalTok{, }\DecValTok{2010}\NormalTok{,}
+\CommentTok{\# list to save the iteration result}
+\NormalTok{hours }\OtherTok{\textless{}{-}} \FunctionTok{list}\NormalTok{()}
+\CommentTok{\# getting the hours of work of all observations}
+\ControlFlowTok{for}\NormalTok{(ayear }\ControlFlowTok{in} \FunctionTok{c}\NormalTok{(}\DecValTok{1979}\SpecialCharTok{:}\DecValTok{1994}\NormalTok{, }\DecValTok{1996}\NormalTok{, }\DecValTok{1998}\NormalTok{, }\DecValTok{2000}\NormalTok{, }\DecValTok{2002}\NormalTok{, }\DecValTok{2004}\NormalTok{, }\DecValTok{2006}\NormalTok{, }\DecValTok{2008}\NormalTok{, }\DecValTok{2010}\NormalTok{,}
                \DecValTok{2012}\NormalTok{, }\DecValTok{2014}\NormalTok{, }\DecValTok{2016}\NormalTok{, }\DecValTok{2018}\NormalTok{)) \{}
-\NormalTok{  hours[[ayear]] <-}\StringTok{ }\KeywordTok{get_hour}\NormalTok{(ayear)}
+\NormalTok{  hours[[ayear]] }\OtherTok{\textless{}{-}} \FunctionTok{get\_hour}\NormalTok{(ayear)}
 \NormalTok{\}}
-\CommentTok{# unlist the hours of work}
-\NormalTok{hours_all <-}\StringTok{ }\KeywordTok{bind_rows}\NormalTok{(}\OperatorTok{!!!}\NormalTok{hours)}
+\CommentTok{\# unlist the hours of work}
+\NormalTok{hours\_all }\OtherTok{\textless{}{-}} \FunctionTok{bind\_rows}\NormalTok{(}\SpecialCharTok{!!!}\NormalTok{hours)}
 \end{Highlighting}
@@
 \begin{Highlighting}[]
-\NormalTok{get_rate <-}\StringTok{ }\ControlFlowTok{function}\NormalTok{(year) \{}
-\NormalTok{  new_data_qnames }\OperatorTok{%>%}
-\StringTok{    }\NormalTok{dplyr}\OperatorTok{::}\KeywordTok{select}\NormalTok{(CASEID_}\DecValTok{1979}\NormalTok{,}
-                  \KeywordTok{starts_with}\NormalTok{(}\StringTok{"HRP"}\NormalTok{) }\OperatorTok{&}
-\StringTok{                    }\KeywordTok{ends_with}\NormalTok{(}\KeywordTok{as.character}\NormalTok{(year))) }\OperatorTok{%>%}
-\StringTok{    }\KeywordTok{pivot_longer}\NormalTok{(}\OperatorTok{!}\NormalTok{CASEID_}\DecValTok{1979}\NormalTok{, }\DataTypeTok{names_to =} \StringTok{"job"}\NormalTok{, }\DataTypeTok{values_to =} \StringTok{"rate_per_hour"}\NormalTok{) }\OperatorTok{%>%}
-\StringTok{    }\KeywordTok{separate}\NormalTok{(}\StringTok{"job"}\NormalTok{, }\KeywordTok{c}\NormalTok{(}\StringTok{"job"}\NormalTok{, }\StringTok{"year"}\NormalTok{), }\DataTypeTok{sep =} \DecValTok{-4}\NormalTok{) }\OperatorTok{%>%}
-\StringTok{    }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{job =} \KeywordTok{paste0}\NormalTok{(}\StringTok{"job_0"}\NormalTok{, }\KeywordTok{substr}\NormalTok{(job, }\DecValTok{4}\NormalTok{, }\DecValTok{4}\NormalTok{))) }\OperatorTok{%>%}
-\StringTok{    }\KeywordTok{rename}\NormalTok{(}\DataTypeTok{id =}\NormalTok{ CASEID_}\DecValTok{1979}\NormalTok{)}
+\NormalTok{get\_rate }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{(year) \{}
+\NormalTok{  new\_data\_qnames }\SpecialCharTok{\%\textgreater{}\%}
+\NormalTok{    dplyr}\SpecialCharTok{::}\FunctionTok{select}\NormalTok{(CASEID\_1979,}
+                  \FunctionTok{starts\_with}\NormalTok{(}\StringTok{"HRP"}\NormalTok{) }\SpecialCharTok{\&}
+                    \FunctionTok{ends\_with}\NormalTok{(}\FunctionTok{as.character}\NormalTok{(year))) }\SpecialCharTok{\%\textgreater{}\%}
+    \FunctionTok{pivot\_longer}\NormalTok{(}\SpecialCharTok{!}\NormalTok{CASEID\_1979, }\AttributeTok{names\_to =} \StringTok{"job"}\NormalTok{, }\AttributeTok{values\_to =} \StringTok{"rate\_per\_hour"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
+    \FunctionTok{separate}\NormalTok{(}\StringTok{"job"}\NormalTok{, }\FunctionTok{c}\NormalTok{(}\StringTok{"job"}\NormalTok{, }\StringTok{"year"}\NormalTok{), }\AttributeTok{sep =} \SpecialCharTok{{-}}\DecValTok{4}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
+    \FunctionTok{mutate}\NormalTok{(}\AttributeTok{job =} \FunctionTok{paste0}\NormalTok{(}\StringTok{"job\_0"}\NormalTok{, }\FunctionTok{substr}\NormalTok{(job, }\DecValTok{4}\NormalTok{, }\DecValTok{4}\NormalTok{))) }\SpecialCharTok{\%\textgreater{}\%}
+    \FunctionTok{rename}\NormalTok{(}\AttributeTok{id =}\NormalTok{ CASEID\_1979)}
 \NormalTok{\}}
-\NormalTok{rates <-}\StringTok{ }\KeywordTok{list}\NormalTok{()}
-\ControlFlowTok{for}\NormalTok{(ayear }\ControlFlowTok{in} \KeywordTok{c}\NormalTok{(}\DecValTok{1979}\OperatorTok{:}\DecValTok{1994}\NormalTok{, }\DecValTok{1996}\NormalTok{, }\DecValTok{1998}\NormalTok{, }\DecValTok{2000}\NormalTok{, }\DecValTok{2002}\NormalTok{, }\DecValTok{2004}\NormalTok{, }\DecValTok{2006}\NormalTok{, }\DecValTok{2008}\NormalTok{, }\DecValTok{2010}\NormalTok{,}
+\NormalTok{rates }\OtherTok{\textless{}{-}} \FunctionTok{list}\NormalTok{()}
+\ControlFlowTok{for}\NormalTok{(ayear }\ControlFlowTok{in} \FunctionTok{c}\NormalTok{(}\DecValTok{1979}\SpecialCharTok{:}\DecValTok{1994}\NormalTok{, }\DecValTok{1996}\NormalTok{, }\DecValTok{1998}\NormalTok{, }\DecValTok{2000}\NormalTok{, }\DecValTok{2002}\NormalTok{, }\DecValTok{2004}\NormalTok{, }\DecValTok{2006}\NormalTok{, }\DecValTok{2008}\NormalTok{, }\DecValTok{2010}\NormalTok{,}
                \DecValTok{2012}\NormalTok{, }\DecValTok{2014}\NormalTok{, }\DecValTok{2016}\NormalTok{, }\DecValTok{2018}\NormalTok{)) \{}
-\NormalTok{  rates[[ayear]] <-}\StringTok{ }\KeywordTok{get_rate}\NormalTok{(ayear)}
+\NormalTok{  rates[[ayear]] }\OtherTok{\textless{}{-}} \FunctionTok{get\_rate}\NormalTok{(ayear)}
 \NormalTok{\}}
-\NormalTok{rates_all <-}\StringTok{ }\KeywordTok{bind_rows}\NormalTok{(}\OperatorTok{!!!}\NormalTok{rates)}
-\CommentTok{# join hours and rates variable}
-\NormalTok{hours_wages <-}\StringTok{ }\KeywordTok{left_join}\NormalTok{(rates_all,}
-\NormalTok{                         hours_all,}
-                         \DataTypeTok{by =} \KeywordTok{c}\NormalTok{(}\StringTok{"id"}\NormalTok{, }\StringTok{"year"}\NormalTok{, }\StringTok{"job"}\NormalTok{)) }\OperatorTok{%>%}
-\StringTok{  }\CommentTok{# set the 0 value in rate_per_hour as NA}
-\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{rate_per_hour =} \KeywordTok{ifelse}\NormalTok{(rate_per_hour }\OperatorTok{==}\StringTok{ }\DecValTok{0}\NormalTok{, }\OtherTok{NA}\NormalTok{,}
-\NormalTok{                                rate_per_hour),}
-         \DataTypeTok{hours_work =} \KeywordTok{ifelse}\NormalTok{(hours_work }\OperatorTok{==}\StringTok{ }\DecValTok{0} \OperatorTok{|}\StringTok{ }\NormalTok{hours_work }\OperatorTok{>}\StringTok{ }\DecValTok{84}\NormalTok{, }\OtherTok{NA}\NormalTok{, hours_work))}
+\NormalTok{rates\_all }\OtherTok{\textless{}{-}} \FunctionTok{bind\_rows}\NormalTok{(}\SpecialCharTok{!!!}\NormalTok{rates)}
+\CommentTok{\# join hours and rates variable}
+\NormalTok{hours\_wages }\OtherTok{\textless{}{-}} \FunctionTok{left\_join}\NormalTok{(rates\_all,}
+\NormalTok{                         hours\_all,}
+                         \AttributeTok{by =} \FunctionTok{c}\NormalTok{(}\StringTok{"id"}\NormalTok{, }\StringTok{"year"}\NormalTok{, }\StringTok{"job"}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%}
+  \CommentTok{\# set the 0 value in rate\_per\_hour as NA}
+  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{rate\_per\_hour =} \FunctionTok{ifelse}\NormalTok{(rate\_per\_hour }\SpecialCharTok{==} \DecValTok{0}\NormalTok{, }\ConstantTok{NA}\NormalTok{,}
+\NormalTok{                                rate\_per\_hour),}
+         \AttributeTok{hours\_work =} \FunctionTok{ifelse}\NormalTok{(hours\_work }\SpecialCharTok{==} \DecValTok{0} \SpecialCharTok{|}\NormalTok{ hours\_work }\SpecialCharTok{\textgreater{}} \DecValTok{84}\NormalTok{, }\ConstantTok{NA}\NormalTok{, hours\_work))}
 \end{Highlighting}
@@
 
-According to Huebner, Vach, and Cessie (2016), Initial Data Analysis (IDA) is the step of inspecting and screening the data after being collected to ensure that the data is clean, valid, and ready to be deployed in the later formal statistical analysis.
+According to Huebner Marianne, Vach, and Cessie (2016), Initial Data Analysis (IDA) is the step of inspecting and screening the data after being collected to ensure that the data is clean, valid, and ready to be deployed in the later formal statistical analysis.
 Moreover, Chatfield (1985) argues that the two main objectives of IDA are data description, which is to assess the structure and the quality of the data, and model formulation without any formal statistical inference.
@@
 
+\begin{figure}
+
+{\centering \includegraphics[width=0.7\linewidth]{figures/sample-plot-1}
+
+}
+
+\caption{Getting a glimpse of the data: random sample of 36 individuals, divided into  twelve subplots. Wages are shown against year. There is a lot of variability in wages.}\label{fig:sample-plot}
+\end{figure}
+
 However, we can not be sure only by observing the summary statistics.
@@
 
-\caption{Two plots showing the distribution of the mean hourly wage.
+\caption{Several plots showing anomalies observed the input data.
+XX Add a couple more plots that reveal the initial problems in the data.
 Plot A portrays the pattern of mean hourly wage of high school cohort from 1979 to 2018 of each ID in US Dollar; Plot B shows the distribution of their minimum, median, and maximum value.
@@
 Some have had flat wages for years but had a sudden increase in one particular year, then it gone down again, while the others experienced an upsurge in their wage, for instance, the IDs in panel 9.
-
-\begin{figure}
-
-{\centering \includegraphics[width=0.7\linewidth]{figures/sample-plot-1}
-
-}
-
-\caption{The mean hourly wages of some random samples are shown in twelve facets, three IDs per facet. It suggests that some IDs had a reasonably fluctuate wages.}\label{fig:sample-plot}
-\end{figure}
 
@@
 However, this method is too computationally and memory expensive, especially for a large data set, like the NLSY79 data.
-Thus, the model for each ID is built utilizing the \texttt{nest} and \texttt{map} function from \texttt{tidyr} (H.
-Wickham 2020b) and \texttt{purrr} (Henry and Wickham 2020), respectively.
+Thus, the model for each ID is built utilizing the \texttt{nest} and \texttt{map} function from \texttt{tidyr} (Wickham 2020b) and \texttt{purrr} (Henry and Wickham 2020), respectively.
 
@@
 
-To minimize the risk of mistakenly identifying an outlier as an ``erroneous outlier'', we simulate some thresholds and study how they affect the data.
+To minimize the risk of mistakenly identifying an outlier as an ``erroneous outlier,'' we simulate some thresholds and study how they affect the data.
 We find that 0.12 is the most reasonable value to be the threshold to minimize that drawback's risk because it still captures the sensible spikes in the data.

### 2021-09-14 paper/scripts/ida.R

@@ -30,34 +30,6 @@ kable(as.array(summary(wages_demog_hs$mean_hourly_wage)),
       col.names = c("Statistics", "Value")) %>%
   kable_styling()
 
-## ---- feature-plot
-spag <- wages_demog_hs %>%
-  ggplot(aes(x = year,
-             y = mean_hourly_wage,
-             group = id)) +
-  geom_line(alpha = 0.1) +
-  ggtitle("A)") +
-  theme(plot.title = element_text(size = 10)) +
-  theme_bw()
-
-wages_demog_hs_tsibble <- as_tsibble(x = wages_demog_hs,
-                                     key = id,
-                                     index = year,
-                                     regular = FALSE)
-wages_three_feat <- wages_demog_hs_tsibble %>%
-  features(mean_hourly_wage,
-           feat_three_num
-  )
-wages_feat_long <- wages_three_feat %>%
-  pivot_longer(c(min, med, max), names_to = "feature", values_to = "value")
-feature <- ggplot(wages_feat_long) +
-  geom_density(aes(x = value, colour = feature, fill = feature), alpha = 0.3) +
-  ggtitle("B)") +
-  theme(plot.title = element_text(size = 10))+
-  theme_bw()
-
-spag + feature
-
 ## ---- high-wages
 wages_high <- filter(wages_demog_hs, mean_hourly_wage > 500) %>%
   as_tibble() %>%
@@ -79,6 +51,11 @@ ggplot(wages_high2) +
   theme_bw()
 
 ## ---- sample-plot
+wages_demog_hs_tsibble <- as_tsibble(x = wages_demog_hs,
+                                     key = id,
+                                     index = year,
+                                     regular = FALSE)
+
 set.seed(20210225)
 ggplot(wages_demog_hs_tsibble,
        aes(x = year,
@@ -89,6 +66,30 @@ ggplot(wages_demog_hs_tsibble,
   theme(axis.text.x = element_text(angle = 10, size = 6)) +
   ylab("mean hourly wage") +
   theme_bw()
+
+## ---- feature-plot
+spag <- wages_demog_hs %>%
+  ggplot(aes(x = year,
+             y = mean_hourly_wage,
+             group = id)) +
+  geom_line(alpha = 0.1) +
+  ggtitle("A)") +
+  theme(plot.title = element_text(size = 10)) +
+  theme_bw()
+
+wages_three_feat <- wages_demog_hs_tsibble %>%
[…]

### 2021-09-14 paper/paper.Rmd
sp

@@
 
-The entire analysis is conducted using `R` [@R] in `rstudio` using these packages: `tidyverse` [@tidyverse], `ggplot2` [@ggplot2], `dplyr` [@dplyr], `readr` [@readr], `tidyr` [@tidyr], `stringr` [@stringr], `purrr` [@purrr], `broom` [@broom], `blorgar` [@brolgar], `patchwork` [@patchwork], `kableExtra` [@kableExtra], `MASS` [@mass], `janitor` [@janitor], `DiagrammeR` [@DiagrammeR], `rsvg` [@rsvg], `webshot` [@webshot], `mgcv` [@mgcv], `tsibble` [@tsibble], and `modelr` [@modelr].
+The entire analysis is conducted using `R` [@R] in `rstudio` using these packages: `tidyverse` [@tidyverse], `ggplot2` [@ggplot2], `dplyr` [@dplyr], `readr` [@readr], `tidyr` [@tidyr], `stringr` [@stringr], `purrr` [@purrr], `broom` [@broom], `brolgar` [@brolgar], `patchwork` [@patchwork], `kableExtra` [@kableExtra], `MASS` [@mass], `janitor` [@janitor], `DiagrammeR` [@DiagrammeR], `rsvg` [@rsvg], `webshot` [@webshot], `mgcv` [@mgcv], `tsibble` [@tsibble], and `modelr` [@modelr].
 The paper are generated using `knitr` [@knitr], `rmarkdown` [@rmarkdown], and `rticles` [@rticles].

### 2021-09-14 paper/paper.tex
sp

@@
 
-\usepackage{booktabs}
-\usepackage{longtable}
-\usepackage{array}
-\usepackage{multirow}
-\usepackage{wrapfig}
-\usepackage{float}
-\usepackage{colortbl}
-\usepackage{pdflscape}
-\usepackage{tabu}
-\usepackage{threeparttable}
-\usepackage{threeparttablex}
-\usepackage[normalem]{ulem}
-\usepackage{makecell}
-\usepackage{xcolor}
-
 \begin{document}
@@
 2019), \texttt{ggplot2} (Wickham 2016), \texttt{dplyr} (Wickham et al.
-2020), \texttt{readr} (Wickham and Hester 2020), \texttt{tidyr} (Wickham 2020b), \texttt{stringr} (Wickham 2019), \texttt{purrr} (Henry and Wickham 2020), \texttt{broom} (Robinson, Hayes, and Couch 2021), \texttt{blorgar} (Tierney, Cook, and Prvan 2020), \texttt{patchwork} (Pedersen 2020), \texttt{kableExtra} (Zhu 2019), \texttt{MASS} (Venables and Ripley 2002), \texttt{janitor} (Firke 2020), \texttt{DiagrammeR} (Iannone 2020), \texttt{rsvg} (Ooms 2020), \texttt{webshot} (Chang 2019), \texttt{mgcv} (Wood 2003), \texttt{tsibble} (Wang, Cook, and Hyndman 2020), and \texttt{modelr} (Wickham 2020a).
+2020), \texttt{readr} (Wickham and Hester 2020), \texttt{tidyr} (Wickham 2020b), \texttt{stringr} (Wickham 2019), \texttt{purrr} (Henry and Wickham 2020), \texttt{broom} (Robinson, Hayes, and Couch 2021), \texttt{brolgar} (Tierney, Cook, and Prvan 2020), \texttt{patchwork} (Pedersen 2020), \texttt{kableExtra} (Zhu 2019), \texttt{MASS} (Venables and Ripley 2002), \texttt{janitor} (Firke 2020), \texttt{DiagrammeR} (Iannone 2020), \texttt{rsvg} (Ooms 2020), \texttt{webshot} (Chang 2019), \texttt{mgcv} (Wood 2003), \texttt{tsibble} (Wang, Cook, and Hyndman 2020), and \texttt{modelr} (Wickham 2020a).
 The paper are generated using \texttt{knitr} (Xie 2014), \texttt{rmarkdown} (Xie, Dervieux, and Riederer 2020), and \texttt{rticles} (Allaire et al.

### 2021-09-14 paper/paper.Rmd

--- 2021-09-14 paper/paper.Rmd
@@
 
-# Data Cleaning {#cleaning}
+# The NLSY79 Data Cleaning {#cleaning}
 
@@
 
-```{r sample-plot, echo = FALSE, fig.cap="Getting a glimpse of the data: random sample of 36 individuals, divided into  twelve subplots.
-Wages are shown against year.
-There is a lot of variability in wages.", fig.height=4, fig.width=5}
-```
-
 However, we can not be sure only by observing the summary statistics.
@@
 
-```{r feature-plot, fig.cap = "Several plots showing anomalies observed the input data.
-XX Add a couple more plots that reveal the initial problems in the data.
+```{r feature-plot, fig.cap = "Two plots showing the distribution of the mean hourly wage.
 Plot A portrays the pattern of mean hourly wage of high school cohort from 1979 to 2018 of each ID in US Dollar; Plot B shows the distribution of their minimum, median, and maximum value.
@@
 Some have had flat wages for years but had a sudden increase in one particular year, then it gone down again, while the others experienced an upsurge in their wage, for instance, the IDs in panel 9.
+
+```{r sample-plot, echo = FALSE, fig.cap="The mean hourly wages of some random samples are shown in twelve facets, three IDs per facet.
+It suggests that some IDs had a reasonably fluctuate wages.", fig.height=4, fig.width=5}
+```
 

### 2021-09-14 paper/paper.tex

--- 2021-09-14 paper/paper.tex
@@
 \hypertarget{cleaning}{%
-\section{Data Cleaning}\label{cleaning}}
+\section{The NLSY79 Data Cleaning}\label{cleaning}}
 
@@
 
-\begin{figure}
-
-{\centering \includegraphics[width=0.7\linewidth]{figures/sample-plot-1}
-
-}
-
-\caption{Getting a glimpse of the data: random sample of 36 individuals, divided into  twelve subplots. Wages are shown against year. There is a lot of variability in wages.}\label{fig:sample-plot}
-\end{figure}
-
 However, we can not be sure only by observing the summary statistics.
@@
 
-\caption{Several plots showing anomalies observed the input data.
-XX Add a couple more plots that reveal the initial problems in the data.
+\caption{Two plots showing the distribution of the mean hourly wage.
 Plot A portrays the pattern of mean hourly wage of high school cohort from 1979 to 2018 of each ID in US Dollar; Plot B shows the distribution of their minimum, median, and maximum value.
@@
 Some have had flat wages for years but had a sudden increase in one particular year, then it gone down again, while the others experienced an upsurge in their wage, for instance, the IDs in panel 9.
+
+\begin{figure}
+
+{\centering \includegraphics[width=0.7\linewidth]{figures/sample-plot-1}
+
+}
+
+\caption{The mean hourly wages of some random samples are shown in twelve facets, three IDs per facet. It suggests that some IDs had a reasonably fluctuate wages.}\label{fig:sample-plot}
+\end{figure}
 

### 2021-09-14 paper/scripts/ida.R
Change yowie::demog_nlsy79 -> full_demographic.

@@ -1,6 +1,6 @@
 ## ---- age-table
-age_table <- yowie::demog_nlsy79 %>%
-group_by(age_1979) %>%
+age_table <- full_demographics %>%
+  group_by(age_1979) %>%
   count(age_1979)
 
 kable(age_table,
@@ -9,7 +9,7 @@ kable(age_table,
   kable_styling()
 
 ## ---- gender-race-table
-gender_race_table <- yowie::demog_nlsy79 %>%
+gender_race_table <- full_demographics %>%
   tabyl(gender, race) %>%
   adorn_totals(c("row", "col")) %>%
   adorn_percentages("row") %>%
@@ -235,7 +235,6 @@ wages <- wages_demog_hs %>%
 
 # Create a data set for demographic variables
 demog_nlsy79 <- full_demographics %>%
-  mutate(age_1979 = 1979 - (dob_year + 1900)) %>%
   dplyr::select(id,
                 age_1979,
                 gender,

### 2021-09-14 paper/scripts/data-preprocessing.R
Add age_1979 to data preprocessing.

@@ -118,7 +118,8 @@ highest_year <- demog_education %>%
 ## ---- full-demog
 full_demographics <- full_join(dob_tidy, demog_tidy, by = "CASEID_1979") %>%
   full_join(highest_year, by = "CASEID_1979") %>%
-  rename("id" = "CASEID_1979")
+  rename("id" = "CASEID_1979") %>%
+  mutate(age_1979 = 1979 - (dob_year + 1900))
 
 ## ---- tidy-hours
 # make a list for years where we used the "QES-52A"

### 2021-09-14 paper/scripts/ida.R
Add plots to feature-plot chunck

@@ -38,17 +38,18 @@ wages_high <- filter(wages_demog_hs, mean_hourly_wage > 500) %>%
 wages_high2 <- wages_demog_hs %>%
   filter(id %in% wages_high$id)
 
-ggplot(wages_high2) +
+plot_high <- ggplot(filter(wages_high2, id == 39)) +
   geom_line(aes(x = year,
                 y = mean_hourly_wage)) +
   geom_point(aes(x = year,
                  y = mean_hourly_wage),
              size = 0.5,
              alpha = 0.5) +
-  facet_wrap(~id, scales = "free_y") +
-  theme(axis.text.x = element_text(angle = 10, size = 6)) +
+  theme(axis.text.x = element_text(angle = 10, size = 6),
+        plot.title = element_text(size = 10)) +
   ylab("mean hourly wage") +
-  theme_bw()
+  theme_bw() +
+  ggtitle("D)")
 
 ## ---- sample-plot
 wages_demog_hs_tsibble <- as_tsibble(x = wages_demog_hs,
@@ -75,7 +76,8 @@ spag <- wages_demog_hs %>%
   geom_line(alpha = 0.1) +
   ggtitle("A)") +
   theme(plot.title = element_text(size = 10)) +
-  theme_bw()
+  theme_bw() +
+  ylab("mean hourly wage")
 
 wages_three_feat <- wages_demog_hs_tsibble %>%
   features(mean_hourly_wage,
@@ -86,10 +88,18 @@ wages_feat_long <- wages_three_feat %>%
 feature <- ggplot(wages_feat_long) +
   geom_density(aes(x = value, colour = feature, fill = feature), alpha = 0.3) +
   ggtitle("B)") +
-  theme(plot.title = element_text(size = 10))+
-  theme_bw()
+  theme_bw() +
+  theme(plot.title = element_text(size = 10))
 
-spag + feature
+feature_bp <- ggplot(wages_feat_long, aes(y=value, fill = feature, color = feature)) +
+  geom_boxplot() +
+  theme_bw() +
+  ggtitle("C)") +
+  theme(legend.position = "none",
+        plot.title = element_text(size = 10))
+
+spag + feature + feature_bp + plot_high + plot_layout(nrow = 1, guides = "collect") &
+  theme(legend.position = "bottom")
 
 ## ---- rlm
 # nest the data by id to build a robust linear model
[…]

### 2021-09-16 paper/paper-jsdse/paper-jsdse.Rmd
html + jsdse template

---
title: "The Journey from Wild to Textbook Data: A Case Study from the National Longitudinal Survey of Youth"
authors:
  - name: Dewi Amaliah
    thanks: ""
    department: Department of Econometrics and Business Statistics
    affiliation: Monash University
    location: Clayton, VIC 3800
    email: dama0007@student.monash.edu
  - name: Dianne Cook
    department: Department of Econometrics and Business Statistics
    affiliation: Monash University
    location: Clayton, VIC 3800
    email: dicook@monash.edu
  - name: Emi Tanaka
    department: Department of Econometrics and Business Statistics
    affiliation: Monash University
    location: Clayton, VIC 3800
    email: emi.tanaka@monash.edu
  - name: Kate Hyde
    thanks: ""
    department: Department of Econometrics and Business Statistics
    affiliation: Monash University
    location: Clayton, VIC 3800
    email: ""
  - name: Nicholas Tierney
    thanks: ""
    department: ""
    affiliation: ""
    location: ""
    email: ""
bibliography: ../references.bib
biblio-style: unsrt
#preamble: >
output: 
  bookdown::pdf_document2:
    template: TAS-template-for-rmd.tex
    keep_tex: true
date: "`r Sys.Date()`"
always_allow_html: true
---

```{r, child = "../paper.Rmd"}
```

### 2021-09-16 paper/paper-jsdse/paper-jsdse.tex
html + jsdse template

--- 2021-09-14 paper/paper.tex
@@
-\documentclass{article}
-
-\usepackage{arxiv}
-
-\usepackage[utf8]{inputenc} % allow utf-8 input
-\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
-\usepackage{lmodern}        % https://github.com/rstudio/rticles/issues/343
-\usepackage{hyperref}       % hyperlinks
-\usepackage{url}            % simple URL typesetting
-\usepackage{booktabs}       % professional-quality tables
-\usepackage{amsfonts}       % blackboard math symbols
-\usepackage{nicefrac}       % compact symbols for 1/2, etc.
-\usepackage{microtype}      % microtypography
-\usepackage{lipsum}
-\usepackage{graphicx}
-
-\title{The Journey from Wild to Textbook Data: A Case Study from the National Longitudinal Survey of Youth}
-
-\author{
-    Dewi Amaliah
-   \\
-    Department of Econometrics and Business Statistics \\
-    Monash University \\
-  Clayton, VIC 3800 \\
-  \texttt{\href{mailto:dama0007@student.monash.edu}{\nolinkurl{dama0007@student.monash.edu}}} \\
-   \And
-    Dianne Cook
-   \\
-    Department of Econometrics and Business Statistics \\
-    Monash University \\
-  Clayton, VIC 3800 \\
-  \texttt{\href{mailto:dicook@monash.edu}{\nolinkurl{dicook@monash.edu}}} \\
-   \And
-    Emi Tanaka
-   \\
-    Department of Econometrics and Business Statistics \\
-    Monash University \\
-  Clayton, VIC 3800 \\
-  \texttt{\href{mailto:emi.tanaka@monash.edu}{\nolinkurl{emi.tanaka@monash.edu}}} \\
-   \And
-    Kate Hyde
-   \\
-    Department of Econometrics and Business Statistics \\
-    Monash University \\
-  Clayton, VIC 3800 \\
-  \texttt{} \\
-   \And
-    Nicholas Tierney
-   \\
-     \\
-   \\
-  \texttt{} \\
-  }
+\documentclass[12pt]{article}
+\usepackage{amsmath}
+\usepackage{graphicx,psfrag,epsf}
+\usepackage{enumerate}
+\usepackage{natbib}
+\usepackage{url} % not crucial - just used below for the URL
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
+
+%\pdfminorversion=4
+% NOTE: To produce blinded version, replace "0" with "1" below.
+\newcommand{\blind}{0}
+
+% need this for pandoc %
+\newlength{\cslhangindent}
+\setlength{\cslhangindent}{1.5em}
+\newlength{\csllabelwidth}
+\setlength{\csllabelwidth}{3em}
+\newenvironment{CSLReferences}[3] % #1 hanging-ident, #2 entry spacing
+ {% don't indent paragraphs
+  \setlength{\parindent}{0pt}
+  % turn on hanging indent if param 1 is 1
+  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
+  % set entry spacing
+  \ifnum #2 > 0
+  \setlength{\parskip}{#2\baselineskip}
+  \fi
+ }%
+ {}
+\usepackage{calc} % for \widthof, \maxof
+\newcommand{\CSLBlock}[1]{#1\hfill\break}
+\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\maxof{\widthof{#1}}{\csllabelwidth}}{#1}}
+\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth}{#1}}
+\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
+
+% DON'T change margins - should be 1 inch all around.
+\addtolength{\oddsidemargin}{-.5in}%
+\addtolength{\evensidemargin}{-.5in}%
+\addtolength{\textwidth}{1in}%
+\addtolength{\textheight}{1.3in}%
+\addtolength{\topmargin}{-.8in}%
+
+%-------------------------------------------------------------------
+% R MARKDOWN
+%-------------------------------------------------------------------
+\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
+
+\def\spacingset#1{\renewcommand{\baselinestretch}%
+{#1}\small\normalsize} \spacingset{1}
 
@@
 
-% Pandoc citation processing
-\newlength{\csllabelwidth}
-\setlength{\csllabelwidth}{3em}
-\newlength{\cslhangindent}
-\setlength{\cslhangindent}{1.5em}
-% for Pandoc 2.8 to 2.10.1
-\newenvironment{cslreferences}%
-  {}%
-  {\par}
-% For Pandoc 2.11+
-\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
- {% don't indent paragraphs
-  \setlength{\parindent}{0pt}
-  % turn on hanging indent if param 1 is 1
-  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
-  % set entry spacing
-  \ifnum #2 > 0
-  \setlength{\parskip}{#2\baselineskip}
-  \fi
- }%
- {}
-\usepackage{calc} % for calculating minipage widths
-\newcommand{\CSLBlock}[1]{#1\hfill\break}
-\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
-\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
-\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
-
 \begin{document}
+
+%\bibliographystyle{natbib}
+
+\def\spacingset#1{\renewcommand{\baselinestretch}%
+{#1}\small\normalsize} \spacingset{1}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\if0\blind
+{
+  \title{\bf The Journey from Wild to Textbook Data: A Case Study from the National Longitudinal Survey of Youth}
+  \author{Author 1\thanks{
+    The authors gratefully acknowledge \textit{please remember to list all relevant funding sources in the unblinded version}}\hspace{.2cm}\\
+    Department of YYY, University of XXX\\
+    and \\
+    Author 2 \\
+    Department of ZZZ, University of WWW}
   \maketitle
-
-\def\tightlist{}
-
+} \fi
+
+\if1\blind
+{
+  \bigskip
+  \bigskip
+  \bigskip
+  \begin{center}
+    {\LARGE\bf Title}
+\end{center}
+  \medskip
+} \fi
+
+\bigskip
 \begin{abstract}
@@
 \end{abstract}
+
+\noindent%
+{\it Keywords:}  3 to 6 keywords, that do not appear in the title
+\vfill
+
+\newpage
+\spacingset{1.45} % DON'T change the spacing!
 
@@
 
-\bibliographystyle{unsrt}
-\bibliography{references.bib}
+\bibliographystyle{Chicago}
 

### 2021-09-16 paper/paper.Rmd
Edit feature plot.

@@
 
-The next step is that explore the mean hourly wage data.
-In this case, we only explore the wages data of up to $12^{th}$ grade.
-Table \ref{tab:summarytable} shows that the median of the overall wage of the cohort is only 7.2, while the mean is 11.87.
-It indicates that the data might contains a lot of extreme values.
-
-```{r summarytable}
-```
+The next step is that explore the mean hourly wage data. In this case, we only explore the wages data of up to $12^{th}$ grade.  Here, we use visualisation techniques to do IDA.
 
@@
 
-However, we can not be sure only by observing the summary statistics.
-Hence, we use visualisation techniques to investigate this matter.
+We take 36 samples randomly from the data and plot them, as shown in Figure \ref{fig:sample-plot}.
+It implies that these respondents have a lot of variability in wages, for example, the IDs in panel numbers 5, 7, and 11.
+The plot also implies that the samples have a different pattern of mean hourly wages.
+Some have had flat wages for years but had a sudden increase in one particular year, then it gone down again, while the others experienced an upsurge in their wage, for instance, the IDs in panel 9.
+
+Further, we investigate the summary statistics of wages. 
 Figure \ref{fig:feature-plot} conveys a problem in the mean hourly wage values.
@@
 In Figure \ref{fig:feature-plot} B, we barely see any difference in the minimum, median, and maximum value of the wages since the distribution is heavily skewed to the right.
+The problem is more clearly conveyed in Figure \ref{fig:feature-plot} C where there is no extreme values in the minimum and median values of wages but the maximum values show some anomalies.
+In Figure \ref{fig:feature-plot} D, we show the example of anomaly in one observation.
+This respondent only experienced one point of extremely high wages.
+Thus, we suspect that these high values are erroneous resulted from a data entry error.
+
+```{r summarytable, eval = FALSE}
+```
+
+```{r high-wages}
+```
 
 ```{r feature-plot, fig.cap = "Several plots showing anomalies observed the input data.
-XX Add a couple more plots that reveal the initial problems in the data.
-Plot A portrays the pattern of mean hourly wage of high school cohort from 1979 to 2018 of each ID in US Dollar; Plot B shows the distribution of their minimum, median, and maximum value.
-We can see that some IDs had an extremely high of wages and it made the distribution of the three features is extremely skewed.", fig.width=6, fig.height=4}
+Plot A portrays the pattern of mean hourly wage of high school cohort from 1979 to 2018 of each ID in US Dollar.
+We can see some extremely high wages; Plot B shows the distribution of their minimum, median, and maximum value.
+We can also see that some IDs had an extremely high of wages and it made the distribution of the three features is extremely skewed; Plot C shows the extreme values more clearly; Plot D shows an ID with a suspected erronous wages.", out.width="90%"}
 ```
-
-In Figure \ref{fig:high-wages}, we plot some respondents with a high value of mean hourly wages.
-We filter all of the IDs who earned more than US$ 500 per hour on average.
-We find that these respondents only experienced one point of extremely high wages.
-Thus, we suspect that these high values are erroneous resulted from a data entry error.
-
-```{r high-wages, fig.cap= "6 out of 45 IDs with extremely high mean hourly wage. Most of the IDs only have one point of high wage."}
-```
-
-Further, we take 36 samples randomly from the data and plot them, as shown in Figure \ref{fig:sample-plot}.
-It implies that not only that some observations earn extremely high figures of wages, but some also have reasonably fluctuated wages, for example, the IDs in panel numbers 5, 7, and 11.
-The plot also implies that the samples have a different pattern of mean hourly wages.
-Some have had flat wages for years but had a sudden increase in one particular year, then it gone down again, while the others experienced an upsurge in their wage, for instance, the IDs in panel 9.
 

### 2021-09-16 paper/paper.tex
Edit feature plot.

--- 2021-09-14 paper/paper.tex
@@
 % Pandoc citation processing
-\newlength{\csllabelwidth}
-\setlength{\csllabelwidth}{3em}
-\newlength{\cslhangindent}
-\setlength{\cslhangindent}{1.5em}
-% for Pandoc 2.8 to 2.10.1
-\newenvironment{cslreferences}%
-  {}%
-  {\par}
-% For Pandoc 2.11+
-\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
- {% don't indent paragraphs
-  \setlength{\parindent}{0pt}
-  % turn on hanging indent if param 1 is 1
-  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
-  % set entry spacing
-  \ifnum #2 > 0
-  \setlength{\parskip}{#2\baselineskip}
-  \fi
- }%
- {}
-\usepackage{calc} % for calculating minipage widths
-\newcommand{\CSLBlock}[1]{#1\hfill\break}
-\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
-\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
-\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
 
@@
 However, open data is often referred to as what we might call ``wild data'' because it requires substantial cleaning and tidying to tame it into textbook shape.
-Huebner Marianne, Vach, and Cessie (2016) emphasize that making the data cleaning process accountable and transparent is imperative.
-Documenting the data cleaning is essential for the integrity of downstream statistical analyses and model building (M.
-Huebner et al.
+Huebner, Vach, and Cessie (2016) emphasize that making the data cleaning process accountable and transparent is imperative.
+Documenting the data cleaning is essential for the integrity of downstream statistical analyses and model building (Huebner et al.
 2020).
@@
 There are few research papers that document the data cleaning (Wickham 2014).
-Furthermore, the decisions made in this stage often go unreported in the sense that IDA is often performed in an unplanned and unstructured way and is only shared among restricted parties (M.
-Huebner et al.
+Furthermore, the decisions made in this stage often go unreported in the sense that IDA is often performed in an unplanned and unstructured way and is only shared among restricted parties (Huebner et al.
 2020).
@@
 Consequently, the data should be tidied and wrangled first to extract the demographic and employment variables we want to put in the final data set.
-We mainly used \texttt{tidyr} (Wickham 2020b), \texttt{dplyr} (Wickham et al.
-2020), and \texttt{stringr} (Wickham 2019) to do this job.
+We mainly used \texttt{tidyr} (H.
+Wickham 2020b), \texttt{dplyr} (H.
+Wickham, François, et al.
+2020), and \texttt{stringr} (H.
+Wickham 2019) to do this job.
 
@@
 \begin{Highlighting}[]
-\CommentTok{\#tidy the date of birth data}
-\NormalTok{dob\_tidy }\OtherTok{\textless{}{-}}\NormalTok{ new\_data\_qnames }\SpecialCharTok{\%\textgreater{}\%}
-\NormalTok{  dplyr}\SpecialCharTok{::}\FunctionTok{select}\NormalTok{(CASEID\_1979,}
-                \FunctionTok{starts\_with}\NormalTok{(}\StringTok{"Q1{-}3\_A\textasciitilde{}"}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{dob\_year =} \FunctionTok{case\_when}\NormalTok{(}
-    \CommentTok{\# if the years recorded in both sets match, take 79 data}
-    \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}Y\_1979}\StringTok{\textasciigrave{}} \SpecialCharTok{==} \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}Y\_1981}\StringTok{\textasciigrave{}} \SpecialCharTok{\textasciitilde{}} \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}Y\_1979}\StringTok{\textasciigrave{}}\NormalTok{,}
-    \CommentTok{\# if the year in the 81 set is missing, take the 79 data}
-    \FunctionTok{is.na}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}Y\_1981}\StringTok{\textasciigrave{}}\NormalTok{) }\SpecialCharTok{\textasciitilde{}} \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}Y\_1979}\StringTok{\textasciigrave{}}\NormalTok{,}
-    \CommentTok{\# if the sets don\textquotesingle{}t match for dob year, take the 79 data}
-    \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}Y\_1979}\StringTok{\textasciigrave{}} \SpecialCharTok{!=} \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}Y\_1981}\StringTok{\textasciigrave{}} \SpecialCharTok{\textasciitilde{}} \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}Y\_1979}\StringTok{\textasciigrave{}}\NormalTok{),}
-    \AttributeTok{dob\_month =} \FunctionTok{case\_when}\NormalTok{(}
-      \CommentTok{\# if months recorded in both sets match, take 79 data}
-      \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}M\_1979}\StringTok{\textasciigrave{}} \SpecialCharTok{==} \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}M\_1981}\StringTok{\textasciigrave{}} \SpecialCharTok{\textasciitilde{}} \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}M\_1979}\StringTok{\textasciigrave{}}\NormalTok{,}
-      \CommentTok{\# if month in 81 set is missing, take the 79 data}
-      \FunctionTok{is.na}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}M\_1981}\StringTok{\textasciigrave{}}\NormalTok{) }\SpecialCharTok{\textasciitilde{}} \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}M\_1979}\StringTok{\textasciigrave{}}\NormalTok{,}
-      \CommentTok{\# if sets don\textquotesingle{}t match for dob month, take the 79 data}
-      \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}M\_1979}\StringTok{\textasciigrave{}} \SpecialCharTok{!=} \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}M\_1981}\StringTok{\textasciigrave{}} \SpecialCharTok{\textasciitilde{}} \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}M\_1979}\StringTok{\textasciigrave{}}\NormalTok{),}
-    \CommentTok{\# flag if there is a conflict between dob recorded in 79 and 81}
-    \AttributeTok{dob\_conflict =} \FunctionTok{case\_when}\NormalTok{(}
-\NormalTok{      (}\StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}M\_1979}\StringTok{\textasciigrave{}} \SpecialCharTok{!=} \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}M\_1981}\StringTok{\textasciigrave{}}\NormalTok{) }\SpecialCharTok{\&} \SpecialCharTok{!}\FunctionTok{is.na}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}M\_1981}\StringTok{\textasciigrave{}}\NormalTok{)}
-      \SpecialCharTok{\textasciitilde{}} \ConstantTok{TRUE}\NormalTok{,}
-\NormalTok{      (}\StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}Y\_1979}\StringTok{\textasciigrave{}} \SpecialCharTok{!=} \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}Y\_1981}\StringTok{\textasciigrave{}}\NormalTok{) }\SpecialCharTok{\&} \SpecialCharTok{!}\FunctionTok{is.na}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}Y\_1981}\StringTok{\textasciigrave{}}\NormalTok{)}
-      \SpecialCharTok{\textasciitilde{}} \ConstantTok{TRUE}\NormalTok{,}
-\NormalTok{      (}\StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}Y\_1979}\StringTok{\textasciigrave{}} \SpecialCharTok{==} \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}Y\_1981}\StringTok{\textasciigrave{}}\NormalTok{) }\SpecialCharTok{\&}
-\NormalTok{        (}\StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}M\_1979}\StringTok{\textasciigrave{}} \SpecialCharTok{==} \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}M\_1981}\StringTok{\textasciigrave{}}\NormalTok{) }\SpecialCharTok{\textasciitilde{}} \ConstantTok{FALSE}\NormalTok{,}
-      \FunctionTok{is.na}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}M\_1981}\StringTok{\textasciigrave{}}\NormalTok{) }\SpecialCharTok{|} \FunctionTok{is.na}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}Y\_1981}\StringTok{\textasciigrave{}}\NormalTok{) }\SpecialCharTok{\textasciitilde{}} \ConstantTok{FALSE}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%}
-\NormalTok{  dplyr}\SpecialCharTok{::}\FunctionTok{select}\NormalTok{(CASEID\_1979,}
-\NormalTok{                dob\_month,}
-\NormalTok{                dob\_year,}
-\NormalTok{                dob\_conflict)}
+\CommentTok{#tidy the date of birth data}
+\NormalTok{dob_tidy <-}\StringTok{ }\NormalTok{new_data_qnames }\OperatorTok{%>%}
+\StringTok{  }\NormalTok{dplyr}\OperatorTok{::}\KeywordTok{select}\NormalTok{(CASEID_}\DecValTok{1979}\NormalTok{,}
+                \KeywordTok{starts_with}\NormalTok{(}\StringTok{"Q1-3_A~"}\NormalTok{)) }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{dob_year =} \KeywordTok{case_when}\NormalTok{(}
+    \CommentTok{# if the years recorded in both sets match, take 79 data}
+    \StringTok{`}\DataTypeTok{Q1-3_A~Y_1979}\StringTok{`} \OperatorTok{==}\StringTok{ `}\DataTypeTok{Q1-3_A~Y_1981}\StringTok{`} \OperatorTok{~}\StringTok{ `}\DataTypeTok{Q1-3_A~Y_1979}\StringTok{`}\NormalTok{,}
+    \CommentTok{# if the year in the 81 set is missing, take the 79 data}
+    \KeywordTok{is.na}\NormalTok{(}\StringTok{`}\DataTypeTok{Q1-3_A~Y_1981}\StringTok{`}\NormalTok{) }\OperatorTok{~}\StringTok{ `}\DataTypeTok{Q1-3_A~Y_1979}\StringTok{`}\NormalTok{,}
+    \CommentTok{# if the sets don't match for dob year, take the 79 data}
+    \StringTok{`}\DataTypeTok{Q1-3_A~Y_1979}\StringTok{`} \OperatorTok{!=}\StringTok{ `}\DataTypeTok{Q1-3_A~Y_1981}\StringTok{`} \OperatorTok{~}\StringTok{ `}\DataTypeTok{Q1-3_A~Y_1979}\StringTok{`}\NormalTok{),}
+    \DataTypeTok{dob_month =} \KeywordTok{case_when}\NormalTok{(}
+      \CommentTok{# if months recorded in both sets match, take 79 data}
+      \StringTok{`}\DataTypeTok{Q1-3_A~M_1979}\StringTok{`} \OperatorTok{==}\StringTok{ `}\DataTypeTok{Q1-3_A~M_1981}\StringTok{`} \OperatorTok{~}\StringTok{ `}\DataTypeTok{Q1-3_A~M_1979}\StringTok{`}\NormalTok{,}
+      \CommentTok{# if month in 81 set is missing, take the 79 data}
+      \KeywordTok{is.na}\NormalTok{(}\StringTok{`}\DataTypeTok{Q1-3_A~M_1981}\StringTok{`}\NormalTok{) }\OperatorTok{~}\StringTok{ `}\DataTypeTok{Q1-3_A~M_1979}\StringTok{`}\NormalTok{,}
+      \CommentTok{# if sets don't match for dob month, take the 79 data}
+      \StringTok{`}\DataTypeTok{Q1-3_A~M_1979}\StringTok{`} \OperatorTok{!=}\StringTok{ `}\DataTypeTok{Q1-3_A~M_1981}\StringTok{`} \OperatorTok{~}\StringTok{ `}\DataTypeTok{Q1-3_A~M_1979}\StringTok{`}\NormalTok{),}
+    \CommentTok{# flag if there is a conflict between dob recorded in 79 and 81}
+    \DataTypeTok{dob_conflict =} \KeywordTok{case_when}\NormalTok{(}
+\NormalTok{      (}\StringTok{`}\DataTypeTok{Q1-3_A~M_1979}\StringTok{`} \OperatorTok{!=}\StringTok{ `}\DataTypeTok{Q1-3_A~M_1981}\StringTok{`}\NormalTok{) }\OperatorTok{&}\StringTok{ }\OperatorTok{!}\KeywordTok{is.na}\NormalTok{(}\StringTok{`}\DataTypeTok{Q1-3_A~M_1981}\StringTok{`}\NormalTok{)}
+      \OperatorTok{~}\StringTok{ }\OtherTok{TRUE}\NormalTok{,}
+\NormalTok{      (}\StringTok{`}\DataTypeTok{Q1-3_A~Y_1979}\StringTok{`} \OperatorTok{!=}\StringTok{ `}\DataTypeTok{Q1-3_A~Y_1981}\StringTok{`}\NormalTok{) }\OperatorTok{&}\StringTok{ }\OperatorTok{!}\KeywordTok{is.na}\NormalTok{(}\StringTok{`}\DataTypeTok{Q1-3_A~Y_1981}\StringTok{`}\NormalTok{)}
+      \OperatorTok{~}\StringTok{ }\OtherTok{TRUE}\NormalTok{,}
+\NormalTok{      (}\StringTok{`}\DataTypeTok{Q1-3_A~Y_1979}\StringTok{`} \OperatorTok{==}\StringTok{ `}\DataTypeTok{Q1-3_A~Y_1981}\StringTok{`}\NormalTok{) }\OperatorTok{&}
+\StringTok{        }\NormalTok{(}\StringTok{`}\DataTypeTok{Q1-3_A~M_1979}\StringTok{`} \OperatorTok{==}\StringTok{ `}\DataTypeTok{Q1-3_A~M_1981}\StringTok{`}\NormalTok{) }\OperatorTok{~}\StringTok{ }\OtherTok{FALSE}\NormalTok{,}
+      \KeywordTok{is.na}\NormalTok{(}\StringTok{`}\DataTypeTok{Q1-3_A~M_1981}\StringTok{`}\NormalTok{) }\OperatorTok{|}\StringTok{ }\KeywordTok{is.na}\NormalTok{(}\StringTok{`}\DataTypeTok{Q1-3_A~Y_1981}\StringTok{`}\NormalTok{) }\OperatorTok{~}\StringTok{ }\OtherTok{FALSE}\NormalTok{)) }\OperatorTok{%>%}
+\StringTok{  }\NormalTok{dplyr}\OperatorTok{::}\KeywordTok{select}\NormalTok{(CASEID_}\DecValTok{1979}\NormalTok{,}
+\NormalTok{                dob_month,}
+\NormalTok{                dob_year,}
+\NormalTok{                dob_conflict)}
 \end{Highlighting}
@@
 \begin{Highlighting}[]
-\CommentTok{\# make a list for years where we used the "QES{-}52A"}
-\NormalTok{year\_A }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(}\DecValTok{1979}\SpecialCharTok{:}\DecValTok{1987}\NormalTok{, }\DecValTok{1993}\NormalTok{)}
-\CommentTok{\#function to get the hour of work}
-\NormalTok{get\_hour }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{(year)\{}
-  \ControlFlowTok{if}\NormalTok{(year }\SpecialCharTok{\%in\%}\NormalTok{ year\_A)\{}
-\NormalTok{    temp }\OtherTok{\textless{}{-}}\NormalTok{ new\_data\_qnames }\SpecialCharTok{\%\textgreater{}\%}
-\NormalTok{      dplyr}\SpecialCharTok{::}\FunctionTok{select}\NormalTok{(CASEID\_1979,}
-                    \FunctionTok{starts\_with}\NormalTok{(}\StringTok{"QES{-}52A"}\NormalTok{) }\SpecialCharTok{\&}
-                      \FunctionTok{ends\_with}\NormalTok{(}\FunctionTok{as.character}\NormalTok{(year)))\}}
+\CommentTok{# make a list for years where we used the "QES-52A"}
+\NormalTok{year_A <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\DecValTok{1979}\OperatorTok{:}\DecValTok{1987}\NormalTok{, }\DecValTok{1993}\NormalTok{)}
+\CommentTok{#function to get the hour of work}
+\NormalTok{get_hour <-}\StringTok{ }\ControlFlowTok{function}\NormalTok{(year)\{}
+  \ControlFlowTok{if}\NormalTok{(year }\OperatorTok{%in%}\StringTok{ }\NormalTok{year_A)\{}
+\NormalTok{    temp <-}\StringTok{ }\NormalTok{new_data_qnames }\OperatorTok{%>%}
+\StringTok{      }\NormalTok{dplyr}\OperatorTok{::}\KeywordTok{select}\NormalTok{(CASEID_}\DecValTok{1979}\NormalTok{,}
+                    \KeywordTok{starts_with}\NormalTok{(}\StringTok{"QES-52A"}\NormalTok{) }\OperatorTok{&}
+\StringTok{                      }\KeywordTok{ends_with}\NormalTok{(}\KeywordTok{as.character}\NormalTok{(year)))\}}
   \ControlFlowTok{else}\NormalTok{\{}
-\NormalTok{    temp }\OtherTok{\textless{}{-}}\NormalTok{ new\_data\_qnames }\SpecialCharTok{\%\textgreater{}\%}
-\NormalTok{      dplyr}\SpecialCharTok{::}\FunctionTok{select}\NormalTok{(CASEID\_1979,}
-                    \FunctionTok{starts\_with}\NormalTok{(}\StringTok{"QES{-}52D"}\NormalTok{) }\SpecialCharTok{\&}
-                      \FunctionTok{ends\_with}\NormalTok{(}\FunctionTok{as.character}\NormalTok{(year)))\}}
-\NormalTok{  temp }\SpecialCharTok{\%\textgreater{}\%}
-    \FunctionTok{pivot\_longer}\NormalTok{(}\SpecialCharTok{!}\NormalTok{CASEID\_1979,}
-                 \AttributeTok{names\_to =} \StringTok{"job"}\NormalTok{,}
-                 \AttributeTok{values\_to =} \StringTok{"hours\_work"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
-    \FunctionTok{separate}\NormalTok{(}\StringTok{"job"}\NormalTok{, }\FunctionTok{c}\NormalTok{(}\StringTok{"job"}\NormalTok{, }\StringTok{"year"}\NormalTok{), }\AttributeTok{sep =} \SpecialCharTok{{-}}\DecValTok{4}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
-    \FunctionTok{mutate}\NormalTok{(}\AttributeTok{job =} \FunctionTok{paste0}\NormalTok{(}\StringTok{"job\_"}\NormalTok{, }\FunctionTok{substr}\NormalTok{(job, }\DecValTok{9}\NormalTok{, }\DecValTok{10}\NormalTok{))) }\SpecialCharTok{\%\textgreater{}\%}
-    \FunctionTok{rename}\NormalTok{(}\AttributeTok{id =}\NormalTok{ CASEID\_1979)}
+\NormalTok{    temp <-}\StringTok{ }\NormalTok{new_data_qnames }\OperatorTok{%>%}
+\StringTok{      }\NormalTok{dplyr}\OperatorTok{::}\KeywordTok{select}\NormalTok{(CASEID_}\DecValTok{1979}\NormalTok{,}
+                    \KeywordTok{starts_with}\NormalTok{(}\StringTok{"QES-52D"}\NormalTok{) }\OperatorTok{&}
+\StringTok{                      }\KeywordTok{ends_with}\NormalTok{(}\KeywordTok{as.character}\NormalTok{(year)))\}}
+\NormalTok{  temp }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{pivot_longer}\NormalTok{(}\OperatorTok{!}\NormalTok{CASEID_}\DecValTok{1979}\NormalTok{,}
+                 \DataTypeTok{names_to =} \StringTok{"job"}\NormalTok{,}
+                 \DataTypeTok{values_to =} \StringTok{"hours_work"}\NormalTok{) }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{separate}\NormalTok{(}\StringTok{"job"}\NormalTok{, }\KeywordTok{c}\NormalTok{(}\StringTok{"job"}\NormalTok{, }\StringTok{"year"}\NormalTok{), }\DataTypeTok{sep =} \DecValTok{-4}\NormalTok{) }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{job =} \KeywordTok{paste0}\NormalTok{(}\StringTok{"job_"}\NormalTok{, }\KeywordTok{substr}\NormalTok{(job, }\DecValTok{9}\NormalTok{, }\DecValTok{10}\NormalTok{))) }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{rename}\NormalTok{(}\DataTypeTok{id =}\NormalTok{ CASEID_}\DecValTok{1979}\NormalTok{)}
 \NormalTok{\}}
-\CommentTok{\# list to save the iteration result}
-\NormalTok{hours }\OtherTok{\textless{}{-}} \FunctionTok{list}\NormalTok{()}
-\CommentTok{\# getting the hours of work of all observations}
-\ControlFlowTok{for}\NormalTok{(ayear }\ControlFlowTok{in} \FunctionTok{c}\NormalTok{(}\DecValTok{1979}\SpecialCharTok{:}\DecValTok{1994}\NormalTok{, }\DecValTok{1996}\NormalTok{, }\DecValTok{1998}\NormalTok{, }\DecValTok{2000}\NormalTok{, }\DecValTok{2002}\NormalTok{, }\DecValTok{2004}\NormalTok{, }\DecValTok{2006}\NormalTok{, }\DecValTok{2008}\NormalTok{, }\DecValTok{2010}\NormalTok{,}
+\CommentTok{# list to save the iteration result}
+\NormalTok{hours <-}\StringTok{ }\KeywordTok{list}\NormalTok{()}
+\CommentTok{# getting the hours of work of all observations}
+\ControlFlowTok{for}\NormalTok{(ayear }\ControlFlowTok{in} \KeywordTok{c}\NormalTok{(}\DecValTok{1979}\OperatorTok{:}\DecValTok{1994}\NormalTok{, }\DecValTok{1996}\NormalTok{, }\DecValTok{1998}\NormalTok{, }\DecValTok{2000}\NormalTok{, }\DecValTok{2002}\NormalTok{, }\DecValTok{2004}\NormalTok{, }\DecValTok{2006}\NormalTok{, }\DecValTok{2008}\NormalTok{, }\DecValTok{2010}\NormalTok{,}
                \DecValTok{2012}\NormalTok{, }\DecValTok{2014}\NormalTok{, }\DecValTok{2016}\NormalTok{, }\DecValTok{2018}\NormalTok{)) \{}
-\NormalTok{  hours[[ayear]] }\OtherTok{\textless{}{-}} \FunctionTok{get\_hour}\NormalTok{(ayear)}
+\NormalTok{  hours[[ayear]] <-}\StringTok{ }\KeywordTok{get_hour}\NormalTok{(ayear)}
 \NormalTok{\}}
-\CommentTok{\# unlist the hours of work}
-\NormalTok{hours\_all }\OtherTok{\textless{}{-}} \FunctionTok{bind\_rows}\NormalTok{(}\SpecialCharTok{!!!}\NormalTok{hours)}
+\CommentTok{# unlist the hours of work}
+\NormalTok{hours_all <-}\StringTok{ }\KeywordTok{bind_rows}\NormalTok{(}\OperatorTok{!!!}\NormalTok{hours)}
 \end{Highlighting}
@@
 \begin{Highlighting}[]
-\NormalTok{get\_rate }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{(year) \{}
-\NormalTok{  new\_data\_qnames }\SpecialCharTok{\%\textgreater{}\%}
-\NormalTok{    dplyr}\SpecialCharTok{::}\FunctionTok{select}\NormalTok{(CASEID\_1979,}
-                  \FunctionTok{starts\_with}\NormalTok{(}\StringTok{"HRP"}\NormalTok{) }\SpecialCharTok{\&}
-                    \FunctionTok{ends\_with}\NormalTok{(}\FunctionTok{as.character}\NormalTok{(year))) }\SpecialCharTok{\%\textgreater{}\%}
-    \FunctionTok{pivot\_longer}\NormalTok{(}\SpecialCharTok{!}\NormalTok{CASEID\_1979, }\AttributeTok{names\_to =} \StringTok{"job"}\NormalTok{, }\AttributeTok{values\_to =} \StringTok{"rate\_per\_hour"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
-    \FunctionTok{separate}\NormalTok{(}\StringTok{"job"}\NormalTok{, }\FunctionTok{c}\NormalTok{(}\StringTok{"job"}\NormalTok{, }\StringTok{"year"}\NormalTok{), }\AttributeTok{sep =} \SpecialCharTok{{-}}\DecValTok{4}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
-    \FunctionTok{mutate}\NormalTok{(}\AttributeTok{job =} \FunctionTok{paste0}\NormalTok{(}\StringTok{"job\_0"}\NormalTok{, }\FunctionTok{substr}\NormalTok{(job, }\DecValTok{4}\NormalTok{, }\DecValTok{4}\NormalTok{))) }\SpecialCharTok{\%\textgreater{}\%}
-    \FunctionTok{rename}\NormalTok{(}\AttributeTok{id =}\NormalTok{ CASEID\_1979)}
+\NormalTok{get_rate <-}\StringTok{ }\ControlFlowTok{function}\NormalTok{(year) \{}
+\NormalTok{  new_data_qnames }\OperatorTok{%>%}
+\StringTok{    }\NormalTok{dplyr}\OperatorTok{::}\KeywordTok{select}\NormalTok{(CASEID_}\DecValTok{1979}\NormalTok{,}
+                  \KeywordTok{starts_with}\NormalTok{(}\StringTok{"HRP"}\NormalTok{) }\OperatorTok{&}
+\StringTok{                    }\KeywordTok{ends_with}\NormalTok{(}\KeywordTok{as.character}\NormalTok{(year))) }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{pivot_longer}\NormalTok{(}\OperatorTok{!}\NormalTok{CASEID_}\DecValTok{1979}\NormalTok{, }\DataTypeTok{names_to =} \StringTok{"job"}\NormalTok{, }\DataTypeTok{values_to =} \StringTok{"rate_per_hour"}\NormalTok{) }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{separate}\NormalTok{(}\StringTok{"job"}\NormalTok{, }\KeywordTok{c}\NormalTok{(}\StringTok{"job"}\NormalTok{, }\StringTok{"year"}\NormalTok{), }\DataTypeTok{sep =} \DecValTok{-4}\NormalTok{) }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{job =} \KeywordTok{paste0}\NormalTok{(}\StringTok{"job_0"}\NormalTok{, }\KeywordTok{substr}\NormalTok{(job, }\DecValTok{4}\NormalTok{, }\DecValTok{4}\NormalTok{))) }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{rename}\NormalTok{(}\DataTypeTok{id =}\NormalTok{ CASEID_}\DecValTok{1979}\NormalTok{)}
 \NormalTok{\}}
-\NormalTok{rates }\OtherTok{\textless{}{-}} \FunctionTok{list}\NormalTok{()}
-\ControlFlowTok{for}\NormalTok{(ayear }\ControlFlowTok{in} \FunctionTok{c}\NormalTok{(}\DecValTok{1979}\SpecialCharTok{:}\DecValTok{1994}\NormalTok{, }\DecValTok{1996}\NormalTok{, }\DecValTok{1998}\NormalTok{, }\DecValTok{2000}\NormalTok{, }\DecValTok{2002}\NormalTok{, }\DecValTok{2004}\NormalTok{, }\DecValTok{2006}\NormalTok{, }\DecValTok{2008}\NormalTok{, }\DecValTok{2010}\NormalTok{,}
+\NormalTok{rates <-}\StringTok{ }\KeywordTok{list}\NormalTok{()}
+\ControlFlowTok{for}\NormalTok{(ayear }\ControlFlowTok{in} \KeywordTok{c}\NormalTok{(}\DecValTok{1979}\OperatorTok{:}\DecValTok{1994}\NormalTok{, }\DecValTok{1996}\NormalTok{, }\DecValTok{1998}\NormalTok{, }\DecValTok{2000}\NormalTok{, }\DecValTok{2002}\NormalTok{, }\DecValTok{2004}\NormalTok{, }\DecValTok{2006}\NormalTok{, }\DecValTok{2008}\NormalTok{, }\DecValTok{2010}\NormalTok{,}
                \DecValTok{2012}\NormalTok{, }\DecValTok{2014}\NormalTok{, }\DecValTok{2016}\NormalTok{, }\DecValTok{2018}\NormalTok{)) \{}
-\NormalTok{  rates[[ayear]] }\OtherTok{\textless{}{-}} \FunctionTok{get\_rate}\NormalTok{(ayear)}
+\NormalTok{  rates[[ayear]] <-}\StringTok{ }\KeywordTok{get_rate}\NormalTok{(ayear)}
 \NormalTok{\}}
-\NormalTok{rates\_all }\OtherTok{\textless{}{-}} \FunctionTok{bind\_rows}\NormalTok{(}\SpecialCharTok{!!!}\NormalTok{rates)}
-\CommentTok{\# join hours and rates variable}
-\NormalTok{hours\_wages }\OtherTok{\textless{}{-}} \FunctionTok{left\_join}\NormalTok{(rates\_all,}
-\NormalTok{                         hours\_all,}
-                         \AttributeTok{by =} \FunctionTok{c}\NormalTok{(}\StringTok{"id"}\NormalTok{, }\StringTok{"year"}\NormalTok{, }\StringTok{"job"}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%}
-  \CommentTok{\# set the 0 value in rate\_per\_hour as NA}
-  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{rate\_per\_hour =} \FunctionTok{ifelse}\NormalTok{(rate\_per\_hour }\SpecialCharTok{==} \DecValTok{0}\NormalTok{, }\ConstantTok{NA}\NormalTok{,}
-\NormalTok{                                rate\_per\_hour),}
-         \AttributeTok{hours\_work =} \FunctionTok{ifelse}\NormalTok{(hours\_work }\SpecialCharTok{==} \DecValTok{0} \SpecialCharTok{|}\NormalTok{ hours\_work }\SpecialCharTok{\textgreater{}} \DecValTok{84}\NormalTok{, }\ConstantTok{NA}\NormalTok{, hours\_work))}
+\NormalTok{rates_all <-}\StringTok{ }\KeywordTok{bind_rows}\NormalTok{(}\OperatorTok{!!!}\NormalTok{rates)}
+\CommentTok{# join hours and rates variable}
+\NormalTok{hours_wages <-}\StringTok{ }\KeywordTok{left_join}\NormalTok{(rates_all,}
+\NormalTok{                         hours_all,}
+                         \DataTypeTok{by =} \KeywordTok{c}\NormalTok{(}\StringTok{"id"}\NormalTok{, }\StringTok{"year"}\NormalTok{, }\StringTok{"job"}\NormalTok{)) }\OperatorTok{%>%}
+\StringTok{  }\CommentTok{# set the 0 value in rate_per_hour as NA}
+\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{rate_per_hour =} \KeywordTok{ifelse}\NormalTok{(rate_per_hour }\OperatorTok{==}\StringTok{ }\DecValTok{0}\NormalTok{, }\OtherTok{NA}\NormalTok{,}
+\NormalTok{                                rate_per_hour),}
+         \DataTypeTok{hours_work =} \KeywordTok{ifelse}\NormalTok{(hours_work }\OperatorTok{==}\StringTok{ }\DecValTok{0} \OperatorTok{|}\StringTok{ }\NormalTok{hours_work }\OperatorTok{>}\StringTok{ }\DecValTok{84}\NormalTok{, }\OtherTok{NA}\NormalTok{, hours_work))}
 \end{Highlighting}
@@
 
-According to Huebner Marianne, Vach, and Cessie (2016), Initial Data Analysis (IDA) is the step of inspecting and screening the data after being collected to ensure that the data is clean, valid, and ready to be deployed in the later formal statistical analysis.
+According to Huebner, Vach, and Cessie (2016), Initial Data Analysis (IDA) is the step of inspecting and screening the data after being collected to ensure that the data is clean, valid, and ready to be deployed in the later formal statistical analysis.
 Moreover, Chatfield (1985) argues that the two main objectives of IDA are data description, which is to assess the structure and the quality of the data, and model formulation without any formal statistical inference.
@@
 
-The next step is that explore the mean hourly wage data.
-In this case, we only explore the wages data of up to \(12^{th}\) grade.
-Table \ref{tab:summarytable} shows that the median of the overall wage of the cohort is only 7.2, while the mean is 11.87.
-It indicates that the data might contains a lot of extreme values.
-
-\begin{table}
-
-\caption{\label{tab:summarytable}Summary Statistics of Wages of High School Data}
-\centering
-\begin{tabular}[t]{l|r}
-\hline
-Statistics & Value\\
-\hline
-Min. & 0.01000\\
-\hline
-1st Qu. & 4.50000\\
-\hline
-Median & 7.20000\\
-\hline
-Mean & 11.86593\\
-\hline
-3rd Qu. & 11.74000\\
-\hline
-Max. & 33400.00000\\
-\hline
-\end{tabular}
-\end{table}
+The next step is that explore the mean hourly wage data. In this case, we only explore the wages data of up to \(12^{th}\) grade. Here, we use visualisation techniques to do IDA.
 
@@
 
-However, we can not be sure only by observing the summary statistics.
-Hence, we use visualisation techniques to investigate this matter.
+We take 36 samples randomly from the data and plot them, as shown in Figure \ref{fig:sample-plot}.
+It implies that these respondents have a lot of variability in wages, for example, the IDs in panel numbers 5, 7, and 11.
+The plot also implies that the samples have a different pattern of mean hourly wages.
+Some have had flat wages for years but had a sudden increase in one particular year, then it gone down again, while the others experienced an upsurge in their wage, for instance, the IDs in panel 9.
+
+Further, we investigate the summary statistics of wages.
 Figure \ref{fig:feature-plot} conveys a problem in the mean hourly wage values.
@@
 In Figure \ref{fig:feature-plot} B, we barely see any difference in the minimum, median, and maximum value of the wages since the distribution is heavily skewed to the right.
+The problem is more clearly conveyed in Figure \ref{fig:feature-plot} C where there is no extreme values in the minimum and median values of wages but the maximum values show some anomalies.
+In Figure \ref{fig:feature-plot} D, we show the example of anomaly in one observation.
+This respondent only experienced one point of extremely high wages.
+Thus, we suspect that these high values are erroneous resulted from a data entry error.
 
@@
 
-{\centering \includegraphics[width=0.7\linewidth]{figures/feature-plot-1}
+{\centering \includegraphics[width=0.9\linewidth]{figures/feature-plot-1}
 
@@
 \caption{Several plots showing anomalies observed the input data.
-XX Add a couple more plots that reveal the initial problems in the data.
-Plot A portrays the pattern of mean hourly wage of high school cohort from 1979 to 2018 of each ID in US Dollar; Plot B shows the distribution of their minimum, median, and maximum value.
-We can see that some IDs had an extremely high of wages and it made the distribution of the three features is extremely skewed.}\label{fig:feature-plot}
+Plot A portrays the pattern of mean hourly wage of high school cohort from 1979 to 2018 of each ID in US Dollar.
+We can see some extremely high wages; Plot B shows the distribution of their minimum, median, and maximum value.
+We can also see that some IDs had an extremely high of wages and it made the distribution of the three features is extremely skewed; Plot C shows the extreme values more clearly; Plot D shows an ID with a suspected erronous wages.}\label{fig:feature-plot}
 \end{figure}
-
-In Figure \ref{fig:high-wages}, we plot some respondents with a high value of mean hourly wages.
-We filter all of the IDs who earned more than US\$ 500 per hour on average.
-We find that these respondents only experienced one point of extremely high wages.
-Thus, we suspect that these high values are erroneous resulted from a data entry error.
-
-\begin{figure}
-
-{\centering \includegraphics[width=0.7\linewidth]{figures/high-wages-1}
-
-}
-
-\caption{6 out of 45 IDs with extremely high mean hourly wage. Most of the IDs only have one point of high wage.}\label{fig:high-wages}
-\end{figure}
-
-Further, we take 36 samples randomly from the data and plot them, as shown in Figure \ref{fig:sample-plot}.
-It implies that not only that some observations earn extremely high figures of wages, but some also have reasonably fluctuated wages, for example, the IDs in panel numbers 5, 7, and 11.
-The plot also implies that the samples have a different pattern of mean hourly wages.
-Some have had flat wages for years but had a sudden increase in one particular year, then it gone down again, while the others experienced an upsurge in their wage, for instance, the IDs in panel 9.
 
@@
 However, this method is too computationally and memory expensive, especially for a large data set, like the NLSY79 data.
-Thus, the model for each ID is built utilizing the \texttt{nest} and \texttt{map} function from \texttt{tidyr} (Wickham 2020b) and \texttt{purrr} (Henry and Wickham 2020), respectively.
+Thus, the model for each ID is built utilizing the \texttt{nest} and \texttt{map} function from \texttt{tidyr} (H.
+Wickham 2020b) and \texttt{purrr} (Henry and Wickham 2020), respectively.
 
@@
 
-To minimize the risk of mistakenly identifying an outlier as an ``erroneous outlier,'' we simulate some thresholds and study how they affect the data.
+To minimize the risk of mistakenly identifying an outlier as an ``erroneous outlier'', we simulate some thresholds and study how they affect the data.
 We find that 0.12 is the most reasonable value to be the threshold to minimize that drawback's risk because it still captures the sensible spikes in the data.
@@
 It implies that the fluctuation can still be observed in the data after the treatment.
-However, the large spikes, which are considered ``erroneous outliers,'' are already eliminated from the data.
+However, the large spikes, which are considered ``erroneous outliers'', are already eliminated from the data.
 Hence, the model produces a data set with a more reasonable degree of fluctuation.
@@
 Extreme values in slope are observed in a few observations, while the range of slope is more stable with more observations.
-Accordingly, we trim the data to only include \(5^{th} - 95^{th}\) percentile.
+Accordingly, we trim the data to only include respondents with at least 10 observations.
 
@@
 \end{figure}
+
+Further, we perform a linear regression to see the effect of education on wages experience.
 
@@
 
-The entire analysis is conducted using \texttt{R} (R Core Team 2020) in \texttt{rstudio} using these packages: \texttt{tidyverse} (Wickham et al.
-2019), \texttt{ggplot2} (Wickham 2016), \texttt{dplyr} (Wickham et al.
-2020), \texttt{readr} (Wickham and Hester 2020), \texttt{tidyr} (Wickham 2020b), \texttt{stringr} (Wickham 2019), \texttt{purrr} (Henry and Wickham 2020), \texttt{broom} (Robinson, Hayes, and Couch 2021), \texttt{blorgar} (Tierney, Cook, and Prvan 2020), \texttt{patchwork} (Pedersen 2020), \texttt{kableExtra} (Zhu 2019), \texttt{MASS} (Venables and Ripley 2002), \texttt{janitor} (Firke 2020), \texttt{DiagrammeR} (Iannone 2020), \texttt{rsvg} (Ooms 2020), \texttt{webshot} (Chang 2019), \texttt{mgcv} (Wood 2003), \texttt{tsibble} (Wang, Cook, and Hyndman 2020), and \texttt{modelr} (Wickham 2020a).
+The entire analysis is conducted using \texttt{R} (R Core Team 2020) in \texttt{rstudio} using these packages: \texttt{tidyverse} (H.
+Wickham, Averick, et al.
+2019), \texttt{ggplot2} (Wickham 2016), \texttt{dplyr} (H.
+Wickham, François, et al.
+2020), \texttt{readr} (H.
+Wickham and Hester 2020), \texttt{tidyr} (H.
+Wickham 2020b), \texttt{stringr} (H.
+Wickham 2019), \texttt{purrr} (Henry and Wickham 2020), \texttt{broom} (Robinson, Hayes, and Couch 2021), \texttt{brolgar} (Tierney, Cook, and Prvan 2020), \texttt{patchwork} (Pedersen 2020), \texttt{kableExtra} (Zhu 2019), \texttt{MASS} (Venables and Ripley 2002), \texttt{janitor} (Firke 2020), \texttt{DiagrammeR} (Iannone 2020), \texttt{rsvg} (Ooms 2020), \texttt{webshot} (Chang 2019), \texttt{mgcv} (Wood 2003), \texttt{tsibble} (Wang, Cook, and Hyndman 2020), and \texttt{modelr} (H.
+Wickham 2020a).
 The paper are generated using \texttt{knitr} (Xie 2014), \texttt{rmarkdown} (Xie, Dervieux, and Riederer 2020), and \texttt{rticles} (Allaire et al.
@@
 \hypertarget{refs}{}
-\begin{CSLReferences}{1}{0}
 \leavevmode\hypertarget{ref-rticles}{}%
@@
 2021.
-\emph{Rticles: Article Formats for r Markdown}.
+\emph{Rticles: Article Formats for R Markdown}.
 \url{https://CRAN.R-project.org/package=rticles}.
@@
 \leavevmode\hypertarget{ref-shiny}{}%
-Chang, Winston, Joe Cheng, JJ Allaire, Yihui Xie, and Jonathan McPherson. 2020. \emph{Shiny: Web Application Framework for r}. \url{https://CRAN.R-project.org/package=shiny}.
+Chang, Winston, Joe Cheng, JJ Allaire, Yihui Xie, and Jonathan McPherson. 2020. \emph{Shiny: Web Application Framework for R}. \url{https://CRAN.R-project.org/package=shiny}.
 
 \leavevmode\hypertarget{ref-Chatfield1985TIEo}{}%
-Chatfield, C. 1985. {``The Initial Examination of Data.''} \emph{Journal of the Royal Statistical Society. Series A. General} 148 (3): 214--53.
+Chatfield, C. 1985. ``The Initial Examination of Data.'' \emph{Journal of the Royal Statistical Society. Series A. General} 148 (3): 214--53.
 
@@
 2017.
-{``Using the National Longitudinal Surveys of Youth (NLSY) to Conduct Life Course Analyses.''} In \emph{Handbook of Life Course Health Development}, edited by Richard M.
+``Using the National Longitudinal Surveys of Youth (Nlsy) to Conduct Life Course Analyses.'' In \emph{Handbook of Life Course Health Development}, edited by Richard M.
 Lerner Neal Halfon Christoper B.
@@
 Cham: Springer.
-https://doi.org/\url{https://doi.org/10.1007/978-3-319-47143-3_23}.
+\url{https://doi.org/https://doi.org/10.1007/978-3-319-47143-3_23}.
 
@@
 2016.
-{``A Systematic Approach to Initial Data Analysis Is Good Research Practice.''} \emph{The Journal of Thoracic and Cardiovascular Surgery} 151 (1): 25--27.
+``A Systematic Approach to Initial Data Analysis Is Good Research Practice.'' \emph{The Journal of Thoracic and Cardiovascular Surgery} 151 (1): 25--27.
 
@@
 2020.
-{``Hidden Analyses: A Review of Reporting Practice and Recommendations for More Transparent Reporting of Initial Data Analyses.''} \emph{BMC Medical Research Methodology} 20 (1): 61--61.
+``Hidden Analyses: A Review of Reporting Practice and Recommendations for More Transparent Reporting of Initial Data Analyses.'' \emph{BMC Medical Research Methodology} 20 (1): 61--61.
 
@@
 \leavevmode\hypertarget{ref-KollerManuel2016rARP}{}%
-Koller, Manuel. 2016. {``Robustlmm: An r Package for Robust Estimation of Linear Mixed-Effects Models.''} \emph{Journal of Statistical Software} 75 (6): 1--24.
+Koller, Manuel. 2016. ``Robustlmm: An R Package for Robust Estimation of Linear Mixed-Effects Models.'' \emph{Journal of Statistical Software} 75 (6): 1--24.
 
 \leavevmode\hypertarget{ref-OECD}{}%
-OECD. n.d. {``Gender Wage Gap.''} https://data.oecd.org/earnwage/gender-wage-gap.htm.
+OECD. n.d. ``Gender Wage Gap.'' https://data.oecd.org/earnwage/gender-wage-gap.htm.
 
 \leavevmode\hypertarget{ref-rsvg}{}%
-Ooms, Jeroen. 2020. \emph{Rsvg: Render SVG Images into PDF, PNG, PostScript, or Bitmap Arrays}. \url{https://CRAN.R-project.org/package=rsvg}.
+Ooms, Jeroen. 2020. \emph{Rsvg: Render Svg Images into Pdf, Png, Postscript, or Bitmap Arrays}. \url{https://CRAN.R-project.org/package=rsvg}.
 
 \leavevmode\hypertarget{ref-opendata}{}%
-Open Knowledge Foundation. 2021. {``Open Definition. Defining Open in Open Data, Open Content, and Open Knowledge.''} 2021. \url{http://opendefinition.org/od/2.1/en/}.
+Open Knowledge Foundation. 2021. ``Open Definition. Defining Open in Open Data, Open Content, and Open Knowledge.'' 2021. \url{http://opendefinition.org/od/2.1/en/}.
 
@@
 2001.
-{``Data Watch: The National Longitudinal Surveys.''} \emph{The Journal of Economic Perspectives} 15 (2): 239--53.
+``Data Watch: The National Longitudinal Surveys.'' \emph{The Journal of Economic Perspectives} 15 (2): 239--53.
 
@@
 \leavevmode\hypertarget{ref-nlsy79}{}%
-The U.S. Bureau of Labor Statistics. n.d. {``National Longitudinal Survey of Youth 1979.''} Available at \url{https://www.nlsinfo.org/content/cohorts/nlsy79} (2021/25/02).
+The U.S. Bureau of Labor Statistics. n.d. ``National Longitudinal Survey of Youth 1979.'' Available at \url{https://www.nlsinfo.org/content/cohorts/nlsy79} (2021/25/02).
 
 \leavevmode\hypertarget{ref-brolgar}{}%
-Tierney, Nicholas, Di Cook, and Tania Prvan. 2020. \emph{Brolgar: BRowse over Longitudinal Data Graphically and Analytically in r}. \url{https://github.com/njtierney/brolgar}.
+Tierney, Nicholas, Di Cook, and Tania Prvan. 2020. \emph{Brolgar: BRowse over Longitudinal Data Graphically and Analytically in R}. \url{https://github.com/njtierney/brolgar}.
 
@@
 \leavevmode\hypertarget{ref-rlm}{}%
-UCLA: Statistical Consulting Group. 2021. {``Robust Regression \textbar{} r Data Analysis Examples.''} February 2021. \url{https://stats.idre.ucla.edu/r/dae/robust-regression/}.
+UCLA: Statistical Consulting Group. 2021. ``Robust Regression \textbar{} R Data Analysis Examples.'' February 2021. \url{https://stats.idre.ucla.edu/r/dae/robust-regression/}.
 
 \leavevmode\hypertarget{ref-validate}{}%
-van der Loo, Mark P. J., and Edwin de Jonge. 2021. {``Data Validation Infrastructure for {R}.''} \emph{Journal of Statistical Software} 97 (10): 1--31. \url{https://doi.org/10.18637/jss.v097.i10}.
+van der Loo, Mark P. J., and Edwin de Jonge. 2021. ``Data Validation Infrastructure for R.'' \emph{Journal of Statistical Software} 97 (10): 1--31. \url{https://doi.org/10.18637/jss.v097.i10}.
 
 \leavevmode\hypertarget{ref-mass}{}%
-Venables, W. N., and B. D. Ripley. 2002. \emph{Modern Applied Statistics with s}. Fourth. New York: Springer. \url{http://www.stats.ox.ac.uk/pub/MASS4}.
+Venables, W. N., and B. D. Ripley. 2002. \emph{Modern Applied Statistics with S}. Fourth. New York: Springer. \url{http://www.stats.ox.ac.uk/pub/MASS4}.
 
@@
 2020.
-{``A New Tidy Data Structure to Support Exploration and Modeling of Temporal Data.''} \emph{Journal of Computational and Graphical Statistics} 29 (3): 466--78.
+``A New Tidy Data Structure to Support Exploration and Modeling of Temporal Data.'' \emph{Journal of Computational and Graphical Statistics} 29 (3): 466--78.
 \url{https://doi.org/10.1080/10618600.2019.1695624}.
@@
 \leavevmode\hypertarget{ref-WickhamHadley2014TD}{}%
-Wickham, Hadley. 2014. {``Tidy Data.''} \emph{Journal of Statistical Software} 59 (10): 1--23.
+Wickham, Hadley. 2014. ``Tidy Data.'' \emph{Journal of Statistical Software} 59 (10): 1--23.
 
@@
 2019.
-{``Welcome to the {tidyverse}.''} \emph{Journal of Open Source Software} 4 (43): 1686.
+``Welcome to the tidyverse.'' \emph{Journal of Open Source Software} 4 (43): 1686.
 \url{https://doi.org/10.21105/joss.01686}.
@@
 \leavevmode\hypertarget{ref-mgcv}{}%
-Wood, S. N. 2003. {``Thin-Plate Regression Splines.''} \emph{Journal of the Royal Statistical Society (B)} 65 (1): 95--114.
+Wood, S. N. 2003. ``Thin-Plate Regression Splines.'' \emph{Journal of the Royal Statistical Society (B)} 65 (1): 95--114.
 
@@
 2014.
-{``Knitr: A Comprehensive Tool for Reproducible Research in {R}.''} In \emph{Implementing Reproducible Computational Research}, edited by Victoria Stodden, Friedrich Leisch, and Roger D.
+``Knitr: A Comprehensive Tool for Reproducible Research in R.'' In \emph{Implementing Reproducible Computational Research}, edited by Victoria Stodden, Friedrich Leisch, and Roger D.
 Peng.
@@
 \leavevmode\hypertarget{ref-kableExtra}{}%
-Zhu, Hao. 2019. \emph{kableExtra: Construct Complex Table with 'Kable' and Pipe Syntax}. \url{https://CRAN.R-project.org/package=kableExtra}.
-
-\end{CSLReferences}
+Zhu, Hao. 2019. \emph{KableExtra: Construct Complex Table with 'Kable' and Pipe Syntax}. \url{https://CRAN.R-project.org/package=kableExtra}.
 

### 2021-09-16 paper/scripts/ida.R

@@ -45,11 +45,11 @@ plot_high <- ggplot(filter(wages_high2, id == 39)) +
                  y = mean_hourly_wage),
              size = 0.5,
              alpha = 0.5) +
-  theme(axis.text.x = element_text(angle = 10, size = 6),
-        plot.title = element_text(size = 10)) +
   ylab("mean hourly wage") +
   theme_bw() +
-  ggtitle("D)")
+  ggtitle("D)") +
+  theme(axis.text.x = element_text(angle = 10, size = 6),
+        plot.title = element_text(size = 10))
 
 ## ---- sample-plot
 wages_demog_hs_tsibble <- as_tsibble(x = wages_demog_hs,
@@ -75,9 +75,10 @@ spag <- wages_demog_hs %>%
              group = id)) +
   geom_line(alpha = 0.1) +
   ggtitle("A)") +
-  theme(plot.title = element_text(size = 10)) +
   theme_bw() +
-  ylab("mean hourly wage")
+  ylab("mean hourly wage") +
+  theme(plot.title = element_text(size = 10),
+        axis.text.x = element_text(angle = 10, size = 6))
 
 wages_three_feat <- wages_demog_hs_tsibble %>%
   features(mean_hourly_wage,
@@ -89,14 +90,16 @@ feature <- ggplot(wages_feat_long) +
   geom_density(aes(x = value, colour = feature, fill = feature), alpha = 0.3) +
   ggtitle("B)") +
   theme_bw() +
-  theme(plot.title = element_text(size = 10))
+  theme(plot.title = element_text(size = 10),
+        axis.text.x = element_text(angle = 10, size = 6),
+        legend.position = "none")
 
 feature_bp <- ggplot(wages_feat_long, aes(y=value, fill = feature, color = feature)) +
   geom_boxplot() +
   theme_bw() +
   ggtitle("C)") +
-  theme(legend.position = "none",
-        plot.title = element_text(size = 10))
+  theme(plot.title = element_text(size = 10),
+        axis.text.x = element_text(angle = 10, size = 6))
 
 spag + feature + feature_bp + plot_high + plot_layout(nrow = 1, guides = "collect") &
   theme(legend.position = "bottom")

### 2021-09-16 paper/paper-jsdse/paper-jsdse.Rmd
author fixed

@@
   - name: Dewi Amaliah
-    thanks: ""
+    thanks: "Test"
     department: Department of Econometrics and Business Statistics
@@
 
-```{r, child = "../paper.Rmd"}
+```{r, child = "../paper.Rmd", eval = FALSE}
 ```

### 2021-09-16 paper/paper.Rmd
move abstract to YAML

--- 2021-09-14 paper/paper.Rmd
@@
     email: ""
+abstract: > 
+ NSLY79 is a prominent open data source that has been an important support for educational purposes and multidisciplinary research on longitudinal data.
+Subsets of this data can be found in numerous textbooks and research articles.
+However, the steps and decisions taken to get from the original data to these subsets is never clearly articulated.
+This article describes our journey when trying to re-create a textbook example data set from the original database.
+Ultimately, the goal is to provide the resources so that refreshing data subsets can be done on a regular basis, especially when updates are made to the database.
+Thus, this paper demonstrates the process of initial data analysis (IDA) -- tidying, cleaning, and documenting the process -- to make data available for text book examples or research.
+A brief exploratory analysis of the resulting data is also conducted.
+Three new data sets, and the code to produce them are provided in an accompanying open source R package, called `yowie`.
+As a result of this process, some recommendations are also made for the NLSY79 curators for incorporating data quality checks, and providing more convenient samples of the data to potential users.
 bibliography: references.bib
@@
 ```
-
-NSLY79 is a prominent open data source that has been an important support for educational purposes and multidisciplinary research on longitudinal data.
-Subsets of this data can be found in numerous textbooks and research articles.
-However, the steps and decisions taken to get from the original data to these subsets is never clearly articulated.
-This article describes our journey when trying to re-create a textbook example data set from the original database.
-Ultimately, the goal is to provide the resources so that refreshing data subsets can be done on a regular basis, especially when updates are made to the database.
-Thus, this paper demonstrates the process of initial data analysis (IDA) -- tidying, cleaning, and documenting the process -- to make data available for text book examples or research.
-A brief exploratory analysis of the resulting data is also conducted.
-Three new data sets, and the code to produce them are provided in an accompanying open source R package, called `yowie`.
-As a result of this process, some recommendations are also made for the NLSY79 curators for incorporating data quality checks, and providing more convenient samples of the data to potential users.
 

### 2021-09-16 paper/paper-jsdse/paper-jsdse.tex

@@
 
-\usepackage{color}
-\usepackage{fancyvrb}
-\newcommand{\VerbBar}{|}
-\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
-\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
-% Add ',fontsize=\small' for more characters per line
-\usepackage{framed}
-\definecolor{shadecolor}{RGB}{248,248,248}
-\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
-\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
-\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
-\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
-\newcommand{\BuiltInTok}[1]{#1}
-\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
-\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
-\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
-\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
-\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
-\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
-\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
-\newcommand{\ExtensionTok}[1]{#1}
-\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
-\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
-\newcommand{\ImportTok}[1]{#1}
-\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
-\newcommand{\NormalTok}[1]{#1}
-\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
-\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
-\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
-\newcommand{\RegionMarkerTok}[1]{#1}
-\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
-\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
-\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
-\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
-\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
-\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-
 \begin{document}
@@
   \title{\bf The Journey from Wild to Textbook Data: A Case Study from the National Longitudinal Survey of Youth}
-  \author{Author 1\thanks{
-    The authors gratefully acknowledge \textit{please remember to list all relevant funding sources in the unblinded version}}\hspace{.2cm}\\
-    Department of YYY, University of XXX\\
-    and \\
-    Author 2 \\
-    Department of ZZZ, University of WWW}
+  \author{
+            Dewi Amaliah\thanks{Test}\\
+      Department of Econometrics and Business Statistics, Monash University, \\
+            Dianne Cook\\
+      Department of Econometrics and Business Statistics, Monash University, \\
+            Emi Tanaka\\
+      Department of Econometrics and Business Statistics, Monash University, \\
+            Kate Hyde\\
+      Department of Econometrics and Business Statistics, Monash University, \\
+            Nicholas Tierney\\
+      , , }
         \maketitle
@@
 
-NSLY79 is a prominent open data source that has been an important support for educational purposes and multidisciplinary research on longitudinal data.
-Subsets of this data can be found in numerous textbooks and research articles.
-However, the steps and decisions taken to get from the original data to these subsets is never clearly articulated.
-This article describes our journey when trying to re-create a textbook example data set from the original database.
-Ultimately, the goal is to provide the resources so that refreshing data subsets can be done on a regular basis, especially when updates are made to the database.
-Thus, this paper demonstrates the process of initial data analysis (IDA) -- tidying, cleaning, and documenting the process -- to make data available for text book examples or research.
-A brief exploratory analysis of the resulting data is also conducted.
-Three new data sets, and the code to produce them are provided in an accompanying open source R package, called \texttt{yowie}.
-As a result of this process, some recommendations are also made for the NLSY79 curators for incorporating data quality checks, and providing more convenient samples of the data to potential users.
-
-\textbf{Keywords}: Data cleaning; Data tidying; Longitudinal data; NLSY79; Open data; Initial data analysis; Outlier detection; Robust linear regression
-
-\hypertarget{intro}{%
-\section{Introduction}\label{intro}}
-
-``Open data'' is data that is freely accessible, modifiable, and shareable by anyone for any purpose (Open Knowledge Foundation 2021).
-This type of data can be useful as example data in statistical textbooks and for research purposes.
-However, open data is often referred to as what we might call ``wild data'' because it requires substantial cleaning and tidying to tame it into textbook shape.
-Huebner Marianne, Vach, and Cessie (2016) emphasize that making the data cleaning process accountable and transparent is imperative.
-Documenting the data cleaning is essential for the integrity of downstream statistical analyses and model building (M.
-Huebner et al.
-2020).
-
-Data cleaning is part of what is called ``initial data analysis (IDA)'' (Chatfield 1985).
-The other IDA steps are to explore the data properties and check the assumption of modeling to document and report the process for the later formal analysis.
-Hence, exploratory data analysis, which is defined by Tukey (1977) as a detective work to get the clue from data, either numerically or graphically, before confirmatory data analysis is performed, encompasses IDA.
-Dasu and Johnson (2003) say that data cleaning and exploration, without naming it as IDA, is a difficult task and consumes 80\% of the data mining task.
-
-Despite its importance, this IDA stage is often undervalued and neglected (Chatfield 1985).
-There are few research papers that document the data cleaning (Wickham 2014).
-Furthermore, the decisions made in this stage often go unreported in the sense that IDA is often performed in an unplanned and unstructured way and is only shared among restricted parties (M.
-Huebner et al.
-2020).
-
-This paper aims to demonstrate the steps of IDA, tidying, cleaning, and documenting the process, for a prominent open data source, National Longitudinal Survey of Youth (NLSY) 1979, henceforth referred to NSLY79.
-This data has been playing an important role in research in various disciplines, including but are not limited to economics, sociology, education, public policy, and public health for more than a quarter of the century (Pergamit et al.
-2001).
-In addition, The National Longitudinal Survey is considered a survey with high retention rates and carefully designed, making it suitable for a life course research (Pergamit et al.
-2001) and (Cooksey 2017).
-According to Cooksey (2017), thousand of articles, and hundreds of book chapters and monographs has incorporated the NLSY data.
-Moreover, the NLSY79 is considered the most widely used and most important cohort in the NLSY79 data sets (Pergamit et al.
-2001).
-
-Singer and Willett (2003) used the wages and other variables of high school dropouts from the NLSY79 data as an example data set to illustrate longitudinal data modeling.
-Our aim is to refresh this textbook data to be more up to date.
-Here, we investigate the process of getting from the raw NLSY79 to this textbook data set.
-However, we cannot create the exact same data set as published in their book since we do not have the information of what age threshold they used to determine the high school dropouts.
-
-This paper is structured to have 5 sections.
-Section \ref{database} describes the original data source.
-Section \ref{cleaning} presents the steps of cleaning the data, including getting and tidying the data from the NLSY79 and initial data analysis to find and treat the anomalies in the data.
-We also gave examples of exploratory data analysis using the clean data in Section \ref{eda}.
-Finally, Section \ref{summary} summarises the paper.
-
-\hypertarget{database}{%
-\section{The NLSY79}\label{database}}
-
-\hypertarget{database-1}{%
-\subsection{Database}\label{database-1}}
-
-The NLSY79 is a longitudinal survey held by the U.S Bureau of Labor Statistics that follows the lives of a sample of American youth and born between 1957-1964 (The U.S.
-Bureau of Labor Statistics, n.d.).
-The cohort originally included 12,686 respondents ages 14-22 when first interviewed in 1979.
-It was comprised of Blacks, Hispanics, economically disadvantaged non-Black non-Hispanics, and youth in the military.
-In 1984 and 1990, two sub-samples were dropped from the interview; the dropped subjects are the 1,079 members of the military sample and 1,643 members of the economically disadvantaged non-Black non-Hispanics, respectively.
-Hence, 9,964 respondents remain in the eligible samples.
-The surveys were conducted annually from 1979 to 1994 and biennially thereafter.
-Data are now available from Round 1 (1979 survey year) to Round 28 (2018 survey year).
-
-Although the main focus area of the NLSY is labor and employment, the NLSY also covers several other topics, including education; training and achievement; household, geography, and contextual variables; dating, marriage, cohabitation; sexual activity, pregnancy, and fertility; children; income, assets and program participation; health; attitudes and expectations; and crime and substance use.
-
-There are two ways to conduct the interview of the NLSY79, which are face-to-face interviews or by telephone.
-In recent survey years, more than 90 percent of respondents were interviewed by telephone (Cooksey 2017).
-
-\hypertarget{target-data}{%
-\subsection{Target Data}\label{target-data}}
-
-As mentioned in Section \ref{intro}, this paper aims to refresh the NLSY79 data as used in Singer and Willet's textbook (Singer and Willett (2003)).
-This data set contains the information of yearly mean hourly wages from the NLSY79 cohort, with education and race as covariates.
-This data set represents the measurement of male high-school dropouts, aged from 14 to 17 years old when first time measured.
-Accordingly, to get this data, we have two additional target data, NSLY79 cohort's demographic profiles and wages data of people who completed education up to high school.
-
-Since there is no information on high school dropouts in Singer and Willet's book, we define high school dropouts as respondents who only completed either \(9^{th}\), \(10^{th}\), \(11^{th}\) grade, or who completed \(12^{th}\) when their age is at least 19 years old (means that these IDs being dropped out in a certain year, but they returned back to high school to complete it).
-
-\hypertarget{cleaning}{%
-\section{The NLSY79 Data Cleaning}\label{cleaning}}
-
-\hypertarget{getting-and-tidying-the-data}{%
-\subsection{Getting and Tidying the Data}\label{getting-and-tidying-the-data}}
-
-The NLSY79 data are stored in a \href{https://www.nlsinfo.org/content/cohorts/nlsy79/get-data}{database} and could be downloaded by variables.
-Several variables are available for download and could be browsed by index.
-For the wages data set, we only extracted these variables:
-
-\textbf{Education, Training \& Achievement Scores}
-
-\begin{itemize}
-\tightlist
-\item
-  Education -\textgreater{} Summary measures -\textgreater{} All schools -\textgreater{} By year -\textgreater{} Highest grade completed
-
-  \begin{itemize}
-  \tightlist
-  \item
-    Downloaded all of variables in Highest grade completed.
-  \end{itemize}
-\end{itemize}
-
-\textbf{Employment}
-
-\begin{itemize}
-\tightlist
-\item
-  Summary measures -\textgreater{} By job -\textgreater{} Hours worked and Hourly wages
-
-  \begin{itemize}
-  \tightlist
-  \item
-    Downloaded all of the variables in Hours worked.
-  \item
-    Downloaded all of the variables in Hourly wages
-    Both hours worked and hourly wages are recorded by the job, up to five jobs for each id/subject.
-  \end{itemize}
-\end{itemize}
-
-\textbf{Household, Geography \& Contextual Variables}
-
-\begin{itemize}
-\tightlist
-\item
-  Context -\textgreater{} Summary measures -\textgreater{} Basic demographics
-
-  \begin{itemize}
-  \tightlist
-  \item
-    Downloaded year and month of birth, race, and sex variables.
-    There are two versions of the year and month of birth, i.e., 1979 and 1981 data. We downloaded these two versions.
-  \end{itemize}
-\end{itemize}
-
-The downloaded data set came in a csv (NLSY79.csv) and dat (NLSY79.dat) format. We only used the .dat format. It also came along with these files:
-
-\begin{itemize}
-\tightlist
-\item
-  NLSY79.NLSY79: Tagset of variables that can be uploaded to the website to recreate the data set.
-\item
-  NLSY79.R: R script provided automatically by the database for reading the data into R and converting the variables' names and label into something more sensible.
-We utilized this code to do an initial data tidying.
-It produced two data set, \texttt{categories\_qnames} (the observations are stored in categorical/interval values) and \texttt{new\_data\_qnames} (the observations are stored in integer form).
-\end{itemize}
-
-According to Wickham (2014), tidy data sets comply with three rules.
-Firstly, each variable forms a column.
-Secondly, each observation forms a row.
-Lastly, each type of observational unit forms a table.
-Unfortunately, the \texttt{new\_data\_qnames} did not meet these requirements in the way that data for each year and job for the same respondent is stored as one variable.
-Hence, the data contains a huge number of columns (686 columns).
-The example of the untidy data set is displayed in Table \ref{tab:untidy-data}.
-The table intended to display the hourly rate of each respondent by job (HRP1 to HRP5) and by year (1979 and 1980).
-The table implies that the column headers are values of the year and job, not variable names.
-Consequently, the data should be tidied and wrangled first to extract the demographic and employment variables we want to put in the final data set.
-We mainly used \texttt{tidyr} (Wickham 2020b), \texttt{dplyr} (Wickham et al.
-2020), and \texttt{stringr} (Wickham 2019) to do this job.
-
-\begin{table}
-
-\caption{\label{tab:untidy-data}The Untidy Form of the NLSY79 Raw Data}
-\centering
-\begin{tabular}[t]{r|r|r|r|r|r|r}
-\hline
-CASEID\_1979 & HRP1\_1979 & HRP2\_1979 & HRP3\_1979 & HRP4\_1979 & HRP5\_1979 & HRP1\_1980\\
-\hline
-1 & 328 & NA & NA & NA & NA & NA\\
-\hline
-2 & 385 & NA & NA & NA & NA & 457\\
-\hline
-3 & 365 & NA & 275 & NA & NA & 397\\
-\hline
-4 & NA & NA & NA & NA & NA & NA\\
-\hline
-5 & 310 & 375 & NA & NA & NA & 333\\
-\hline
-6 & NA & NA & NA & 250 & NA & 275\\
-\hline
-\end{tabular}
-\end{table}
-
-\hypertarget{tidying-demographic-variables}{%
-\subsubsection{Tidying demographic variables}\label{tidying-demographic-variables}}
-
-\texttt{Age\ in\ 1979}, \texttt{gender}, \texttt{race}, \texttt{highest\ grade\ completed}, and the \texttt{year\ when\ the\ highest\ grade\ completed} are the variables that we want to use in the cleaned data set.
-
-Age in 1979 are derived from year and month of birth variables in the raw data.
-The variables have two versions, which are the 1979 version and the 1981 version.
-We only used the 1979 data and did a consistency check of those years and flag the inconsistent observations.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\CommentTok{\#tidy the date of birth data}
-\NormalTok{dob\_tidy }\OtherTok{\textless{}{-}}\NormalTok{ new\_data\_qnames }\SpecialCharTok{\%\textgreater{}\%}
-\NormalTok{  dplyr}\SpecialCharTok{::}\FunctionTok{select}\NormalTok{(CASEID\_1979,}
-                \FunctionTok{starts\_with}\NormalTok{(}\StringTok{"Q1{-}3\_A\textasciitilde{}"}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{dob\_year =} \FunctionTok{case\_when}\NormalTok{(}
-    \CommentTok{\# if the years recorded in both sets match, take 79 data}
-    \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}Y\_1979}\StringTok{\textasciigrave{}} \SpecialCharTok{==} \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}Y\_1981}\StringTok{\textasciigrave{}} \SpecialCharTok{\textasciitilde{}} \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}Y\_1979}\StringTok{\textasciigrave{}}\NormalTok{,}
-    \CommentTok{\# if the year in the 81 set is missing, take the 79 data}
-    \FunctionTok{is.na}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}Y\_1981}\StringTok{\textasciigrave{}}\NormalTok{) }\SpecialCharTok{\textasciitilde{}} \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}Y\_1979}\StringTok{\textasciigrave{}}\NormalTok{,}
-    \CommentTok{\# if the sets don\textquotesingle{}t match for dob year, take the 79 data}
-    \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}Y\_1979}\StringTok{\textasciigrave{}} \SpecialCharTok{!=} \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}Y\_1981}\StringTok{\textasciigrave{}} \SpecialCharTok{\textasciitilde{}} \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}Y\_1979}\StringTok{\textasciigrave{}}\NormalTok{),}
-    \AttributeTok{dob\_month =} \FunctionTok{case\_when}\NormalTok{(}
-      \CommentTok{\# if months recorded in both sets match, take 79 data}
-      \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}M\_1979}\StringTok{\textasciigrave{}} \SpecialCharTok{==} \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}M\_1981}\StringTok{\textasciigrave{}} \SpecialCharTok{\textasciitilde{}} \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}M\_1979}\StringTok{\textasciigrave{}}\NormalTok{,}
-      \CommentTok{\# if month in 81 set is missing, take the 79 data}
-      \FunctionTok{is.na}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}M\_1981}\StringTok{\textasciigrave{}}\NormalTok{) }\SpecialCharTok{\textasciitilde{}} \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}M\_1979}\StringTok{\textasciigrave{}}\NormalTok{,}
-      \CommentTok{\# if sets don\textquotesingle{}t match for dob month, take the 79 data}
-      \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}M\_1979}\StringTok{\textasciigrave{}} \SpecialCharTok{!=} \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}M\_1981}\StringTok{\textasciigrave{}} \SpecialCharTok{\textasciitilde{}} \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}M\_1979}\StringTok{\textasciigrave{}}\NormalTok{),}
-    \CommentTok{\# flag if there is a conflict between dob recorded in 79 and 81}
-    \AttributeTok{dob\_conflict =} \FunctionTok{case\_when}\NormalTok{(}
-\NormalTok{      (}\StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}M\_1979}\StringTok{\textasciigrave{}} \SpecialCharTok{!=} \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}M\_1981}\StringTok{\textasciigrave{}}\NormalTok{) }\SpecialCharTok{\&} \SpecialCharTok{!}\FunctionTok{is.na}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}M\_1981}\StringTok{\textasciigrave{}}\NormalTok{)}
-      \SpecialCharTok{\textasciitilde{}} \ConstantTok{TRUE}\NormalTok{,}
-\NormalTok{      (}\StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}Y\_1979}\StringTok{\textasciigrave{}} \SpecialCharTok{!=} \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}Y\_1981}\StringTok{\textasciigrave{}}\NormalTok{) }\SpecialCharTok{\&} \SpecialCharTok{!}\FunctionTok{is.na}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}Y\_1981}\StringTok{\textasciigrave{}}\NormalTok{)}
-      \SpecialCharTok{\textasciitilde{}} \ConstantTok{TRUE}\NormalTok{,}
-\NormalTok{      (}\StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}Y\_1979}\StringTok{\textasciigrave{}} \SpecialCharTok{==} \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}Y\_1981}\StringTok{\textasciigrave{}}\NormalTok{) }\SpecialCharTok{\&}
-\NormalTok{        (}\StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}M\_1979}\StringTok{\textasciigrave{}} \SpecialCharTok{==} \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}M\_1981}\StringTok{\textasciigrave{}}\NormalTok{) }\SpecialCharTok{\textasciitilde{}} \ConstantTok{FALSE}\NormalTok{,}
-      \FunctionTok{is.na}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}M\_1981}\StringTok{\textasciigrave{}}\NormalTok{) }\SpecialCharTok{|} \FunctionTok{is.na}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}Y\_1981}\StringTok{\textasciigrave{}}\NormalTok{) }\SpecialCharTok{\textasciitilde{}} \ConstantTok{FALSE}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%}
-\NormalTok{  dplyr}\SpecialCharTok{::}\FunctionTok{select}\NormalTok{(CASEID\_1979,}
-\NormalTok{                dob\_month,}
-\NormalTok{                dob\_year,}
-\NormalTok{                dob\_conflict)}
-\end{Highlighting}
-\end{Shaded}
-
-For \texttt{gender} and \texttt{race}, we only rename the column names.
-It is worth noting that using these two variable needs special attention.
-Gender, as reported in the data, only has two categories, which is recognised today as inadequate.
-Gender is not binary.
-Further, race is as reported in the database.
-When doing analysis with this variable, one should keep in mind that the purpose is to study racism rather than race.
-
-The next step is tidying the \texttt{highest\ grade\ completed} variable.
-This variable came with several version in each year.
-We choose the revised May data because it seemed to have less missing and presumably has been checked.
-However, there is no revised May data for 2012, 2014, 2016, and 2018.
-Thus, we just use the ordinary May data for these years.
-The code for tidying this variable is provided in the supplementary materials.
-
-Further, \texttt{highest\ grade\ completed} is measured and could be updated in each period of the survey.
-Hence, we only used the highest grade completed ever and derived the year when it is completed.
-The code for tidying this variable is also available in the supplementary materials.
-
-Finally, we get all of the demographic profiles of the NLSY79 cohort. We then save this data as \texttt{demog\_nlsy79}.
-
-\hypertarget{tidying-employment-variables}{%
-\subsubsection{Tidying employment variables}\label{tidying-employment-variables}}
-
-The employment data comprises of three variables, i.e., \texttt{total\ hours\ of\ work\ per\ week}, \texttt{number\ of\ jobs\ that\ an\ individual\ has}, and \texttt{mean\ hourly\ wage}.
-
-\texttt{hours\ worked\ per\ week} initially has only one version per job, no choice from 1979 to 1987 (QES-52A).
-From 1988 onward, when we had more options, we chose the variable for total hours, including time spent working from home (QES-52D).
-However, in 1993, this variable did not have all the five D variables (the first one and the last one were missing), so we used QES-52A variable instead.
-In addition, 2008 only had jobs 1-4 for the QES-52D variable (whereas the other years had 1-5), so we use these.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\CommentTok{\# make a list for years where we used the "QES{-}52A"}
-\NormalTok{year\_A }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(}\DecValTok{1979}\SpecialCharTok{:}\DecValTok{1987}\NormalTok{, }\DecValTok{1993}\NormalTok{)}
-\CommentTok{\#function to get the hour of work}
-\NormalTok{get\_hour }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{(year)\{}
-  \ControlFlowTok{if}\NormalTok{(year }\SpecialCharTok{\%in\%}\NormalTok{ year\_A)\{}
-\NormalTok{    temp }\OtherTok{\textless{}{-}}\NormalTok{ new\_data\_qnames }\SpecialCharTok{\%\textgreater{}\%}
-\NormalTok{      dplyr}\SpecialCharTok{::}\FunctionTok{select}\NormalTok{(CASEID\_1979,}
-                    \FunctionTok{starts\_with}\NormalTok{(}\StringTok{"QES{-}52A"}\NormalTok{) }\SpecialCharTok{\&}
-                      \FunctionTok{ends\_with}\NormalTok{(}\FunctionTok{as.character}\NormalTok{(year)))\}}
-  \ControlFlowTok{else}\NormalTok{\{}
-\NormalTok{    temp }\OtherTok{\textless{}{-}}\NormalTok{ new\_data\_qnames }\SpecialCharTok{\%\textgreater{}\%}
-\NormalTok{      dplyr}\SpecialCharTok{::}\FunctionTok{select}\NormalTok{(CASEID\_1979,}
-                    \FunctionTok{starts\_with}\NormalTok{(}\StringTok{"QES{-}52D"}\NormalTok{) }\SpecialCharTok{\&}
-                      \FunctionTok{ends\_with}\NormalTok{(}\FunctionTok{as.character}\NormalTok{(year)))\}}
-\NormalTok{  temp }\SpecialCharTok{\%\textgreater{}\%}
-    \FunctionTok{pivot\_longer}\NormalTok{(}\SpecialCharTok{!}\NormalTok{CASEID\_1979,}
-                 \AttributeTok{names\_to =} \StringTok{"job"}\NormalTok{,}
-                 \AttributeTok{values\_to =} \StringTok{"hours\_work"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
-    \FunctionTok{separate}\NormalTok{(}\StringTok{"job"}\NormalTok{, }\FunctionTok{c}\NormalTok{(}\StringTok{"job"}\NormalTok{, }\StringTok{"year"}\NormalTok{), }\AttributeTok{sep =} \SpecialCharTok{{-}}\DecValTok{4}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
-    \FunctionTok{mutate}\NormalTok{(}\AttributeTok{job =} \FunctionTok{paste0}\NormalTok{(}\StringTok{"job\_"}\NormalTok{, }\FunctionTok{substr}\NormalTok{(job, }\DecValTok{9}\NormalTok{, }\DecValTok{10}\NormalTok{))) }\SpecialCharTok{\%\textgreater{}\%}
-    \FunctionTok{rename}\NormalTok{(}\AttributeTok{id =}\NormalTok{ CASEID\_1979)}
-\NormalTok{\}}
-\CommentTok{\# list to save the iteration result}
-\NormalTok{hours }\OtherTok{\textless{}{-}} \FunctionTok{list}\NormalTok{()}
-\CommentTok{\# getting the hours of work of all observations}
-\ControlFlowTok{for}\NormalTok{(ayear }\ControlFlowTok{in} \FunctionTok{c}\NormalTok{(}\DecValTok{1979}\SpecialCharTok{:}\DecValTok{1994}\NormalTok{, }\DecValTok{1996}\NormalTok{, }\DecValTok{1998}\NormalTok{, }\DecValTok{2000}\NormalTok{, }\DecValTok{2002}\NormalTok{, }\DecValTok{2004}\NormalTok{, }\DecValTok{2006}\NormalTok{, }\DecValTok{2008}\NormalTok{, }\DecValTok{2010}\NormalTok{,}
-               \DecValTok{2012}\NormalTok{, }\DecValTok{2014}\NormalTok{, }\DecValTok{2016}\NormalTok{, }\DecValTok{2018}\NormalTok{)) \{}
-\NormalTok{  hours[[ayear]] }\OtherTok{\textless{}{-}} \FunctionTok{get\_hour}\NormalTok{(ayear)}
-\NormalTok{\}}
-\CommentTok{\# unlist the hours of work}
-\NormalTok{hours\_all }\OtherTok{\textless{}{-}} \FunctionTok{bind\_rows}\NormalTok{(}\SpecialCharTok{!!!}\NormalTok{hours)}
-\end{Highlighting}
-\end{Shaded}
-
-The same way was also deployed to tidy the rate of wage by year and by ID.
-The difference is that the hourly rate has only one version of each year.
-The hours of work and the hourly rate are then joined to calculate the number of jobs that a respondent has and their mean hourly wage.
-Some observations have 0 in their hourly rate and work hours, which is considered an invalid value.
-Thus, these observations are set to be NA.
-Besides, there are some observations with unusual hours of work.
-Here, we also censored the observations with \textgreater{} 84 hours of work a week to be NA.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{get\_rate }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{(year) \{}
-\NormalTok{  new\_data\_qnames }\SpecialCharTok{\%\textgreater{}\%}
-\NormalTok{    dplyr}\SpecialCharTok{::}\FunctionTok{select}\NormalTok{(CASEID\_1979,}
-                  \FunctionTok{starts\_with}\NormalTok{(}\StringTok{"HRP"}\NormalTok{) }\SpecialCharTok{\&}
-                    \FunctionTok{ends\_with}\NormalTok{(}\FunctionTok{as.character}\NormalTok{(year))) }\SpecialCharTok{\%\textgreater{}\%}
-    \FunctionTok{pivot\_longer}\NormalTok{(}\SpecialCharTok{!}\NormalTok{CASEID\_1979, }\AttributeTok{names\_to =} \StringTok{"job"}\NormalTok{, }\AttributeTok{values\_to =} \StringTok{"rate\_per\_hour"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
-    \FunctionTok{separate}\NormalTok{(}\StringTok{"job"}\NormalTok{, }\FunctionTok{c}\NormalTok{(}\StringTok{"job"}\NormalTok{, }\StringTok{"year"}\NormalTok{), }\AttributeTok{sep =} \SpecialCharTok{{-}}\DecValTok{4}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
-    \FunctionTok{mutate}\NormalTok{(}\AttributeTok{job =} \FunctionTok{paste0}\NormalTok{(}\StringTok{"job\_0"}\NormalTok{, }\FunctionTok{substr}\NormalTok{(job, }\DecValTok{4}\NormalTok{, }\DecValTok{4}\NormalTok{))) }\SpecialCharTok{\%\textgreater{}\%}
-    \FunctionTok{rename}\NormalTok{(}\AttributeTok{id =}\NormalTok{ CASEID\_1979)}
-\NormalTok{\}}
-\NormalTok{rates }\OtherTok{\textless{}{-}} \FunctionTok{list}\NormalTok{()}
-\ControlFlowTok{for}\NormalTok{(ayear }\ControlFlowTok{in} \FunctionTok{c}\NormalTok{(}\DecValTok{1979}\SpecialCharTok{:}\DecValTok{1994}\NormalTok{, }\DecValTok{1996}\NormalTok{, }\DecValTok{1998}\NormalTok{, }\DecValTok{2000}\NormalTok{, }\DecValTok{2002}\NormalTok{, }\DecValTok{2004}\NormalTok{, }\DecValTok{2006}\NormalTok{, }\DecValTok{2008}\NormalTok{, }\DecValTok{2010}\NormalTok{,}
-               \DecValTok{2012}\NormalTok{, }\DecValTok{2014}\NormalTok{, }\DecValTok{2016}\NormalTok{, }\DecValTok{2018}\NormalTok{)) \{}
-\NormalTok{  rates[[ayear]] }\OtherTok{\textless{}{-}} \FunctionTok{get\_rate}\NormalTok{(ayear)}
-\NormalTok{\}}
-\NormalTok{rates\_all }\OtherTok{\textless{}{-}} \FunctionTok{bind\_rows}\NormalTok{(}\SpecialCharTok{!!!}\NormalTok{rates)}
-\CommentTok{\# join hours and rates variable}
-\NormalTok{hours\_wages }\OtherTok{\textless{}{-}} \FunctionTok{left\_join}\NormalTok{(rates\_all,}
-\NormalTok{                         hours\_all,}
-                         \AttributeTok{by =} \FunctionTok{c}\NormalTok{(}\StringTok{"id"}\NormalTok{, }\StringTok{"year"}\NormalTok{, }\StringTok{"job"}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%}
-  \CommentTok{\# set the 0 value in rate\_per\_hour as NA}
-  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{rate\_per\_hour =} \FunctionTok{ifelse}\NormalTok{(rate\_per\_hour }\SpecialCharTok{==} \DecValTok{0}\NormalTok{, }\ConstantTok{NA}\NormalTok{,}
-\NormalTok{                                rate\_per\_hour),}
-         \AttributeTok{hours\_work =} \FunctionTok{ifelse}\NormalTok{(hours\_work }\SpecialCharTok{==} \DecValTok{0} \SpecialCharTok{|}\NormalTok{ hours\_work }\SpecialCharTok{\textgreater{}} \DecValTok{84}\NormalTok{, }\ConstantTok{NA}\NormalTok{, hours\_work))}
-\end{Highlighting}
-\end{Shaded}
-
-Since our ultimate goal is to calculate the mean hourly wage, the number of jobs is calculated based on the availability of the \texttt{rate\_per\_hour} information.
-For example, the number of jobs of ID 1, based on \texttt{hours\_work}, is 2.
-However, since the information of hourly rate of \texttt{job\_02} is not available, the number of jobs is considered as 1.
-
-Further, we calculated the mean hourly wage for each ID in each year using a weighted mean with the hours of work as the weight.
-However, there is a lot of missing value in \texttt{hours\_work} variable.
-In that case, we only calculated the mean hourly wage based on the arithmetic/regular mean method.
-Hence, we created a new variable to flag whether the mean hourly wage is weighted or regular.
-Additionally, if an ID only had one job, we directly used their hourly wages information and flagged it as arithmetic mean.
-
-\begin{verbatim}
-#> # A tibble: 10 x 6
-#>       id  year mean_hourly_wage total_hours number_of_jobs is_wm
-#>    <int> <dbl>            <dbl>       <int>          <dbl> <lgl>
-#>  1     1  1979             3.28          38              1 FALSE
-#>  2     1  1981             3.61          NA              1 FALSE
-#>  3     2  1979             3.85          35              1 FALSE
-#>  4     2  1980             4.57          NA              1 FALSE
-#>  5     2  1981             5.14          NA              1 FALSE
-#>  6     2  1982             5.71          35              1 FALSE
-#>  7     2  1983             5.71          NA              1 FALSE
-#>  8     2  1984             5.14          NA              1 FALSE
-#>  9     2  1985             7.71          NA              1 FALSE
-#> 10     2  1986             7.69          NA              1 FALSE
-\end{verbatim}
-
-The employment and demographic variables are then joined.
-We also filtered the data to only have the cohort who completed the education up to 12th grade and participated at least five rounds in the survey and save it to an object called \texttt{wages}.
-
-\hypertarget{initial-data-analysis}{%
-\subsection{Initial Data Analysis}\label{initial-data-analysis}}
-
-According to Huebner Marianne, Vach, and Cessie (2016), Initial Data Analysis (IDA) is the step of inspecting and screening the data after being collected to ensure that the data is clean, valid, and ready to be deployed in the later formal statistical analysis.
-Moreover, Chatfield (1985) argues that the two main objectives of IDA are data description, which is to assess the structure and the quality of the data, and model formulation without any formal statistical inference.
-
-In this paper, we conduct an IDA or a preliminary data analysis to assess the consistency of the data with the cohort information that the NLSY provides.
-In addition, we also aim to find the anomaly in the wages values using this approach.
-We mainly use graphical summaries to do the IDA using \texttt{ggplot2}(Wickham 2016) and \texttt{brolgar} (Tierney, Cook, and Prvan 2020).
-
-As stated previously, the respondents' ages ranged from 12 to 22 when first interviewed in 1979.
-Hence, we validate whether all of the respondents were in this range.
-Additionally, the \href{https://www.nlsinfo.org/content/cohorts/nlsy79/intro-to-the-sample/nlsy79-sample-introduction}{NLSY} also provides the number of the survey cohort by their gender (6,403 males and 6,283 females) and race (7,510 Non-Black/Non-Hispanic; 3,174 Black; 2,002 Hispanic).
-To validate this, we used the \texttt{demog\_nlsy79}, i.e., the data with the survey years 1979 sample.
-Table \ref{tab:age-table} and Table \ref{tab:gender-race-table} suggest that the demographic data we had is consistent with the sample information in the database.
-
-\begin{table}
-
-\caption{\label{tab:age-table}Age Distribution of the NLSY79 samples}
-\centering
-\begin{tabular}[t]{r|r}
-\hline
-Age & Number of Sample\\
-\hline
-15 & 1265\\
-\hline
-16 & 1550\\
-\hline
-17 & 1600\\
-\hline
-18 & 1530\\
-\hline
-19 & 1662\\
-\hline
-20 & 1722\\
-\hline
-21 & 1677\\
-\hline
-22 & 1680\\
-\hline
-\end{tabular}
-\end{table}
-
-\begin{table}
-
-\caption{\label{tab:gender-race-table}Gender and Race Distribution of the NLSY79 Samples}
-\centering
-\begin{tabular}[t]{l|l|l|l|l}
-\hline
-\multicolumn{1}{c|}{ } & \multicolumn{3}{c|}{Race} & \multicolumn{1}{c}{ } \\
-\cline{2-4}
-Gender & Hispanic & Black & Non-Black, Non-Hispanic & Total\\
-\hline
-Male & 1000 (15.62\%) & 1613 (25.19\%) & 3790 (59.19\%) & 6403 (100.00\%)\\
-\hline
-Female & 1002 (15.95\%) & 1561 (24.84\%) & 3720 (59.21\%) & 6283 (100.00\%)\\
-\hline
-Total & 2002 (15.78\%) & 3174 (25.02\%) & 7510 (59.20\%) & 12686 (100.00\%)\\
-\hline
-\end{tabular}
-\end{table}
-
-The next step is that explore the mean hourly wage data.
-In this case, we only explore the wages data of up to \(12^{th}\) grade.
-Table \ref{tab:summarytable} shows that the median of the overall wage of the cohort is only 7.2, while the mean is 11.87.
-It indicates that the data might contains a lot of extreme values.
-
-\begin{table}
-
-\caption{\label{tab:summarytable}Summary Statistics of Wages of High School Data}
-\centering
-\begin{tabular}[t]{l|r}
-\hline
-Statistics & Value\\
-\hline
-Min. & 0.01000\\
-\hline
-1st Qu. & 4.50000\\
-\hline
-Median & 7.20000\\
-\hline
-Mean & 11.86593\\
-\hline
-3rd Qu. & 11.74000\\
-\hline
-Max. & 33400.00000\\
-\hline
-\end{tabular}
-\end{table}
-
-However, we can not be sure only by observing the summary statistics.
-Hence, we use visualisation techniques to investigate this matter.
-Figure \ref{fig:feature-plot} conveys a problem in the mean hourly wage values.
-Figure \ref{fig:feature-plot} A shows that some observations have an exceptionally high figure of wages, even more than US\$10,000 per hour.
-In Figure \ref{fig:feature-plot} B, we barely see any difference in the minimum, median, and maximum value of the wages since the distribution is heavily skewed to the right.
-
-\begin{figure}
-
-{\centering \includegraphics[width=0.7\linewidth]{figures/feature-plot-1}
-
-}
-
-\caption{Two plots showing the distribution of the mean hourly wage.
-Plot A portrays the pattern of mean hourly wage of high school cohort from 1979 to 2018 of each ID in US Dollar; Plot B shows the distribution of their minimum, median, and maximum value.
-We can see that some IDs had an extremely high of wages and it made the distribution of the three features is extremely skewed.}\label{fig:feature-plot}
-\end{figure}
-
-In Figure \ref{fig:high-wages}, we plot some respondents with a high value of mean hourly wages.
-We filter all of the IDs who earned more than US\$ 500 per hour on average.
-We find that these respondents only experienced one point of extremely high wages.
-Thus, we suspect that these high values are erroneous resulted from a data entry error.
-
-\begin{figure}
-
-{\centering \includegraphics[width=0.7\linewidth]{figures/high-wages-1}
-
-}
-
-\caption{6 out of 45 IDs with extremely high mean hourly wage. Most of the IDs only have one point of high wage.}\label{fig:high-wages}
-\end{figure}
-
-Further, we take 36 samples randomly from the data and plot them, as shown in Figure \ref{fig:sample-plot}.
-It implies that not only that some observations earn extremely high figures of wages, but some also have reasonably fluctuated wages, for example, the IDs in panel numbers 5, 7, and 11.
-The plot also implies that the samples have a different pattern of mean hourly wages.
-Some have had flat wages for years but had a sudden increase in one particular year, then it gone down again, while the others experienced an upsurge in their wage, for instance, the IDs in panel 9.
-
-\begin{figure}
-
-{\centering \includegraphics[width=0.7\linewidth]{figures/sample-plot-1}
-
-}
-
-\caption{The mean hourly wages of some random samples are shown in twelve facets, three IDs per facet. It suggests that some IDs had a reasonably fluctuate wages.}\label{fig:sample-plot}
-\end{figure}
-
-The anomalies are also found in the total hours of work, where some observations work for 420 hours a week in total.
-According to Pergamit et al.
-(2001), one of the flaws of the NLSY79 employment data is that the NLSY79 collects the information of the working hours since the last interview.
-Thus, it might be challenging for the respondents to track the within-job hours' changes between survey years, especially for the respondents with fluctuated working hours or seasonal jobs.
-It even has been more challenging since 1994, where the respondents had to recall two years periods.
-This shortcoming might also contribute to the fluctuation of one's wages data.
-
-\hypertarget{replacing-extreme-values}{%
-\subsubsection{Replacing extreme values}\label{replacing-extreme-values}}
-
-As part of the IDA, which is the model formulation, we build a robust linear regression model to treat the extreme values in the data.
-Robust linear regression yields an estimation robust to the influence of noise or contamination (Koller 2016).
-It also aims to detect the contamination by weighting each observation based on how ``well-behaved'' they are, known as robustness weight.
-Observations with lower robustness weight are suggested as an outlier by this method (Koller 2016).
-
-Since we work with longitudinal data, we build the model for each ID instead of the overall data.
-The robust mixed model could be the best model to be employed in this case.
-However, this method is too computationally and memory expensive, especially for a large data set, like the NLSY79 data.
-Thus, the model for each ID is built utilizing the \texttt{nest} and \texttt{map} function from \texttt{tidyr} (Wickham 2020b) and \texttt{purrr} (Henry and Wickham 2020), respectively.
-
-We build the model using the \texttt{rlm} function from \texttt{MASS} package (Venables and Ripley 2002).
-We set the \texttt{mean\_hourly\_wage} and \texttt{year} as the dependent and predictor, respectively.
-Furthermore, we use M-Estimation with Huber weighting, where the observation with a slight residual gets a weight of 1, while the larger the residual, the smaller the weight (less than 1) (UCLA: Statistical Consulting Group 2021).
-However, the challenging part of detecting the anomaly using the robustness weight is determining the weight threshold in which the observations are considered outliers.
-Moreover, it should be noted that not all the outliers are due to an error.
-Instead, it might be that one had reasonably increasing or decreasing wages in a particular period.
-
-To minimize the risk of mistakenly identifying an outlier as an ``erroneous outlier,'' we simulate some thresholds and study how they affect the data.
-We find that 0.12 is the most reasonable value to be the threshold to minimize that drawback's risk because it still captures the sensible spikes in the data.
-In other words, we keep maintaining the natural variability of the wages while minimizing anomalies because of the error in the data recording.
-After deciding the threshold, we impute the observations whose weights are less than 0.12 with the models' predicted value.
-We then flag those observations in a new variable called \texttt{is\_pred}.
-
-Figure \ref{fig:compare-plot} shows the mean hourly wage before and after the extreme values are replaced.
-It implies that the fluctuation can still be observed in the data after the treatment.
-However, the large spikes, which are considered ``erroneous outliers,'' are already eliminated from the data.
-Hence, the model produces a data set with a more reasonable degree of fluctuation.
-
-\begin{figure}
-
-{\centering \includegraphics[width=0.7\linewidth]{figures/compare-plot-1}
-
-}
-
-\caption{Comparison between the original and the treated mean hourly wage.
-The orange line portray the original value of mean hourly wage, while the turquoise line display the mean hourly wages value after the extreme values imputed with the robust linear model's prediction value.
-We can see that some extreme spikes has been reduced by the model.}\label{fig:compare-plot}
-\end{figure}
-
-Further, \ref{fig:comppict} A shows that after eliminating the extreme values, the highest value has decreased to be around \$350.
-The spikes are still observed but not as extreme as the original data set.
-In Figure \ref{fig:comppict} B), we plot the three features of mean hourly wages, namely the minimum, median, and maximum value, transformed to log scale.
-The plot implies that the skewness is slightly negative.
-We also see that the three features are overlapped each other.
-It indicates that some ID's minimum wages are higher than some ID's maximum wages.
-
-\begin{figure}
-
-{\centering \includegraphics[width=0.7\linewidth]{figures/comppict-1}
-
-}
-
-\caption{The distribution of the mean hourly wage after the extreme values are replaced.
-Plot A portrays the pattern of mean hourly wage of high school cohort from 1979 to 2018 of each ID in US Dollar; Plot B shows the distribution of their minimum, median, and maximum value transformed to log10 scale.
-We can see that some observations still had reasonbaly higher wages than the others.
-Also, some IDs' have a minimum wages that is higher than others' maximum wages.}\label{fig:comppict}
-\end{figure}
-
-Finally, we save the imputed data and set the appropriate data type for the variables.
-As our target data is the mean hourly wage of the high-school dropouts, we then subset the high-school graduate data set to have only the male-high school dropout data.
-We then make these three data sets and their processing documentation publicly available through an R data container package called \texttt{yowie}.
-The complete flow from the raw data to these data set is displayed in Figure \ref{fig:flow-chart}.
-
-\begin{figure}
-
-{\centering \includegraphics[width=0.7\linewidth]{figures/flow-chart-1}
-
-}
-
-\caption{The stages of data filtering from the raw data to get three datasets contained yowie, n means the number of ID, while n\_obs means the number of observations.}\label{fig:flow-chart}
-\end{figure}
-
-\hypertarget{eda}{%
-\section{Exploratory Data Analysis}\label{eda}}
-
-This section will provide an example of how data sets could be used to do exploratory data analysis as in textbook data.
-We are interested in examining how the respondents experience changes in their wages based on their demographic characteristics.
-Thus, we use \texttt{wages} data instead of the refreshed \texttt{wages\_hs\_do} as it only has male respondents.
-
-The statistics we use in measuring the change in wages experience is the slope of the wages of each respondent.
-Using \texttt{key\_slope} function in \texttt{brolgar} (Tierney, Cook, and Prvan 2020), we perform linear regression with the log of mean hourly wage as the response variable and year as the predictor.
-Since we do modeling, it is important to look at the number of observations from each respondent.
-Figure \ref{fig:slope-nobs-plot} shows that the number of observations for each individual varies.
-Moreover, each number of observations seems to have an impact on the slope.
-Extreme values in slope are observed in a few observations, while the range of slope is more stable with more observations.
-Accordingly, we trim the data to only include \(5^{th} - 95^{th}\) percentile.
-
-\begin{figure}
-
-{\centering \includegraphics[width=0.7\linewidth]{figures/slope-nobs-plot-1}
-
-}
-
-\caption{A) Distribution of observations.
-The numbers of rounds varies among respondents.
-Most of respondents participated in 28 rounds of survey.
-B) Distribution of wages slope by number of observations.
-The slopes' range are more stable with more observations.}\label{fig:slope-nobs-plot}
-\end{figure}
-
-We then create some visualizations of five respondents with the highest slope and five respondents with the lowest slope, henceforth referred to as top-5 slopes and bottom-5 slopes, respectively.
-We plot them based on their gender and education.
-Figure \ref{fig:slope-gender} shows top-5 slopes are dominated by males, while females dominate bottom-5 slopes.
-However, as this figure only shows a small portion of the data, we further inspect the gender wage gap.
-
-\begin{figure}
-
-{\centering \includegraphics[width=0.7\linewidth]{figures/slope-gender-1}
-
-}
-
-\caption{Wages slope over the years based on gender.
-Most of respondents who experienced high wages slope are males.
-In contrast, respondents who experienced lower wages slope are mostly females.}\label{fig:slope-gender}
-\end{figure}
-
-According to OECD (n.d.), the gender pay gap is the difference between the median wage of males and females relative to the median wage of males.
-We find that, as shown in \ref{fig:gap-gender}, over the years, females tend to earn fewer than males.
-For example, in 2018, females earned about 78 cents for every \$1 earned by males.
-The widest gender gaps were observed in the 1990s.
-
-\begin{figure}
-
-{\centering \includegraphics[width=0.9\linewidth]{figures/gap-gender-1}
-
-}
-
-\caption{A) Number of observations by gender and by year.
-There are more males participated in survey than females.
-The number of observations decreased over the years.
-B) Gender wage gap or the median wages of females relatives to median wages of males.
-Females tend to earn fewer than males.
-For example, in 2018 females earn about 78 cents for every \$1 earned by males.
-Note that this is a crude measure as gap is calculated not based on other variable, such as experience and job class.}\label{fig:gap-gender}
-\end{figure}
-
-According to their highest education completed, both top-5 and bottom-5 slopes are dominated by people who completed high school.
-This is understandable since the respondents are mostly completed high school.
-
-\begin{figure}
-
-{\centering \includegraphics[width=0.7\linewidth]{figures/slope-educ-1}
-
-}
-
-\caption{Wages slope over the years based on highest education completed. People who are in top-5 and bottom-5 wages are dominated by people who completed high school.}\label{fig:slope-educ}
-\end{figure}
-
-\hypertarget{summary}{%
-\section{Summary}\label{summary}}
-
-This paper has performed stages to make open data suitable for textbook data or make it ready for research.
-In the first stage, we showed the steps performed to get the data from the NLSY79 database.
-Since the data format is untidy, we showed how the data had been tidied.
-After that, we conducted an initial data analysis to investigate and screen the quality of the data.
-We found and fixed the anomalous observations in the data set using the robust linear regression model with its predicted values.
-We also performed an example of exploratory data analysis using the cleaned data set.
-
-This paper has also demonstrated data cleaning documentation by providing all of the codes in performing data tidying and initial data analysis.
-Thus, this paper provided an opportunity to continuously refresh the textbook data whenever the updated data is published in the NLSY79 database.
-It could be done by following the documentation of the code that is provided in this paper.
-
-Moreover, the documentation also includes how we generated the robustness weight and how we decided the threshold of anomalous observations.
-It is also documented along with the flag of whether an observation is an imputed value or not.
-Accordingly, if somebody wishes to make another decision, it can be done by making a small change in the code provided.
-Further, this paper is also supplemented by a \texttt{shiny} (Chang et al.
-2020) app as a simulation tool to customize the weight threshold.
-
-Finally, this paper implies that data providers should design a database that is able to produce tidy data sets.
-A data provider should also check for data anomalies before the data publishing or at least provides a set of rules or threshold values.
-For example, in this case, is the threshold of reasonable wages.
-This will greatly support the data users to validate and set the same understanding of which data are considered outliers.
-Moreover, providing validation rules would facilitate any established data validation tool, such as \texttt{validate} (van der Loo and de Jonge 2021) package.
-In this case, we cannot use this handy validation package due to the absence of validation rules.
-
-\hypertarget{acknowledgements}{%
-\section{Acknowledgements}\label{acknowledgements}}
-
-We would like to thank Aarathy Babu for the insight and discussion during the writing of this paper.
-
-The entire analysis is conducted using \texttt{R} (R Core Team 2020) in \texttt{rstudio} using these packages: \texttt{tidyverse} (Wickham et al.
-2019), \texttt{ggplot2} (Wickham 2016), \texttt{dplyr} (Wickham et al.
-2020), \texttt{readr} (Wickham and Hester 2020), \texttt{tidyr} (Wickham 2020b), \texttt{stringr} (Wickham 2019), \texttt{purrr} (Henry and Wickham 2020), \texttt{broom} (Robinson, Hayes, and Couch 2021), \texttt{blorgar} (Tierney, Cook, and Prvan 2020), \texttt{patchwork} (Pedersen 2020), \texttt{kableExtra} (Zhu 2019), \texttt{MASS} (Venables and Ripley 2002), \texttt{janitor} (Firke 2020), \texttt{DiagrammeR} (Iannone 2020), \texttt{rsvg} (Ooms 2020), \texttt{webshot} (Chang 2019), \texttt{mgcv} (Wood 2003), \texttt{tsibble} (Wang, Cook, and Hyndman 2020), and \texttt{modelr} (Wickham 2020a).
-The paper are generated using \texttt{knitr} (Xie 2014), \texttt{rmarkdown} (Xie, Dervieux, and Riederer 2020), and \texttt{rticles} (Allaire et al.
-2021).
-
-\hypertarget{supplementary-materials}{%
-\section{Supplementary Materials}\label{supplementary-materials}}
-
-\begin{itemize}
-\item
-  \textbf{Codes} : R script to reproduce data tidying and cleaning are available in this \href{https://github.com/numbats/yowie/blob/master/data-raw/data_preprocessing.R}{Github Repository}.
-\item
-  \textbf{R Package \texttt{yowie}}:\texttt{yowie} is a data container R package that contains 3 datasets, namely the high school mean hourly wage data, high school dropouts mean hourly wage data, and demographic data of the NLSY79 cohort.
-This package could be accessed \href{https://github.com/numbats/yowie}{here}.
-\item
-  \textbf{shiny app}: An web interactive \texttt{shiny} app to run a simulation to customize the weight threshold.
-This app could be accessed \href{https://github.com/numbats/summer-wages-refresh/tree/main/app}{here}.
-\end{itemize}
-
-\hypertarget{references}{%
-\section*{References}\label{references}}
-\addcontentsline{toc}{section}{References}
-
-\hypertarget{refs}{}
-\begin{CSLReferences}{1}{0}
-\leavevmode\hypertarget{ref-rticles}{}%
-Allaire, JJ, Yihui Xie, R Foundation, Hadley Wickham, Journal of Statistical Software, Ramnath Vaidyanathan, Association for Computing Machinery, et al.
-2021.
-\emph{Rticles: Article Formats for r Markdown}.
-\url{https://CRAN.R-project.org/package=rticles}.
-
-\leavevmode\hypertarget{ref-webshot}{}%
-Chang, Winston. 2019. \emph{Webshot: Take Screenshots of Web Pages}. \url{https://CRAN.R-project.org/package=webshot}.
-
-\leavevmode\hypertarget{ref-shiny}{}%
-Chang, Winston, Joe Cheng, JJ Allaire, Yihui Xie, and Jonathan McPherson. 2020. \emph{Shiny: Web Application Framework for r}. \url{https://CRAN.R-project.org/package=shiny}.
-
-\leavevmode\hypertarget{ref-Chatfield1985TIEo}{}%
-Chatfield, C. 1985. {``The Initial Examination of Data.''} \emph{Journal of the Royal Statistical Society. Series A. General} 148 (3): 214--53.
-
-\leavevmode\hypertarget{ref-eliznlsy}{}%
-Cooksey, Elizabeth C.
-2017.
-{``Using the National Longitudinal Surveys of Youth (NLSY) to Conduct Life Course Analyses.''} In \emph{Handbook of Life Course Health Development}, edited by Richard M.
-Lerner Neal Halfon Christoper B.
-Forrest, 561--77.
-Cham: Springer.
-https://doi.org/\url{https://doi.org/10.1007/978-3-319-47143-3_23}.
-
-\leavevmode\hypertarget{ref-DasuTamraparni2003Edma}{}%
-Dasu, Tamraparni, and Theodore Johnson. 2003. \emph{Exploratory Data Mining and Data Cleaning}. Wiley Series in Probability and Statistics. Hoboken: WILEY.
-
-\leavevmode\hypertarget{ref-janitor}{}%
-Firke, Sam. 2020. \emph{Janitor: Simple Tools for Examining and Cleaning Dirty Data}. \url{https://CRAN.R-project.org/package=janitor}.
-
-\leavevmode\hypertarget{ref-purrr}{}%
-Henry, Lionel, and Hadley Wickham. 2020. \emph{Purrr: Functional Programming Tools}. \url{https://CRAN.R-project.org/package=purrr}.
-
-\leavevmode\hypertarget{ref-HuebnerMariannePhD2016Asat}{}%
-Huebner, Marianne, Werner Vach, and Saskia le Cessie.
-2016.
-{``A Systematic Approach to Initial Data Analysis Is Good Research Practice.''} \emph{The Journal of Thoracic and Cardiovascular Surgery} 151 (1): 25--27.
-
-\leavevmode\hypertarget{ref-HuebnerMarianne2020Haar}{}%
-Huebner, Marianne, Werner Vach, Saskia le Cessie, Carsten Oliver Schmidt, and Lara Lusa.
-2020.
-{``Hidden Analyses: A Review of Reporting Practice and Recommendations for More Transparent Reporting of Initial Data Analyses.''} \emph{BMC Medical Research Methodology} 20 (1): 61--61.
-
-\leavevmode\hypertarget{ref-DiagrammeR}{}%
-Iannone, Richard. 2020. \emph{DiagrammeR: Graph/Network Visualization}. \url{https://CRAN.R-project.org/package=DiagrammeR}.
-
-\leavevmode\hypertarget{ref-KollerManuel2016rARP}{}%
-Koller, Manuel. 2016. {``Robustlmm: An r Package for Robust Estimation of Linear Mixed-Effects Models.''} \emph{Journal of Statistical Software} 75 (6): 1--24.
-
-\leavevmode\hypertarget{ref-OECD}{}%
-OECD. n.d. {``Gender Wage Gap.''} https://data.oecd.org/earnwage/gender-wage-gap.htm.
-
-\leavevmode\hypertarget{ref-rsvg}{}%
-Ooms, Jeroen. 2020. \emph{Rsvg: Render SVG Images into PDF, PNG, PostScript, or Bitmap Arrays}. \url{https://CRAN.R-project.org/package=rsvg}.
-
-\leavevmode\hypertarget{ref-opendata}{}%
-Open Knowledge Foundation. 2021. {``Open Definition. Defining Open in Open Data, Open Content, and Open Knowledge.''} 2021. \url{http://opendefinition.org/od/2.1/en/}.
-
-\leavevmode\hypertarget{ref-patchwork}{}%
-Pedersen, Thomas Lin. 2020. \emph{Patchwork: The Composer of Plots}. \url{https://CRAN.R-project.org/package=patchwork}.
-
-\leavevmode\hypertarget{ref-MichaelRPergamit2001DWTN}{}%
-Pergamit, Michael R., Charles R.
-Pierret, Donna S.
-Rothstein, and Jonathan R.
-Veum.
-2001.
-{``Data Watch: The National Longitudinal Surveys.''} \emph{The Journal of Economic Perspectives} 15 (2): 239--53.
-
-\leavevmode\hypertarget{ref-R}{}%
-R Core Team. 2020. \emph{R: A Language and Environment for Statistical Computing}. Vienna, Austria: R Foundation for Statistical Computing. \url{https://www.R-project.org/}.
-
-\leavevmode\hypertarget{ref-broom}{}%
-Robinson, David, Alex Hayes, and Simon Couch. 2021. \emph{Broom: Convert Statistical Objects into Tidy Tibbles}. \url{https://CRAN.R-project.org/package=broom}.
-
-\leavevmode\hypertarget{ref-SingerJudithD2003Alda}{}%
-Singer, Judith D, and John B Willett. 2003. \emph{Applied Longitudinal Data Analysis: Modeling Change and Event Occurrence}. Oxford u.a: Oxford Univ. Pr.
-
-\leavevmode\hypertarget{ref-nlsy79}{}%
-The U.S. Bureau of Labor Statistics. n.d. {``National Longitudinal Survey of Youth 1979.''} Available at \url{https://www.nlsinfo.org/content/cohorts/nlsy79} (2021/25/02).
-
-\leavevmode\hypertarget{ref-brolgar}{}%
-Tierney, Nicholas, Di Cook, and Tania Prvan. 2020. \emph{Brolgar: BRowse over Longitudinal Data Graphically and Analytically in r}. \url{https://github.com/njtierney/brolgar}.
-
-\leavevmode\hypertarget{ref-tukey}{}%
-Tukey, John W. (John Wilder). 1977. \emph{Exploratory Data Analysis}. Addison-Wesley Series in Behavioral Science. Reading, Mass.: Addison-Wesley Pub. Co.
-
-\leavevmode\hypertarget{ref-rlm}{}%
-UCLA: Statistical Consulting Group. 2021. {``Robust Regression \textbar{} r Data Analysis Examples.''} February 2021. \url{https://stats.idre.ucla.edu/r/dae/robust-regression/}.
-
-\leavevmode\hypertarget{ref-validate}{}%
-van der Loo, Mark P. J., and Edwin de Jonge. 2021. {``Data Validation Infrastructure for {R}.''} \emph{Journal of Statistical Software} 97 (10): 1--31. \url{https://doi.org/10.18637/jss.v097.i10}.
-
-\leavevmode\hypertarget{ref-mass}{}%
-Venables, W. N., and B. D. Ripley. 2002. \emph{Modern Applied Statistics with s}. Fourth. New York: Springer. \url{http://www.stats.ox.ac.uk/pub/MASS4}.
-
-\leavevmode\hypertarget{ref-tsibble}{}%
-Wang, Earo, Dianne Cook, and Rob J Hyndman.
-2020.
-{``A New Tidy Data Structure to Support Exploration and Modeling of Temporal Data.''} \emph{Journal of Computational and Graphical Statistics} 29 (3): 466--78.
-\url{https://doi.org/10.1080/10618600.2019.1695624}.
-
-\leavevmode\hypertarget{ref-WickhamHadley2014TD}{}%
-Wickham, Hadley. 2014. {``Tidy Data.''} \emph{Journal of Statistical Software} 59 (10): 1--23.
-
-\leavevmode\hypertarget{ref-ggplot2}{}%
----------. 2016. \emph{Ggplot2: Elegant Graphics for Data Analysis}. Springer-Verlag New York. \url{https://ggplot2.tidyverse.org}.
-
-\leavevmode\hypertarget{ref-stringr}{}%
----------. 2019. \emph{Stringr: Simple, Consistent Wrappers for Common String Operations}. \url{https://CRAN.R-project.org/package=stringr}.
-
-\leavevmode\hypertarget{ref-modelr}{}%
----------. 2020a. \emph{Modelr: Modelling Functions That Work with the Pipe}. \url{https://CRAN.R-project.org/package=modelr}.
-
-\leavevmode\hypertarget{ref-tidyr}{}%
----------. 2020b. \emph{Tidyr: Tidy Messy Data}. \url{https://CRAN.R-project.org/package=tidyr}.
-
-\leavevmode\hypertarget{ref-tidyverse}{}%
-Wickham, Hadley, Mara Averick, Jennifer Bryan, Winston Chang, Lucy D'Agostino McGowan, Romain François, Garrett Grolemund, et al.
-2019.
-{``Welcome to the {tidyverse}.''} \emph{Journal of Open Source Software} 4 (43): 1686.
-\url{https://doi.org/10.21105/joss.01686}.
-
-\leavevmode\hypertarget{ref-dplyr}{}%
-Wickham, Hadley, Romain François, Lionel Henry, and Kirill Müller. 2020. \emph{Dplyr: A Grammar of Data Manipulation}. \url{https://CRAN.R-project.org/package=dplyr}.
-
-\leavevmode\hypertarget{ref-readr}{}%
-Wickham, Hadley, and Jim Hester. 2020. \emph{Readr: Read Rectangular Text Data}. \url{https://CRAN.R-project.org/package=readr}.
-
-\leavevmode\hypertarget{ref-mgcv}{}%
-Wood, S. N. 2003. {``Thin-Plate Regression Splines.''} \emph{Journal of the Royal Statistical Society (B)} 65 (1): 95--114.
-
-\leavevmode\hypertarget{ref-knitr}{}%
-Xie, Yihui.
-2014.
-{``Knitr: A Comprehensive Tool for Reproducible Research in {R}.''} In \emph{Implementing Reproducible Computational Research}, edited by Victoria Stodden, Friedrich Leisch, and Roger D.
-Peng.
-Chapman; Hall/CRC.
-\url{http://www.crcpress.com/product/isbn/9781466561595}.
-
-\leavevmode\hypertarget{ref-rmarkdown}{}%
-Xie, Yihui, Christophe Dervieux, and Emily Riederer. 2020. \emph{R Markdown Cookbook}. Boca Raton, Florida: Chapman; Hall/CRC. \url{https://bookdown.org/yihui/rmarkdown-cookbook}.
-
-\leavevmode\hypertarget{ref-kableExtra}{}%
-Zhu, Hao. 2019. \emph{kableExtra: Construct Complex Table with 'Kable' and Pipe Syntax}. \url{https://CRAN.R-project.org/package=kableExtra}.
-
-\end{CSLReferences}
-
 \bibliographystyle{Chicago}

### 2021-09-16 paper/paper.Rmd
abstract

--- 2021-09-16 paper/paper.Rmd
@@
 abstract: > 
- NSLY79 is a prominent open data source that has been an important support for educational purposes and multidisciplinary research on longitudinal data.
+ The National Longitudinal Survey of Youth (NLSY79) is a prominent open data source that has been important for educational purposes and multidisciplinary research on longitudinal data.
 Subsets of this data can be found in numerous textbooks and research articles.
-However, the steps and decisions taken to get from the original data to these subsets is never clearly articulated.
-This article describes our journey when trying to re-create a textbook example data set from the original database.
-Ultimately, the goal is to provide the resources so that refreshing data subsets can be done on a regular basis, especially when updates are made to the database.
-Thus, this paper demonstrates the process of initial data analysis (IDA) -- tidying, cleaning, and documenting the process -- to make data available for text book examples or research.
-A brief exploratory analysis of the resulting data is also conducted.
+However, the steps and decisions taken to get from the raw data to the textbook data is never clearly articulated.
+This article describes our journey when trying to re-create a textbook example data set from the original database, with the goal being to refresh the textbook data more regularly.
+Thus, this paper demonstrates the process -- extracting, tidying, cleaning, and exploring with documentation -- to make refreshed data available for education or research.
 Three new data sets, and the code to produce them are provided in an accompanying open source R package, called `yowie`.
@@
 #preamble: >
+header-includes:
+  - \usepackage{tcolorbox}
 output: 
@@
 
-```{r, load-pkgs}
+```{r load-pkgs}
 ```
@@
 
-This paper aims to demonstrate the steps of IDA,  tidying, cleaning, and documenting the process, for a prominent open data source, National Longitudinal Survey of Youth (NLSY) 1979, henceforth referred to NSLY79.
+This paper aims to demonstrate the steps of IDA,  tidying, cleaning, and documenting the process, for a prominent open data source, National Longitudinal Survey of Youth (NLSY) 1979, henceforth referred to NLSY79.
 This data has been playing an important role in research in various disciplines, including but are not limited to economics, sociology, education, public policy, and public health for more than a quarter of the century [@MichaelRPergamit2001DWTN].
@@
 
+[FILL] We should include how many variables there are in the data so people get a grasp of the extensive coverage of the variables.
+
 Although the main focus area of the NLSY is labor and employment, the NLSY also covers several other topics, including education; training and achievement; household, geography, and contextual variables; dating, marriage, cohabitation; sexual activity, pregnancy, and fertility; children; income, assets and program participation; health; attitudes and expectations; and crime and substance use.
@@
 This data set represents the measurement of male high-school dropouts, aged from 14 to 17 years old when first time measured.
-Accordingly, to get this data, we have two additional target data, NSLY79 cohort's demographic profiles and wages data of people who completed education up to high school.
+Accordingly, to get this data, we have two additional target data, NLSY79 cohort's demographic profiles and wages data of people who completed education up to high school.
 
@@
 
+\begin{tcolorbox}[width=\textwidth,
+                  boxsep=5pt,
+                  left=5pt,
+                  right=5pt,
+                  top=5pt,
+                  ]
 **Education, Training & Achievement Scores**
@@
 
+\end{tcolorbox}
+
 The downloaded data set came in a csv (NLSY79.csv) and dat (NLSY79.dat) format. We only used the .dat format. It also came along with these files:
@@
 
-```{r, raw-data}
+```{r raw-data}
 ```

### 2021-09-16 paper/paper.tex
abstract

--- 2021-09-14 paper/paper.tex
@@
 \begin{abstract}
-
-\end{abstract}
-
-NSLY79 is a prominent open data source that has been an important support for educational purposes and multidisciplinary research on longitudinal data.
+The National Longitudinal Survey of Youth (NLSY79) is a prominent open data source that has been important for educational purposes and multidisciplinary research on longitudinal data.
 Subsets of this data can be found in numerous textbooks and research articles.
-However, the steps and decisions taken to get from the original data to these subsets is never clearly articulated.
-This article describes our journey when trying to re-create a textbook example data set from the original database.
-Ultimately, the goal is to provide the resources so that refreshing data subsets can be done on a regular basis, especially when updates are made to the database.
-Thus, this paper demonstrates the process of initial data analysis (IDA) -- tidying, cleaning, and documenting the process -- to make data available for text book examples or research.
-A brief exploratory analysis of the resulting data is also conducted.
+However, the steps and decisions taken to get from the raw data to the textbook data is never clearly articulated.
+This article describes our journey when trying to re-create a textbook example data set from the original database, with the goal being to refresh the textbook data more regularly.
+Thus, this paper demonstrates the process -- extracting, tidying, cleaning, and exploring with documentation -- to make refreshed data available for education or research.
 Three new data sets, and the code to produce them are provided in an accompanying open source R package, called \texttt{yowie}.
 As a result of this process, some recommendations are also made for the NLSY79 curators for incorporating data quality checks, and providing more convenient samples of the data to potential users.
+\end{abstract}
 
@@
 
-This paper aims to demonstrate the steps of IDA, tidying, cleaning, and documenting the process, for a prominent open data source, National Longitudinal Survey of Youth (NLSY) 1979, henceforth referred to NSLY79.
+This paper aims to demonstrate the steps of IDA, tidying, cleaning, and documenting the process, for a prominent open data source, National Longitudinal Survey of Youth (NLSY) 1979, henceforth referred to NLSY79.
 This data has been playing an important role in research in various disciplines, including but are not limited to economics, sociology, education, public policy, and public health for more than a quarter of the century (Pergamit et al.
@@
 
-\begin{figure}
-
-{\centering \includegraphics[width=0.7\linewidth]{figures/high-wages-1}
-
-}
-
-\caption{6 out of 45 IDs with extremely high mean hourly wage. Most of the IDs only have one point of high wage.}\label{fig:high-wages}
-\end{figure}
-
 Further, we take 36 samples randomly from the data and plot them, as shown in Figure \ref{fig:sample-plot}.

### 2021-09-16 paper/paper.Rmd
color box for getting data

--- 2021-09-16 paper/paper.Rmd
@@
   There are two versions of the year and month of birth, i.e., 1979 and 1981 data. We downloaded these two versions.
+
+\end{tcolorbox}
 

### 2021-09-16 paper/paper.tex

@@
 
+\usepackage{booktabs}
+\usepackage{longtable}
+\usepackage{array}
+\usepackage{multirow}
+\usepackage{wrapfig}
+\usepackage{float}
+\usepackage{colortbl}
+\usepackage{pdflscape}
+\usepackage{tabu}
+\usepackage{threeparttable}
+\usepackage{threeparttablex}
+\usepackage[normalem]{ulem}
+\usepackage{makecell}
+\usepackage{xcolor}
+
 \begin{document}
@@
 \begin{abstract}
-The National Longitudinal Survey of Youth (NLSY79) is a prominent open data source that has been important for educational purposes and multidisciplinary research on longitudinal data.
+
+\end{abstract}
+
+NSLY79 is a prominent open data source that has been an important support for educational purposes and multidisciplinary research on longitudinal data.
 Subsets of this data can be found in numerous textbooks and research articles.
-However, the steps and decisions taken to get from the raw data to the textbook data is never clearly articulated.
-This article describes our journey when trying to re-create a textbook example data set from the original database, with the goal being to refresh the textbook data more regularly.
-Thus, this paper demonstrates the process -- extracting, tidying, cleaning, and exploring with documentation -- to make refreshed data available for education or research.
+However, the steps and decisions taken to get from the original data to these subsets is never clearly articulated.
+This article describes our journey when trying to re-create a textbook example data set from the original database.
+Ultimately, the goal is to provide the resources so that refreshing data subsets can be done on a regular basis, especially when updates are made to the database.
+Thus, this paper demonstrates the process of initial data analysis (IDA) -- tidying, cleaning, and documenting the process -- to make data available for text book examples or research.
+A brief exploratory analysis of the resulting data is also conducted.
 Three new data sets, and the code to produce them are provided in an accompanying open source R package, called \texttt{yowie}.
 As a result of this process, some recommendations are also made for the NLSY79 curators for incorporating data quality checks, and providing more convenient samples of the data to potential users.
-\end{abstract}
 
@@
 
-This paper aims to demonstrate the steps of IDA, tidying, cleaning, and documenting the process, for a prominent open data source, National Longitudinal Survey of Youth (NLSY) 1979, henceforth referred to NLSY79.
+This paper aims to demonstrate the steps of IDA, tidying, cleaning, and documenting the process, for a prominent open data source, National Longitudinal Survey of Youth (NLSY) 1979, henceforth referred to NSLY79.
 This data has been playing an important role in research in various disciplines, including but are not limited to economics, sociology, education, public policy, and public health for more than a quarter of the century (Pergamit et al.
@@
 This data set represents the measurement of male high-school dropouts, aged from 14 to 17 years old when first time measured.
-Accordingly, to get this data, we have two additional target data, NLSY79 cohort's demographic profiles and wages data of people who completed education up to high school.
+Accordingly, to get this data, we have two additional target data, NSLY79 cohort's demographic profiles and wages data of people who completed education up to high school.
 

### 2021-09-16 paper/paper.Rmd
working on intro

--- 2021-09-16 paper/paper.Rmd
@@
 Documenting the data cleaning is essential for the integrity of downstream statistical analyses and model building [@HuebnerMarianne2020Haar].
-
-Data cleaning is part of what is called "initial data analysis (IDA)" [@Chatfield1985TIEo].
-The other IDA steps are to explore the data properties and check the assumption of modeling to document and report the process for the later formal analysis.
-Hence, exploratory data analysis, which is defined by @tukey as a detective work to get the clue from data, either numerically or graphically, before confirmatory data analysis is performed, encompasses IDA.
+Data cleaning can be considered to be a part of what is called "initial data analysis (IDA)" [@Chatfield1985TIEo].
+In IDA one would also explore the data, especially to check if the dta is consistent with assumptions required for modeling.
+This is also related to exploratory data analysis, which @tukey defines as detective work to learn from data, either numerically or graphically, before confirmatory data analysis is performed, encompasses IDA.
 @DasuTamraparni2003Edma say that data cleaning and exploration, without naming it as IDA, is a difficult task and consumes 80% of the data mining task.
@@
                   ]
-**Education, Training & Achievement Scores**
+**Education, Training \& Achievement Scores**
 
@@
 
-**Household, Geography & Contextual Variables**
+**Household, Geography \& Contextual Variables**
 

### 2021-09-16 paper/paper.tex
working on intro

--- 2021-09-16 paper/paper.tex
@@
 
+\usepackage{tcolorbox}
+
 \begin{document}
@@
 2020).
-
-Data cleaning is part of what is called ``initial data analysis (IDA)'' (Chatfield 1985).
-The other IDA steps are to explore the data properties and check the assumption of modeling to document and report the process for the later formal analysis.
-Hence, exploratory data analysis, which is defined by Tukey (1977) as a detective work to get the clue from data, either numerically or graphically, before confirmatory data analysis is performed, encompasses IDA.
+Data cleaning can be considered to be a part of what is called ``initial data analysis (IDA)'' (Chatfield 1985).
+In IDA one would also explore the data, especially to check if the dta is consistent with assumptions required for modeling.
+This is also related to exploratory data analysis, which Tukey (1977) defines as detective work to learn from data, either numerically or graphically, before confirmatory data analysis is performed, encompasses IDA.
 Dasu and Johnson (2003) say that data cleaning and exploration, without naming it as IDA, is a difficult task and consumes 80\% of the data mining task.
@@
 
+{[}FILL{]} We should include how many variables there are in the data so people get a grasp of the extensive coverage of the variables.
+
 Although the main focus area of the NLSY is labor and employment, the NLSY also covers several other topics, including education; training and achievement; household, geography, and contextual variables; dating, marriage, cohabitation; sexual activity, pregnancy, and fertility; children; income, assets and program participation; health; attitudes and expectations; and crime and substance use.
@@
 
-\textbf{Education, Training \& Achievement Scores}
+\begin{tcolorbox}[width=\textwidth,
+                  boxsep=5pt,
+                  left=5pt,
+                  right=5pt,
+                  top=5pt,
+                  ]
+**Education, Training \& Achievement Scores**
 
-\begin{itemize}
-\tightlist
-\item
-  Education -\textgreater{} Summary measures -\textgreater{} All schools -\textgreater{} By year -\textgreater{} Highest grade completed
+- Education -> Summary measures -> All schools -> By year -> Highest grade completed
+  - Downloaded all of variables in Highest grade completed.
 
-  \begin{itemize}
-  \tightlist
-  \item
-    Downloaded all of variables in Highest grade completed.
-  \end{itemize}
-\end{itemize}
+**Employment**
 
-\textbf{Employment}
+- Summary measures -> By job -> Hours worked and Hourly wages
+  - Downloaded all of the variables in Hours worked.
+  - Downloaded all of the variables in Hourly wages
+  Both hours worked and hourly wages are recorded by the job, up to five jobs for each id/subject.
 
-\begin{itemize}
-\tightlist
-\item
-  Summary measures -\textgreater{} By job -\textgreater{} Hours worked and Hourly wages
+**Household, Geography \& Contextual Variables**
 
-  \begin{itemize}
-  \tightlist
-  \item
-    Downloaded all of the variables in Hours worked.
-  \item
-    Downloaded all of the variables in Hourly wages
-    Both hours worked and hourly wages are recorded by the job, up to five jobs for each id/subject.
-  \end{itemize}
-\end{itemize}
+- Context -> Summary measures -> Basic demographics 
+  - Downloaded year and month of birth, race, and sex variables.
+  There are two versions of the year and month of birth, i.e., 1979 and 1981 data. We downloaded these two versions.
 
-\textbf{Household, Geography \& Contextual Variables}
-
-\begin{itemize}
-\tightlist
-\item
-  Context -\textgreater{} Summary measures -\textgreater{} Basic demographics
-
-  \begin{itemize}
-  \tightlist
-  \item
-    Downloaded year and month of birth, race, and sex variables.
-    There are two versions of the year and month of birth, i.e., 1979 and 1981 data. We downloaded these two versions.
-  \end{itemize}
-\end{itemize}
+\end{tcolorbox}
 

### 2021-09-16 paper/paper-jsdse/paper-jsdse.tex
figure update

--- 2021-09-16 paper/paper-jsdse/paper-jsdse.tex
@@
 \usepackage{natbib}
+\usepackage{tcolorbox}
 \usepackage{url} % not crucial - just used below for the URL
@@
 
-NSLY79 is a prominent open data source that has been an important support for educational purposes and multidisciplinary research on longitudinal data.
-Subsets of this data can be found in numerous textbooks and research articles.
-However, the steps and decisions taken to get from the original data to these subsets is never clearly articulated.
-This article describes our journey when trying to re-create a textbook example data set from the original database.
-Ultimately, the goal is to provide the resources so that refreshing data subsets can be done on a regular basis, especially when updates are made to the database.
-Thus, this paper demonstrates the process of initial data analysis (IDA) -- tidying, cleaning, and documenting the process -- to make data available for text book examples or research.
-A brief exploratory analysis of the resulting data is also conducted.
-Three new data sets, and the code to produce them are provided in an accompanying open source R package, called \texttt{yowie}.
-As a result of this process, some recommendations are also made for the NLSY79 curators for incorporating data quality checks, and providing more convenient samples of the data to potential users.
-
 \textbf{Keywords}: Data cleaning; Data tidying; Longitudinal data; NLSY79; Open data; Initial data analysis; Outlier detection; Robust linear regression
@@
 
-\textbf{Education, Training \& Achievement Scores}
+\begin{tcolorbox}[width=\textwidth,
+                  boxsep=5pt,
+                  left=5pt,
+                  right=5pt,
+                  top=5pt,
+                  ]
+Education, Training and Achievement Scores
 
-\begin{itemize}
-\tightlist
-\item
-  Education -\textgreater{} Summary measures -\textgreater{} All schools -\textgreater{} By year -\textgreater{} Highest grade completed
+- Education -> Summary measures -> All schools -> By year -> Highest grade completed
+  - Downloaded all of variables in Highest grade completed.
 
-  \begin{itemize}
-  \tightlist
-  \item
-    Downloaded all of variables in Highest grade completed.
-  \end{itemize}
-\end{itemize}
+Employment
 
-\textbf{Employment}
+- Summary measures -> By job -> Hours worked and Hourly wages
+  - Downloaded all of the variables in Hours worked.
+  - Downloaded all of the variables in Hourly wages
+  Both hours worked and hourly wages are recorded by the job, up to five jobs for each id/subject.
 
-\begin{itemize}
-\tightlist
-\item
-  Summary measures -\textgreater{} By job -\textgreater{} Hours worked and Hourly wages
+Household, Geography and Contextual Variables
 
-  \begin{itemize}
-  \tightlist
-  \item
-    Downloaded all of the variables in Hours worked.
-  \item
-    Downloaded all of the variables in Hourly wages
-    Both hours worked and hourly wages are recorded by the job, up to five jobs for each id/subject.
-  \end{itemize}
-\end{itemize}
+- Context -> Summary measures -> Basic demographics 
+  - Downloaded year and month of birth, race, and sex variables.
+  There are two versions of the year and month of birth, i.e., 1979 and 1981 data. We downloaded these two versions.
 
-\textbf{Household, Geography \& Contextual Variables}
-
-\begin{itemize}
-\tightlist
-\item
-  Context -\textgreater{} Summary measures -\textgreater{} Basic demographics
-
-  \begin{itemize}
-  \tightlist
-  \item
-    Downloaded year and month of birth, race, and sex variables.
-    There are two versions of the year and month of birth, i.e., 1979 and 1981 data. We downloaded these two versions.
-  \end{itemize}
-\end{itemize}
+\end{tcolorbox}
 

### 2021-09-16 paper/paper.Rmd
figure update

--- 2021-09-16 paper/paper.Rmd
@@
                cache.path = "paper_cache/",
+               fig.retina = 2,
                fig.path = "figures/",
                comment = "#>",
-               fig.align = 'center', 
-               out.width = "70%")
+               fig.align = 'center')
 
@@
                   ]
-**Education, Training & Achievement Scores**
+Education, Training and Achievement Scores
 
@@
 
-**Employment**
+Employment
 
@@
 
-**Household, Geography & Contextual Variables**
+Household, Geography and Contextual Variables
 

### 2021-09-16 paper/paper.Rmd

--- 2021-09-16 paper/paper.Rmd
@@
                   ]
-**Education, Training \& Achievement Scores**
+Education, Training and Achievement Scores
 
@@
 
-**Household, Geography \& Contextual Variables**
+Household, Geography and Contextual Variables
 

### 2021-09-16 paper/paper.Rmd

--- 2021-09-16 paper/paper.Rmd
@@
                cache.path = "paper_cache/",
-               fig.retina = 2,
                fig.path = "figures/",
                comment = "#>",
-               fig.align = 'center')
+               fig.align = 'center', 
+               out.width = "70%")
 
@@
                   ]
-Education, Training and Achievement Scores
+**Education, Training & Achievement Scores**
 
@@
 
-Employment
+**Employment**
 
@@
 
-Household, Geography and Contextual Variables
+**Household, Geography & Contextual Variables**
 

### 2021-09-16 paper/paper-jsdse/paper-jsdse.tex
messing with figures in the cleaning section

@@
 
-{\centering \includegraphics[width=0.7\linewidth]{figures/sample-plot-1}
+{\centering \includegraphics[width=1\linewidth]{figures/sample-plot-1}
 
@@
 
-\caption{Getting a glimpse of the data: random sample of 36 individuals, divided into  twelve subplots. Wages are shown against year. There is a lot of variability in wages.}\label{fig:sample-plot}
+\caption{A glimpse of the refreshed data.
+Longitudinal profiles of wages from 1979-2018 are shown for a random sample of 36 individuals, divided into twelve subplots.
+Most individuals here experienced increasing wages over time, although there is considerable fluctuation from one year to another.}\label{fig:sample-plot}
 \end{figure}
@@
 
-{\centering \includegraphics[width=0.7\linewidth]{figures/feature-plot-1}
+{\centering \includegraphics[width=432px]{figures/feature-plot-1}
 
@@
 
-\caption{Several plots showing anomalies observed the input data.
-XX Add a couple more plots that reveal the initial problems in the data.
-Plot A portrays the pattern of mean hourly wage of high school cohort from 1979 to 2018 of each ID in US Dollar; Plot B shows the distribution of their minimum, median, and maximum value.
-We can see that some IDs had an extremely high of wages and it made the distribution of the three features is extremely skewed.}\label{fig:feature-plot}
+\caption{Summary plots to check the cleaned data reveal more cleaning is necessary: longitudinal profiles of wages for all individuals 1979-2018 (A), boxplots of minimum, median, and maximum wages of each individual (B), and one individual with an unusual wage relative to their years of data (C).
+Some values of hourly wages are unbelievable, and some individuals have extemely unusual wages in some years.}\label{fig:feature-plot}
 \end{figure}
@@
 
-{\centering \includegraphics[width=0.7\linewidth]{figures/compare-plot-1}
+{\centering \includegraphics[width=432px]{figures/compare-plot-1}
 
@@
 
-{\centering \includegraphics[width=0.7\linewidth]{figures/comppict-1}
+{\centering \includegraphics[width=432px]{figures/fixed-feature-plot-1}
 
@@
 
-\caption{The distribution of the mean hourly wage after the extreme values are replaced.
-Plot A portrays the pattern of mean hourly wage of high school cohort from 1979 to 2018 of each ID in US Dollar; Plot B shows the distribution of their minimum, median, and maximum value transformed to log10 scale.
-We can see that some observations still had reasonbaly higher wages than the others.
-Also, some IDs' have a minimum wages that is higher than others' maximum wages.}\label{fig:comppict}
+\caption{Re-make of the summary plots of the fully processed data suggest that the data is now in a reasonable state: longitudinal profiles of wages for all individuals 1979-2018 (A), boxplots of minimum, median, and maximum wages of each individual (B), and one individual with an unusual wage relative to their years of data (C).
+}\label{fig:fixed-feature-plot}
 \end{figure}
@@
 
-{\centering \includegraphics[width=0.7\linewidth]{figures/flow-chart-1}
+{\centering \includegraphics[width=468px]{figures/flow-chart-1}
 
@@
 
-{\centering \includegraphics[width=0.7\linewidth]{figures/slope-nobs-plot-1}
+{\centering \includegraphics[width=468px]{figures/slope-nobs-plot-1}
 
@@
 
-{\centering \includegraphics[width=0.7\linewidth]{figures/slope-gender-1}
+{\centering \includegraphics[width=468px]{figures/slope-gender-1}
 
@@
 
-{\centering \includegraphics[width=0.7\linewidth]{figures/slope-educ-1}
+{\centering \includegraphics[width=468px]{figures/slope-educ-1}
 

### 2021-09-16 paper/paper.Rmd
messing with figures in the cleaning section

--- 2021-09-16 paper/paper.Rmd
@@
 
-```{r sample-plot, echo = FALSE, fig.cap="Getting a glimpse of the data: random sample of 36 individuals, divided into  twelve subplots.
-Wages are shown against year.
-There is a lot of variability in wages.", fig.height=4, fig.width=5}
+```{r sample-plot, echo = FALSE, fig.cap="A glimpse of the refreshed data.
+Longitudinal profiles of wages from 1979-2018 are shown for a random sample of 36 individuals, divided into twelve subplots.
+Most individuals here experienced increasing wages over time, although there is considerable fluctuation from one year to another.", fig.height=4, fig.width=5, out.width="100%"}
 ```
@@
 
-```{r feature-plot, fig.cap = "Several plots showing anomalies observed the input data.
-XX Add a couple more plots that reveal the initial problems in the data.
-Plot A portrays the pattern of mean hourly wage of high school cohort from 1979 to 2018 of each ID in US Dollar; Plot B shows the distribution of their minimum, median, and maximum value.
-We can see that some IDs had an extremely high of wages and it made the distribution of the three features is extremely skewed.", fig.width=6, fig.height=4}
+```{r feature-plot, fig.cap = "Summary plots to check the cleaned data reveal more cleaning is necessary: longitudinal profiles of wages for all individuals 1979-2018 (A), boxplots of minimum, median, and maximum wages of each individual (B), and one individual with an unusual wage relative to their years of data (C).
+Some values of hourly wages are unbelievable, and some individuals have extemely unusual wages in some years.", fig.width=6, fig.height=2}
 ```
@@
 
-```{r high-wages, fig.cap= "6 out of 45 IDs with extremely high mean hourly wage. Most of the IDs only have one point of high wage."}
+```{r high-wages, fig.cap= "6 out of 45 IDs with extremely high mean hourly wage. Most of the IDs only have one point of high wage.", eval=FALSE}
 ```
@@
 The orange line portray the original value of mean hourly wage, while the turquoise line display the mean hourly wages value after the extreme values imputed with the robust linear model's prediction value.
-We can see that some extreme spikes has been reduced by the model.", fig.height=5, fig.width=6}
+We can see that some extreme spikes has been reduced by the model.", fig.height=5, fig.width=6, outwidth="100%"}
 ```
@@
 We can see that some observations still had reasonbaly higher wages than the others.
-Also, some IDs' have a minimum wages that is higher than others' maximum wages.", fig.width=6, fig.height=4}
+Also, some IDs' have a minimum wages that is higher than others' maximum wages.", fig.width=6, fig.height=4, eval=FALSE}
+```
+
+```{r fixed-feature-plot, fig.cap = "Re-make of the summary plots of the fully processed data suggest that the data is now in a reasonable state: longitudinal profiles of wages for all individuals 1979-2018 (A), boxplots of minimum, median, and maximum wages of each individual (B), and one individual with an unusual wage relative to their years of data (C).
+", fig.width=6, fig.height=2}
 ```

### 2021-09-16 paper/paper.tex
messing with figures in the cleaning section

@@
 \usepackage{tcolorbox}
+\usepackage{booktabs}
+\usepackage{longtable}
+\usepackage{array}
+\usepackage{multirow}
+\usepackage{wrapfig}
+\usepackage{float}
+\usepackage{colortbl}
+\usepackage{pdflscape}
+\usepackage{tabu}
+\usepackage{threeparttable}
+\usepackage{threeparttablex}
+\usepackage[normalem]{ulem}
+\usepackage{makecell}
+\usepackage{xcolor}
 

### 2021-09-16 paper/scripts/ida.R

@@ -43,8 +43,12 @@ ggplot(wages_demog_hs_tsibble,
            group = id)) +
   geom_line(alpha = 0.7) +
   facet_sample() +
+  scale_x_continuous("",
+                     breaks = seq(1980, 2020, 10),
+                     labels = c("80", "90", "00", "10", "20"),
+                     minor_breaks = seq(1980, 2020, 5)) +
   theme(axis.text.x = element_text(angle = 10, size = 6)) +
-  ylab("mean hourly wage") +
+  ylab("Hourly wage ($)") +
   theme_bw()
 
 ## ---- feature-plot
@@ -60,27 +64,38 @@ spag <- wages_demog_hs %>%
              y = mean_hourly_wage,
              group = id)) +
   geom_line(alpha = 0.1) +
-  ggtitle("A)") +
+  scale_x_continuous("",
+                     breaks = seq(1980, 2020, 10),
+                     labels = c("80", "90", "00", "10", "20"),
+                     minor_breaks = seq(1980, 2020, 5)) +
+  ggtitle("A") +
   theme(plot.title = element_text(size = 10)) +
   theme_bw() +
-  ylab("mean hourly wage")
+  ylab("Hourly wage ($)")
 
 wages_three_feat <- wages_demog_hs_tsibble %>%
   features(mean_hourly_wage,
            feat_three_num
   )
 wages_feat_long <- wages_three_feat %>%
-  pivot_longer(c(min, med, max), names_to = "feature", values_to = "value")
+  pivot_longer(c(min, med, max),
+               names_to = "feature", values_to = "value") %>%
+  mutate(feature = factor(feature, levels = c("min", "med", "max")))
+
 feature <- ggplot(wages_feat_long) +
   geom_density(aes(x = value, colour = feature, fill = feature), alpha = 0.3) +
-  ggtitle("B)") +
+  ggtitle("B") +
   theme_bw() +
   theme(plot.title = element_text(size = 10))
 
-feature_bp <- ggplot(wages_feat_long, aes(y=value, fill = feature, color = feature)) +
+feature_bp <- ggplot(wages_feat_long,
+                     aes(y=value, x = feature,
+                         fill = feature, color = feature)) +
   geom_boxplot() +
   theme_bw() +
-  ggtitle("C)") +
+  #ggtitle("C") +
+  ggtitle("B")  +
+  ylab("Hourly wage ($)") +
   theme(legend.position = "none",
[…]

### 2021-09-16 paper/paper.Rmd
added id to individual subject plot

@@
 
-```{r feature-plot, fig.cap = "Summary plots to check the cleaned data reveal more cleaning is necessary: longitudinal profiles of wages for all individuals 1979-2018 (A), boxplots of minimum, median, and maximum wages of each individual (B), and one individual with an unusual wage relative to their years of data (C).
+```{r feature-plot, fig.cap = "Summary plots to check the cleaned data reveal more cleaning is necessary: longitudinal profiles of wages for all individuals 1979-2018 (A), boxplots of minimum, median, and maximum wages of each individual (B), and one individual (id=39) with an unusual wage relative to their years of data (C).
 Some values of hourly wages are unbelievable, and some individuals have extemely unusual wages in some years.", fig.width=6, fig.height=2}

### 2021-09-16 paper/scripts/ida.R

@@ -106,6 +106,7 @@ plot_high <- ggplot(filter(wages_high2, id == 39)) +
                  y = mean_hourly_wage),
              size = 0.5,
              alpha = 0.5) +
+  annotate("text", x = 2010, y = 1000, label = "id: 39") +
   scale_x_continuous("",
                      breaks = seq(1980, 2020, 10),
                      labels = c("80", "90", "00", "10", "20"),
@@ -323,6 +324,7 @@ plot_high <- ggplot(filter(wages_high2, id == 39)) +
                  y = mean_hourly_wage),
              size = 0.5,
              alpha = 0.5) +
+  annotate("text", x = 2010, y = 1000, label = "id: 39") +
   scale_x_continuous("",
                      breaks = seq(1980, 2020, 10),
                      labels = c("80", "90", "00", "10", "20"),

### 2021-09-16 paper/paper-jsdse/paper-jsdse.tex

@@
 
-\caption{Summary plots to check the cleaned data reveal more cleaning is necessary: longitudinal profiles of wages for all individuals 1979-2018 (A), boxplots of minimum, median, and maximum wages of each individual (B), and one individual with an unusual wage relative to their years of data (C).
+\caption{Summary plots to check the cleaned data reveal more cleaning is necessary: longitudinal profiles of wages for all individuals 1979-2018 (A), boxplots of minimum, median, and maximum wages of each individual (B), and one individual (id=39) with an unusual wage relative to their years of data (C).
 Some values of hourly wages are unbelievable, and some individuals have extemely unusual wages in some years.}\label{fig:feature-plot}

### 2021-09-16 paper/paper.Rmd

@@
 
-The next step is that explore the mean hourly wage data.
-In this case, we only explore the wages data of up to $12^{th}$ grade.
-Table \ref{tab:summarytable} shows that the median of the overall wage of the cohort is only 7.2, while the mean is 11.87.
-It indicates that the data might contains a lot of extreme values.
+The next step is that explore the mean hourly wage data. In this case, we only explore the wages data of up to $12^{th}$ grade.  Here, we use visualisation techniques to do IDA.
 
@@
 
-However, we can not be sure only by observing the summary statistics.
-Hence, we use visualisation techniques to investigate this matter.
-Figure \ref{fig:feature-plot} conveys a problem in the mean hourly wage values.
-Figure \ref{fig:feature-plot} A shows that some observations have an exceptionally high figure of wages, even more than US$10,000 per hour.
-In Figure \ref{fig:feature-plot} B, we barely see any difference in the minimum, median, and maximum value of the wages since the distribution is heavily skewed to the right.
+We take 36 samples randomly from the data and plot them, as shown in Figure \ref{fig:sample-plot}.
+It implies that these respondents have a lot of variability in wages, for example, the IDs in panel numbers 5, 7, and 11.
+The plot also implies that the samples have a different pattern of mean hourly wages.
+Some have had flat wages for years but had a sudden increase in one particular year, then it gone down again, while the others experienced an upsurge in their wage, for instance, the IDs in panel 9.
 
@@
 
-In Figure \ref{fig:high-wages}, we plot some respondents with a high value of mean hourly wages.
-We filter all of the IDs who earned more than US$ 500 per hour on average.
-We find that these respondents only experienced one point of extremely high wages.
-Thus, we suspect that these high values are erroneous resulted from a data entry error.
+```{r high-wages}
+```
 
@@
 ```
-
-Further, we take 36 samples randomly from the data and plot them, as shown in Figure \ref{fig:sample-plot}.
-It implies that not only that some observations earn extremely high figures of wages, but some also have reasonably fluctuated wages, for example, the IDs in panel numbers 5, 7, and 11.
-The plot also implies that the samples have a different pattern of mean hourly wages.
-Some have had flat wages for years but had a sudden increase in one particular year, then it gone down again, while the others experienced an upsurge in their wage, for instance, the IDs in panel 9.
 

### 2021-09-16 paper/scripts/data-preprocessing.R
change objects name, remove wages_demog_hs.

@@ -229,10 +229,10 @@ wages_demog <- wages_demog %>%
   mutate(years_in_workforce = year - yr_hgc) %>%
   mutate(age_1979 = 1979 - (dob_year + 1900))
 # filter only the id with high school education
-wages_demog_hs <- wages_demog  %>% filter(grepl("GRADE", hgc))
+wages_before <- wages_demog  %>% filter(grepl("GRADE", hgc))
 # calculate the number of observation
-keep_me <- wages_demog_hs %>%
+keep_me <- wages_before %>%
   count(id) %>%
   filter(n > 4)
-wages_demog_hs <- wages_demog_hs %>%
+wages_before <- wages_before %>%
   filter(id %in% keep_me$id)

### 2021-09-16 paper/scripts/ida.R

@@ -25,19 +25,19 @@ kable(gender_race_table,
   add_header_above(c(" " = 1, "Race" = 3, " " = 1))
 
 ## ---- summarytable
-kable(as.array(summary(wages_demog_hs$mean_hourly_wage)),
+kable(as.array(summary(wages_before$mean_hourly_wage)),
       caption = "Summary Statistics of Wages of High School Data",
       col.names = c("Statistics", "Value")) %>%
   kable_styling()
 
 ## ---- sample-plot
-wages_demog_hs_tsibble <- as_tsibble(x = wages_demog_hs,
+wages_before_tsibble <- as_tsibble(x = wages_before,
                                      key = id,
                                      index = year,
                                      regular = FALSE)
 
 set.seed(20210225)
-ggplot(wages_demog_hs_tsibble,
+ggplot(wages_before_tsibble,
        aes(x = year,
            y = mean_hourly_wage,
            group = id)) +
@@ -52,14 +52,14 @@ ggplot(wages_demog_hs_tsibble,
   theme_bw()
 
 ## ---- feature-plot
-wages_high <- filter(wages_demog_hs, mean_hourly_wage > 500) %>%
+wages_high <- filter(wages_before, mean_hourly_wage > 500) %>%
   as_tibble() %>%
   head(n = 6)
 
-wages_high2 <- wages_demog_hs %>%
+wages_high2 <- wages_before %>%
   filter(id %in% wages_high$id)
 
-spag <- wages_demog_hs %>%
+spag <- wages_before %>%
   ggplot(aes(x = year,
              y = mean_hourly_wage,
              group = id)) +
@@ -73,7 +73,7 @@ spag <- wages_demog_hs %>%
   theme_bw() +
   ylab("Hourly wage ($)")
 
-wages_three_feat <- wages_demog_hs_tsibble %>%
+wages_three_feat <- wages_before_tsibble %>%
   features(mean_hourly_wage,
            feat_three_num
   )
@@ -132,7 +132,7 @@ spag + feature_bp + plot_high +
 
 ## ---- rlm
 # nest the data by id to build a robust linear model
-by_id <- wages_demog_hs %>%
+by_id <- wages_before %>%
   dplyr::select(id, year, mean_hourly_wage) %>%
   group_by(id) %>%
   nest()
@@ -183,17 +183,17 @@ wages_rlm_dat <- id_aug_w %>%
 
 # join back the `wages_rlm_dat` to `wages_demog_hs`
 
[…]

### 2021-09-16 paper/scripts/ida.R
Change histogram to use boxplot in the cleaned data.

@@ -245,7 +245,19 @@ feature2 <- ggplot(wages_feat_long_rlm) +
   xlab("value (log10)") +
   theme_bw()
 
-spag2 + feature2
+feature2_bp <- ggplot(wages_feat_long_rlm,
+                     aes(y=value, x = feature,
+                         fill = feature, color = feature)) +
+  geom_boxplot() +
+  theme_bw() +
+  #ggtitle("C") +
+  ggtitle("B")  +
+  ylab("Hourly wage ($)") +
+  theme(legend.position = "none",
+        plot.title = element_text(size = 10))
+
+
+spag2 + feature2_bp
 
 
 ## ---- fixed-feature-plot

### 2021-09-16 paper/scripts/ida.R
Add individual plot in clean data as comparison.

@@ -256,8 +256,26 @@ feature2_bp <- ggplot(wages_feat_long_rlm,
   theme(legend.position = "none",
         plot.title = element_text(size = 10))
 
+plot_high_after <- ggplot(filter(wages_cleaned, id == 39)) +
+  geom_line(aes(x = year,
+                y = wages_rlm)) +
+  geom_point(aes(x = year,
+                 y = wages_rlm),
+             size = 0.5,
+             alpha = 0.5) +
+  annotate("text", x = 2010, y = 1000, label = "id: 39") +
+  scale_x_continuous("",
+                     breaks = seq(1980, 2020, 10),
+                     labels = c("80", "90", "00", "10", "20"),
+                     minor_breaks = seq(1980, 2020, 5)) +
+  theme(axis.text.x = element_text(angle = 10, size = 6),
+        plot.title = element_text(size = 10)) +
+  ylab("Hourly wage ($)") +
+  theme_bw() +
+  ggtitle("C")
 
-spag2 + feature2_bp
+
+spag2 + feature2_bp + plot_high_after
 
 
 ## ---- fixed-feature-plot

### 2021-09-19 paper/paper-jsdse/paper-jsdse.Rmd
fixed keywords to YAML, modified JSDSE template to accommodate keywords, fixed abstract, fixed tcolorbox to div fence

--- 2021-09-16 paper/paper-jsdse/paper-jsdse.Rmd
@@
 ---
-title: "The Journey from Wild to Textbook Data: A Case Study from the National Longitudinal Survey of Youth"
 authors:
   - name: Dewi Amaliah
-    thanks: ""
+    thanks: "Test"
     department: Department of Econometrics and Business Statistics
@@
 
+```{r, include = FALSE}
+paper <- here::here("paper/paper.Rmd")
+meta <- rmarkdown::yaml_front_matter(paper)
+```
+
+---
+title: "`r meta$title`"
+abstract: >
+  `r meta$abstract`
+keywords: `r meta$keywords`
+---
+
 ```{r, child = "../paper.Rmd"}

### 2021-09-19 paper/paper-jsdse/paper-jsdse.tex
fixed keywords to YAML, modified JSDSE template to accommodate keywords, fixed abstract, fixed tcolorbox to div fence

@@
 \begin{abstract}
-
+The National Longitudinal Survey of Youth (NLSY79) is a prominent open data source that has been important for educational purposes and multidisciplinary research on longitudinal data.
+Subsets of this data can be found in numerous textbooks and research articles.
+However, the steps and decisions taken to get from the raw data to the textbook data is never clearly articulated.
+This article describes our journey when trying to re-create a textbook example data set from the original database, with the goal being to refresh the textbook data more regularly.
+Thus, this paper demonstrates the process -- extracting, tidying, cleaning, and exploring with documentation -- to make refreshed data available for education or research.
+Three new data sets, and the code to produce them are provided in an accompanying open source R package, called \texttt{yowie}.
+As a result of this process, some recommendations are also made for the NLSY79 curators for incorporating data quality checks, and providing more convenient samples of the data to potential users.
 \end{abstract}
@@
 \noindent%
-{\it Keywords:}  3 to 6 keywords, that do not appear in the title
+{\it Keywords:}  Data cleaning; Data tidying; Longitudinal data; NLSY79; Open data; Initial data analysis; Outlier detection; Robust linear regression
 \vfill
@@
 \spacingset{1.45} % DON'T change the spacing!
-
-\textbf{Keywords}: Data cleaning; Data tidying; Longitudinal data; NLSY79; Open data; Initial data analysis; Outlier detection; Robust linear regression
 
@@
 
+In this section, we outline the steps taken to the download the raw data from the database provided at {[}CITE{]}. Specifically, we only extract a subset of the variables in the data base.
+
 \hypertarget{getting-and-tidying-the-data}{%
@@
 
-\begin{tcolorbox}[width=\textwidth,
-                  boxsep=5pt,
-                  left=5pt,
-                  right=5pt,
-                  top=5pt,
-                  ]
+\begin{tcolorbox}[label="box:data"]
+
 Education, Training and Achievement Scores
 
-- Education -> Summary measures -> All schools -> By year -> Highest grade completed
-  - Downloaded all of variables in Highest grade completed.
+\begin{itemize}
+\tightlist
+\item
+  Education -\textgreater{} Summary measures -\textgreater{} All schools -\textgreater{} By year -\textgreater{} Highest grade completed
+
+  \begin{itemize}
+  \tightlist
+  \item
+    Downloaded all of variables in Highest grade completed.
+  \end{itemize}
+\end{itemize}
 
@@
 
-- Summary measures -> By job -> Hours worked and Hourly wages
-  - Downloaded all of the variables in Hours worked.
-  - Downloaded all of the variables in Hourly wages
+\begin{itemize}
+\tightlist
+\item
+  Summary measures -\textgreater{} By job -\textgreater{} Hours worked and Hourly wages
+
+  \begin{itemize}
+  \tightlist
+  \item
+    Downloaded all of the variables in Hours worked.
+  \item
+    Downloaded all of the variables in Hourly wages
     Both hours worked and hourly wages are recorded by the job, up to five jobs for each id/subject.
+  \end{itemize}
+\end{itemize}
 
@@
 
-- Context -> Summary measures -> Basic demographics 
-  - Downloaded year and month of birth, race, and sex variables.
+\begin{itemize}
+\tightlist
+\item
+  Context -\textgreater{} Summary measures -\textgreater{} Basic demographics
+
+  \begin{itemize}
+  \tightlist
+  \item
+    Downloaded year and month of birth, race, and sex variables.
     There are two versions of the year and month of birth, i.e., 1979 and 1981 data. We downloaded these two versions.
+  \end{itemize}
+\end{itemize}
 
@@
 
-The next step is that explore the mean hourly wage data.
-In this case, we only explore the wages data of up to \(12^{th}\) grade.
-Table \ref{tab:summarytable} shows that the median of the overall wage of the cohort is only 7.2, while the mean is 11.87.
-It indicates that the data might contains a lot of extreme values.
+The next step is that explore the mean hourly wage data. In this case, we only explore the wages data of up to \(12^{th}\) grade. Here, we use visualisation techniques to do IDA.
 
@@
 
-However, we can not be sure only by observing the summary statistics.
-Hence, we use visualisation techniques to investigate this matter.
-Figure \ref{fig:feature-plot} conveys a problem in the mean hourly wage values.
-Figure \ref{fig:feature-plot} A shows that some observations have an exceptionally high figure of wages, even more than US\$10,000 per hour.
-In Figure \ref{fig:feature-plot} B, we barely see any difference in the minimum, median, and maximum value of the wages since the distribution is heavily skewed to the right.
+We take 36 samples randomly from the data and plot them, as shown in Figure \ref{fig:sample-plot}.
+It implies that these respondents have a lot of variability in wages, for example, the IDs in panel numbers 5, 7, and 11.
+The plot also implies that the samples have a different pattern of mean hourly wages.
+Some have had flat wages for years but had a sudden increase in one particular year, then it gone down again, while the others experienced an upsurge in their wage, for instance, the IDs in panel 9.
 
@@
 \end{figure}
-
-In Figure \ref{fig:high-wages}, we plot some respondents with a high value of mean hourly wages.
-We filter all of the IDs who earned more than US\$ 500 per hour on average.
-We find that these respondents only experienced one point of extremely high wages.
-Thus, we suspect that these high values are erroneous resulted from a data entry error.
-
-Further, we take 36 samples randomly from the data and plot them, as shown in Figure \ref{fig:sample-plot}.
-It implies that not only that some observations earn extremely high figures of wages, but some also have reasonably fluctuated wages, for example, the IDs in panel numbers 5, 7, and 11.
-The plot also implies that the samples have a different pattern of mean hourly wages.
-Some have had flat wages for years but had a sudden increase in one particular year, then it gone down again, while the others experienced an upsurge in their wage, for instance, the IDs in panel 9.
 

### 2021-09-19 paper/paper.Rmd
fixed keywords to YAML, modified JSDSE template to accommodate keywords, fixed abstract, fixed tcolorbox to div fence

@@
   - \usepackage{tcolorbox}
+  - \newtcolorbox{tbox}{width=\textwidth,boxsep=5pt}
 output: 
@@
 always_allow_html: true
+keywords: "Data cleaning; Data tidying; Longitudinal data; NLSY79; Open data; Initial data analysis; Outlier detection; Robust linear regression"
 ---
@@
 ```
-
-**Keywords**: Data cleaning; Data tidying; Longitudinal data; NLSY79; Open data; Initial data analysis; Outlier detection; Robust linear regression
 
@@
 
+In this section, we outline the steps taken to the download the raw data from the database provided at [CITE]. Specifically, we only extract a subset of the variables in the data base.
+
 ## Getting and Tidying the Data
@@
 
-\begin{tcolorbox}[width=\textwidth,
-                  boxsep=5pt,
-                  left=5pt,
-                  right=5pt,
-                  top=5pt,
-                  ]
+::: {.tcolorbox data-latex="[label=\"box:data\"]" }
+
 Education, Training and Achievement Scores
@@
 
-\end{tcolorbox}
+:::
 

### 2021-09-19 paper/paper.tex
fixed keywords to YAML, modified JSDSE template to accommodate keywords, fixed abstract, fixed tcolorbox to div fence

@@
 % For Pandoc 2.11+
-\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
+\newenvironment{CSLReferences}[3] % #1 hanging-ident, #2 entry spacing
  {% don't indent paragraphs
@@
 \newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
-\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
+\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}}
 \newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
@@
 \usepackage{tcolorbox}
+\newtcolorbox{tbox}{width=\textwidth,boxsep=5pt}
 \usepackage{booktabs}
@@
 
-\begin{tcolorbox}[width=\textwidth,
-                  boxsep=5pt,
-                  left=5pt,
-                  right=5pt,
-                  top=5pt,
-                  ]
 Education, Training and Achievement Scores
 
-- Education -> Summary measures -> All schools -> By year -> Highest grade completed
-  - Downloaded all of variables in Highest grade completed.
+\begin{itemize}
+\tightlist
+\item
+  Education -\textgreater{} Summary measures -\textgreater{} All schools -\textgreater{} By year -\textgreater{} Highest grade completed
+
+  \begin{itemize}
+  \tightlist
+  \item
+    Downloaded all of variables in Highest grade completed.
+  \end{itemize}
+\end{itemize}
 
@@
 
-- Context -> Summary measures -> Basic demographics 
-  - Downloaded year and month of birth, race, and sex variables.
+\begin{itemize}
+\tightlist
+\item
+  Context -\textgreater{} Summary measures -\textgreater{} Basic demographics
+
+  \begin{itemize}
+  \tightlist
+  \item
+    Downloaded year and month of birth, race, and sex variables.
     There are two versions of the year and month of birth, i.e., 1979 and 1981 data. We downloaded these two versions.
-
-\end{tcolorbox}
+  \end{itemize}
+\end{itemize}
 
@@
 
-\caption{Summary plots to check the cleaned data reveal more cleaning is necessary: longitudinal profiles of wages for all individuals 1979-2018 (A), boxplots of minimum, median, and maximum wages of each individual (B), and one individual with an unusual wage relative to their years of data (C).
+\caption{Summary plots to check the cleaned data reveal more cleaning is necessary: longitudinal profiles of wages for all individuals 1979-2018 (A), boxplots of minimum, median, and maximum wages of each individual (B), and one individual (id=39) with an unusual wage relative to their years of data (C).
 Some values of hourly wages are unbelievable, and some individuals have extemely unusual wages in some years.}\label{fig:feature-plot}
 \end{figure}
-
-In Figure \ref{fig:high-wages}, we plot some respondents with a high value of mean hourly wages.
-We filter all of the IDs who earned more than US\$ 500 per hour on average.
-We find that these respondents only experienced one point of extremely high wages.
-Thus, we suspect that these high values are erroneous resulted from a data entry error.
-
-Further, we take 36 samples randomly from the data and plot them, as shown in Figure \ref{fig:sample-plot}.
-It implies that not only that some observations earn extremely high figures of wages, but some also have reasonably fluctuated wages, for example, the IDs in panel numbers 5, 7, and 11.
-The plot also implies that the samples have a different pattern of mean hourly wages.
-Some have had flat wages for years but had a sudden increase in one particular year, then it gone down again, while the others experienced an upsurge in their wage, for instance, the IDs in panel 9.
 

### 2021-09-19 paper/paper-jsdse/paper-jsdse.tex
box get data cleaned up

@@
 
-In this section, we outline the steps taken to the download the raw data from the database provided at {[}CITE{]}. Specifically, we only extract a subset of the variables in the data base.
+In this section, we outline the steps taken to the download the raw data from the database provided at {[}CITE: \url{https://www.nlsinfo.org/content/cohorts/nlsy79/get-data}{]}.
+Specifically, we only extract a subset of the variables in the database as shown in \ref{box:data}.
 
@@
 
-\begin{tcolorbox}[label="box:data"]
-
-Education, Training and Achievement Scores
+\begin{tcolorbox}[label="box:data", title="Getting data"]
 
 \begin{itemize}
-\tightlist
-\item
-  Education -\textgreater{} Summary measures -\textgreater{} All schools -\textgreater{} By year -\textgreater{} Highest grade completed
-
+\item[$\triangleright$] Education, Training and Achievement Scores
 \begin{itemize}
-  \tightlist
-  \item
-    Downloaded all of variables in Highest grade completed.
+\item[$\triangleright$] Education $\triangleright$ Summary measures $\triangleright$ All schools $\triangleright$ By year $\triangleright$ Highest grade completed
+\begin{itemize}
+\item Downloaded all of variables in Highest grade completed.
 \end{itemize}
 \end{itemize}
-
-Employment
-
+\item[$\triangleright$] Employment
 \begin{itemize}
-\tightlist
-\item
-  Summary measures -\textgreater{} By job -\textgreater{} Hours worked and Hourly wages
-
+\item[$\triangleright$] Summary measures $\triangleright$ By job $\triangleright$ Hours worked and Hourly wages
 \begin{itemize}
-  \tightlist
-  \item
-    Downloaded all of the variables in Hours worked.
-  \item
-    Downloaded all of the variables in Hourly wages
+\item Downloaded all of the variables in Hours worked.
+\item Downloaded all of the variables in Hourly wages
   Both hours worked and hourly wages are recorded by the job, up to five jobs for each id/subject. 
@@
 \end{itemize}
-
-Household, Geography and Contextual Variables
-
+\item[$\triangleright$] Household, Geography and Contextual Variables
 \begin{itemize}
-\tightlist
-\item
-  Context -\textgreater{} Summary measures -\textgreater{} Basic demographics
-
+\item[$\triangleright$] Context $\triangleright$ Summary measures $\triangleright$ Basic demographics 
 \begin{itemize}
-  \tightlist
-  \item
-    Downloaded year and month of birth, race, and sex variables.
-    There are two versions of the year and month of birth, i.e., 1979 and 1981 data. We downloaded these two versions.
+\item Downloaded year and month of birth, race, and sex variables. There are two versions of the year and month of birth, i.e., 1979 and 1981 data. We downloaded these two versions.
+\end{itemize}
 \end{itemize}

### 2021-09-19 paper/paper.Rmd
box get data cleaned up

--- 2021-09-16 paper/paper.Rmd
@@
   - \usepackage{tcolorbox}
+  - \usepackage{enumitem}
 output: 
@@
 
+In this section, we outline the steps taken to the download the raw data from the database provided at [CITE: https://www.nlsinfo.org/content/cohorts/nlsy79/get-data].
+Specifically, we only extract a subset of the variables in the database as shown in \ref{box:data}.
+
 ## Getting and Tidying the Data
@@
 
-\begin{tcolorbox}[width=\textwidth,
-                  boxsep=5pt,
-                  left=5pt,
-                  right=5pt,
-                  top=5pt,
-                  ]
-Education, Training and Achievement Scores
+::: {.tcolorbox data-latex="[label=\"box:data\", title=\"Getting data\"]" }
 
-- Education -> Summary measures -> All schools -> By year -> Highest grade completed
-  - Downloaded all of variables in Highest grade completed.
+\begin{itemize}
+\item[$\triangleright$] Education, Training and Achievement Scores
+\begin{itemize}
+\item[$\triangleright$] Education $\triangleright$ Summary measures $\triangleright$ All schools $\triangleright$ By year $\triangleright$ Highest grade completed
+\begin{itemize}
+\item Downloaded all of variables in Highest grade completed.
+\end{itemize}
+\end{itemize}
+\item[$\triangleright$] Employment
+\begin{itemize}
+\item[$\triangleright$] Summary measures $\triangleright$ By job $\triangleright$ Hours worked and Hourly wages
+\begin{itemize}
+\item Downloaded all of the variables in Hours worked.
+\item Downloaded all of the variables in Hourly wages
+  Both hours worked and hourly wages are recorded by the job, up to five jobs for each id/subject. 
+\end{itemize}
+\end{itemize}
+\item[$\triangleright$] Household, Geography and Contextual Variables
+\begin{itemize}
+\item[$\triangleright$] Context $\triangleright$ Summary measures $\triangleright$ Basic demographics 
+\begin{itemize}
+\item Downloaded year and month of birth, race, and sex variables. There are two versions of the year and month of birth, i.e., 1979 and 1981 data. We downloaded these two versions.
+\end{itemize}
+\end{itemize}
+\end{itemize}
 
-Employment
-
-- Summary measures -> By job -> Hours worked and Hourly wages
-  - Downloaded all of the variables in Hours worked.
-  - Downloaded all of the variables in Hourly wages
-  Both hours worked and hourly wages are recorded by the job, up to five jobs for each id/subject.
-
-Household, Geography and Contextual Variables
-
-- Context -> Summary measures -> Basic demographics 
-  - Downloaded year and month of birth, race, and sex variables.
-  There are two versions of the year and month of birth, i.e., 1979 and 1981 data. We downloaded these two versions.
-
-\end{tcolorbox}
+:::
 

### 2021-09-19 paper/paper-jsdse/paper-jsdse.Rmd
Intro clean up

@@
 #preamble: >
-header-includes:
-  - \usepackage{tcolorbox}
 output: 

### 2021-09-19 paper/paper-jsdse/paper-jsdse.tex
Intro clean up

@@
 \usepackage{tcolorbox}
+\usepackage{fontawesome}
 \usepackage{url} % not crucial - just used below for the URL
@@
 Data cleaning can be considered to be a part of what is called ``initial data analysis (IDA)'' (Chatfield 1985).
-In IDA one would also explore the data, especially to check if the dta is consistent with assumptions required for modeling.
-This is also related to exploratory data analysis, which Tukey (1977) defines as detective work to learn from data, either numerically or graphically, before confirmatory data analysis is performed, encompasses IDA.
+In IDA one would also explore the data, especially to check if the data is consistent with assumptions required for modeling.
+This is also related to exploratory data analysis, which Tukey (1977) defines as detective work to learn from data, either numerically or graphically, before confirmatory data analysis is performed, which encompasses IDA.
 Dasu and Johnson (2003) say that data cleaning and exploration, without naming it as IDA, is a difficult task and consumes 80\% of the data mining task.
@@
 2001).
-In addition, The National Longitudinal Survey is considered a survey with high retention rates and carefully designed, making it suitable for a life course research (Pergamit et al.
-2001) and (Cooksey 2017).
+In addition, The NLSY is considered a survey with high retention rates and carefully designed, making it suitable for a life course research (Pergamit et al.
+2001; Cooksey 2017).
 According to Cooksey (2017), thousand of articles, and hundreds of book chapters and monographs has incorporated the NLSY data.
@@
 Bureau of Labor Statistics, n.d.).
-The cohort originally included 12,686 respondents ages 14-22 when first interviewed in 1979.
+The cohort originally included 12,686 respondents aged 14-22 when first interviewed in 1979.
 It was comprised of Blacks, Hispanics, economically disadvantaged non-Black non-Hispanics, and youth in the military.
@@
 
-In this section, we outline the steps taken to the download the raw data from the database provided at {[}CITE: \url{https://www.nlsinfo.org/content/cohorts/nlsy79/get-data}{]}.
+In this section, we outline the steps taken to the download the raw data from the database provided at \url{https://www.nlsinfo.org/content/cohorts/nlsy79/get-data} (The U.S.
+Bureau of Labor Statistics, n.d.).
 Specifically, we only extract a subset of the variables in the database as shown in \ref{box:data}.
@@
 
-\begin{tcolorbox}[label="box:data", title="Getting data"]
+\begin{tcolorbox}[label="box:data", title=Getting the data]
+
+\faDatabase National Longitudinal Survey of Youth 1979
 
@@
 \begin{itemize}
-\item Downloaded all of variables in Highest grade completed.
+\item[\faCloudDownload] Downloaded all of variables in Highest grade completed.
 \end{itemize}
@@
 \begin{itemize}
-\item Downloaded all of the variables in Hours worked.
-\item Downloaded all of the variables in Hourly wages
+\item[\faCloudDownload] Downloaded all of the variables in Hours worked.
+\item[\faCloudDownload] Downloaded all of the variables in Hourly wages
   Both hours worked and hourly wages are recorded by the job, up to five jobs for each id/subject. 
@@
 \begin{itemize}
-\item Downloaded year and month of birth, race, and sex variables. There are two versions of the year and month of birth, i.e., 1979 and 1981 data. We downloaded these two versions.
+\item[\faCloudDownload] Downloaded year and month of birth, race, and sex variables. There are two versions of the year and month of birth, i.e., 1979 and 1981 data. We downloaded these two versions.
 \end{itemize}

### 2021-09-19 paper/paper.Rmd
Intro clean up

@@
   - \usepackage{tcolorbox}
-  - \usepackage{enumitem}
+  - \usepackage{fontawesome}
 output: 
@@
 Data cleaning can be considered to be a part of what is called "initial data analysis (IDA)" [@Chatfield1985TIEo].
-In IDA one would also explore the data, especially to check if the dta is consistent with assumptions required for modeling.
-This is also related to exploratory data analysis, which @tukey defines as detective work to learn from data, either numerically or graphically, before confirmatory data analysis is performed, encompasses IDA.
+In IDA one would also explore the data, especially to check if the data is consistent with assumptions required for modeling.
+This is also related to exploratory data analysis, which @tukey defines as detective work to learn from data, either numerically or graphically, before confirmatory data analysis is performed, which encompasses IDA.
 @DasuTamraparni2003Edma say that data cleaning and exploration, without naming it as IDA, is a difficult task and consumes 80% of the data mining task.
@@
 This data has been playing an important role in research in various disciplines, including but are not limited to economics, sociology, education, public policy, and public health for more than a quarter of the century [@MichaelRPergamit2001DWTN].
-In addition, The National Longitudinal Survey is considered a survey with high retention rates and carefully designed, making it suitable for a life course research [@MichaelRPergamit2001DWTN] and [@eliznlsy].
+In addition, The NLSY is considered a survey with high retention rates and carefully designed, making it suitable for a life course research [@MichaelRPergamit2001DWTN; @eliznlsy].
 According to @eliznlsy, thousand of articles, and hundreds of book chapters and monographs has incorporated the NLSY data.
@@
 The NLSY79 is a longitudinal survey held by the U.S Bureau of Labor Statistics that follows the lives of a sample of American youth and born between 1957-1964 [@nlsy79].
-The cohort originally included 12,686 respondents ages 14-22 when first interviewed in 1979.
+The cohort originally included 12,686 respondents aged 14-22 when first interviewed in 1979.
 It was comprised of Blacks, Hispanics, economically disadvantaged non-Black non-Hispanics, and youth in the military.
@@
 
-In this section, we outline the steps taken to the download the raw data from the database provided at [CITE: https://www.nlsinfo.org/content/cohorts/nlsy79/get-data].
+In this section, we outline the steps taken to the download the raw data from the database provided at \url{https://www.nlsinfo.org/content/cohorts/nlsy79/get-data} [@nlsy79].
 Specifically, we only extract a subset of the variables in the database as shown in \ref{box:data}.
@@
 
-::: {.tcolorbox data-latex="[label=\"box:data\", title=\"Getting data\"]" }
+::: {.tcolorbox data-latex="[label=\"box:data\", title=Getting the data]" }
+\faDatabase National Longitudinal Survey of Youth 1979
 

### 2021-09-20 paper/paper.Rmd
Modify chunk label so that Figure 4 does not display same plot fig.2

@@
 
-Further, \ref{fig:comppict} A shows that after eliminating the extreme values, the highest value has decreased to be around $350.
+Further, \ref{fig:fixed-feature-plot} A shows that after eliminating the extreme values, the highest value has decreased to be around $350.
 The spikes are still observed but not as extreme as the original data set.
-In Figure \ref{fig:comppict} B), we plot the three features of mean hourly wages, namely the minimum, median, and maximum value, transformed to log scale.
+In Figure \ref{fixed-feature-plot} B, we plot the three features of mean hourly wages, namely the minimum, median, and maximum value, transformed to log scale.
 The plot implies that the skewness is slightly negative.
@@
 It indicates that some ID's minimum wages are higher than some ID's maximum wages.
-
-```{r comppict, fig.cap = "The distribution of the mean hourly wage after the extreme values are replaced.
-Plot A portrays the pattern of mean hourly wage of high school cohort from 1979 to 2018 of each ID in US Dollar; Plot B shows the distribution of their minimum, median, and maximum value transformed to log10 scale.
-We can see that some observations still had reasonbaly higher wages than the others.
-Also, some IDs' have a minimum wages that is higher than others' maximum wages.", fig.width=6, fig.height=4, eval=FALSE}
-```
 

### 2021-09-20 paper/paper.tex
Modify chunk label so that Figure 4 does not display same plot fig.2

@@
 % Pandoc citation processing
-\newlength{\csllabelwidth}
-\setlength{\csllabelwidth}{3em}
-\newlength{\cslhangindent}
-\setlength{\cslhangindent}{1.5em}
-% for Pandoc 2.8 to 2.10.1
-\newenvironment{cslreferences}%
-  {}%
-  {\par}
-% For Pandoc 2.11+
-\newenvironment{CSLReferences}[3] % #1 hanging-ident, #2 entry spacing
- {% don't indent paragraphs
-  \setlength{\parindent}{0pt}
-  % turn on hanging indent if param 1 is 1
-  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
-  % set entry spacing
-  \ifnum #2 > 0
-  \setlength{\parskip}{#2\baselineskip}
-  \fi
- }%
- {}
-\usepackage{calc} % for calculating minipage widths
-\newcommand{\CSLBlock}[1]{#1\hfill\break}
-\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
-\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}}
-\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
 
 \usepackage{tcolorbox}
-\newtcolorbox{tbox}{width=\textwidth,boxsep=5pt}
+\usepackage{fontawesome}
 \usepackage{booktabs}
@@
 
-\textbf{Keywords}: Data cleaning; Data tidying; Longitudinal data; NLSY79; Open data; Initial data analysis; Outlier detection; Robust linear regression
+\keywords{
+    Data cleaning; Data tidying; Longitudinal data; NLSY79; Open data; Initial data analysis; Outlier detection; Robust linear regression
+  }
 
@@
 However, open data is often referred to as what we might call ``wild data'' because it requires substantial cleaning and tidying to tame it into textbook shape.
-Huebner Marianne, Vach, and Cessie (2016) emphasize that making the data cleaning process accountable and transparent is imperative.
-Documenting the data cleaning is essential for the integrity of downstream statistical analyses and model building (M.
-Huebner et al.
+Huebner, Vach, and Cessie (2016) emphasize that making the data cleaning process accountable and transparent is imperative.
+Documenting the data cleaning is essential for the integrity of downstream statistical analyses and model building (Huebner et al.
 2020).
 Data cleaning can be considered to be a part of what is called ``initial data analysis (IDA)'' (Chatfield 1985).
-In IDA one would also explore the data, especially to check if the dta is consistent with assumptions required for modeling.
-This is also related to exploratory data analysis, which Tukey (1977) defines as detective work to learn from data, either numerically or graphically, before confirmatory data analysis is performed, encompasses IDA.
+In IDA one would also explore the data, especially to check if the data is consistent with assumptions required for modeling.
+This is also related to exploratory data analysis, which Tukey (1977) defines as detective work to learn from data, either numerically or graphically, before confirmatory data analysis is performed, which encompasses IDA.
 Dasu and Johnson (2003) say that data cleaning and exploration, without naming it as IDA, is a difficult task and consumes 80\% of the data mining task.
@@
 
+In this section, we outline the steps taken to the download the raw data from the database provided at \url{https://www.nlsinfo.org/content/cohorts/nlsy79/get-data} (The U.S.
+Bureau of Labor Statistics, n.d.).
+Specifically, we only extract a subset of the variables in the database as shown in \ref{box:data}.
+
 \hypertarget{getting-and-tidying-the-data}{%
@@
 
-Education, Training and Achievement Scores
+\begin{tcolorbox}[label="box:data", title=Getting the data]
+
+\faDatabase National Longitudinal Survey of Youth 1979
 
 \begin{itemize}
-\tightlist
-\item
-  Education -\textgreater{} Summary measures -\textgreater{} All schools -\textgreater{} By year -\textgreater{} Highest grade completed
-
+\item[$\triangleright$] Education, Training and Achievement Scores
 \begin{itemize}
-  \tightlist
-  \item
-    Downloaded all of variables in Highest grade completed.
+\item[$\triangleright$] Education $\triangleright$ Summary measures $\triangleright$ All schools $\triangleright$ By year $\triangleright$ Highest grade completed
+\begin{itemize}
+\item[\faCloudDownload] Downloaded all of variables in Highest grade completed.
+\end{itemize}
+\end{itemize}
+\item[$\triangleright$] Employment
+\begin{itemize}
+\item[$\triangleright$] Summary measures $\triangleright$ By job $\triangleright$ Hours worked and Hourly wages
+\begin{itemize}
+\item[\faCloudDownload] Downloaded all of the variables in Hours worked.
+\item[\faCloudDownload] Downloaded all of the variables in Hourly wages
+  Both hours worked and hourly wages are recorded by the job, up to five jobs for each id/subject. 
+\end{itemize}
+\end{itemize}
+\item[$\triangleright$] Household, Geography and Contextual Variables
+\begin{itemize}
+\item[$\triangleright$] Context $\triangleright$ Summary measures $\triangleright$ Basic demographics 
+\begin{itemize}
+\item[\faCloudDownload] Downloaded year and month of birth, race, and sex variables. There are two versions of the year and month of birth, i.e., 1979 and 1981 data. We downloaded these two versions.
+\end{itemize}
 \end{itemize}
@@
 
-Employment
-
-\begin{itemize}
-\tightlist
-\item
-  Summary measures -\textgreater{} By job -\textgreater{} Hours worked and Hourly wages
-
-  \begin{itemize}
-  \tightlist
-  \item
-    Downloaded all of the variables in Hours worked.
-  \item
-    Downloaded all of the variables in Hourly wages
-    Both hours worked and hourly wages are recorded by the job, up to five jobs for each id/subject.
-  \end{itemize}
-\end{itemize}
-
-Household, Geography and Contextual Variables
-
-\begin{itemize}
-\tightlist
-\item
-  Context -\textgreater{} Summary measures -\textgreater{} Basic demographics
-
-  \begin{itemize}
-  \tightlist
-  \item
-    Downloaded year and month of birth, race, and sex variables.
-    There are two versions of the year and month of birth, i.e., 1979 and 1981 data. We downloaded these two versions.
-  \end{itemize}
-\end{itemize}
+\end{tcolorbox}
 
@@
 
-Further, \ref{fig:comppict} A shows that after eliminating the extreme values, the highest value has decreased to be around \$350.
+Further, \ref{fig:fixed-feature-plot} A shows that after eliminating the extreme values, the highest value has decreased to be around \$350.
 The spikes are still observed but not as extreme as the original data set.
-In Figure \ref{fig:comppict} B), we plot the three features of mean hourly wages, namely the minimum, median, and maximum value, transformed to log scale.
+In Figure \ref{fixed-feature-plot} B, we plot the three features of mean hourly wages, namely the minimum, median, and maximum value, transformed to log scale.
 The plot implies that the skewness is slightly negative.
@@
 
-The entire analysis is conducted using \texttt{R} (R Core Team 2020) in \texttt{rstudio} using these packages: \texttt{tidyverse} (Wickham et al.
-2019), \texttt{ggplot2} (Wickham 2016), \texttt{dplyr} (Wickham et al.
-2020), \texttt{readr} (Wickham and Hester 2020), \texttt{tidyr} (Wickham 2020b), \texttt{stringr} (Wickham 2019), \texttt{purrr} (Henry and Wickham 2020), \texttt{broom} (Robinson, Hayes, and Couch 2021), \texttt{brolgar} (Tierney, Cook, and Prvan 2020), \texttt{patchwork} (Pedersen 2020), \texttt{kableExtra} (Zhu 2019), \texttt{MASS} (Venables and Ripley 2002), \texttt{janitor} (Firke 2020), \texttt{DiagrammeR} (Iannone 2020), \texttt{rsvg} (Ooms 2020), \texttt{webshot} (Chang 2019), \texttt{mgcv} (Wood 2003), \texttt{tsibble} (Wang, Cook, and Hyndman 2020), and \texttt{modelr} (Wickham 2020a).
+The entire analysis is conducted using \texttt{R} (R Core Team 2020) in \texttt{rstudio} using these packages: \texttt{tidyverse} (H.
+Wickham, Averick, et al.
+2019), \texttt{ggplot2} (Wickham 2016), \texttt{dplyr} (H.
+Wickham, François, et al.
+2020), \texttt{readr} (H.
+Wickham and Hester 2020), \texttt{tidyr} (H.
+Wickham 2020b), \texttt{stringr} (H.
+Wickham 2019), \texttt{purrr} (Henry and Wickham 2020), \texttt{broom} (Robinson, Hayes, and Couch 2021), \texttt{brolgar} (Tierney, Cook, and Prvan 2020), \texttt{patchwork} (Pedersen 2020), \texttt{kableExtra} (Zhu 2019), \texttt{MASS} (Venables and Ripley 2002), \texttt{janitor} (Firke 2020), \texttt{DiagrammeR} (Iannone 2020), \texttt{rsvg} (Ooms 2020), \texttt{webshot} (Chang 2019), \texttt{mgcv} (Wood 2003), \texttt{tsibble} (Wang, Cook, and Hyndman 2020), and \texttt{modelr} (H.
+Wickham 2020a).
 The paper are generated using \texttt{knitr} (Xie 2014), \texttt{rmarkdown} (Xie, Dervieux, and Riederer 2020), and \texttt{rticles} (Allaire et al.

### 2021-09-20 paper/scripts/ida.R

@@ -217,7 +217,7 @@ ggplot(wages_compare) +
   theme(axis.text.x = element_text(angle = 10, size = 5),
         legend.position = "bottom")
 
-## ---- comppict
+## ---- fixed-feature-plot
 spag2 <- wages_cleaned %>%
   ggplot(aes(x = year,
              y = wages_rlm,
@@ -276,81 +276,6 @@ plot_high_after <- ggplot(filter(wages_cleaned, id == 39)) +
 
 
 spag2 + feature2_bp + plot_high_after
-
-
-## ---- fixed-feature-plot
-wages_high <- filter(wages_before, mean_hourly_wage > 500) %>%
-  as_tibble() %>%
-  head(n = 6)
-
-wages_high2 <- wages_before %>%
-  filter(id %in% wages_high$id)
-
-spag <- wages_before %>%
-  ggplot(aes(x = year,
-             y = mean_hourly_wage,
-             group = id)) +
-  geom_line(alpha = 0.1) +
-  scale_x_continuous("",
-                     breaks = seq(1980, 2020, 10),
-                     labels = c("80", "90", "00", "10", "20"),
-                     minor_breaks = seq(1980, 2020, 5)) +
-  ggtitle("A") +
-  theme(plot.title = element_text(size = 10)) +
-  theme_bw() +
-  ylab("Hourly wage ($)")
-
-wages_three_feat <- wages_before_tsibble %>%
-  features(mean_hourly_wage,
-           feat_three_num
-  )
-wages_feat_long <- wages_three_feat %>%
-  pivot_longer(c(min, med, max),
-               names_to = "feature", values_to = "value") %>%
-  mutate(feature = factor(feature, levels = c("min", "med", "max")))
-
-feature <- ggplot(wages_feat_long) +
-  geom_density(aes(x = value, colour = feature, fill = feature), alpha = 0.3) +
-  ggtitle("B") +
-  theme_bw() +
-  theme(plot.title = element_text(size = 10))
-
-feature_bp <- ggplot(wages_feat_long,
-                     aes(y=value, x = feature,
-                         fill = feature, color = feature)) +
-  geom_boxplot() +
-  theme_bw() +
-  #ggtitle("C") +
-  ggtitle("B")  +
-  ylab("Hourly wage ($)") +
-  theme(legend.position = "none",
-        plot.title = element_text(size = 10))
-
-plot_high <- ggplot(filter(wages_high2, id == 39)) +
-  geom_line(aes(x = year,
[…]

### 2021-09-20 paper/paper-jsdse/paper-jsdse.tex
Edit the flowcharts.

@@
 
-{\centering \includegraphics[width=468px]{figures/flow-chart-1}
+{\centering \includegraphics[width=0.93\linewidth]{/Users/deeamaliah/Documents/summer_project/summer-wages-refresh/paper/figures/flowchart}
 

### 2021-09-20 paper/paper.Rmd
Edit the flowcharts.

@@
 
-```{r flow-chart, fig.cap="The stages of data filtering from the raw data to get three datasets contained yowie, n means the number of ID, while n\\_obs means the number of observations."}
+```{r flow-chart, fig.cap="The stages of data filtering from the raw data to get three datasets contained yowie, n means the number of ID, while n\\_obs means the number of observations.", out.width="93%"}
 ```

### 2021-09-20 paper/scripts/data-preprocessing.R

@@ -16,6 +16,7 @@ library(lme4)
 library(modelr)
 library(forcats)
 library(yowie)
+library(knitr)
 
 
 ## ---- raw-data

### 2021-09-20 paper/scripts/ida.R

@@ -357,30 +357,31 @@ n_obs_do_pred = filter(wages_hs_do, is_pred == TRUE) %>%
   nrow()
 
 ### ---- flow-chart
-grViz("digraph flowchart {
-      node [fontname = Helvetica, shape = rectangle]
-      tab1 [label = '@@1']
-      tab2 [label = '@@2']
-      tab3 [label = '@@3']
-      tab4 [label = '@@4']
-      tab5 [label = '@@5']
-      tab6 [label = '@@6']
-      tab7 [label = '@@7']
-
-      tab1 -> tab2;
-      tab1 -> tab3;
-      tab1 -> tab4 -> tab5 -> tab6;
-      tab5 -> tab7;
-      tab6 -> tab7;
-      tab2 -> tab7
-      }
-
-      [1]: paste0('NLSY79, n = ', a)
-      [2]: paste0('Extract demographic variable, n = ', b)
-      [3]: paste0('Exclude ', a - d, ' IDs whose hourly rate is missing')
-      [4]: paste0('Eligible ID, n = ', d)
-      [5]: paste0('Cohort whose hgc is up to 12th grade and \\n participated at least 5 years in the survey, \\n n = ', n_hgc, ', n_obs = ', n_obs_hgc, ' \\n (', n_obs_hgc_pred, ' observations are predicted value)')
-      [6]: paste0('High school dropouts cohort, \\n n = ', n_do, ', n_obs = ', n_obs_do, ' \\n (', n_obs_do_pred, ' observations are predicted value)')
-      [7]: 'yowie Package'
-      ",
-            height = 700)
+include_graphics(here::here("paper/figures/flowchart.png"))
+# grViz("digraph flowchart {
+#       node [fontname = Helvetica, shape = rectangle]
+#       tab1 [label = '@@1']
+#       tab2 [label = '@@2']
+#       tab3 [label = '@@3']
+#       tab4 [label = '@@4']
+#       tab5 [label = '@@5']
+#       tab6 [label = '@@6']
+#       tab7 [label = '@@7']
+#
+#       tab1 -> tab2;
+#       tab1 -> tab3;
+#       tab1 -> tab4 -> tab5 -> tab6;
+#       tab5 -> tab7;
+#       tab6 -> tab7;
+#       tab2 -> tab7
+#       }
+#
+#       [1]: paste0('NLSY79, n = ', a)
+#       [2]: paste0('Extract demographic variable, n = ', b)
[…]

### 2021-09-20 paper/scripts/ida.R
Tidy plots.

@@ -43,13 +43,14 @@ ggplot(wages_before_tsibble,
            group = id)) +
   geom_line(alpha = 0.7) +
   facet_sample() +
-  scale_x_continuous("",
+  scale_x_continuous("Year",
                      breaks = seq(1980, 2020, 10),
                      labels = c("80", "90", "00", "10", "20"),
                      minor_breaks = seq(1980, 2020, 5)) +
+  theme_bw() +
   theme(axis.text.x = element_text(angle = 10, size = 6)) +
-  ylab("Hourly wage ($)") +
-  theme_bw()
+  ylab("Hourly wage ($)")
+
 
 ## ---- feature-plot
 wages_high <- filter(wages_before, mean_hourly_wage > 500) %>%
@@ -64,14 +65,14 @@ spag <- wages_before %>%
              y = mean_hourly_wage,
              group = id)) +
   geom_line(alpha = 0.1) +
-  scale_x_continuous("",
+  scale_x_continuous("Year",
                      breaks = seq(1980, 2020, 10),
                      labels = c("80", "90", "00", "10", "20"),
                      minor_breaks = seq(1980, 2020, 5)) +
   ggtitle("A") +
-  theme(plot.title = element_text(size = 10)) +
   theme_bw() +
-  ylab("Hourly wage ($)")
+  ylab("Hourly wage ($)") +
+  theme(plot.title = element_text(size = 10))
 
 wages_three_feat <- wages_before_tsibble %>%
   features(mean_hourly_wage,
@@ -98,6 +99,7 @@ feature_bp <- ggplot(wages_feat_long,
   #ggtitle("C") +
   ggtitle("B")  +
   ylab("Hourly wage ($)") +
+  xlab("Feature") +
   theme(legend.position = "none",
         plot.title = element_text(size = 10))
 
@@ -109,15 +111,15 @@ plot_high <- ggplot(filter(wages_high2, id == 39)) +
              size = 0.5,
              alpha = 0.5) +
   annotate("text", x = 2010, y = 1000, label = "id: 39") +
-  scale_x_continuous("",
+  scale_x_continuous("Year",
                      breaks = seq(1980, 2020, 10),
                      labels = c("80", "90", "00", "10", "20"),
                      minor_breaks = seq(1980, 2020, 5)) +
-  theme(axis.text.x = element_text(angle = 10, size = 6),
[…]

### 2021-09-20 R/old_eda_paper.Rmd
Edit eda section.

--- 2021-09-20 paper/paper.Rmd
@@
 ---
-title: "The Journey from Wild to Textbook Data: A Case Study from the National Longitudinal Survey of Youth"
-authors:
-  - name: Dewi Amaliah
-    thanks: ""
-    department: Department of Econometrics and Business Statistics
-    affiliation: Monash University
-    location: Clayton, VIC 3800
-    email: dama0007@student.monash.edu
-  - name: Dianne Cook
-    department: Department of Econometrics and Business Statistics
-    affiliation: Monash University
-    location: Clayton, VIC 3800
-    email: dicook@monash.edu
-  - name: Emi Tanaka
-    department: Department of Econometrics and Business Statistics
-    affiliation: Monash University
-    location: Clayton, VIC 3800
-    email: emi.tanaka@monash.edu
-  - name: Kate Hyde
-    thanks: ""
-    department: Department of Econometrics and Business Statistics
-    affiliation: Monash University
-    location: Clayton, VIC 3800
-    email: ""
-  - name: Nicholas Tierney
-    thanks: ""
-    department: ""
-    affiliation: ""
-    location: ""
-    email: ""
-abstract: > 
- The National Longitudinal Survey of Youth (NLSY79) is a prominent open data source that has been important for educational purposes and multidisciplinary research on longitudinal data.
-Subsets of this data can be found in numerous textbooks and research articles.
-However, the steps and decisions taken to get from the raw data to the textbook data is never clearly articulated.
-This article describes our journey when trying to re-create a textbook example data set from the original database, with the goal being to refresh the textbook data more regularly.
-Thus, this paper demonstrates the process -- extracting, tidying, cleaning, and exploring with documentation -- to make refreshed data available for education or research.
-Three new data sets, and the code to produce them are provided in an accompanying open source R package, called `yowie`.
-As a result of this process, some recommendations are also made for the NLSY79 curators for incorporating data quality checks, and providing more convenient samples of the data to potential users.
-bibliography: references.bib
-biblio-style: unsrt
-#preamble: >
-header-includes:
-  - \usepackage{tcolorbox}
-  - \usepackage{fontawesome}
-output: 
-  #rticles::arxiv_article:
-    #keep_tex: false
-  bookdown::pdf_book:
-    base_format: rticles::arxiv_article
-date: "24/02/2021"
-always_allow_html: true
-keywords: "Data cleaning; Data tidying; Longitudinal data; NLSY79; Open data; Initial data analysis; Outlier detection; Robust linear regression"
+title: "old_eda_paper"
+author: "Dewi"
+date: "21/09/2021"
+output: html_document
 ---
 
-```{r setup, echo = FALSE, cache = FALSE, include = FALSE}
-library(knitr)
-opts_chunk$set(echo = FALSE, warning = FALSE, 
-               message = FALSE,
-               cache = TRUE, 
-               cache.path = "paper_cache/",
-               fig.retina = 2,
-               fig.path = "figures/",
-               comment = "#>",
-               fig.align = 'center')
-
-read_chunk(here::here("paper/scripts/data-preprocessing.R"))
-read_chunk(here::here("paper/scripts/ida.R"))
-read_chunk(here::here("paper/scripts/eda.R"))
-```
-
-```{r load-pkgs}
-```
-
-# Introduction {#intro}
-
-"Open data" is data that is freely accessible, modifiable, and shareable by anyone for any purpose [@opendata].
-This type of data can be useful as example data in statistical textbooks and for research purposes.
-However, open data is often referred to as what we might call "wild data" because it requires substantial cleaning and tidying to tame it into textbook shape.
-@HuebnerMariannePhD2016Asat emphasize that making the data cleaning process accountable and transparent is imperative.
-Documenting the data cleaning is essential for the integrity of downstream statistical analyses and model building [@HuebnerMarianne2020Haar].
-Data cleaning can be considered to be a part of what is called "initial data analysis (IDA)" [@Chatfield1985TIEo].
-In IDA one would also explore the data, especially to check if the data is consistent with assumptions required for modeling.
-This is also related to exploratory data analysis, which @tukey defines as detective work to learn from data, either numerically or graphically, before confirmatory data analysis is performed, which encompasses IDA.
-@DasuTamraparni2003Edma say that data cleaning and exploration, without naming it as IDA, is a difficult task and consumes 80% of the data mining task.
-
-Despite its importance, this IDA stage is often undervalued and neglected [@Chatfield1985TIEo].
-There are few research papers that document the data cleaning [@WickhamHadley2014TD].
-Furthermore, the decisions made in this stage often go unreported in the sense that IDA is often performed in an unplanned and unstructured way and is only shared among restricted parties [@HuebnerMarianne2020Haar].
-
-This paper aims to demonstrate the steps of IDA, tidying, cleaning, and documenting the process, for a prominent open data source, National Longitudinal Survey of Youth (NLSY) 1979, henceforth referred to NLSY79.
-This data has been playing an important role in research in various disciplines, including but are not limited to economics, sociology, education, public policy, and public health for more than a quarter of the century [@MichaelRPergamit2001DWTN].
-In addition, The NLSY is considered a survey with high retention rates and carefully designed, making it suitable for a life course research [@MichaelRPergamit2001DWTN; @eliznlsy].
-According to @eliznlsy, thousand of articles, and hundreds of book chapters and monographs has incorporated the NLSY data.
-Moreover, the NLSY79 is considered the most widely used and most important cohort in the NLSY79 data sets [@MichaelRPergamit2001DWTN].
-
-@SingerJudithD2003Alda used the wages and other variables of high school dropouts from the NLSY79 data as an example data set to illustrate longitudinal data modeling.
-Our aim is to refresh this textbook data to be more up to date.
-Here, we investigate the process of getting from the raw NLSY79 to this textbook data set.
-However, we cannot create the exact same data set as published in their book since we do not have the information of what age threshold they used to determine the high school dropouts.
-
-This paper is structured to have 5 sections.
-Section \@ref(database) describes the original data source.
-Section \@ref(cleaning) presents the steps of cleaning the data, including getting and tidying the data from the NLSY79 and initial data analysis to find and treat the anomalies in the data.
-We also gave examples of exploratory data analysis using the clean data in Section \@ref(eda).
-Finally, Section \@ref(summary) summarises the paper.
-
-# The NLSY79 {#database}
-
-## Database
-
-The NLSY79 is a longitudinal survey held by the U.S Bureau of Labor Statistics that follows the lives of a sample of American youth and born between 1957-1964 [@nlsy79].
-The cohort originally included 12,686 respondents aged 14-22 when first interviewed in 1979.
-It was comprised of Blacks, Hispanics, economically disadvantaged non-Black non-Hispanics, and youth in the military.
-In 1984 and 1990, two sub-samples were dropped from the interview; the dropped subjects are the 1,079 members of the military sample and 1,643 members of the economically disadvantaged non-Black non-Hispanics, respectively.
-Hence, 9,964 respondents remain in the eligible samples.
-The surveys were conducted annually from 1979 to 1994 and biennially thereafter.
-Data are now available from Round 1 (1979 survey year) to Round 28 (2018 survey year).
-
-[FILL] We should include how many variables there are in the data so people get a grasp of the extensive coverage of the variables.
-
-Although the main focus area of the NLSY is labor and employment, the NLSY also covers several other topics, including education; training and achievement; household, geography, and contextual variables; dating, marriage, cohabitation; sexual activity, pregnancy, and fertility; children; income, assets and program participation; health; attitudes and expectations; and crime and substance use.
-
-There are two ways to conduct the interview of the NLSY79, which are face-to-face interviews or by telephone.
-In recent survey years, more than 90 percent of respondents were interviewed by telephone [@eliznlsy].
-
-## Target Data
-
-As mentioned in Section \@ref(intro), this paper aims to refresh the NLSY79 data as used in Singer and Willet's textbook (@SingerJudithD2003Alda).
-This data set contains the information of yearly mean hourly wages from the NLSY79 cohort, with education and race as covariates.
-This data set represents the measurement of male high-school dropouts, aged from 14 to 17 years old when first time measured.
-Accordingly, to get this data, we have two additional target data, NLSY79 cohort's demographic profiles and wages data of people who completed education up to high school.
-
-Since there is no information on high school dropouts in Singer and Willet's book, we define high school dropouts as respondents who only completed either $9^{th}$, $10^{th}$, $11^{th}$ grade, or who completed $12^{th}$ when their age is at least 19 years old (means that these IDs being dropped out in a certain year, but they returned back to high school to complete it).
-
-# Data Cleaning {#cleaning}
-
-In this section, we outline the steps taken to the download the raw data from the database provided at \url{https://www.nlsinfo.org/content/cohorts/nlsy79/get-data} [@nlsy79].
-Specifically, we only extract a subset of the variables in the database as shown in \ref{box:data}.
-
-## Getting and Tidying the Data
-
-The NLSY79 data are stored in a [database](https://www.nlsinfo.org/content/cohorts/nlsy79/get-data) and could be downloaded by variables.
-Several variables are available for download and could be browsed by index.
-For the wages data set, we only extracted these variables:
-
-::: {.tcolorbox data-latex="[label=\"box:data\", title=Getting the data]" }
-\faDatabase National Longitudinal Survey of Youth 1979
-
-\begin{itemize}
-\item[$\triangleright$] Education, Training and Achievement Scores
-\begin{itemize}
-\item[$\triangleright$] Education $\triangleright$ Summary measures $\triangleright$ All schools $\triangleright$ By year $\triangleright$ Highest grade completed
-\begin{itemize}
-\item[\faCloudDownload] Downloaded all of variables in Highest grade completed.
-\end{itemize}
-\end{itemize}
-\item[$\triangleright$] Employment
-\begin{itemize}
-\item[$\triangleright$] Summary measures $\triangleright$ By job $\triangleright$ Hours worked and Hourly wages
-\begin{itemize}
-\item[\faCloudDownload] Downloaded all of the variables in Hours worked.
-\item[\faCloudDownload] Downloaded all of the variables in Hourly wages
-  Both hours worked and hourly wages are recorded by the job, up to five jobs for each id/subject. 
-\end{itemize}
-\end{itemize}
-\item[$\triangleright$] Household, Geography and Contextual Variables
-\begin{itemize}
-\item[$\triangleright$] Context $\triangleright$ Summary measures $\triangleright$ Basic demographics 
-\begin{itemize}
-\item[\faCloudDownload] Downloaded year and month of birth, race, and sex variables. There are two versions of the year and month of birth, i.e., 1979 and 1981 data. We downloaded these two versions.
-\end{itemize}
-\end{itemize}
-\end{itemize}
-
-:::
-
-The downloaded data set came in a csv (NLSY79.csv) and dat (NLSY79.dat) format. We only used the .dat format. It also came along with these files:
-
-- NLSY79.NLSY79: Tagset of variables that can be uploaded to the website to recreate the data set.
-- NLSY79.R: R script provided automatically by the database for reading the data into R and converting the variables' names and label into something more sensible.
-We utilized this code to do an initial data tidying.
-It produced two data set,  `categories_qnames` (the observations are stored in categorical/interval values) and `new_data_qnames` (the observations are stored in integer form).
-
-```{r raw-data}
-```
-
-According to @WickhamHadley2014TD, tidy data sets comply with three rules.
-Firstly, each variable forms a column.
-Secondly, each observation forms a row.
-Lastly, each type of observational unit forms a table.
-Unfortunately, the `new_data_qnames` did not meet these requirements in the way that data for each year and job for the same respondent is stored as one variable.
-Hence, the data contains a huge number of columns (686 columns).
-The example of the untidy data set is displayed in Table \ref{tab:untidy-data}.
-The table intended to display the hourly rate of each respondent by job (HRP1 to HRP5) and by year (1979 and 1980).
-The table implies that the column headers are values of the year and job, not variable names.
-Consequently, the data should be tidied and wrangled first to extract the demographic and employment variables we want to put in the final data set.
-We mainly used `tidyr` [@tidyr], `dplyr` [@dplyr], and `stringr` [@stringr] to do this job.
-
-```{r untidy-data}
-```
-
-### Tidying demographic variables
-
-`Age in 1979`, `gender`, `race`, `highest grade completed`, and the `year when the highest grade completed` are the variables that we want to use in the cleaned data set.
-
-Age in 1979 are derived from year and month of birth variables in the raw data.
-The variables have two versions, which are the 1979 version and the 1981 version.
-We only used the 1979 data and did a consistency check of those years and flag the inconsistent observations.
-
-```{r dob-tidy, echo=TRUE}
-```
-
-For `gender` and `race`, we only rename the column names.
-It is worth noting that using these two variable needs special attention.
-Gender, as reported in the data, only has two categories, which is recognised today as inadequate.
-Gender is not binary.
-Further, race is as reported in the database.
-When doing analysis with this variable, one should keep in mind that the purpose is to study racism rather than race.
-
-```{r demog-tidy}
-```
-
-The next step is tidying the `highest grade completed` variable.
-This variable came with several version in each year.
-We choose the revised May data because it seemed to have less missing and presumably has been checked.
-However, there is no revised May data for 2012, 2014, 2016, and 2018.
-Thus, we just use the ordinary May data for these years.
-The code for tidying this variable is provided in the supplementary materials.
-
-```{r tidy-grade}
-```
-
-Further, `highest grade completed` is measured and could be updated in each period of the survey.
-Hence, we only used the highest grade completed ever and derived the year when it is completed.
-The code for tidying this variable is also available in the supplementary materials.
-
-```{r tidy-hgc}
-```
-
-Finally, we get all of the demographic profiles of the NLSY79 cohort. We then save this data as `demog_nlsy79`.
-
-```{r full-demog}
-```
-
-### Tidying employment variables
-
-The employment data comprises of three variables, i.e., `total hours of work per week`, `number of jobs that an individual has`, and `mean hourly wage`.
-
-`hours worked per week` initially has only one version per job, no choice from 1979 to 1987 (QES-52A).
-From 1988 onward, when we had more options, we chose the variable for total hours, including time spent working from home (QES-52D).
-However, in 1993, this variable did not have all the five D variables (the first one and the last one were missing), so we used QES-52A variable instead.
-In addition, 2008 only had jobs 1-4 for the QES-52D variable (whereas the other years had 1-5), so we use these.
-
-```{r tidy-hours, echo=TRUE}
-```
-
-The same way was also deployed to tidy the rate of wage by year and by ID.
-The difference is that the hourly rate has only one version of each year.
-The hours of work and the hourly rate are then joined to calculate the number of jobs that a respondent has and their mean hourly wage.
-Some observations have 0 in their hourly rate and work hours, which is considered an invalid value.
-Thus, these observations are set to be NA.
-Besides, there are some observations with unusual hours of work.
-Here, we also censored the observations with > 84 hours of work a week to be NA.
-
-```{r tidy-rate, echo = TRUE}
-```
-
-Since our ultimate goal is to calculate the mean hourly wage, the number of jobs is calculated based on the availability of the `rate_per_hour` information.
-For example, the number of jobs of ID 1, based on `hours_work`, is 2.
-However, since the information of hourly rate of `job_02` is not available, the number of jobs is considered as 1.
-
-Further, we calculated the mean hourly wage for each ID in each year using a weighted mean with the hours of work as the weight.
-However, there is a lot of missing value in `hours_work` variable.
-In that case, we only calculated the mean hourly wage based on the arithmetic/regular mean method.
-Hence, we created a new variable to flag whether the mean hourly wage is weighted or regular.
-Additionally, if an ID only had one job, we directly used their hourly wages information and flagged it as arithmetic mean.
-
-```{r tidy-nojob}
-```
-
-The employment and demographic variables are then joined.
-We also filtered the data to only have the cohort who completed the education up to 12th grade and participated at least five rounds in the survey and save it to an object called `wages`.
-
-```{r wages-demog-hs}
-```
-
-## Initial Data Analysis
-
-According to @HuebnerMariannePhD2016Asat, Initial Data Analysis (IDA) is the step of inspecting and screening the data after being collected to ensure that the data is clean, valid, and ready to be deployed in the later formal statistical analysis.
-Moreover, @Chatfield1985TIEo argues that the two main objectives of IDA are data description, which is to assess the structure and the quality of the data, and model formulation without any formal statistical inference.
-
-In this paper, we conduct an IDA or a preliminary data analysis to assess the consistency of the data with the cohort information that the NLSY provides.
-In addition, we also aim to find the anomaly in the wages values using this approach.
-We mainly use graphical summaries to do the IDA using `ggplot2`[@ggplot2] and `brolgar` [@brolgar].
-
-As stated previously, the respondents' ages ranged from 12 to 22 when first interviewed in 1979.
-Hence, we validate whether all of the respondents were in this range.
-Additionally, the [NLSY](https://www.nlsinfo.org/content/cohorts/nlsy79/intro-to-the-sample/nlsy79-sample-introduction) also provides the number of the survey cohort by their gender (6,403 males and 6,283 females) and race (7,510 Non-Black/Non-Hispanic; 3,174 Black; 2,002 Hispanic).
-To validate this, we used the `demog_nlsy79`, i.e., the data with the survey years 1979 sample.
-Table \ref{tab:age-table} and Table \ref{tab:gender-race-table} suggest that the demographic data we had is consistent with the sample information in the database.
-
-```{r age-table}
-```
-
-```{r gender-race-table}
-```
-
-The next step is that explore the mean hourly wage data. In this case, we only explore the wages data of up to $12^{th}$ grade.  Here, we use visualisation techniques to do IDA.
-
-```{r summarytable}
-```
-
-```{r sample-plot, echo = FALSE, fig.cap="A glimpse of the refreshed data.
-Longitudinal profiles of wages from 1979-2018 are shown for a random sample of 36 individuals, divided into twelve subplots.
-Most individuals here experienced increasing wages over time, although there is considerable fluctuation from one year to another.", fig.height=4, fig.width=5, out.width="100%"}
-```
-
-We take 36 samples randomly from the data and plot them, as shown in Figure \ref{fig:sample-plot}.
-It implies that these respondents have a lot of variability in wages, for example, the IDs in panel numbers 5, 7, and 11.
-The plot also implies that the samples have a different pattern of mean hourly wages.
-Some have had flat wages for years but had a sudden increase in one particular year, then it gone down again, while the others experienced an upsurge in their wage, for instance, the IDs in panel 9.
-
-```{r feature-plot, fig.cap = "Summary plots to check the cleaned data reveal more cleaning is necessary: longitudinal profiles of wages for all individuals 1979-2018 (A), boxplots of minimum, median, and maximum wages of each individual (B), and one individual (id=39) with an unusual wage relative to their years of data (C).
-Some values of hourly wages are unbelievable, and some individuals have extemely unusual wages in some years.", fig.width=6, fig.height=2}
-```
-
-```{r high-wages}
-```
-
-```{r high-wages, fig.cap= "6 out of 45 IDs with extremely high mean hourly wage. Most of the IDs only have one point of high wage.", eval=FALSE}
-```
-
-The anomalies are also found in the total hours of work, where some observations work for 420 hours a week in total.
-According to @MichaelRPergamit2001DWTN, one of the flaws of the NLSY79 employment data is that the NLSY79 collects the information of the working hours since the last interview.
-Thus, it might be challenging for the respondents to track the within-job hours' changes between survey years, especially for the respondents with fluctuated working hours or seasonal jobs.
-It even has been more challenging since 1994, where the respondents had to recall two years periods.
-This shortcoming might also contribute to the fluctuation of one's wages data.
-
-### Replacing extreme values
-
-As part of the IDA, which is the model formulation, we build a robust linear regression model to treat the extreme values in the data.
-Robust linear regression yields an estimation robust to the influence of noise or contamination [@KollerManuel2016rARP].
-It also aims to detect the contamination by weighting each observation based on how "well-behaved" they are, known as robustness weight.
-Observations with lower robustness weight are suggested as an outlier by this method [@KollerManuel2016rARP].
-
-Since we work with longitudinal data, we build the model for each ID instead of the overall data.
-The robust mixed model could be the best model to be employed in this case.
-However, this method is too computationally and memory expensive, especially for a large data set, like the NLSY79 data.
-Thus, the model for each ID is built utilizing the `nest` and `map` function from `tidyr` [@tidyr] and `purrr` [@purrr], respectively.
-
-We build the model using the `rlm` function from `MASS` package [@mass].
-We set the `mean_hourly_wage` and `year` as the dependent and predictor, respectively.
-Furthermore, we use M-Estimation with Huber weighting, where the observation with a slight residual gets a weight of 1, while the larger the residual, the smaller the weight (less than 1) [@rlm].
-However, the challenging part of detecting the anomaly using the robustness weight is determining the weight threshold in which the observations are considered outliers.
-Moreover, it should be noted that not all the outliers are due to an error.
-Instead, it might be that one had reasonably increasing or decreasing wages in a particular period.
-
-To minimize the risk of mistakenly identifying an outlier as an "erroneous outlier", we simulate some thresholds and study how they affect the data.
-We find that 0.12 is the most reasonable value to be the threshold to minimize that drawback's risk because it still captures the sensible spikes in the data.
-In other words, we keep maintaining the natural variability of the wages while minimizing anomalies because of the error in the data recording.
-After deciding the threshold, we impute the observations whose weights are less than 0.12 with the models' predicted value.
-We then flag those observations in a new variable called `is_pred`.
-
-```{r rlm, eval=FALSE}
-```
-
-Figure \ref{fig:compare-plot} shows the mean hourly wage before and after the extreme values are replaced.
-It implies that the fluctuation can still be observed in the data after the treatment.
-However, the large spikes, which are considered "erroneous outliers", are already eliminated from the data.
-Hence, the model produces a data set with a more reasonable degree of fluctuation.
-
-```{r compare-data}
-```
-
-```{r compare-plot, fig.cap="Comparison between the original and the treated mean hourly wage.
-The orange line portray the original value of mean hourly wage, while the turquoise line display the mean hourly wages value after the extreme values imputed with the robust linear model's prediction value.
-We can see that some extreme spikes has been reduced by the model.", fig.height=5, fig.width=6, outwidth="100%"}
-```
-
-Further, \ref{fig:fixed-feature-plot} A shows that after eliminating the extreme values, the highest value has decreased to be around $350.
-The spikes are still observed but not as extreme as the original data set.
-In Figure \ref{fixed-feature-plot} B, we plot the three features of mean hourly wages, namely the minimum, median, and maximum value, transformed to log scale.
-The plot implies that the skewness is slightly negative.
-We also see that the three features are overlapped each other.
-It indicates that some ID's minimum wages are higher than some ID's maximum wages.
-
-```{r fixed-feature-plot, fig.cap = "Re-make of the summary plots of the fully processed data suggest that the data is now in a reasonable state: longitudinal profiles of wages for all individuals 1979-2018 (A), boxplots of minimum, median, and maximum wages of each individual (B), and one individual with an unusual wage relative to their years of data (C).
-", fig.width=6, fig.height=2}
-```
-
-Finally, we save the imputed data and set the appropriate data type for the variables.
-As our target data is the mean hourly wage of the high-school dropouts, we then subset the high-school graduate data set to have only the male-high school dropout data.
-We then make these three data sets and their processing documentation publicly available through an R data container package called `yowie`.
-The complete flow from the raw data to these data set is displayed in Figure \ref{fig:flow-chart}.
-
-```{r save-data}
-```
-
-```{r nrow}
-```
-
-```{r flow-chart, fig.cap="The stages of data filtering from the raw data to get three datasets contained yowie, n means the number of ID, while n\\_obs means the number of observations."}
+```{r setup, include=FALSE}
+knitr::opts_chunk$set(echo = TRUE)
 ```
@@
 Further, we perform a linear regression to see the effect of education on wages experience.
-
-# Summary {#summary}
-
-This paper has performed stages to make open data suitable for textbook data or make it ready for research.
-In the first stage, we showed the steps performed to get the data from the NLSY79 database.
-Since the data format is untidy, we showed how the data had been tidied.
-After that, we conducted an initial data analysis to investigate and screen the quality of the data.
-We found and fixed the anomalous observations in the data set using the robust linear regression model with its predicted values.
-We also performed an example of exploratory data analysis using the cleaned data set.
-
-This paper has also demonstrated data cleaning documentation by providing all of the codes in performing data tidying and initial data analysis.
-Thus, this paper provided an opportunity to continuously refresh the textbook data whenever the updated data is published in the NLSY79 database.
-It could be done by following the documentation of the code that is provided in this paper.
-
-Moreover, the documentation also includes how we generated the robustness weight and how we decided the threshold of anomalous observations.
-It is also documented along with the flag of whether an observation is an imputed value or not.
-Accordingly, if somebody wishes to make another decision, it can be done by making a small change in the code provided.
-Further, this paper is also supplemented by a `shiny`  [@shiny] app as a simulation tool to customize the weight threshold.
-
-Finally, this paper implies that data providers should design a database that is able to produce tidy data sets.
-A data provider should also check for data anomalies before the data publishing or at least provides a set of rules or threshold values.
-For example, in this case, is the threshold of reasonable wages.
-This will greatly support the data users to validate and set the same understanding of which data are considered outliers.
-Moreover, providing validation rules would facilitate any established data validation tool, such as `validate` [@validate] package.
-In this case, we cannot use this handy validation package due to the absence of validation rules.
-
-# Acknowledgements
-
-We would like to thank Aarathy Babu for the insight and discussion during the writing of this paper.
-
-The entire analysis is conducted using `R` [@R] in `rstudio` using these packages: `tidyverse` [@tidyverse], `ggplot2` [@ggplot2], `dplyr` [@dplyr], `readr` [@readr], `tidyr` [@tidyr], `stringr` [@stringr], `purrr` [@purrr], `broom` [@broom], `brolgar` [@brolgar], `patchwork` [@patchwork], `kableExtra` [@kableExtra], `MASS` [@mass], `janitor` [@janitor], `DiagrammeR` [@DiagrammeR], `rsvg` [@rsvg], `webshot` [@webshot], `mgcv` [@mgcv], `tsibble` [@tsibble], and `modelr` [@modelr].
-The paper are generated using `knitr` [@knitr], `rmarkdown` [@rmarkdown], and `rticles` [@rticles].
-
-# Supplementary Materials
-
-- **Codes** : R script to reproduce data tidying and cleaning are available in this [Github Repository](https://github.com/numbats/yowie/blob/master/data-raw/data_preprocessing.R).
-
-- **R Package `yowie`**:`yowie` is a data container R package that contains 3 datasets, namely the high school mean hourly wage data, high school dropouts mean hourly wage data, and demographic data of the NLSY79 cohort.
-This package could be accessed [here](https://github.com/numbats/yowie).
-
-- **shiny app**: An web interactive `shiny` app to run a simulation to customize the weight threshold. This app could be accessed [here](https://github.com/numbats/summer-wages-refresh/tree/main/app).
-
-# References

### 2021-09-20 paper/paper.Rmd
Tidy plots.
Edit eda section.

@@
 
+## Recap
+
 ```{r flow-chart, fig.cap="The stages of data filtering from the raw data to get three datasets contained yowie, n means the number of ID, while n\\_obs means the number of observations.", out.width="93%"}
@@
 
-# Exploratory Data Analysis {#eda}
-
-This section will provide an example of how data sets could be used to do exploratory data analysis as in textbook data.
-We are interested in examining how the respondents experience changes in their wages based on their demographic characteristics.
-Thus, we use `wages` data instead of the refreshed `wages_hs_do` as it only has male respondents.
-
-The statistics we use in measuring the change in wages experience is the slope of the wages of each respondent. 
-Using `key_slope` function in `brolgar` [@brolgar], we perform linear regression with the log of mean hourly wage as the response variable and year as the predictor.
-Since we do modeling, it is important to look at the number of observations from each respondent.
-Figure \ref{fig:slope-nobs-plot} shows that the number of observations for each individual varies.
-Moreover, each number of observations seems to have an impact on the slope.
-Extreme values in slope are observed in a few observations, while the range of slope is more stable with more observations.
-Accordingly, we trim the data to only include respondents with at least 10 observations.
-
-```{r nobs-plot}
-```
-
-```{r slope}
-```
-
-```{r slope-nobs}
-```
-
-```{r slope-nobs-plot, fig.cap = "A) Distribution of observations.
-The numbers of rounds varies among respondents.
-Most of respondents participated in 28 rounds of survey.
-B) Distribution of wages slope by number of observations.
-The slopes' range are more stable with more observations."}
-```
-
-```{r trim-slope}
-```
-
-We then create some visualizations of five respondents with the highest slope and five respondents with the lowest slope, henceforth referred to as top-5 slopes and bottom-5 slopes, respectively.
-We plot them based on their gender and education.
-Figure \ref{fig:slope-gender} shows top-5 slopes are dominated by males, while females dominate bottom-5 slopes.
-However, as this figure only shows a small portion of the data, we further inspect the gender wage gap.
-
-```{r slope-gender, fig.cap = "Wages slope over the years based on gender.
-Most of respondents who experienced high wages slope are males.
-In contrast, respondents who experienced lower wages slope are mostly females."}
-```
-
-According to @OECD, the gender pay gap is the difference between the median wage of males and females relative to the median wage of males.
-We find that, as shown in \ref{fig:gap-gender}, over the years, females tend to earn fewer than males.
-For example, in 2018, females earned about 78 cents for every $1 earned by males.
-The widest gender gaps were observed in the 1990s.
-However, as more males participated in the survey than females from year to year, these gaps could be affected by imbalance number of observations.
-
-```{r gap-gender, fig.cap="A) Number of observations by gender and by year.
-There are more males participated in survey than females.
-The number of observations decreased over the years.
-B) Gender wage gap or the median wages of females relatives to median wages of males.
-Females tend to earn fewer than males.
-For example, in 2018 females earn about 78 cents for every \\$1 earned by males.
-Note that this is a crude measure as gap is calculated not based on other variable, such as experience and job class.", out.width="90%"}
-```
-
-According to their highest education completed, both top-5 and bottom-5 slopes are dominated by people who completed high school.
-This is understandable since the respondents are mostly completed high school.
-
-```{r slope-educ, fig.cap = "Wages slope over the years based on highest education completed. People who are in top-5 and bottom-5 wages are dominated by people who completed high school."}
-```
-
-Further, we perform a linear regression to see the effect of education on wages experience.
+# Exploratory Data Analysis
 

### 2021-09-20 R/eda_old_paper.R

@@ -1,0 +1,146 @@
+## ---- nobs-plot
+wages_obs <- yowie::wages %>%
+  features(wage, n_obs) %>%
+  arrange(n_obs)
+
+wages_obs_p <- ggplot(wages_obs, aes(x = n_obs)) +
+  geom_bar() +
+  xlab("number of observations") +
+  theme_bw() +
+  ggtitle("A)")
+
+## ---- slope
+wages <- yowie::wages %>%
+  mutate(year0 = year - 1979,
+         ln_wages = log(wage))
+
+wages_slope <- key_slope(wages, ln_wages ~ year0) %>%
+  left_join(yowie::demog_nlsy79, by = "id") %>%
+  mutate(hgc_regroup = fct_collapse(hgc, `ungraded` = "UNGRADED",
+                                    `8TH grade or less` = c("1ST GRADE", "3RD GRADE",
+                                                            "4TH GRADE", "5TH GRADE",
+                                                            "6TH GRADE", "7TH GRADE",
+                                                            "8TH GRADE"),
+                                    `9TH to 11TH grade` = c("9TH GRADE",
+                                                            "10TH GRADE",
+                                                            "11TH GRADE"),
+                                    `12TH grade` = "12TH GRADE")) %>%
+  mutate(hgc_regroup = factor(hgc_regroup, levels = c("UNGRADED",
+                                                      "8TH grade or less",
+                                                      "9TH to 11TH grade",
+                                                      "12TH grade")))
+## ---- slope-nobs
+wages_obs_slope <- left_join(wages_obs, wages_slope, by = "id")
+wages_obs_slop_p <- ggplot(wages_obs_slope, aes(x = n_obs, y = .slope_year0)) +
+  geom_point() +
+  geom_jitter() +
+  xlab("number of observations") +
+  ylab("slope") +
+  theme_bw() +
+  ggtitle("B)")
+
+## ---- slope-nobs-plot
+wages_obs_p + wages_obs_slop_p
+
+## ---- trim-slope
+slope_trimmed <- wages_obs_slope %>%
+  filter(n_obs >= 10)
+
+## ---- slope-gender
+slope_bottom_5<- slope_trimmed %>%
+  arrange(.slope_year0) %>%
[…]

### 2021-09-20 paper/scripts/eda.R

@@ -1,146 +1,30 @@
-## ---- nobs-plot
-wages_obs <- yowie::wages %>%
-  features(wage, n_obs) %>%
-  arrange(n_obs)
+## ---- sw_wages
 
-wages_obs_p <- ggplot(wages_obs, aes(x = n_obs)) +
-  geom_bar() +
-  xlab("number of observations") +
+sw_wages <- brolgar::wages %>%
+  ggplot(aes(x = xp,
+             y = ln_wages,
+             group = id)) +
+  geom_line(alpha = 0.1) +
+  geom_smooth(se = FALSE, color = "red") +
+  ggtitle("A") +
   theme_bw() +
-  ggtitle("A)")
+  ylab("ln(Hourly wage) ($)") +
+  theme(plot.title = element_text(size = 10))
 
-## ---- slope
-wages <- yowie::wages %>%
-  mutate(year0 = year - 1979,
-         ln_wages = log(wage))
+sw_wages
 
-wages_slope <- key_slope(wages, ln_wages ~ year0) %>%
-  left_join(yowie::demog_nlsy79, by = "id") %>%
-  mutate(hgc_regroup = fct_collapse(hgc, `ungraded` = "UNGRADED",
-                                    `8TH grade or less` = c("1ST GRADE", "3RD GRADE",
-                                                            "4TH GRADE", "5TH GRADE",
-                                                            "6TH GRADE", "7TH GRADE",
-                                                            "8TH GRADE"),
-                                    `9TH to 11TH grade` = c("9TH GRADE",
-                                                            "10TH GRADE",
-                                                            "11TH GRADE"),
-                                    `12TH grade` = "12TH GRADE")) %>%
-  mutate(hgc_regroup = factor(hgc_regroup, levels = c("UNGRADED",
-                                                      "8TH grade or less",
-                                                      "9TH to 11TH grade",
-                                                      "12TH grade")))
-## ---- slope-nobs
-wages_obs_slope <- left_join(wages_obs, wages_slope, by = "id")
-wages_obs_slop_p <- ggplot(wages_obs_slope, aes(x = n_obs, y = .slope_year0)) +
-  geom_point() +
-  geom_jitter() +
[…]

### 2021-09-20 paper/paper-jsdse/paper-jsdse.tex
fixed begin{document} not found issue

--- 2021-09-19 paper/paper-jsdse/paper-jsdse.tex
@@
 Bureau of Labor Statistics, n.d.).
-Specifically, we only extract a subset of the variables in the database as shown in \ref{box:data}.
+Specifically, we only extract a subset of the variables in the database as described in Section\textasciitilde{}\ref{sec:getdata}.
 
-\hypertarget{getting-and-tidying-the-data}{%
-\subsection{Getting and Tidying the Data}\label{getting-and-tidying-the-data}}
+\hypertarget{getdata}{%
+\subsection{Getting and Tidying the Data}\label{getdata}}
 
@@
 
-\faDatabase National Longitudinal Survey of Youth 1979
+\faDatabase~National Longitudinal Survey of Youth 1979
 
@@
 
+\hypertarget{recap}{%
+\subsection{Recap}\label{recap}}
+
 \begin{figure}
 
-{\centering \includegraphics[width=468px]{figures/flow-chart-1}
+{\centering \includegraphics[width=0.93\linewidth]{/Users/etan0038/Dropbox/repo/summer-wages-refresh/paper/figures/flowchart}
 
@@
 
-\hypertarget{eda}{%
-\section{Exploratory Data Analysis}\label{eda}}
-
-This section will provide an example of how data sets could be used to do exploratory data analysis as in textbook data.
-We are interested in examining how the respondents experience changes in their wages based on their demographic characteristics.
-Thus, we use \texttt{wages} data instead of the refreshed \texttt{wages\_hs\_do} as it only has male respondents.
-
-The statistics we use in measuring the change in wages experience is the slope of the wages of each respondent.
-Using \texttt{key\_slope} function in \texttt{brolgar} (Tierney, Cook, and Prvan 2020), we perform linear regression with the log of mean hourly wage as the response variable and year as the predictor.
-Since we do modeling, it is important to look at the number of observations from each respondent.
-Figure \ref{fig:slope-nobs-plot} shows that the number of observations for each individual varies.
-Moreover, each number of observations seems to have an impact on the slope.
-Extreme values in slope are observed in a few observations, while the range of slope is more stable with more observations.
-Accordingly, we trim the data to only include respondents with at least 10 observations.
-
-\begin{figure}
-
-{\centering \includegraphics[width=468px]{figures/slope-nobs-plot-1}
-
-}
-
-\caption{A) Distribution of observations.
-The numbers of rounds varies among respondents.
-Most of respondents participated in 28 rounds of survey.
-B) Distribution of wages slope by number of observations.
-The slopes' range are more stable with more observations.}\label{fig:slope-nobs-plot}
-\end{figure}
-
-We then create some visualizations of five respondents with the highest slope and five respondents with the lowest slope, henceforth referred to as top-5 slopes and bottom-5 slopes, respectively.
-We plot them based on their gender and education.
-Figure \ref{fig:slope-gender} shows top-5 slopes are dominated by males, while females dominate bottom-5 slopes.
-However, as this figure only shows a small portion of the data, we further inspect the gender wage gap.
-
-\begin{figure}
-
-{\centering \includegraphics[width=468px]{figures/slope-gender-1}
-
-}
-
-\caption{Wages slope over the years based on gender.
-Most of respondents who experienced high wages slope are males.
-In contrast, respondents who experienced lower wages slope are mostly females.}\label{fig:slope-gender}
-\end{figure}
-
-According to OECD (n.d.), the gender pay gap is the difference between the median wage of males and females relative to the median wage of males.
-We find that, as shown in \ref{fig:gap-gender}, over the years, females tend to earn fewer than males.
-For example, in 2018, females earned about 78 cents for every \$1 earned by males.
-The widest gender gaps were observed in the 1990s.
-However, as more males participated in the survey than females from year to year, these gaps could be affected by imbalance number of observations.
-
-\begin{figure}
-
-{\centering \includegraphics[width=0.9\linewidth]{figures/gap-gender-1}
-
-}
-
-\caption{A) Number of observations by gender and by year.
-There are more males participated in survey than females.
-The number of observations decreased over the years.
-B) Gender wage gap or the median wages of females relatives to median wages of males.
-Females tend to earn fewer than males.
-For example, in 2018 females earn about 78 cents for every \$1 earned by males.
-Note that this is a crude measure as gap is calculated not based on other variable, such as experience and job class.}\label{fig:gap-gender}
-\end{figure}
-
-According to their highest education completed, both top-5 and bottom-5 slopes are dominated by people who completed high school.
-This is understandable since the respondents are mostly completed high school.
-
-\begin{figure}
-
-{\centering \includegraphics[width=468px]{figures/slope-educ-1}
-
-}
-
-\caption{Wages slope over the years based on highest education completed. People who are in top-5 and bottom-5 wages are dominated by people who completed high school.}\label{fig:slope-educ}
-\end{figure}
-
-Further, we perform a linear regression to see the effect of education on wages experience.
+\hypertarget{exploratory-data-analysis}{%
+\section{Exploratory Data Analysis}\label{exploratory-data-analysis}}
 
@@
 
-\leavevmode\hypertarget{ref-OECD}{}%
-OECD. n.d. {``Gender Wage Gap.''} https://data.oecd.org/earnwage/gender-wage-gap.htm.
-
 \leavevmode\hypertarget{ref-rsvg}{}%

### 2021-09-20 paper/paper.tex
fixed begin{document} not found issue

--- 2021-09-19 paper/paper.tex
@@
 \usepackage{tcolorbox}
-\newtcolorbox{tbox}{width=\textwidth,boxsep=5pt}
+\usepackage{fontawesome}
 \usepackage{booktabs}
@@
 
-\hypertarget{getting-and-tidying-the-data}{%
-\subsection{Getting and Tidying the Data}\label{getting-and-tidying-the-data}}
+In this section, we outline the steps taken to the download the raw data from the database provided at \url{https://www.nlsinfo.org/content/cohorts/nlsy79/get-data} (The U.S.
+Bureau of Labor Statistics, n.d.).
+Specifically, we only extract a subset of the variables in the database as described in Section\textasciitilde{}\ref{sec:getdata}.
+
+\hypertarget{getdata}{%
+\subsection{Getting and Tidying the Data}\label{getdata}}
 
@@
 
-Education, Training and Achievement Scores
+\begin{tcolorbox}[label="box:data", title=Getting the data]
+
+\faDatabase~National Longitudinal Survey of Youth 1979
 
 \begin{itemize}
-\tightlist
-\item
-  Education -\textgreater{} Summary measures -\textgreater{} All schools -\textgreater{} By year -\textgreater{} Highest grade completed
-
+\item[$\triangleright$] Education, Training and Achievement Scores
 \begin{itemize}
-  \tightlist
-  \item
-    Downloaded all of variables in Highest grade completed.
+\item[$\triangleright$] Education $\triangleright$ Summary measures $\triangleright$ All schools $\triangleright$ By year $\triangleright$ Highest grade completed
+\begin{itemize}
+\item[\faCloudDownload] Downloaded all of variables in Highest grade completed.
+\end{itemize}
+\end{itemize}
+\item[$\triangleright$] Employment
+\begin{itemize}
+\item[$\triangleright$] Summary measures $\triangleright$ By job $\triangleright$ Hours worked and Hourly wages
+\begin{itemize}
+\item[\faCloudDownload] Downloaded all of the variables in Hours worked.
+\item[\faCloudDownload] Downloaded all of the variables in Hourly wages
+  Both hours worked and hourly wages are recorded by the job, up to five jobs for each id/subject. 
+\end{itemize}
+\end{itemize}
+\item[$\triangleright$] Household, Geography and Contextual Variables
+\begin{itemize}
+\item[$\triangleright$] Context $\triangleright$ Summary measures $\triangleright$ Basic demographics 
+\begin{itemize}
+\item[\faCloudDownload] Downloaded year and month of birth, race, and sex variables. There are two versions of the year and month of birth, i.e., 1979 and 1981 data. We downloaded these two versions.
+\end{itemize}
 \end{itemize}

### 2021-09-21 paper/scripts/eda.R
Compare singer-willet and hs-do data.

@@ -1,6 +1,9 @@
 ## ---- sw_wages
 
-sw_wages <- brolgar::wages %>%
+sw <- brolgar::wages %>%
+  mutate(id = as.factor(id))
+
+sw_wages <- sw %>%
   ggplot(aes(x = xp,
              y = ln_wages,
              group = id)) +
@@ -9,22 +12,39 @@ sw_wages <- brolgar::wages %>%
   ggtitle("A") +
   theme_bw() +
   ylab("ln(Hourly wage) ($)") +
+  xlab("Experience (years)") +
   theme(plot.title = element_text(size = 10))
-
-sw_wages
-
 
 ## ---- do_refreshed
 
-do_ref <- yowie::wages_hs_do %>%
+do <- yowie::wages_hs_do
+
+do_ref <- do %>%
   ggplot(aes(x = year-1979,
              y = log10(wage),
              group = id)) +
   geom_line(alpha = 0.1) +
-  ggtitle("A") +
+  ggtitle("B") +
   theme_bw() +
   ylab("ln(Hourly wage) ($)") +
+  xlab("Experience (years)")
   theme(plot.title = element_text(size = 10))
 
-do_ref
+
+## --- compare-sw-do
+
+sw_nid <- length(key_rows(sw))
+do_nid <- length(key_rows(do))
+
+sw_tib <- as_tibble(sw)
+do_tib <- as_tibble(do)
+unique(sw$id)
+
+
+# check common id in data set
+
+
+
+
+
 

### 2021-09-21 paper/scripts/eda.R
Added things in comparison.

@@ -1,47 +1,149 @@
 ## ---- sw_wages
 
-sw <- brolgar::wages %>%
-  mutate(id = as.factor(id))
+sw <- brolgar::wages
 
 sw_wages <- sw %>%
   ggplot(aes(x = xp,
-             y = ln_wages,
-             group = id)) +
-  geom_line(alpha = 0.1) +
-  geom_smooth(se = FALSE, color = "red") +
+             y = ln_wages)) +
+  geom_line(aes(group = id),alpha = 0.1) +
+  geom_smooth(se = FALSE) +
   ggtitle("A") +
   theme_bw() +
   ylab("ln(Hourly wage) ($)") +
   xlab("Experience (years)") +
-  theme(plot.title = element_text(size = 10))
+  theme(plot.title = element_text(size = 10)) +
+  ylim(-4, 6)
 
 ## ---- do_refreshed
 
-do <- yowie::wages_hs_do
+do <- yowie::wages_hs_do %>%
+  mutate(exp = year-1979,
+         lnwage = log(wage))
 
 do_ref <- do %>%
-  ggplot(aes(x = year-1979,
-             y = log10(wage),
-             group = id)) +
-  geom_line(alpha = 0.1) +
+  ggplot(aes(x = exp,
+             y = lnwage)) +
+  geom_line(aes(group = id), alpha = 0.1) +
+  geom_smooth(se = FALSE) +
   ggtitle("B") +
   theme_bw() +
   ylab("ln(Hourly wage) ($)") +
-  xlab("Experience (years)")
-  theme(plot.title = element_text(size = 10))
+  xlab("Experience (years)") +
+  theme(plot.title = element_text(size = 10)) +
+  ylim(-4, 6)
+
+sw_wages + do_ref
+
+# the two plots are not comparable
+# since the id who are in Singer-Willet and the id in do-refreshed migth be different
 
 
 ## --- compare-sw-do
 
-sw_nid <- length(key_rows(sw))
-do_nid <- length(key_rows(do))
+sw_id <- as_tibble(sw) %>%
+  group_by(id) %>%
+  count() %>%
+  dplyr::select(id)
 
-sw_tib <- as_tibble(sw)
-do_tib <- as_tibble(do)
-unique(sw$id)
-
+do_id <- as_tibble(do) %>%
+  group_by(id) %>%
+  mutate(id = as.numeric(id)) %>%
+  count() %>%
+  dplyr::select(id)
 
[…]

### 2021-09-21 paper/paper.Rmd
fixed begin{document} not found issue
replace table 1 as missing value figure 1

@@
   - \usepackage{fontawesome}
+  - \usepackage{color, colortbl}
 output: 
@@
 library(knitr)
-opts_chunk$set(echo = FALSE, warning = FALSE, 
+opts_chunk$set(echo = FALSE, 
+               warning = FALSE, 
                message = FALSE,
@@
 In this section, we outline the steps taken to the download the raw data from the database provided at \url{https://www.nlsinfo.org/content/cohorts/nlsy79/get-data} [@nlsy79].
-Specifically, we only extract a subset of the variables in the database as shown in \ref{box:data}.
+Specifically, we only extract a subset of the variables in the database as described in Section~\ref{getdata}.
 
-## Getting and Tidying the Data
+## Getting and Tidying the Data {#getdata}
 
@@
 ::: {.tcolorbox data-latex="[label=\"box:data\", title=Getting the data]" }
-\faDatabase National Longitudinal Survey of Youth 1979
+\faDatabase\ National Longitudinal Survey of Youth 1979
 
@@
 
-The downloaded data set came in a csv (NLSY79.csv) and dat (NLSY79.dat) format. We only used the .dat format. It also came along with these files:
+The downloaded data set came with the following set of files:
 
-- NLSY79.NLSY79: Tagset of variables that can be uploaded to the website to recreate the data set.
-- NLSY79.R: R script provided automatically by the database for reading the data into R and converting the variables' names and label into something more sensible.
+- `NLSY79.csv`: csv format
+- `NLSY79.dat`: dat format
+- `NLSY79.NLSY79`: Tagset of variables that can be uploaded to the website to recreate the data set.
+- `NLSY79.R`: R script provided automatically by the database for reading the data into R and converting the variables' names and label into something more sensible.
 We utilized this code to do an initial data tidying.
@@
 
-```{r untidy-data}
+```{r untidy-data, fig.cap="The missing value structure of the NLSY79 raw data."}
 ```

### 2021-09-21 paper/paper.tex
replace table 1 as missing value figure 1

@@
 \usepackage{fontawesome}
+\usepackage{color, colortbl}
 \usepackage{booktabs}
@@
 Bureau of Labor Statistics, n.d.).
-Specifically, we only extract a subset of the variables in the database as described in Section\textasciitilde{}\ref{sec:getdata}.
+Specifically, we only extract a subset of the variables in the database as described in Section\textasciitilde{}\ref{getdata}.
 
@@
 
-The downloaded data set came in a csv (NLSY79.csv) and dat (NLSY79.dat) format. We only used the .dat format. It also came along with these files:
+The downloaded data set came with the following set of files:
 
@@
 \item
-  NLSY79.NLSY79: Tagset of variables that can be uploaded to the website to recreate the data set.
+  \texttt{NLSY79.csv}: csv format
 \item
-  NLSY79.R: R script provided automatically by the database for reading the data into R and converting the variables' names and label into something more sensible.
+  \texttt{NLSY79.dat}: dat format
+\item
+  \texttt{NLSY79.NLSY79}: Tagset of variables that can be uploaded to the website to recreate the data set.
+\item
+  \texttt{NLSY79.R}: R script provided automatically by the database for reading the data into R and converting the variables' names and label into something more sensible.
 We utilized this code to do an initial data tidying.
@@
 
-\begin{table}
+\begin{figure}
 
-\caption{\label{tab:untidy-data}The Untidy Form of the NLSY79 Raw Data}
-\centering
-\begin{tabular}[t]{r|r|r|r|r|r|r}
-\hline
-CASEID\_1979 & HRP1\_1979 & HRP2\_1979 & HRP3\_1979 & HRP4\_1979 & HRP5\_1979 & HRP1\_1980\\
-\hline
-1 & 328 & NA & NA & NA & NA & NA\\
-\hline
-2 & 385 & NA & NA & NA & NA & 457\\
-\hline
-3 & 365 & NA & 275 & NA & NA & 397\\
-\hline
-4 & NA & NA & NA & NA & NA & NA\\
-\hline
-5 & 310 & 375 & NA & NA & NA & 333\\
-\hline
-6 & NA & NA & NA & 250 & NA & 275\\
-\hline
-\end{tabular}
-\end{table}
+{\centering \includegraphics[width=468px]{figures/untidy-data-1}
+
+}
+
+\caption{The missing value structure of the NLSY79 raw data.}\label{fig:untidy-data}
+\end{figure}
 

### 2021-09-21 paper/scripts/data-preprocessing.R

@@ -17,25 +17,17 @@ library(modelr)
 library(forcats)
 library(yowie)
 library(knitr)
-
+select <- dplyr::select
 
 ## ---- raw-data
 source(here::here("data-raw/NLSY79/NLSY79.R"))
 
 ## ---- untidy-data
-untidy_demo <- new_data_qnames %>%
-  as_tibble() %>%
-  # in 2018, the variable's name is Q3-4_2018, instead of HGC_2018
-  rename(HGC_2018 = `Q3-4_2018`) %>%
-  dplyr::select(CASEID_1979,
-                starts_with("HRP") &
-                  ends_with(c("1979", "1980", "1981", "1982", "1983"))) %>%
-  dplyr::select(1:7) %>%
-  head()
-
-kable(untidy_demo,
-      caption = "The Untidy Form of the NLSY79 Raw Data") %>%
-  kable_styling()
+new_data_qnames %>%
+  select(CASEID_1979,
+         starts_with("HRP") &
+           ends_with(c("1979", "1980", "1981", "1982", "1983"))) %>%
+  visdat::vis_miss()
 
 ## ---- dob-tidy
 #tidy the date of birth data

### 2021-09-21 paper/scripts/data-preprocessing.R
clean up scripts

@@ -17,7 +17,7 @@ library(modelr)
 library(forcats)
 library(yowie)
 library(knitr)
-select <- dplyr::select
+select <- select
 
 ## ---- raw-data
 source(here::here("data-raw/NLSY79/NLSY79.R"))
@@ -32,42 +32,42 @@ new_data_qnames %>%
 ## ---- dob-tidy
 #tidy the date of birth data
 dob_tidy <- new_data_qnames %>%
-  dplyr::select(CASEID_1979,
+  select(CASEID_1979,
                 starts_with("Q1-3_A~")) %>%
   mutate(dob_year = case_when(
-    # if the years recorded in both sets match, take 79 data
-    `Q1-3_A~Y_1979` == `Q1-3_A~Y_1981` ~ `Q1-3_A~Y_1979`,
-    # if the year in the 81 set is missing, take the 79 data
-    is.na(`Q1-3_A~Y_1981`) ~ `Q1-3_A~Y_1979`,
-    # if the sets don't match for dob year, take the 79 data
-    `Q1-3_A~Y_1979` != `Q1-3_A~Y_1981` ~ `Q1-3_A~Y_1979`),
-    dob_month = case_when(
-      # if months recorded in both sets match, take 79 data
-      `Q1-3_A~M_1979` == `Q1-3_A~M_1981` ~ `Q1-3_A~M_1979`,
-      # if month in 81 set is missing, take the 79 data
-      is.na(`Q1-3_A~M_1981`) ~ `Q1-3_A~M_1979`,
-      # if sets don't match for dob month, take the 79 data
-      `Q1-3_A~M_1979` != `Q1-3_A~M_1981` ~ `Q1-3_A~M_1979`),
-    # flag if there is a conflict between dob recorded in 79 and 81
-    dob_conflict = case_when(
-      (`Q1-3_A~M_1979` != `Q1-3_A~M_1981`) & !is.na(`Q1-3_A~M_1981`)
-      ~ TRUE,
-      (`Q1-3_A~Y_1979` != `Q1-3_A~Y_1981`) & !is.na(`Q1-3_A~Y_1981`)
-      ~ TRUE,
-      (`Q1-3_A~Y_1979` == `Q1-3_A~Y_1981`) &
-        (`Q1-3_A~M_1979` == `Q1-3_A~M_1981`) ~ FALSE,
-      is.na(`Q1-3_A~M_1981`) | is.na(`Q1-3_A~Y_1981`) ~ FALSE)) %>%
-  dplyr::select(CASEID_1979,
[…]

### 2021-09-21 paper/scripts/eda.R

@@ -43,13 +43,13 @@ sw_wages + do_ref
 sw_id <- as_tibble(sw) %>%
   group_by(id) %>%
   count() %>%
-  dplyr::select(id)
+  select(id)
 
 do_id <- as_tibble(do) %>%
   group_by(id) %>%
   mutate(id = as.numeric(id)) %>%
   count() %>%
-  dplyr::select(id)
+  select(id)
 
 # check common id in data set
 sw_do <- inner_join(sw_id, do_id, by = "id")

### 2021-09-21 paper/scripts/ida.R

@@ -5,7 +5,9 @@ age_table <- full_demographics %>%
 
 kable(age_table,
       caption = "Age Distribution of the NLSY79 samples",
-      col.names = c("Age", "Number of Sample")) %>%
+      col.names = c("Age", "Number of Sample"),
+      booktabs = TRUE,
+      linesep = "") %>%
   kable_styling()
 
 ## ---- gender-race-table
@@ -15,19 +17,22 @@ gender_race_table <- full_demographics %>%
   adorn_percentages("row") %>%
   adorn_pct_formatting(digits = 2) %>%
   adorn_ns(position = "front") %>%
-  mutate(gender = ifelse(gender == "MALE", "Male",
-                         ifelse(gender == "FEMALE", "Female", "Total")))
+  mutate(gender = str_to_title(gender))
 
 kable(gender_race_table,
       caption = "Gender and Race Distribution of the NLSY79 Samples",
-      col.names = c("Gender", "Hispanic", "Black", "Non-Black, Non-Hispanic", "Total")) %>%
+      col.names = c("Gender", "Hispanic", "Black", "Non-Black, Non-Hispanic", "Total"),
+      bookdtabs = TRUE,
+      linesep = "") %>%
   kable_styling() %>%
   add_header_above(c(" " = 1, "Race" = 3, " " = 1))
 
 ## ---- summarytable
 kable(as.array(summary(wages_before$mean_hourly_wage)),
       caption = "Summary Statistics of Wages of High School Data",
-      col.names = c("Statistics", "Value")) %>%
+      col.names = c("Statistics", "Value"),
+      booktabs = TRUE,
+      linesep = "") %>%
   kable_styling()
 
 ## ---- sample-plot
@@ -131,7 +136,7 @@ spag + feature_bp + plot_high +
 ## ---- rlm
 # nest the data by id to build a robust linear model
 by_id <- wages_before %>%
-  dplyr::select(id, year, mean_hourly_wage) %>%
+  select(id, year, mean_hourly_wage) %>%
   group_by(id) %>%
   nest()
 
@@ -153,11 +158,11 @@ id_w <- id_rlm %>%
                    x$w
                  })) %>%
   unnest(w) %>%
-  dplyr::select(w)
+  select(w)
 
 # bind the property of each observation with their weight
[…]

### 2021-09-21 paper/scripts/data-preprocessing.R
select fix

@@ -17,7 +17,7 @@ library(modelr)
 library(forcats)
 library(yowie)
 library(knitr)
-select <- select
+select <- dplyr::select
 
 ## ---- raw-data
 source(here::here("data-raw/NLSY79/NLSY79.R"))

### 2021-09-21 paper/paper.tex
script fixes

@@
 \NormalTok{dob\_tidy }\OtherTok{\textless{}{-}}\NormalTok{ new\_data\_qnames }\SpecialCharTok{\%\textgreater{}\%}
-\NormalTok{  dplyr}\SpecialCharTok{::}\FunctionTok{select}\NormalTok{(CASEID\_1979,}
+  \FunctionTok{select}\NormalTok{(CASEID\_1979,}
                 \FunctionTok{starts\_with}\NormalTok{(}\StringTok{"Q1{-}3\_A\textasciitilde{}"}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%}
@@
                       \FunctionTok{is.na}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}M\_1981}\StringTok{\textasciigrave{}}\NormalTok{) }\SpecialCharTok{|} \FunctionTok{is.na}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}Y\_1981}\StringTok{\textasciigrave{}}\NormalTok{) }\SpecialCharTok{\textasciitilde{}} \ConstantTok{FALSE}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%}
-\NormalTok{  dplyr}\SpecialCharTok{::}\FunctionTok{select}\NormalTok{(CASEID\_1979,}
+  \FunctionTok{select}\NormalTok{(CASEID\_1979,}
 \NormalTok{         dob\_month,}
@@
 \NormalTok{    temp }\OtherTok{\textless{}{-}}\NormalTok{ new\_data\_qnames }\SpecialCharTok{\%\textgreater{}\%}
-\NormalTok{      dplyr}\SpecialCharTok{::}\FunctionTok{select}\NormalTok{(CASEID\_1979,}
+      \FunctionTok{select}\NormalTok{(CASEID\_1979,}
                     \FunctionTok{starts\_with}\NormalTok{(}\StringTok{"QES{-}52A"}\NormalTok{) }\SpecialCharTok{\&}
@@
 \NormalTok{    temp }\OtherTok{\textless{}{-}}\NormalTok{ new\_data\_qnames }\SpecialCharTok{\%\textgreater{}\%}
-\NormalTok{      dplyr}\SpecialCharTok{::}\FunctionTok{select}\NormalTok{(CASEID\_1979,}
+      \FunctionTok{select}\NormalTok{(CASEID\_1979,}
                     \FunctionTok{starts\_with}\NormalTok{(}\StringTok{"QES{-}52D"}\NormalTok{) }\SpecialCharTok{\&}
@@
 \NormalTok{\}}
-\CommentTok{\# list to save the iteration result}
-\NormalTok{hours }\OtherTok{\textless{}{-}} \FunctionTok{list}\NormalTok{()}
+
 \CommentTok{\# getting the hours of work of all observations}
-\ControlFlowTok{for}\NormalTok{(ayear }\ControlFlowTok{in} \FunctionTok{c}\NormalTok{(}\DecValTok{1979}\SpecialCharTok{:}\DecValTok{1994}\NormalTok{, }\DecValTok{1996}\NormalTok{, }\DecValTok{1998}\NormalTok{, }\DecValTok{2000}\NormalTok{, }\DecValTok{2002}\NormalTok{, }\DecValTok{2004}\NormalTok{, }\DecValTok{2006}\NormalTok{, }\DecValTok{2008}\NormalTok{, }\DecValTok{2010}\NormalTok{,}
-               \DecValTok{2012}\NormalTok{, }\DecValTok{2014}\NormalTok{, }\DecValTok{2016}\NormalTok{, }\DecValTok{2018}\NormalTok{)) \{}
-\NormalTok{  hours[[ayear]] }\OtherTok{\textless{}{-}} \FunctionTok{get\_hour}\NormalTok{(ayear)}
-\NormalTok{\}}
-\CommentTok{\# unlist the hours of work}
-\NormalTok{hours\_all }\OtherTok{\textless{}{-}} \FunctionTok{bind\_rows}\NormalTok{(}\SpecialCharTok{!!!}\NormalTok{hours)}
+\NormalTok{years }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(}\DecValTok{1979}\SpecialCharTok{:}\DecValTok{1994}\NormalTok{, }\FunctionTok{seq}\NormalTok{(}\DecValTok{1996}\NormalTok{, }\DecValTok{2018}\NormalTok{, }\AttributeTok{by =} \DecValTok{2}\NormalTok{))}
+\NormalTok{hours\_all }\OtherTok{\textless{}{-}} \FunctionTok{map\_dfr}\NormalTok{(years, get\_hour)}
 \end{Highlighting}
@@
 \NormalTok{  new\_data\_qnames }\SpecialCharTok{\%\textgreater{}\%}
-\NormalTok{    dplyr}\SpecialCharTok{::}\FunctionTok{select}\NormalTok{(CASEID\_1979,}
+    \FunctionTok{select}\NormalTok{(CASEID\_1979,}
                   \FunctionTok{starts\_with}\NormalTok{(}\StringTok{"HRP"}\NormalTok{) }\SpecialCharTok{\&}
@@
 \NormalTok{\}}
-\NormalTok{rates }\OtherTok{\textless{}{-}} \FunctionTok{list}\NormalTok{()}
-\ControlFlowTok{for}\NormalTok{(ayear }\ControlFlowTok{in} \FunctionTok{c}\NormalTok{(}\DecValTok{1979}\SpecialCharTok{:}\DecValTok{1994}\NormalTok{, }\DecValTok{1996}\NormalTok{, }\DecValTok{1998}\NormalTok{, }\DecValTok{2000}\NormalTok{, }\DecValTok{2002}\NormalTok{, }\DecValTok{2004}\NormalTok{, }\DecValTok{2006}\NormalTok{, }\DecValTok{2008}\NormalTok{, }\DecValTok{2010}\NormalTok{,}
-               \DecValTok{2012}\NormalTok{, }\DecValTok{2014}\NormalTok{, }\DecValTok{2016}\NormalTok{, }\DecValTok{2018}\NormalTok{)) \{}
-\NormalTok{  rates[[ayear]] }\OtherTok{\textless{}{-}} \FunctionTok{get\_rate}\NormalTok{(ayear)}
-\NormalTok{\}}
-\NormalTok{rates\_all }\OtherTok{\textless{}{-}} \FunctionTok{bind\_rows}\NormalTok{(}\SpecialCharTok{!!!}\NormalTok{rates)}
+\NormalTok{rates\_all }\OtherTok{\textless{}{-}} \FunctionTok{map\_dfr}\NormalTok{(years, get\_rate)}
 \CommentTok{\# join hours and rates variable}
@@
 \centering
-\begin{tabular}[t]{r|r}
-\hline
+\begin{tabular}[t]{rr}
+\toprule
 Age & Number of Sample\\
-\hline
+\midrule
 15 & 1265\\
-\hline
 16 & 1550\\
-\hline
 17 & 1600\\
-\hline
 18 & 1530\\
-\hline
 19 & 1662\\
-\hline
 20 & 1722\\
-\hline
 21 & 1677\\
-\hline
 22 & 1680\\
-\hline
+\bottomrule
 \end{tabular}
@@
 \centering
-\begin{tabular}[t]{l|l|l|l|l}
-\hline
-\multicolumn{1}{c|}{ } & \multicolumn{3}{c|}{Race} & \multicolumn{1}{c}{ } \\
-\cline{2-4}
+\begin{tabular}[t]{lllll}
+\toprule
+\multicolumn{1}{c}{ } & \multicolumn{3}{c}{Race} & \multicolumn{1}{c}{ } \\
+\cmidrule(l{3pt}r{3pt}){2-4}
 Gender & Hispanic & Black & Non-Black, Non-Hispanic & Total\\
-\hline
+\midrule
 Male & 1000 (15.62\%) & 1613 (25.19\%) & 3790 (59.19\%) & 6403 (100.00\%)\\
-\hline
 Female & 1002 (15.95\%) & 1561 (24.84\%) & 3720 (59.21\%) & 6283 (100.00\%)\\
-\hline
 Total & 2002 (15.78\%) & 3174 (25.02\%) & 7510 (59.20\%) & 12686 (100.00\%)\\
-\hline
+\bottomrule
 \end{tabular}
@@
 \centering
-\begin{tabular}[t]{l|r}
-\hline
+\begin{tabular}[t]{lr}
+\toprule
 Statistics & Value\\
-\hline
+\midrule
 Min. & 0.01000\\
-\hline
 1st Qu. & 4.50000\\
-\hline
 Median & 7.20000\\
-\hline
 Mean & 11.86593\\
-\hline
 3rd Qu. & 11.74000\\
-\hline
 Max. & 33400.00000\\
-\hline
+\bottomrule
 \end{tabular}

### 2021-09-21 paper/scripts/ida.R

@@ -22,7 +22,7 @@ gender_race_table <- full_demographics %>%
 kable(gender_race_table,
       caption = "Gender and Race Distribution of the NLSY79 Samples",
       col.names = c("Gender", "Hispanic", "Black", "Non-Black, Non-Hispanic", "Total"),
-      bookdtabs = TRUE,
+      booktabs = TRUE,
       linesep = "") %>%
   kable_styling() %>%
   add_header_above(c(" " = 1, "Race" = 3, " " = 1))

### 2021-09-21 paper/scripts/eda.R
takeaways from eda.

@@ -16,6 +16,12 @@ sw_wages <- sw %>%
 
 ## ---- do_refreshed
 
+# singer-willet did not stated how they calculate experience
+# in database, there is work experience topic
+# (https://www.nlsinfo.org/content/cohorts/nlsy79/topical-guide/employment/work-experience)
+# but there is no variable explicitly mention about experience.
+# in this case, we approximate it with year-1979
+
 do <- yowie::wages_hs_do %>%
   mutate(exp = year-1979,
          lnwage = log(wage))
@@ -35,7 +41,7 @@ do_ref <- do %>%
 sw_wages + do_ref
 
 # the two plots are not comparable
-# since the id who are in Singer-Willet and the id in do-refreshed migth be different
+# since the id who are in Singer-Willet and the id in do-refreshed might be different
 
 
 ## --- compare-sw-do
@@ -43,13 +49,13 @@ sw_wages + do_ref
 sw_id <- as_tibble(sw) %>%
   group_by(id) %>%
   count() %>%
-  select(id)
+  dplyr::select(id)
 
 do_id <- as_tibble(do) %>%
   group_by(id) %>%
   mutate(id = as.numeric(id)) %>%
   count() %>%
-  select(id)
+  dplyr::select(id)
 
 # check common id in data set
 sw_do <- inner_join(sw_id, do_id, by = "id")
@@ -92,9 +98,9 @@ sw_wages_agree + do_ref_agree
 
 ## ---- compare-period
 
+
 # filter the xp in refreshed data in order to get the same experience with Singer-Willet data,
 # which is 13 years of experience
-
 do_agree_cutxp <- do_agree %>%
   filter(exp <= 13)
 
@@ -108,7 +114,7 @@ summary(do_agree_cutxp$lnwage)
 
 # plotting
 
-sw_wages_agree_exp <- sw_agree %>%
+sw_wages_agree_exp <- sw_agree_xpedit %>%
   ggplot(aes(x = xp,
              y = ln_wages)) +
   geom_line(aes(group = id),alpha = 0.1) +
@@ -138,13 +144,27 @@ sw_wages_agree_exp + do_ref_agree_cutxp
 
 ## ---- compare-id
 
-# Takes some id and compare them
-
-
 
 # Takeaways:
 
+# we don't know how exactly the criteria of drop out,
+# resulted view agreements of id being dropped out in the two dataset.
 
+# with the same id, the ln(wages) didn't agree each other, probably due to:
+
+# we don't have experience variable in the original data base.
[…]

### 2021-09-21 paper/scripts/eda.R
Add takeaways from eda.

@@ -49,13 +49,13 @@ sw_wages + do_ref
 sw_id <- as_tibble(sw) %>%
   group_by(id) %>%
   count() %>%
-  dplyr::select(id)
+  select(id)
 
 do_id <- as_tibble(do) %>%
   group_by(id) %>%
   mutate(id = as.numeric(id)) %>%
   count() %>%
-  dplyr::select(id)
+  select(id)
 
 # check common id in data set
 sw_do <- inner_join(sw_id, do_id, by = "id")

### 2021-09-21 paper/paper.Rmd
Add recap.

@@
 Finally, we save the imputed data and set the appropriate data type for the variables.
-As our target data is the mean hourly wage of the high-school dropouts, we then subset the high-school graduate data set to have only the male-high school dropout data.
-We then make these three data sets and their processing documentation publicly available through an R data container package called `yowie`.
-The complete flow from the raw data to these data set is displayed in Figure \ref{fig:flow-chart}.
 
@@
 ## Recap
+
+In this section, we have demonstrate a workflow from getting the data from the NSLY79 database until a ready-analysed data.
+Referring to @LooMarkvander2018Sdcw, this section shows  part of statistical value chain, which is data cleaning from raw, input, and valid data.
+The raw demographic and employment data from the database are not in a tidy format.
+Hence, we perform data wrangling to get the input data.
+We inspect the input data to find eligible observations to the target data.
+One of the target data is wages data of respondents who completed up to 12th grade and participated at least in 5 rounds of survey.
+This needs us to join the extracted demographic and employment data.
+
+Further, using IDA, we inspect on the existence of anomalies/extreme values in the data using the weight from robust linear model. We then replace these extreme values using their predicted value. 
+From this stage forward, we assume that we already have valid data.
+We also subset the data to get the wages of high school dropouts.
+We then make these three data sets and the cleaning workflow documentation publicly available through an R data container package called `yowie`.
+A visualisation of these process is displayed in  Figure \ref{fig:flow-chart}.
 

### 2021-09-21 paper/paper-jsdse/paper-jsdse.tex
fixed merge conflict

@@
 
-This paper aims to demonstrate the steps of IDA, tidying, cleaning, and documenting the process, for a prominent open data source, National Longitudinal Survey of Youth (NLSY) 1979, henceforth referred to NLSY79.
+This paper demonstrates the steps of IDA -- tidying, cleaning, and documenting the process -- for a prominent open data source, National Longitudinal Survey of Youth (NLSY) 1979, henceforth referred to NLSY79.
 This data has been playing an important role in research in various disciplines, including but are not limited to economics, sociology, education, public policy, and public health for more than a quarter of the century (Pergamit et al.
@@
 Singer and Willett (2003) used the wages and other variables of high school dropouts from the NLSY79 data as an example data set to illustrate longitudinal data modeling.
-Our aim is to refresh this textbook data to be more up to date.
+Our aim is to refresh this textbook data with data from .
 Here, we investigate the process of getting from the raw NLSY79 to this textbook data set.
@@
 
-\caption{Re-make of the summary plots of the fully processed data suggest that the data is now in a reasonable state: longitudinal profiles of wages for all individuals 1979-2018 (A), boxplots of minimum, median, and maximum wages of each individual (B), and one individual with an unusual wage relative to their years of data (C).
+\caption{Re-make of the summary plots of the fully processed data suggest that it is now in a reasonable state: longitudinal profiles of wages for all individuals 1979-2018 (A), boxplots of minimum, median, and maximum wages of each individual (B), and one individual with an unusual wage relative to their years of data (C).
 }\label{fig:fixed-feature-plot}
@@
 Finally, we save the imputed data and set the appropriate data type for the variables.
-As our target data is the mean hourly wage of the high-school dropouts, we then subset the high-school graduate data set to have only the male-high school dropout data.
-We then make these three data sets and their processing documentation publicly available through an R data container package called \texttt{yowie}.
-The complete flow from the raw data to these data set is displayed in Figure \ref{fig:flow-chart}.
 
@@
 
+In this section, we have demonstrate a workflow from getting the data from the NSLY79 database until a ready-analysed data.
+Referring to Loo (2018), this section shows part of statistical value chain, which is data cleaning from raw, input, and valid data.
+The raw demographic and employment data from the database are not in a tidy format.
+Hence, we perform data wrangling to get the input data.
+We inspect the input data to find eligible observations to the target data.
+One of the target data is wages data of respondents who completed up to 12th grade and participated at least in 5 rounds of survey.
+This needs us to join the extracted demographic and employment data.
+
+Further, using IDA, we inspect on the existence of anomalies/extreme values in the data using the weight from robust linear model. We then replace these extreme values using their predicted value.
+From this stage forward, we assume that we already have valid data.
+We also subset the data to get the wages of high school dropouts.
+We then make these three data sets and the cleaning workflow documentation publicly available through an R data container package called \texttt{yowie}.
+A visualisation of these process is displayed in Figure \ref{fig:flow-chart}.
+
 \begin{figure}
 
-{\centering \includegraphics[width=0.93\linewidth]{/Users/etan0038/Dropbox/repo/summer-wages-refresh/paper/figures/flowchart}
+{\centering \includegraphics[width=0.93\linewidth]{/Users/cookd/numbats/summer-wages-refresh/paper/figures/flowchart}
 
@@
 
+\leavevmode\hypertarget{ref-LooMarkvander2018Sdcw}{}%
+Loo, Mark van der. 2018. \emph{Statistical Data Cleaning with Applications in r}.
+
 \leavevmode\hypertarget{ref-rsvg}{}%

### 2021-09-21 paper/paper.Rmd
fixed merge conflict

@@
 
-This paper aims to demonstrate the steps of IDA, tidying, cleaning, and documenting the process, for a prominent open data source, National Longitudinal Survey of Youth (NLSY) 1979, henceforth referred to NLSY79.
+This paper demonstrates the steps of IDA -- tidying, cleaning, and documenting the process -- for a prominent open data source, National Longitudinal Survey of Youth (NLSY) 1979, henceforth referred to NLSY79.
 This data has been playing an important role in research in various disciplines, including but are not limited to economics, sociology, education, public policy, and public health for more than a quarter of the century [@MichaelRPergamit2001DWTN].
@@
 @SingerJudithD2003Alda used the wages and other variables of high school dropouts from the NLSY79 data as an example data set to illustrate longitudinal data modeling.
-Our aim is to refresh this textbook data to be more up to date.
+Our aim is to refresh this textbook data with data from .
 Here, we investigate the process of getting from the raw NLSY79 to this textbook data set.
@@
 
-```{r fixed-feature-plot, fig.cap = "Re-make of the summary plots of the fully processed data suggest that the data is now in a reasonable state: longitudinal profiles of wages for all individuals 1979-2018 (A), boxplots of minimum, median, and maximum wages of each individual (B), and one individual with an unusual wage relative to their years of data (C).
+```{r fixed-feature-plot, fig.cap = "Re-make of the summary plots of the fully processed data suggest that it is now in a reasonable state: longitudinal profiles of wages for all individuals 1979-2018 (A), boxplots of minimum, median, and maximum wages of each individual (B), and one individual with an unusual wage relative to their years of data (C).
 ", fig.width=6, fig.height=2}

### 2021-09-21 paper/paper.tex
fixed merge conflict

@@
 % For Pandoc 2.11+
-\newenvironment{CSLReferences}[3] % #1 hanging-ident, #2 entry spacing
+\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
  {% don't indent paragraphs
@@
 \newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
-\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}}
+\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
 \newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

### 2021-09-21 paper/paper.Rmd
finished with intro

@@
 However, open data is often referred to as what we might call "wild data" because it requires substantial cleaning and tidying to tame it into textbook shape.
-@HuebnerMariannePhD2016Asat emphasize that making the data cleaning process accountable and transparent is imperative.
-Documenting the data cleaning is essential for the integrity of downstream statistical analyses and model building [@HuebnerMarianne2020Haar].
+@HuebnerMariannePhD2016Asat emphasize that making the data cleaning process accountable and transparent is imperative and essential for the integrity of downstream statistical analyses and model building [@HuebnerMarianne2020Haar].
 Data cleaning can be considered to be a part of what is called "initial data analysis (IDA)" [@Chatfield1985TIEo].
 In IDA one would also explore the data, especially to check if the data is consistent with assumptions required for modeling.
-This is also related to exploratory data analysis, which @tukey defines as detective work to learn from data, either numerically or graphically, before confirmatory data analysis is performed, which encompasses IDA.
+This is also related to exploratory data analysis (EDA), coined by @tukey with a focus on learning from data.
+EDA can be considered to encompass IDA.
 @DasuTamraparni2003Edma say that data cleaning and exploration, without naming it as IDA, is a difficult task and consumes 80% of the data mining task.
 
-Despite its importance, this IDA stage is often undervalued and neglected [@Chatfield1985TIEo].
+Despite it's importance, this IDA stage is often undervalued and neglected [@Chatfield1985TIEo].
 There are few research papers that document the data cleaning [@WickhamHadley2014TD].
@@
 
-This paper demonstrates the steps of IDA -- tidying, cleaning, and documenting the process -- for a prominent open data source, National Longitudinal Survey of Youth (NLSY) 1979, henceforth referred to NLSY79.
+This paper demonstrates the steps of IDA, and documents the process, for a prominent open data source, National Longitudinal Survey of Youth 1979, henceforth referred to NLSY79.
 This data has been playing an important role in research in various disciplines, including but are not limited to economics, sociology, education, public policy, and public health for more than a quarter of the century [@MichaelRPergamit2001DWTN].
-In addition, The NLSY is considered a survey with high retention rates and carefully designed, making it suitable for a life course research [@MichaelRPergamit2001DWTN; @eliznlsy].
-According to @eliznlsy, thousand of articles, and hundreds of book chapters and monographs has incorporated the NLSY data.
-Moreover, the NLSY79 is considered the most widely used and most important cohort in the NLSY79 data sets [@MichaelRPergamit2001DWTN].
+In addition, this is considered a carefully designed longitudinal survey with high retention rates, making it suitable for life course research [@MichaelRPergamit2001DWTN; @eliznlsy].
+According to @eliznlsy, thousands of articles, and hundreds of book chapters and monographs have utilized this data.
+Moreover, the NLSY79 is considered the most widely used and most important cohort in the survey data [@MichaelRPergamit2001DWTN].
 
 @SingerJudithD2003Alda used the wages and other variables of high school dropouts from the NLSY79 data as an example data set to illustrate longitudinal data modeling.
-Our aim is to refresh this textbook data with data from .
-Here, we investigate the process of getting from the raw NLSY79 to this textbook data set.
-However, we cannot create the exact same data set as published in their book since we do not have the information of what age threshold they used to determine the high school dropouts.
+Our aim is to refresh this textbook data with data from 1994 through to the latest data reported in 2018.
+Here, we investigate the process of getting from the raw NLSY79 data to a textbook data set as similar as possible to that provided by @SingerJudithD2003Alda.
+<!--However, we cannot create the exact same data set as published in their book since we do not have the information of what age threshold they used to determine the high school dropouts.-->
 
-This paper is structured to have 5 sections.
-Section \@ref(database) describes the original data source.
-Section \@ref(cleaning) presents the steps of cleaning the data, including getting and tidying the data from the NLSY79 and initial data analysis to find and treat the anomalies in the data.
-We also gave examples of exploratory data analysis using the clean data in Section \@ref(eda).
-Finally, Section \@ref(summary) summarises the paper.
+Our process of cleaning builds heavily on the tidyverse approach [@tidyverse].
+The data is first organised into "tidy data" [@WickhamHadley2014TD] and then further wrangled using the data pipeline and split-apply-combine approach [@plyr].
+The resulting data is provided in a new R package called `yowie` which includes the code so that the process is reproducible, and could be used to further refresh the data as new records are made available in the NLSY79 database.
+
+This paper is structured in the following way.
+Section \@ref(database) describes the NLSY79 data source.
+Section \@ref(cleaning) presents the steps of cleaning the data, including getting and tidying the data from the NLSY79 and IDA to find and repair anomalies.
+Our final subset is compared the old textbook subset in Section \@ref(comare).
+Finally, Section \@ref(summary) summarizes the contribution and makes recommendations for the NLSY79 data curators.
 

### 2021-09-21 paper/paper.tex
finished with intro

@@
 However, open data is often referred to as what we might call ``wild data'' because it requires substantial cleaning and tidying to tame it into textbook shape.
-Huebner Marianne, Vach, and Cessie (2016) emphasize that making the data cleaning process accountable and transparent is imperative.
-Documenting the data cleaning is essential for the integrity of downstream statistical analyses and model building (M.
+Huebner Marianne, Vach, and Cessie (2016) emphasize that making the data cleaning process accountable and transparent is imperative and essential for the integrity of downstream statistical analyses and model building (M.
 Huebner et al.
@@
 In IDA one would also explore the data, especially to check if the data is consistent with assumptions required for modeling.
-This is also related to exploratory data analysis, which Tukey (1977) defines as detective work to learn from data, either numerically or graphically, before confirmatory data analysis is performed, which encompasses IDA.
+This is also related to exploratory data analysis (EDA), coined by Tukey (1977) with a focus on learning from data.
+EDA can be considered to encompass IDA.
 Dasu and Johnson (2003) say that data cleaning and exploration, without naming it as IDA, is a difficult task and consumes 80\% of the data mining task.
 
-Despite its importance, this IDA stage is often undervalued and neglected (Chatfield 1985).
+Despite it's importance, this IDA stage is often undervalued and neglected (Chatfield 1985).
 There are few research papers that document the data cleaning (Wickham 2014).
@@
 
-This paper demonstrates the steps of IDA -- tidying, cleaning, and documenting the process -- for a prominent open data source, National Longitudinal Survey of Youth (NLSY) 1979, henceforth referred to NLSY79.
+This paper demonstrates the steps of IDA, and documents the process, for a prominent open data source, National Longitudinal Survey of Youth 1979, henceforth referred to NLSY79.
 This data has been playing an important role in research in various disciplines, including but are not limited to economics, sociology, education, public policy, and public health for more than a quarter of the century (Pergamit et al.
 2001).
-In addition, The NLSY is considered a survey with high retention rates and carefully designed, making it suitable for a life course research (Pergamit et al.
+In addition, this is considered a carefully designed longitudinal survey with high retention rates, making it suitable for life course research (Pergamit et al.
 2001; Cooksey 2017).
-According to Cooksey (2017), thousand of articles, and hundreds of book chapters and monographs has incorporated the NLSY data.
-Moreover, the NLSY79 is considered the most widely used and most important cohort in the NLSY79 data sets (Pergamit et al.
+According to Cooksey (2017), thousands of articles, and hundreds of book chapters and monographs have utilized this data.
+Moreover, the NLSY79 is considered the most widely used and most important cohort in the survey data (Pergamit et al.
 2001).
@@
 Singer and Willett (2003) used the wages and other variables of high school dropouts from the NLSY79 data as an example data set to illustrate longitudinal data modeling.
-Our aim is to refresh this textbook data with data from .
-Here, we investigate the process of getting from the raw NLSY79 to this textbook data set.
-However, we cannot create the exact same data set as published in their book since we do not have the information of what age threshold they used to determine the high school dropouts.
+Our aim is to refresh this textbook data with data from 1994 through to the latest data reported in 2018.
+Here, we investigate the process of getting from the raw NLSY79 data to a textbook data set as similar as possible to that provided by Singer and Willett (2003).
 
-This paper is structured to have 5 sections.
-Section \ref{database} describes the original data source.
-Section \ref{cleaning} presents the steps of cleaning the data, including getting and tidying the data from the NLSY79 and initial data analysis to find and treat the anomalies in the data.
-We also gave examples of exploratory data analysis using the clean data in Section \ref{eda}.
-Finally, Section \ref{summary} summarises the paper.
+Our process of cleaning builds heavily on the tidyverse approach (Wickham et al.
+2019).
+The data is first organised into ``tidy data'' (Wickham 2014) and then further wrangled using the data pipeline and split-apply-combine approach (Wickham 2011).
+The resulting data is provided in a new R package called \texttt{yowie} which includes the code so that the process is reproducible, and could be used to further refresh the data as new records are made available in the NLSY79 database.
+
+This paper is structured in the following way.
+Section \ref{database} describes the NLSY79 data source.
+Section \ref{cleaning} presents the steps of cleaning the data, including getting and tidying the data from the NLSY79 and IDA to find and repair anomalies.
+Our final subset is compared the old textbook subset in Section \ref{comare}.
+Finally, Section \ref{summary} summarises the contribution and makes recommendations for the NLSY79 data curators.
 
@@
 
+\leavevmode\hypertarget{ref-plyr}{}%
+Wickham, Hadley. 2011. {``The Split-Apply-Combine Strategy for Data Analysis.''} \emph{Journal of Statistical Software, Articles} 40 (1): 1--29. \url{https://doi.org/10.18637/jss.v040.i01}.
+
 \leavevmode\hypertarget{ref-WickhamHadley2014TD}{}%
-Wickham, Hadley. 2014. {``Tidy Data.''} \emph{Journal of Statistical Software} 59 (10): 1--23.
+---------. 2014. {``Tidy Data.''} \emph{Journal of Statistical Software} 59 (10): 1--23.
 

### 2021-09-22 paper/paper-jsdse/paper-jsdse.tex
Add reference and modify comparison plot.

--- 2021-09-20 paper/paper-jsdse/paper-jsdse.tex
@@
 However, open data is often referred to as what we might call ``wild data'' because it requires substantial cleaning and tidying to tame it into textbook shape.
-Huebner, Vach, and Cessie (2016) emphasize that making the data cleaning process accountable and transparent is imperative.
-Documenting the data cleaning is essential for the integrity of downstream statistical analyses and model building (Huebner et al.
+Huebner, Vach, and Cessie (2016) emphasize that making the data cleaning process accountable and transparent is imperative and essential for the integrity of downstream statistical analyses and model building (Huebner et al.
 2020).
@@
 
-This paper aims to demonstrate the steps of IDA, tidying, cleaning, and documenting the process, for a prominent open data source, National Longitudinal Survey of Youth (NLSY) 1979, henceforth referred to NLSY79.
+This paper demonstrates the steps of IDA, and documents the process, for a prominent open data source, National Longitudinal Survey of Youth 1979, henceforth referred to NLSY79.
 This data has been playing an important role in research in various disciplines, including but are not limited to economics, sociology, education, public policy, and public health for more than a quarter of the century (Pergamit et al.
 2001).
-In addition, The NLSY is considered a survey with high retention rates and carefully designed, making it suitable for a life course research (Pergamit et al.
+In addition, this is considered a carefully designed longitudinal survey with high retention rates, making it suitable for life course research (Pergamit et al.
 2001; Cooksey 2017).
-According to Cooksey (2017), thousand of articles, and hundreds of book chapters and monographs has incorporated the NLSY data.
-Moreover, the NLSY79 is considered the most widely used and most important cohort in the NLSY79 data sets (Pergamit et al.
+According to Cooksey (2017), thousands of articles, and hundreds of book chapters and monographs have utilized this data.
+Moreover, the NLSY79 is considered the most widely used and most important cohort in the survey data (Pergamit et al.
 2001).
@@
 Singer and Willett (2003) used the wages and other variables of high school dropouts from the NLSY79 data as an example data set to illustrate longitudinal data modeling.
-Our aim is to refresh this textbook data to be more up to date.
-Here, we investigate the process of getting from the raw NLSY79 to this textbook data set.
-However, we cannot create the exact same data set as published in their book since we do not have the information of what age threshold they used to determine the high school dropouts.
+Our aim is to refresh this textbook data with data from 1994 through to the latest data reported in 2018.
+Here, we investigate the process of getting from the raw NLSY79 data to a textbook data set as similar as possible to that provided by Singer and Willett (2003).
 
-This paper is structured to have 5 sections.
-Section \ref{database} describes the original data source.
-Section \ref{cleaning} presents the steps of cleaning the data, including getting and tidying the data from the NLSY79 and initial data analysis to find and treat the anomalies in the data.
-We also gave examples of exploratory data analysis using the clean data in Section \ref{eda}.
-Finally, Section \ref{summary} summarises the paper.
+Our process of cleaning builds heavily on the tidyverse approach (H.
+Wickham, Averick, et al.
+2019).
+The data is first organised into ``tidy data'' (Wickham 2014) and then further wrangled using the data pipeline and split-apply-combine approach (Wickham 2011).
+The resulting data is provided in a new R package called \texttt{yowie} which includes the code so that the process is reproducible, and could be used to further refresh the data as new records are made available in the NLSY79 database.
+
+This paper is structured in the following way.
+Section \ref{database} describes the NLSY79 data source.
+Section \ref{cleaning} presents the steps of cleaning the data, including getting and tidying the data from the NLSY79 and IDA to find and repair anomalies.
+Our final subset is compared the old textbook subset in Section \ref{comare}.
+Finally, Section \ref{summary} summarizes the contribution and makes recommendations for the NLSY79 data curators.
 
@@
 Bureau of Labor Statistics, n.d.).
-Specifically, we only extract a subset of the variables in the database as shown in \ref{box:data}.
+Specifically, we only extract a subset of the variables in the database as described in Section\textasciitilde{}\ref{getdata}.
 
-\hypertarget{getting-and-tidying-the-data}{%
-\subsection{Getting and Tidying the Data}\label{getting-and-tidying-the-data}}
+\hypertarget{getdata}{%
+\subsection{Getting and Tidying the Data}\label{getdata}}
 
@@
 \NormalTok{dob_tidy <-}\StringTok{ }\NormalTok{new_data_qnames }\OperatorTok{%>%}
-\StringTok{  }\NormalTok{dplyr}\OperatorTok{::}\KeywordTok{select}\NormalTok{(CASEID_}\DecValTok{1979}\NormalTok{,}
+\StringTok{  }\KeywordTok{select}\NormalTok{(CASEID_}\DecValTok{1979}\NormalTok{,}
                 \KeywordTok{starts_with}\NormalTok{(}\StringTok{"Q1-3_A~"}\NormalTok{)) }\OperatorTok{%>%}
@@
                       \KeywordTok{is.na}\NormalTok{(}\StringTok{`}\DataTypeTok{Q1-3_A~M_1981}\StringTok{`}\NormalTok{) }\OperatorTok{|}\StringTok{ }\KeywordTok{is.na}\NormalTok{(}\StringTok{`}\DataTypeTok{Q1-3_A~Y_1981}\StringTok{`}\NormalTok{) }\OperatorTok{~}\StringTok{ }\OtherTok{FALSE}\NormalTok{)) }\OperatorTok{%>%}
-\StringTok{  }\NormalTok{dplyr}\OperatorTok{::}\KeywordTok{select}\NormalTok{(CASEID_}\DecValTok{1979}\NormalTok{,}
+\StringTok{  }\KeywordTok{select}\NormalTok{(CASEID_}\DecValTok{1979}\NormalTok{,}
 \NormalTok{         dob_month,}
@@
 \NormalTok{    temp <-}\StringTok{ }\NormalTok{new_data_qnames }\OperatorTok{%>%}
-\StringTok{      }\NormalTok{dplyr}\OperatorTok{::}\KeywordTok{select}\NormalTok{(CASEID_}\DecValTok{1979}\NormalTok{,}
+\StringTok{      }\KeywordTok{select}\NormalTok{(CASEID_}\DecValTok{1979}\NormalTok{,}
                     \KeywordTok{starts_with}\NormalTok{(}\StringTok{"QES-52A"}\NormalTok{) }\OperatorTok{&}
@@
 \NormalTok{    temp <-}\StringTok{ }\NormalTok{new_data_qnames }\OperatorTok{%>%}
-\StringTok{      }\NormalTok{dplyr}\OperatorTok{::}\KeywordTok{select}\NormalTok{(CASEID_}\DecValTok{1979}\NormalTok{,}
+\StringTok{      }\KeywordTok{select}\NormalTok{(CASEID_}\DecValTok{1979}\NormalTok{,}
                     \KeywordTok{starts_with}\NormalTok{(}\StringTok{"QES-52D"}\NormalTok{) }\OperatorTok{&}
@@
 \NormalTok{\}}
-\CommentTok{# list to save the iteration result}
-\NormalTok{hours <-}\StringTok{ }\KeywordTok{list}\NormalTok{()}
+
 \CommentTok{# getting the hours of work of all observations}
-\ControlFlowTok{for}\NormalTok{(ayear }\ControlFlowTok{in} \KeywordTok{c}\NormalTok{(}\DecValTok{1979}\OperatorTok{:}\DecValTok{1994}\NormalTok{, }\DecValTok{1996}\NormalTok{, }\DecValTok{1998}\NormalTok{, }\DecValTok{2000}\NormalTok{, }\DecValTok{2002}\NormalTok{, }\DecValTok{2004}\NormalTok{, }\DecValTok{2006}\NormalTok{, }\DecValTok{2008}\NormalTok{, }\DecValTok{2010}\NormalTok{,}
-               \DecValTok{2012}\NormalTok{, }\DecValTok{2014}\NormalTok{, }\DecValTok{2016}\NormalTok{, }\DecValTok{2018}\NormalTok{)) \{}
-\NormalTok{  hours[[ayear]] <-}\StringTok{ }\KeywordTok{get_hour}\NormalTok{(ayear)}
-\NormalTok{\}}
-\CommentTok{# unlist the hours of work}
-\NormalTok{hours_all <-}\StringTok{ }\KeywordTok{bind_rows}\NormalTok{(}\OperatorTok{!!!}\NormalTok{hours)}
+\NormalTok{years <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\DecValTok{1979}\OperatorTok{:}\DecValTok{1994}\NormalTok{, }\KeywordTok{seq}\NormalTok{(}\DecValTok{1996}\NormalTok{, }\DecValTok{2018}\NormalTok{, }\DataTypeTok{by =} \DecValTok{2}\NormalTok{))}
+\NormalTok{hours_all <-}\StringTok{ }\KeywordTok{map_dfr}\NormalTok{(years, get_hour)}
 \end{Highlighting}
@@
 \NormalTok{  new_data_qnames }\OperatorTok{%>%}
-\StringTok{    }\NormalTok{dplyr}\OperatorTok{::}\KeywordTok{select}\NormalTok{(CASEID_}\DecValTok{1979}\NormalTok{,}
+\StringTok{    }\KeywordTok{select}\NormalTok{(CASEID_}\DecValTok{1979}\NormalTok{,}
                   \KeywordTok{starts_with}\NormalTok{(}\StringTok{"HRP"}\NormalTok{) }\OperatorTok{&}
@@
 \NormalTok{\}}
-\NormalTok{rates <-}\StringTok{ }\KeywordTok{list}\NormalTok{()}
-\ControlFlowTok{for}\NormalTok{(ayear }\ControlFlowTok{in} \KeywordTok{c}\NormalTok{(}\DecValTok{1979}\OperatorTok{:}\DecValTok{1994}\NormalTok{, }\DecValTok{1996}\NormalTok{, }\DecValTok{1998}\NormalTok{, }\DecValTok{2000}\NormalTok{, }\DecValTok{2002}\NormalTok{, }\DecValTok{2004}\NormalTok{, }\DecValTok{2006}\NormalTok{, }\DecValTok{2008}\NormalTok{, }\DecValTok{2010}\NormalTok{,}
-               \DecValTok{2012}\NormalTok{, }\DecValTok{2014}\NormalTok{, }\DecValTok{2016}\NormalTok{, }\DecValTok{2018}\NormalTok{)) \{}
-\NormalTok{  rates[[ayear]] <-}\StringTok{ }\KeywordTok{get_rate}\NormalTok{(ayear)}
-\NormalTok{\}}
-\NormalTok{rates_all <-}\StringTok{ }\KeywordTok{bind_rows}\NormalTok{(}\OperatorTok{!!!}\NormalTok{rates)}
+\NormalTok{rates_all <-}\StringTok{ }\KeywordTok{map_dfr}\NormalTok{(years, get_rate)}
 \CommentTok{# join hours and rates variable}
@@
 Finally, we save the imputed data and set the appropriate data type for the variables.
-As our target data is the mean hourly wage of the high-school dropouts, we then subset the high-school graduate data set to have only the male-high school dropout data.
-We then make these three data sets and their processing documentation publicly available through an R data container package called \texttt{yowie}.
-The complete flow from the raw data to these data set is displayed in Figure \ref{fig:flow-chart}.
+
+\hypertarget{recap}{%
+\subsection{Recap}\label{recap}}
+
+In this section, we have demonstrate a workflow from getting the data from the NSLY79 database until a ready-analysed data.
+Referring to Loo (2018), this section shows part of statistical value chain, which is data cleaning from raw, input, and valid data.
+The raw demographic and employment data from the database are not in a tidy format.
+Hence, we perform data wrangling to get the input data.
+We inspect the input data to find eligible observations to the target data.
+One of the target data is wages data of respondents who completed up to 12th grade and participated at least in 5 rounds of survey.
+This needs us to join the extracted demographic and employment data.
+
+Further, using IDA, we inspect on the existence of anomalies/extreme values in the data using the weight from robust linear model. We then replace these extreme values using their predicted value.
+From this stage forward, we assume that we already have valid data.
+We also subset the data to get the wages of high school dropouts.
+We then make these three data sets and the cleaning workflow documentation publicly available through an R data container package called \texttt{yowie}.
+A visualisation of these process is displayed in Figure \ref{fig:flow-chart}.
 
@@
 
-\leavevmode\hypertarget{ref-OECD}{}%
-OECD. n.d. ``Gender Wage Gap.'' https://data.oecd.org/earnwage/gender-wage-gap.htm.
+\leavevmode\hypertarget{ref-LooMarkvander2018Sdcw}{}%
+Loo, Mark van der. 2018. \emph{Statistical Data Cleaning with Applications in R}.
 
@@
 
+\leavevmode\hypertarget{ref-plyr}{}%
+Wickham, Hadley. 2011. ``The Split-Apply-Combine Strategy for Data Analysis.'' \emph{Journal of Statistical Software, Articles} 40 (1): 1--29. \url{https://doi.org/10.18637/jss.v040.i01}.
+
 \leavevmode\hypertarget{ref-WickhamHadley2014TD}{}%
-Wickham, Hadley. 2014. ``Tidy Data.'' \emph{Journal of Statistical Software} 59 (10): 1--23.
+---------. 2014. ``Tidy Data.'' \emph{Journal of Statistical Software} 59 (10): 1--23.
 

### 2021-09-22 paper/scripts/data-preprocessing.R
data processing

@@ -26,7 +26,7 @@ source(here::here("data-raw/NLSY79/NLSY79.R"))
 new_data_qnames %>%
   select(CASEID_1979,
          starts_with("HRP") &
-           ends_with(c("1979", "1980", "1981", "1982", "1983"))) %>%
+           ends_with(c("1979", "1980"))) %>%
   visdat::vis_miss()
 
 ## ---- dob-tidy

### 2021-09-22 paper/paper.tex

@@
 However, open data is often referred to as what we might call ``wild data'' because it requires substantial cleaning and tidying to tame it into textbook shape.
-Huebner Marianne, Vach, and Cessie (2016) emphasize that making the data cleaning process accountable and transparent is imperative and essential for the integrity of downstream statistical analyses and model building (M.
-Huebner et al.
+Huebner, Vach, and Cessie (2016) emphasize that making the data cleaning process accountable and transparent is imperative and essential for the integrity of downstream statistical analyses and model building (Huebner et al.
 2020).
@@
 
-Our process of cleaning builds heavily on the tidyverse approach (Wickham et al.
+Our process of cleaning builds heavily on the tidyverse approach (H.
+Wickham, Averick, et al.
 2019).
@@
 Our final subset is compared the old textbook subset in Section \ref{comare}.
-Finally, Section \ref{summary} summarises the contribution and makes recommendations for the NLSY79 data curators.
+Finally, Section \ref{summary} summarizes the contribution and makes recommendations for the NLSY79 data curators.
 
@@
 \begin{Highlighting}[]
-\CommentTok{\#tidy the date of birth data}
-\NormalTok{dob\_tidy }\OtherTok{\textless{}{-}}\NormalTok{ new\_data\_qnames }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{select}\NormalTok{(CASEID\_1979,}
-                \FunctionTok{starts\_with}\NormalTok{(}\StringTok{"Q1{-}3\_A\textasciitilde{}"}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{dob\_year =} \FunctionTok{case\_when}\NormalTok{(}
-                      \CommentTok{\# if the years recorded in both sets match, take 79 data}
-                      \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}Y\_1979}\StringTok{\textasciigrave{}} \SpecialCharTok{==} \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}Y\_1981}\StringTok{\textasciigrave{}} \SpecialCharTok{\textasciitilde{}} \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}Y\_1979}\StringTok{\textasciigrave{}}\NormalTok{,}
-                      \CommentTok{\# if the year in the 81 set is missing, take the 79 data}
-                      \FunctionTok{is.na}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}Y\_1981}\StringTok{\textasciigrave{}}\NormalTok{) }\SpecialCharTok{\textasciitilde{}} \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}Y\_1979}\StringTok{\textasciigrave{}}\NormalTok{,}
-                      \CommentTok{\# if the sets don\textquotesingle{}t match for dob year, take the 79 data}
-                      \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}Y\_1979}\StringTok{\textasciigrave{}} \SpecialCharTok{!=} \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}Y\_1981}\StringTok{\textasciigrave{}} \SpecialCharTok{\textasciitilde{}} \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}Y\_1979}\StringTok{\textasciigrave{}}\NormalTok{),}
-         \AttributeTok{dob\_month =} \FunctionTok{case\_when}\NormalTok{(}
-                      \CommentTok{\# if months recorded in both sets match, take 79 data}
-                      \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}M\_1979}\StringTok{\textasciigrave{}} \SpecialCharTok{==} \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}M\_1981}\StringTok{\textasciigrave{}} \SpecialCharTok{\textasciitilde{}} \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}M\_1979}\StringTok{\textasciigrave{}}\NormalTok{,}
-                      \CommentTok{\# if month in 81 set is missing, take the 79 data}
-                      \FunctionTok{is.na}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}M\_1981}\StringTok{\textasciigrave{}}\NormalTok{) }\SpecialCharTok{\textasciitilde{}} \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}M\_1979}\StringTok{\textasciigrave{}}\NormalTok{,}
-                      \CommentTok{\# if sets don\textquotesingle{}t match for dob month, take the 79 data}
-                      \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}M\_1979}\StringTok{\textasciigrave{}} \SpecialCharTok{!=} \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}M\_1981}\StringTok{\textasciigrave{}} \SpecialCharTok{\textasciitilde{}} \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}M\_1979}\StringTok{\textasciigrave{}}\NormalTok{),}
-         \CommentTok{\# flag if there is a conflict between dob recorded in 79 and 81}
-         \AttributeTok{dob\_conflict =} \FunctionTok{case\_when}\NormalTok{(}
-\NormalTok{                      (}\StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}M\_1979}\StringTok{\textasciigrave{}} \SpecialCharTok{!=} \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}M\_1981}\StringTok{\textasciigrave{}}\NormalTok{) }\SpecialCharTok{\&} \SpecialCharTok{!}\FunctionTok{is.na}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}M\_1981}\StringTok{\textasciigrave{}}\NormalTok{)}
-                      \SpecialCharTok{\textasciitilde{}} \ConstantTok{TRUE}\NormalTok{,}
-\NormalTok{                      (}\StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}Y\_1979}\StringTok{\textasciigrave{}} \SpecialCharTok{!=} \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}Y\_1981}\StringTok{\textasciigrave{}}\NormalTok{) }\SpecialCharTok{\&} \SpecialCharTok{!}\FunctionTok{is.na}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}Y\_1981}\StringTok{\textasciigrave{}}\NormalTok{)}
-                      \SpecialCharTok{\textasciitilde{}} \ConstantTok{TRUE}\NormalTok{,}
-\NormalTok{                      (}\StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}Y\_1979}\StringTok{\textasciigrave{}} \SpecialCharTok{==} \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}Y\_1981}\StringTok{\textasciigrave{}}\NormalTok{) }\SpecialCharTok{\&}
-\NormalTok{                        (}\StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}M\_1979}\StringTok{\textasciigrave{}} \SpecialCharTok{==} \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}M\_1981}\StringTok{\textasciigrave{}}\NormalTok{) }\SpecialCharTok{\textasciitilde{}} \ConstantTok{FALSE}\NormalTok{,}
-                      \FunctionTok{is.na}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}M\_1981}\StringTok{\textasciigrave{}}\NormalTok{) }\SpecialCharTok{|} \FunctionTok{is.na}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}Y\_1981}\StringTok{\textasciigrave{}}\NormalTok{) }\SpecialCharTok{\textasciitilde{}} \ConstantTok{FALSE}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{select}\NormalTok{(CASEID\_1979,}
-\NormalTok{         dob\_month,}
-\NormalTok{         dob\_year,}
-\NormalTok{         dob\_conflict)}
+\CommentTok{#tidy the date of birth data}
+\NormalTok{dob_tidy <-}\StringTok{ }\NormalTok{new_data_qnames }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{select}\NormalTok{(CASEID_}\DecValTok{1979}\NormalTok{,}
+                \KeywordTok{starts_with}\NormalTok{(}\StringTok{"Q1-3_A~"}\NormalTok{)) }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{dob_year =} \KeywordTok{case_when}\NormalTok{(}
+                      \CommentTok{# if the years recorded in both sets match, take 79 data}
+                      \StringTok{`}\DataTypeTok{Q1-3_A~Y_1979}\StringTok{`} \OperatorTok{==}\StringTok{ `}\DataTypeTok{Q1-3_A~Y_1981}\StringTok{`} \OperatorTok{~}\StringTok{ `}\DataTypeTok{Q1-3_A~Y_1979}\StringTok{`}\NormalTok{,}
+                      \CommentTok{# if the year in the 81 set is missing, take the 79 data}
+                      \KeywordTok{is.na}\NormalTok{(}\StringTok{`}\DataTypeTok{Q1-3_A~Y_1981}\StringTok{`}\NormalTok{) }\OperatorTok{~}\StringTok{ `}\DataTypeTok{Q1-3_A~Y_1979}\StringTok{`}\NormalTok{,}
+                      \CommentTok{# if the sets don't match for dob year, take the 79 data}
+                      \StringTok{`}\DataTypeTok{Q1-3_A~Y_1979}\StringTok{`} \OperatorTok{!=}\StringTok{ `}\DataTypeTok{Q1-3_A~Y_1981}\StringTok{`} \OperatorTok{~}\StringTok{ `}\DataTypeTok{Q1-3_A~Y_1979}\StringTok{`}\NormalTok{),}
+         \DataTypeTok{dob_month =} \KeywordTok{case_when}\NormalTok{(}
+                      \CommentTok{# if months recorded in both sets match, take 79 data}
+                      \StringTok{`}\DataTypeTok{Q1-3_A~M_1979}\StringTok{`} \OperatorTok{==}\StringTok{ `}\DataTypeTok{Q1-3_A~M_1981}\StringTok{`} \OperatorTok{~}\StringTok{ `}\DataTypeTok{Q1-3_A~M_1979}\StringTok{`}\NormalTok{,}
+                      \CommentTok{# if month in 81 set is missing, take the 79 data}
+                      \KeywordTok{is.na}\NormalTok{(}\StringTok{`}\DataTypeTok{Q1-3_A~M_1981}\StringTok{`}\NormalTok{) }\OperatorTok{~}\StringTok{ `}\DataTypeTok{Q1-3_A~M_1979}\StringTok{`}\NormalTok{,}
+                      \CommentTok{# if sets don't match for dob month, take the 79 data}
+                      \StringTok{`}\DataTypeTok{Q1-3_A~M_1979}\StringTok{`} \OperatorTok{!=}\StringTok{ `}\DataTypeTok{Q1-3_A~M_1981}\StringTok{`} \OperatorTok{~}\StringTok{ `}\DataTypeTok{Q1-3_A~M_1979}\StringTok{`}\NormalTok{),}
+         \CommentTok{# flag if there is a conflict between dob recorded in 79 and 81}
+         \DataTypeTok{dob_conflict =} \KeywordTok{case_when}\NormalTok{(}
+\NormalTok{                      (}\StringTok{`}\DataTypeTok{Q1-3_A~M_1979}\StringTok{`} \OperatorTok{!=}\StringTok{ `}\DataTypeTok{Q1-3_A~M_1981}\StringTok{`}\NormalTok{) }\OperatorTok{&}\StringTok{ }\OperatorTok{!}\KeywordTok{is.na}\NormalTok{(}\StringTok{`}\DataTypeTok{Q1-3_A~M_1981}\StringTok{`}\NormalTok{)}
+                      \OperatorTok{~}\StringTok{ }\OtherTok{TRUE}\NormalTok{,}
+\NormalTok{                      (}\StringTok{`}\DataTypeTok{Q1-3_A~Y_1979}\StringTok{`} \OperatorTok{!=}\StringTok{ `}\DataTypeTok{Q1-3_A~Y_1981}\StringTok{`}\NormalTok{) }\OperatorTok{&}\StringTok{ }\OperatorTok{!}\KeywordTok{is.na}\NormalTok{(}\StringTok{`}\DataTypeTok{Q1-3_A~Y_1981}\StringTok{`}\NormalTok{)}
+                      \OperatorTok{~}\StringTok{ }\OtherTok{TRUE}\NormalTok{,}
+\NormalTok{                      (}\StringTok{`}\DataTypeTok{Q1-3_A~Y_1979}\StringTok{`} \OperatorTok{==}\StringTok{ `}\DataTypeTok{Q1-3_A~Y_1981}\StringTok{`}\NormalTok{) }\OperatorTok{&}
+\StringTok{                        }\NormalTok{(}\StringTok{`}\DataTypeTok{Q1-3_A~M_1979}\StringTok{`} \OperatorTok{==}\StringTok{ `}\DataTypeTok{Q1-3_A~M_1981}\StringTok{`}\NormalTok{) }\OperatorTok{~}\StringTok{ }\OtherTok{FALSE}\NormalTok{,}
+                      \KeywordTok{is.na}\NormalTok{(}\StringTok{`}\DataTypeTok{Q1-3_A~M_1981}\StringTok{`}\NormalTok{) }\OperatorTok{|}\StringTok{ }\KeywordTok{is.na}\NormalTok{(}\StringTok{`}\DataTypeTok{Q1-3_A~Y_1981}\StringTok{`}\NormalTok{) }\OperatorTok{~}\StringTok{ }\OtherTok{FALSE}\NormalTok{)) }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{select}\NormalTok{(CASEID_}\DecValTok{1979}\NormalTok{,}
+\NormalTok{         dob_month,}
+\NormalTok{         dob_year,}
+\NormalTok{         dob_conflict)}
 \end{Highlighting}
@@
 \begin{Highlighting}[]
-\CommentTok{\# make a list for years where we used the "QES{-}52A"}
-\NormalTok{year\_A }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(}\DecValTok{1979}\SpecialCharTok{:}\DecValTok{1987}\NormalTok{, }\DecValTok{1993}\NormalTok{)}
-\CommentTok{\#function to get the hour of work}
-\NormalTok{get\_hour }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{(year)\{}
-  \ControlFlowTok{if}\NormalTok{(year }\SpecialCharTok{\%in\%}\NormalTok{ year\_A)\{}
-\NormalTok{    temp }\OtherTok{\textless{}{-}}\NormalTok{ new\_data\_qnames }\SpecialCharTok{\%\textgreater{}\%}
-      \FunctionTok{select}\NormalTok{(CASEID\_1979,}
-                    \FunctionTok{starts\_with}\NormalTok{(}\StringTok{"QES{-}52A"}\NormalTok{) }\SpecialCharTok{\&}
-                      \FunctionTok{ends\_with}\NormalTok{(}\FunctionTok{as.character}\NormalTok{(year)))\}}
+\CommentTok{# make a list for years where we used the "QES-52A"}
+\NormalTok{year_A <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\DecValTok{1979}\OperatorTok{:}\DecValTok{1987}\NormalTok{, }\DecValTok{1993}\NormalTok{)}
+\CommentTok{#function to get the hour of work}
+\NormalTok{get_hour <-}\StringTok{ }\ControlFlowTok{function}\NormalTok{(year)\{}
+  \ControlFlowTok{if}\NormalTok{(year }\OperatorTok{%in%}\StringTok{ }\NormalTok{year_A)\{}
+\NormalTok{    temp <-}\StringTok{ }\NormalTok{new_data_qnames }\OperatorTok{%>%}
+\StringTok{      }\KeywordTok{select}\NormalTok{(CASEID_}\DecValTok{1979}\NormalTok{,}
+                    \KeywordTok{starts_with}\NormalTok{(}\StringTok{"QES-52A"}\NormalTok{) }\OperatorTok{&}
+\StringTok{                      }\KeywordTok{ends_with}\NormalTok{(}\KeywordTok{as.character}\NormalTok{(year)))\}}
   \ControlFlowTok{else}\NormalTok{\{}
-\NormalTok{    temp }\OtherTok{\textless{}{-}}\NormalTok{ new\_data\_qnames }\SpecialCharTok{\%\textgreater{}\%}
-      \FunctionTok{select}\NormalTok{(CASEID\_1979,}
-                    \FunctionTok{starts\_with}\NormalTok{(}\StringTok{"QES{-}52D"}\NormalTok{) }\SpecialCharTok{\&}
-                      \FunctionTok{ends\_with}\NormalTok{(}\FunctionTok{as.character}\NormalTok{(year)))\}}
-\NormalTok{  temp }\SpecialCharTok{\%\textgreater{}\%}
-    \FunctionTok{pivot\_longer}\NormalTok{(}\SpecialCharTok{!}\NormalTok{CASEID\_1979,}
-                 \AttributeTok{names\_to =} \StringTok{"job"}\NormalTok{,}
-                 \AttributeTok{values\_to =} \StringTok{"hours\_work"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
-    \FunctionTok{separate}\NormalTok{(}\StringTok{"job"}\NormalTok{, }\FunctionTok{c}\NormalTok{(}\StringTok{"job"}\NormalTok{, }\StringTok{"year"}\NormalTok{), }\AttributeTok{sep =} \SpecialCharTok{{-}}\DecValTok{4}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
-    \FunctionTok{mutate}\NormalTok{(}\AttributeTok{job =} \FunctionTok{paste0}\NormalTok{(}\StringTok{"job\_"}\NormalTok{, }\FunctionTok{substr}\NormalTok{(job, }\DecValTok{9}\NormalTok{, }\DecValTok{10}\NormalTok{))) }\SpecialCharTok{\%\textgreater{}\%}
-    \FunctionTok{rename}\NormalTok{(}\AttributeTok{id =}\NormalTok{ CASEID\_1979)}
+\NormalTok{    temp <-}\StringTok{ }\NormalTok{new_data_qnames }\OperatorTok{%>%}
+\StringTok{      }\KeywordTok{select}\NormalTok{(CASEID_}\DecValTok{1979}\NormalTok{,}
+                    \KeywordTok{starts_with}\NormalTok{(}\StringTok{"QES-52D"}\NormalTok{) }\OperatorTok{&}
+\StringTok{                      }\KeywordTok{ends_with}\NormalTok{(}\KeywordTok{as.character}\NormalTok{(year)))\}}
+\NormalTok{  temp }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{pivot_longer}\NormalTok{(}\OperatorTok{!}\NormalTok{CASEID_}\DecValTok{1979}\NormalTok{,}
+                 \DataTypeTok{names_to =} \StringTok{"job"}\NormalTok{,}
+                 \DataTypeTok{values_to =} \StringTok{"hours_work"}\NormalTok{) }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{separate}\NormalTok{(}\StringTok{"job"}\NormalTok{, }\KeywordTok{c}\NormalTok{(}\StringTok{"job"}\NormalTok{, }\StringTok{"year"}\NormalTok{), }\DataTypeTok{sep =} \DecValTok{-4}\NormalTok{) }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{job =} \KeywordTok{paste0}\NormalTok{(}\StringTok{"job_"}\NormalTok{, }\KeywordTok{substr}\NormalTok{(job, }\DecValTok{9}\NormalTok{, }\DecValTok{10}\NormalTok{))) }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{rename}\NormalTok{(}\DataTypeTok{id =}\NormalTok{ CASEID_}\DecValTok{1979}\NormalTok{)}
 \NormalTok{\}}
 
-\CommentTok{\# getting the hours of work of all observations}
-\NormalTok{years }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(}\DecValTok{1979}\SpecialCharTok{:}\DecValTok{1994}\NormalTok{, }\FunctionTok{seq}\NormalTok{(}\DecValTok{1996}\NormalTok{, }\DecValTok{2018}\NormalTok{, }\AttributeTok{by =} \DecValTok{2}\NormalTok{))}
-\NormalTok{hours\_all }\OtherTok{\textless{}{-}} \FunctionTok{map\_dfr}\NormalTok{(years, get\_hour)}
+\CommentTok{# getting the hours of work of all observations}
+\NormalTok{years <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\DecValTok{1979}\OperatorTok{:}\DecValTok{1994}\NormalTok{, }\KeywordTok{seq}\NormalTok{(}\DecValTok{1996}\NormalTok{, }\DecValTok{2018}\NormalTok{, }\DataTypeTok{by =} \DecValTok{2}\NormalTok{))}
+\NormalTok{hours_all <-}\StringTok{ }\KeywordTok{map_dfr}\NormalTok{(years, get_hour)}
 \end{Highlighting}
@@
 \begin{Highlighting}[]
-\NormalTok{get\_rate }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{(year) \{}
-\NormalTok{  new\_data\_qnames }\SpecialCharTok{\%\textgreater{}\%}
-    \FunctionTok{select}\NormalTok{(CASEID\_1979,}
-                  \FunctionTok{starts\_with}\NormalTok{(}\StringTok{"HRP"}\NormalTok{) }\SpecialCharTok{\&}
-                    \FunctionTok{ends\_with}\NormalTok{(}\FunctionTok{as.character}\NormalTok{(year))) }\SpecialCharTok{\%\textgreater{}\%}
-    \FunctionTok{pivot\_longer}\NormalTok{(}\SpecialCharTok{!}\NormalTok{CASEID\_1979, }\AttributeTok{names\_to =} \StringTok{"job"}\NormalTok{, }\AttributeTok{values\_to =} \StringTok{"rate\_per\_hour"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
-    \FunctionTok{separate}\NormalTok{(}\StringTok{"job"}\NormalTok{, }\FunctionTok{c}\NormalTok{(}\StringTok{"job"}\NormalTok{, }\StringTok{"year"}\NormalTok{), }\AttributeTok{sep =} \SpecialCharTok{{-}}\DecValTok{4}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
-    \FunctionTok{mutate}\NormalTok{(}\AttributeTok{job =} \FunctionTok{paste0}\NormalTok{(}\StringTok{"job\_0"}\NormalTok{, }\FunctionTok{substr}\NormalTok{(job, }\DecValTok{4}\NormalTok{, }\DecValTok{4}\NormalTok{))) }\SpecialCharTok{\%\textgreater{}\%}
-    \FunctionTok{rename}\NormalTok{(}\AttributeTok{id =}\NormalTok{ CASEID\_1979)}
+\NormalTok{get_rate <-}\StringTok{ }\ControlFlowTok{function}\NormalTok{(year) \{}
+\NormalTok{  new_data_qnames }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{select}\NormalTok{(CASEID_}\DecValTok{1979}\NormalTok{,}
+                  \KeywordTok{starts_with}\NormalTok{(}\StringTok{"HRP"}\NormalTok{) }\OperatorTok{&}
+\StringTok{                    }\KeywordTok{ends_with}\NormalTok{(}\KeywordTok{as.character}\NormalTok{(year))) }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{pivot_longer}\NormalTok{(}\OperatorTok{!}\NormalTok{CASEID_}\DecValTok{1979}\NormalTok{, }\DataTypeTok{names_to =} \StringTok{"job"}\NormalTok{, }\DataTypeTok{values_to =} \StringTok{"rate_per_hour"}\NormalTok{) }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{separate}\NormalTok{(}\StringTok{"job"}\NormalTok{, }\KeywordTok{c}\NormalTok{(}\StringTok{"job"}\NormalTok{, }\StringTok{"year"}\NormalTok{), }\DataTypeTok{sep =} \DecValTok{-4}\NormalTok{) }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{job =} \KeywordTok{paste0}\NormalTok{(}\StringTok{"job_0"}\NormalTok{, }\KeywordTok{substr}\NormalTok{(job, }\DecValTok{4}\NormalTok{, }\DecValTok{4}\NormalTok{))) }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{rename}\NormalTok{(}\DataTypeTok{id =}\NormalTok{ CASEID_}\DecValTok{1979}\NormalTok{)}
 \NormalTok{\}}
-\NormalTok{rates\_all }\OtherTok{\textless{}{-}} \FunctionTok{map\_dfr}\NormalTok{(years, get\_rate)}
-\CommentTok{\# join hours and rates variable}
-\NormalTok{hours\_wages }\OtherTok{\textless{}{-}} \FunctionTok{left\_join}\NormalTok{(rates\_all,}
-\NormalTok{                         hours\_all,}
-                         \AttributeTok{by =} \FunctionTok{c}\NormalTok{(}\StringTok{"id"}\NormalTok{, }\StringTok{"year"}\NormalTok{, }\StringTok{"job"}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%}
-  \CommentTok{\# set the 0 value in rate\_per\_hour as NA}
-  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{rate\_per\_hour =} \FunctionTok{ifelse}\NormalTok{(rate\_per\_hour }\SpecialCharTok{==} \DecValTok{0}\NormalTok{, }\ConstantTok{NA}\NormalTok{,}
-\NormalTok{                                rate\_per\_hour),}
-         \AttributeTok{hours\_work =} \FunctionTok{ifelse}\NormalTok{(hours\_work }\SpecialCharTok{==} \DecValTok{0} \SpecialCharTok{|}\NormalTok{ hours\_work }\SpecialCharTok{\textgreater{}} \DecValTok{84}\NormalTok{, }\ConstantTok{NA}\NormalTok{, hours\_work))}
+\NormalTok{rates_all <-}\StringTok{ }\KeywordTok{map_dfr}\NormalTok{(years, get_rate)}
+\CommentTok{# join hours and rates variable}
+\NormalTok{hours_wages <-}\StringTok{ }\KeywordTok{left_join}\NormalTok{(rates_all,}
+\NormalTok{                         hours_all,}
+                         \DataTypeTok{by =} \KeywordTok{c}\NormalTok{(}\StringTok{"id"}\NormalTok{, }\StringTok{"year"}\NormalTok{, }\StringTok{"job"}\NormalTok{)) }\OperatorTok{%>%}
+\StringTok{  }\CommentTok{# set the 0 value in rate_per_hour as NA}
+\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{rate_per_hour =} \KeywordTok{ifelse}\NormalTok{(rate_per_hour }\OperatorTok{==}\StringTok{ }\DecValTok{0}\NormalTok{, }\OtherTok{NA}\NormalTok{,}
+\NormalTok{                                rate_per_hour),}
+         \DataTypeTok{hours_work =} \KeywordTok{ifelse}\NormalTok{(hours_work }\OperatorTok{==}\StringTok{ }\DecValTok{0} \OperatorTok{|}\StringTok{ }\NormalTok{hours_work }\OperatorTok{>}\StringTok{ }\DecValTok{84}\NormalTok{, }\OtherTok{NA}\NormalTok{, hours_work))}
 \end{Highlighting}
@@
 
-{\centering \includegraphics[width=0.93\linewidth]{/Users/cookd/numbats/summer-wages-refresh/paper/figures/flowchart}
+{\centering \includegraphics[width=0.93\linewidth]{/Users/deeamaliah/Documents/summer_project/summer-wages-refresh/paper/figures/flowchart}
 
@@
 \leavevmode\hypertarget{ref-KollerManuel2016rARP}{}%
-Koller, Manuel. 2016. {``Robustlmm: An r Package for Robust Estimation of Linear Mixed-Effects Models.''} \emph{Journal of Statistical Software} 75 (6): 1--24.
+Koller, Manuel. 2016. ``Robustlmm: An R Package for Robust Estimation of Linear Mixed-Effects Models.'' \emph{Journal of Statistical Software} 75 (6): 1--24.
 
 \leavevmode\hypertarget{ref-LooMarkvander2018Sdcw}{}%
-Loo, Mark van der. 2018. \emph{Statistical Data Cleaning with Applications in r}.
+Loo, Mark van der. 2018. \emph{Statistical Data Cleaning with Applications in R}.
 
 \leavevmode\hypertarget{ref-rsvg}{}%
-Ooms, Jeroen. 2020. \emph{Rsvg: Render SVG Images into PDF, PNG, PostScript, or Bitmap Arrays}. \url{https://CRAN.R-project.org/package=rsvg}.
+Ooms, Jeroen. 2020. \emph{Rsvg: Render Svg Images into Pdf, Png, Postscript, or Bitmap Arrays}. \url{https://CRAN.R-project.org/package=rsvg}.
 
 \leavevmode\hypertarget{ref-opendata}{}%
-Open Knowledge Foundation. 2021. {``Open Definition. Defining Open in Open Data, Open Content, and Open Knowledge.''} 2021. \url{http://opendefinition.org/od/2.1/en/}.
+Open Knowledge Foundation. 2021. ``Open Definition. Defining Open in Open Data, Open Content, and Open Knowledge.'' 2021. \url{http://opendefinition.org/od/2.1/en/}.
 
@@
 \leavevmode\hypertarget{ref-plyr}{}%
-Wickham, Hadley. 2011. {``The Split-Apply-Combine Strategy for Data Analysis.''} \emph{Journal of Statistical Software, Articles} 40 (1): 1--29. \url{https://doi.org/10.18637/jss.v040.i01}.
+Wickham, Hadley. 2011. ``The Split-Apply-Combine Strategy for Data Analysis.'' \emph{Journal of Statistical Software, Articles} 40 (1): 1--29. \url{https://doi.org/10.18637/jss.v040.i01}.
 
 \leavevmode\hypertarget{ref-WickhamHadley2014TD}{}%
----------. 2014. {``Tidy Data.''} \emph{Journal of Statistical Software} 59 (10): 1--23.
+---------. 2014. ``Tidy Data.'' \emph{Journal of Statistical Software} 59 (10): 1--23.
 

### 2021-09-22 paper/paper.Rmd
target data description cleaned up

@@
 Section \@ref(cleaning) presents the steps of cleaning the data, including getting and tidying the data from the NLSY79 and IDA to find and repair anomalies.
-Our final subset is compared the old textbook subset in Section \@ref(comare).
+Our final subset is compared the old textbook subset in Section \@ref(compare).
 Finally, Section \@ref(summary) summarizes the contribution and makes recommendations for the NLSY79 data curators.
@@
 
-[FILL] We should include how many variables there are in the data so people get a grasp of the extensive coverage of the variables.
-
 Although the main focus area of the NLSY is labor and employment, the NLSY also covers several other topics, including education; training and achievement; household, geography, and contextual variables; dating, marriage, cohabitation; sexual activity, pregnancy, and fertility; children; income, assets and program participation; health; attitudes and expectations; and crime and substance use.
@@
 
-As mentioned in Section \@ref(intro), this paper aims to refresh the NLSY79 data as used in Singer and Willet's textbook (@SingerJudithD2003Alda).
-This data set contains the information of yearly mean hourly wages from the NLSY79 cohort, with education and race as covariates.
-This data set represents the measurement of male high-school dropouts, aged from 14 to 17 years old when first time measured.
-Accordingly, to get this data, we have two additional target data, NLSY79 cohort's demographic profiles and wages data of people who completed education up to high school.
+This paper aims, as mentioned in Section \@ref(intro), to refresh the NLSY79 data used in @SingerJudithD2003Alda.
+This data set contains the information of yearly mean hourly wages from the NLSY79 subcohort, with education and race as covariates.
+The subcohort correspond to male high-school dropouts that participated in the study from 14 to 17 years old.
+Accordingly, to refresh this data, we make use of NLSY79 cohort's demographic profiles and wages data of people who completed education up to high school.
 
-Since there is no information on high school dropouts in Singer and Willet's book, we define high school dropouts as respondents who only completed either $9^{th}$, $10^{th}$, $11^{th}$ grade, or who completed $12^{th}$ when their age is at least 19 years old (means that these IDs being dropped out in a certain year, but they returned back to high school to complete it).
+Since no specific criteria on high school dropouts was defined
+in @SingerJudithD2003Alda, we define high school dropouts as respondents who only completed either 9th to 11th grade, or who completed 12th when their age is at least 19 years old (meaning that the individual may have dropped out of high school but returned at a later time to complete high school education).
 
@@
 
-```{r dob-tidy, echo=TRUE}
+```{r dob-tidy}
 ```
@@
 
-```{r tidy-hours, echo=TRUE}
+```{r tidy-hours}
 ```
@@
 
-```{r tidy-rate, echo = TRUE}
+```{r tidy-rate}
 ```
@@
 
-# Exploratory Data Analysis
+# Comparison with the Textbook Data {#compare}
 

### 2021-09-22 paper/paper.tex
target data description cleaned up

--- 2021-09-21 paper/paper.tex
@@
 
-\usepackage{color}
-\usepackage{fancyvrb}
-\newcommand{\VerbBar}{|}
-\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
-\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
-% Add ',fontsize=\small' for more characters per line
-\usepackage{framed}
-\definecolor{shadecolor}{RGB}{248,248,248}
-\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
-\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
-\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
-\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
-\newcommand{\BuiltInTok}[1]{#1}
-\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
-\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
-\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
-\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
-\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
-\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
-\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
-\newcommand{\ExtensionTok}[1]{#1}
-\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
-\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
-\newcommand{\ImportTok}[1]{#1}
-\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
-\newcommand{\NormalTok}[1]{#1}
-\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
-\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
-\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
-\newcommand{\RegionMarkerTok}[1]{#1}
-\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
-\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
-\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
-\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
-\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
-\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-
 % Pandoc citation processing
@@
 Section \ref{cleaning} presents the steps of cleaning the data, including getting and tidying the data from the NLSY79 and IDA to find and repair anomalies.
-Our final subset is compared the old textbook subset in Section \ref{comare}.
-Finally, Section \ref{summary} summarises the contribution and makes recommendations for the NLSY79 data curators.
+Our final subset is compared the old textbook subset in Section \ref{compare}.
+Finally, Section \ref{summary} summarizes the contribution and makes recommendations for the NLSY79 data curators.
 
@@
 
-{[}FILL{]} We should include how many variables there are in the data so people get a grasp of the extensive coverage of the variables.
-
 Although the main focus area of the NLSY is labor and employment, the NLSY also covers several other topics, including education; training and achievement; household, geography, and contextual variables; dating, marriage, cohabitation; sexual activity, pregnancy, and fertility; children; income, assets and program participation; health; attitudes and expectations; and crime and substance use.
@@
 
-As mentioned in Section \ref{intro}, this paper aims to refresh the NLSY79 data as used in Singer and Willet's textbook (Singer and Willett (2003)).
-This data set contains the information of yearly mean hourly wages from the NLSY79 cohort, with education and race as covariates.
-This data set represents the measurement of male high-school dropouts, aged from 14 to 17 years old when first time measured.
-Accordingly, to get this data, we have two additional target data, NLSY79 cohort's demographic profiles and wages data of people who completed education up to high school.
+This paper aims, as mentioned in Section \ref{intro}, to refresh the NLSY79 data used in Singer and Willett (2003).
+This data set contains the information of yearly mean hourly wages from the NLSY79 subcohort, with education and race as covariates.
+The subcohort correspond to male high-school dropouts that participated in the study from 14 to 17 years old.
+Accordingly, to refresh this data, we make use of NLSY79 cohort's demographic profiles and wages data of people who completed education up to high school.
 
-Since there is no information on high school dropouts in Singer and Willet's book, we define high school dropouts as respondents who only completed either \(9^{th}\), \(10^{th}\), \(11^{th}\) grade, or who completed \(12^{th}\) when their age is at least 19 years old (means that these IDs being dropped out in a certain year, but they returned back to high school to complete it).
+Since no specific criteria on high school dropouts was defined
+in Singer and Willett (2003), we define high school dropouts as respondents who only completed either 9th to 11th grade, or who completed 12th when their age is at least 19 years old (meaning that the individual may have dropped out of high school but returned at a later time to complete high school education).
 
@@
 
-\begin{Shaded}
-\begin{Highlighting}[]
-\CommentTok{\#tidy the date of birth data}
-\NormalTok{dob\_tidy }\OtherTok{\textless{}{-}}\NormalTok{ new\_data\_qnames }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{select}\NormalTok{(CASEID\_1979,}
-                \FunctionTok{starts\_with}\NormalTok{(}\StringTok{"Q1{-}3\_A\textasciitilde{}"}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{dob\_year =} \FunctionTok{case\_when}\NormalTok{(}
-                      \CommentTok{\# if the years recorded in both sets match, take 79 data}
-                      \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}Y\_1979}\StringTok{\textasciigrave{}} \SpecialCharTok{==} \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}Y\_1981}\StringTok{\textasciigrave{}} \SpecialCharTok{\textasciitilde{}} \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}Y\_1979}\StringTok{\textasciigrave{}}\NormalTok{,}
-                      \CommentTok{\# if the year in the 81 set is missing, take the 79 data}
-                      \FunctionTok{is.na}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}Y\_1981}\StringTok{\textasciigrave{}}\NormalTok{) }\SpecialCharTok{\textasciitilde{}} \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}Y\_1979}\StringTok{\textasciigrave{}}\NormalTok{,}
-                      \CommentTok{\# if the sets don\textquotesingle{}t match for dob year, take the 79 data}
-                      \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}Y\_1979}\StringTok{\textasciigrave{}} \SpecialCharTok{!=} \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}Y\_1981}\StringTok{\textasciigrave{}} \SpecialCharTok{\textasciitilde{}} \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}Y\_1979}\StringTok{\textasciigrave{}}\NormalTok{),}
-         \AttributeTok{dob\_month =} \FunctionTok{case\_when}\NormalTok{(}
-                      \CommentTok{\# if months recorded in both sets match, take 79 data}
-                      \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}M\_1979}\StringTok{\textasciigrave{}} \SpecialCharTok{==} \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}M\_1981}\StringTok{\textasciigrave{}} \SpecialCharTok{\textasciitilde{}} \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}M\_1979}\StringTok{\textasciigrave{}}\NormalTok{,}
-                      \CommentTok{\# if month in 81 set is missing, take the 79 data}
-                      \FunctionTok{is.na}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}M\_1981}\StringTok{\textasciigrave{}}\NormalTok{) }\SpecialCharTok{\textasciitilde{}} \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}M\_1979}\StringTok{\textasciigrave{}}\NormalTok{,}
-                      \CommentTok{\# if sets don\textquotesingle{}t match for dob month, take the 79 data}
-                      \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}M\_1979}\StringTok{\textasciigrave{}} \SpecialCharTok{!=} \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}M\_1981}\StringTok{\textasciigrave{}} \SpecialCharTok{\textasciitilde{}} \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}M\_1979}\StringTok{\textasciigrave{}}\NormalTok{),}
-         \CommentTok{\# flag if there is a conflict between dob recorded in 79 and 81}
-         \AttributeTok{dob\_conflict =} \FunctionTok{case\_when}\NormalTok{(}
-\NormalTok{                      (}\StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}M\_1979}\StringTok{\textasciigrave{}} \SpecialCharTok{!=} \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}M\_1981}\StringTok{\textasciigrave{}}\NormalTok{) }\SpecialCharTok{\&} \SpecialCharTok{!}\FunctionTok{is.na}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}M\_1981}\StringTok{\textasciigrave{}}\NormalTok{)}
-                      \SpecialCharTok{\textasciitilde{}} \ConstantTok{TRUE}\NormalTok{,}
-\NormalTok{                      (}\StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}Y\_1979}\StringTok{\textasciigrave{}} \SpecialCharTok{!=} \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}Y\_1981}\StringTok{\textasciigrave{}}\NormalTok{) }\SpecialCharTok{\&} \SpecialCharTok{!}\FunctionTok{is.na}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}Y\_1981}\StringTok{\textasciigrave{}}\NormalTok{)}
-                      \SpecialCharTok{\textasciitilde{}} \ConstantTok{TRUE}\NormalTok{,}
-\NormalTok{                      (}\StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}Y\_1979}\StringTok{\textasciigrave{}} \SpecialCharTok{==} \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}Y\_1981}\StringTok{\textasciigrave{}}\NormalTok{) }\SpecialCharTok{\&}
-\NormalTok{                        (}\StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}M\_1979}\StringTok{\textasciigrave{}} \SpecialCharTok{==} \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}M\_1981}\StringTok{\textasciigrave{}}\NormalTok{) }\SpecialCharTok{\textasciitilde{}} \ConstantTok{FALSE}\NormalTok{,}
-                      \FunctionTok{is.na}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}M\_1981}\StringTok{\textasciigrave{}}\NormalTok{) }\SpecialCharTok{|} \FunctionTok{is.na}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}Y\_1981}\StringTok{\textasciigrave{}}\NormalTok{) }\SpecialCharTok{\textasciitilde{}} \ConstantTok{FALSE}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{select}\NormalTok{(CASEID\_1979,}
-\NormalTok{         dob\_month,}
-\NormalTok{         dob\_year,}
-\NormalTok{         dob\_conflict)}
-\end{Highlighting}
-\end{Shaded}
-
 For \texttt{gender} and \texttt{race}, we only rename the column names.
@@
 
-\begin{Shaded}
-\begin{Highlighting}[]
-\CommentTok{\# make a list for years where we used the "QES{-}52A"}
-\NormalTok{year\_A }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(}\DecValTok{1979}\SpecialCharTok{:}\DecValTok{1987}\NormalTok{, }\DecValTok{1993}\NormalTok{)}
-\CommentTok{\#function to get the hour of work}
-\NormalTok{get\_hour }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{(year)\{}
-  \ControlFlowTok{if}\NormalTok{(year }\SpecialCharTok{\%in\%}\NormalTok{ year\_A)\{}
-\NormalTok{    temp }\OtherTok{\textless{}{-}}\NormalTok{ new\_data\_qnames }\SpecialCharTok{\%\textgreater{}\%}
-      \FunctionTok{select}\NormalTok{(CASEID\_1979,}
-                    \FunctionTok{starts\_with}\NormalTok{(}\StringTok{"QES{-}52A"}\NormalTok{) }\SpecialCharTok{\&}
-                      \FunctionTok{ends\_with}\NormalTok{(}\FunctionTok{as.character}\NormalTok{(year)))\}}
-  \ControlFlowTok{else}\NormalTok{\{}
-\NormalTok{    temp }\OtherTok{\textless{}{-}}\NormalTok{ new\_data\_qnames }\SpecialCharTok{\%\textgreater{}\%}
-      \FunctionTok{select}\NormalTok{(CASEID\_1979,}
-                    \FunctionTok{starts\_with}\NormalTok{(}\StringTok{"QES{-}52D"}\NormalTok{) }\SpecialCharTok{\&}
-                      \FunctionTok{ends\_with}\NormalTok{(}\FunctionTok{as.character}\NormalTok{(year)))\}}
-\NormalTok{  temp }\SpecialCharTok{\%\textgreater{}\%}
-    \FunctionTok{pivot\_longer}\NormalTok{(}\SpecialCharTok{!}\NormalTok{CASEID\_1979,}
-                 \AttributeTok{names\_to =} \StringTok{"job"}\NormalTok{,}
-                 \AttributeTok{values\_to =} \StringTok{"hours\_work"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
-    \FunctionTok{separate}\NormalTok{(}\StringTok{"job"}\NormalTok{, }\FunctionTok{c}\NormalTok{(}\StringTok{"job"}\NormalTok{, }\StringTok{"year"}\NormalTok{), }\AttributeTok{sep =} \SpecialCharTok{{-}}\DecValTok{4}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
-    \FunctionTok{mutate}\NormalTok{(}\AttributeTok{job =} \FunctionTok{paste0}\NormalTok{(}\StringTok{"job\_"}\NormalTok{, }\FunctionTok{substr}\NormalTok{(job, }\DecValTok{9}\NormalTok{, }\DecValTok{10}\NormalTok{))) }\SpecialCharTok{\%\textgreater{}\%}
-    \FunctionTok{rename}\NormalTok{(}\AttributeTok{id =}\NormalTok{ CASEID\_1979)}
-\NormalTok{\}}
-
-\CommentTok{\# getting the hours of work of all observations}
-\NormalTok{years }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(}\DecValTok{1979}\SpecialCharTok{:}\DecValTok{1994}\NormalTok{, }\FunctionTok{seq}\NormalTok{(}\DecValTok{1996}\NormalTok{, }\DecValTok{2018}\NormalTok{, }\AttributeTok{by =} \DecValTok{2}\NormalTok{))}
-\NormalTok{hours\_all }\OtherTok{\textless{}{-}} \FunctionTok{map\_dfr}\NormalTok{(years, get\_hour)}
-\end{Highlighting}
-\end{Shaded}
-
 The same way was also deployed to tidy the rate of wage by year and by ID.
@@
 Here, we also censored the observations with \textgreater{} 84 hours of work a week to be NA.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{get\_rate }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{(year) \{}
-\NormalTok{  new\_data\_qnames }\SpecialCharTok{\%\textgreater{}\%}
-    \FunctionTok{select}\NormalTok{(CASEID\_1979,}
-                  \FunctionTok{starts\_with}\NormalTok{(}\StringTok{"HRP"}\NormalTok{) }\SpecialCharTok{\&}
-                    \FunctionTok{ends\_with}\NormalTok{(}\FunctionTok{as.character}\NormalTok{(year))) }\SpecialCharTok{\%\textgreater{}\%}
-    \FunctionTok{pivot\_longer}\NormalTok{(}\SpecialCharTok{!}\NormalTok{CASEID\_1979, }\AttributeTok{names\_to =} \StringTok{"job"}\NormalTok{, }\AttributeTok{values\_to =} \StringTok{"rate\_per\_hour"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
-    \FunctionTok{separate}\NormalTok{(}\StringTok{"job"}\NormalTok{, }\FunctionTok{c}\NormalTok{(}\StringTok{"job"}\NormalTok{, }\StringTok{"year"}\NormalTok{), }\AttributeTok{sep =} \SpecialCharTok{{-}}\DecValTok{4}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
-    \FunctionTok{mutate}\NormalTok{(}\AttributeTok{job =} \FunctionTok{paste0}\NormalTok{(}\StringTok{"job\_0"}\NormalTok{, }\FunctionTok{substr}\NormalTok{(job, }\DecValTok{4}\NormalTok{, }\DecValTok{4}\NormalTok{))) }\SpecialCharTok{\%\textgreater{}\%}
-    \FunctionTok{rename}\NormalTok{(}\AttributeTok{id =}\NormalTok{ CASEID\_1979)}
-\NormalTok{\}}
-\NormalTok{rates\_all }\OtherTok{\textless{}{-}} \FunctionTok{map\_dfr}\NormalTok{(years, get\_rate)}
-\CommentTok{\# join hours and rates variable}
-\NormalTok{hours\_wages }\OtherTok{\textless{}{-}} \FunctionTok{left\_join}\NormalTok{(rates\_all,}
-\NormalTok{                         hours\_all,}
-                         \AttributeTok{by =} \FunctionTok{c}\NormalTok{(}\StringTok{"id"}\NormalTok{, }\StringTok{"year"}\NormalTok{, }\StringTok{"job"}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%}
-  \CommentTok{\# set the 0 value in rate\_per\_hour as NA}
-  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{rate\_per\_hour =} \FunctionTok{ifelse}\NormalTok{(rate\_per\_hour }\SpecialCharTok{==} \DecValTok{0}\NormalTok{, }\ConstantTok{NA}\NormalTok{,}
-\NormalTok{                                rate\_per\_hour),}
-         \AttributeTok{hours\_work =} \FunctionTok{ifelse}\NormalTok{(hours\_work }\SpecialCharTok{==} \DecValTok{0} \SpecialCharTok{|}\NormalTok{ hours\_work }\SpecialCharTok{\textgreater{}} \DecValTok{84}\NormalTok{, }\ConstantTok{NA}\NormalTok{, hours\_work))}
-\end{Highlighting}
-\end{Shaded}
 
@@
 
-{\centering \includegraphics[width=0.93\linewidth]{/Users/cookd/numbats/summer-wages-refresh/paper/figures/flowchart}
+{\centering \includegraphics[width=0.93\linewidth]{/Users/etan0038/Dropbox/repo/summer-wages-refresh/paper/figures/flowchart}
 
@@
 
-\hypertarget{exploratory-data-analysis}{%
-\section{Exploratory Data Analysis}\label{exploratory-data-analysis}}
+\hypertarget{compare}{%
+\section{Comparison with the Textbook Data}\label{compare}}
 

### 2021-09-23 paper/scripts/eda.R
Only display one comparison plot.

@@ -37,14 +37,13 @@ do_ref <- do %>%
   theme(plot.title = element_text(size = 10)) +
   ylim(-4, 6)
 
-sw_wages + do_ref
+# sw_wages + do_ref
 
 # the two plots are not comparable
 # since the id who are in Singer-Willet and the id in do-refreshed might be different
 
 
 ## --- compare-sw-do
-
 sw_id <- as_tibble(sw) %>%
   group_by(id) %>%
   count() %>%
@@ -91,13 +90,11 @@ do_ref_agree <- do_agree %>%
   ylim(-3, 5)
 
 
-sw_wages_agree + do_ref_agree
+#sw_wages_agree + do_ref_agree
 # even though it already has the same id, it still has different period measured
 # hence the following using the same period apprx.
 
 ## ---- compare-period
-
-
 # filter the xp in refreshed data in order to get the same experience with Singer-Willet data,
 # which is 13 years of experience
 do_agree_cutxp <- do_agree %>%

### 2021-09-23 paper/paper.Rmd
Add crossreference for target data.
Add explanation in comparison section.

@@
 
-## Target Data
+## Target Data {#target}
 
@@
 
+```{r sw_wages}
+```
+
+```{r do_refreshed}
+```
+
+```{r compare-sw-do}
+```
+
+As mentioned previously, the aim of this paper is to refresh the high school dropouts wages used in textbook [@SingerJudithD2003Alda] as similar as possible.
+Hence, in this section, we will compare how close the `wages_hs_do`, henceforth refreshed data, to the textbook data.
+It is also worth noting that the textbook data is reused from @MurnaneRichardJ1999DMDB’s paper.
+Hence, it is not directly extracted from the NLSY79 database.
+@SingerJudithD2003Alda do not state the time frame covered by the data.
+However, from @MurnaneRichardJ1999DMDB’s paper, we find that the textbook data cover the survey period from 1979-1994.
+Hence, for comparison purposes, we filter the refreshed data only until 1994.
+
+In the absence of highs chool dropouts criteria in the textbook data, we create a clearer criteria of high school dropouts as mentioned in Section \@ref(#target).
+Consequently, we do not have the same ID of high-school dropouts as in the textbook data.
+The textbook data has `r nrow(sw_id)` IDs included, while the refreshed data has `r nrow(do_id)` IDs.
+Furthermore, there are only `r nrow(sw_do)` IDs included both in the textbook and the refreshed data.
+After observing @MurnaneRichardJ1999DMDB’s paper, it is mentioned that the data is randomly drawn from the low income subsample of the NLSY.
+This is probably the reason why, even using some criteria, we cannot match the dropouts in the textbook and the refreshed data.
+
+Further, the temporal variable used in the textbook data is experience, instead of year.
+The only explanation we found regarding this variable is that it is derived from the years of IDs’ experience since their first job.
+@SingerJudithD2003Alda stated that this variable reflects the specific moment in each ID labor force history associated with each wage observation.
+Since it is not clearly articulated, we decide to keep the temporal variable in the refreshed data as year.
+Accordingly, in this comparison, we only compare the same ID but not the same temporal variable.
+
+Regarding the wages, the textbook data express the wages as inflation adjusted with 1990 dollars as constant [@SingerJudithD2003Alda].
+However, it is not stated how exactly the wages are adjusted.
+Note that in the refreshed data we publish in `yowie`, we do not adjust the data with inflation because the time frame covered ranged from 1979-2018.
+Hence, we allow flexibility to the data user to adjust, not adjust,  or adjust it with different base years.
+For comparison purposes, we adjust the wages with the average Consumer Price Index For All Urban Consumers (CPI-U) we got from @fed.
+
+```{r plotting-sw-do, fig.cap= "Textbook data (A) and Refreshed data (B). The textbook data and refreshed data is not exactly the same because it has different temporal variables."}
+```
+
+Figure \ref{fig:plotting-sw-do} shows the comparison between the textbook and the refreshed data.
+Even though it has different temporal variables and possibly different CPI indicators used in inflation adjustment, we can still capture some of the patterns in the refreshed data.
+
+From this comparison, we learn that we cannot make an apple to apple comparison and produce the exact same data as the textbook due to the absence of clearly articulated data extraction and derivation process.
+Hence, the takeaway is that it is imperative to disclose how the data derived from the initial source to make data analysis reproducible.
+This is why in this paper, we show the practice of reproducible and responsible data cleaning workflow to make sure the longitudinal data from a continuous survey can be refreshed.
+
 # Summary {#summary}

### 2021-09-23 paper/paper-jsdse/paper-jsdse.tex
tidy data section

--- 2021-09-21 paper/paper-jsdse/paper-jsdse.tex
@@
 {#1}\small\normalsize} \spacingset{1}
-
-\usepackage{color}
-\usepackage{fancyvrb}
-\newcommand{\VerbBar}{|}
-\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
-\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
-% Add ',fontsize=\small' for more characters per line
-\usepackage{framed}
-\definecolor{shadecolor}{RGB}{248,248,248}
-\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
-\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
-\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
-\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
-\newcommand{\BuiltInTok}[1]{#1}
-\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
-\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
-\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
-\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
-\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
-\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
-\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
-\newcommand{\ExtensionTok}[1]{#1}
-\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
-\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
-\newcommand{\ImportTok}[1]{#1}
-\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
-\newcommand{\NormalTok}[1]{#1}
-\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
-\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
-\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
-\newcommand{\RegionMarkerTok}[1]{#1}
-\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
-\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
-\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
-\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
-\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
-\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
 
@@
 Singer and Willett (2003) used the wages and other variables of high school dropouts from the NLSY79 data as an example data set to illustrate longitudinal data modeling.
-Our aim is to refresh this textbook data with data from .
-Here, we investigate the process of getting from the raw NLSY79 to this textbook data set.
-However, we cannot create the exact same data set as published in their book since we do not have the information of what age threshold they used to determine the high school dropouts.
+Our aim is to refresh this textbook data with data from 1994 through to the latest data reported in 2018.
+Here, we investigate the process of getting from the raw NLSY79 data to a textbook data set as similar as possible to that provided by Singer and Willett (2003).
 
-This paper is structured to have 5 sections.
-Section \ref{database} describes the original data source.
-Section \ref{cleaning} presents the steps of cleaning the data, including getting and tidying the data from the NLSY79 and initial data analysis to find and treat the anomalies in the data.
-We also gave examples of exploratory data analysis using the clean data in Section \ref{eda}.
-Finally, Section \ref{summary} summarises the paper.
+Our process of cleaning builds heavily on the tidyverse approach (Wickham et al.
+2019).
+The data is first organised into ``tidy data'' (Wickham 2014) and then further wrangled using the data pipeline and split-apply-combine approach (Wickham 2011).
+The resulting data is provided in a new R package called \texttt{yowie} which includes the code so that the process is reproducible, and could be used to further refresh the data as new records are made available in the NLSY79 database.
+
+This paper is structured in the following way.
+Section \ref{database} describes the NLSY79 data source.
+Section \ref{cleaning} presents the steps of cleaning the data, including getting and tidying the data from the NLSY79 and IDA to find and repair anomalies.
+Our final subset is compared the old textbook subset in Section \ref{compare}.
+Finally, Section \ref{summary} summarizes the contribution and makes recommendations for the NLSY79 data curators.
 
@@
 
-\hypertarget{target-data}{%
-\subsection{Target Data}\label{target-data}}
+\hypertarget{target}{%
+\subsection{Target Data}\label{target}}
 
-As mentioned in Section \ref{intro}, this paper aims to refresh the NLSY79 data as used in Singer and Willet's textbook (Singer and Willett (2003)).
-This data set contains the information of yearly mean hourly wages from the NLSY79 cohort, with education and race as covariates.
-This data set represents the measurement of male high-school dropouts, aged from 14 to 17 years old when first time measured.
-Accordingly, to get this data, we have two additional target data, NLSY79 cohort's demographic profiles and wages data of people who completed education up to high school.
+This paper aims, as mentioned in Section \ref{intro}, to refresh the NLSY79 data used in Singer and Willett (2003).
+This data set contains the information of yearly mean hourly wages from the NLSY79 subcohort, with education and race as covariates.
+The subcohort correspond to male high-school dropouts that participated in the study from 14 to 17 years old.
+Accordingly, to refresh this data, we make use of NLSY79 cohort's demographic profiles and wages data of people who completed education up to high school.
 
-Since there is no information on high school dropouts in Singer and Willet's book, we define high school dropouts as respondents who only completed either \(9^{th}\), \(10^{th}\), \(11^{th}\) grade, or who completed \(12^{th}\) when their age is at least 19 years old (means that these IDs being dropped out in a certain year, but they returned back to high school to complete it).
+Since no specific criteria on high school dropouts was defined
+in Singer and Willett (2003), we define high school dropouts as respondents who only completed either 9th to 11th grade, or who completed 12th when their age is at least 19 years old (meaning that the individual may have dropped out of high school but returned at a later time to complete high school education).
 
@@
 
-The NLSY79 data are stored in a \href{https://www.nlsinfo.org/content/cohorts/nlsy79/get-data}{database} and could be downloaded by variables.
-Several variables are available for download and could be browsed by index.
-For the wages data set, we only extracted these variables:
+The NLSY79 data contains a large scope of variables but we limit the scope of the data to demographic profiles and wages data.
+To extract the relevant component of the data, we went to the NLSY79 database website at \url{https://www.nlsinfo.org/content/cohorts/nlsy79/get-data} and navigated as described below.
 
@@
 
-According to Wickham (2014), tidy data sets comply with three rules.
-Firstly, each variable forms a column.
-Secondly, each observation forms a row.
-Lastly, each type of observational unit forms a table.
+According to Wickham (2014), tidy data sets comply with three rules: (i) each variable forms a column, (ii) each observation forms a row, and (iii) each type of observational unit forms a table.
+The input data, \texttt{new\_data\_qnames}, is organised such that each row corresponds to an individual.
+As respondents can have multiple jobs at specific years, the column names, such as \texttt{HRP1\_1980}, such as contain the information about the job number and the year.
+The input data consequently has a large number of columns ()
+
 Unfortunately, the \texttt{new\_data\_qnames} did not meet these requirements in the way that data for each year and job for the same respondent is stored as one variable.
@@
 We only used the 1979 data and did a consistency check of those years and flag the inconsistent observations.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\CommentTok{\#tidy the date of birth data}
-\NormalTok{dob\_tidy }\OtherTok{\textless{}{-}}\NormalTok{ new\_data\_qnames }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{select}\NormalTok{(CASEID\_1979,}
-                \FunctionTok{starts\_with}\NormalTok{(}\StringTok{"Q1{-}3\_A\textasciitilde{}"}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{dob\_year =} \FunctionTok{case\_when}\NormalTok{(}
-                      \CommentTok{\# if the years recorded in both sets match, take 79 data}
-                      \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}Y\_1979}\StringTok{\textasciigrave{}} \SpecialCharTok{==} \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}Y\_1981}\StringTok{\textasciigrave{}} \SpecialCharTok{\textasciitilde{}} \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}Y\_1979}\StringTok{\textasciigrave{}}\NormalTok{,}
-                      \CommentTok{\# if the year in the 81 set is missing, take the 79 data}
-                      \FunctionTok{is.na}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}Y\_1981}\StringTok{\textasciigrave{}}\NormalTok{) }\SpecialCharTok{\textasciitilde{}} \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}Y\_1979}\StringTok{\textasciigrave{}}\NormalTok{,}
-                      \CommentTok{\# if the sets don\textquotesingle{}t match for dob year, take the 79 data}
-                      \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}Y\_1979}\StringTok{\textasciigrave{}} \SpecialCharTok{!=} \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}Y\_1981}\StringTok{\textasciigrave{}} \SpecialCharTok{\textasciitilde{}} \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}Y\_1979}\StringTok{\textasciigrave{}}\NormalTok{),}
-         \AttributeTok{dob\_month =} \FunctionTok{case\_when}\NormalTok{(}
-                      \CommentTok{\# if months recorded in both sets match, take 79 data}
-                      \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}M\_1979}\StringTok{\textasciigrave{}} \SpecialCharTok{==} \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}M\_1981}\StringTok{\textasciigrave{}} \SpecialCharTok{\textasciitilde{}} \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}M\_1979}\StringTok{\textasciigrave{}}\NormalTok{,}
-                      \CommentTok{\# if month in 81 set is missing, take the 79 data}
-                      \FunctionTok{is.na}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}M\_1981}\StringTok{\textasciigrave{}}\NormalTok{) }\SpecialCharTok{\textasciitilde{}} \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}M\_1979}\StringTok{\textasciigrave{}}\NormalTok{,}
-                      \CommentTok{\# if sets don\textquotesingle{}t match for dob month, take the 79 data}
-                      \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}M\_1979}\StringTok{\textasciigrave{}} \SpecialCharTok{!=} \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}M\_1981}\StringTok{\textasciigrave{}} \SpecialCharTok{\textasciitilde{}} \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}M\_1979}\StringTok{\textasciigrave{}}\NormalTok{),}
-         \CommentTok{\# flag if there is a conflict between dob recorded in 79 and 81}
-         \AttributeTok{dob\_conflict =} \FunctionTok{case\_when}\NormalTok{(}
-\NormalTok{                      (}\StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}M\_1979}\StringTok{\textasciigrave{}} \SpecialCharTok{!=} \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}M\_1981}\StringTok{\textasciigrave{}}\NormalTok{) }\SpecialCharTok{\&} \SpecialCharTok{!}\FunctionTok{is.na}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}M\_1981}\StringTok{\textasciigrave{}}\NormalTok{)}
-                      \SpecialCharTok{\textasciitilde{}} \ConstantTok{TRUE}\NormalTok{,}
-\NormalTok{                      (}\StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}Y\_1979}\StringTok{\textasciigrave{}} \SpecialCharTok{!=} \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}Y\_1981}\StringTok{\textasciigrave{}}\NormalTok{) }\SpecialCharTok{\&} \SpecialCharTok{!}\FunctionTok{is.na}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}Y\_1981}\StringTok{\textasciigrave{}}\NormalTok{)}
-                      \SpecialCharTok{\textasciitilde{}} \ConstantTok{TRUE}\NormalTok{,}
-\NormalTok{                      (}\StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}Y\_1979}\StringTok{\textasciigrave{}} \SpecialCharTok{==} \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}Y\_1981}\StringTok{\textasciigrave{}}\NormalTok{) }\SpecialCharTok{\&}
-\NormalTok{                        (}\StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}M\_1979}\StringTok{\textasciigrave{}} \SpecialCharTok{==} \StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}M\_1981}\StringTok{\textasciigrave{}}\NormalTok{) }\SpecialCharTok{\textasciitilde{}} \ConstantTok{FALSE}\NormalTok{,}
-                      \FunctionTok{is.na}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}M\_1981}\StringTok{\textasciigrave{}}\NormalTok{) }\SpecialCharTok{|} \FunctionTok{is.na}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Q1{-}3\_A\textasciitilde{}Y\_1981}\StringTok{\textasciigrave{}}\NormalTok{) }\SpecialCharTok{\textasciitilde{}} \ConstantTok{FALSE}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{select}\NormalTok{(CASEID\_1979,}
-\NormalTok{         dob\_month,}
-\NormalTok{         dob\_year,}
-\NormalTok{         dob\_conflict)}
-\end{Highlighting}
-\end{Shaded}
 

### 2021-09-23 paper/paper.Rmd
tidy data section

--- 2021-09-22 paper/paper.Rmd
@@
 
-The NLSY79 data are stored in a [database](https://www.nlsinfo.org/content/cohorts/nlsy79/get-data) and could be downloaded by variables.
-Several variables are available for download and could be browsed by index.
-For the wages data set, we only extracted these variables:
+The NLSY79 data contains a large scope of variables but we limit the scope of the data to demographic profiles and wages data.
+To extract the relevant component of the data, we went to the NLSY79 database website at \url{https://www.nlsinfo.org/content/cohorts/nlsy79/get-data} and navigated as described below.
 
@@
 \faDatabase\ National Longitudinal Survey of Youth 1979 
-
 \begin{itemize}
@@
 
-According to @WickhamHadley2014TD, tidy data sets comply with three rules.
-Firstly, each variable forms a column.
-Secondly, each observation forms a row.
-Lastly, each type of observational unit forms a table.
-Unfortunately, the `new_data_qnames` did not meet these requirements in the way that data for each year and job for the same respondent is stored as one variable.
+According to @WickhamHadley2014TD, tidy data sets comply with three rules: (i) each variable forms a column, (ii) each observation forms a row, and (iii) each type of observational unit forms a table.
+The input data, `new_data_qnames`, is organised such that each row corresponds to an individual.
+As respondents can have multiple jobs at specific years, the column names, such as `HRP1_1980`, such as contain the information about the job number and the year.
+The input data consequently has a large number of columns (686 to be specific).
+
+[REMOVE] Unfortunately, the `new_data_qnames` did not meet these requirements in the way that data for each year and job for the same respondent is stored as one variable.
 Hence, the data contains a huge number of columns (686 columns).

### 2021-09-23 paper/paper.tex
tidy data section

@@
 
-\hypertarget{target-data}{%
-\subsection{Target Data}\label{target-data}}
+\hypertarget{target}{%
+\subsection{Target Data}\label{target}}
 
@@
 
-According to Wickham (2014), tidy data sets comply with three rules.
-Firstly, each variable forms a column.
-Secondly, each observation forms a row.
-Lastly, each type of observational unit forms a table.
-Unfortunately, the \texttt{new\_data\_qnames} did not meet these requirements in the way that data for each year and job for the same respondent is stored as one variable.
+According to Wickham (2014), tidy data sets comply with three rules: (i) each variable forms a column, (ii) each observation forms a row, and (iii) each type of observational unit forms a table.
+The input data, \texttt{new\_data\_qnames}, is organised such that each row corresponds to an individual.
+As respondents can have multiple jobs at specific years, the column names, such as \texttt{HRP1\_1980}, such as contain the information about the job number and the year.
+The input data consequently has a large number of columns (686 to be specific).
+
+{[}REMOVE{]} Unfortunately, the \texttt{new\_data\_qnames} did not meet these requirements in the way that data for each year and job for the same respondent is stored as one variable.
 Hence, the data contains a huge number of columns (686 columns).

### 2021-09-23 paper/paper-jsdse/paper-jsdse.tex
booktabs package added

@@
 \usepackage{fontawesome}
+\usepackage{booktabs}
 \usepackage{url} % not crucial - just used below for the URL
@@
 As respondents can have multiple jobs at specific years, the column names, such as \texttt{HRP1\_1980}, such as contain the information about the job number and the year.
-The input data consequently has a large number of columns ()
+The input data consequently has a large number of columns (686 to be specific).
 
-Unfortunately, the \texttt{new\_data\_qnames} did not meet these requirements in the way that data for each year and job for the same respondent is stored as one variable.
+{[}REMOVE{]} Unfortunately, the \texttt{new\_data\_qnames} did not meet these requirements in the way that data for each year and job for the same respondent is stored as one variable.
 Hence, the data contains a huge number of columns (686 columns).
@@
 
+As mentioned previously, the aim of this paper is to refresh the high school dropouts wages used in textbook (Singer and Willett 2003) as similar as possible.
+Hence, in this section, we will compare how close the \texttt{wages\_hs\_do}, henceforth refreshed data, to the textbook data.
+It is also worth noting that the textbook data is reused from Murnane, Willett, and Boudett (1999)'s paper.
+Hence, it is not directly extracted from the NLSY79 database.
+Singer and Willett (2003) do not state the time frame covered by the data.
+However, from Murnane, Willett, and Boudett (1999)'s paper, we find that the textbook data cover the survey period from 1979-1994.
+Hence, for comparison purposes, we filter the refreshed data only until 1994.
+
+In the absence of highs chool dropouts criteria in the textbook data, we create a clearer criteria of high school dropouts as mentioned in Section @ref(\#target).
+Consequently, we do not have the same ID of high-school dropouts as in the textbook data.
+The textbook data has 888 IDs included, while the refreshed data has 1184 IDs.
+Furthermore, there are only 94 IDs included both in the textbook and the refreshed data.
+After observing Murnane, Willett, and Boudett (1999)'s paper, it is mentioned that the data is randomly drawn from the low income subsample of the NLSY.
+This is probably the reason why, even using some criteria, we cannot match the dropouts in the textbook and the refreshed data.
+
+Further, the temporal variable used in the textbook data is experience, instead of year.
+The only explanation we found regarding this variable is that it is derived from the years of IDs' experience since their first job.
+Singer and Willett (2003) stated that this variable reflects the specific moment in each ID labor force history associated with each wage observation.
+Since it is not clearly articulated, we decide to keep the temporal variable in the refreshed data as year.
+Accordingly, in this comparison, we only compare the same ID but not the same temporal variable.
+
+Regarding the wages, the textbook data express the wages as inflation adjusted with 1990 dollars as constant (Singer and Willett 2003).
+However, it is not stated how exactly the wages are adjusted.
+Note that in the refreshed data we publish in \texttt{yowie}, we do not adjust the data with inflation because the time frame covered ranged from 1979-2018.
+Hence, we allow flexibility to the data user to adjust, not adjust, or adjust it with different base years.
+For comparison purposes, we adjust the wages with the average Consumer Price Index For All Urban Consumers (CPI-U) we got from (\textbf{fed?}).
+
+\begin{figure}
+
+{\centering \includegraphics[width=468px]{figures/plotting-sw-do-1}
+
+}
+
+\caption{Textbook data (A) and Refreshed data (B). The textbook data and refreshed data is not exactly the same because it has different temporal variables.}\label{fig:plotting-sw-do}
+\end{figure}
+
+Figure \ref{fig:plotting-sw-do} shows the comparison between the textbook and the refreshed data.
+Even though it has different temporal variables and possibly different CPI indicators used in inflation adjustment, we can still capture some of the patterns in the refreshed data.
+
+From this comparison, we learn that we cannot make an apple to apple comparison and produce the exact same data as the textbook due to the absence of clearly articulated data extraction and derivation process.
+Hence, the takeaway is that it is imperative to disclose how the data derived from the initial source to make data analysis reproducible.
+This is why in this paper, we show the practice of reproducible and responsible data cleaning workflow to make sure the longitudinal data from a continuous survey can be refreshed.
+
 \hypertarget{summary}{%
@@
 
+\leavevmode\hypertarget{ref-MurnaneRichardJ1999DMDB}{}%
+Murnane, Richard J, John B Willett, and Kathryn Parker Boudett.
+1999.
+{``Do Male Dropouts Benefit from Obtaining a GED, Postsecondary Education, and Training?''} \emph{Evaluation Review} 23 (5): 475--503.
+
 \leavevmode\hypertarget{ref-rsvg}{}%

### 2021-09-23 paper/paper-jsdse/paper-jsdse.tex
lots of changes

--- 2021-09-23 paper/paper-jsdse/paper-jsdse.tex
@@
   \author{
-            Dewi Amaliah\thanks{Test}\\
-      Department of Econometrics and Business Statistics, Monash University, \\
+            Dewi Amaliah\thanks{Corresponding author, \href{mailto:dlamaleeah@gmail.com}{\nolinkurl{dlamaleeah@gmail.com}}}\\
+      Dept of Econometrics and Business Statistics, Monash University, \\
             Dianne Cook\\
-      Department of Econometrics and Business Statistics, Monash University, \\
+      Dept of Econometrics and Business Statistics, Monash University, \\
             Emi Tanaka\\
-      Department of Econometrics and Business Statistics, Monash University, \\
+      Dept of Econometrics and Business Statistics, Monash University, \\
             Kate Hyde\\
-      Department of Econometrics and Business Statistics, Monash University, \\
+      Dept of Econometrics and Business Statistics, Monash University, \\
             Nicholas Tierney\\
-      , , }
+      Telethon Kids Institute, Perth, WA, }
         \maketitle
@@
 
-\begin{figure}
-
-{\centering \includegraphics[width=468px]{figures/untidy-data-1}
-
-}
-
-\caption{The missing value structure of the NLSY79 raw data.}\label{fig:untidy-data}
-\end{figure}
+\begin{verbatim}
+#> Rows: 12,686
+#> Columns: 11
+#> $ CASEID_1979 <int> 1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13, 14, 1~
+#> $ HRP1_1979   <int> 328, 385, 365, NA, 310, NA, NA, NA, 214, NA, 337~
+#> $ HRP2_1979   <int> NA, NA, NA, NA, 375, NA, NA, NA, NA, NA, 300, NA~
+#> $ HRP3_1979   <int> NA, NA, 275, NA, NA, NA, NA, NA, NA, NA, NA, NA,~
+#> $ HRP4_1979   <int> NA, NA, NA, NA, NA, 250, NA, NA, NA, NA, NA, NA,~
+#> $ HRP5_1979   <int> NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, ~
+#> $ HRP1_1980   <int> NA, 457, 397, NA, 333, 275, 300, 394, 200, 318, ~
+#> $ HRP2_1980   <int> NA, NA, 367, NA, NA, NA, NA, NA, NA, NA, NA, NA,~
+#> $ HRP3_1980   <int> NA, NA, 380, NA, NA, NA, 290, NA, NA, NA, NA, NA~
+#> $ HRP4_1980   <int> NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, ~
+#> $ HRP5_1980   <int> NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, ~
+\end{verbatim}
 
@@
 
-\begin{table}
-
-\caption{\label{tab:summarytable}Summary Statistics of Wages of High School Data}
-\centering
-\begin{tabular}[t]{lr}
-\toprule
-Statistics & Value\\
-\midrule
-Min. & 0.01000\\
-1st Qu. & 4.50000\\
-Median & 7.20000\\
-Mean & 11.86593\\
-3rd Qu. & 11.74000\\
-Max. & 33400.00000\\
-\bottomrule
-\end{tabular}
-\end{table}
-
 \begin{figure}
@@
 
-{\centering \includegraphics[width=0.93\linewidth]{/Users/etan0038/Dropbox/repo/summer-wages-refresh/paper/figures/flowchart}
+{\centering \includegraphics[width=0.93\linewidth]{/Users/cookd/numbats/summer-wages-refresh/paper/figures/flowchart}
 
@@
 \hypertarget{compare}{%
-\section{Comparison with the Textbook Data}\label{compare}}
+\section{Comparison of refreshed with the original data}\label{compare}}
+
+Now is the time to see how close we have come to the original textbook data.
+The original set contains wages of high school dropouts wages (Singer and Willett 2003) from 1979 through to 1994.
+The refreshed data set for comparison is also wages on high school drop outs, over 1979 to 2018.
+The original set is available in the R package \texttt{brolgar} and the refreshed data is available in the R package \texttt{yowie}.
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{figures/plotting-sw-do-1}
+
+}
+
+\caption{Comparison of original textbook example (A) with refreshed data (B).
+The original data was inflation-adjusted to 1990 prices and the individual's time of collection was converted to a length of experience in the workforce, which make it difficult to precisely compare the two sets.}\label{fig:plotting-sw-do}
+\end{figure}
+
+There are two aspects of the original data that make a direct comparison difficult.
+The time variable provided is ``experience in the workforce.'' It is not clear how this is calculated.
+One would expect that there is a record of the day the individual first started a job, and this is used to adjust the year of collection.
+We have not been able to find this information in the database.
+Secondly, wages were inflation-adjusted to 1990 prices, which is not done for the refreshed data.
+
+Figure \ref{fig:plotting-sw-do} shows the longitudinal profiles of the two sets.
+(A direct ID matching is not possible.) There are 888 individuals in the original and 1,188 individuals in the refreshed data.
+This suggests some individuals were removed in the original data.
+Overall, the profiles show some similarity.
 

### 2021-09-23 paper/paper.Rmd
lots of changes

@@
                fig.align = 'center')
-
 read_chunk(here::here("paper/scripts/data-preprocessing.R"))
@@
 read_chunk(here::here("paper/scripts/eda.R"))
+```
+
+```{r theme}
+theme_set(theme_bw(base_size = 18) +
+            theme(plot.background = element_rect(fill = 'transparent', colour = NA), axis.line.y = element_line(color = "black", linetype = "solid"),
+                  plot.title.position = "plot",
+                  plot.title = element_text(size = 24),
+                  panel.background  = element_rect(fill = 'transparent', colour = NA),
+                  legend.background = element_rect(fill = 'transparent', colour = NA),
+                  legend.key        = element_rect(fill = 'transparent', colour = NA)
+                  ) )
 ```
@@
 
-```{r untidy-data, fig.cap="The missing value structure of the NLSY79 raw data."}
+```{r untidy-data, fig.cap="The structure of the NLSY79 raw data on wages for up to five jobs in any year."}
 ```
@@
 
-```{r summarytable}
+```{r summarytable, eval=FALSE}
 ```
@@
 
-# Comparison with the Textbook Data {#compare}
+# Comparison of refreshed with the original data {#compare}
 
@@
 
-As mentioned previously, the aim of this paper is to refresh the high school dropouts wages used in textbook [@SingerJudithD2003Alda] as similar as possible.
-Hence, in this section, we will compare how close the `wages_hs_do`, henceforth refreshed data, to the textbook data.
+Now is the time to see how close we have come to the original textbook data.
+The original set contains wages of high school dropouts wages [@SingerJudithD2003Alda] from 1979 through to 1994.
+The refreshed data set for comparison is also wages on high school drop outs, over 1979 to 2018.
+The original set is available in the R package `brolgar` and the refreshed data is available in the R package `yowie`.
+
+```{r plotting-sw-do, fig.cap = "Comparison of original textbook example (A) with refreshed data (B).
+The original data was inflation-adjusted to 1990 prices and the individual's time of collection was converted to a length of experience in the workforce, which make it difficult to precisely compare the two sets.", fig.width = 9, fig.height = 5, out.width = "100%"}
+```
+
+There are two aspects of the original data that make a direct comparison difficult.
+The time variable provided is "experience in the workforce".
+It is not clear how this is calculated.
+One would expect that there is a record of the day the individual first started a job, and this is used to adjust the year of collection.
+We have not been able to find this information in the database.
+Secondly, wages were inflation-adjusted to 1990 prices, which is not done for the refreshed data.
+
+Figure \ref{fig:plotting-sw-do} shows the longitudinal profiles of the two sets.
+(A direct ID matching is not possible.) There are 888 individuals in the original and 1,188 individuals in the refreshed data.
+This suggests some individuals were removed in the original data.
+Overall, the profiles show some similarity.
+
+<!-- Hence, in this section, we will compare how close the `wages_hs_do`, henceforth refreshed data, to the textbook data.
 It is also worth noting that the textbook data is reused from @MurnaneRichardJ1999DMDB’s paper.
@@
 For comparison purposes, we adjust the wages with the average Consumer Price Index For All Urban Consumers (CPI-U) we got from @fed.
+-->
 
-```{r plotting-sw-do, fig.cap= "Textbook data (A) and Refreshed data (B). The textbook data and refreshed data is not exactly the same because it has different temporal variables."}
-```
-
-Figure \ref{fig:plotting-sw-do} shows the comparison between the textbook and the refreshed data.
+<!-- Figure \ref{fig:plotting-sw-do} shows the comparison between the textbook and the refreshed data.
 Even though it has different temporal variables and possibly different CPI indicators used in inflation adjustment, we can still capture some of the patterns in the refreshed data.
@@
 Hence, the takeaway is that it is imperative to disclose how the data derived from the initial source to make data analysis reproducible.
-This is why in this paper, we show the practice of reproducible and responsible data cleaning workflow to make sure the longitudinal data from a continuous survey can be refreshed.
+This is why in this paper, we show the practice of reproducible and responsible data cleaning workflow to make sure the longitudinal data from a continuous survey can be refreshed.-->
 

### 2021-09-23 paper/scripts/data-preprocessing.R

@@ -23,11 +23,13 @@ select <- dplyr::select
 source(here::here("data-raw/NLSY79/NLSY79.R"))
 
 ## ---- untidy-data
+options(width=70)
 new_data_qnames %>%
   select(CASEID_1979,
          starts_with("HRP") &
            ends_with(c("1979", "1980"))) %>%
-  visdat::vis_miss()
+  glimpse()
+  #visdat::vis_miss()
 
 ## ---- dob-tidy
 #tidy the date of birth data

### 2021-09-23 paper/scripts/eda.R

@@ -6,13 +6,13 @@ sw_wages <- sw %>%
   ggplot(aes(x = xp,
              y = ln_wages)) +
   geom_line(aes(group = id),alpha = 0.1) +
-  geom_smooth(se = FALSE) +
-  ggtitle("A") +
+  geom_smooth(se = FALSE, colour = "pink") +
+  ggtitle("A original") +
   theme_bw() +
-  ylab("ln(Hourly wage) ($)") +
-  xlab("Experience (years)") +
+  ylab("Hourly wage ($, natural log)") +
+  xlab("Experience in years") +
   theme(plot.title = element_text(size = 10)) +
-  ylim(-4, 6)
+  ylim(0, 5)
 
 ## ---- do_refreshed
 
@@ -32,13 +32,13 @@ do_ref <- do %>%
   ggplot(aes(x = year,
              y = lnwage)) +
   geom_line(aes(group = id), alpha = 0.1) +
-  geom_smooth(se = FALSE) +
-  ggtitle("B") +
+  geom_smooth(se = FALSE, colour = "pink") +
+  ggtitle("B refreshed data") +
   theme_bw() +
-  ylab("ln(Hourly wage) ($)") +
-  xlab("Year") +
+  ylab("Hourly wage ($, natural log)") +
+  xlab("Year of data collection") +
   theme(plot.title = element_text(size = 10)) +
-  ylim(-4, 6)
+  ylim(0, 5)
 
 # sw_wages + do_ref
 
@@ -64,6 +64,27 @@ sw_do <- inner_join(sw_id, do_id, by = "id")
 # only have 94 id that agree upon each other
 sw_agree <- filter(sw, id %in% sw_do$id)
 do_agree <- filter(do, id %in% sw_do$id)
+
+sw_do_sample <- sample(sw_do$id, 12)
+
+sw_agree_sample <- sw_agree %>%
+  filter(id %in% sw_do_sample) %>%
+  select(id, xp, ln_wages) %>%
+  as_tibble()
+do_agree_sample <- do_agree %>%
+  mutate(id = as.integer(as.character(id))) %>%
+  filter(id %in% sw_do_sample) %>%
+  select(id, year, lnwage) %>%
+  as_tibble()
+
+agree_ref <- ggplot() +
+  geom_line(data=do_agree_sample,
+            aes(x = year-1979,
+                y = lnwage), colour = "black") +
+  geom_line(data=sw_agree_sample,
[…]

### 2021-09-23 paper/scripts/ida.R

@@ -222,7 +222,9 @@ ggplot(wages_compare) +
   guides(linetype = FALSE) +
   xlab("Year") +
   ylab("Hourly wage ($)") +
-  scale_color_hue(labels = c("Before", "After"))
+  scale_colour_brewer("", palette = "Dark2", direction = -1,
+                      labels = c("Before", "After"))
+#  scale_color_hue(labels = c("Before", "After"))
 
 ## ---- fixed-feature-plot
 spag2 <- wages_cleaned %>%

### 2021-09-23 paper/paper-jsdse/paper-jsdse.tex
section headings

@@
 \hypertarget{target}{%
-\subsection{Target Data}\label{target}}
+\subsection{Target data}\label{target}}
 
@@
 \hypertarget{cleaning}{%
-\section{Data Cleaning}\label{cleaning}}
+\section{Data cleaning}\label{cleaning}}
 
@@
 \hypertarget{getdata}{%
-\subsection{Getting and Tidying the Data}\label{getdata}}
+\subsection{Getting and tidying the data}\label{getdata}}
 
@@
 \hypertarget{initial-data-analysis}{%
-\subsection{Initial Data Analysis}\label{initial-data-analysis}}
+\subsection{Initial data analysis}\label{initial-data-analysis}}
 
@@
   \textbf{shiny app}: An web interactive \texttt{shiny} app to run a simulation to customize the weight threshold.
-This app could be accessed \href{https://github.com/numbats/summer-wages-refresh/tree/main/app}{here}.
+This app could be accessed \href{https://github.com/numbats/wages-refresh/tree/main/app}{here}.
 \end{itemize}

### 2021-09-23 paper/paper.Rmd
section headings
tidy data section cleaned up

@@
 
-## Target Data {#target}
+## Target data {#target}
 
@@
 
-# Data Cleaning {#cleaning}
+# Data cleaning {#cleaning}
 
@@
 
-## Getting and Tidying the Data {#getdata}
+## Getting and tidying the data {#getdata}
 
@@
 
+The raw data, `new_data_qnames`, is organised such that each row corresponds to an individual.
+As respondents can have multiple jobs at specific years, the column names, such as `HRP1_1979`, `HRP2_1979`, `HRP1_1980` and `HRP2_1980`, contain the information about the job number up to 5 (`HRP1` = job 1, `HRP2` = job 2) and the year.
+The raw data consequently has a large number of columns (686 to be specific).
+The values in the cell under the variables that begin with `HRP` correspond to the hourly wage in dollars.
+A glimpse of this data output is shown below.
+
+```{r untidy-data, fig.cap="The structure of the NLSY79 raw data on wages for up to five jobs in any year."}
+```
+
 According to @WickhamHadley2014TD, tidy data sets comply with three rules: (i) each variable forms a column, (ii) each observation forms a row, and (iii) each type of observational unit forms a table.
-The input data, `new_data_qnames`, is organised such that each row corresponds to an individual.
-As respondents can have multiple jobs at specific years, the column names, such as `HRP1_1980`, such as contain the information about the job number and the year.
-The input data consequently has a large number of columns (686 to be specific).
+The raw data does not comply the rules of the tidy data, thus we wrangle the data in form required by making use of the `tidyverse` suite of packages [@tidyverse], which include the main set of packages needed, namely `tidyr` [@tidyr], `dplyr` [@dplyr], and `stringr` [@stringr].
+The tidy data form, where we have a variable that indicates the individual ID, year, job number, wage in dollars and demographic variables, is important for downstream analysis, such as visualisation and modelling that will be performed later.
 
@@
 We mainly used `tidyr` [@tidyr], `dplyr` [@dplyr], and `stringr` [@stringr] to do this job.
-
-```{r untidy-data, fig.cap="The structure of the NLSY79 raw data on wages for up to five jobs in any year."}
-```
 
@@
 
-## Initial Data Analysis
+## Initial data analysis
 
@@
 
-- **shiny app**: An web interactive `shiny` app to run a simulation to customize the weight threshold. This app could be accessed [here](https://github.com/numbats/summer-wages-refresh/tree/main/app).
+- **shiny app**: An web interactive `shiny` app to run a simulation to customize the weight threshold. This app could be accessed [here](https://github.com/numbats/wages-refresh/tree/main/app).
 

### 2021-09-23 paper/paper.tex
tidy data section cleaned up

@@
 
+The raw data, \texttt{new\_data\_qnames}, is organised such that each row corresponds to an individual.
+As respondents can have multiple jobs at specific years, the column names, such as \texttt{HRP1\_1979}, \texttt{HRP2\_1979}, \texttt{HRP1\_1980} and \texttt{HRP2\_1980}, contain the information about the job number up to 5 (\texttt{HRP1} = job 1, \texttt{HRP2} = job 2) and the year.
+The raw data consequently has a large number of columns (686 to be specific).
+The values in the cell under the variables that begin with \texttt{HRP} correspond to the hourly wage in dollars.
+A glimpse of this data output is shown below.
+
+\begin{verbatim}
+#> 'data.frame':    12686 obs. of  686 variables:
+#>  $ CASEID_1979                  : int  1 2 3 4 5 6 7 8 9 10 ...
+#>  $ HRP1_1979                    : int  328 385 365 NA 310 NA NA NA 214 NA ...
+#>  $ HRP2_1979                    : int  NA NA NA NA 375 NA NA NA NA NA ...
+#>  $ HRP3_1979                    : int  NA NA 275 NA NA NA NA NA NA NA ...
+#>  $ HRP4_1979                    : int  NA NA NA NA NA 250 NA NA NA NA ...
+#>  $ HRP5_1979                    : int  NA NA NA NA NA NA NA NA NA NA ...
+#>  $ HRP1_1980                    : int  NA 457 397 NA 333 275 300 394 200 318 ...
+#>  $ HRP2_1980                    : int  NA NA 367 NA NA NA NA NA NA NA ...
+#>  $ HRP3_1980                    : int  NA NA 380 NA NA NA 290 NA NA NA ...
+#>  $ HRP4_1980                    : int  NA NA NA NA NA NA NA NA NA NA ...
+#>   [list output truncated]
+\end{verbatim}
+
 According to Wickham (2014), tidy data sets comply with three rules: (i) each variable forms a column, (ii) each observation forms a row, and (iii) each type of observational unit forms a table.
-The input data, \texttt{new\_data\_qnames}, is organised such that each row corresponds to an individual.
-As respondents can have multiple jobs at specific years, the column names, such as \texttt{HRP1\_1980}, such as contain the information about the job number and the year.
-The input data consequently has a large number of columns (686 to be specific).
+The raw data does not comply the rules of the tidy data, thus we wrangle the data in form required by making use of the \texttt{tidyverse} suite of packages (Wickham et al.
+2019), which include the main set of packages needed, namely \texttt{tidyr} (Wickham 2020b), \texttt{dplyr} (Wickham et al.
+2020), and \texttt{stringr} (Wickham 2019).
+The tidy data form, where we have a variable that indicates the individual ID, year, job number, wage in dollars and demographic variables, is important for downstream analysis, such as visualisation and modelling that will be performed later.
 
@@
 2020), and \texttt{stringr} (Wickham 2019) to do this job.
-
-\begin{figure}
-
-{\centering \includegraphics[width=468px]{figures/untidy-data-1}
-
-}
-
-\caption{The missing value structure of the NLSY79 raw data.}\label{fig:untidy-data}
-\end{figure}
 
@@
 The next step is that explore the mean hourly wage data. In this case, we only explore the wages data of up to \(12^{th}\) grade. Here, we use visualisation techniques to do IDA.
-
-\begin{table}
-
-\caption{\label{tab:summarytable}Summary Statistics of Wages of High School Data}
-\centering
-\begin{tabular}[t]{lr}
-\toprule
-Statistics & Value\\
-\midrule
-Min. & 0.01000\\
-1st Qu. & 4.50000\\
-Median & 7.20000\\
-Mean & 11.86593\\
-3rd Qu. & 11.74000\\
-Max. & 33400.00000\\
-\bottomrule
-\end{tabular}
-\end{table}
 

### 2021-09-23 paper/paper.Rmd
Modify flowchart.

--- 2021-09-23 paper/paper.Rmd
@@
 The spikes are still observed but not as extreme as the original data set.
-In Figure \ref{fixed-feature-plot} B, we plot the three features of mean hourly wages, namely the minimum, median, and maximum value, transformed to log scale.
-The plot implies that the skewness is slightly negative.
-We also see that the three features are overlapped each other.
-It indicates that some ID's minimum wages are higher than some ID's maximum wages.
+In Figure \ref{fig:fixed-feature-plot} B, we plot the three features of mean hourly wages, namely the minimum, median, and maximum value.
+We still can see some extreme values in maximum wages, but consider it as a natural variability of the data.
+In \ref{fig:fixed-feature-plot} C, we can see that after imputing the extreme value in ID=39, we can see how the the wages change over the years more clearly.
 
@@
 
-```{r flow-chart, fig.cap="The stages of data filtering from the raw data to get three datasets contained yowie, n means the number of ID, while n\\_obs means the number of observations.", out.width="93%"}
+```{r flow-chart, fig.cap="The stages of data cleaning from the raw data to get three datasets contained yowie.
+n individuals means the number of respondents included in each stage, while n observations means the number of rows in the data.
+The color represented the stage of data cleaning in statistical value chain (van der Loo \\& de Jonge, 2018).
+Pink, blue, and green represents the raw, input, and valid data, respectively.", out.width="93%"}
 ```

### 2021-09-23 paper/paper.tex
Modify flowchart.

--- 2021-09-23 paper/paper.tex
@@
 Consequently, the data should be tidied and wrangled first to extract the demographic and employment variables we want to put in the final data set.
-We mainly used \texttt{tidyr} (Wickham 2020b), \texttt{dplyr} (Wickham et al.
-2020), and \texttt{stringr} (Wickham 2019) to do this job.
+We mainly used \texttt{tidyr} (H.
+Wickham 2020b), \texttt{dplyr} (H.
+Wickham, François, et al.
+2020), and \texttt{stringr} (H.
+Wickham 2019) to do this job.
 
-\begin{figure}
-
-{\centering \includegraphics[width=468px]{figures/untidy-data-1}
-
-}
-
-\caption{The missing value structure of the NLSY79 raw data.}\label{fig:untidy-data}
-\end{figure}
+\begin{verbatim}
+#> Rows: 12,686
+#> Columns: 11
+#> $ CASEID_1979 <int> 1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13, 14, 1~
+#> $ HRP1_1979   <int> 328, 385, 365, NA, 310, NA, NA, NA, 214, NA, 337~
+#> $ HRP2_1979   <int> NA, NA, NA, NA, 375, NA, NA, NA, NA, NA, 300, NA~
+#> $ HRP3_1979   <int> NA, NA, 275, NA, NA, NA, NA, NA, NA, NA, NA, NA,~
+#> $ HRP4_1979   <int> NA, NA, NA, NA, NA, 250, NA, NA, NA, NA, NA, NA,~
+#> $ HRP5_1979   <int> NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, ~
+#> $ HRP1_1980   <int> NA, 457, 397, NA, 333, 275, 300, 394, 200, 318, ~
+#> $ HRP2_1980   <int> NA, NA, 367, NA, NA, NA, NA, NA, NA, NA, NA, NA,~
+#> $ HRP3_1980   <int> NA, NA, 380, NA, NA, NA, 290, NA, NA, NA, NA, NA~
+#> $ HRP4_1980   <int> NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, ~
+#> $ HRP5_1980   <int> NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, ~
+\end{verbatim}
 
@@
 \hypertarget{initial-data-analysis}{%
-\subsection{Initial Data Analysis}\label{initial-data-analysis}}
+\subsection{Initial data analysis}\label{initial-data-analysis}}
 
-According to Huebner Marianne, Vach, and Cessie (2016), Initial Data Analysis (IDA) is the step of inspecting and screening the data after being collected to ensure that the data is clean, valid, and ready to be deployed in the later formal statistical analysis.
+According to Huebner, Vach, and Cessie (2016), Initial Data Analysis (IDA) is the step of inspecting and screening the data after being collected to ensure that the data is clean, valid, and ready to be deployed in the later formal statistical analysis.
 Moreover, Chatfield (1985) argues that the two main objectives of IDA are data description, which is to assess the structure and the quality of the data, and model formulation without any formal statistical inference.
@@
 In this section, we have demonstrate a workflow from getting the data from the NSLY79 database until a ready-analysed data.
-Referring to Loo (2018), this section shows part of statistical value chain, which is data cleaning from raw, input, and valid data.
+Referring to Loo and Jonge (2018), this section shows part of statistical value chain, which is data cleaning from raw, input, and valid data.
 The raw demographic and employment data from the database are not in a tidy format.
@@
 
-\caption{The stages of data filtering from the raw data to get three datasets contained yowie, n means the number of ID, while n\_obs means the number of observations.}\label{fig:flow-chart}
+\caption{The stages of data cleaning from the raw data to get three datasets contained yowie.
+n individuals means the number of respondents included in each stage, while n observations means the number of rows in the data.
+The color represented the stage of data cleaning in statistical value chain (van der Loo \& de Jonge, 2018).
+Pink, blue, and green represents the raw, input, and valid data, respectively.}\label{fig:flow-chart}
 \end{figure}
@@
 There are two aspects of the original data that make a direct comparison difficult.
-The time variable provided is ``experience in the workforce.'' It is not clear how this is calculated.
+The time variable provided is ``experience in the workforce''.
+It is not clear how this is calculated.
 One would expect that there is a record of the day the individual first started a job, and this is used to adjust the year of collection.
@@
 \leavevmode\hypertarget{ref-KollerManuel2016rARP}{}%
-Koller, Manuel. 2016. {``Robustlmm: An r Package for Robust Estimation of Linear Mixed-Effects Models.''} \emph{Journal of Statistical Software} 75 (6): 1--24.
+Koller, Manuel. 2016. ``Robustlmm: An R Package for Robust Estimation of Linear Mixed-Effects Models.'' \emph{Journal of Statistical Software} 75 (6): 1--24.
 
 \leavevmode\hypertarget{ref-LooMarkvander2018Sdcw}{}%
-Loo, Mark van der. 2018. \emph{Statistical Data Cleaning with Applications in r}.
+Loo, Mark van der, and Edwin de Jonge. 2018. \emph{Statistical Data Cleaning with Applications in R}.
 
 \leavevmode\hypertarget{ref-rsvg}{}%
-Ooms, Jeroen. 2020. \emph{Rsvg: Render SVG Images into PDF, PNG, PostScript, or Bitmap Arrays}. \url{https://CRAN.R-project.org/package=rsvg}.
+Ooms, Jeroen. 2020. \emph{Rsvg: Render Svg Images into Pdf, Png, Postscript, or Bitmap Arrays}. \url{https://CRAN.R-project.org/package=rsvg}.
 
 \leavevmode\hypertarget{ref-opendata}{}%
-Open Knowledge Foundation. 2021. {``Open Definition. Defining Open in Open Data, Open Content, and Open Knowledge.''} 2021. \url{http://opendefinition.org/od/2.1/en/}.
+Open Knowledge Foundation. 2021. ``Open Definition. Defining Open in Open Data, Open Content, and Open Knowledge.'' 2021. \url{http://opendefinition.org/od/2.1/en/}.
 

### 2021-09-23 paper/paper.Rmd
statistical value chain described in data cleaning process

@@
 @HuebnerMariannePhD2016Asat emphasize that making the data cleaning process accountable and transparent is imperative and essential for the integrity of downstream statistical analyses and model building [@HuebnerMarianne2020Haar].
-Data cleaning can be considered to be a part of what is called "initial data analysis (IDA)" [@Chatfield1985TIEo].
+Data cleaning can be considered to be a part of what is called "initial data analysis" (IDA) [@Chatfield1985TIEo].
 In IDA one would also explore the data, especially to check if the data is consistent with assumptions required for modeling.
@@
 
-The NLSY79 is a longitudinal survey held by the U.S Bureau of Labor Statistics that follows the lives of a sample of American youth and born between 1957-1964 [@nlsy79].
+The NLSY79 is a longitudinal survey held by the U.S Bureau of Labor Statistics that follows the lives of a sample of American youth born between 1957-1964 [@nlsy79].
 The cohort originally included 12,686 respondents aged 14-22 when first interviewed in 1979.
-It was comprised of Blacks, Hispanics, economically disadvantaged non-Black non-Hispanics, and youth in the military.
+It comprised of Blacks, Hispanics, economically disadvantaged non-Black non-Hispanics, and youth in the military.
 In 1984 and 1990, two sub-samples were dropped from the interview; the dropped subjects are the 1,079 members of the military sample and 1,643 members of the economically disadvantaged non-Black non-Hispanics, respectively.
@@
 The surveys were conducted annually from 1979 to 1994 and biennially thereafter.
-Data are now available from Round 1 (1979 survey year) to Round 28 (2018 survey year).
+Data are currently available from Round 1 (1979 survey year) to Round 28 (2018 survey year).
 
@@
 
-In this section, we outline the steps taken to the download the raw data from the database provided at \url{https://www.nlsinfo.org/content/cohorts/nlsy79/get-data} [@nlsy79].
-Specifically, we only extract a subset of the variables in the database as described in Section~\ref{getdata}.
+@LooMarkvander2018Sdcw describe the notion of a "statistical value chain" where the production stages of data cleaning process are earmarked as raw data (data from the data source), input data (data organised with correct type and identified variables) and valid data (data that faithfully represent the variables).
+In this section, we outline the steps to download the raw data (Section \@ref(getdata)) and then tidying the raw data into input data, specifically for wages data (Section \@ref(tidywages)), demographic variables (Section \@ref(tidydemog)) and the employment variables (Section \@ref(tidyemp)), so that the resulting input data can be used downstream for validating the data as described in Section \@ref(ida).
 
-## Getting and tidying the data {#getdata}
+## Getting the data {#getdata}
 
-The NLSY79 data contains a large scope of variables but we limit the scope of the data to demographic profiles and wages data.
-To extract the relevant component of the data, we went to the NLSY79 database website at \url{https://www.nlsinfo.org/content/cohorts/nlsy79/get-data} and navigated as described below.
+The NLSY79 data contains a large scope of variables but for our aim, the scope required is limisted to demographic profiles and wages data.
+More specifically, we went to the NLSY79 database website at \url{https://www.nlsinfo.org/content/cohorts/nlsy79/get-data} and navigated as described below.
 
@@
 \begin{itemize}
-\item[\faCloudDownload] Downloaded all of the variables in Hours worked.
-\item[\faCloudDownload] Downloaded all of the variables in Hourly wages
+\item[\faCloudDownload] We downloaded all of the variables in Hours worked.
+\item[\faCloudDownload] We downloaded all of the variables in Hourly wages
   Both hours worked and hourly wages are recorded by the job, up to five jobs for each id/subject. 
@@
 
-- `NLSY79.csv`: csv format 
-- `NLSY79.dat`: dat format
+- `NLSY79.csv`: csv format of the data
+- `NLSY79.dat`: dat format of the data
 - `NLSY79.NLSY79`: Tagset of variables that can be uploaded to the website to recreate the data set.
@@
 ```
+
+### Tidying the wages data {#tidywages}
 
@@
 
-### Tidying demographic variables
+### Tidying demographic variables {#tidydemog}
 
@@
 
-### Tidying employment variables
+### Tidying employment variables {#tidyemp}
 
@@
 
-## Initial data analysis
+## Initial data analysis {#ida}
 
-According to @HuebnerMariannePhD2016Asat, Initial Data Analysis (IDA) is the step of inspecting and screening the data after being collected to ensure that the data is clean, valid, and ready to be deployed in the later formal statistical analysis.
+According to @HuebnerMariannePhD2016Asat, initial data analysis (IDA) is the step of inspecting and screening the data after being collected to ensure that the data is clean, valid, and ready to be deployed in the later formal statistical analysis.
 Moreover, @Chatfield1985TIEo argues that the two main objectives of IDA are data description, which is to assess the structure and the quality of the data, and model formulation without any formal statistical inference.
@@
 In addition, we also aim to find the anomaly in the wages values using this approach.
-We mainly use graphical summaries to do the IDA using `ggplot2`[@ggplot2] and `brolgar` [@brolgar].
+We mainly use graphical summaries to do the IDA using `ggplot2` [@ggplot2] and `brolgar` [@brolgar].
 

### 2021-09-23 paper/paper.tex
statistical value chain described in data cleaning process

--- 2021-09-23 paper/paper.tex
@@
 2020).
-Data cleaning can be considered to be a part of what is called ``initial data analysis (IDA)'' (Chatfield 1985).
+Data cleaning can be considered to be a part of what is called ``initial data analysis'' (IDA) (Chatfield 1985).
 In IDA one would also explore the data, especially to check if the data is consistent with assumptions required for modeling.
@@
 
-The NLSY79 is a longitudinal survey held by the U.S Bureau of Labor Statistics that follows the lives of a sample of American youth and born between 1957-1964 (The U.S.
+The NLSY79 is a longitudinal survey held by the U.S Bureau of Labor Statistics that follows the lives of a sample of American youth born between 1957-1964 (The U.S.
 Bureau of Labor Statistics, n.d.).
 The cohort originally included 12,686 respondents aged 14-22 when first interviewed in 1979.
-It was comprised of Blacks, Hispanics, economically disadvantaged non-Black non-Hispanics, and youth in the military.
+It comprised of Blacks, Hispanics, economically disadvantaged non-Black non-Hispanics, and youth in the military.
 In 1984 and 1990, two sub-samples were dropped from the interview; the dropped subjects are the 1,079 members of the military sample and 1,643 members of the economically disadvantaged non-Black non-Hispanics, respectively.
@@
 
-In this section, we outline the steps taken to the download the raw data from the database provided at \url{https://www.nlsinfo.org/content/cohorts/nlsy79/get-data} (The U.S.
-Bureau of Labor Statistics, n.d.).
-Specifically, we only extract a subset of the variables in the database as described in Section\textasciitilde{}\ref{getdata}.
+Loo and Jonge (2018) describe the notion of a ``statistical value chain'' where the production stages of data cleaning process are earmarked as raw data (data from the data source), input data (data organised with correct type and identified variables) and valid data (data that faithfully represent the variables).
+In this section, we outline the steps to download the raw data (Section \ref{getdata}) and then tidying the raw data into input data, specifically for wages data (Section \ref{tidywages}), demographic variables (Section \ref{tidydemog}) and the employment variables (Section \ref{tidyemp}), so that the resulting input data can be used downstream for validating the data as described in Section \ref{ida}.
 
 \hypertarget{getdata}{%
-\subsection{Getting and tidying the data}\label{getdata}}
+\subsection{Getting the data}\label{getdata}}
 
-The NLSY79 data contains a large scope of variables but we limit the scope of the data to demographic profiles and wages data.
-To extract the relevant component of the data, we went to the NLSY79 database website at \url{https://www.nlsinfo.org/content/cohorts/nlsy79/get-data} and navigated as described below.
+The NLSY79 data contains a large scope of variables but for our aim, the scope required is limisted to demographic profiles and wages data.
+More specifically, we went to the NLSY79 database website at \url{https://www.nlsinfo.org/content/cohorts/nlsy79/get-data} and navigated as described below.
 
@@
 \item
-  \texttt{NLSY79.csv}: csv format
+  \texttt{NLSY79.csv}: csv format of the data
 \item
-  \texttt{NLSY79.dat}: dat format
+  \texttt{NLSY79.dat}: dat format of the data
 \item
@@
 \end{itemize}
+
+\hypertarget{tidywages}{%
+\subsubsection{Tidying the wages data}\label{tidywages}}
 
@@
 
-\hypertarget{tidying-demographic-variables}{%
-\subsubsection{Tidying demographic variables}\label{tidying-demographic-variables}}
+\hypertarget{tidydemog}{%
+\subsubsection{Tidying demographic variables}\label{tidydemog}}
 
@@
 
-\hypertarget{tidying-employment-variables}{%
-\subsubsection{Tidying employment variables}\label{tidying-employment-variables}}
+\hypertarget{tidyemp}{%
+\subsubsection{Tidying employment variables}\label{tidyemp}}
 
@@
 
-\hypertarget{initial-data-analysis}{%
-\subsection{Initial data analysis}\label{initial-data-analysis}}
+\hypertarget{ida}{%
+\subsection{Initial data analysis}\label{ida}}
 
-According to Huebner Marianne, Vach, and Cessie (2016), Initial Data Analysis (IDA) is the step of inspecting and screening the data after being collected to ensure that the data is clean, valid, and ready to be deployed in the later formal statistical analysis.
+According to Huebner Marianne, Vach, and Cessie (2016), initial data analysis (IDA) is the step of inspecting and screening the data after being collected to ensure that the data is clean, valid, and ready to be deployed in the later formal statistical analysis.
 Moreover, Chatfield (1985) argues that the two main objectives of IDA are data description, which is to assess the structure and the quality of the data, and model formulation without any formal statistical inference.
@@
 In addition, we also aim to find the anomaly in the wages values using this approach.
-We mainly use graphical summaries to do the IDA using \texttt{ggplot2}(Wickham 2016) and \texttt{brolgar} (Tierney, Cook, and Prvan 2020).
+We mainly use graphical summaries to do the IDA using \texttt{ggplot2} (Wickham 2016) and \texttt{brolgar} (Tierney, Cook, and Prvan 2020).
 
@@
 \leavevmode\hypertarget{ref-LooMarkvander2018Sdcw}{}%
-Loo, Mark van der. 2018. \emph{Statistical Data Cleaning with Applications in r}.
+Loo, Mark van der, and Edwin de Jonge. 2018. \emph{Statistical Data Cleaning with Applications in r}.
 

### 2021-09-24 paper/paper.tex
cleaned up tables + its captions

@@
 
-\caption{\label{tab:age-table}Age Distribution of the NLSY79 samples}
+\caption{\label{tab:age-table}The frequency table of the age at the start of the survey (?CHECK) in the full NSLY79 data}
 \centering
@@
 \toprule
-Age & Number of Sample\\
+Age & Number of individuals\\
 \midrule
-15 & 1265\\
-16 & 1550\\
-17 & 1600\\
-18 & 1530\\
-19 & 1662\\
-20 & 1722\\
-21 & 1677\\
-22 & 1680\\
+15 & 1,265\\
+16 & 1,550\\
+17 & 1,600\\
+18 & 1,530\\
+19 & 1,662\\
+20 & 1,722\\
+21 & 1,677\\
+22 & 1,680\\
 \bottomrule
@@
 
-\caption{\label{tab:gender-race-table}Gender and Race Distribution of the NLSY79 Samples}
+\caption{\label{tab:gender-race-table}The contingency table for gender and race for the full NLSY79 data.}
 \centering
-\begin{tabular}[t]{lllll}
+\begin{tabular}[t]{lrrrr}
 \toprule
@@
 \midrule
-Male & 1000 (15.62\%) & 1613 (25.19\%) & 3790 (59.19\%) & 6403 (100.00\%)\\
-Female & 1002 (15.95\%) & 1561 (24.84\%) & 3720 (59.21\%) & 6283 (100.00\%)\\
-Total & 2002 (15.78\%) & 3174 (25.02\%) & 7510 (59.20\%) & 12686 (100.00\%)\\
+Male & 1,000 (15.62\%) & 1,613 (25.19\%) & 3,790 (59.19\%) & 6,403 (100.00\%)\\
+Female & 1,002 (15.95\%) & 1,561 (24.84\%) & 3,720 (59.21\%) & 6,283 (100.00\%)\\
+\midrule
+Total & 2,002 (15.78\%) & 3,174 (25.02\%) & 7,510 (59.20\%) & 12,686 (100.00\%)\\
 \bottomrule

### 2021-09-24 paper/scripts/ida.R

@@ -3,34 +3,46 @@ age_table <- full_demographics %>%
   group_by(age_1979) %>%
   count(age_1979)
 
-kable(age_table,
-      caption = "Age Distribution of the NLSY79 samples",
-      col.names = c("Age", "Number of Sample"),
+age_table %>%
+  mutate(n = scales::comma(n)) %>%
+  kable(
+      caption = "The frequency table of the age at the start of the survey (?CHECK) in the full NSLY79 data",
+      col.names = c("Age", "Number of individuals"),
       booktabs = TRUE,
-      linesep = "") %>%
+      linesep = "",
+      align = "r") %>%
   kable_styling()
 
 ## ---- gender-race-table
+#' @param x a tabyl with frequency table in "core" attribute
+freq_formatting <- function(x) {
+  attr(x, "core") %>%
+    adorn_totals(c("row", "col")) %>%
+    mutate(across(where(is.numeric), ~scales::comma(.x, 1)))
+}
+
 gender_race_table <- full_demographics %>%
   tabyl(gender, race) %>%
   adorn_totals(c("row", "col")) %>%
   adorn_percentages("row") %>%
   adorn_pct_formatting(digits = 2) %>%
-  adorn_ns(position = "front") %>%
+  adorn_ns(., position = "front", ns = freq_formatting(.)) %>%
   mutate(gender = str_to_title(gender))
 
 kable(gender_race_table,
-      caption = "Gender and Race Distribution of the NLSY79 Samples",
+      caption = "The contingency table for gender and race for the full NLSY79 data.",
       col.names = c("Gender", "Hispanic", "Black", "Non-Black, Non-Hispanic", "Total"),
       booktabs = TRUE,
-      linesep = "") %>%
-  kable_styling() %>%
+      linesep = "",
+      align = "lrrrr") %>%
+  kable_classic() %>%
+  row_spec(2, hline_after = TRUE) %>%
   add_header_above(c(" " = 1, "Race" = 3, " " = 1))
 
 ## ---- summarytable
 kable(as.array(summary(wages_before$mean_hourly_wage)),
-      caption = "Summary Statistics of Wages of High School Data",
-      col.names = c("Statistics", "Value"),
+      caption = "Summary statistics of the hourly wages from the high school dropout subcohort of NLSY79 data",
+      col.names = c("Statistics", "Hourly wage ($)"),
       booktabs = TRUE,
       linesep = "") %>%
   kable_styling()

### 2021-09-25 paper/paper-jsdse/paper-jsdse.tex
tweaking plots and captions

@@
 
+\hypertarget{tidywages}{%
+\subsubsection{Tidying the wages data}\label{tidywages}}
+
+The raw data, \texttt{new\_data\_qnames}, is organised such that each row corresponds to an individual.
+As respondents can have multiple jobs at specific years, the column names, such as \texttt{HRP1\_1979}, \texttt{HRP2\_1979}, \texttt{HRP1\_1980} and \texttt{HRP2\_1980}, contain the information about the job number up to 5 (\texttt{HRP1} = job 1, \texttt{HRP2} = job 2) and the year.
+The raw data consequently has a large number of columns (686 to be specific).
+The values in the cell under the variables that begin with \texttt{HRP} correspond to the hourly wage in dollars.
+A glimpse of this data output is shown below.
+
+\begin{verbatim}
+#> 'data.frame':    12686 obs. of  686 variables:
+#>  $ CASEID_1979                  : int  1 2 3 4 5 6 7 8 9 10 ...
+#>  $ HRP1_1979                    : int  328 385 365 NA 310 NA NA NA 214 NA ...
+#>  $ HRP2_1979                    : int  NA NA NA NA 375 NA NA NA NA NA ...
+#>  $ HRP3_1979                    : int  NA NA 275 NA NA NA NA NA NA NA ...
+#>  $ HRP4_1979                    : int  NA NA NA NA NA 250 NA NA NA NA ...
+#>  $ HRP5_1979                    : int  NA NA NA NA NA NA NA NA NA NA ...
+#>  $ HRP1_1980                    : int  NA 457 397 NA 333 275 300 394 200 318 ...
+#>  $ HRP2_1980                    : int  NA NA 367 NA NA NA NA NA NA NA ...
+#>  $ HRP3_1980                    : int  NA NA 380 NA NA NA 290 NA NA NA ...
+#>  $ HRP4_1980                    : int  NA NA NA NA NA NA NA NA NA NA ...
+#>   [list output truncated]
+\end{verbatim}
+
 According to Wickham (2014), tidy data sets comply with three rules: (i) each variable forms a column, (ii) each observation forms a row, and (iii) each type of observational unit forms a table.
-The input data, \texttt{new\_data\_qnames}, is organised such that each row corresponds to an individual.
-As respondents can have multiple jobs at specific years, the column names, such as \texttt{HRP1\_1980}, such as contain the information about the job number and the year.
-The input data consequently has a large number of columns (686 to be specific).
+The raw data does not comply the rules of the tidy data, thus we wrangle the data in form required by making use of the \texttt{tidyverse} suite of packages (Wickham et al.
+2019), which include the main set of packages needed, namely \texttt{tidyr} (Wickham 2020b), \texttt{dplyr} (Wickham et al.
+2020), and \texttt{stringr} (Wickham 2019).
+The tidy data form, where we have a variable that indicates the individual ID, year, job number, wage in dollars and demographic variables, is important for downstream analysis, such as visualisation and modelling that will be performed later.
 
@@
 
-\begin{verbatim}
-#> Rows: 12,686
-#> Columns: 11
-#> $ CASEID_1979 <int> 1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13, 14, 1~
-#> $ HRP1_1979   <int> 328, 385, 365, NA, 310, NA, NA, NA, 214, NA, 337~
-#> $ HRP2_1979   <int> NA, NA, NA, NA, 375, NA, NA, NA, NA, NA, 300, NA~
-#> $ HRP3_1979   <int> NA, NA, 275, NA, NA, NA, NA, NA, NA, NA, NA, NA,~
-#> $ HRP4_1979   <int> NA, NA, NA, NA, NA, 250, NA, NA, NA, NA, NA, NA,~
-#> $ HRP5_1979   <int> NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, ~
-#> $ HRP1_1980   <int> NA, 457, 397, NA, 333, 275, 300, 394, 200, 318, ~
-#> $ HRP2_1980   <int> NA, NA, 367, NA, NA, NA, NA, NA, NA, NA, NA, NA,~
-#> $ HRP3_1980   <int> NA, NA, 380, NA, NA, NA, 290, NA, NA, NA, NA, NA~
-#> $ HRP4_1980   <int> NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, ~
-#> $ HRP5_1980   <int> NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, ~
-\end{verbatim}
-
-\hypertarget{tidying-demographic-variables}{%
-\subsubsection{Tidying demographic variables}\label{tidying-demographic-variables}}
+\hypertarget{tidydemog}{%
+\subsubsection{Tidying demographic variables}\label{tidydemog}}
 
@@
 
-\caption{A glimpse of the refreshed data.
-Longitudinal profiles of wages from 1979-2018 are shown for a random sample of 36 individuals, divided into twelve subplots.
-Most individuals here experienced increasing wages over time, although there is considerable fluctuation from one year to another.}\label{fig:sample-plot}
+\caption{Longitudinal profiles of wages for a random sample of 20 individuals in the refreshed data.
+Most, but not all, individuals here experienced increasing wages over time, and several have experienced considerable fluctuation in wages.
+Some individuals are only measured for a short period.}\label{fig:sample-plot}
 \end{figure}
@@
 
-{\centering \includegraphics[width=432px]{figures/feature-plot-1}
+{\centering \includegraphics[width=1\linewidth]{figures/feature-plot-1}
 
@@
 
-{\centering \includegraphics[width=432px]{figures/compare-plot-1}
+{\centering \includegraphics[width=1\linewidth]{figures/compare-plot-1}
 
@@
 
-\caption{Comparison between the original and the treated mean hourly wage.
-The orange line portray the original value of mean hourly wage, while the turquoise line display the mean hourly wages value after the extreme values imputed with the robust linear model's prediction value.
-We can see that some extreme spikes has been reduced by the model.}\label{fig:compare-plot}
+\caption{Comparison between the original (black dots) and the corrected (solid grey) mean hourly wage for a sample of individuals.
+A robust linear model prediction was used to correct mean hourly wages value.
+We can see that some extreme spikes, corresponding to implausible wages, have been repleaced with values more similar to wages in neighboring years, but otherwise the profiles are not changed.}\label{fig:compare-plot}
 \end{figure}
@@
 
-{\centering \includegraphics[width=432px]{figures/fixed-feature-plot-1}
+{\centering \includegraphics[width=1\linewidth]{figures/fixed-feature-plot-1}
 
@@
 
-{\centering \includegraphics[width=0.93\linewidth]{/Users/cookd/numbats/summer-wages-refresh/paper/figures/flowchart}
+{\centering \includegraphics[width=0.85\linewidth]{/Users/cookd/numbats/summer-wages-refresh/paper/figures/flowchart}
 

### 2021-09-25 paper/paper.Rmd
tweaking plots and captions

@@
 
-```{r sample-plot, echo = FALSE, fig.cap="A glimpse of the refreshed data.
-Longitudinal profiles of wages from 1979-2018 are shown for a random sample of 36 individuals, divided into twelve subplots.
-Most individuals here experienced increasing wages over time, although there is considerable fluctuation from one year to another.", fig.height=4, fig.width=5, out.width="100%"}
+```{r sample-plot, echo = FALSE, fig.cap="Longitudinal profiles of wages for a random sample of 20 individuals in the refreshed data.
+Most, but not all, individuals here experienced increasing wages over time, and several have experienced considerable fluctuation in wages.
+Some individuals are only measured for a short period.", fig.height=8, fig.width=10, out.width="100%"}
 ```
@@
 ```{r feature-plot, fig.cap = "Summary plots to check the cleaned data reveal more cleaning is necessary: longitudinal profiles of wages for all individuals 1979-2018 (A), boxplots of minimum, median, and maximum wages of each individual (B), and one individual (id=39) with an unusual wage relative to their years of data (C).
-Some values of hourly wages are unbelievable, and some individuals have extemely unusual wages in some years.", fig.width=6, fig.height=2}
+Some values of hourly wages are unbelievable, and some individuals have extemely unusual wages in some years.", fig.width=10, fig.height=4, out.width="100%"}
 ```
@@
 
-```{r compare-plot, fig.cap="Comparison between the original and the treated mean hourly wage.
-The orange line portray the original value of mean hourly wage, while the turquoise line display the mean hourly wages value after the extreme values imputed with the robust linear model's prediction value.
-We can see that some extreme spikes has been reduced by the model.", fig.height=5, fig.width=6, outwidth="100%"}
+```{r compare-plot, fig.cap="Comparison between the original (black dots) and the corrected (solid grey) mean hourly wage for a sample of individuals.
+A robust linear model prediction was used to correct mean hourly wages value.
+We can see that some extreme spikes, corresponding to implausible wages, have been repleaced with values more similar to wages in neighboring years, but otherwise the profiles are not changed.", fig.height=8, fig.width=10, out.width="100%"}
 ```
@@
 ```{r fixed-feature-plot, fig.cap = "Re-make of the summary plots of the fully processed data suggest that it is now in a reasonable state: longitudinal profiles of wages for all individuals 1979-2018 (A), boxplots of minimum, median, and maximum wages of each individual (B), and one individual with an unusual wage relative to their years of data (C).
-", fig.width=6, fig.height=2}
+", fig.width=10, fig.height=4, out.width="100%"}
 ```
@@
 The color represented the stage of data cleaning in statistical value chain (van der Loo \\& de Jonge, 2018).
-Pink, blue, and green represents the raw, input, and valid data, respectively.", out.width="93%"}
+Pink, blue, and green represents the raw, input, and valid data, respectively.", out.width="85%"}
 ```

### 2021-09-25 paper/scripts/ida.R

@@ -59,13 +59,13 @@ ggplot(wages_before_tsibble,
            y = mean_hourly_wage,
            group = id)) +
   geom_line(alpha = 0.7) +
-  facet_sample() +
+  facet_sample(n_per_facet = 1, n_facets = 20) +
   scale_x_continuous("Year",
                      breaks = seq(1980, 2020, 10),
                      labels = c("80", "90", "00", "10", "20"),
                      minor_breaks = seq(1980, 2020, 5)) +
   theme_bw() +
-  theme(axis.text.x = element_text(angle = 10, size = 6)) +
+  #theme(axis.text.x = element_text(angle = 10, size = 12)) +
   ylab("Hourly wage ($)")
 
 
@@ -81,15 +81,15 @@ spag <- wages_before %>%
   ggplot(aes(x = year,
              y = mean_hourly_wage,
              group = id)) +
-  geom_line(alpha = 0.1) +
+  geom_line(alpha = 0.5) +
   scale_x_continuous("Year",
                      breaks = seq(1980, 2020, 10),
                      labels = c("80", "90", "00", "10", "20"),
                      minor_breaks = seq(1980, 2020, 5)) +
   ggtitle("A") +
   theme_bw() +
-  ylab("Hourly wage ($)") +
-  theme(plot.title = element_text(size = 10))
+  ylab("Hourly wage ($)") #+
+  #theme(plot.title = element_text(size = 10))
 
 wages_three_feat <- wages_before_tsibble %>%
   features(mean_hourly_wage,
@@ -104,8 +104,8 @@ feature <- ggplot(wages_feat_long) +
   geom_density(aes(x = value, colour = feature, fill = feature), alpha = 0.3) +
   ggtitle("B") +
   theme_bw() +
-  theme(plot.title = element_text(size = 10),
-        axis.text.x = element_text(angle = 10, size = 6),
+  theme(#plot.title = element_text(size = 10),
+        #axis.text.x = element_text(angle = 10, size = 6),
         legend.position = "none")
 
 feature_bp <- ggplot(wages_feat_long,
@@ -117,8 +117,8 @@ feature_bp <- ggplot(wages_feat_long,
   ggtitle("B")  +
   ylab("Hourly wage ($)") +
   xlab("Feature") +
-  theme(legend.position = "none",
-        plot.title = element_text(size = 10))
+  theme(legend.position = "none") #,
+  #      plot.title = element_text(size = 10))
 
 plot_high <- ggplot(filter(wages_high2, id == 39)) +
[…]

### 2021-09-25 paper/paper-jsdse/paper-jsdse.tex
racism not race

@@
 
-Singer and Willett (2003) used the wages and other variables of high school dropouts from the NLSY79 data as an example data set to illustrate longitudinal data modeling.
+Singer and Willett (2003) used the wages and other variables of high school dropouts from the NLSY79 data as an example data set to illustrate longitudinal data modeling of wages on workforce experience, with covariates education and race.
 Our aim is to refresh this textbook data with data from 1994 through to the latest data reported in 2018.
 Here, we investigate the process of getting from the raw NLSY79 data to a textbook data set as similar as possible to that provided by Singer and Willett (2003).
+We should also note that race is a variable in the original data set, and for the purposes of studying racism, not race(T.
+1998), it also provided with the refreshed data.
 
@@
 A robust linear model prediction was used to correct mean hourly wages value.
-We can see that some extreme spikes, corresponding to implausible wages, have been repleaced with values more similar to wages in neighboring years, but otherwise the profiles are not changed.}\label{fig:compare-plot}
+We can see that some extreme spikes, corresponding to implausible wages, have been replaced with values more similar to wages in neighboring years, but otherwise the profiles are not changed.
+Some spikes might remain when wage vaues are plausible.}\label{fig:compare-plot}
 \end{figure}
@@
 
+\leavevmode\hypertarget{ref-racismnotrace}{}%
+T., Fullilove M. 1998. {``Comment: Abandoning "Race" as a Variable in Public Health Research--an Idea Whose Time Has Come.''} \emph{American Journal of Public Health} 88 (9): 1297--98.
+
 \leavevmode\hypertarget{ref-nlsy79}{}%

### 2021-09-25 paper/paper.Rmd
racism not race

@@
 
-@SingerJudithD2003Alda used the wages and other variables of high school dropouts from the NLSY79 data as an example data set to illustrate longitudinal data modeling.
+@SingerJudithD2003Alda used the wages and other variables of high school dropouts from the NLSY79 data as an example data set to illustrate longitudinal data modeling of wages on workforce experience, with covariates education and race.
 Our aim is to refresh this textbook data with data from 1994 through to the latest data reported in 2018.
 Here, we investigate the process of getting from the raw NLSY79 data to a textbook data set as similar as possible to that provided by @SingerJudithD2003Alda.
+We should also note that race is a variable in the original data set, and for the purposes of studying racism, not race[@racismnotrace], it also provided with the refreshed data.
 <!--However, we cannot create the exact same data set as published in their book since we do not have the information of what age threshold they used to determine the high school dropouts.-->
@@
 A robust linear model prediction was used to correct mean hourly wages value.
-We can see that some extreme spikes, corresponding to implausible wages, have been repleaced with values more similar to wages in neighboring years, but otherwise the profiles are not changed.", fig.height=8, fig.width=10, out.width="100%"}
+We can see that some extreme spikes, corresponding to implausible wages, have been replaced with values more similar to wages in neighboring years, but otherwise the profiles are not changed.
+Some spikes might remain when wage vaues are plausible.", fig.height=8, fig.width=10, out.width="100%"}
 ```

### 2021-09-25 paper/paper-jsdse/paper-jsdse.tex
racism not race

@@
 Here, we investigate the process of getting from the raw NLSY79 data to a textbook data set as similar as possible to that provided by Singer and Willett (2003).
-We should also note that race is a variable in the original data set, and for the purposes of studying racism, not race(T.
-1998), it also provided with the refreshed data.
+We should also note that race is a variable in the original data set, and for compatibility it is also provided with the refreshed data, for the purposes of studying racism, not race (Fullilove 1998),
 
@@
 
+\leavevmode\hypertarget{ref-racismnotrace}{}%
+Fullilove, M. T. 1998. {``Comment: Abandoning "Race" as a Variable in Public Health Research--an Idea Whose Time Has Come.''} \emph{American Journal of Public Health} 88 (9): 1297--98.
+
 \leavevmode\hypertarget{ref-purrr}{}%
@@
 Singer, Judith D, and John B Willett. 2003. \emph{Applied Longitudinal Data Analysis: Modeling Change and Event Occurrence}. Oxford u.a: Oxford Univ. Pr.
-
-\leavevmode\hypertarget{ref-racismnotrace}{}%
-T., Fullilove M. 1998. {``Comment: Abandoning "Race" as a Variable in Public Health Research--an Idea Whose Time Has Come.''} \emph{American Journal of Public Health} 88 (9): 1297--98.
 

### 2021-09-25 paper/paper.Rmd
racism not race

@@
 Here, we investigate the process of getting from the raw NLSY79 data to a textbook data set as similar as possible to that provided by @SingerJudithD2003Alda.
-We should also note that race is a variable in the original data set, and for the purposes of studying racism, not race[@racismnotrace], it also provided with the refreshed data.
-<!--However, we cannot create the exact same data set as published in their book since we do not have the information of what age threshold they used to determine the high school dropouts.-->
+We should also note that race is a variable in the original data set, and for compatibility it is also provided with the refreshed data, for the purposes of studying racism, not race [@racismnotrace],  <!--However, we cannot create the exact same data set as published in their book since we do not have the information of what age threshold they used to determine the high school dropouts.-->
 

### 2021-09-26 paper/paper.Rmd
expanded and corrected description in getting the data section

@@
 The NLSY79 data contains a large scope of variables but for our aim, the scope required is limisted to demographic profiles and wages data.
-More specifically, we went to the NLSY79 database website at \url{https://www.nlsinfo.org/content/cohorts/nlsy79/get-data} and navigated as described below.
+More specifically, we went to the NLSY79 database website at \url{https://www.nlsinfo.org/content/cohorts/nlsy79/get-data}, clicked on the direct link to NSLY79 data and navigated as described below.
 
-::: {.tcolorbox data-latex="[label=\"box:data\", title=Getting the data]" }
-\faDatabase\ National Longitudinal Survey of Youth 1979 
+::: {.tcolorbox data-latex="[label=\"box:data\", title=Navigating the data source to download the raw data]" }
+\faDatabase\ NLSY79 (\url{https://www.nlsinfo.org/investigator/pages/search?s=NLSY79})  
+\vspace{1mm}
+\faCheck\ The CASEID will be always be selected.  
+\vspace{1mm}
+\faCheck\ The 3 recommended demographic variable (sample ID, race and sex) were selected.  
+\vspace{1mm}
+For the remaining variables, we went to the "Variable Search" tab and select variables as follows
 \begin{itemize}
@@
 \begin{itemize}
-\item[\faCloudDownload] Downloaded all of variables in Highest grade completed.
+\item[\faCheck] All 80 variables in Highest grade completed were selected.
 \end{itemize}
@@
 \begin{itemize}
-\item[$\triangleright$] Summary measures $\triangleright$ By job $\triangleright$ Hours worked and Hourly wages
+\item[$\triangleright$] Summary measures $\triangleright$ By job
 \begin{itemize}
-\item[\faCloudDownload] We downloaded all of the variables in Hours worked.
-\item[\faCloudDownload] We downloaded all of the variables in Hourly wages
-  Both hours worked and hourly wages are recorded by the job, up to five jobs for each id/subject. 
+\item[$\triangleright$] Hours worked  
+\begin{itemize}
+\item[\faCheck] All 442 primary variables in Hours worked were selected.
+\end{itemize}
+\item[$\triangleright$] Hourly wages
+\begin{itemize}
+\item[\faCheck] All 156 variables in Hourly wages were selected.
+\end{itemize}
 \end{itemize}
@@
 \begin{itemize}
-\item[$\triangleright$] Context $\triangleright$ Summary measures $\triangleright$ Basic demographics 
+\item[$\triangleright$] Context $\triangleright$ Summary measures $\triangleright$ Basic demographics $\triangleright$ Date of birth
 \begin{itemize}
-\item[\faCloudDownload] Downloaded year and month of birth, race, and sex variables. There are two versions of the year and month of birth, i.e., 1979 and 1981 data. We downloaded these two versions.
+\item[\faCheck] All 4 variables in Date of birth were selected. 
 \end{itemize}
@@
 \end{itemize}
-
+\begin{itemize}
+\item[\faCloudDownload] To download all 686 variables selected, we then navigate to the tab "Save / Download" then select the tab "Advanced Download".
+We select the R Source code, Codebook of selected variables, Short Description File and Comma-delimited datafile of selected variables with Reference Number as column headers.
+We name the filename "NLSY79" and press the download button.
+\end{itemize}
 :::
 
-The downloaded data set came with the following set of files:
+The downloaded data set came as a zip file when extracted contains the following set of files:
 
-- `NLSY79.csv`: csv format of the data
-- `NLSY79.dat`: dat format of the data
+- `NLSY79.csv`: csv format of the response data.
+- `NLSY79.dat`: dat format of the response data.
 - `NLSY79.NLSY79`: Tagset of variables that can be uploaded to the website to recreate the data set.
-- `NLSY79.R`: R script provided automatically by the database for reading the data into R and converting the variables' names and label into something more sensible.
-We utilized this code to do an initial data tidying.
-It produced two data set,  `categories_qnames` (the observations are stored in categorical/interval values) and `new_data_qnames` (the observations are stored in integer form).
+- `NLSY79.R`: R script for reading the data into R and converting the variables' names and label into something more sensible.
+We utilize this code to do an initial data tidying which produces two data set,  `categories_qnames` (where the observations are stored in categorical/interval values) and `new_data_qnames` (the observations are stored in integer form).
 

### 2021-09-26 paper/paper.tex
expanded and corrected description in getting the data section

@@
 
-Singer and Willett (2003) used the wages and other variables of high school dropouts from the NLSY79 data as an example data set to illustrate longitudinal data modeling.
+Singer and Willett (2003) used the wages and other variables of high school dropouts from the NLSY79 data as an example data set to illustrate longitudinal data modeling of wages on workforce experience, with covariates education and race.
 Our aim is to refresh this textbook data with data from 1994 through to the latest data reported in 2018.
 Here, we investigate the process of getting from the raw NLSY79 data to a textbook data set as similar as possible to that provided by Singer and Willett (2003).
+We should also note that race is a variable in the original data set, and for compatibility it is also provided with the refreshed data, for the purposes of studying racism, not race (Fullilove 1998),
 
@@
 The NLSY79 data contains a large scope of variables but for our aim, the scope required is limisted to demographic profiles and wages data.
-More specifically, we went to the NLSY79 database website at \url{https://www.nlsinfo.org/content/cohorts/nlsy79/get-data} and navigated as described below.
+More specifically, we went to the NLSY79 database website at \url{https://www.nlsinfo.org/content/cohorts/nlsy79/get-data}, clicked on the direct link to NSLY79 data and navigated as described below.
 
-\begin{tcolorbox}[label="box:data", title=Getting the data]
+\begin{tcolorbox}[label="box:data", title=Navigating the data source to download the raw data]
 
-\faDatabase~National Longitudinal Survey of Youth 1979
+\faDatabase~NLSY79 (\url{https://www.nlsinfo.org/investigator/pages/search?s=NLSY79})\\
+\vspace{1mm}
+\faCheck~The CASEID will be always be selected.\\
+\vspace{1mm}
+\faCheck~The 3 recommended demographic variable (sample ID, race and sex) were selected.\\
+\vspace{1mm}
+For the remaining variables, we went to the ``Variable Search'' tab and select variables as follows
 
@@
 \begin{itemize}
-\item[$\triangleright$] Context $\triangleright$ Summary measures $\triangleright$ Basic demographics 
+\item[$\triangleright$] Context $\triangleright$ Summary measures $\triangleright$ Basic demographics $\triangleright$ Date of birth
 \begin{itemize}
-\item[\faCloudDownload] Downloaded year and month of birth, race, and sex variables. There are two versions of the year and month of birth, i.e., 1979 and 1981 data. We downloaded these two versions.
+\item[\faCheck] All 4 variables in Date of birth were selected. 
 \end{itemize}
 \end{itemize}
+\end{itemize}
+\begin{itemize}
+\item[\faCloudDownload] To download all 686 variables selected, we then navigate to the tab "Save / Download" then select the tab "Advanced Download".
+We select the R Source code, Codebook of selected variables, Short Description File and Comma-delimited datafile of selected variables with Reference Number as column headers.
+We name the filename "NLSY79" and press the download button.
 \end{itemize}
@@
 
-The downloaded data set came with the following set of files:
+The downloaded data set came as a zip file when extracted contains the following set of files:
 
@@
 \item
-  \texttt{NLSY79.csv}: csv format of the data
+  \texttt{NLSY79.csv}: csv format of the response data.
 \item
-  \texttt{NLSY79.dat}: dat format of the data
+  \texttt{NLSY79.dat}: dat format of the response data.
 \item
@@
 \item
-  \texttt{NLSY79.R}: R script provided automatically by the database for reading the data into R and converting the variables' names and label into something more sensible.
-We utilized this code to do an initial data tidying.
-It produced two data set, \texttt{categories\_qnames} (the observations are stored in categorical/interval values) and \texttt{new\_data\_qnames} (the observations are stored in integer form).
+  \texttt{NLSY79.R}: R script for reading the data into R and converting the variables' names and label into something more sensible.
+We utilize this code to do an initial data tidying which produces two data set, \texttt{categories\_qnames} (where the observations are stored in categorical/interval values) and \texttt{new\_data\_qnames} (the observations are stored in integer form).
 \end{itemize}
@@
 
-\caption{Comparison between the original and the treated mean hourly wage.
-The orange line portray the original value of mean hourly wage, while the turquoise line display the mean hourly wages value after the extreme values imputed with the robust linear model's prediction value.
-We can see that some extreme spikes has been reduced by the model.}\label{fig:compare-plot}
+\caption{Comparison between the original (black dots) and the corrected (solid grey) mean hourly wage for a sample of individuals.
+A robust linear model prediction was used to correct mean hourly wages value.
+We can see that some extreme spikes, corresponding to implausible wages, have been replaced with values more similar to wages in neighboring years, but otherwise the profiles are not changed.
+Some spikes might remain when wage vaues are plausible.}\label{fig:compare-plot}
 \end{figure}
@@
 
-{\centering \includegraphics[width=0.93\linewidth]{/Users/etan0038/Dropbox/repo/summer-wages-refresh/paper/figures/flowchart}
+{\centering \includegraphics[width=0.85\linewidth]{/Users/etan0038/Dropbox/repo/summer-wages-refresh/paper/figures/flowchart}
 

### 2021-09-26 paper/paper-jsdse/paper-jsdse.tex
changes to target data and data cleaning

@@
 
-This paper aims, as mentioned in Section \ref{intro}, to refresh the NLSY79 data used in Singer and Willett (2003).
-This data set contains the information of yearly mean hourly wages from the NLSY79 subcohort, with education and race as covariates.
-The subcohort correspond to male high-school dropouts that participated in the study from 14 to 17 years old.
-Accordingly, to refresh this data, we make use of NLSY79 cohort's demographic profiles and wages data of people who completed education up to high school.
-
-Since no specific criteria on high school dropouts was defined
-in Singer and Willett (2003), we define high school dropouts as respondents who only completed either 9th to 11th grade, or who completed 12th when their age is at least 19 years old (meaning that the individual may have dropped out of high school but returned at a later time to complete high school education).
+The NLSY79 data used in Singer and Willett (2003).
+contains the longitudinal measurements on yearly mean hourly wages with years of workforce experience, and demographic variables education and race, from 1979 through to 1994.
+In addition, the cohort is restricted to male high-school dropouts who first participated in the study at age 14-17 years.
+Thus the target data set is to collect the same variables for the extended time frame of 1994 through to 2018, the most recent year reported.
+.
 
@@
 
-The downloaded data set came with the following set of files:
+The downloaded data set comes as a zip file, containing the following set of files:
 
@@
 According to Wickham (2014), tidy data sets comply with three rules: (i) each variable forms a column, (ii) each observation forms a row, and (iii) each type of observational unit forms a table.
-The raw data does not comply the rules of the tidy data, thus we wrangle the data in form required by making use of the \texttt{tidyverse} suite of packages (Wickham et al.
-2019), which include the main set of packages needed, namely \texttt{tidyr} (Wickham 2020b), \texttt{dplyr} (Wickham et al.
-2020), and \texttt{stringr} (Wickham 2019).
-The tidy data form, where we have a variable that indicates the individual ID, year, job number, wage in dollars and demographic variables, is important for downstream analysis, such as visualisation and modelling that will be performed later.
-
-{[}REMOVE{]} Unfortunately, the \texttt{new\_data\_qnames} did not meet these requirements in the way that data for each year and job for the same respondent is stored as one variable.
-Hence, the data contains a huge number of columns (686 columns).
-The example of the untidy data set is displayed in Table \ref{tab:untidy-data}.
-The table intended to display the hourly rate of each respondent by job (HRP1 to HRP5) and by year (1979 and 1980).
-The table implies that the column headers are values of the year and job, not variable names.
-Consequently, the data should be tidied and wrangled first to extract the demographic and employment variables we want to put in the final data set.
-We mainly used \texttt{tidyr} (Wickham 2020b), \texttt{dplyr} (Wickham et al.
-2020), and \texttt{stringr} (Wickham 2019) to do this job.
+The raw data does not comply these rules, thus we re-arrange and wrangle the data into tidy data form, columns corresponding to individual ID, year, job number, wage in dollars and the demographic variables.
+This is done using the \texttt{tidyverse} suite of packages (Wickham et al.
+2019), \texttt{tidyr} (Wickham 2020b) to pivot the data into long form, \texttt{dplyr} (Wickham et al.
+2020), and \texttt{stringr} (Wickham 2019) for creating new variables, and levels of factors by text wrangling.
+The long form of the data makes it possible to do these data transformations efficiently, and it is an intermediate step towards the final target data.
 

### 2021-09-26 paper/paper.Rmd
changes to target data and data cleaning

@@
 
-This paper aims, as mentioned in Section \@ref(intro), to refresh the NLSY79 data used in @SingerJudithD2003Alda.
-This data set contains the information of yearly mean hourly wages from the NLSY79 subcohort, with education and race as covariates.
-The subcohort correspond to male high-school dropouts that participated in the study from 14 to 17 years old.
-Accordingly, to refresh this data, we make use of NLSY79 cohort's demographic profiles and wages data of people who completed education up to high school.
+The NLSY79 data used in @SingerJudithD2003Alda.
+contains the longitudinal measurements on yearly mean hourly wages with years of workforce experience, and demographic variables education and race, from 1979 through to 1994.
+In addition, the cohort is restricted to male high-school dropouts who first participated in the study at age 14-17 years.
+Thus the target data set is to collect the same variables for the extended time frame of 1994 through to 2018, the most recent year reported.
+.
 
-Since no specific criteria on high school dropouts was defined
+<!-- Since no specific criteria on high school dropouts was defined
 in @SingerJudithD2003Alda, we define high school dropouts as respondents who only completed either 9th to 11th grade, or who completed 12th when their age is at least 19 years old (meaning that the individual may have dropped out of high school but returned at a later time to complete high school education).
+XXX This should probably go further down in the flow.
+-->
 
@@
 
-The downloaded data set came as a zip file when extracted contains the following set of files:
+The downloaded data set comes as a zip file, containing the following set of files:
 
@@
 According to @WickhamHadley2014TD, tidy data sets comply with three rules: (i) each variable forms a column, (ii) each observation forms a row, and (iii) each type of observational unit forms a table.
-The raw data does not comply the rules of the tidy data, thus we wrangle the data in form required by making use of the `tidyverse` suite of packages [@tidyverse], which include the main set of packages needed, namely `tidyr` [@tidyr], `dplyr` [@dplyr], and `stringr` [@stringr].
+The raw data does not comply these rules, thus we re-arrange and wrangle the data into tidy data form, columns corresponding to individual ID, year, job number, wage in dollars and the demographic variables.
+This is done using the `tidyverse` suite of packages [@tidyverse], `tidyr` [@tidyr] to pivot the data into long form, `dplyr` [@dplyr], and `stringr` [@stringr] for creating new variables, and levels of factors by text wrangling.
+The long form of the data makes it possible to do these data transformations efficiently, and it is an intermediate step towards the final target data.
+
+<!-- required by making use of the `tidyverse` suite of packages [@tidyverse], which include the main set of packages needed, namely `tidyr` [@tidyr], `dplyr` [@dplyr], and `stringr` [@stringr].
 The tidy data form, where we have a variable that indicates the individual ID, year, job number, wage in dollars and demographic variables, is important for downstream analysis, such as visualisation and modelling that will be performed later.
@@
 Consequently, the data should be tidied and wrangled first to extract the demographic and employment variables we want to put in the final data set.
-We mainly used `tidyr` [@tidyr], `dplyr` [@dplyr], and `stringr` [@stringr] to do this job.
+We mainly used `tidyr` [@tidyr], `dplyr` [@dplyr], and `stringr` [@stringr] to do this job.-->
 

### 2021-09-26 paper/paper.tex
comparison plots

@@
 
-{\centering \includegraphics[width=1\linewidth]{figures/compare-plot-1}
+{\centering \includegraphics[width=0.9\linewidth]{figures/compare-plot-1}
 
@@
 
-{\centering \includegraphics[width=0.85\linewidth]{/Users/etan0038/Dropbox/repo/summer-wages-refresh/paper/figures/flowchart}
+{\centering \includegraphics[width=0.85\linewidth]{/Users/cookd/numbats/summer-wages-refresh/paper/figures/flowchart}
 

### 2021-09-26 paper/scripts/data-preprocessing.R

@@ -17,6 +17,7 @@ library(modelr)
 library(forcats)
 library(yowie)
 library(knitr)
+library(ggthemes)
 select <- dplyr::select
 
 ## ---- raw-data

### 2021-09-26 paper/scripts/eda.R

@@ -5,14 +5,34 @@ sw <- brolgar::wages
 sw_wages <- sw %>%
   ggplot(aes(x = xp,
              y = ln_wages)) +
-  geom_line(aes(group = id),alpha = 0.1) +
+  geom_line(aes(group = id), alpha = 0.1) +
   geom_smooth(se = FALSE, colour = "pink") +
   ggtitle("A original") +
   theme_bw() +
   ylab("Hourly wage ($, natural log)") +
   xlab("Experience in years") +
-  theme(plot.title = element_text(size = 10)) +
+  #theme(plot.title = element_text(size = 10)) +
   ylim(0, 5)
+
+sw_wages_mod <- sw %>%
+  as_tibble() %>%
+  mutate(hgc = ifelse(high_grade < 9, "8TH", "12TH")) %>%
+  mutate(race = case_when(black == 1 ~ "black",
+                          hispanic == 1 ~ "hispanic",
+                          TRUE ~ "white")) %>%
+  mutate(race = factor(race)) %>%
+  ggplot(aes(x = xp,
+             y = ln_wages,
+             linetype = hgc)) +
+  geom_smooth(method = "lm", se = FALSE,
+              colour = "black") +
+  ggtitle("A original") +
+  scale_linetype("") +
+  theme_bw() +
+  ylab("Hourly wage ($, natural log)") +
+  xlab("Experience in years") +
+  theme(legend.position = "bottom",
+        legend.direction = "horizontal")
 
 ## ---- do_refreshed
 
@@ -25,8 +45,8 @@ sw_wages <- sw %>%
 # filter do data up to 1994 since the textbook data only covers that period.
 
 do <- yowie::wages_hs_do %>%
-  mutate(lnwage = log(wage)) %>%
-  filter(year < 1995)
+  mutate(lnwage = log(wage)) #%>%
+  #filter(year < 1995)
 
 do_ref <- do %>%
   ggplot(aes(x = year,
@@ -39,6 +59,21 @@ do_ref <- do %>%
   xlab("Year of data collection") +
   theme(plot.title = element_text(size = 10)) +
   ylim(0, 5)
+
+do_ref_mod <- do %>%
+  mutate(hgc12 = ifelse(hgc_i < 12, "BELOW 12TH", "12TH")) %>%
+  ggplot(aes(x = year,
+             y = lnwage,
+             linetype = hgc12)) +
+  geom_smooth(method = "lm", se = FALSE,
[…]

### 2021-09-26 paper/paper-jsdse/paper-jsdse.tex
comparison plots
comparison

@@
 
-In this section, we have demonstrate a workflow from getting the data from the NSLY79 database until a ready-analysed data.
-Referring to Loo and Jonge (2018), this section shows part of statistical value chain, which is data cleaning from raw, input, and valid data.
-The raw demographic and employment data from the database are not in a tidy format.
-Hence, we perform data wrangling to get the input data.
-We inspect the input data to find eligible observations to the target data.
-One of the target data is wages data of respondents who completed up to 12th grade and participated at least in 5 rounds of survey.
-This needs us to join the extracted demographic and employment data.
-
-Further, using IDA, we inspect on the existence of anomalies/extreme values in the data using the weight from robust linear model. We then replace these extreme values using their predicted value.
-From this stage forward, we assume that we already have valid data.
-We also subset the data to get the wages of high school dropouts.
-We then make these three data sets and the cleaning workflow documentation publicly available through an R data container package called \texttt{yowie}.
-A visualisation of these process is displayed in Figure \ref{fig:flow-chart}.
+Figure \ref{fig:flow-chart} summarizes the steps taken to go from raw to input to valid data (Loo and Jonge 2018) to create a refreshed wages data set.
 
@@
 
+\begin{verbatim}
+#> # A tibble: 3 x 2
+#>   race         n
+#>   <fct>    <int>
+#> 1 black      246
+#> 2 hispanic   204
+#> 3 white      438
+\end{verbatim}
+
+\begin{verbatim}
+#> # A tibble: 2 x 2
+#>   hgc       n
+#>   <chr> <int>
+#> 1 12TH    522
+#> 2 8TH     366
+\end{verbatim}
+
+\begin{verbatim}
+#> # A tibble: 3 x 2
+#>   race                        n
+#>   <fct>                   <int>
+#> 1 HISPANIC                  219
+#> 2 BLACK                     346
+#> 3 NON-BLACK, NON-HISPANIC   623
+\end{verbatim}
+
+\begin{verbatim}
+#> # A tibble: 2 x 2
+#>   hgc12          n
+#>   <chr>      <int>
+#> 1 12TH         967
+#> 2 BELOW 12TH   221
+\end{verbatim}
+
 There are two aspects of the original data that make a direct comparison difficult.
@@
 
-Figure \ref{fig:plotting-sw-do} shows the longitudinal profiles of the two sets.
+Figure \ref{fig:plotting-sw-do} shows a comparison of the two sets.
 (A direct ID matching is not possible.) There are 888 individuals in the original and 1,188 individuals in the refreshed data.
 This suggests some individuals were removed in the original data.
-Overall, the profiles show some similarity.
+In addition, in the original set the racial break down was 246 black, 204 hispanic and 438 white, while in tne refreshed data there are 346 black, 219 hispanic and 623 white, so the proportions have not been exactly reproduced.
+On education, in the refreshed data there are very few individuals with less than a 12th grade education, which varies from the original.
+In the original data there were 366 individuals with up to 8th grade and 522 9th-12th grade, as compared to 967 with 12th grade in the refreshed data.
 

### 2021-09-26 paper/paper.Rmd
comparison plots
comparison

@@
 
+```{r load-pkgs}
+```
+
 ```{r theme}
@@
                   ) )
-```
-
-```{r load-pkgs}
 ```
@@
 We can see that some extreme spikes, corresponding to implausible wages, have been replaced with values more similar to wages in neighboring years, but otherwise the profiles are not changed.
-Some spikes might remain when wage vaues are plausible.", fig.height=8, fig.width=10, out.width="100%"}
+Some spikes might remain when wage vaues are plausible.", fig.height=8, fig.width=10, out.width="90%"}
 ```
@@
 
-In this section, we have demonstrate a workflow from getting the data from the NSLY79 database until a ready-analysed data.
+Figure \ref{fig:flow-chart} summarizes the steps taken to go from raw to input to valid data [@LooMarkvander2018Sdcw] to create a refreshed wages data set.
+
+<!--To summarise, the  a workflow from getting the data from the NSLY79 database until a ready-analysed data.
 Referring to @LooMarkvander2018Sdcw, this section shows  part of statistical value chain, which is data cleaning from raw, input, and valid data.
@@
 We then make these three data sets and the cleaning workflow documentation publicly available through an R data container package called `yowie`.
-A visualisation of these process is displayed in  Figure \ref{fig:flow-chart}.
+A visualisation of these process is displayed in  Figure \ref{fig:flow-chart}.-->
 

### 2021-09-26 paper/scripts/eda.R

@@ -152,6 +152,25 @@ do_ref_agree <- do_agree %>%
 # sw_wages + do_ref
 sw_wages_mod + do_ref_mod
 
+# Summaries
+sw2 <- sw %>%
+  as_tibble() %>%
+  mutate(hgc = ifelse(high_grade < 9, "8TH", "12TH")) %>%
+  mutate(race = case_when(black == 1 ~ "black",
+                          hispanic == 1 ~ "hispanic",
+                          TRUE ~ "white")) %>%
+  mutate(race = factor(race))
+sw2 %>% select(id, race) %>% distinct() %>% count(race)
+sw2 %>% select(id, hgc) %>% distinct() %>% count(hgc)
+
+do2 <- do %>%
+  as_tibble() %>%
+  mutate(hgc12 = ifelse(hgc_i < 12, "BELOW 12TH", "12TH"))
+do2 %>% select(id, race) %>% distinct() %>% count(race)
+do2 %>% select(id, hgc12) %>% distinct() %>% count(hgc12)
+
+
+
 # Takeaways:
 
 # we don't know how exactly the criteria of drop out,

### 2021-09-26 paper/paper-jsdse/paper-jsdse.tex
Conclusions

@@
 
-\begin{verbatim}
-#> # A tibble: 3 x 2
-#>   race         n
-#>   <fct>    <int>
-#> 1 black      246
-#> 2 hispanic   204
-#> 3 white      438
-\end{verbatim}
-
-\begin{verbatim}
-#> # A tibble: 2 x 2
-#>   hgc       n
-#>   <chr> <int>
-#> 1 12TH    522
-#> 2 8TH     366
-\end{verbatim}
-
-\begin{verbatim}
-#> # A tibble: 3 x 2
-#>   race                        n
-#>   <fct>                   <int>
-#> 1 HISPANIC                  219
-#> 2 BLACK                     346
-#> 3 NON-BLACK, NON-HISPANIC   623
-\end{verbatim}
-
-\begin{verbatim}
-#> # A tibble: 2 x 2
-#>   hgc12          n
-#>   <chr>      <int>
-#> 1 12TH         967
-#> 2 BELOW 12TH   221
-\end{verbatim}
-
 There are two aspects of the original data that make a direct comparison difficult.
@@
 
-This paper has performed stages to make open data suitable for textbook data or make it ready for research.
+This paper has described the stages to take a particular open data set and make it a textbook data set, ready for the classroom or research.
 In the first stage, we showed the steps performed to get the data from the NLSY79 database.
-Since the data format is untidy, we showed how the data had been tidied.
-After that, we conducted an initial data analysis to investigate and screen the quality of the data.
-We found and fixed the anomalous observations in the data set using the robust linear regression model with its predicted values.
-We also performed an example of exploratory data analysis using the cleaned data set.
+The data format needed converting to tidy format, and this was described.
+After that, an initial data analysis was conducted to investigate and screen the quality of the data.
+We found and fixed the anomalous observations in wages using a robust linear regression model.
+Finally a comparison is made between the original and refreshed data sets.
 
-This paper has also demonstrated data cleaning documentation by providing all of the codes in performing data tidying and initial data analysis.
-Thus, this paper provided an opportunity to continuously refresh the textbook data whenever the updated data is published in the NLSY79 database.
-It could be done by following the documentation of the code that is provided in this paper.
+The data cleaning process is documented and the code has been made available. These provide the opportunity to again refresh the textbook data as updated data is published in the NLSY79 database.
+Determining the appropriate robustness weight to threshold the anomalous observations is documented, and a \texttt{shiny} (Chang et al. 2020) app is provided to assist.
+The current subset is made available in a new R package, called yowie.
 
-Moreover, the documentation also includes how we generated the robustness weight and how we decided the threshold of anomalous observations.
-It is also documented along with the flag of whether an observation is an imputed value or not.
-Accordingly, if somebody wishes to make another decision, it can be done by making a small change in the code provided.
-Further, this paper is also supplemented by a \texttt{shiny} (Chang et al.
-2020) app as a simulation tool to customize the weight threshold.
+Various difficulties were encountered in trying to refresh the data, which include:
 
-Finally, this paper implies that data providers should design a database that is able to produce tidy data sets.
-A data provider should also check for data anomalies before the data publishing or at least provides a set of rules or threshold values.
-For example, in this case, is the threshold of reasonable wages.
-This will greatly support the data users to validate and set the same understanding of which data are considered outliers.
-Moreover, providing validation rules would facilitate any established data validation tool, such as \texttt{validate} (van der Loo and de Jonge 2021) package.
-In this case, we cannot use this handy validation package due to the absence of validation rules.
+\begin{itemize}
+\tightlist
+\item
+  determining which records should be downloaded from the database.
+\item
+  there are many errors in the data, e.g.~hourly wages greater than \$30,000 per hour, and there was no information on how these were handled in the original data.
+\item
+  there is no explicit variable in the database recording high school dropout, which means we needed to compare date of 12th grade with the individual's age.
+\item
+  calculating experience in the workforce, would require knowing the time of first job within the first year the individual was recorded.
+\end{itemize}
+
+Ultimately, the refreshed data is reasonably similar to the original, but unsatisfactorily far from it.
+The last step required would be to inflation-adjust wages, but this is better to do with each wave of new data added, so that it is relative to the last date in the data.
+
+Finally, for the data providers we recommend that a validation system with clear rules is added on data entry, and that alternative output formats, such as a tidy format would help users make better use of the resource.
+The problem with many of the wages records is that there are implausible values, or confusion on how to record wages for multiple jobs.
+These values could be validated with simple checks at data entry.
+Providing an open data resource also is accompanied with the responsibility that the data, especially data that is as valuable as this, is reliable.
+Users need to be able to trust the data.
 
@@
 
-\leavevmode\hypertarget{ref-validate}{}%
-van der Loo, Mark P. J., and Edwin de Jonge. 2021. {``Data Validation Infrastructure for {R}.''} \emph{Journal of Statistical Software} 97 (10): 1--31. \url{https://doi.org/10.18637/jss.v097.i10}.
-
 \leavevmode\hypertarget{ref-mass}{}%

### 2021-09-26 paper/paper.Rmd
Conclusions

@@
 
+```{r summaries, eval=FALSE}
+```
+
 <!-- Hence, in this section, we will compare how close the `wages_hs_do`, henceforth refreshed data, to the textbook data.
@@
 
-This paper has performed stages to make open data suitable for textbook data or make it ready for research.
+This paper has described the stages to take a particular open data set and make it a textbook data set, ready for the classroom or research.
 In the first stage, we showed the steps performed to get the data from the NLSY79 database.
-Since the data format is untidy, we showed how the data had been tidied.
-After that, we conducted an initial data analysis to investigate and screen the quality of the data.
-We found and fixed the anomalous observations in the data set using the robust linear regression model with its predicted values.
-We also performed an example of exploratory data analysis using the cleaned data set.
+The data format needed converting to tidy format, and this was described.
+After that, an initial data analysis was conducted to investigate and screen the quality of the data.
+We found and fixed the anomalous observations in wages using a robust linear regression model.
+Finally a comparison is made between the original and refreshed data sets.
 
-This paper has also demonstrated data cleaning documentation by providing all of the codes in performing data tidying and initial data analysis.
-Thus, this paper provided an opportunity to continuously refresh the textbook data whenever the updated data is published in the NLSY79 database.
-It could be done by following the documentation of the code that is provided in this paper.
+The data cleaning process is documented and the code has been made available. These provide the opportunity to again refresh the textbook data as updated data is published in the NLSY79 database.
+Determining the appropriate robustness weight to threshold the anomalous observations is documented, and a `shiny`  [@shiny] app is provided to assist. 
+The current subset is made available in a new R package, called yowie.
 
-Moreover, the documentation also includes how we generated the robustness weight and how we decided the threshold of anomalous observations.
-It is also documented along with the flag of whether an observation is an imputed value or not.
-Accordingly, if somebody wishes to make another decision, it can be done by making a small change in the code provided.
-Further, this paper is also supplemented by a `shiny`  [@shiny] app as a simulation tool to customize the weight threshold.
+Various difficulties were encountered in trying to refresh the data, which include:
 
-Finally, this paper implies that data providers should design a database that is able to produce tidy data sets.
+- determining which records should be downloaded from the database.
+- there are many errors in the data, e.g. hourly wages greater than $30,000 per hour, and there was no information on how these were handled in the original data.
+- there is no explicit variable in the database recording high school dropout, which means we needed to compare date of 12th grade with the individual's age.
+- calculating experience in the workforce, would require knowing the time of first job within the first year the individual was recorded.
+
+Ultimately, the refreshed data is reasonably similar to the original, but unsatisfactorily far from it.
+The last step required would be to inflation-adjust wages, but this is better to do with each wave of new data added, so that it is relative to the last date in the data.
+
+Finally, for the data providers we recommend that a validation system with clear rules is added on data entry, and that alternative output formats, such as a tidy format would help users make better use of the resource.
+The problem with many of the wages records is that there are implausible values, or confusion on how to record wages for multiple jobs.
+These values could be validated with simple checks at data entry.
+Providing an open data resource also is accompanied with the responsibility that the data, especially data that is as valuable as this, is reliable.
+Users need to be able to trust the data.
+
+<!--this paper implies that data providers should design a database that is able to produce tidy data sets.
 A data provider should also check for data anomalies before the data publishing or at least provides a set of rules or threshold values.
@@
 In this case, we cannot use this handy validation package due to the absence of validation rules.
+-->
 

### 2021-09-26 paper/scripts/eda.R

@@ -152,7 +152,7 @@ do_ref_agree <- do_agree %>%
 # sw_wages + do_ref
 sw_wages_mod + do_ref_mod
 
-# Summaries
+# ---- summaries
 sw2 <- sw %>%
   as_tibble() %>%
   mutate(hgc = ifelse(high_grade < 9, "8TH", "12TH")) %>%

### 2021-09-26 paper/paper-jsdse/paper-jsdse.tex
Fix reference.

@@
 
-Loo and Jonge (2018) describe the notion of a ``statistical value chain'' where the production stages of data cleaning process are earmarked as raw data (data from the data source), input data (data organised with correct type and identified variables) and valid data (data that faithfully represent the variables).
+van der Loo and de Jonge (2018) describe the notion of a ``statistical value chain'' where the production stages of data cleaning process are earmarked as raw data (data from the data source), input data (data organised with correct type and identified variables) and valid data (data that faithfully represent the variables).
 In this section, we outline the steps to download the raw data (Section \ref{getdata}) and then tidying the raw data into input data, specifically for wages data (Section \ref{tidywages}), demographic variables (Section \ref{tidydemog}) and the employment variables (Section \ref{tidyemp}), so that the resulting input data can be used downstream for validating the data as described in Section \ref{ida}.
@@
 
-The NLSY79 data contains a large scope of variables but for our aim, the scope required is limisted to demographic profiles and wages data.
+The NLSY79 data contains a large scope of variables but for our aim, the scope required is limited to demographic profiles and wages data.
 More specifically, we went to the NLSY79 database website at \url{https://www.nlsinfo.org/content/cohorts/nlsy79/get-data}, clicked on the direct link to NSLY79 data and navigated as described below.
@@
 The raw data does not comply these rules, thus we re-arrange and wrangle the data into tidy data form, columns corresponding to individual ID, year, job number, wage in dollars and the demographic variables.
-This is done using the \texttt{tidyverse} suite of packages (Wickham et al.
-2019), \texttt{tidyr} (Wickham 2020b) to pivot the data into long form, \texttt{dplyr} (Wickham et al.
-2020), and \texttt{stringr} (Wickham 2019) for creating new variables, and levels of factors by text wrangling.
+This is done using the \texttt{tidyverse} suite of packages (H.
+Wickham, Averick, et al.
+2019), \texttt{tidyr} (H.
+Wickham 2020b) to pivot the data into long form, \texttt{dplyr} (H.
+Wickham, François, et al.
+2020), and \texttt{stringr} (H.
+Wickham 2019) for creating new variables, and levels of factors by text wrangling.
 The long form of the data makes it possible to do these data transformations efficiently, and it is an intermediate step towards the final target data.
@@
 
-According to Huebner Marianne, Vach, and Cessie (2016), initial data analysis (IDA) is the step of inspecting and screening the data after being collected to ensure that the data is clean, valid, and ready to be deployed in the later formal statistical analysis.
+According to Huebner, Vach, and Cessie (2016), initial data analysis (IDA) is the step of inspecting and screening the data after being collected to ensure that the data is clean, valid, and ready to be deployed in the later formal statistical analysis.
 Moreover, Chatfield (1985) argues that the two main objectives of IDA are data description, which is to assess the structure and the quality of the data, and model formulation without any formal statistical inference.
@@
 
-Figure \ref{fig:flow-chart} summarizes the steps taken to go from raw to input to valid data (Loo and Jonge 2018) to create a refreshed wages data set.
+Figure \ref{fig:flow-chart} summarizes the steps taken to go from raw to input to valid data (van der Loo and de Jonge 2018) to create a refreshed wages data set.
 
@@
 
-{\centering \includegraphics[width=0.85\linewidth]{/Users/cookd/numbats/summer-wages-refresh/paper/figures/flowchart}
+{\centering \includegraphics[width=0.85\linewidth]{/Users/deeamaliah/Documents/summer_project/summer-wages-refresh/paper/figures/flowchart}
 
@@
 This suggests some individuals were removed in the original data.
-In addition, in the original set the racial break down was 246 black, 204 hispanic and 438 white, while in tne refreshed data there are 346 black, 219 hispanic and 623 white, so the proportions have not been exactly reproduced.
+In addition, in the original set the racial break down was 246 black, 204 hispanic and 438 white, while in the refreshed data there are 346 black, 219 hispanic and 623 white, so the proportions have not been exactly reproduced.
 On education, in the refreshed data there are very few individuals with less than a 12th grade education, which varies from the original.
@@
 \leavevmode\hypertarget{ref-racismnotrace}{}%
-Fullilove, M. T. 1998. {``Comment: Abandoning "Race" as a Variable in Public Health Research--an Idea Whose Time Has Come.''} \emph{American Journal of Public Health} 88 (9): 1297--98.
+Fullilove, M. T. 1998. ``Comment: Abandoning "Race" as a Variable in Public Health Research--an Idea Whose Time Has Come.'' \emph{American Journal of Public Health} 88 (9): 1297--8.
 
@@
 \leavevmode\hypertarget{ref-KollerManuel2016rARP}{}%
-Koller, Manuel. 2016. {``Robustlmm: An r Package for Robust Estimation of Linear Mixed-Effects Models.''} \emph{Journal of Statistical Software} 75 (6): 1--24.
-
-\leavevmode\hypertarget{ref-LooMarkvander2018Sdcw}{}%
-Loo, Mark van der, and Edwin de Jonge. 2018. \emph{Statistical Data Cleaning with Applications in r}.
+Koller, Manuel. 2016. ``Robustlmm: An R Package for Robust Estimation of Linear Mixed-Effects Models.'' \emph{Journal of Statistical Software} 75 (6): 1--24.
 
 \leavevmode\hypertarget{ref-rsvg}{}%
-Ooms, Jeroen. 2020. \emph{Rsvg: Render SVG Images into PDF, PNG, PostScript, or Bitmap Arrays}. \url{https://CRAN.R-project.org/package=rsvg}.
+Ooms, Jeroen. 2020. \emph{Rsvg: Render Svg Images into Pdf, Png, Postscript, or Bitmap Arrays}. \url{https://CRAN.R-project.org/package=rsvg}.
 
 \leavevmode\hypertarget{ref-opendata}{}%
-Open Knowledge Foundation. 2021. {``Open Definition. Defining Open in Open Data, Open Content, and Open Knowledge.''} 2021. \url{http://opendefinition.org/od/2.1/en/}.
+Open Knowledge Foundation. 2021. ``Open Definition. Defining Open in Open Data, Open Content, and Open Knowledge.'' 2021. \url{http://opendefinition.org/od/2.1/en/}.
 
@@
 \leavevmode\hypertarget{ref-rlm}{}%
-UCLA: Statistical Consulting Group. 2021. {``Robust Regression \textbar{} r Data Analysis Examples.''} February 2021. \url{https://stats.idre.ucla.edu/r/dae/robust-regression/}.
+UCLA: Statistical Consulting Group. 2021. ``Robust Regression \textbar{} R Data Analysis Examples.'' February 2021. \url{https://stats.idre.ucla.edu/r/dae/robust-regression/}.
+
+\leavevmode\hypertarget{ref-LooMarkvander2018Sdcw}{}%
+van der Loo, Mark, and Edwin de Jonge. 2018. \emph{Statistical Data Cleaning with Applications in R}.
 
 \leavevmode\hypertarget{ref-mass}{}%
-Venables, W. N., and B. D. Ripley. 2002. \emph{Modern Applied Statistics with s}. Fourth. New York: Springer. \url{http://www.stats.ox.ac.uk/pub/MASS4}.
+Venables, W. N., and B. D. Ripley. 2002. \emph{Modern Applied Statistics with S}. Fourth. New York: Springer. \url{http://www.stats.ox.ac.uk/pub/MASS4}.
 

### 2021-09-26 paper/paper.Rmd
Fix reference.

@@
 Some have had flat wages for years but had a sudden increase in one particular year, then it gone down again, while the others experienced an upsurge in their wage, for instance, the IDs in panel 9.
+However, when checking the summary plots (Figure \ref{fig:feature-plot}), we found that some observations had exceptionally high wages.
+Some of them, for example respondents in Figure \ref {fig:feature-plot} C had experienced unusual wages only in certain year.
 

### 2021-09-26 paper/paper.tex
Fix reference.

@@
 
-In this section, we have demonstrate a workflow from getting the data from the NSLY79 database until a ready-analysed data.
-Referring to Loo and Jonge (2018), this section shows part of statistical value chain, which is data cleaning from raw, input, and valid data.
-The raw demographic and employment data from the database are not in a tidy format.
-Hence, we perform data wrangling to get the input data.
-We inspect the input data to find eligible observations to the target data.
-One of the target data is wages data of respondents who completed up to 12th grade and participated at least in 5 rounds of survey.
-This needs us to join the extracted demographic and employment data.
-
-Further, using IDA, we inspect on the existence of anomalies/extreme values in the data using the weight from robust linear model. We then replace these extreme values using their predicted value.
-From this stage forward, we assume that we already have valid data.
-We also subset the data to get the wages of high school dropouts.
-We then make these three data sets and the cleaning workflow documentation publicly available through an R data container package called \texttt{yowie}.
-A visualisation of these process is displayed in Figure \ref{fig:flow-chart}.
+Figure \ref{fig:flow-chart} summarizes the steps taken to go from raw to input to valid data (van der Loo and de Jonge 2018) to create a refreshed wages data set.
 
@@
 
-Figure \ref{fig:plotting-sw-do} shows the longitudinal profiles of the two sets.
+Figure \ref{fig:plotting-sw-do} shows a comparison of the two sets.
 (A direct ID matching is not possible.) There are 888 individuals in the original and 1,188 individuals in the refreshed data.
 This suggests some individuals were removed in the original data.
-Overall, the profiles show some similarity.
+In addition, in the original set the racial break down was 246 black, 204 hispanic and 438 white, while in the refreshed data there are 346 black, 219 hispanic and 623 white, so the proportions have not been exactly reproduced.
+On education, in the refreshed data there are very few individuals with less than a 12th grade education, which varies from the original.
+In the original data there were 366 individuals with up to 8th grade and 522 9th-12th grade, as compared to 967 with 12th grade in the refreshed data.
 
@@
 \leavevmode\hypertarget{ref-rlm}{}%
-UCLA: Statistical Consulting Group. 2021. {``Robust Regression \textbar{} r Data Analysis Examples.''} February 2021. \url{https://stats.idre.ucla.edu/r/dae/robust-regression/}.
+UCLA: Statistical Consulting Group. 2021. ``Robust Regression \textbar{} R Data Analysis Examples.'' February 2021. \url{https://stats.idre.ucla.edu/r/dae/robust-regression/}.
 
-\leavevmode\hypertarget{ref-validate}{}%
-van der Loo, Mark P. J., and Edwin de Jonge. 2021. {``Data Validation Infrastructure for {R}.''} \emph{Journal of Statistical Software} 97 (10): 1--31. \url{https://doi.org/10.18637/jss.v097.i10}.
+\leavevmode\hypertarget{ref-LooMarkvander2018Sdcw}{}%
+van der Loo, Mark, and Edwin de Jonge. 2018. \emph{Statistical Data Cleaning with Applications in R}.
 
 \leavevmode\hypertarget{ref-mass}{}%
-Venables, W. N., and B. D. Ripley. 2002. \emph{Modern Applied Statistics with s}. Fourth. New York: Springer. \url{http://www.stats.ox.ac.uk/pub/MASS4}.
+Venables, W. N., and B. D. Ripley. 2002. \emph{Modern Applied Statistics with S}. Fourth. New York: Springer. \url{http://www.stats.ox.ac.uk/pub/MASS4}.
 

### 2021-09-26 paper/scripts/ida.R

@@ -271,7 +271,8 @@ wages_three_feat_rlm <- wages_hs2020_rlm %>%
            feat_three_num
   )
 wages_feat_long_rlm <- wages_three_feat_rlm %>%
-  pivot_longer(c(min, med, max), names_to = "feature", values_to = "value")
+  pivot_longer(c(min, med, max), names_to = "feature", values_to = "value") %>%
+  mutate(feature = factor(feature, levels = c("min", "med", "max")))
 
 feature2 <- ggplot(wages_feat_long_rlm) +
   geom_density(aes(x = value, colour = feature, fill = feature), alpha = 0.3) +

### 2021-09-27 paper/paper.tex
Fix error in flowchart's caption.

@@
 \caption{The stages of data cleaning from the raw data to get three datasets contained yowie.
-n individuals means the number of respondents included in each stage, while n observations means the number of rows in the data.
+\# of individuals means the number of respondents included in each stage, while \# of observations means the number of rows in the data.
 The color represented the stage of data cleaning in statistical value chain (van der Loo \& de Jonge, 2018).

### 2021-09-27 paper/paper.Rmd
Modify flowchart.
Fix error in flowchart's caption.
Edit flowchart's caption.

@@
 ```{r flow-chart, fig.cap="The stages of data cleaning from the raw data to get three datasets contained yowie.
-n individuals means the number of respondents included in each stage, while n observations means the number of rows in the data.
-The color represented the stage of data cleaning in statistical value chain (van der Loo \\& de Jonge, 2018).
+\\# of individuals means the number of respondents included in each stage, while \\# of observations means the number of rows in the data.
+The color represents the stage of data cleaning in statistical value chain (van der Loo \\& de Jonge, 2018).
 Pink, blue, and green represents the raw, input, and valid data, respectively.", out.width="85%"}

### 2021-09-28 app/app.Rmd
Modify the app.

---
title: "Customize Your Threshold"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
runtime: shiny
---

```{r setup, include=FALSE}
library(flexdashboard)
library(tidyverse)
library(plotly)
library(shiny)
library(shinyWidgets)
```

```{r data}
wages_hs <- readRDS(here::here("app/wages_hs.rds")) %>%
    dplyr::select(-`X1.x`, -`X1.y`)
```

Inputs {.sidebar}
-----------------------------------------------------------------------
<br/>

<style>
div.blue { background-color:#FADBD8; border-radius: 10px; padding: 10px;}
</style>
<div class = "blue">

```{r}
p(icon("info-circle", lib = "font-awesome"), strong("About"))
```

The wages data in `yowie` package uses the weight of robust linear model to detect anomalies and imputes those anomalies with its fitted values. The threshold of weight used is 0.12 to maintain the natural variability of the data. Here, you can customize your own threshold.

</div>

<br/>

```{r}
sliderInput("th", "Threshold:",
            min = 0.01,
            max = 1,
            value = 5)
```

<br/>

Row {data-height=400}
-------------------------------------

### Spaghetti plots

```{r}

thres_data_all <- reactive({
  temp_data <- wages_hs %>%
    mutate(wages_rlm = ifelse(w < input$th  & .fitted >= 0, .fitted,
                                      mean_hourly_wage)) %>%
    mutate(is_pred = ifelse(w < input$th & .fitted >= 0, TRUE, FALSE))

  wages_compare <- temp_data %>%
    select(id, year, mean_hourly_wage, wages_rlm) %>%
    rename(original = mean_hourly_wage,
           imputed = wages_rlm) %>%
    pivot_longer(c(-id, -year), names_to = "type", values_to = "wages") %>%
    mutate(type = factor(type, levels = c("original", "imputed")))
  })

renderPlot({
  ggplot(thres_data_all()) +
    geom_line(aes(x = year,
                  y = wages,
                  group = id),
              alpha = 0.1) +
    theme_bw() +
    facet_wrap(~type, scales = "free_y") +
    theme(strip.text = element_text(size = 12))
  })
```

### Summary Statistics

```{r}
max_ori <- max(wages_hs$mean_hourly_wage)
min_ori <- round(min(wages_hs$mean_hourly_wage), 2)
med_ori <- round(median(wages_hs$mean_hourly_wage), 2)
mean_ori <- round(mean(wages_hs$mean_hourly_wage), 2)
```

```{r}
pred <- reactive({
  temp_data <- wages_hs %>%
    mutate(wages_rlm = ifelse(w < input$th  & .fitted >= 0, .fitted,
                                      mean_hourly_wage)) %>%
    mutate(is_pred = ifelse(w < input$th & .fitted >= 0, TRUE, FALSE))
  })
```

<style>
div.blue { background-color:#FADBD8; border-radius: 10px; padding: 10px;}
</style>
<div class = "blue">

```{r}
p(icon("info-circle", lib = "font-awesome"), "Summary statistics of wages in imputed data")
renderText({paste0("Minimum = ", round(min(pred()$wages_rlm), 2), " (Original = 0.01)")})
renderText({paste0("Median = ", round(median(pred()$wages_rlm), 2), " (Original = 7.2)")})
renderText({paste0("Mean = ", round(mean(pred()$wages_rlm), 2), " (Original = 11.87)")})
renderText({paste0("Maximum = ", round(max(pred()$wages_rlm), 2), " (Original = 33400)")})
```

</div>

### Box

```{r}
renderValueBox({
  n_pred <- round(nrow(filter(pred(), is_pred == TRUE))/nrow(pred())*100,2)
  valueBox(
    value = n_pred,
    caption = "percentage of imputed observations",
    icon = "fa-area-chart"
  )
})

```

Row {data-height=650}
-------------------------------------

### Individual plots

```{r}
actionButton("resample", "Sample individuals to display")
```

```{r}

compare_data <- eventReactive(input$resample, {
  sample_id <- sample(unique(wages_hs$id), 16)
  sample <- subset(wages_hs, id %in% sample_id)
})

thres_data <- reactive({
  temp_data <- compare_data() %>%
    mutate(wages_rlm = ifelse(w < input$th  & .fitted >= 0, .fitted,
                                      mean_hourly_wage)) %>%
    mutate(is_pred = ifelse(w < input$th & .fitted >= 0, TRUE, FALSE))

  wages_compare <- temp_data %>%
    select(id, year, mean_hourly_wage, wages_rlm) %>%
    rename(original = mean_hourly_wage,
           imputed = wages_rlm) %>%
    pivot_longer(c(-id, -year), names_to = "type", values_to = "wages") %>%
    mutate(type = factor(type, levels = c("original", "imputed")))
  })

renderPlot({
  ggplot(thres_data()) +
    geom_line(aes(x = year,
                  y = wages,
                  colour = type,
                  linetype = type),
              alpha = 1,
              size = 1) +
    geom_point(aes(x = year,
                   y = wages,
                   colour = type),
               alpha = 0.5,
               size = 1) +
    scale_color_manual(values = c("#229954", "#DC7633")) +
    theme_bw() +
    theme(axis.text.x = element_text(angle = 10, size = 5),
          legend.position = "bottom",
          legend.text = element_text(size = 12),
          legend.key.size = unit(6,"line")) +
    facet_wrap(~id, scales = "free_y")
  })

```

### 2021-09-28 app/app.R

@@ -1,223 +1,0 @@
-#
-# This is a Shiny web application. You can run the application by clicking
-# the 'Run App' button above.
-#
-# Find out more about building applications with Shiny here:
-#
-#    http://shiny.rstudio.com/
-#
-
-library(shiny)
-library(shinythemes)
-library(tidyverse)
-
-
-# LOAD THE DATA
-
-wages_hs <- readRDS(here::here("app/wages_hs.rds")) %>%
-    dplyr::select(-`X1.x`, -`X1.y`)
-
-
-## Make it separate tab for these three plots
-
-## Add progress bar for how long will it take to recalculate threshold and resample the ids
-
-ui <- bootstrapPage(
-    navbarPage(theme = shinytheme("cerulean"),
-               collapsible = TRUE,
-               # Application title
-               "Customize Your Threshold",
-               tabPanel("Set the weight threshold",
-                        sidebarLayout(
-                            sidebarPanel(width = 4,
-                                   sliderInput("th",
-                                               "Set the weight threshold:",
-                                               min = 0.01,
-                                               max = 1,
-                                               value = 5),
-                                   br(),
-                                   br(),
-                                   actionButton("resample",
-                                             "Compare individuals")
-
-                            ),
-
-                            # Show a plot of the samples of indivudial before and after
-                            mainPanel(width = 8,
-                                      plotOutput("compar_plot",
-                                              height = 500)
-                                      # h3("Before"),
-                                      # plotOutput("spag1"),
-                                      # br(),
-                                      # textOutput("summary_before"),
-                                      # textOutput("n_obs"),
-                                      # textOutput("n_id"),
-                                      #
-                                      # h3("After"),
-                                      # plotOutput("spag2"),
-                                      # br(),
-                                      # textOutput("summary_after"),
-                                      # textOutput("n_obs_pred"),
-                                      # textOutput("n_id_pred")
-
-                            )
-                        )),
-               tabPanel(
-                   "Inspect the Potential Outliers",
-                   column(
-                       width = 12,
[…]

### 2021-09-28 paper/paper.Rmd
Edit the bullet point in conclusion based on isse #8.

@@
 - determining which records should be downloaded from the database.
-- there are many errors in the data, e.g. hourly wages greater than $30,000 per hour, and there was no information on how these were handled in the original data.
+- there are many errors in the data, e.g.
+hourly wages greater than $30,000 per hour.
+Further for original data, wages greater than 75 dollars were set to be missing.
+However, in recent context, this value might be too low to be set as the maximum threshold.
 - there is no explicit variable in the database recording high school dropout, which means we needed to compare date of 12th grade with the individual's age.
@@
 
-- **Codes** : R script to reproduce data tidying and cleaning are available in this [Github Repository](https://github.com/numbats/yowie/blob/master/data-raw/data_preprocessing.R).
+- **Codes** : R script to reproduce data tidying and cleaning are available in this [page](https://numbats.github.io/yowie/articles/process-data.html).
 
@@
 
-- **shiny app**: An web interactive `shiny` app to run a simulation to customize the weight threshold. This app could be accessed [here](https://github.com/numbats/wages-refresh/tree/main/app).
+- **shiny app**: An web interactive `shiny` app to customize the weight threshold. This app could be accessed [here](https://github.com/numbats/wages-refresh/tree/main/app).
 
